### 2024-05-03 MCR_Algae_Data_Cleaning.Rmd
adding in the data cleaning rmd and raw data file.

---
title: "MCR_Algae_Data_Cleaning"
author: "Noam Altman-Kurosaki and Lauren Enright"
date: "2024-05-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls()) # clean up
library(knitr)
library(dplyr)
library(plyr)
library(car)
library(reshape)
library(reshape2)
library(tidyr)
library(effects)
library(devtools)
opts_chunk$set(comment="  ",
               collapse=TRUE, 
               echo=FALSE,
               #fig.asp=1/gr,
               fig.height=8,
               fig.width = 10,
               dev="png",
               warning=TRUE
               )
opts_knit$set(eval.after = "fig.cap")

```

```{r data cleaning}
library(here)

data <- read.csv(here("data", "MCR_LTER_Annual_Survey_Benthic_Cover_20231211.csv"), stringsAsFactors = F) # replacements don't work when strings are factors
str(data)
levels(as.factor(data$Taxonomy_Substrate_Functional_Group)) # 86 levels
data$Site_habitat <- as.factor( paste(data$Site, data$Habitat))

#### REPLACE OUTDATED TAXA

# Caulerpa peltata is outdated --> Caulerpa racemosa
data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Caulerpa peltata"] <- "Caulerpa racemosa"

# Boodlea kaeneana --> Cladophoropsis membranacea
data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Boodlea kaeneana"] <- 
  "Cladophoropsis membranacea"

# Cladophoropsis luxurians --> Cladophora fuliginosa
data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Cladophoropsis luxurians"] <-
  "Cladophora fuliginosa"

# Neogoniolithon frutescens --> Neogoniolithon brassica-florida
data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Neogoniolithon frutescens"] <- 
  "Neogoniolithon brassica-florida"

# remove old levels
data <- droplevels(data)
levels(as.factor(data$Taxonomy_Substrate_Functional_Group)) # 84 levels

```

```{r QA QC - checking quadrats}

quad_level_summary <- data %>%
  dplyr::group_by(Year, Location, Site, Habitat, Site_habitat, Transect, Quadrat) %>%
  dplyr::summarize(sum = sum(Percent_Cover))

non100_quad_level <- quad_level_summary %>%
  dplyr::filter(sum != 100)

#99 observations that do not equal 100

#how many are no data (-1)
quad_level_summary %>%
  dplyr::filter(sum == -1)

#14

mistakes_quad_level <- quad_level_summary %>%
  dplyr::filter(sum != 100) %>%
  dplyr::filter(sum != -1)

#remaining 85 have something weird happening! Will send to Bob for updates. For now, 
# will filter these out

unique(quad_level_summary$sum)

#write out these 85

#write.csv(mistakes_quad_level, "mistakes_quad_level.csv", row.names = FALSE)
```

```{r subsetting the data }

### CREATE DATAFRAME THAT IS JUST ALGAE

not_algae <- c("Ascidian", "Bare Space", "Coral", "Coral Rubble", "Corallimorpharia", "Heteractis sp.", "Millepora platyphylla", "No data", "Sand", "Sacrophyton sp.", "Shell Debris", "Soft Coral", "Sponge", "Tridacna sp.", "Zooxanthellae")

### CREATE DATAFRAME THAT IS JUST MACROALGAE
# Maintaining mat-forming cyanos even though they're not macroalgae in the traditional sense
# "Coelothrix irregularis", "Symploca hydnoides"

not_macro <- c("Algal Turf", "Crustose Corallines", "Cyanophyta", "Damselfish Turf", "Lithophyllum kotschyanum", "Neogoniolithon brassica-florida", "Sporolithon sp." )

```

```{r wide - with all data}

comb_wide <- cast(data, Year + Location + Site + Habitat + Site_habitat + Transect + Quadrat ~ Taxonomy_Substrate_Functional_Group, fun.aggregate = mean, value = "Percent_Cover") # this is the correct number of rows given the missing data

#change NAs to 0s

is.nan.data.frame <- function(x){
  do.call(cbind, lapply(x, is.nan))}

comb_wide[is.nan.data.frame(comb_wide)] <- 0

```

```{r QAQC for the wide dataset}

comb_wide_quad_sum <- comb_wide %>%
  mutate(row_sums = rowSums(dplyr::select(., 8:91)))

non100_wide_quad_level <- comb_wide_quad_sum %>%
  dplyr::filter(row_sums != 100)

#also 99 observations that do not equal 100

#how many are no data (-1)
comb_wide_quad_sum %>%
  dplyr::filter(row_sums == -1)

#14

mistakes_wide_quad_level <- comb_wide_quad_sum %>%
  dplyr::filter(row_sums != 100) %>%
  dplyr::filter(row_sums != -1)

#remaining 85 have something weird happening! Will send to Bob for updates. For now, 
# will filter these out

#want to make sure these 85 are the same we found using the long data

setdiff(mistakes_wide_quad_level$Location, mistakes_quad_level$Location)

setdiff(mistakes_quad_level$Location, mistakes_wide_quad_level$Location)

#yay, they are the same 85 
```

```{r filter the wide dataset}

#we want to filter out 2005 and 2006 - these were "ramp up years"
#also want to filter out rows where the sum does not = 0 
wide_alldata_filtered <- comb_wide_quad_sum %>%
  dplyr::filter(Year != 2005 & Year != 2006) %>%
  dplyr::filter(row_sums == 100)

write.csv(wide_alldata_filtered, "allbenthicdata_cleaned.csv", row.names = FALSE)

#we also want a copy that only includes taxa we care about 
#not_algae character string created by Noam in Chunk 4 (Subsetting the data) and includes 14 
#categories  -- these categories will be listed in a separate document as well

wide_algaeonly_filtered <- wide_alldata_filtered[ , -which(names(wide_alldata_filtered) 
                                                           %in% not_algae)]

#need to still remove the row_sums column for algae only

wide_algaeonly_filtered <- wide_algaeonly_filtered %>%
  dplyr::select(-row_sums)

write.csv(wide_algaeonly_filtered, "algaeonly_cleaned.csv", row.names = FALSE)

```

## Anything below this line was to help identify missing data, overall taxa levels, etc. and does not need to be run to get a cleaned dataframe

```{r subsetting the data }

### CREATE DATAFRAME THAT IS JUST ALGAE
not_algae <- c("Ascidian", "Bare Space", "Coral", "Coral Rubble", "Corallimorpharia", "Heteractis sp.", "Millepora platyphylla", "No data", "Sand", "Sacrophyton sp.", "Shell Debris", "Soft Coral", "Sponge", "Tridacna sp.", "Zooxanthellae")

algae_df <- data[ ! data$Taxonomy_Substrate_Functional_Group %in% not_algae, ]
algae_df <- droplevels(algae_df)
levels(as.factor(algae_df$Taxonomy_Substrate_Functional_Group )) # 70 levels

### CREATE DATAFRAME THAT IS JUST MACROALGAE
# Maintaining mat-forming cyanos even though they're not macroalgae in the traditional sense
# "Coelothrix irregularis", "Symploca hydnoides" 
not_macro <- c("Algal Turf", "Crustose Corallines", "Cyanophyta", "Damselfish Turf", "Lithophyllum kotschyanum", "Neogoniolithon brassica-florida", "Sporolithon sp." )

macro_df <- algae_df[! algae_df$Taxonomy_Substrate_Functional_Group %in% not_macro, ]
macro_df <- droplevels(macro_df)
levels(as.factor(macro_df$Taxonomy_Substrate_Functional_Group )) # 63 levels

# NOTE: WHICH DO WE WANT TO WORK WITH?
# MY VOTE IS "algae_df"

```

```{r create wide df - algae only}
# using algae data
# I don't think the date of the transect matters. Dropping that.

algae_wide <- cast(algae_df, Year + Location + Site + Habitat + Site_habitat + Transect + Quadrat ~ Taxonomy_Substrate_Functional_Group, fun.aggregate = mean, value = "Percent_Cover")
# replace NaN in dataframe
# function from https://stackoverflow.com/questions/18142117/how-to-replace-nan-value-with-zero-in-a-huge-data-frame

is.nan.data.frame <- function(x){
  do.call(cbind, lapply(x, is.nan))}

algae_wide[is.nan.data.frame(algae_wide)] <- 0

length(levels(as.factor(algae_wide$Year))) # 19 years of data collection
length(levels(algae_wide$Site_habitat)) # 24 sites
# 5 transects and 10 quads per transect
# should be 22800 rows, so clearly some years are missing sites

```

```{r missing data part 1}

### NOTE - USED CHAT GPT FOR THIS NEXT PART
# Step 1: Create a contingency table of counts for each combination of year and site
year_site_counts <- table(algae_wide$Year, algae_wide$Site_habitat)

# Step 2: Identify rows with zero counts (indicating missing sites)
missing_years <- rownames(year_site_counts)[rowSums(year_site_counts == 0) > 0] # 2005, 2006, 2021 missing data

# should be 19200 rows if we remove those years
nrow(algae_wide[! algae_wide$Year %in% missing_years,]) # 19186... what are those 14 datapoints??

# Step 3: Extract the years corresponding to missing sites
for (year in missing_years) {
  # Filter the dataframe for the current missing year
  missing_year_data <- subset(algae_wide, Year == year)

  # Get the unique sites present in the missing year data
  present_sites <- unique(missing_year_data$Site_habitat)

  # Get all possible sites
  all_sites <- unique(algae_wide$Site_habitat)

  # Identify missing sites
  missing_sites <- setdiff(all_sites, present_sites)

  # Print missing sites for the current year
  cat("Year", year, "is missing data from sites:", paste(missing_sites, collapse = ", "), "\n")
}

# try to find what else is missing
algae_subset <- algae_wide[! algae_wide$Year %in% missing_years,]

site_transect_counts <- table(algae_subset$Site_habitat, algae_subset$Transect)
site_quadrat_counts <- table(algae_subset$Site_habitat, algae_subset$Quadrat)

# Step 2: Identify sites with missing transects or quadrats
missing_transect_sites <- rownames(site_transect_counts)[rowSums(site_transect_counts == 0) > 0]
missing_quadrat_sites <- rownames(site_quadrat_counts)[rowSums(site_quadrat_counts == 0) > 0]

# Step 3: Print out sites with missing transects or quadrats
print(paste("Sites with missing transects:", missing_transect_sites))
print(paste("Sites with missing quadrats:", missing_quadrat_sites))
 # Hm, nothing missing
```

```{r missing data part 2}
# create a dataframe with all possible combinations of year, site, habitat, transect, quadrat
year <- unique(data$Year)
sites <- unique(data$Site)
habitat <- unique(data$Habitat)
transect <- unique(data$Transect)
quadrat <- unique(data$Quadrat)
all_comb <- expand_grid(year, sites, habitat, transect, quadrat)
all_comb$comb <- paste(all_comb$year, all_comb$sites, all_comb$habitat, all_comb$transect, all_comb$quadrat)

# create a vector with the combinations that are actually present in the data
data$comb <- paste(data$Year, data$Site, data$Habitat, data$Transect, data$Quadrat) 
data_comb <- unique(data$comb)

missing_quadrats <- setdiff(all_comb$comb, data_comb)
length(missing_quadrats) # 1800 missing quadrats, consistent with the missing sites/years above
# extra missing values might be from quadrats with no algae in wide dataframe?

```

```{r wide - with all data}

comb_wide <- cast(data, Year + Location + Site + Habitat + Site_habitat + Transect + Quadrat ~ Taxonomy_Substrate_Functional_Group, fun.aggregate = mean, value = "Percent_Cover") # this is the correct number of rows given the missing data

#change NAs to 0s

is.nan.data.frame <- function(x){
  do.call(cbind, lapply(x, is.nan))}

comb_wide[is.nan.data.frame(comb_wide)] <- 0

# remove non-algae columns with subset
algae_wide_2 <- comb_wide[ , -which(names(comb_wide) %in% not_algae)]

```

### 2024-05-03 MCR_Algae_Data_Cleaning.Rmd
updated cleaning to identify which transects are missing quads after filtering

@@
 
+#do we have data for all transects?
+
+comb_wide_quad_sum %>%
+  dplyr::group_by(Year, Site) %>%
+  dplyr::summarise(count = n())
+
+#should be 200 per year per site (50 at each habitat * 4 habitats)
+#2005 and 2006 are one large collective survey
+#2021 sites are at 100 (did not do forereef habitats due to COVID)
+
+#do all quads = 100 ?
+
 non100_wide_quad_level <- comb_wide_quad_sum %>%
@@
 
+comb_wide_quad_sum %>%
+ dplyr::filter(row_sums != 100) %>% dplyr::filter(Year == 2020)
+
 #remaining 85 have something weird happening! Will send to Bob for updates. For now, 
@@
 #yay, they are the same 85
+
+```
+```{r}
+
+test <- comb_wide_quad_sum %>%
+  dplyr::group_by(Year, Site, Habitat, Transect) %>%
+  dplyr::summarise(sum = sum(Quadrat))
+
+test <- comb_wide_quad_sum %>%
+  dplyr::group_by(Year, Site, Habitat) %>%
+  dplyr::summarise(sum = sum(Transect))
+
+test <- comb_wide_quad_sum %>%
+  dplyr::group_by(Year, Site) %>%
+  dplyr::summarise(count = n())
+
+unique(test$sum)
+
+test <- comb_wide_quad_sum %>%
+  dplyr::group_by(Year) %>%
+  dplyr::summarise(sum = sum(Transect))
+
+test <- wide_alldata_filtered %>%
+  dplyr::group_by(Year, Site, Habitat) %>%
+  dplyr::summarise(count = n())
+
+nique(comb_wide_quad_sum$Year)
+
+sum(test$sum)
+
+(1*10) + (2*10) + (3*10) + (4*10) + (5*10)
+
 ```
@@
 
-## Anything below this line was to help identify missing data, overall taxa levels, etc. and does not need to be run to get a cleaned dataframe
+```{r filtered data QA QC}
+# you can use this to see which sites lost quads during filtering process 
+#should be 50 per year/site/habitat combo
+
+wide_alldata_filtered %>%
+  dplyr::group_by(Year, Site, Habitat) %>%
+  dplyr::summarise(count = n())
+
+wide_alldata_filtered %>%
+  dplyr::group_by(Year, Site) %>%
+  dplyr::summarise(count = n())
+
+```
+
+## Anything below this line was to help identify missing data, overall taxa levels, etc. and does not necessarily need to be run to get a cleaned dataframe
 
@@
 # Step 1: Create a contingency table of counts for each combination of year and site
-year_site_counts <- table(algae_wide$Year, algae_wide$Site_habitat)
+year_site_counts <- table(comb_wide$Year, comb_wide$Site_habitat)
 
@@
 ```
-
-```{r wide - with all data}
-
-comb_wide <- cast(data, Year + Location + Site + Habitat + Site_habitat + Transect + Quadrat ~ Taxonomy_Substrate_Functional_Group, fun.aggregate = mean, value = "Percent_Cover") # this is the correct number of rows given the missing data
-
-#change NAs to 0s
-
-is.nan.data.frame <- function(x){
-  do.call(cbind, lapply(x, is.nan))}
-
-comb_wide[is.nan.data.frame(comb_wide)] <- 0
-
-# remove non-algae columns with subset
-algae_wide_2 <- comb_wide[ , -which(names(comb_wide) %in% not_algae)]
-
-```

### 2024-05-07 MCR_Algae_Data_Cleaning.Rmd
removing lines of code that were just tests and unimportant

@@
 ```
-```{r}
-
-test <- comb_wide_quad_sum %>%
-  dplyr::group_by(Year, Site, Habitat, Transect) %>%
-  dplyr::summarise(sum = sum(Quadrat))
-
-test <- comb_wide_quad_sum %>%
-  dplyr::group_by(Year, Site, Habitat) %>%
-  dplyr::summarise(sum = sum(Transect))
-
-test <- comb_wide_quad_sum %>%
-  dplyr::group_by(Year, Site) %>%
-  dplyr::summarise(count = n())
-
-unique(test$sum)
-
-test <- comb_wide_quad_sum %>%
-  dplyr::group_by(Year) %>%
-  dplyr::summarise(sum = sum(Transect))
-
-test <- wide_alldata_filtered %>%
-  dplyr::group_by(Year, Site, Habitat) %>%
-  dplyr::summarise(count = n())
-
-nique(comb_wide_quad_sum$Year)
-
-sum(test$sum)
-
-(1*10) + (2*10) + (3*10) + (4*10) + (5*10)
-
-```
 

### 2024-05-29 noam_data_exploration_v1.Rmd
Add files via upload

---
title: "noam_data_exploration_v1"
author: "Noam Altman-Kurosaki"
date: "2024-05-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(vegan)
library(glmmTMB)
library(car)
library(effects)
library(DHARMa)
library(stringr)
library(MuMIn)
```

```{r }
data <- read.csv(here::here("data", "algaeonly_cleaned.csv"), stringsAsFactors = F) # replacements don't work when strings are factors

str(data)

#pivot long

long_data <- pivot_longer(
  data = data,
  cols = 8:77,
  names_to = "Taxa",
  values_to = "Percent_cover")

transect_level <- long_data %>%
  dplyr::group_by(Year, Site, Habitat, Transect, Taxa) %>%
  summarise(Mean_cover = mean(Percent_cover), 
            SE = sd(Percent_cover)/sqrt(n()))

site_level <- transect_level %>%
  dplyr::group_by(Year, Site, Habitat, Taxa) %>%
  summarise(Mean_site_cover = mean(Mean_cover), 
            Site_se = sd(Mean_cover)/sqrt(n()))
```

```{r diversity indicies, fig.height = 15, fig.width=20}
meta_cols <- colnames(data[1:7]) 
species_cols <- colnames(data[8:77])

wide_transect_level <- data %>%
  dplyr::group_by(Year, Site, Habitat, Site_habitat, Transect) %>%
  summarize(across(all_of(species_cols), mean, na.rm = TRUE))

alpha_diversity <-
  data.frame(
    Year = as.factor(wide_transect_level$Year),
    Site = as.factor(wide_transect_level$Site),
    Habitat = as.factor(wide_transect_level$Habitat),
    Site_habitat = as.factor(wide_transect_level$Site_habitat),
    Transect = as.factor(wide_transect_level$Transect),
    Site_habitat_transect = as.factor(paste(wide_transect_level$Site_habitat, wide_transect_level$Transect)),
    Shannon = diversity(wide_transect_level[5:74], "shannon"),
    Richness = rowSums(wide_transect_level[5:74] > 0)
  )

alpha_diversity_means <- alpha_diversity %>%
  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
  summarise(Richness_mean = mean(Richness), 
            Richness_se = sd(Richness)/sqrt(n()),
            Shannon_mean = mean(Shannon), 
            Shannon_se = sd(Shannon)/sqrt(n()))

site_colours <- c("LTER 1" = "cornflowerblue",
               "LTER 2" = "darkblue",
               "LTER 3" = "yellowgreen",
               "LTER 4" = "forestgreen",
               "LTER 5" = "mediumorchid",
               "LTER 6" = "orange")
habitat_linetypes <- c("Fringing" = "solid",
                       "Backreef" = "longdash",
                       "Outer 10" = "twodash",
                       "Outer 17" = "dotted")
habitat_shapes <- c("Fringing" = 16,
                     "Backreef" = 17,
                     "Outer 10" = 15,
                     "Outer 17" = 18)

(richness_plot <-
  ggplot(data = alpha_diversity_means) +
  geom_point(aes(x = Year, y = Richness_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8) +
  geom_line(aes(x = Year, y = Richness_mean, colour = Site, group = Site_habitat), size = 1.2) +
  geom_errorbar(aes(x = Year, ymin = Richness_mean - Richness_se , ymax = Richness_mean + Richness_se, colour = Site,
                    group = Site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Habitat)+
  labs(title = "a. Richness\n", y = "Richness\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )

(shannon_plot <-
  ggplot(data = alpha_diversity_means) +
  geom_point(aes(x = Year, y = Shannon_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8) +
  geom_line(aes(x = Year, y = Shannon_mean, colour = Site, group = Site_habitat), size = 1.2) +
  geom_errorbar(aes(x = Year, ymin = Shannon_mean - Shannon_se , ymax = Shannon_mean + Shannon_se, colour = Site,
                    group = Site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Habitat)+
  labs(title = "b. Shannon\n", y = "shannon\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )

```

```{r alpha diversity panel graph, fig.height = 30, fig.width= 20}

ggarrange(richness_plot + rremove("x.text"), shannon_plot, common.legend = T,
          nrow = 2)

# seems like patterns in alpha diversity different depending on if you look at richness or shannon diversity 
```

```{r basic alpha diversity stats}
# these violate all sorts of assumptions I'm just messing around
richness_mod <- glmmTMB(Richness ~ Year*Site*Habitat + (1|Site_habitat_transect), data = alpha_diversity, family = poisson())
Anova(richness_mod)

# shannon_mod <- glmmTMB(Shannon ~ Year*Site*Habitat + (1|Site_habitat_transect), data = alpha_diversity, family = Gamma(link = "log")) # 0 values fuck this up. Return to this when not tired. 
```

```{r nmds, fig.height = 15, fig.width = 15}
species_matrix <- as.matrix(wide_transect_level[6:74])
rownames(species_matrix) <- paste(wide_transect_level$Year, wide_transect_level$Site_habitat, wide_transect_level$Transect)
rows_to_remove <- rowSums(species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
species_matrix <- species_matrix[!rows_to_remove,] 
wide_transect_level_subset <- wide_transect_level[!rows_to_remove,] # make sure rows are also removed from dataframe
wide_transect_level_subset$Stratum <- paste(wide_transect_level_subset$Year, "-", wide_transect_level_subset$Site, "-", wide_transect_level_subset$Habitat )

# WARNING - Before you run this, know that it will take a long time
set.seed(1032) 
ord2 <- metaMDS(species_matrix, dist	= "bray",	trymax	= 100,	k	= 2) # looks like it automatically wisconsin transforms now?
ord2 # stress = 0.22
stressplot(ord2)

set.seed(1032)
ord3 <- metaMDS(species_matrix, dist	= "bray",	trymax	= 100,	k	= 3)
ord3 # 0.166
stressplot(ord3)

pro <- procrustes(ord2, ord3) # visually compare w procrustes rotation

plot(pro, cex=1.5)

site.scores <- as.data.frame(vegan::scores(ord3)$sites)  # extract the site scores and convert to a data.frame
# add back in metadata
site.scores$Year <- as.factor(wide_transect_level_subset$Year)
site.scores$Site <- as.factor(wide_transect_level_subset$Site)
site.scores$Habitat <- as.factor(wide_transect_level_subset$Habitat)
site.scores$Site_habitat <- as.factor(wide_transect_level_subset$Site_habitat)
site.scores$Transect <- as.factor(wide_transect_level_subset$Transect)
site.scores$Stratum <- as.factor(wide_transect_level_subset$Stratum)

cent <- aggregate(cbind(NMDS1, NMDS2) ~ Stratum, data = site.scores, FUN = mean) # compute centroids for NMDS1 and NMDS2 
cent$NMDS1 <- as.numeric(cent$NMDS1) # read in as characters for some reason
cent$NMDS2 <- as.numeric(cent$NMDS2)
# Need to split stratum up for plotting later
cent<- 
  cent %>%
    separate(Stratum, into = c("Year", "Site", "Habitat"), sep = "-")
cent$Site_habitat <- paste(cent$Site, cent$Habitat)
# Add back in stratum column
cent$Stratum <-  aggregate(cbind(NMDS1, NMDS2) ~ Stratum, data = site.scores, FUN = mean)$Stratum
# A lot of extra spaces

cent <- cent %>%
  mutate(across(everything(), str_trim))
cent$Year <- as.factor(cent$Year)
cent$Habitat <- as.factor(cent$Habitat)
cent$Site <- as.factor(cent$Site)
cent$Site_habitat <- as.factor(cent$Site_habitat)

species.scores <- as.data.frame(scores(ord3)$species)  # extract the species scores and convert to a data.frame
species.scores$Species <- rownames(species.scores)

ggplot() + 
  geom_point(data= site.scores, aes(x=NMDS1,y=NMDS2, shape = Habitat, colour = Site),
             alpha = 0.25, size=2) + # add the point markers
#  geom_text(data= species.scores,aes(x=NMDS1,y=NMDS2, label = Species), size = 4) +  # add the species labels

  geom_path(data = cent, aes(x = NMDS1, y = NMDS2, group = Site_habitat, colour = Site),
            arrow = arrow(length=unit(0.5,"cm"), ends="last", type = "closed"), size = 1) +
  facet_wrap(~Habitat) +
  scale_linetype_manual(values = habitat_linetypes) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        strip.text.x = element_text(size = 25),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(1.5)), # increase legend text
        legend.title = element_text(size = rel(1.5)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) + # add box around legend
  guides(colour=guide_legend(title.position="top", # make color legend title go above legend
                                     title.hjust =0.5, nrow = 2),
         shape=guide_legend(title.position="top",  # make shape legend title go above legend
                                     title.hjust =0.5,
                            nrow = 2),
         path = guide_legend(title.position="top",  # make shape legend title go above legend
                                     title.hjust =0.5)
         )

# looks like stochasticity increases from outer 17 --> Fringing
# Even in lagoon though it looks like communities return to initial state... except for North Shore LTER sites
# ways to test stochasticity vs. determinism?
# seems like deterministic trajectories, although we're not starting from a clean slate
# accounting for founder effects?

# Want to plot mean distance to centroid, but not sure how to do that with so many factors/interactions...
bray_matrix <- vegdist(species_matrix, method = "bray")
year_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Year)
year_beta_disper #

# plot of dispersion test
plot(year_beta_disper, sub = "") # lol
boxplot(year_beta_disper, main = "Year", xlab = "")

site_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Site)
plot(site_beta_disper, sub = "") # site five stands out
boxplot(site_beta_disper, main = "Site", xlab = "")

habitat_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Habitat)
plot(habitat_beta_disper, sub = "") 
boxplot(habitat_beta_disper, main = "Habitat", xlab = "") # greatest variance in fringing it seems (as expected)

site_habitat_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Site_habitat)
plot(site_habitat_beta_disper, sub = "") 
boxplot(site_habitat_beta_disper, main = "Site_habitat", xlab = "")

```

```{r CV, fig.height = 10, fig.width = 10}
# coefficient of variation
CV<-function(x){
  return(sd(x,na.rm=T)/mean(x,na.rm=T))
}

alpha_diversity$Cover <- rowSums(wide_transect_level[5:74])/100

diversity_stability <- alpha_diversity %>%
  dplyr::group_by(Site, Habitat, Site_habitat) %>%
  summarise(Richness_mean = mean(Richness), 
            Richness_se = sd(Richness)/sqrt(n()),
            Richness_cv = CV(Richness),
            Shannon_mean = mean(Shannon), 
            Shannon_se = sd(Shannon)/sqrt(n()),
            Shannon_cv = CV(Shannon),
            Cover_mean = mean(Cover),
            Cover_se = sd(Cover)/sqrt(n()),
            Cover_cv = CV(Cover))

diversity_mod_1 <- glmmTMB(Cover_cv ~ Richness_mean, family = Gamma(link = "log"), data = diversity_stability)
summary(diversity_mod_1)
plot(allEffects(diversity_mod_1)) # richness reduces cv (i.e. enhances diversity!)

diversity_mod_2 <- glmmTMB(Cover_cv ~ Shannon_mean, family = Gamma(link = "log"), data = diversity_stability)
summary(diversity_mod_2) # similar results

diversity_mod_3 <- glmmTMB(Cover_cv ~ Habitat*Site, family = Gamma(link = "log"), data = diversity_stability)
summary(diversity_mod_3) # christ 
Anova(diversity_mod_3) # all highly sig but what the fuck am I doing here

diversity_mod_4 <- glmmTMB(Cover_cv ~ Habitat, family = Gamma(link = "log"), data = diversity_stability)
summary(diversity_mod_4)
Anova(diversity_mod_4)
plot(allEffects(diversity_mod_4)) # greatest variance in the lagoon (makes sense given beta diversity trends above)

diversity_mod_5 <- glmmTMB(Cover_cv ~ Site, family = Gamma(link = "log"), data = diversity_stability)
summary(diversity_mod_5)
Anova(diversity_mod_5) # first non-significant model! Variance in cover not driven by site
# Also makes sense given nmds plots above
# Maybe site as a random effect similar to Cline and Allgeier 2022?

diversity_mod_6 <- glmmTMB(Cover_cv ~ Shannon_mean*Habitat + (1|Site), family = Gamma(link = "log"), data = diversity_stability)
summary(diversity_mod_6) # negligible variance in random effects
Anova(diversity_mod_6) # Diversity the strongest predictor of "stability" by far!

AICc(diversity_mod_1, diversity_mod_2, diversity_mod_4, diversity_mod_6) # Model with just Shannon diversity best

(richness_cv_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = Richness_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 8) + 
   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

(shannon_cv_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 8) + 
   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

# maybe some heteroskedasticity?
# just eyeballing but it looks like "stability" is highest in LTER 2 Fringe, which also has low diversity
# Possibility for context-dependence. Alternative mechanisms drive stability?
# Functional form vs. portfolio effects?

# Stopping for now
# GOALS: test portfolio effects, look at turnover and temporal beta diversity stats, ways to test beta dispersion

```

### 2024-05-30 noam_data_exploration_v2.Rmd
Add files via upload

--- 2024-05-29 noam_data_exploration_v1.Rmd
@@
                      "Backreef" = 17,
-                     "Outer 10" = 15,
-                     "Outer 17" = 18)
+                     "Outer 10" = 1,
+                     "Outer 17" = 2)
 
@@
 
+# identifying species w/ > 10% cover for plotting later
+# NOTE: Used ChatGPT to generate these next few lines of code but it looked good so I went with it
+columns_to_keep <- apply(species_matrix, 2, function(x) any(x >= 10))
+abundant_species_matrix <- as.data.frame(species_matrix)[ , c(TRUE, columns_to_keep)]
+abundant_species_names <- colnames(abundant_species_matrix)
+
 # WARNING - Before you run this, know that it will take a long time
 set.seed(1032) 
-ord2 <- metaMDS(species_matrix, dist	= "bray",	trymax	= 100,	k	= 2) # looks like it automatically wisconsin transforms now?
-ord2 # stress = 0.22
-stressplot(ord2)
+# ord2 <- metaMDS(species_matrix, dist	= "bray",	trymax	= 100,	k	= 2) # looks like it automatically wisconsin transforms now?
+# ord2 # stress = 0.22
+# stressplot(ord2)
 
@@
 
-pro <- procrustes(ord2, ord3) # visually compare w procrustes rotation
+# pro <- procrustes(ord2, ord3) # visually compare w procrustes rotation
 
-plot(pro, cex=1.5)
+# plot(pro, cex=1.5)
 
@@
 cent <- aggregate(cbind(NMDS1, NMDS2) ~ Stratum, data = site.scores, FUN = mean) # compute centroids for NMDS1 and NMDS2
-cent$NMDS1 <- as.numeric(cent$NMDS1) # read in as characters for some reason
-cent$NMDS2 <- as.numeric(cent$NMDS2)
+
 # Need to split stratum up for plotting later
@@
 cent$Site_habitat <- as.factor(cent$Site_habitat)
-
+cent$NMDS1 <- as.numeric(cent$NMDS1) # read in as characters for some reason
+cent$NMDS2 <- as.numeric(cent$NMDS2)
 species.scores <- as.data.frame(scores(ord3)$species)  # extract the species scores and convert to a data.frame
 species.scores$Species <- rownames(species.scores)
+
+# subset only the species w/ > 10% cover
+abundant.species.scores <- species.scores[species.scores$Species %in% abundant_species_names,]
 
@@
              alpha = 0.25, size=2) + # add the point markers
-#  geom_text(data= species.scores,aes(x=NMDS1,y=NMDS2, label = Species), size = 4) +  # add the species labels
+  geom_text(data= abundant.species.scores,aes(x=NMDS1,y=NMDS2, label = Species), size = 4) +  # add the species labels
 
@@
 
-```{r CV, fig.height = 10, fig.width = 10}
+```{r CV, fig.height = 8, fig.width = 10}
 # coefficient of variation
@@
 diversity_stability <- alpha_diversity %>%
-  dplyr::group_by(Site, Habitat, Site_habitat) %>%
+  dplyr::group_by(Site, Habitat, Site_habitat, Transect) %>%
   summarise(Richness_mean = mean(Richness), 
@@
 summary(diversity_mod_1)
-plot(allEffects(diversity_mod_1)) # richness reduces cv (i.e. enhances diversity!)
+plot(allEffects(diversity_mod_1)) # richness reduces cv (i.e. enhances stability!)
 
@@
 summary(diversity_mod_5)
-Anova(diversity_mod_5) # first non-significant model! Variance in cover not driven by site
-# Also makes sense given nmds plots above
-# Maybe site as a random effect similar to Cline and Allgeier 2022?
+Anova(diversity_mod_5)
 
@@
 
-AICc(diversity_mod_1, diversity_mod_2, diversity_mod_4, diversity_mod_6) # Model with just Shannon diversity best
+diversity_mod_7 <- glmmTMB(Cover_cv ~ Shannon_mean*Site + (1|Site_habitat), family = Gamma(link = "log"), data = diversity_stability)
+summary(diversity_mod_7)
+Anova(diversity_mod_7)
+
+diversity_mod_universal <- glmmTMB(Cover_cv ~ Shannon_mean*Site*Habitat + (1|Site_habitat), family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
+diversity_dredge <- dredge(diversity_mod_universal)
+model.sel(diversity_dredge)
+
+# best model based on dredge:
+diversity_best <-  glmmTMB(Cover_cv ~ Shannon_mean*Site + Site*Habitat + (1|Site_habitat), family = Gamma(link = "log"), data = diversity_stability)
+summary(diversity_best) # very low variance in random effect, which is great
+Anova(diversity_best) # Interactions all significant
+plot(dp <- simulateResiduals(diversity_best)) # some outliers, but otherwise good model fit. I am fine with this.
 
@@
   ggplot(data = diversity_stability) +
-  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 8) + 
+  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
    scale_colour_manual(values = site_colours) +
    scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site) +
   theme_bw() + 
@@
 # just eyeballing but it looks like "stability" is highest in LTER 2 Fringe, which also has low diversity
-# Possibility for context-dependence. Alternative mechanisms drive stability?
 # Functional form vs. portfolio effects?
+# Either way looks sick as shit. Great clustering by habitat within site, and diversity trend is strong in LTER 3 4 5
+# LTER2 the trend is reversed
+# LTER 1 is mostly driven by habitat
+# LTER 6 potentially a wash
 
-# Stopping for now
 # GOALS: test portfolio effects, look at turnover and temporal beta diversity stats, ways to test beta dispersion
@@
 ```
+
+```{r variance vs mean diversity plot, fig.height = 8, fig.width = 10}
+(diversity_mean_cv_plot = 
+  ggplot(data = diversity_stability) +
+  geom_point(aes(x = Shannon_mean, y = Shannon_cv, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+# More diverse regions are also those that fluctuate the least in their diversity
+```
+
+```{r general cover diversity, fig.height = 30, fig.width = 30}
+(diversity_cover_plot = 
+  ggplot(data = alpha_diversity) +
+  geom_point(aes(x = Cover, y = Shannon, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Year) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         ) # ABORT ABORT ABORT
+```

### 2024-06-01 noam_data_exploration_v3.Rmd
Incomplete
Just need to do this because I can't figure out why I can't push so I'm trying this.

--- 2024-05-30 noam_data_exploration_v2.Rmd
@@
 
+```
+
+```{r beta dispersion, fig.height = 15, fig.width=20}
 # Want to plot mean distance to centroid, but not sure how to do that with so many factors/interactions...
-bray_matrix <- vegdist(species_matrix, method = "bray")
-year_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Year)
-year_beta_disper #
+# bray_matrix <- vegdist(species_matrix, method = "bray")
+# year_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Year)
+# year_beta_disper #
 
 # plot of dispersion test
-plot(year_beta_disper, sub = "") # lol
-boxplot(year_beta_disper, main = "Year", xlab = "")
+# plot(year_beta_disper, sub = "") # lol
+# boxplot(year_beta_disper, main = "Year", xlab = "") 
+# 
+# site_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Site)
+# plot(site_beta_disper, sub = "") # site five stands out
+# boxplot(site_beta_disper, main = "Site", xlab = "") 
+# 
+# 
+# habitat_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Habitat)
+# plot(habitat_beta_disper, sub = "") 
+# boxplot(habitat_beta_disper, main = "Habitat", xlab = "") # greatest variance in fringing it seems (as expected)
+# 
+# site_habitat_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Site_habitat)
+# plot(site_habitat_beta_disper, sub = "") 
+# boxplot(site_habitat_beta_disper, main = "Site_habitat", xlab = "") 
+# 
+stratum_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Stratum)
+# plot(site_habitat_beta_disper, sub = "") 
+boxplot(stratum_beta_disper, main = "Site_habitat", xlab = "")
+stratum_distances <- as.data.frame(stratum_beta_disper$distances)
+colnames(stratum_distances)[1] <- "Distance"
+stratum_distances$Stratum <- wide_transect_level_subset$Stratum
+stratum_distances$Transect <- wide_transect_level_subset$Transect
+stratum_distances$Year <- wide_transect_level_subset$Year
+stratum_distances$Site <- wide_transect_level_subset$Site
+stratum_distances$Habitat <- wide_transect_level_subset$Habitat
+stratum_distances$Site_habitat <- wide_transect_level_subset$Site_habitat
 
-site_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Site)
-plot(site_beta_disper, sub = "") # site five stands out
-boxplot(site_beta_disper, main = "Site", xlab = "")
+beta_disper_means <- stratum_distances %>%
+  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
+  summarise(Distance_mean = mean(Distance), 
+            Distance_se = sd(Distance)/sqrt(n()))
 
-habitat_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Habitat)
-plot(habitat_beta_disper, sub = "") 
-boxplot(habitat_beta_disper, main = "Habitat", xlab = "") # greatest variance in fringing it seems (as expected)
-
-site_habitat_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Site_habitat)
-plot(site_habitat_beta_disper, sub = "") 
-boxplot(site_habitat_beta_disper, main = "Site_habitat", xlab = "")
+(Distance_plot <-
+  ggplot(data = beta_disper_means) +
+  geom_point(aes(x = Year, y = Distance_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
+  geom_line(aes(x = Year, y = Distance_mean, colour = Site, group = Site_habitat), size = 1.2) +
+  geom_errorbar(aes(x = Year, ymin = Distance_mean - Distance_se , ymax = Distance_mean + Distance_se, colour = Site,
+                    group = Site_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Habitat)+
+  labs(title = "Distance to centroid\n", y = "Distance\n") +
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+        axis.line.y = element_line(colour = "black"),
+        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        axis.line.x = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(size = 32),
+        legend.key.size = unit(2, "cm"), # increase legend size
+        legend.text = element_text(size = rel(2)), # increase legend text
+        legend.title = element_text(size = rel(2.25)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal',
+        legend.box.background = element_rect(colour = "black") # make legend horizontal
+        )
+    )
 
@@
 # More diverse regions are also those that fluctuate the least in their diversity
+
+(diversity_mean_cv_plot_2 = 
+  ggplot(data = diversity_stability) +
+  geom_point(aes(x = Shannon_mean, y = Shannon_cv, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+#  facet_wrap(~Site) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         ) # awesome relationship
 ```
@@
 ```
+
+```{r turnover, fig.height = 30, fig.width=25}
+library(codyn)
+transect_level$Site_habitat <-  paste(transect_level$Site, transect_level$Habitat)
+transect_level$Stratum <- paste(transect_level$Site, "-", transect_level$Habitat, "-", transect_level$Transect)
+total_turnover<-
+  turnover(
+  df = transect_level,
+  time.var = "Year",
+  species.var = "Taxa",
+  abundance.var = "Mean_cover",
+  replicate.var = "Stratum",
+  metric = "total"
+)
+
+total_turnover<- 
+  total_turnover %>%
+    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
+total_turnover$Site_habitat <- paste(total_turnover$Site, total_turnover$Habitat)
+# Add back in stratum column
+
+# A lot of extra spaces
+
+total_turnover <- total_turnover %>%
+  mutate(across(everything(), str_trim))
+
+str(total_turnover) # everything became characters
+
+total_turnover$total <- as.numeric(total_turnover$total)
+total_turnover$Year <- as.factor(total_turnover$Year)
+total_turnover$Site <- as.factor(total_turnover$Site)
+total_turnover$Habitat <- as.factor(total_turnover$Habitat)
+
+turnover_means <-
+  total_turnover %>%
+  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
+  summarise(Turnover_mean = mean(total), 
+            Turnover_se = sd(total)/sqrt(n()))
+
+(turnover_plot <-
+  ggplot(data = turnover_means) +
+  geom_point(aes(x = Year, y = Turnover_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
+  geom_line(aes(x = Year, y = Turnover_mean, colour = Site, group = Site_habitat), size = 1.2) +
+  geom_errorbar(aes(x = Year, ymin = Turnover_mean - Turnover_se , ymax = Turnover_mean + Turnover_se, colour = Site,
+                    group = Site_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site_habitat, nrow = 6)+
+  labs(title = "a. Turnover\n", y = "Total turnover\n") +
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+        axis.line.y = element_line(colour = "black"),
+        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        axis.line.x = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(size = 32),
+        legend.key.size = unit(2, "cm"), # increase legend size
+        legend.text = element_text(size = rel(2)), # increase legend text
+        legend.title = element_text(size = rel(2.25)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal',
+        legend.box.background = element_rect(colour = "black") # make legend horizontal
+        )
+    )
+# this is useless
+
+# rank shift
+
+rank_shift<-
+  rank_shift(
+  df = transect_level,
+  time.var = "Year",
+  species.var = "Taxa",
+  abundance.var = "Mean_cover",
+  replicate.var = "Stratum"
+)
+
+rank_shift<- 
+  rank_shift %>%
+    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
+rank_shift$Site_habitat <- paste(rank_shift$Site, rank_shift$Habitat)
+# Add back in stratum column
+
+# A lot of extra spaces
+
+rank_shift <- rank_shift %>%
+  mutate(across(everything(), str_trim))
+
+str(rank_shift) # everything became characters
+
+rank_shift$MRS <- as.numeric(rank_shift$MRS)
+# Create a column with the final year from the returned time.var_pair
+rank_shift$Year <- as.numeric(substr(rank_shift$year_pair, 6, 9))
+rank_shift$Site <- as.factor(rank_shift$Site)
+rank_shift$Habitat <- as.factor(rank_shift$Habitat)
+
+rank_shift_means <-
+  rank_shift %>%
+  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
+  summarise(Rank_shift_mean = mean(MRS), 
+            Rank_shift_se = sd(MRS)/sqrt(n()))
+
+(rank_shift_plot <-
+  ggplot(data = rank_shift_means) +
+  geom_point(aes(x = Year, y = Rank_shift_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
+  geom_line(aes(x = Year, y = Rank_shift_mean, colour = Site, group = Site_habitat), size = 1.2) +
+  geom_errorbar(aes(x = Year, ymin = Rank_shift_mean - Rank_shift_se , ymax = Rank_shift_mean + Rank_shift_se, colour = Site,
+                    group = Site_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site_habitat, nrow = 6)+
+  labs(title = "a. rank_shift\n", y = "Total rank_shift\n") +
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+        axis.line.y = element_line(colour = "black"),
+        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        axis.line.x = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(size = 32),
+        legend.key.size = unit(2, "cm"), # increase legend size
+        legend.text = element_text(size = rel(2)), # increase legend text
+        legend.title = element_text(size = rel(2.25)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal',
+        legend.box.background = element_rect(colour = "black") # make legend horizontal
+        )
+    )
+
+# rank shift
+
+rate_change<-
+  rate_change_interval(
+  df = transect_level,
+  time.var = "Year",
+  species.var = "Taxa",
+  abundance.var = "Mean_cover",
+  replicate.var = "Stratum"
+)
+
+rate_change<- 
+  rate_change %>%
+    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
+rate_change$Site_habitat <- paste(rate_change$Site, rate_change$Habitat)
+
+# A lot of extra spaces
+
+rate_change <- rate_change %>%
+  mutate(across(everything(), str_trim))
+
+str(rate_change) # everything became characters
+
+rate_change$distance <- as.numeric(rate_change$distance)
+rate_change$interval <- as.numeric(rate_change$interval)
+rate_change$Site <- as.factor(rate_change$Site)
+rate_change$Habitat <- as.factor(rate_change$Habitat)
+
+ggplot(rate_change, aes(interval, distance, colour = Site, shape = Habitat, group = Site_habitat, linetype = Habitat)) +
+  facet_wrap(~Site) + 
+  geom_point() +
+  scale_colour_manual(values = site_colours) +
+  scale_linetype_manual(values = habitat_linetypes) +
+  stat_smooth(method = "lm", se = T, size = 2) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site)+
+#  labs(y = "Distance\n") +
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+        axis.line.y = element_line(colour = "black"),
+        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        axis.line.x = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(size = 32),
+        legend.key.size = unit(2, "cm"), # increase legend size
+        legend.text = element_text(size = rel(2)), # increase legend text
+        legend.title = element_text(size = rel(2.25)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal',
+        legend.box.background = element_rect(colour = "black") # make legend horizontal
+        )
+```

### 2024-06-12 R/Fish_tidy.R
Adding .R file for cleaning fish data (find big fish data in the drive) and the data files for summarized biomass generated by this script

@@ -1,0 +1,76 @@
+#### Tidy fish data
+### Created by: Callie Stephenson
+### Created on: 6/11/24
+
+library(tidyverse)
+library(here)
+library(fishualize)
+library(ggplot2)
+library(dplyr)
+library(stringr) #to make functional fish groups match taxonomy
+library(tidyr)
+
+fish <- read.csv("data/MCR_LTER_Annual_Fish_Survey_20230615.csv") %>% 
+  filter(!Year == "(100840 rows)")
+fish_functional <- read.csv("data/recharge_fish_fun_groups.csv")
+
+#I was given the fish_functional sheet for updated trophic levels for some species. but it doesn't have all species
+#So, I'm going to make a column for functional_group in fish that has either what it is listed as in the recharge sheet
+#and if not, I'm going to have it take the fine_trophic
+
+#Capitalize the first letter of the functional group to match style of fine trophic
+fish_functional$functional_group <- str_to_title(fish_functional$functional_group)
+
+#Change the taxonomy to have a capital letter only for Genus (the first letter)
+capitalize_first <- function(x) {
+  # Convert the first letter to uppercase and concatenate it with the rest of the string
+  paste0(toupper(substring(x, 1, 1)), substring(x, 2))
+}
+
+#make the Taxonomy joining column
+fish_functional$Taxonomy <- sapply(fish_functional$genus_species, capitalize_first)
+
+#Remove NA values and duplicates
+fish_functional_no_na <- fish_functional[!is.na(fish_functional$functional_group), ] %>% 
+  distinct(Taxonomy, .keep_all = TRUE)
+
+#change around some weirdly-named functional groups
+fish_functional_no_na_fixed <- fish_functional_no_na %>%
+  mutate(functional_group = if_else(functional_group == 'Benthicinvertebrateconsumer', 'Benthic Invertebrate Consumer', functional_group)) %>% 
+  mutate(functional_group = if_else(functional_group == 'Planktivoreexclusively', 'Planktivore_exclusively', functional_group)) %>% 
+  mutate(functional_group = if_else(functional_group == 'Fishscaleconsumer', 'Fish Scale Consumer', functional_group)) %>% 
+  mutate(functional_group = if_else(functional_group == 'Concealedcropper', 'Concealed Cropper', functional_group)) %>% 
[…]

### 2024-06-12 R/Fish_tidy.R
whoops, uploaded it kind of ugly. this just cleans up Fish_tidy so it is a bit more digestible :fish:

@@ -10,30 +10,27 @@ library(dplyr)
 library(stringr) #to make functional fish groups match taxonomy
 library(tidyr)
 
-fish <- read.csv("data/MCR_LTER_Annual_Fish_Survey_20230615.csv") %>% 
-  filter(!Year == "(100840 rows)")
-fish_functional <- read.csv("data/recharge_fish_fun_groups.csv")
+fish <- read.csv("data/MCR_LTER_Annual_Fish_Survey_20230615.csv") %>% #Find this file in google drive, too big for git
+  filter(!Year == "(100840 rows)") #i don't know why it has this but i hate it
+fish_functional <- read.csv("data/recharge_fish_fun_groups.csv") #Also find this in google drive, didn't want it public on git
 
-#I was given the fish_functional sheet for updated trophic levels for some species. but it doesn't have all species
+#I was given the fish_functional sheet for updated trophic levels for some species from recharge, but it doesn't have all species
 #So, I'm going to make a column for functional_group in fish that has either what it is listed as in the recharge sheet
 #and if not, I'm going to have it take the fine_trophic
 
+### Tidy functional sheet to match fine_trophic style: 
 #Capitalize the first letter of the functional group to match style of fine trophic
 fish_functional$functional_group <- str_to_title(fish_functional$functional_group)
-
 #Change the taxonomy to have a capital letter only for Genus (the first letter)
 capitalize_first <- function(x) {
   # Convert the first letter to uppercase and concatenate it with the rest of the string
   paste0(toupper(substring(x, 1, 1)), substring(x, 2))
 }
-
 #make the Taxonomy joining column
 fish_functional$Taxonomy <- sapply(fish_functional$genus_species, capitalize_first)
-
 #Remove NA values and duplicates
 fish_functional_no_na <- fish_functional[!is.na(fish_functional$functional_group), ] %>% 
   distinct(Taxonomy, .keep_all = TRUE)
-
 #change around some weirdly-named functional groups
 fish_functional_no_na_fixed <- fish_functional_no_na %>%
   mutate(functional_group = if_else(functional_group == 'Benthicinvertebrateconsumer', 'Benthic Invertebrate Consumer', functional_group)) %>% 
@@ -42,12 +39,9 @@ fish_functional_no_na_fixed <- fish_functional_no_na %>%
[…]

### 2024-06-12 R/Fish_tidy.R
:fish: In Fish_tidy, I changed backreef and fringing to match the algae dataset, but don't understand the difference between Outer 10 and 17, so FO (forereef for the fish surveys) is still present. I also fixed some places where packages were confusing R so that everything comes through.

@@ -47,20 +47,22 @@ combined_df <- left_join(fish, fish_functional_no_na_fixed[,c("herbivore_status"
 #Ok, so now we need to fix all the species that it didn't know where herbivores before because the herbivore_status was only in the recharge table that only has 116 species...
 herbivore_groups <- c("Cropper", "Browser", "Excavator", "Herbivore/Detritivore", "Brusher", "Scraper", "Concealed Cropper")
 cleaned_fish <- combined_df %>%
-  mutate(herbivore_status = if_else(functional_group %in% herbivore_groups, "Y", "N"))
+  mutate(herbivore_status = if_else(functional_group %in% herbivore_groups, "Y", "N")) %>%
+  mutate(Habitat = if_else(Habitat == "BA", "Backreef", Habitat)) %>% 
+  mutate(Habitat = if_else(Habitat == "FR", "Fringing", Habitat))
 
 #now lets summarize it for actual use:
 biomass_by_functional_group_wide <- cleaned_fish %>%
-  group_by(Year, Site, Habitat, Functional_group) %>%
-  summarize(total_biomass = sum(Biomass, na.rm = TRUE), .groups = 'drop') %>%
-  pivot_wider(names_from = Functional_group, values_from = total_biomass)
+  dplyr::group_by(Year, Site, Habitat, Functional_group) %>%
+  dplyr::summarize(total_biomass = sum(Biomass, na.rm = TRUE), .groups = 'drop') %>%
+  tidyr::pivot_wider(names_from = Functional_group, values_from = total_biomass)
 
 biomass_by_herb_status_wide <- cleaned_fish %>%
-  group_by(Year, Site, Habitat, herbivore_status) %>%
-  summarize(total_biomass = sum(Biomass, na.rm = TRUE), .groups = 'drop') %>%
-  pivot_wider(names_from = herbivore_status, values_from = total_biomass) %>% 
-  rename('Non-Herbivore Biomass' = N) %>% 
-  rename('Herbivore Biomass' = Y)
+  dplyr::group_by(Year, Site, Habitat, herbivore_status) %>%
+  dplyr::summarize(total_biomass = sum(Biomass, na.rm = TRUE), .groups = 'drop') %>%
+  tidyr::pivot_wider(names_from = herbivore_status, values_from = total_biomass) %>% 
+  dplyr::rename('Non-Herbivore Biomass' = N) %>% 
+  dplyr::rename('Herbivore Biomass' = Y)
 
[…]

### 2024-06-12 noam_data_exploration_v3.Rmd
test change
last minute changes before June 2024 meeting

@@
 str(data)
-
+# test
 #pivot long
@@
             arrow = arrow(length=unit(0.5,"cm"), ends="last", type = "closed"), size = 1) +
-  facet_wrap(~Habitat) +
+  facet_wrap(~factor(Habitat, levels=c('Fringing','Backreef','Outer 10','Outer 17'))) +
   scale_linetype_manual(values = habitat_linetypes) +
@@
 # Want to plot mean distance to centroid, but not sure how to do that with so many factors/interactions...
-# bray_matrix <- vegdist(species_matrix, method = "bray")
+ bray_matrix <- vegdist(species_matrix, method = "bray")
 # year_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Year)
@@
   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Habitat)+
+  facet_wrap(~factor(Habitat, levels=c('Fringing','Backreef','Outer 10','Outer 17')))+
   labs(title = "Distance to centroid\n", y = "Distance\n") +
@@
     )
+# beta dispersion increases as we move from forereef to lagoon, lagoon to shore
 
@@
 plot(dp <- simulateResiduals(diversity_best)) # some outliers, but otherwise good model fit. I am fine with this. 
+r.squaredGLMM(diversity_best) # 0.77, 0.77
 
-(richness_cv_plot = 
+(shannon_cv_plot = 
   ggplot(data = diversity_stability) +
-  geom_point(aes(x = Richness_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 8) + 
+  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 8) + 
    scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
+    stat_smooth(aes(x = Shannon_mean, y = Shannon_cv),method = "lm", se = T, size = 2) + 
   theme_bw() + 
@@
 
-(shannon_cv_plot = 
+(shannon_cv_plot_2 = 
   ggplot(data = diversity_stability) +
@@
   facet_wrap(~Site) +
+
   theme_bw() + 
@@
    scale_shape_manual(values = habitat_shapes) +
+
 #  facet_wrap(~Site) +
@@
 transect_level$Stratum <- paste(transect_level$Site, "-", transect_level$Habitat, "-", transect_level$Transect)
+
+# the turnover function calculates turnover as the proportion or species that either appear or disappear between timepoints
 total_turnover<-
@@
     )
-# this is useless
+#
+
+# examine appearance and disappearances independently 
+# refers to the proportion of species that appeared or disappeared relative to total species pool between two time points
+
+appearance <-
+  turnover(
+  df = transect_level,
+  time.var = "Year",
+  species.var = "Taxa",
+  abundance.var = "Mean_cover",
+  replicate.var = "Stratum",
+  metric = "appearance"
+)
+
+appearance<- 
+  appearance %>%
+    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
+appearance$Site_habitat <- paste(appearance$Site, appearance$Habitat)
+# Add back in stratum column
+
+# A lot of extra spaces
+
+appearance <- appearance %>%
+  mutate(across(everything(), str_trim))
+
+str(appearance) # everything became characters
+
+appearance$appearance <- as.numeric(appearance$appearance)
+appearance$Year <- as.factor(appearance$Year)
+appearance$Site <- as.factor(appearance$Site)
+appearance$Habitat <- as.factor(appearance$Habitat)
+
+appearance_means <-
+  appearance %>%
+  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
+  summarise(Appearance_mean = mean(appearance), 
+            Appearance_se = sd(appearance)/sqrt(n()))
+
+(appearance_plot <-
+  ggplot(data = appearance_means) +
+  geom_point(aes(x = Year, y = Appearance_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
+  geom_line(aes(x = Year, y = Appearance_mean, colour = Site, group = Site_habitat), size = 1.2) +
+  geom_errorbar(aes(x = Year, ymin = Appearance_mean - Appearance_se , ymax = Appearance_mean + Appearance_se, colour = Site,
+                    group = Site_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site_habitat, nrow = 6)+
+  labs(title = "b. Appearances\n", y = "Total turnover\n") +
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+        axis.line.y = element_line(colour = "black"),
+        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        axis.line.x = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(size = 32),
+        legend.key.size = unit(2, "cm"), # increase legend size
+        legend.text = element_text(size = rel(2)), # increase legend text
+        legend.title = element_text(size = rel(2.25)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal',
+        legend.box.background = element_rect(colour = "black") # make legend horizontal
+        )
+    )
+#
+
+disappearance <-
+  turnover(
+  df = transect_level,
+  time.var = "Year",
+  species.var = "Taxa",
+  abundance.var = "Mean_cover",
+  replicate.var = "Stratum",
+  metric = "disappearance"
+)
+
+disappearance<- 
+  disappearance %>%
+    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
+disappearance$Site_habitat <- paste(disappearance$Site, disappearance$Habitat)
+# Add back in stratum column
+
+# A lot of extra spaces
+
+disappearance <- disappearance %>%
+  mutate(across(everything(), str_trim))
+
+str(disappearance) # everything became characters
+
+disappearance$disappearance <- as.numeric(disappearance$disappearance)
+disappearance$Year <- as.factor(disappearance$Year)
+disappearance$Site <- as.factor(disappearance$Site)
+disappearance$Habitat <- as.factor(disappearance$Habitat)
+
+disappearance_means <-
+  disappearance %>%
+  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
+  summarise(Disappearance_mean = mean(disappearance), 
+            Disappearance_se = sd(disappearance)/sqrt(n()))
+
+(disappearance_plot <-
+  ggplot(data = disappearance_means) +
+  geom_point(aes(x = Year, y = Disappearance_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
+  geom_line(aes(x = Year, y = Disappearance_mean, colour = Site, group = Site_habitat), size = 1.2) +
+  geom_errorbar(aes(x = Year, ymin = Disappearance_mean - Disappearance_se , ymax = Disappearance_mean + Disappearance_se, colour = Site,
+                    group = Site_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site_habitat, nrow = 6)+
+  labs(title = "c. Disappearance\n", y = "Total turnover\n") +
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+        axis.line.y = element_line(colour = "black"),
+        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        axis.line.x = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(size = 32),
+        legend.key.size = unit(2, "cm"), # increase legend size
+        legend.text = element_text(size = rel(2)), # increase legend text
+        legend.title = element_text(size = rel(2.25)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal',
+        legend.box.background = element_rect(colour = "black") # make legend horizontal
+        )
+    )
 
@@
 
-# rank shift
+# rate change
+rate_change_table<-
+  rate_change(
+  df = transect_level,
+  time.var = "Year",
+  species.var = "Taxa",
+  abundance.var = "Mean_cover",
+  replicate.var = "Stratum"
+)
+# kable(rate_change_table)
 
@@
         legend.box.background = element_rect(colour = "black") # make legend horizontal
-        )
+        ) # I don't think these are linear relationships so the smoother isn't terriibly helpful
 ```

### 2024-06-13 R/Fish_tidy.R
:fish: dataset FO -> Outer 10 to make congruent with algae dataset

@@ -49,7 +49,8 @@ herbivore_groups <- c("Cropper", "Browser", "Excavator", "Herbivore/Detritivore"
 cleaned_fish <- combined_df %>%
   mutate(herbivore_status = if_else(functional_group %in% herbivore_groups, "Y", "N")) %>%
   mutate(Habitat = if_else(Habitat == "BA", "Backreef", Habitat)) %>% 
-  mutate(Habitat = if_else(Habitat == "FR", "Fringing", Habitat))
+  mutate(Habitat = if_else(Habitat == "FR", "Fringing", Habitat)) %>% 
+  mutate(Habitat = if_else(Habitat == "FO", "Outer 10", Habitat))
 
 #now lets summarize it for actual use:
 biomass_by_functional_group_wide <- cleaned_fish %>%

### 2024-07-02 julianna_data_exploration.Rmd
Made a script to do some exploration

---
title: "Pre working group data exploration: Julianna"
output: html_document
date: "2024-07-02"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load packages
require(here)
require(tidyverse)
```

# Bring in the data

```{r}

```

### 2024-07-02 mcr_algal_data_cleaning_final.Rmd
use this cleaning script with the raw data

--- 2024-05-07 MCR_Algae_Data_Cleaning.Rmd
@@
 
+data %>% 
+  filter(Taxonomy_Substrate_Functional_Group == "No data")
+
 #99 observations that do not equal 100
@@
 
+#21000 obs of 91 variables
+
 #change NAs to 0s
@@
 comb_wide[is.nan.data.frame(comb_wide)] <- 0
+
+#21000 obs of 91 variables
 
@@
 
-#also 99 observations that do not equal 100
+#also 99 observations that do not equal 100, which matches what we found in the long data
 
@@
 
+```{r dealing with the 99s in the wide dataset}
+#LE talked to Hillary and we found that of the 85 rows that have sums that do not = 100, 49 of them are because a "-1" was accidentally added to the no data column, even though data was taken.
+(100 percent cover - 1 = 99).
+This happened only in 2017.
+Hillary will fix this eventually on EDI, but for now I am going to manually remove these -1 from the data set so we can continue with our analyses
+
+#show this error in the data
+
+no_data_mistakes <- mistakes_wide_quad_level %>%
+  filter(row_sums == 99)
+
+unique(no_data_mistakes$`No data`)
+
+#remove the -1 from the No data column
+
+comb_wide_quad_sum$`No data`[comb_wide_quad_sum$row_sums == 99 & comb_wide_quad_sum$`No data` == -1] <- 0
+
+#check to see if it worked
+
+cleaned_sumrows99 <- comb_wide_quad_sum %>%
+  filter(row_sums == 99)
+
+unique(cleaned_sumrows99$`No data`)
+
+#yay it works
+
+```
+
+#of our 85 quads that do not equal 100, 49 have been accounted for.
+23 of them are from 2020, either Backreef or Fringing, which we know were averaged.
+This leaves only 13 quads with questions remaining.
+
+```{r make new row sums column}
+#need to update row sum now that fixes have been made
+comb_wide_quad_sum_final <- comb_wide_quad_sum %>%
+  dplyr::rename(row_sum_old = row_sums) %>%
+  mutate(row_sums = rowSums(dplyr::select(., 8:91))) %>%
+  select(-row_sum_old)
+
+#confirm 99 is gone 
+unique(comb_wide_quad_sum_final$row_sums)
+
+#whooo
+
+```
+
 ```{r filter the wide dataset}
@@
 #also want to filter out rows where the sum does not = 0 
-wide_alldata_filtered <- comb_wide_quad_sum %>%
+wide_alldata_filtered <- comb_wide_quad_sum_final %>%
   dplyr::filter(Year != 2005 & Year != 2006) %>%
@@
 
+#should be 19750 obs
+
+#writing one file so column names will work with Noam's code
 write.csv(wide_alldata_filtered, "allbenthicdata_cleaned.csv", row.names = FALSE)
+
+#clean names so column names are not so hard to work with
+
+wide_alldata_filtered_cleaned_column_names <- 
+  wide_alldata_filtered %>%
+  janitor::clean_names()
+
+write.csv(wide_alldata_filtered_cleaned_column_names, "allbenthicdata_cleaned_colnames.csv", row.names = FALSE)
 
@@
 write.csv(wide_algaeonly_filtered, "algaeonly_cleaned.csv", row.names = FALSE)
+
+#clean names so column names are not so hard to work with
+
+wide_algaeonly_cleaned_column_names <- 
+  wide_algaeonly_filtered %>%
+  janitor::clean_names()
+
+write.csv(wide_algaeonly_cleaned_column_names, "algaeonly_cleaned_colnames.csv", row.names = FALSE)
 

### 2024-07-03 julianna_data_exploration.Rmd
Getting rid of my rogue R proj which I think(?) was a mistake
Checking pull status
Smol amount of code

@@
 # load packages
-require(here)
-require(tidyverse)
+require(here) # relative file paths
+require(tidyverse) # data wrangling
+require(vegan) # community/diversity analyses
 ```
@@
 
+These are the most recent file versions done by Lauren as of July 2, 2024
+
 ```{r}
+allbenthos <- read_csv(here("data", "allbenthicdata_cleaned_colnames.csv")) 
+algae <- read_csv(here("data", "algaeonly_cleaned_colnames.csv"))
+```
 
+# Calculate algal diversity metrics
+
+Make a spp matrix to use with vegan
+
+```{r}
+# matrix of just species and their relative percent covers
+algae %>% 
+  select(-year, -location, -site, -habitat, -site_habitat, -transect, -quadrat) -> sppMat
+
+# metadata info that is in the same order (can re-join after)
+algae %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat) -> sppMeta
 ```
+
+```{r}
+diversity(algae)
+```
+
+# Join algae and coral data
+
+```{r}
+allbenthos %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, coral) %>% 
+  full_join(algae) -> coralAlg
+```

### 2024-07-03 mcr_algal_data_cleaning_final.Rmd
adding in QA/QC step
Julianna reviewed cleaning code--looks great! :)

@@
 output: html_document
+editor_options: 
+  chunk_output_type: console
 ---
@@
 rm(list=ls()) # clean up
+library(tidyverse)
 library(knitr)
@@
 opts_knit$set(eval.after = "fig.cap") 
+library(here)
 
@@
 ```{r data cleaning}
-library(here)
+data <- read.csv(here("data", "MCR_LTER_Annual_Survey_Benthic_Cover_20231211.csv"), stringsAsFactors = F) # replacements don't work when strings are factors
 
-data <- read.csv(here("data", "MCR_LTER_Annual_Survey_Benthic_Cover_20231211.csv"), stringsAsFactors = F) # replacements don't work when strings are factors
 str(data)
+
 levels(as.factor(data$Taxonomy_Substrate_Functional_Group)) # 86 levels
@@
 
+# can see that there are 99 observations that do not equal 100
+dim(non100_quad_level)
+
+# can look at which ones have no data 
 data %>% 
   filter(Taxonomy_Substrate_Functional_Group == "No data")
-
-#99 observations that do not equal 100
 
@@
 
+Just to check, Julianna is going to try this a different way (no mean function--that shouldn't be necessary, right?). Can delete this next chunk if you want:
+
+```{r}
+data %>% 
+  # doing this Date removal because there is actually one that has multiple dates for the quadrat... must be a typo? If you don't filter date, get an extra row
+  select(-Date) %>% 
+  pivot_wider(names_from = Taxonomy_Substrate_Functional_Group, 
+              values_from = Percent_Cover) %>% 
+  replace(is.na(.), 0) -> testWide
+
+# check
+summary(testWide) 
+summary(comb_wide) # eyeball perusal suggests they look identical
+```
+
 ```{r QAQC for the wide dataset}
@@
 #do we have data for all transects?
+
+comb_wide_quad_sum %>%
+  dplyr::group_by(Year, Site, Habitat) %>%
+  dplyr::summarise(count = n()) # 50 quadrats per habitat/year/site (this is correct)
 
@@
 comb_wide_quad_sum %>%
-  dplyr::filter(row_sums == -1)
+  dplyr::filter(row_sums == -1) %>% dim()
 
@@
 
+# grab the mistake quadrats 
 mistakes_wide_quad_level <- comb_wide_quad_sum %>%
@@
 comb_wide_quad_sum %>%
- dplyr::filter(row_sums != 100) %>% dplyr::filter(Year == 2020)
+  dplyr::filter(row_sums != 100) %>% 
+  dplyr::filter(Year == 2020) %>% dim()
 
@@
 #should be 19750 obs
+dim(wide_alldata_filtered)
 
@@
   janitor::clean_names()
+
+# Julianna is adding this code here IF we want it:
+wide_alldata_filtered_cleaned_column_names %>% 
+  mutate(location = gsub(" ", "_", str_to_lower(location)),
+         site = gsub(" ", "_", str_to_lower(site)),
+         habitat = gsub(" ", "_", str_to_lower(habitat)),
+         site_habitat = gsub(" ", "_", str_to_lower(site_habitat)),
+         ) # would want to save this then do it for the algae dataset as well
 
@@
   dplyr::summarise(count = n())
+
+#should be 200 per year/site combo
 

### 2024-07-03 noam_data_exploration_v3.Rmd
Added "dplyr::" before summarise so code works consistently

@@
   dplyr::group_by(Year, Site, Habitat, Transect, Taxa) %>%
-  summarise(Mean_cover = mean(Percent_cover), 
+  dplyr::summarise(Mean_cover = mean(Percent_cover), 
             SE = sd(Percent_cover)/sqrt(n()))
@@
   dplyr::group_by(Year, Site, Habitat, Taxa) %>%
-  summarise(Mean_site_cover = mean(Mean_cover), 
+  dplyr::summarise(Mean_site_cover = mean(Mean_cover), 
             Site_se = sd(Mean_cover)/sqrt(n()))
@@
   dplyr::group_by(Year, Site, Habitat, Site_habitat, Transect) %>%
-  summarize(across(all_of(species_cols), mean, na.rm = TRUE))
+  dplyr::summarize(across(all_of(species_cols), mean, na.rm = TRUE))
 
@@
   dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
-  summarise(Richness_mean = mean(Richness), 
+  dplyr::summarise(Richness_mean = mean(Richness), 
             Richness_se = sd(Richness)/sqrt(n()),
@@
   dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
-  summarise(Distance_mean = mean(Distance), 
+  dplyr::summarise(Distance_mean = mean(Distance), 
             Distance_se = sd(Distance)/sqrt(n()))
@@
   dplyr::group_by(Site, Habitat, Site_habitat, Transect) %>%
-  summarise(Richness_mean = mean(Richness), 
+  dplyr::summarise(Richness_mean = mean(Richness), 
             Richness_se = sd(Richness)/sqrt(n()),
@@
   dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
-  summarise(Turnover_mean = mean(total), 
+  dplyr::summarise(Turnover_mean = mean(total), 
             Turnover_se = sd(total)/sqrt(n()))
@@
   dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
-  summarise(Appearance_mean = mean(appearance), 
+  dplyr::summarise(Appearance_mean = mean(appearance), 
             Appearance_se = sd(appearance)/sqrt(n()))
@@
   dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
-  summarise(Disappearance_mean = mean(disappearance), 
+  dplyr::summarise(Disappearance_mean = mean(disappearance), 
             Disappearance_se = sd(disappearance)/sqrt(n()))
@@
   dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
-  summarise(Rank_shift_mean = mean(MRS), 
+  dplyr::summarise(Rank_shift_mean = mean(MRS), 
             Rank_shift_se = sd(MRS)/sqrt(n()))

### 2024-07-03 julianna_data_exploration.Rmd
Quick test
One more test
Will the CSVs make it? Time will tell!
Deleted large CSVs
Exploratory diversity analyses with coral cover

--- 2024-07-02 julianna_data_exploration.Rmd
@@
 ---
-title: "Pre working group data exploration: Julianna"
+title: "Pre working group data exploration: Mainly wrt coral cover"
 output: html_document
@@
 
+# clean up
+rm(list=ls())
+
 # load packages
-require(here)
-require(tidyverse)
+require(here) # relative file paths
+require(tidyverse) # data wrangling
+require(vegan) # community/diversity analyses
+require(lme4) # glmer()
+require(emmeans) # emmeans()
 ```
@@
 
+These are the most recent file versions done by Lauren as of July 2, 2024
+
 ```{r}
+allbenthos <- read_csv(here("data", "allbenthicdata_cleaned_colnames.csv")) 
+algae <- read_csv(here("data", "algaeonly_cleaned_colnames.csv"))
+```
+
+# Algal alpha diversity
+
+Make a spp matrix to use with vegan
+
+```{r}
+# matrix of just species and their relative percent covers in the same order as the algae tibble
+algae %>% 
+  select(-year, -location, -site, -habitat, -site_habitat, -transect, -quadrat) -> sppMat
+```
+
+```{r}
+algae %>% 
+  # grab the metadata info that is in the same order (can re-join after)
+  select(year, location, site, habitat, site_habitat, transect, quadrat) %>% 
+  mutate(Richness = specnumber(sppMat),
+         Shannon = diversity(sppMat, "shannon"),
+         Richness = specnumber(sppMat)) -> algDiv
+```
+
+## Plot
+
+Broken out by year
+
+```{r}
+algDiv %>% 
+  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
+  ggplot(aes(x = habitat, y = Shannon, group = habitat)) +
+  geom_boxplot() +
+  facet_grid(site~year) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
+```
+
+Or without:
+
+```{r}
+algDiv %>% 
+  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
+  ggplot(aes(x = habitat, y = Shannon, group = habitat)) +
+  geom_boxplot() +
+  facet_wrap(~site) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust=1))
+```
+
+Even simpler:
+
+```{r}
+# shannon
+algDiv %>% 
+  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
+  ggplot(aes(x = habitat, y = Shannon, group = habitat)) +
+  geom_boxplot() +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust=1))
+
+# richness
+algDiv %>% 
+  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
+  ggplot(aes(x = Richness, group = habitat, fill = habitat)) +
+  geom_histogram() +
+  theme_bw() +
+  facet_wrap(~habitat) +
+  theme(axis.text.x = element_text(angle = 45, hjust=1))
 
 ```
+
+Richness on top of each other
+
+```{r}
+algDiv %>% 
+  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
+  ggplot(aes(x = Richness, group = habitat, fill = habitat)) +
+  geom_bar(position = "dodge") +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust=1))
+```
+
+# Compare algal alpha diversity and coral cover
+
+Make a joined dataframe for comparisons
+
+```{r}
+allbenthos %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, coral) %>% 
+  full_join(algDiv) %>% 
+  # add column for each transect ID
+  rowwise() %>% 
+  mutate(transect_id = paste0(site, habitat, year, transect)) -> coralAlgDiv  
+```
+
+## Plot shannon
+
+Plot Shannon diversity across sites and habitats--what's up with this structure? Emailed Jamie and asked Lauren. TBD
+
+```{r}
+coralAlgDiv %>% 
+  ggplot(aes(x = coral, y = Shannon, color = year)) +
+  geom_point() + 
+  facet_grid(habitat~site) +
+  theme_classic()
+```
+
+Try by year:
+
+```{r}
+coralAlgDiv %>% 
+  filter(habitat == "backreef") %>% 
+  ggplot(aes(x = coral, y = Shannon, color = as.factor(transect))) +
+  geom_point() + 
+  facet_grid(year~site) +
+  theme_classic()
+```
+
+## Plot richness
+
+Plot richness across habitats and years
+
+As points:
+
+```{r}
+coralAlgDiv %>% 
+  ggplot(aes(x = coral, y = Richness, color = year, group = Richness)) +
+  geom_jitter() + 
+  facet_grid(habitat~site) +
+  theme_classic()
+```
+
+```{r}
+coralAlgDiv %>% 
+  ggplot(aes(x = Richness, y = coral, color = year, group = Richness)) +
+  geom_boxplot() + 
+  facet_grid(year ~ habitat) +
+  theme_classic()
+```
+
+Plot richness across sites and habitats
+
+```{r}
+coralAlgDiv %>% 
+  ggplot(aes(x = Richness, y = coral, color = year, group = Richness)) +
+  geom_boxplot() + 
+  facet_grid(site ~ habitat) +
+  theme_classic()
+```
+
+## Make a bad richness model
+
+Dumb model--need to think more about random effects and figure out time component, but instructive to see there may be a relationship
+
+```{r}
+m1 <- glmer(Richness ~ scale(coral) + habitat + (1|site) + (1|transect_id) + (1|year), 
+            family = poisson(link = "log"), data = coralAlgDiv)
+  plot(m1)
+  summary(m1)
+  emmeans(m1, ~ habitat)
+```
+
+# Beta diversity
+
+## Distance matrix
+
+We get an error message if we run vegdist right away saying that we have empty rows and that their dissimilarities may be meaningless. Let's look to see if any rows have no algae at all:
+
+```{r}
+sppMat %>% 
+  mutate(AlgSum = rowSums(across(where(is.numeric)))) %>% 
+  filter(AlgSum == 0) %>% dim() # 672 with no algae at all
+```
+
+Need to run this without zero observations, so we'll exclude quadrats with no algae and make a species matrix
+
+```{r}
+# spp matrix
+algae %>% 
+  mutate(AlgSum = rowSums(across(where(is.numeric)))) %>% 
+  filter(AlgSum != 0) %>% 
+  select(-year, -location, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgSum) -> sppMat0
+
+# metadata matrix
+algae %>% 
+  mutate(AlgSum = rowSums(across(where(is.numeric)))) %>% 
+  filter(AlgSum != 0) %>%
+  select(year, location, site, habitat, site_habitat, transect, quadrat) -> sppMeta0
+
+```
+
+Calculate Bray Curtis distances
+
+```{r}
+brayMat <- vegdist(sppMat0, "bray") # 1 is maximum dissimilarity
+```
+
+Calculate dispersion (this is using Noam's code)
+
+```{r}
+siteHabBetaDisper <- vegan::betadisper(bray_matrix, sppMeta0$site_habitat)
+
+# plot of dispersion test
+plot(year_beta_disper, sub = "") # lol
+boxplot(year_beta_disper, main = "Year", xlab = "") 
+```

### 2024-07-04 julianna_data_exploration.Rmd
Did a bunch of exploration to figure out what the bottom bound structure is in the Shannon algal vs. coral/turf/algae plots. Believe it has to do with instances when you have one dominant thing (e.g., turf algae) and then increase the proportion of a second alga incrementally at 4% intervals

@@
 allbenthos <- read_csv(here("data", "allbenthicdata_cleaned_colnames.csv")) 
-algae <- read_csv(here("data", "algaeonly_cleaned_colnames.csv"))
+algae <- read_csv(here("data", "algaeonly_cleaned_colnames.csv")) %>% 
+  # going to run everything without algal turf
+  select(-algal_turf)
 ```
@@
          Shannon = diversity(sppMat, "shannon"),
+        # Evenness = diversity(sppMat, "ev")
          Richness = specnumber(sppMat)) -> algDiv
@@
 
+# Compare algal and coral cover
+
+```{r}
+# get ranges to calculate over
+startAlg <- grep("acanthophora_spicifera", colnames(algae))
+endAlg <- grep("valonia_ventricosa", colnames(algae))
+
+algae %>% 
+  mutate(AlgSum = rowSums(algae[startAlg:endAlg])) -> algSum
+
+algSum %>% 
+  full_join(allbenthos) %>% 
+  ggplot(aes(x = coral, y = AlgSum, color = year)) +
+  geom_point() +
+  facet_grid(site~habitat)
+
+```
+
 # Compare algal alpha diversity and coral cover
@@
   rowwise() %>% 
-  mutate(transect_id = paste0(site, habitat, year, transect)) -> coralAlgDiv  
+  mutate(transect_id = paste0(site, habitat, "_", year, "_", transect)) -> coralAlgDiv  
 ```
 
-## Plot shannon
+## Plot Shannon
 
@@
 coralAlgDiv %>% 
-  ggplot(aes(x = coral, y = Shannon, color = year)) +
-  geom_point() + 
+  ggplot(aes(x = coral, y = Shannon, color = as.factor(Richness))) +
+  geom_point(alpha = 0.5) + 
   facet_grid(habitat~site) +
-  theme_classic()
+  theme_classic() + 
+  geom_hline(yintercept = 0.693, color = "red", linetype = "dashed") # this is when two species are equally distributed (half half)--not sure why that would be
 ```
@@
   theme_classic()
+```
+
+Or look at Shannon vs. the percent cover of algae in the plot
+
+```{r}
+coralAlgDiv %>% 
+  full_join(algSum) %>% 
+  ggplot(aes(x = AlgSum, y = Shannon, color = as.factor(Richness))) +
+  geom_point(alpha = 0.5) + 
+  facet_grid(habitat~site) +
+  theme_classic() + 
+  geom_hline(yintercept = 0.693, color = "red", linetype = "dashed")
+```
+
+What about this when Richness == 2?
+
+```{r}
+coralAlgDiv %>% 
+  full_join(algSum) %>% view()
+  filter(Richness == 2) %>% 
+  ggplot(aes(x = AlgSum, y = Shannon, color = as.factor(Richness))) +
+  geom_point(alpha = 0.5) + 
+  facet_grid(habitat~site) +
+  theme_classic() + 
+  geom_hline(yintercept = 0.693, color = "red", linetype = "dashed")
+```
+
+Could this be because the only way to get a high enough imbalance/unevenness is having a large enough algal N
+
+What does it look like relative to turf algae?
+
+```{r}
+coralAlgDiv %>% 
+  full_join(algSum) %>% 
+  filter(Richness == 2) %>% 
+  ggplot(aes(x = algal_turf, y = Shannon, color = as.factor(Richness))) +
+  geom_point(alpha = 0.5) + 
+  facet_grid(habitat~site) +
+  theme_classic() + 
+  geom_hline(yintercept = 0.693, color = "red", linetype = "dashed")
+```
+
+Nic: are these lines all turf but 1 line, all turf but 2 points line, all turf but 3 points line?
+
+```{r}
+coralAlgDiv %>% 
+  full_join(algSum) %>% 
+  filter(Richness == 2) %>% 
+  mutate(Nonturf = AlgSum - algal_turf) %>% select(Nonturf) %>% unique() %>% view()
+  ggplot(aes(x = algal_turf, y = Shannon, color = as.factor(Nonturf))) +
+  geom_point(alpha = 0.5) + 
+  facet_grid(habitat~site) +
+  theme_classic() + 
+  geom_hline(yintercept = 0.693, color = "red", linetype = "dashed")
+```
+
+^This seems to be true here.
+It seems like lines have to do with how much of the non-turf you have.
+They only do 25 points and multiply by 4 so everything is a multiple of 4.
+But then how can we have a 2 there???
+
+## Problem solve Shannon
+
+Trying to instead calculate Shannon manually with 100 as the denominator?
+
+```{r}
+shannon100 <- function(x, equalize.groups = FALSE, base = exp(1)) {
+  x <- drop(as.matrix(x))
+  if (length(dim(x)) > 1) {
+        x <- sweep(x, MARGIN = 1, 100, FUN = "/") # MARGIN = 1 I think means to do this rowwise; FUN = "/" means to divide by 100 (replaced the row sum with 100 here)
+    }
+    else {
+        x <- x/(100 <- sum(x)) # replaced total with 100 here
+    }
+  x <- -x * log(x, base)
+  if (length(dim(x)) > 1) 
+        H <- apply(x, MARGIN = 1, sum, na.rm = TRUE)
+    else H <- sum(x, na.rm = TRUE)
+  H
+}
+
+# add this to the df
+algae %>% 
+  # grab the metadata info that is in the same order (can re-join after)
+  select(year, location, site, habitat, site_habitat, transect, quadrat) %>% 
+  mutate(Richness = specnumber(sppMat),
+         Shannon = diversity(sppMat, "shannon"),
+         Shannon100 = shannon100(sppMat),
+         Richness = specnumber(sppMat)) -> algDiv100
+```
+
+Plot
+
+```{r}
+algDiv100 %>% 
+  ggplot(aes(x = Shannon100, y = Shannon)) +
+  geom_point() +
+  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") 
+```
+
+Re-make old plot
+
+```{r}
+# re-join with coral
+allbenthos %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, coral) %>% 
+  full_join(algDiv100) %>% 
+  # add column for each transect ID
+  rowwise() %>% 
+  mutate(transect_id = paste0(site, habitat, "_", year, "_", transect)) -> coralAlgDiv100
+
+# plot  
+coralAlgDiv100 %>% 
+  ggplot(aes(x = coral, y = Shannon, color = year)) +
+  geom_point() + 
+  facet_grid(habitat~site) +
+  theme_classic() # NOPE very much still there 
 ```
@@
             family = poisson(link = "log"), data = coralAlgDiv)
-  plot(m1)
+  plot(m1) # I think this looks pretty weird right?
   summary(m1)

### 2024-07-09 julianna_data_exploration.Rmd
Did more analyses; worked towards Deron's vision for "figure 1"

--- 2024-07-03 julianna_data_exploration.Rmd
@@
 require(emmeans) # emmeans()
+require(viridis) # color palette
+require(MoMAColors) # another color palette
 ```
@@
 allbenthos <- read_csv(here("data", "allbenthicdata_cleaned_colnames.csv")) 
-algae <- read_csv(here("data", "algaeonly_cleaned_colnames.csv"))
+algae <- read_csv(here("data", "algaeonly_cleaned_colnames.csv")) %>% 
+  mutate(island_side = case_when(site == "lter_1" ~ "north",
+                                 site == "lter_2" ~ "north",
+                                 site == "lter_3" ~ "east",
+                                 site == "lter_4" ~ "east",
+                                 site == "lter_5" ~ "west",
+                                 site == "lter_6" ~ "west",
+                                 TRUE ~ "UdINGLEbUTT"))
+  # going to run everything without algal turf
+  # select(-algal_turf)
+
+macroalgae <- algae %>% 
+  select(-algal_turf, -crustose_corallines, -cyanophyta, -damselfish_turf, -lithophyllum_kotschyanum,
+         -neogoniolithon_brassica_florida, -sporolithon_sp) # got this from the cleaning code--should Peysonellia be included?
 ```
@@
 algae %>% 
-  select(-year, -location, -site, -habitat, -site_habitat, -transect, -quadrat) -> sppMat
+  select(-year, -location, -site, -island_side, -habitat, -site_habitat, -transect, -quadrat) -> sppMat
 ```
@@
   # grab the metadata info that is in the same order (can re-join after)
-  select(year, location, site, habitat, site_habitat, transect, quadrat) %>% 
+  select(year, location, site, island_side, habitat, site_habitat, transect, quadrat) %>% 
   mutate(Richness = specnumber(sppMat),
          Shannon = diversity(sppMat, "shannon"),
-         Richness = specnumber(sppMat)) -> algDiv
+         Evenness = Shannon/log(Richness), # I'm pretty sure this is right but double check (Pielou's evenness)
+         Richness = specnumber(sppMat)) %>% 
+  # clean for graphing
+  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17")),
+         island_side = factor(island_side, levels = c("north", "east", "west"))) -> algDiv
 ```
@@
   theme(axis.text.x = element_text(angle = 45, hjust=1))
+```
+
+Island side:
+
+```{r}
+algDiv %>% 
+  ggplot(aes(x = island_side, y = Shannon, group = island_side, fill = island_side)) +
+  geom_boxplot() +
+  facet_wrap(~habitat) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust=1))
+
 ```
@@
 
+Broken up by island side
+
+```{r}
+algDiv %>% 
+  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
+  ggplot(aes(x = Richness, group = habitat, fill = habitat)) +
+  geom_bar(position = "dodge") +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+  scale_fill_viridis_d() +
+  facet_wrap(~island_side)
+```
+
+# Compare algal and coral cover
+
+```{r}
+# get ranges to calculate over
+startAlg <- grep("acanthophora_spicifera", colnames(algae))
+endAlg <- grep("valonia_ventricosa", colnames(algae))
+
+algae %>% 
+  mutate(AlgSum = rowSums(algae[startAlg:endAlg])) -> algSum
+
+algSum %>% 
+  full_join(allbenthos) %>% 
+  ggplot(aes(x = coral, y = AlgSum, color = year)) +
+  geom_point() +
+  facet_grid(site~habitat)
+
+```
+
 # Compare algal alpha diversity and coral cover
@@
   rowwise() %>% 
-  mutate(transect_id = paste0(site, habitat, year, transect)) -> coralAlgDiv  
+  mutate(transect_id = paste0(site, habitat, "_", year, "_", transect)) %>% 
+  # clean for graphing
+  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17")),
+         island_side = factor(island_side, levels = c("north", "east", "west"))) -> coralAlgDiv  
 ```
 
-## Plot shannon
+## Plot Shannon
 
-Plot Shannon diversity across sites and habitats--what's up with this structure? Emailed Jamie and asked Lauren. TBD
+Plot Shannon diversity across sites and habitats--what's up with this structure?
 
@@
 ```
+
+Or look at Shannon vs. the percent cover of algae in the plot
+
+```{r}
+coralAlgDiv %>% 
+  full_join(algSum) %>% 
+  ggplot(aes(x = AlgSum, y = Shannon, color = as.factor(Richness))) +
+  geom_point(alpha = 0.5) + 
+  facet_grid(habitat~site) +
+  theme_classic() + 
+  geom_hline(yintercept = 0.693, color = "red", linetype = "dashed")
+```
+
+What about this when Richness == 2?
+
+```{r}
+coralAlgDiv %>% 
+  full_join(algSum) %>% 
+  filter(Richness == 2) %>% 
+  ggplot(aes(x = AlgSum, y = Shannon, color = as.factor(Richness))) +
+  geom_point(alpha = 0.5) + 
+  facet_grid(habitat~site) +
+  theme_classic() + 
+  geom_hline(yintercept = 0.693, color = "red", linetype = "dashed")
+```
+
+Could this be because the only way to get a high enough imbalance/unevenness is having a large enough algal N
+
+What does it look like relative to turf algae?
+
+```{r}
+coralAlgDiv %>% 
+  full_join(algSum) %>% 
+  filter(Richness == 2) %>% 
+  ggplot(aes(x = algal_turf, y = Shannon, color = as.factor(Richness))) +
+  geom_point(alpha = 0.5) + 
+  facet_grid(habitat~site) +
+  theme_classic() + 
+  geom_hline(yintercept = 0.693, color = "red", linetype = "dashed")
+```
+
+Nic: are these lines all turf but 1 line, all turf but 2 points line, all turf but 3 points line?
+
+```{r}
+coralAlgDiv %>% 
+  full_join(algSum) %>% 
+  filter(Richness == 2) %>% 
+  mutate(Nonturf = AlgSum - algal_turf) %>% 
+  ggplot(aes(x = AlgSum, y = Shannon, color = as.factor(Nonturf))) +
+  geom_point(alpha = 0.5) + 
+  facet_grid(habitat~site) +
+  theme_classic() + 
+  geom_hline(yintercept = 0.693, color = "red", linetype = "dashed")
+```
+
+^This seems to be true here.
+It seems like lines have to do with how much of the non-turf you have.
+They only do 25 points and multiply by 4 so everything is a multiple of 4.
+But then how can we have a 2 there???
+Answer: 2020 averaging
 
@@
 coralAlgDiv %>% 
-  ggplot(aes(x = Richness, y = coral, color = year, group = Richness)) +
+  ggplot(aes(x = Richness, y = coral, fill = habitat, group = Richness)) +
   geom_boxplot() + 
-  facet_grid(site ~ habitat) +
+  facet_grid(island_side ~ habitat) +
+  scale_fill_viridis_d() +
   theme_classic()
@@
 ```{r}
-m1 <- glmer(Richness ~ scale(coral) + habitat + (1|site) + (1|transect_id) + (1|year), 
+m1 <- glmer(Richness ~ scale(coral) + habitat + island_side + (1|site) + (1|transect_id) + (1|year), 
             family = poisson(link = "log"), data = coralAlgDiv)
-  plot(m1)
+  plot(m1) # I think this looks pretty weird right?
   summary(m1)
   emmeans(m1, ~ habitat)
+```
+
+## Plot evenness
+
+Plot evennes across habitats and years
+
+As points:
+
+```{r}
+coralAlgDiv %>% 
+  ggplot(aes(x = coral, y = Evenness, color = year, group = Evenness)) +
+  geom_jitter() + 
+  facet_grid(habitat~site) +
+  theme_classic()
+```
+
+# Plot changes of algae, coral, etc. through time
+
+```{r make a big joined tibble}
+startMAlg <- grep("acanthophora_spicifera", colnames(macroalgae))
+endMAlg <- grep("valonia_ventricosa", colnames(macroalgae))
+
+macroalgae %>% 
+  mutate(macroalgae_sum = rowSums(macroalgae[startMAlg:endMAlg])) %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side, macroalgae_sum) %>% 
+  full_join(coralAlgDiv) %>% 
+  full_join(algae) %>% 
+  rowwise() %>% 
+  mutate(turf_sum = sum(algal_turf, damselfish_turf),
+         CCA_sum = sum(lithophyllum_kotschyanum, crustose_corallines, neogoniolithon_brassica_florida, sporolithon_sp)) %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side, turf_sum, CCA_sum, coral, macroalgae_sum) %>% 
+  # pivot key columns longer
+  pivot_longer(cols = c(turf_sum, CCA_sum, coral, macroalgae_sum),
+               names_to = "substrate",
+               values_to = "sum_cover") %>% 
+  # summarize and group quadrats by transects
+  group_by(year, site, habitat, site_habitat, transect, island_side, substrate) %>% 
+  summarise(avg_cov = mean(sum_cover)) -> transect4mean
+
+```
+
+## Quadrat level island-habitat
+
+```{r make a df for island side/habitat}
+macroalgae %>% 
+  mutate(macroalgae_sum = rowSums(macroalgae[startMAlg:endMAlg])) %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side, macroalgae_sum) %>% 
+  full_join(coralAlgDiv) %>% 
+  full_join(algae) %>% 
+  rowwise() %>% 
+  mutate(turf_sum = sum(algal_turf, damselfish_turf),
+         CCA_sum = sum(lithophyllum_kotschyanum, crustose_corallines, neogoniolithon_brassica_florida, sporolithon_sp)) %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side, turf_sum, CCA_sum, coral, macroalgae_sum) %>% 
+  # pivot key columns longer
+  pivot_longer(cols = c(turf_sum, CCA_sum, coral, macroalgae_sum),
+               names_to = "substrate",
+               values_to = "sum_cover") %>% 
+  # summarize and group quadrats by island/habitat
+  group_by(year, habitat, island_side, substrate) %>% 
+  summarise(mean_cov = mean(sum_cover),
+            sd_cov = sd(sum_cover, na.rm = TRUE),
+            n = n(),
+            CI95 = 1.96*(sd_cov)/sqrt(n)) %>% 
+  # relevel for graphing
+  mutate(island_side = factor(island_side, levels = c("north", "east", "west")),
+         habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) -> quadHabIS
+
+```
+
+```{r plot cover through time}
+quadHabIS %>% 
+  ggplot(aes(x = year, y = mean_cov, group = substrate, fill = substrate)) +
+  geom_ribbon(aes(ymin = mean_cov -CI95, 
+                    ymax = mean_cov + CI95),
+              alpha  = 0.7) +
+  geom_line() +
+  theme_bw() +
+  scale_fill_manual(values = c(
+    moma.colors("Klein", type = "discrete")[2], # CCA
+    moma.colors("Klein", type = "discrete")[1], # coral
+    moma.colors("Klein", type = "discrete")[5], # macro
+    moma.colors("Klein", type = "discrete")[4] # turf
+  )) +
+  facet_grid(island_side ~ habitat) +
+  ylab("Mean cover (%)") +
+  xlab("Year")
+```
+
+## Transect-level island/habitat
+
+```{r make a df for island side/habitat}
+transect4mean %>% 
+  group_by(year, habitat, island_side, substrate) %>% 
+  summarise(mean_cov = mean(avg_cov),
+            sd_cov = sd(avg_cov, na.rm = TRUE),
+            n = n(),
+            CI95 = 1.96*(sd_cov)/sqrt(n)) %>% 
+  # relevel for graphing
+  mutate(island_side = factor(island_side, levels = c("north", "east", "west")),
+         habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) -> transectHabIS
+
+```
+
+```{r plot cover through time}
+# this color palette:
+#moma.colors("Abbott", type = "discrete")
+
+transectHabIS %>% 
+  ggplot(aes(x = year, y = mean_cov, group = substrate, fill = substrate)) +
+  geom_ribbon(aes(ymin = mean_cov -CI95, 
+                    ymax = mean_cov + CI95),
+              alpha  = 0.7) +
+  geom_line() +
+  theme_bw() +
+  scale_fill_manual(values = c(
+    moma.colors("Klein", type = "discrete")[2], # CCA
+    moma.colors("Klein", type = "discrete")[1], # coral
+    moma.colors("Klein", type = "discrete")[5], # macro
+    moma.colors("Klein", type = "discrete")[4] # turf
+  )) +
+  facet_grid(island_side ~ habitat) +
+  ylab("Mean cover (%)") +
+  xlab("Year")
+```
+
+## Transect-level habitat
+
+```{r make a df for island side/habitat}
+transect4mean %>% 
+  group_by(year, habitat, substrate) %>% 
+  summarise(mean_cov = mean(avg_cov),
+            sd_cov = sd(avg_cov, na.rm = TRUE),
+            n = n(),
+            CI95 = 1.96*(sd_cov)/sqrt(n)) %>% 
+  # relevel for graphing
+  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) -> transectHab
+
+```
+
+```{r plot cover through time}
+# this color palette:
+#moma.colors("Abbott", type = "discrete")
+
+transectHab %>% 
+  ggplot(aes(x = year, y = mean_cov, group = substrate, fill = substrate)) +
+  geom_ribbon(aes(ymin = mean_cov -CI95, 
+                    ymax = mean_cov + CI95),
+              alpha  = 0.7) +
+  geom_line() +
+  theme_bw() +
+  scale_fill_manual(values = c(
+    moma.colors("Klein", type = "discrete")[2], # CCA
+    moma.colors("Klein", type = "discrete")[1], # coral
+    moma.colors("Klein", type = "discrete")[5], # macro
+    moma.colors("Klein", type = "discrete")[4] # turf
+  )) +
+  facet_wrap(~habitat) +
+  ylab("Mean cover (%)") +
+  xlab("Year")
+```
+
+## Transect level island side
+
+```{r make a df for island side}
+transect4mean %>% 
+  group_by(year, island_side, substrate) %>% 
+  summarise(mean_cov = mean(avg_cov),
+            sd_cov = sd(avg_cov, na.rm = TRUE),
+            n = n(),
+            CI95 = 1.96*(sd_cov)/sqrt(n)) %>% 
+  # relevel for graphing
+  mutate(island_side = factor(island_side, levels = c("north", "east", "west"))) -> transectIS
+
+```
+
+```{r plot cover through time}
+
+transectIS %>% 
+  ggplot(aes(x = year, y = mean_cov, group = substrate, fill = substrate)) +
+  geom_ribbon(aes(ymin = mean_cov -CI95, 
+                    ymax = mean_cov + CI95),
+              alpha  = 0.7) +
+  geom_line() +
+  theme_bw() +
+  scale_fill_manual(values = c(
+    moma.colors("Klein", type = "discrete")[2], # CCA
+    moma.colors("Klein", type = "discrete")[1], # coral
+    moma.colors("Klein", type = "discrete")[5], # macro
+    moma.colors("Klein", type = "discrete")[4] # turf
+  )) +
+  facet_grid(~island_side) +
+  ylab("Mean cover (%)") +
+  xlab("Year")
+```
+
+# Examine most abundant algae
+
+Reformat so that the data are easier to summarize:
+
+```{r}
+# get ranges to calculate over
+startAlg <- grep("acanthophora_spicifera", colnames(algae))
+endAlg <- grep("valonia_ventricosa", colnames(algae))
+
+algae %>% 
+  pivot_longer(cols = colnames(algae)[startAlg:endAlg],
+               names_to = "taxa",
+               values_to = "percent") -> algaeLong
+```
+
+Get some summary statistics
+
+## Across all time, by habitat:
+
+```{r}
+algaeLong %>% 
+  group_by(site, habitat, year, transect, island_side, taxa) %>% # get transect summaries
+  summarize(Mean_cov = mean(percent), 
+            Num_occur = n(), 
+            sd_cov = sd(percent)) -> algTaxTrans
+
+```
+
+Plot
+
+```{r}
+algTaxTrans %>% 
+  ggplot(aes(x = reorder(taxa, -Mean_cov), y = Mean_cov, fill = habitat)) +
+  geom_boxplot() + 
+  scale_fill_viridis_d() +
+  facet_wrap(~habitat) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+With just macroalgae:
+
+```{r}
+algTaxTrans %>% 
+  filter(!
+taxa %in% c("algal_turf", "crustose_corallines", "cyanophyta", "damselfish_turf", "lithophyllum_kotschyanum", "neogoniolithon_brassica_florida", "sporolithon_sp" )) %>% # this list taken from the cleaning code
+  ggplot(aes(x = reorder(taxa, -Mean_cov), y = Mean_cov, fill = habitat)) +
+  geom_boxplot() + 
+  scale_fill_viridis_d() +
+  facet_wrap(~habitat) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+Not by habitat
+
+```{r}
+algTaxTrans %>% 
+  filter(!
+taxa %in% c("algal_turf", "crustose_corallines", "cyanophyta", "damselfish_turf", "lithophyllum_kotschyanum", "neogoniolithon_brassica_florida", "sporolithon_sp" )) %>% # this list taken from the cleaning code
+  ggplot(aes(x = reorder(taxa, -Mean_cov), y = Mean_cov)) +
+  geom_boxplot() + 
+  scale_fill_viridis_d() +
+  #facet_wrap(~habitat) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 ```
@@
 ```{r}
+startAlg <- grep("acanthophora_spicifera", colnames(algae))
+endAlg <- grep("valonia_ventricosa", colnames(algae))
+
 # spp matrix
 algae %>% 
-  mutate(AlgSum = rowSums(across(where(is.numeric)))) %>% 
+  mutate(AlgSum = rowSums(algae[startAlg:endAlg])) %>% 
   filter(AlgSum != 0) %>% 
-  select(-year, -location, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgSum) -> sppMat0
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgSum) -> sppMat0
 
@@
 algae %>% 
-  mutate(AlgSum = rowSums(across(where(is.numeric)))) %>% 
+  mutate(AlgSum = rowSums(algae[startAlg:endAlg])) %>% 
   filter(AlgSum != 0) %>%
-  select(year, location, site, habitat, site_habitat, transect, quadrat) -> sppMeta0
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
+  # make a year-habitat column for betadisper
+  mutate(hab_year = paste0(habitat, "_", year)) -> sppMeta0
 
@@
 
+- Talking with Lauren, decided want to be smart about which groupings we do for `betadisper()`. Priorities are:
+  1. Year-habitat as the group: include turf/cyano/everything
+  2. Island side - year - habitat
+
 ```{r}
-siteHabBetaDisper <- vegan::betadisper(bray_matrix, sppMeta0$site_habitat)
+habBetaDisper <- vegan::betadisper(brayMat, sppMeta0$hab_year)
 
 # plot of dispersion test
-plot(year_beta_disper, sub = "") # lol
-boxplot(year_beta_disper, main = "Year", xlab = "") 
+plot(habBetaDisper, sub = "") # lol
+boxplot(habBetaDisper, main = "Habitat", xlab = "") 
 ```

### 2024-07-10 julianna_data_exploration.Rmd
Started playing around with SACs

@@
 require(MoMAColors) # another color palette
+require(patchwork) # putting figures together
+require(BiodiversityR)
+require(ggsci) # scientific jounral and sci-fi themed color palette
 ```
@@
   select(-year, -location, -site, -island_side, -habitat, -site_habitat, -transect, -quadrat) -> sppMat
+
+# metadata matrix
+algae %>% 
+  mutate(AlgSum = rowSums(algae[startAlg:endAlg])) %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
+  # make a year-habitat column for betadisper
+  mutate(hab_year = paste0(habitat, "_", year),
+         island_side_hab_year = paste0(island_side, "_", hab_year)) -> sppMeta
 ```
@@
   geom_point() +
-  facet_grid(site~habitat)
+  facet_grid(site~habitat) +
+  ylab("Algal cover (%)") +
+  xlab("Coral cover (%)")
 
@@
   geom_jitter() + 
-  facet_grid(habitat~site) +
+  facet_grid(habitat~island_side) +
   theme_classic()
+```
+
+As boxplots
+
+```{r}
+coralAlgDiv %>% 
+  ggplot(aes(x = Richness, y = coral, color = island_side, group = Richness)) +
+  geom_boxplot() + 
+  facet_grid(habitat~island_side) +
+  scale_color_manual(values = c("darkblue", "darkred", "darkgreen")) +
+  theme_classic()
+
 ```
@@
 
+# Species accummulation curves
+
+## Set special theme
+
+From https://rpubs.com/Roeland-KINDT/694021
+
+```{r}
+BioR.theme <- theme(
+        panel.background = element_blank(),
+        panel.border = element_blank(),
+        panel.grid = element_blank(),
+        axis.line = element_line("gray25"),
+        text = element_text(size = 12),
+        axis.text = element_text(size = 10, colour = "gray25"),
+        axis.title = element_text(size = 14, colour = "gray25"),
+        legend.title = element_text(size = 14),
+        legend.text = element_text(size = 14),
+        legend.key = element_blank())
+```
+
+## By habitat (all time)
+
+Noam suggested doing this on July 10, 2024 in response to the richness plots. To learn how to do this, I used this tutorial: https://rpubs.com/Roeland-KINDT/694021
+
+```{r accumpcomp}
+# use accumcomp() to calculate accumulated species richness
+# conditioned = FALSE switches estimation of standard deviation to a formula based on the estimation of extrapolated species richness
+# returns data on species accumulations separately for each different level of a factor
+accum.Hab <- accumcomp(sppMat, # spp matrix
+                     y = sppMeta, # metadata df (same length)
+                     factor = 'habitat', # factor to group by in the metadata file
+    method='exact', conditioned = FALSE, plotit = FALSE)
+```
+
+To plot, use accumcomp.long() to get it into long/ggplot format
+
+```{r make long}
+accum.Hablong <- accumcomp.long(accum.Hab, ci=NA, 
+                              label.freq=2000) # lebel.freq adjusts the labelit variabel to show how often a symbol will be plotted on the SAC
+```
+
+Plot
+
+```{r plot}
+accum.Hablong %>% 
+  mutate(Grouping = factor(Grouping, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
+  ggplot(aes(x = Sites, y = Richness, ymax = UPR, ymin = LWR)) +
+  scale_x_continuous(expand=c(0, 1), sec.axis = dup_axis(labels=NULL, name=NULL)) + # this moves it so zero is right at the origin
+  scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) + # this makes it more like a box
+  geom_line(aes(colour=Grouping), linewidth=2) + # adds line
+  #geom_point(data=subset(accum.Hablong, labelit==TRUE), 
+  #             aes(colour=Grouping, shape=Grouping), size=.5) +
+  geom_ribbon(aes(colour=Grouping), alpha=0.2, show.legend=FALSE) + 
+  BioR.theme +
+  scale_colour_viridis_d(direction = -1) +
+  labs(x = "Quadrats", y = "Richness", colour = "Haitat", shape = "Haitat")
+```
+
+## By island side (all time)
+
+```{r accumpcomp}
+# use accumcomp() to calculate accumulated species richness
+# conditioned = FALSE switches estimation of standard deviation to a formula based on the estimation of extrapolated species richness
+# returns data on species accumulations separately for each different level of a factor
+accum.IS <- accumcomp(sppMat, # spp matrix
+                     y = sppMeta, # metadata df (same length)
+                     factor = 'island_side', # factor to group by in the metadata file
+    method='exact', conditioned = FALSE, plotit = FALSE)
+```
+
+To plot, use accumcomp.long() to get it into long/ggplot format
+
+```{r make long}
+accum.ISlong <- accumcomp.long(accum.IS, ci=NA, 
+                              label.freq=2000) # lebel.freq adjusts the labelit variabel to show how often a symbol will be plotted on the SAC
+```
+
+Plot
+
+```{r plot}
+accum.ISlong %>% 
+  mutate(Grouping = factor(Grouping, levels = c("north", "east", "west"))) %>% 
+  ggplot(aes(x = Sites, y = Richness, ymax = UPR, ymin = LWR)) +
+  scale_x_continuous(expand=c(0, 1), sec.axis = dup_axis(labels=NULL, name=NULL)) + # this moves it so zero is right at the origin
+  scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) + # this makes it more like a box
+  geom_line(aes(colour=Grouping), linewidth=2) + # adds line
+  #geom_point(data=subset(accum.Hablong, labelit==TRUE), 
+  #             aes(colour=Grouping, shape=Grouping), size=.5) +
+  geom_ribbon(aes(colour=Grouping), alpha=0.2, show.legend=FALSE) + 
+  BioR.theme +
+  scale_color_manual(values = c("darkblue", "darkred", "darkgreen")) +
+  labs(x = "Quadrats", y = "Richness", colour = "Haitat", shape = "Haitat")
+```
+
+## By habitat-time
+
+Noam suggested doing this on July 10, 2024 in response to the richness plots. To learn how to do this, I used this tutorial: https://rpubs.com/Roeland-KINDT/694021
+
+```{r accumpcomp}
+# use accumcomp() to calculate accumulated species richness
+# conditioned = FALSE switches estimation of standard deviation to a formula based on the estimation of extrapolated species richness
+# returns data on species accumulations separately for each different level of a factor
+accum.HabYr <- accumcomp(sppMat, # spp matrix
+                     y = sppMeta, # metadata df (same length)
+                     factor = 'hab_year', # factor to group by in the metadata file
+    method='exact', conditioned = FALSE, plotit = FALSE)
+```
+
+To plot, use accumcomp.long() to get it into long/ggplot format
+
+```{r make long}
+accum.HabYrlong <- accumcomp.long(accum.HabYr, ci=NA, 
+                              label.freq=2000) %>% # lebel.freq adjusts the labelit variabel to show how often a symbol will be plotted on the SAC
+  # get metadata back
+  rowwise() %>% 
+  mutate(Year = (str_split(Grouping, pattern = "_")[[1]])[length(str_split(Grouping, pattern = "_")[[1]])]) %>% 
+  mutate(Habitat = case_when(
+    str_split(Grouping, pattern = "_")[[1]][[2]] == "17" ~ "outer_17",
+    str_split(Grouping, pattern = "_")[[1]][[2]] == "10" ~ "outer_10",
+    str_split(Grouping, pattern = "_")[[1]][[1]] == "fringing" ~ "fringing",
+    str_split(Grouping, pattern = "_")[[1]][[1]] == "backreef" ~ "backreef",
+    TRUE ~ "Error"
+    )) %>% 
+  mutate(Habitat = factor(Habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) 
+```
+
+Plot
+
+```{r plot}
+accum.HabYrlong %>% 
+  ggplot(aes(x = Sites, y = Richness, ymax = UPR, ymin = LWR)) +
+  scale_x_continuous(expand=c(0, 1), sec.axis = dup_axis(labels=NULL, name=NULL)) + # this moves it so zero is right at the origin
+  scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) + # this makes it more like a box
+  geom_line(aes(colour=Year), linewidth=2) + # adds line
+  #geom_point(data=subset(accum.Hablong, labelit==TRUE), 
+  #             aes(colour=Grouping, shape=Grouping), size=.5) +
+  #geom_ribbon(aes(colour=Year), alpha=0.2, show.legend=FALSE) + 
+  BioR.theme +
+  scale_colour_viridis_d(direction = -1) +
+  facet_wrap(~Habitat) +
+  labs(x = "Quadrats", y = "Richness", colour = "Haitat", shape = "Haitat")
+```
+
+## By habitat-time-island side
+
+```{r accumpcomp}
+# use accumcomp() to calculate accumulated species richness
+# conditioned = FALSE switches estimation of standard deviation to a formula based on the estimation of extrapolated species richness
+# returns data on species accumulations separately for each different level of a factor
+accum.HabISYr <- accumcomp(sppMat, # spp matrix
+                     y = sppMeta, # metadata df (same length)
+                     factor = 'island_side_hab_year', # factor to group by in the metadata file
+    method='exact', conditioned = FALSE, plotit = FALSE)
+```
+
+To plot, use accumcomp.long() to get it into long/ggplot format
+
+```{r make long}
+accum.HabISYrlong <- accumcomp.long(accum.HabISYr, ci=NA, 
+                              label.freq=2000) %>% # lebel.freq adjusts the labelit variabel to show how often a symbol will be plotted on the SAC
+  # get metadata back
+  rowwise() %>% 
+  mutate(Year = (str_split(Grouping, pattern = "_")[[1]])[length(str_split(Grouping, pattern = "_")[[1]])]) %>% 
+  mutate(Habitat = case_when(
+    str_split(Grouping, pattern = "_")[[1]][[3]] == "17" ~ "outer_17",
+    str_split(Grouping, pattern = "_")[[1]][[3]] == "10" ~ "outer_10",
+    str_split(Grouping, pattern = "_")[[1]][[2]] == "fringing" ~ "fringing",
+    str_split(Grouping, pattern = "_")[[1]][[2]] == "backreef" ~ "backreef",
+    TRUE ~ "Error"
+    )) %>% 
+  mutate(island_side = str_split(Grouping, pattern = "_")[[1]][[1]]) %>% 
+  mutate(Habitat = factor(Habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) 
+```
+
+Plot
+
+```{r plot}
+accum.HabISYrlong %>% 
+  ggplot(aes(x = Sites, y = Richness, ymax = UPR, ymin = LWR)) +
+  scale_x_continuous(expand=c(0, 1), sec.axis = dup_axis(labels=NULL, name=NULL)) + # this moves it so zero is right at the origin
+  scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) + # this makes it more like a box
+  geom_line(aes(colour=Year), linewidth=1) + # adds line
+  #geom_point(data=subset(accum.Hablong, labelit==TRUE), 
+  #             aes(colour=Grouping, shape=Grouping), size=.5) +
+  #geom_ribbon(aes(colour=Year), alpha=0.2, show.legend=FALSE) + 
+  BioR.theme +
+  scale_colour_viridis_d(direction = -1) +
+  facet_grid(island_side~Habitat) +
+  labs(x = "Quadrats", y = "Richness")
+```
+
 # Plot changes of algae, coral, etc. through time
@@
                names_to = "substrate",
-               values_to = "sum_cover") %>% 
+               values_to = "sum_cover") -> quad4sum
+
+quad4sum %>% 
   # summarize and group quadrats by island/habitat
@@
   xlab("Year")
+```
+
+## Quadrat-level habitat
+
+```{r make a df for island side/habitat}
+quad4sum %>% 
+  # summarize and group quadrats by island/habitat
+  group_by(year, habitat, substrate) %>% 
+  summarise(mean_cov = mean(sum_cover),
+            sd_cov = sd(sum_cover, na.rm = TRUE),
+            n = n(),
+            CI95 = 1.96*(sd_cov)/sqrt(n)) %>% 
+  # relevel for graphing
+  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
+  ggplot(aes(x = year, y = mean_cov, group = substrate, fill = substrate)) +
+  geom_ribbon(aes(ymin = mean_cov -CI95, 
+                    ymax = mean_cov + CI95),
+              alpha  = 0.7) +
+  geom_line() +
+  theme_bw() +
+  scale_fill_manual(values = c(
+    moma.colors("Klein", type = "discrete")[2], # CCA
+    moma.colors("Klein", type = "discrete")[1], # coral
+    moma.colors("Klein", type = "discrete")[5], # macro
+    moma.colors("Klein", type = "discrete")[4] # turf
+  )) +
+  facet_wrap(~habitat, nrow = 1) +
+  ylab("Mean cover (%)") +
+  xlab("Year")
+
 ```
@@
   )) +
-  facet_wrap(~habitat) +
+  facet_wrap(~habitat, nrow = 1) +
   ylab("Mean cover (%)") +
@@
   facet_grid(~island_side) +
+  ylab("Mean cover (%)") +
+  xlab("Year")
+```
+
+## Quadrat-level island side
+
+```{r}
+quad4sum %>% 
+  # summarize and group quadrats by island/habitat
+  group_by(year, island_side, substrate) %>% 
+  summarise(mean_cov = mean(sum_cover),
+            sd_cov = sd(sum_cover, na.rm = TRUE),
+            n = n(),
+            CI95 = 1.96*(sd_cov)/sqrt(n)) %>% 
+  # relevel for graphing
+  mutate(island_side = factor(island_side, levels = c("north", "east", "west"))) %>% 
+  ggplot(aes(x = year, y = mean_cov, group = substrate, fill = substrate)) +
+  geom_ribbon(aes(ymin = mean_cov -CI95, 
+                    ymax = mean_cov + CI95),
+              alpha  = 0.7) +
+  geom_line() +
+  theme_bw() +
+  scale_fill_manual(values = c(
+    moma.colors("Klein", type = "discrete")[2], # CCA
+    moma.colors("Klein", type = "discrete")[1], # coral
+    moma.colors("Klein", type = "discrete")[5], # macro
+    moma.colors("Klein", type = "discrete")[4] # turf
+  )) +
+  facet_wrap(~island_side, nrow = 1) +
+  ylab("Mean cover (%)") +
+  xlab("Year")
+```
+
+## Plot by substrate with habitat as the facet
+
+By transect
+
+```{r plot}
+transectHabIS %>% 
+  ggplot(aes(x = year, y = mean_cov, group = habitat, fill = habitat)) +
+  geom_ribbon(aes(ymin = mean_cov -CI95, 
+                    ymax = mean_cov + CI95),
+              alpha  = 0.7) +
+  geom_line() +
+  theme_bw() +
+  scale_fill_viridis_d(direction = -1) +
+  facet_grid(substrate ~ island_side) +
+  ylab("Mean cover (%)") +
+  xlab("Year")
+```
+
+By quadrat
+
+```{r}
+quad4sum %>% 
+  # summarize and group quadrats by island/habitat
+  group_by(year, habitat, island_side, substrate) %>% 
+  summarise(mean_cov = mean(sum_cover),
+            sd_cov = sd(sum_cover, na.rm = TRUE),
+            n = n(),
+            CI95 = 1.96*(sd_cov)/sqrt(n)) %>% 
+  # relevel for graphing
+  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17")),
+         island_side = factor(island_side, levels = c("north", "east", "west"))) %>% 
+  ggplot(aes(x = year, y = mean_cov, group = habitat, fill = habitat)) +
+  geom_ribbon(aes(ymin = mean_cov -CI95, 
+                    ymax = mean_cov + CI95),
+              alpha  = 0.7) +
+  geom_line() +
+  theme_bw() +
+  scale_fill_viridis_d(direction = -1) +
+  facet_grid(substrate ~ island_side) +
   ylab("Mean cover (%)") +
@@
 
-## Distance matrix
+## Distance matrix: habitat-year
 
@@
 
-```{r}
+```{r look for zeros}
 sppMat %>% 
@@
 
-```{r}
+```{r get spp mat df}
 startAlg <- grep("acanthophora_spicifera", colnames(algae))
@@
 
+Start with just one habitat at a time
+
+### Fringing
+
+```{r get fringe mat}
+startAlg <- grep("acanthophora_spicifera", colnames(algae))
+endAlg <- grep("valonia_ventricosa", colnames(algae))
+
+# spp matrix
+algae %>% 
+  mutate(AlgSum = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(AlgSum != 0) %>% 
+  filter(habitat == "fringing") %>% 
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgSum) -> sppMat0fring
+
+# metadata matrix
+algae %>% 
+  mutate(AlgSum = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(AlgSum != 0) %>%
+   filter(habitat == "fringing") %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
+  # make a year-habitat column for betadisper
+  mutate(hab_year = paste0(habitat, "_", year)) -> sppMeta0fring
+```
+
 Calculate Bray Curtis distances
 
-```{r}
-brayMat <- vegdist(sppMat0, "bray") # 1 is maximum dissimilarity
+```{r calculate distances}
+brayMatFring <- vegdist(sppMat0fring, "bray") # 1 is maximum dissimilarity
 ```
@@
 
-```{r}
-habBetaDisper <- vegan::betadisper(brayMat, sppMeta0$hab_year)
+```{r calculate beta dispersion}
+yrBetaDisperFring <- vegan::betadisper(brayMatFring, sppMeta0fring$year) # warning: some squared distances are negative and changed to zero (ask Jamie??)
 
 # plot of dispersion test
-plot(habBetaDisper, sub = "") # lol
-boxplot(habBetaDisper, main = "Habitat", xlab = "") 
+plot(yrBetaDisperFring, sub = "") 
+boxplot(yrBetaDisperFring, main = "Fringing reef", xlab = "") 
 ```
+
+### Backreef
+
+```{r get backreef df}
+startAlg <- grep("acanthophora_spicifera", colnames(algae))
+endAlg <- grep("valonia_ventricosa", colnames(algae))
+
+# spp matrix
+algae %>% 
+  mutate(AlgSum = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(AlgSum != 0) %>% 
+  filter(habitat == "backreef") %>% 
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgSum) -> sppMat0back
+
+# metadata matrix
+algae %>% 
+  mutate(AlgSum = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(AlgSum != 0) %>%
+   filter(habitat == "backreef") %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
+  # make a year-habitat column for betadisper
+  mutate(hab_year = paste0(habitat, "_", year)) -> sppMeta0back
+```
+
+Calculate Bray Curtis distances
+
+```{r calculate distance}
+brayMatBack <- vegdist(sppMat0back, "bray") # 1 is maximum dissimilarity
+```
+
+Calculate beta disperion
+
+```{r calculate beta dispersion}
+yrBetaDisperBack <- vegan::betadisper(brayMatBack, sppMeta0back$year)
+
+# plot of dispersion test
+plot(yrBetaDisperBack, sub = "") # lol
+boxplot(yrBetaDisperBack, main = "Backreef", xlab = "") 
+```
+
+### Outer 10
+
+```{r get outer 10 df}
+startAlg <- grep("acanthophora_spicifera", colnames(algae))
+endAlg <- grep("valonia_ventricosa", colnames(algae))
+
+# spp matrix
+algae %>% 
+  mutate(AlgSum = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(AlgSum != 0) %>% 
+  filter(habitat == "outer_10") %>% 
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgSum) -> sppMat0outer10
+
+# metadata matrix
+algae %>% 
+  mutate(AlgSum = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(AlgSum != 0) %>%
+   filter(habitat == "outer_10") %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
+  # make a year-habitat column for betadisper
+  mutate(hab_year = paste0(habitat, "_", year)) -> sppMeta0outer10
+```
+
+Calculate Bray Curtis distances
+
+```{r calculate distance}
+brayMatOuter10 <- vegdist(sppMat0outer10, "bray") # 1 is maximum dissimilarity
+```
+
+Calculate beta disperion
+
+```{r calculate beta dispersion}
+yrBetaDisperOuter10 <- vegan::betadisper(brayMatOuter10, sppMeta0outer10$year) # error: some squared distances are negative and changed to zero
+
+# plot of dispersion test
+plot(yrBetaDisperOuter10, sub = "") # lol
+boxplot(yrBetaDisperOuter10, main = "Outer 10", xlab = "") 
+```
+
+### Outer 17
+
+```{r get outer 17 df}
+startAlg <- grep("acanthophora_spicifera", colnames(algae))
+endAlg <- grep("valonia_ventricosa", colnames(algae))
+
+# spp matrix
+algae %>% 
+  mutate(AlgSum = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(AlgSum != 0) %>% 
+  filter(habitat == "outer_17") %>% 
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgSum) -> sppMat0outer17
+
+# metadata matrix
+algae %>% 
+  mutate(AlgSum = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(AlgSum != 0) %>%
+   filter(habitat == "outer_17") %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
+  # make a year-habitat column for betadisper
+  mutate(hab_year = paste0(habitat, "_", year)) -> sppMeta0outer17
+```
+
+Calculate Bray Curtis distances
+
+```{r calculate distance}
+brayMatOuter17 <- vegdist(sppMat0outer17, "bray") # 1 is maximum dissimilarity
+```
+
+Calculate beta disperion
+
+```{r calculate beta dispersion}
+yrBetaDisperOuter17 <- vegan::betadisper(brayMatOuter17, sppMeta0outer17$year) # error: some squared distances are negative and changed to zero
+
+# plot of dispersion test
+plot(yrBetaDisperOuter17, sub = "") # lol
+boxplot(yrBetaDisperOuter17, main = "Outer 17", xlab = "") 
+```
+
+## Plot together
+
+How??
+
+```{r}
+```
+
+### Save dispersions
+
+```{r}
+# save(yrBetaDisperFring, file = here("data/fringingyrdispersion.Rdata"))
+# save(yrBetaDisperBack, file = here("data/backyrdispersion.Rdata"))
+# save(yrBetaDisperOuter10, file = here("data/outer10yrdispersion.Rdata"))
+# save(yrBetaDisperOuter17, file = here("data/outer17yrdispersion.Rdata"))
+```
+
+### Load dispersions
+
+```{r}
+load(here("data/fringingyrdispersion.Rdata"))
+load(here("data/backyrdispersion.Rdata"))
+load(here("data/outer10yrdispersion.Rdata"))
+load(here("data/outer17yrdispersion.Rdata"))
+```

### 2024-07-10 julianna_data_exploration.Rmd
Broke beta diversity up by island side as well as habitat + year

@@
   mutate(island_side = str_split(Grouping, pattern = "_")[[1]][[1]]) %>% 
-  mutate(Habitat = factor(Habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) 
+  mutate(Habitat = factor(Habitat, levels = c("fringing", "backreef", "outer_10", "outer_17")),
+         island_side = factor(island_side, levels = c("north", "east", "west"))) 
 ```
@@
   scale_colour_viridis_d(direction = -1) +
-  facet_grid(island_side~Habitat) +
+  facet_grid(island_side ~ Habitat) +
   labs(x = "Quadrats", y = "Richness")
@@
   # make a year-habitat column for betadisper
-  mutate(hab_year = paste0(habitat, "_", year)) -> sppMeta0
+  mutate(hab_year = paste0(habitat, "_", year),
+         island_side_hab_year = paste0(island_side, hab_year)) -> sppMeta0
 
@@
   # make a year-habitat column for betadisper
-  mutate(hab_year = paste0(habitat, "_", year)) -> sppMeta0fring
+  mutate(hab_year = paste0(habitat, "_", year),
+         island_side_hab_year = paste0(island_side, hab_year),
+         island_side_year = paste0(island_side, year)) -> sppMeta0fring
 ```
@@
   # make a year-habitat column for betadisper
-  mutate(hab_year = paste0(habitat, "_", year)) -> sppMeta0back
+  mutate(hab_year = paste0(habitat, "_", year),
+         island_side_hab_year = paste0(island_side, hab_year),
+         island_side_year = paste0(island_side, year)) -> sppMeta0back
 ```
@@
   # make a year-habitat column for betadisper
-  mutate(hab_year = paste0(habitat, "_", year)) -> sppMeta0outer10
+  mutate(hab_year = paste0(habitat, "_", year),
+         island_side_hab_year = paste0(island_side, hab_year),
+         island_side_year = paste0(island_side, year)) -> sppMeta0outer10
 ```
@@
   # make a year-habitat column for betadisper
-  mutate(hab_year = paste0(habitat, "_", year)) -> sppMeta0outer17
+  mutate(hab_year = paste0(habitat, "_", year),
+         island_side_hab_year = paste0(island_side, hab_year),
+         island_side_year = paste0(island_side, year)) -> sppMeta0outer17
 ```
@@
 
-## Plot together
-
-How??
-
-```{r}
+## Join together
+
+Attach back to metadata
+
+```{r join distances with metadata}
+sppMeta0fring %>% 
+  mutate(distance_centriod = yrBetaDisperFring$distances) -> fringeDist
+
+sppMeta0back %>% 
+  mutate(distance_centriod = yrBetaDisperBack$distances) -> backDist
+
+sppMeta0outer10 %>% 
+  mutate(distance_centriod = yrBetaDisperOuter10$distances) -> outer10Dist
+
+sppMeta0outer17 %>% 
+  mutate(distance_centriod = yrBetaDisperOuter17$distances) -> outer17Dist
+```
+
+```{r join together}
+# make sure this looks good
+colnames(fringeDist) == colnames(backDist) 
+colnames(backDist) == colnames(outer10Dist) 
+colnames(outer10Dist) == colnames(outer17Dist)
+
+# join together
+betaDHabYr <- rbind(fringeDist, backDist, outer10Dist, outer17Dist)
+```
+
+Confirm these produce the same plots
+
+```{r sanity check}
+betaDHabYr %>% 
+  ggplot(aes(x = year, y = distance_centriod, group = year)) +
+  geom_boxplot() +
+  facet_wrap(~habitat)
+```
+
+Add coral data
+
+```{r}
+allbenthos %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, coral) %>% 
+  # inner so that we don't add back in quadrats with 0% algae
+  inner_join(betaDHabYr) -> betaDHabYrCoral
+```
+
+## Plot with coral
+
+```{r scatter plot}
+betaDHabYrCoral %>% 
+  ggplot(aes(x = coral, y = distance_centriod)) +
+  geom_point() +
+  facet_grid(habitat ~ year)
+```
+
+```{r boxplot}
+# get mean coral cover
+betaDHabYrCoral %>% 
+  group_by(year, habitat) %>% 
+  summarize(Mean_coral = mean(coral)) %>% 
+  full_join(betaDHabYrCoral) %>% 
+  ggplot(aes(x = year, y = distance_centriod, group = year)) +
+  geom_boxplot() +
+  # Custom the Y scales:
+  scale_y_continuous(
+    # Features of the first axis
+    name = "Algal beta diversity",
+    # Add a second axis and specify its features
+    sec.axis = sec_axis( trans=~.*100, name="Avg coral cover")
+  ) +
+  geom_line(aes(group = 1, x = year, y = Mean_coral/100), size = 2, color = "red") +
+  facet_wrap(~habitat) 
+```
+
+```{r}
+betaDHabYrCoral %>% 
+  group_by(year, habitat) %>% 
+  summarize(Mean_coral = mean(coral)) %>% 
+  full_join(betaDHabYrCoral) %>% 
+  ggplot(aes(group = 1, x = year, y = Mean_coral, group = year)) +
+  geom_line()  +
+  facet_wrap(~habitat) 
 ```
@@
 ```
+
+## Dispersion by island side-year-habitat
+
+Calculate beta dispersion (do it in groups for the sake of the computer)
+
+```{r calculate beta dispersion by habitat}
+yrISBetaDisperFring <- vegan::betadisper(brayMatFring, sppMeta0fring$island_side_year) # error: some squared distances are negative and changed to zero
+# save(yrISBetaDisperFring, file = here("data/fringingyrISdispersion.Rdata"))
+
+yrISBetaDisperBack <- vegan::betadisper(brayMatBack, sppMeta0back$island_side_year)  # error: some squared distances are negative and changed to zero
+# save(yrISBetaDisperBack, file = here("data/backyrISdispersion.Rdata"))
+
+yrISBetaDisperOuter10 <- vegan::betadisper(brayMatOuter10, sppMeta0outer10$island_side_year) # error: some squared distances are negative and changed to zero
+# save(yrISBetaDisperOuter10, file = here("data/outer10yrISdispersion.Rdata"))
+
+yrISBetaDisperOuter17 <- vegan::betadisper(brayMatOuter17, sppMeta0outer17$island_side_year) # error: some squared distances are negative and changed to zero
+# save(yrISBetaDisperOuter17, file = here("data/outer17yrISdispersion.Rdata"))
+```
+
+### Load dispersions
+
+```{r}
+load(here("data/fringingyrISdispersion.Rdata"))
+load(here("data/backyrISdispersion.Rdata"))
+load(here("data/outer10yrISdispersion.Rdata"))
+load(here("data/outer17yrISdispersion.Rdata"))
+```
+
+### Join together
+
+```{r get all distances}
+# get distances
+sppMeta0fring %>% 
+  mutate(distance_centriod = yrISBetaDisperFring$distances) -> fringeDistIS
+
+sppMeta0back %>% 
+  mutate(distance_centriod = yrISBetaDisperBack$distances) -> backDistIS
+
+sppMeta0outer10 %>% 
+  mutate(distance_centriod = yrISBetaDisperOuter10$distances) -> outer10DistIS
+
+sppMeta0outer17 %>% 
+  mutate(distance_centriod = yrISBetaDisperOuter17$distances) -> outer17DistIS
+```
+
+```{r join together}
+# make sure this looks good
+colnames(fringeDistIS) == colnames(backDistIS) 
+colnames(backDistIS) == colnames(outer10DistIS) 
+colnames(outer10DistIS) == colnames(outer17DistIS)
+
+# join together
+betaDHabYrIS <- rbind(fringeDistIS, backDistIS, outer10DistIS, outer17DistIS)
+
+# add coral
+allbenthos %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, coral) %>% 
+  # inner so that we don't add back in quadrats with 0% algae
+  inner_join(betaDHabYrIS) -> betaDHabYrISCoral
+```
+
+### Plot
+
+```{r without coral}
+betaDHabYrISCoral %>% 
+  mutate(island_side = factor(island_side, levels = c("north", "east", "west")),
+        habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
+  ggplot(aes(x = year, y = distance_centriod, group = year)) +
+  geom_boxplot() +
+  facet_grid(island_side~habitat)
+```
+
+With coral
+
+```{r scatter plot}
+betaDHabYrISCoral %>% 
+  mutate(island_side = factor(island_side, levels = c("north", "east", "west")),
+        habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
+  ggplot(aes(x = coral, y = distance_centriod)) +
+  geom_point() +
+  facet_grid(habitat ~ year)
+```
+
+```{r boxplot}
+# get mean coral cover
+betaDHabYrISCoral %>% 
+  group_by(year, habitat, island_side) %>% 
+  summarize(Mean_coral = mean(coral)) %>% 
+  full_join(betaDHabYrCoral) %>% 
+  mutate(island_side = factor(island_side, levels = c("north", "east", "west")),
+        habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
+  ggplot(aes(x = year, y = distance_centriod, group = year)) +
+  geom_boxplot() +
+  # Custom the Y scales:
+  scale_y_continuous(
+    # Features of the first axis
+    name = "Algal beta diversity",
+    # Add a second axis and specify its features
+    sec.axis = sec_axis( trans=~.*100, name="Avg coral cover")
+  ) +
+  geom_line(aes(group = 1, x = year, y = Mean_coral/100), size = 2, color = "red") +
+  facet_grid(island_side~habitat) 
+```

### 2024-07-16 noam_data_exploration_v3.Rmd
test

@@
 editor_options: 
-  chunk_output_type: console
+  chunk_output_type: inline
 ---
@@
 knitr::opts_chunk$set(echo = TRUE)
-
+rm(list = ls())
 library(here)
@@
 library(MuMIn)
-```
-
-```{r }
-data <- read.csv(here::here("data", "algaeonly_cleaned.csv"), stringsAsFactors = F) # replacements don't work when strings are factors
+library(codyn)
+library(nlme)
+library(mgcv)
+library(emmeans)
+```
+
+```{r import and format data}
+# starting with all data now because of 0 algae transects/quadrats
+data <- read.csv(here::here("data", "allbenthicdata_cleaned.csv"), stringsAsFactors = F)
+
+not_algae <- c("Ascidian", "Bare.Space", "Coral", "Coral.Rubble", "Corallimorpharia", "Heteractis.sp.", "Millepora.platyphylla", "No.data", "Sand", "Sacrophyton.sp.", "Shell.Debris", "Soft.Coral", "Sponge", "Tridacna.sp.", "Zooxanthellae", "row_sums")
+
+data <- data[ , -which(names(data) %in% not_algae)]
+
+# adding island side - adapted from Julianna's code
+data <- data %>% mutate(Island_side = case_when(Site == "LTER 1" ~ "North",
+                                 Site == "LTER 2" ~ "North",
+                                 Site == "LTER 3" ~ "East",
+                                 Site == "LTER 4" ~ "East",
+                                 Site == "LTER 5" ~ "West",
+                                 Site == "LTER 6" ~ "West"))
+data$Island_side_habitat <- paste(data$Island_side, data$Habitat)
 
 str(data)
-# test
+
+meta_cols <- c(colnames(data[1:7] ), "Island_side", "Island_side_habitat")
+species_cols <- colnames(data[8:77])
+
 #pivot long 
-
 long_data <- pivot_longer(
@@
 
-site_level <- transect_level %>%
-  dplyr::group_by(Year, Site, Habitat, Taxa) %>%
-  dplyr::summarise(Mean_site_cover = mean(Mean_cover), 
-            Site_se = sd(Mean_cover)/sqrt(n()))
+# site_level <- transect_level %>%
+#   dplyr::group_by(Year, Site, Habitat, Taxa) %>%
+#   dplyr::summarise(Mean_site_cover = mean(Mean_cover), 
+#             Site_se = sd(Mean_cover)/sqrt(n()))
 ```
@@
 ```{r diversity indicies, fig.height = 15, fig.width=20}
-meta_cols <- colnames(data[1:7]) 
-species_cols <- colnames(data[8:77])
 
 wide_transect_level <- data %>%
-  dplyr::group_by(Year, Site, Habitat, Site_habitat, Transect) %>%
+  dplyr::group_by(Year, Island_side, Island_side_habitat, Site, Habitat, Site_habitat, Transect) %>%
   dplyr::summarize(across(all_of(species_cols), mean, na.rm = TRUE))
@@
     Year = as.factor(wide_transect_level$Year),
+    Island_side = as.factor(wide_transect_level$Island_side),
+    Island_side_habitat = as.factor(wide_transect_level$Island_side_habitat),
     Site = as.factor(wide_transect_level$Site),
@@
     Site_habitat_transect = as.factor(paste(wide_transect_level$Site_habitat, wide_transect_level$Transect)),
-    Shannon = diversity(wide_transect_level[5:74], "shannon"),
-    Richness = rowSums(wide_transect_level[5:74] > 0)
+    Shannon = diversity(wide_transect_level[ , -which(names(wide_transect_level) %in% meta_cols)], "shannon"),
+    Richness = rowSums(wide_transect_level[ , -which(names(wide_transect_level) %in% meta_cols)] > 0)
   )
@@
 alpha_diversity_means <- alpha_diversity %>%
-  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
+  dplyr::group_by(Year, Island_side, Island_side_habitat, Site, Habitat, Site_habitat) %>%
   dplyr::summarise(Richness_mean = mean(Richness), 
@@
                "LTER 6" = "orange")
+island_side_colours <- c("North" = "darkblue",
+                         "East" = "mediumorchid",
+                         "West" = "orange")
 habitat_linetypes <- c("Fringing" = "solid",
@@
   ggplot(data = alpha_diversity_means) +
-  geom_point(aes(x = Year, y = Richness_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8) +
-  geom_line(aes(x = Year, y = Richness_mean, colour = Site, group = Site_habitat), size = 1.2) +
-  geom_errorbar(aes(x = Year, ymin = Richness_mean - Richness_se , ymax = Richness_mean + Richness_se, colour = Site,
-                    group = Site_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = site_colours) +
+  geom_point(aes(x = Year, y = Richness_mean, colour = Island_side, shape = Habitat, group = Island_side_habitat), size = 8) +
+  geom_line(aes(x = Year, y = Richness_mean, colour = Island_side, group = Island_side_habitat), size = 1.2) +
+  geom_errorbar(aes(x = Year, ymin = Richness_mean - Richness_se , ymax = Richness_mean + Richness_se, colour = Island_side,
+                    group = Island_side_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = island_side_colours) +
   scale_shape_manual(values = habitat_shapes) +
@@
   ggplot(data = alpha_diversity_means) +
-  geom_point(aes(x = Year, y = Shannon_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8) +
-  geom_line(aes(x = Year, y = Shannon_mean, colour = Site, group = Site_habitat), size = 1.2) +
-  geom_errorbar(aes(x = Year, ymin = Shannon_mean - Shannon_se , ymax = Shannon_mean + Shannon_se, colour = Site,
-                    group = Site_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = site_colours) +
+  geom_point(aes(x = Year, y = Shannon_mean, colour = Island_side, shape = Habitat, group = Island_side_habitat), size = 8) +
+  geom_line(aes(x = Year, y = Shannon_mean, colour = Island_side, group = Island_side_habitat), size = 1.2) +
+  geom_errorbar(aes(x = Year, ymin = Shannon_mean - Shannon_se , ymax = Shannon_mean + Shannon_se, colour = Island_side,
+                    group = Island_side_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = island_side_colours) +
   scale_shape_manual(values = habitat_shapes) +
@@
 
-```{r basic alpha diversity stats}
-# these violate all sorts of assumptions I'm just messing around
-richness_mod <- glmmTMB(Richness ~ Year*Site*Habitat + (1|Site_habitat_transect), data = alpha_diversity, family = poisson())
-Anova(richness_mod)
-
-# shannon_mod <- glmmTMB(Shannon ~ Year*Site*Habitat + (1|Site_habitat_transect), data = alpha_diversity, family = Gamma(link = "log")) # 0 values fuck this up. Return to this when not tired. 
-```
-
-```{r nmds, fig.height = 15, fig.width = 15}
-species_matrix <- as.matrix(wide_transect_level[6:74])
-rownames(species_matrix) <- paste(wide_transect_level$Year, wide_transect_level$Site_habitat, wide_transect_level$Transect)
-rows_to_remove <- rowSums(species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
-species_matrix <- species_matrix[!rows_to_remove,] 
-wide_transect_level_subset <- wide_transect_level[!rows_to_remove,] # make sure rows are also removed from dataframe
-wide_transect_level_subset$Stratum <- paste(wide_transect_level_subset$Year, "-", wide_transect_level_subset$Site, "-", wide_transect_level_subset$Habitat )
-
-# identifying species w/ > 10% cover for plotting later
-# NOTE: Used ChatGPT to generate these next few lines of code but it looked good so I went with it
-columns_to_keep <- apply(species_matrix, 2, function(x) any(x >= 10))
-abundant_species_matrix <- as.data.frame(species_matrix)[ , c(TRUE, columns_to_keep)]
-abundant_species_names <- colnames(abundant_species_matrix)
-
-# WARNING - Before you run this, know that it will take a long time
-set.seed(1032) 
-# ord2 <- metaMDS(species_matrix, dist	= "bray",	trymax	= 100,	k	= 2) # looks like it automatically wisconsin transforms now?
-# ord2 # stress = 0.22
-# stressplot(ord2)
-
-set.seed(1032)
-ord3 <- metaMDS(species_matrix, dist	= "bray",	trymax	= 100,	k	= 3)
-ord3 # 0.166
-stressplot(ord3)
-
-# pro <- procrustes(ord2, ord3) # visually compare w procrustes rotation
-
-# plot(pro, cex=1.5)
-
-site.scores <- as.data.frame(vegan::scores(ord3)$sites)  # extract the site scores and convert to a data.frame
-# add back in metadata
-site.scores$Year <- as.factor(wide_transect_level_subset$Year)
-site.scores$Site <- as.factor(wide_transect_level_subset$Site)
-site.scores$Habitat <- as.factor(wide_transect_level_subset$Habitat)
-site.scores$Site_habitat <- as.factor(wide_transect_level_subset$Site_habitat)
-site.scores$Transect <- as.factor(wide_transect_level_subset$Transect)
-site.scores$Stratum <- as.factor(wide_transect_level_subset$Stratum)
-
-cent <- aggregate(cbind(NMDS1, NMDS2) ~ Stratum, data = site.scores, FUN = mean) # compute centroids for NMDS1 and NMDS2
-
-# Need to split stratum up for plotting later
-cent<- 
-  cent %>%
-    separate(Stratum, into = c("Year", "Site", "Habitat"), sep = "-")
-cent$Site_habitat <- paste(cent$Site, cent$Habitat)
-# Add back in stratum column
-cent$Stratum <-  aggregate(cbind(NMDS1, NMDS2) ~ Stratum, data = site.scores, FUN = mean)$Stratum
-# A lot of extra spaces
-
-cent <- cent %>%
-  mutate(across(everything(), str_trim))
-cent$Year <- as.factor(cent$Year)
-cent$Habitat <- as.factor(cent$Habitat)
-cent$Site <- as.factor(cent$Site)
-cent$Site_habitat <- as.factor(cent$Site_habitat)
-cent$NMDS1 <- as.numeric(cent$NMDS1) # read in as characters for some reason
-cent$NMDS2 <- as.numeric(cent$NMDS2)
-species.scores <- as.data.frame(scores(ord3)$species)  # extract the species scores and convert to a data.frame
-species.scores$Species <- rownames(species.scores)
-
-# subset only the species w/ > 10% cover
-abundant.species.scores <- species.scores[species.scores$Species %in% abundant_species_names,]
-
-ggplot() + 
-  geom_point(data= site.scores, aes(x=NMDS1,y=NMDS2, shape = Habitat, colour = Site),
-             alpha = 0.25, size=2) + # add the point markers
-  geom_text(data= abundant.species.scores,aes(x=NMDS1,y=NMDS2, label = Species), size = 4) +  # add the species labels
-
-  geom_path(data = cent, aes(x = NMDS1, y = NMDS2, group = Site_habitat, colour = Site),
-            arrow = arrow(length=unit(0.5,"cm"), ends="last", type = "closed"), size = 1) +
-  facet_wrap(~factor(Habitat, levels=c('Fringing','Backreef','Outer 10','Outer 17'))) +
-  scale_linetype_manual(values = habitat_linetypes) +
-  scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        strip.text.x = element_text(size = 25),
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.key.size = unit(2, "cm"), # increase legend size
-        legend.text = element_text(size = rel(1.5)), # increase legend text
-        legend.title = element_text(size = rel(1.5)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black")) + # add box around legend
-  guides(colour=guide_legend(title.position="top", # make color legend title go above legend
-                                     title.hjust =0.5, nrow = 2),
-         shape=guide_legend(title.position="top",  # make shape legend title go above legend
-                                     title.hjust =0.5,
-                            nrow = 2),
-         path = guide_legend(title.position="top",  # make shape legend title go above legend
-                                     title.hjust =0.5)
-         )
-
-# looks like stochasticity increases from outer 17 --> Fringing
-# Even in lagoon though it looks like communities return to initial state... except for North Shore LTER sites
-# ways to test stochasticity vs. determinism?
-# seems like deterministic trajectories, although we're not starting from a clean slate
-# accounting for founder effects?
-
-```
-
-```{r beta dispersion, fig.height = 15, fig.width=20}
-# Want to plot mean distance to centroid, but not sure how to do that with so many factors/interactions...
- bray_matrix <- vegdist(species_matrix, method = "bray")
-# year_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Year)
-# year_beta_disper #
-
-# plot of dispersion test
-# plot(year_beta_disper, sub = "") # lol
-# boxplot(year_beta_disper, main = "Year", xlab = "") 
-# 
-# site_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Site)
-# plot(site_beta_disper, sub = "") # site five stands out
-# boxplot(site_beta_disper, main = "Site", xlab = "") 
-# 
-# 
-# habitat_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Habitat)
-# plot(habitat_beta_disper, sub = "") 
-# boxplot(habitat_beta_disper, main = "Habitat", xlab = "") # greatest variance in fringing it seems (as expected)
-# 
-# site_habitat_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Site_habitat)
-# plot(site_habitat_beta_disper, sub = "") 
-# boxplot(site_habitat_beta_disper, main = "Site_habitat", xlab = "") 
-# 
-stratum_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Stratum)
-# plot(site_habitat_beta_disper, sub = "") 
-boxplot(stratum_beta_disper, main = "Site_habitat", xlab = "")
-stratum_distances <- as.data.frame(stratum_beta_disper$distances)
-colnames(stratum_distances)[1] <- "Distance"
-stratum_distances$Stratum <- wide_transect_level_subset$Stratum
-stratum_distances$Transect <- wide_transect_level_subset$Transect
-stratum_distances$Year <- wide_transect_level_subset$Year
-stratum_distances$Site <- wide_transect_level_subset$Site
-stratum_distances$Habitat <- wide_transect_level_subset$Habitat
-stratum_distances$Site_habitat <- wide_transect_level_subset$Site_habitat
-
-beta_disper_means <- stratum_distances %>%
-  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
-  dplyr::summarise(Distance_mean = mean(Distance), 
-            Distance_se = sd(Distance)/sqrt(n()))
-
-(Distance_plot <-
-  ggplot(data = beta_disper_means) +
-  geom_point(aes(x = Year, y = Distance_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
-  geom_line(aes(x = Year, y = Distance_mean, colour = Site, group = Site_habitat), size = 1.2) +
-  geom_errorbar(aes(x = Year, ymin = Distance_mean - Distance_se , ymax = Distance_mean + Distance_se, colour = Site,
-                    group = Site_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~factor(Habitat, levels=c('Fringing','Backreef','Outer 10','Outer 17')))+
-  labs(title = "Distance to centroid\n", y = "Distance\n") +
+```{r alpha diversity quadrat level, fig.height = 15, fig.width=20}
+alpha_diversity_quad <-
+  data.frame(
+    Year = as.numeric(data$Year),
+    Location = as.factor(data$Location),
+    Island_side = as.factor(data$Island_side),
+    Island_side_habitat = as.factor(data$Island_side_habitat),
+    Site = as.factor(data$Site),
+    Habitat = as.factor(data$Habitat),
+    Site_habitat = as.factor(data$Site_habitat),
+    Transect = as.factor(data$Transect),
+    Site_habitat_transect = as.factor(paste(data$Site_habitat, data$Transect)),
+    Shannon = diversity(data[ , -which(names(data) %in% meta_cols)], "shannon"),
+    Richness = rowSums(data[ , -which(names(data) %in% meta_cols)] > 0)
+  )
+
+alpha_diversity_quad_means <- alpha_diversity_quad %>%
+  dplyr::group_by(Year, Island_side, Island_side_habitat, Site, Habitat, Site_habitat) %>%
+  dplyr::summarise(Richness_mean = mean(Richness), 
+            Richness_se = sd(Richness)/sqrt(n()),
+            Shannon_mean = mean(Shannon), 
+            Shannon_se = sd(Shannon)/sqrt(n()))
+
+(richness_quad_plot <-
+  ggplot(data = alpha_diversity_quad_means) +
+  geom_point(aes(x = Year, y = Richness_mean, colour = Island_side, shape = Habitat, group = Island_side_habitat), size = 8) +
+  geom_line(aes(x = Year, y = Richness_mean, colour = Island_side, group = Island_side_habitat), size = 1.2) +
+  geom_errorbar(aes(x = Year, ymin = Richness_mean - Richness_se , ymax = Richness_mean + Richness_se, colour = Island_side,
+                    group = Island_side_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = island_side_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Habitat)+
+  labs(title = "a. Richness quad level\n", y = "Richness\n") +
   theme_bw() + 
@@
     )
+
+(shannon_quad_plot <-
+  ggplot(data = alpha_diversity_quad_means) +
+  geom_point(aes(x = Year, y = Shannon_mean, colour = Island_side, shape = Habitat, group = Island_side_habitat), size = 8) +
+  geom_line(aes(x = Year, y = Shannon_mean, colour = Island_side, group = Island_side_habitat), size = 1.2) +
+  geom_errorbar(aes(x = Year, ymin = Shannon_mean - Shannon_se , ymax = Shannon_mean + Shannon_se, colour = Island_side,
+                    group = Island_side_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = island_side_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Habitat)+
+  labs(title = "b. Shannon quad level\n", y = "shannon\n") +
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+        axis.line.y = element_line(colour = "black"),
+        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        axis.line.x = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(size = 32),
+        legend.key.size = unit(2, "cm"), # increase legend size
+        legend.text = element_text(size = rel(2)), # increase legend text
+        legend.title = element_text(size = rel(2.25)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal',
+        legend.box.background = element_rect(colour = "black") # make legend horizontal
+        )
+    )
+
+```
+```{r compare quad and transect level plots, fig.height = 30, fig.width= 40}
+ggarrange(richness_plot + rremove("x.text"), richness_quad_plot + rremove("x.text"),
+          shannon_plot,shannon_quad_plot,
+          common.legend = T,
+          nrow = 2, ncol = 2)
+
+```
+
+```{r gls approach}
+
+# FRAMING: What is the general trend in algal alpha diversity over the course of the time series? Increasing, decreasing, neutral?
+
+columns <- c("Location", "Intercept", "Slope") # names of columns for dataframe
+gls_coefficients <- # new dataframe name
+  data.frame(matrix(nrow = length(unique(alpha_diversity_quad$Location)), # 1200 quadrats = 1200 different time series
+                    ncol = length(columns)) # specified to make sure number of columns doesn't exceed what was specified
+             )
+colnames(gls_coefficients) <- columns # change column names to what was specified
+gls_coefficients$Location <- levels(alpha_diversity_quad$Location) # populate Location column with each quad/time series
+
+for(i in 1:length(unique(alpha_diversity_quad$Location)) ){ # for each  quadrat
+  # create a new dataframe for the quadrat of the current iteration
+  newData <- subset(alpha_diversity_quad, Location == levels(alpha_diversity_quad$Location)[i])
+
+  # fit gls for current time series
+  loop_mod <- gls(Richness ~ Year, correlation =corAR1(form=~Year), data = newData)
+
+  # populate appropriate row and column with the relevant coeffcients from gls
+  gls_coefficients[i, 2] <- unname(coef(loop_mod)[1])
+  gls_coefficients[i, 3] <- unname(coef(loop_mod)[2])
+
+}
+
+hist(gls_coefficients$Slope) # looks basically normal, bounded -Inf, Inf
+
+hist(gls_coefficients$Intercept) # same with intercept - although it probably shouldn't be
+
+# Start adding back in relevant metadata - probably a more elegant way to do this!!!
+habitat_vector <- c(rep("Backreef", 50), rep("Fringing", 50), rep("Outer 10", 50), rep("Outer 17", 50))
+transect_sequence <- c(rep(1,10),
+                       rep(2,10),
+                       rep(3,10),
+                       rep(4,10),
+                       rep(5,10))
+extra_meta <- 
+  data.frame(Site = as.factor(c(rep("LTER 1", 200),
+                                rep("LTER 2", 200),
+                                rep("LTER 3", 200),
+                                rep("LTER 4", 200),
+                                rep("LTER 5", 200),
+                                rep("LTER 6", 200)) ),
+             Island_side = as.factor(c(rep("North", 400),
+                                     rep("East", 400),
+                                     rep("West", 400))
+                                     ),
+             Habitat = as.factor(rep(habitat_vector, 6 )),
+             Transect = as.factor(rep(transect_sequence, 240)),
+             Location = levels(alpha_diversity_quad$Location) )
+
+gls_coefficients_merged <- merge(gls_coefficients, extra_meta, by = "Location")
+
+slope_mod <- glmmTMB(Slope ~ Habitat*Island_side + (1|Site/Transect), data = gls_coefficients_merged)
+summary(slope_mod) # pretty low variance in random effects
+plot(gls_res <- simulateResiduals(slope_mod)) # there are issues, but I can't figure out what they are...
+# I think this is due to 1200 obs in model
+hist(residuals(slope_mod))
+plot(predict(slope_mod) ~ residuals(slope_mod))
+Anova(slope_mod)
+#                        Chisq Df Pr(>Chisq)    
+# Island_side           16.075  2  0.0003232 ***
+# Habitat             1301.260  3  < 2.2e-16 ***
+# Island_side:Habitat  799.995  6  < 2.2e-16 ***
+r.squaredGLMM(slope_mod) # 0.16, 0.18
+plot(allEffects(slope_mod))
+# Backreef: richness increases on East and North shore, but relatively constant west side
+# Fringe: richness decreasing on east and west side, but relatvely constant on north shore
+# Outer 10: Richness neutral/decreasing on east side, but neutral/increasing on north and west side
+# Outer 17: Richness decreasing north/east side, but neutral west side
+# Worth noting trends are slight - arguable whether they are biologically meaningful
+
+slope_emm <- emmeans(slope_mod, ~ Habitat*Island_side) # create emmeans object
+emmeans::contrast(slope_emm, "consec", simple = "each", combine = TRUE, adjust = "bonferroni")
+# Note: not sure I want to include this
+```
+
+```{r gls plot, fig.height=20, fig.width=20}
+# not sure how to visualize here... thinking something like Dornelas et al. 2014 Fig 2a
+# I reached out to them about getting their code for this
+
+ggplot(data = gls_coefficients_merged) +
+  geom_histogram(aes(x = Slope), bins = 30) +
+  geom_vline(xintercept = 0, size = 2, linetype = 2) +
+  facet_grid(Habitat ~ Island_side) +
+#  theme_classic() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+#        panel.background = element_blank(),
+        legend.text = element_text(size=20),
+        legend.title = element_text(size=20),
+        axis.line = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
+        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain")
+  )
+
+```
+
+```{r gls intercept, fig.height=20, fig.width=20}
+intercept_mod <- glmmTMB(Intercept ~ Habitat*Island_side + (1|Site/Transect), data = gls_coefficients_merged)
+summary(intercept_mod) # a lot of variance w/i transect, less so w/i site
+plot(gls_int_res <- simulateResiduals(intercept_mod)) # same issues as above
+
+hist(residuals(intercept_mod))
+plot(predict(intercept_mod) ~ residuals(intercept_mod))
+Anova(intercept_mod)
+#                        Chisq Df Pr(>Chisq)    
+# Habitat             1307.838  3  < 2.2e-16 ***
+# Island_side           26.358  2  1.889e-06 ***
+# Habitat:Island_side  789.627  6  < 2.2e-16 ***
+r.squaredGLMM(intercept_mod) # 0.16, 0.18
+plot(allEffects(intercept_mod))
+
+ggplot(data = gls_coefficients_merged) +
+  geom_histogram(aes(x = Intercept), bins = 30) +
+  geom_vline(xintercept = 0, size = 2, linetype = 2) +
+  facet_grid(Habitat ~ Island_side) +
+#  theme_classic() +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+#        panel.background = element_blank(),
+        legend.text = element_text(size=20),
+        legend.title = element_text(size=20),
+        axis.line = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
+        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain")
+  )
+
+```
+
+```{r gam approach}
+# NOTE: Not sure where I'm going with this yet
+# Idea is to look at smoother estimates for each time series to get an idea for the "shape" of the time series
+# Not sure if worthwhile, might skip
+
+# The below takes 17 years to run in and of itself
+# gam_model <- gam(Richness ~ s(Year) + s(Year, by = Habitat) + 
+#                  s(Island_side, bs = "re") + s(Transect, bs = "re") + s(Location, bs = "re"),
+#                  data = alpha_diversity_quad, family = poisson())
+
+gam_columns <- c("Location", "Intercept", "edf") # names of columns for dataframe
+gam_coefficients <- # new dataframe name - same workflow as above
+  data.frame(matrix(nrow = length(unique(alpha_diversity_quad$Location)),
+                    ncol = length(gam_columns))
+             )
+colnames(gam_coefficients) <- gam_columns # change column names to what was specified
+gam_coefficients$Location <- levels(alpha_diversity_quad$Location) # populate Location column with each quad/time series
+
+# same workflow as above
+for(i in 1:length(unique(alpha_diversity_quad$Location)) ){ # for each  quadrat
+  # create a new dataframe for the quadrat of the current iteration
+  newData <- subset(alpha_diversity_quad, Location == levels(alpha_diversity_quad$Location)[i])
+
+  # fit gam for current time series
+  loop_mod <- gamm(Richness ~ s(Year), data = newData, correlation = corCAR1(form = ~ Year), method = "REML", family = poisson)
+  loop_sum <- summary(loop_mod$gam) # store summary data
+  # populate appropriate row and column with the relevant coeffcients from gls
+  gam_coefficients[i, 2] <- unname(loop_sum$p.coeff)
+  gam_coefficients[i, 3] <- unname(loop_sum$edf)
+
+}
+
+hist(gam_coefficients$Intercept)
+hist(gam_coefficients$edf) # generally around 1 except for an outlier
+# LTER 5 Fringing Reef Algae Transect 3 Quad 6 - a quadrat that has almost no algae the entire time series
+# issues with quadrat - smaller sample = more 0s, makes modeling trickier
+hist(subset(gam_coefficients, Location != "LTER 5 Fringing Reef Algae Transect 3 Quad 6")$edf)
+sum(gam_coefficients$edf > 1.01) # all but 18/1200 quadrats can be fit by a smoother with edf < 1.01
+# trends more straightforward than I anticipated?
+# argument for just looking at mean richness and slope?
+```
+
+```{r glmm approach}
+alpha_diversity_quad$Year_factor <- as.factor(alpha_diversity_quad$Year) # for glmmTMB ar1 structure
+# Framing - is adiv generally higher in certain habitats/regions?
+
+richness_mod <- glmmTMB(
+  Richness ~ Island_side*Habitat + (1 | Island_side/Site/Transect/Location) + ar1(Year_factor + 0 | Location),
+  family = poisson,
+  data = alpha_diversity_quad,
+  control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000))
+) # overspecified but going with it for now
+summary(richness_mod) # looks like strong autocorrelation, which makes sense to me
+Anova(richness_mod)
+plot(allEffects(richness_mod))
+
+```
+```{r turnover troubleshooting, fig.height = 30, fig.width=25}
+
+transect_level$Site_habitat <-  paste(transect_level$Site, transect_level$Habitat)
+transect_level$Stratum <- paste(transect_level$Site, "-", transect_level$Habitat, "-", transect_level$Transect)
+
+# long_data$Site_habitat <-  paste(long_data$Site, long_data$Habitat)
+# long_data$Stratum <- paste(long_data$Site, "-", long_data$Habitat, "-", long_data$Transect)
+
+# turnover analyses break when I try to do it at the quadrat level
+# checking if it's because there are quadrats that have effecively 0% algae
+total_cover_per_quadrat_year <- long_data %>%
+  group_by(Location, Year) %>%
+  summarize(total_cover = sum(Percent_cover), .groups = 'drop')
+
+# Filter quadrats where the total percent cover is 0%
+zero_cover_quadrats <- total_cover_per_quadrat_year %>%
+  filter(total_cover == 0)
+
+# Count the number of years each quadrat had 0% cover
+zero_cover_count <- zero_cover_quadrats %>%
+  group_by(Location) %>%
+  summarize(years_with_zero_cover = n())
+
+# Filter quadrats that had 0% cover for multiple years
+quadrats_multiple_years_zero_cover <- zero_cover_count %>%
+  filter(years_with_zero_cover > 1)
+
+length(unique(quadrats_multiple_years_zero_cover$Location)) # 144/1200 quadrats have multiple years with no algal cover
+
+# compare this to the number of transects with 0 algae cover
+total_cover_per_transect_year <- transect_level %>%
+  group_by(Stratum, Year) %>%
+  summarize(total_cover = sum(Mean_cover), .groups = 'drop')
+
+# Filter quadrats where the total percent cover is 0%
+zero_cover_transects <- total_cover_per_transect_year %>%
+  filter(total_cover == 0)
+
+# Count the number of years each quadrat had 0% cover
+zero_cover_transect_count <- zero_cover_transects %>%
+  group_by(Stratum) %>%
+  summarize(years_with_zero_cover = n())
+
+# Filter quadrats that had 0% cover for multiple years
+transects_multiple_years_zero_cover <- zero_cover_transect_count %>%
+  filter(years_with_zero_cover > 1)
+
+# NO transects ever have more than one year with 0 algal cover, as opposed to 144 quadrats
+# Fairly certain this is why the turnover analyses are breaking
+
+# identify quadrats missing timepoints
+years <- unique(long_data$Year)
+quadrats <- unique(long_data$Location)
+
+# Create a dataframe with all combinations of quadrats and years
+all_combinations <- expand.grid(Location = quadrats, Year = years)
+
+# Identify the missing combinations by performing an anti-join with your actual data
+missing_timepoints <- anti_join(all_combinations, long_data, by = c("Location", "Year"))
+
+# Display the missing timepoints
+print(missing_timepoints)
+
+problem_children <- c("LTER 5 Fringing Reef Algae Transect 3 Quad 9", "LTER 5 Fringing Reef Algae Transect 3 Quad 8", "LTER 5 Fringing Reef Algae Transect 3 Quad 6", "LTER 5 Fringing Reef Algae Transect 3 Quad 3", "LTER 5 Fringing Reef Algae Transect 3 Quad 10")
+
+long_data_sub <- long_data[long_data$Location != problem_children, ] # remove problem children
+# still get the same error when I try to run at quadrat level
+# I think it's because of the 0s
+# Error in names(output) <- `*vtmp*` : 
+#  'names' attribute [1] must be the same length as the vector [0]
+```
+
+```{r turnover}
+# the turnover function calculates turnover as the proportion or species that either appear or disappear between timepoints
+total_turnover<-
+  turnover(
+  df = long_data_sub,
+  time.var = "Year",
+  species.var = "Taxa",
+  abundance.var = "Percent_cover",
+  replicate.var = "Location",
+  metric = "total"
+)
+
+total_turnover<- 
+  total_turnover %>%
+    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
+total_turnover$Site_habitat <- paste(total_turnover$Site, total_turnover$Habitat)
+# Add back in stratum column
+
+# A lot of extra spaces
+
+total_turnover <- total_turnover %>%
+  mutate(across(everything(), str_trim))
+
+str(total_turnover) # everything became characters
+
+total_turnover$total <- as.numeric(total_turnover$total)
+total_turnover$Year <- as.factor(total_turnover$Year)
+total_turnover$Site <- as.factor(total_turnover$Site)
+total_turnover$Habitat <- as.factor(total_turnover$Habitat)
+
+turnover_means <-
+  total_turnover %>%
+  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
+  dplyr::summarise(Turnover_mean = mean(total), 
+            Turnover_se = sd(total)/sqrt(n()))
+
+(turnover_plot <-
+  ggplot(data = turnover_means) +
+  geom_point(aes(x = Year, y = Turnover_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
+  geom_line(aes(x = Year, y = Turnover_mean, colour = Site, group = Site_habitat), size = 1.2) +
+  geom_errorbar(aes(x = Year, ymin = Turnover_mean - Turnover_se , ymax = Turnover_mean + Turnover_se, colour = Site,
+                    group = Site_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site_habitat, nrow = 6)+
+  labs(title = "a. Turnover\n", y = "Total turnover\n") +
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+        axis.line.y = element_line(colour = "black"),
+        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        axis.line.x = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(size = 32),
+        legend.key.size = unit(2, "cm"), # increase legend size
+        legend.text = element_text(size = rel(2)), # increase legend text
+        legend.title = element_text(size = rel(2.25)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal',
+        legend.box.background = element_rect(colour = "black") # make legend horizontal
+        )
+    )
+#
+
+# examine appearance and disappearances independently 
+# refers to the proportion of species that appeared or disappeared relative to total species pool between two time points
+
+appearance <-
+  turnover(
+  df = long_data,
+  time.var = "Year",
+  species.var = "Taxa",
+  abundance.var = "Percent_cover",
+  replicate.var = "Location",
+  metric = "appearance"
+)
+
+appearance<- 
+  appearance %>%
+    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
+appearance$Site_habitat <- paste(appearance$Site, appearance$Habitat)
+# Add back in stratum column
+
+# A lot of extra spaces
+
+appearance <- appearance %>%
+  mutate(across(everything(), str_trim))
+
+str(appearance) # everything became characters
+
+appearance$appearance <- as.numeric(appearance$appearance)
+appearance$Year <- as.factor(appearance$Year)
+appearance$Site <- as.factor(appearance$Site)
+appearance$Habitat <- as.factor(appearance$Habitat)
+
+appearance_means <-
+  appearance %>%
+  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
+  dplyr::summarise(Appearance_mean = mean(appearance), 
+            Appearance_se = sd(appearance)/sqrt(n()))
+
+(appearance_plot <-
+  ggplot(data = appearance_means) +
+  geom_point(aes(x = Year, y = Appearance_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
+  geom_line(aes(x = Year, y = Appearance_mean, colour = Site, group = Site_habitat), size = 1.2) +
+  geom_errorbar(aes(x = Year, ymin = Appearance_mean - Appearance_se , ymax = Appearance_mean + Appearance_se, colour = Site,
+                    group = Site_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site_habitat, nrow = 6)+
+  labs(title = "b. Appearances\n", y = "Total turnover\n") +
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+        axis.line.y = element_line(colour = "black"),
+        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        axis.line.x = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(size = 32),
+        legend.key.size = unit(2, "cm"), # increase legend size
+        legend.text = element_text(size = rel(2)), # increase legend text
+        legend.title = element_text(size = rel(2.25)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal',
+        legend.box.background = element_rect(colour = "black") # make legend horizontal
+        )
+    )
+#
+
+disappearance <-
+  turnover(
+  df = long_data,
+  time.var = "Year",
+  species.var = "Taxa",
+  abundance.var = "Mean_cover",
+  replicate.var = "Stratum",
+  metric = "disappearance"
+)
+
+disappearance<- 
+  disappearance %>%
+    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
+disappearance$Site_habitat <- paste(disappearance$Site, disappearance$Habitat)
+# Add back in stratum column
+
+# A lot of extra spaces
+
+disappearance <- disappearance %>%
+  mutate(across(everything(), str_trim))
+
+str(disappearance) # everything became characters
+
+disappearance$disappearance <- as.numeric(disappearance$disappearance)
+disappearance$Year <- as.factor(disappearance$Year)
+disappearance$Site <- as.factor(disappearance$Site)
+disappearance$Habitat <- as.factor(disappearance$Habitat)
+
+disappearance_means <-
+  disappearance %>%
+  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
+  dplyr::summarise(Disappearance_mean = mean(disappearance), 
+            Disappearance_se = sd(disappearance)/sqrt(n()))
+
+(disappearance_plot <-
+  ggplot(data = disappearance_means) +
+  geom_point(aes(x = Year, y = Disappearance_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
+  geom_line(aes(x = Year, y = Disappearance_mean, colour = Site, group = Site_habitat), size = 1.2) +
+  geom_errorbar(aes(x = Year, ymin = Disappearance_mean - Disappearance_se , ymax = Disappearance_mean + Disappearance_se, colour = Site,
+                    group = Site_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site_habitat, nrow = 6)+
+  labs(title = "c. Disappearance\n", y = "Total turnover\n") +
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+        axis.line.y = element_line(colour = "black"),
+        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        axis.line.x = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(size = 32),
+        legend.key.size = unit(2, "cm"), # increase legend size
+        legend.text = element_text(size = rel(2)), # increase legend text
+        legend.title = element_text(size = rel(2.25)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal',
+        legend.box.background = element_rect(colour = "black") # make legend horizontal
+        )
+    )
+
+# rank shift
+
+rank_shift<-
+  rank_shift(
+  df = long_data,
+  time.var = "Year",
+  species.var = "Taxa",
+  abundance.var = "Mean_cover",
+  replicate.var = "Stratum"
+)
+
+rank_shift<- 
+  rank_shift %>%
+    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
+rank_shift$Site_habitat <- paste(rank_shift$Site, rank_shift$Habitat)
+# Add back in stratum column
+
+# A lot of extra spaces
+
+rank_shift <- rank_shift %>%
+  mutate(across(everything(), str_trim))
+
+str(rank_shift) # everything became characters
+
+rank_shift$MRS <- as.numeric(rank_shift$MRS)
+# Create a column with the final year from the returned time.var_pair
+rank_shift$Year <- as.numeric(substr(rank_shift$year_pair, 6, 9))
+rank_shift$Site <- as.factor(rank_shift$Site)
+rank_shift$Habitat <- as.factor(rank_shift$Habitat)
+
+rank_shift_means <-
+  rank_shift %>%
+  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
+  dplyr::summarise(Rank_shift_mean = mean(MRS), 
+            Rank_shift_se = sd(MRS)/sqrt(n()))
+
+(rank_shift_plot <-
+  ggplot(data = rank_shift_means) +
+  geom_point(aes(x = Year, y = Rank_shift_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
+  geom_line(aes(x = Year, y = Rank_shift_mean, colour = Site, group = Site_habitat), size = 1.2) +
+  geom_errorbar(aes(x = Year, ymin = Rank_shift_mean - Rank_shift_se , ymax = Rank_shift_mean + Rank_shift_se, colour = Site,
+                    group = Site_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site_habitat, nrow = 6)+
+  labs(title = "a. rank_shift\n", y = "Total rank_shift\n") +
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+        axis.line.y = element_line(colour = "black"),
+        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        axis.line.x = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(size = 32),
+        legend.key.size = unit(2, "cm"), # increase legend size
+        legend.text = element_text(size = rel(2)), # increase legend text
+        legend.title = element_text(size = rel(2.25)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal',
+        legend.box.background = element_rect(colour = "black") # make legend horizontal
+        )
+    )
+
+# rate change
+rate_change_table<-
+  rate_change(
+  df = long_data,
+  time.var = "Year",
+  species.var = "Taxa",
+  abundance.var = "Mean_cover",
+  replicate.var = "Stratum"
+)
+# kable(rate_change_table)
+
+rate_change<-
+  rate_change_interval(
+  df = long_data,
+  time.var = "Year",
+  species.var = "Taxa",
+  abundance.var = "Mean_cover",
+  replicate.var = "Stratum"
+)
+
+rate_change<- 
+  rate_change %>%
+    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
+rate_change$Site_habitat <- paste(rate_change$Site, rate_change$Habitat)
+
+# A lot of extra spaces
+
+rate_change <- rate_change %>%
+  mutate(across(everything(), str_trim))
+
+str(rate_change) # everything became characters
+
+rate_change$distance <- as.numeric(rate_change$distance)
+rate_change$interval <- as.numeric(rate_change$interval)
+rate_change$Site <- as.factor(rate_change$Site)
+rate_change$Habitat <- as.factor(rate_change$Habitat)
+
+ggplot(rate_change, aes(interval, distance, colour = Site, shape = Habitat, group = Site_habitat, linetype = Habitat)) +
+  facet_wrap(~Site) + 
+  geom_point() +
+  scale_colour_manual(values = site_colours) +
+  scale_linetype_manual(values = habitat_linetypes) +
+  stat_smooth(method = "lm", se = T, size = 2) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site)+
+#  labs(y = "Distance\n") +
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+        axis.line.y = element_line(colour = "black"),
+        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        axis.line.x = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(size = 32),
+        legend.key.size = unit(2, "cm"), # increase legend size
+        legend.text = element_text(size = rel(2)), # increase legend text
+        legend.title = element_text(size = rel(2.25)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal',
+        legend.box.background = element_rect(colour = "black") # make legend horizontal
+        ) # I don't think these are linear relationships so the smoother isn't terriibly helpful
+```
+
+```{r nmds, fig.height = 15, fig.width = 15}
+species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
+species_matrix <- species_matrix/100 # convert from percent to proportion
+rownames(species_matrix) <- paste(data$Year, data$Site_habitat, data$Transect)
+rows_to_remove <- rowSums(species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
+species_matrix <- species_matrix[!rows_to_remove,] 
+data_subset <- data[!rows_to_remove,] # make sure rows are also removed from dataframe
+data_subset$Stratum <- paste(data_subset$Year, "-", data_subset$Site, "-", data_subset$Habitat )
+
+# identifying species w/ > 10% cover for plotting later
+# NOTE: Used ChatGPT to generate these next few lines of code but it looked good so I went with it
+columns_to_keep <- apply(species_matrix, 2, function(x) any(x >= 0.1))
+abundant_species_matrix <- as.data.frame(species_matrix)[ , c(TRUE, columns_to_keep)]
+abundant_species_names <- colnames(abundant_species_matrix)
+
+# My computer can't run this at the quadrat level
+set.seed(1032) 
+ ord2 <- metaMDS(species_matrix, dist	= "bray",	trymax	= 100,	k	= 2) 
+# ord2 # stress = 0.22
+# stressplot(ord2)
+
+set.seed(1032)
+ord3 <- metaMDS(species_matrix, dist	= "bray",	trymax	= 100,	k	= 3)
+ord3 # 0.166
+stressplot(ord3)
+
+# pro <- procrustes(ord2, ord3) # visually compare w procrustes rotation
+
+# plot(pro, cex=1.5)
+
+site.scores <- as.data.frame(vegan::scores(ord3)$sites)  # extract the site scores and convert to a data.frame
+# add back in metadata
+site.scores$Year <- as.factor(data_subset$Year)
+site.scores$Site <- as.factor(data_subset$Site)
+site.scores$Habitat <- as.factor(data_subset$Habitat)
+site.scores$Site_habitat <- as.factor(data_subset$Site_habitat)
+site.scores$Transect <- as.factor(data_subset$Transect)
+site.scores$Stratum <- as.factor(data_subset$Stratum)
+
+cent <- aggregate(cbind(NMDS1, NMDS2) ~ Stratum, data = site.scores, FUN = mean) # compute centroids for NMDS1 and NMDS2
+
+# Need to split stratum up for plotting later
+cent<- 
+  cent %>%
+    separate(Stratum, into = c("Year", "Site", "Habitat"), sep = "-")
+cent$Site_habitat <- paste(cent$Site, cent$Habitat)
+# Add back in stratum column
+cent$Stratum <-  aggregate(cbind(NMDS1, NMDS2) ~ Stratum, data = site.scores, FUN = mean)$Stratum
+# A lot of extra spaces
+
+cent <- cent %>%
+  mutate(across(everything(), str_trim))
+cent$Year <- as.factor(cent$Year)
+cent$Habitat <- as.factor(cent$Habitat)
+cent$Site <- as.factor(cent$Site)
+cent$Site_habitat <- as.factor(cent$Site_habitat)
+cent$NMDS1 <- as.numeric(cent$NMDS1) # read in as characters for some reason
+cent$NMDS2 <- as.numeric(cent$NMDS2)
+species.scores <- as.data.frame(scores(ord3)$species)  # extract the species scores and convert to a data.frame
+species.scores$Species <- rownames(species.scores)
+
+# subset only the species w/ > 10% cover
+abundant.species.scores <- species.scores[species.scores$Species %in% abundant_species_names,]
+
+ggplot() + 
+  geom_point(data= site.scores, aes(x=NMDS1,y=NMDS2, shape = Habitat, colour = Site),
+             alpha = 0.25, size=2) + # add the point markers
+  geom_text(data= abundant.species.scores,aes(x=NMDS1,y=NMDS2, label = Species), size = 4) +  # add the species labels
+
+  geom_path(data = cent, aes(x = NMDS1, y = NMDS2, group = Site_habitat, colour = Site),
+            arrow = arrow(length=unit(0.5,"cm"), ends="last", type = "closed"), size = 1) +
+  facet_wrap(~factor(Habitat, levels=c('Fringing','Backreef','Outer 10','Outer 17'))) +
+  scale_linetype_manual(values = habitat_linetypes) +
+  scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        strip.text.x = element_text(size = 25),
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.key.size = unit(2, "cm"), # increase legend size
+        legend.text = element_text(size = rel(1.5)), # increase legend text
+        legend.title = element_text(size = rel(1.5)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) + # add box around legend
+  guides(colour=guide_legend(title.position="top", # make color legend title go above legend
+                                     title.hjust =0.5, nrow = 2),
+         shape=guide_legend(title.position="top",  # make shape legend title go above legend
+                                     title.hjust =0.5,
+                            nrow = 2),
+         path = guide_legend(title.position="top",  # make shape legend title go above legend
+                                     title.hjust =0.5)
+         )
+
+# looks like stochasticity increases from outer 17 --> Fringing
+# Even in lagoon though it looks like communities return to initial state... except for North Shore LTER sites
+# ways to test stochasticity vs. determinism?
+# seems like deterministic trajectories, although we're not starting from a clean slate
+# accounting for founder effects?
+
+```
+
+```{r beta dispersion, fig.height = 15, fig.width=20}
+# Want to plot mean distance to centroid, but not sure how to do that with so many factors/interactions...
+ bray_matrix <- vegdist(species_matrix, method = "bray")
+# year_beta_disper <- vegan::betadisper(bray_matrix, data_subset$Year)
+# year_beta_disper #
+
+# plot of dispersion test
+# plot(year_beta_disper, sub = "") # lol
+# boxplot(year_beta_disper, main = "Year", xlab = "") 
+# 
+# site_beta_disper <- vegan::betadisper(bray_matrix, data_subset$Site)
+# plot(site_beta_disper, sub = "") # site five stands out
+# boxplot(site_beta_disper, main = "Site", xlab = "") 
+# 
+# 
+# habitat_beta_disper <- vegan::betadisper(bray_matrix, data_subset$Habitat)
+# plot(habitat_beta_disper, sub = "") 
+# boxplot(habitat_beta_disper, main = "Habitat", xlab = "") # greatest variance in fringing it seems (as expected)
+# 
+# site_habitat_beta_disper <- vegan::betadisper(bray_matrix, data_subset$Site_habitat)
+# plot(site_habitat_beta_disper, sub = "") 
+# boxplot(site_habitat_beta_disper, main = "Site_habitat", xlab = "") 
+# 
+stratum_beta_disper <- vegan::betadisper(bray_matrix, data_subset$Stratum)
+# plot(site_habitat_beta_disper, sub = "") 
+boxplot(stratum_beta_disper, main = "Site_habitat", xlab = "")
+stratum_distances <- as.data.frame(stratum_beta_disper$distances)
+colnames(stratum_distances)[1] <- "Distance"
+stratum_distances$Stratum <- data_subset$Stratum
+stratum_distances$Transect <- data_subset$Transect
+stratum_distances$Year <- data_subset$Year
+stratum_distances$Site <- data_subset$Site
+stratum_distances$Habitat <- data_subset$Habitat
+stratum_distances$Site_habitat <- data_subset$Site_habitat
+
+beta_disper_means <- stratum_distances %>%
+  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
+  dplyr::summarise(Distance_mean = mean(Distance), 
+            Distance_se = sd(Distance)/sqrt(n()))
+
+(Distance_plot <-
+  ggplot(data = beta_disper_means) +
+  geom_point(aes(x = Year, y = Distance_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
+  geom_line(aes(x = Year, y = Distance_mean, colour = Site, group = Site_habitat), size = 1.2) +
+  geom_errorbar(aes(x = Year, ymin = Distance_mean - Distance_se , ymax = Distance_mean + Distance_se, colour = Site,
+                    group = Site_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~factor(Habitat, levels=c('Fringing','Backreef','Outer 10','Outer 17')))+
+  labs(title = "Distance to centroid\n", y = "Distance\n") +
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+        axis.line.y = element_line(colour = "black"),
+        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        axis.line.x = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(size = 32),
+        legend.key.size = unit(2, "cm"), # increase legend size
+        legend.text = element_text(size = rel(2)), # increase legend text
+        legend.title = element_text(size = rel(2.25)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal',
+        legend.box.background = element_rect(colour = "black") # make legend horizontal
+        )
+    )
 # beta dispersion increases as we move from forereef to lagoon, lagoon to shore
@@
 ```
-
-```{r turnover, fig.height = 30, fig.width=25}
-library(codyn)
-transect_level$Site_habitat <-  paste(transect_level$Site, transect_level$Habitat)
-transect_level$Stratum <- paste(transect_level$Site, "-", transect_level$Habitat, "-", transect_level$Transect)
-
-# the turnover function calculates turnover as the proportion or species that either appear or disappear between timepoints
-total_turnover<-
-  turnover(
-  df = transect_level,
-  time.var = "Year",
-  species.var = "Taxa",
-  abundance.var = "Mean_cover",
-  replicate.var = "Stratum",
-  metric = "total"
-)
-
-total_turnover<- 
-  total_turnover %>%
-    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
-total_turnover$Site_habitat <- paste(total_turnover$Site, total_turnover$Habitat)
-# Add back in stratum column
-
-# A lot of extra spaces
-
-total_turnover <- total_turnover %>%
-  mutate(across(everything(), str_trim))
-
-str(total_turnover) # everything became characters
-
-total_turnover$total <- as.numeric(total_turnover$total)
-total_turnover$Year <- as.factor(total_turnover$Year)
-total_turnover$Site <- as.factor(total_turnover$Site)
-total_turnover$Habitat <- as.factor(total_turnover$Habitat)
-
-turnover_means <-
-  total_turnover %>%
-  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
-  dplyr::summarise(Turnover_mean = mean(total), 
-            Turnover_se = sd(total)/sqrt(n()))
-
-(turnover_plot <-
-  ggplot(data = turnover_means) +
-  geom_point(aes(x = Year, y = Turnover_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
-  geom_line(aes(x = Year, y = Turnover_mean, colour = Site, group = Site_habitat), size = 1.2) +
-  geom_errorbar(aes(x = Year, ymin = Turnover_mean - Turnover_se , ymax = Turnover_mean + Turnover_se, colour = Site,
-                    group = Site_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site_habitat, nrow = 6)+
-  labs(title = "a. Turnover\n", y = "Total turnover\n") +
-  theme_bw() + 
-  theme(panel.grid.major = element_blank(), 
-        panel.grid.minor = element_blank(),
-        panel.background = element_blank(),
-        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
-        axis.line.y = element_line(colour = "black"),
-        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
-        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
-        axis.line.x = element_line(colour = "black"),
-        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
-        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
-        strip.text.x = element_text(size = 32),
-        legend.key.size = unit(2, "cm"), # increase legend size
-        legend.text = element_text(size = rel(2)), # increase legend text
-        legend.title = element_text(size = rel(2.25)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal',
-        legend.box.background = element_rect(colour = "black") # make legend horizontal
-        )
-    )
-#
-
-# examine appearance and disappearances independently 
-# refers to the proportion of species that appeared or disappeared relative to total species pool between two time points
-
-appearance <-
-  turnover(
-  df = transect_level,
-  time.var = "Year",
-  species.var = "Taxa",
-  abundance.var = "Mean_cover",
-  replicate.var = "Stratum",
-  metric = "appearance"
-)
-
-appearance<- 
-  appearance %>%
-    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
-appearance$Site_habitat <- paste(appearance$Site, appearance$Habitat)
-# Add back in stratum column
-
-# A lot of extra spaces
-
-appearance <- appearance %>%
-  mutate(across(everything(), str_trim))
-
-str(appearance) # everything became characters
-
-appearance$appearance <- as.numeric(appearance$appearance)
-appearance$Year <- as.factor(appearance$Year)
-appearance$Site <- as.factor(appearance$Site)
-appearance$Habitat <- as.factor(appearance$Habitat)
-
-appearance_means <-
-  appearance %>%
-  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
-  dplyr::summarise(Appearance_mean = mean(appearance), 
-            Appearance_se = sd(appearance)/sqrt(n()))
-
-(appearance_plot <-
-  ggplot(data = appearance_means) +
-  geom_point(aes(x = Year, y = Appearance_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
-  geom_line(aes(x = Year, y = Appearance_mean, colour = Site, group = Site_habitat), size = 1.2) +
-  geom_errorbar(aes(x = Year, ymin = Appearance_mean - Appearance_se , ymax = Appearance_mean + Appearance_se, colour = Site,
-                    group = Site_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site_habitat, nrow = 6)+
-  labs(title = "b. Appearances\n", y = "Total turnover\n") +
-  theme_bw() + 
-  theme(panel.grid.major = element_blank(), 
-        panel.grid.minor = element_blank(),
-        panel.background = element_blank(),
-        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
-        axis.line.y = element_line(colour = "black"),
-        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
-        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
-        axis.line.x = element_line(colour = "black"),
-        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
-        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
-        strip.text.x = element_text(size = 32),
-        legend.key.size = unit(2, "cm"), # increase legend size
-        legend.text = element_text(size = rel(2)), # increase legend text
-        legend.title = element_text(size = rel(2.25)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal',
-        legend.box.background = element_rect(colour = "black") # make legend horizontal
-        )
-    )
-#
-
-disappearance <-
-  turnover(
-  df = transect_level,
-  time.var = "Year",
-  species.var = "Taxa",
-  abundance.var = "Mean_cover",
-  replicate.var = "Stratum",
-  metric = "disappearance"
-)
-
-disappearance<- 
-  disappearance %>%
-    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
-disappearance$Site_habitat <- paste(disappearance$Site, disappearance$Habitat)
-# Add back in stratum column
-
-# A lot of extra spaces
-
-disappearance <- disappearance %>%
-  mutate(across(everything(), str_trim))
-
-str(disappearance) # everything became characters
-
-disappearance$disappearance <- as.numeric(disappearance$disappearance)
-disappearance$Year <- as.factor(disappearance$Year)
-disappearance$Site <- as.factor(disappearance$Site)
-disappearance$Habitat <- as.factor(disappearance$Habitat)
-
-disappearance_means <-
-  disappearance %>%
-  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
-  dplyr::summarise(Disappearance_mean = mean(disappearance), 
-            Disappearance_se = sd(disappearance)/sqrt(n()))
-
-(disappearance_plot <-
-  ggplot(data = disappearance_means) +
-  geom_point(aes(x = Year, y = Disappearance_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
-  geom_line(aes(x = Year, y = Disappearance_mean, colour = Site, group = Site_habitat), size = 1.2) +
-  geom_errorbar(aes(x = Year, ymin = Disappearance_mean - Disappearance_se , ymax = Disappearance_mean + Disappearance_se, colour = Site,
-                    group = Site_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site_habitat, nrow = 6)+
-  labs(title = "c. Disappearance\n", y = "Total turnover\n") +
-  theme_bw() + 
-  theme(panel.grid.major = element_blank(), 
-        panel.grid.minor = element_blank(),
-        panel.background = element_blank(),
-        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
-        axis.line.y = element_line(colour = "black"),
-        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
-        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
-        axis.line.x = element_line(colour = "black"),
-        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
-        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
-        strip.text.x = element_text(size = 32),
-        legend.key.size = unit(2, "cm"), # increase legend size
-        legend.text = element_text(size = rel(2)), # increase legend text
-        legend.title = element_text(size = rel(2.25)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal',
-        legend.box.background = element_rect(colour = "black") # make legend horizontal
-        )
-    )
-
-# rank shift
-
-rank_shift<-
-  rank_shift(
-  df = transect_level,
-  time.var = "Year",
-  species.var = "Taxa",
-  abundance.var = "Mean_cover",
-  replicate.var = "Stratum"
-)
-
-rank_shift<- 
-  rank_shift %>%
-    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
-rank_shift$Site_habitat <- paste(rank_shift$Site, rank_shift$Habitat)
-# Add back in stratum column
-
-# A lot of extra spaces
-
-rank_shift <- rank_shift %>%
-  mutate(across(everything(), str_trim))
-
-str(rank_shift) # everything became characters
-
-rank_shift$MRS <- as.numeric(rank_shift$MRS)
-# Create a column with the final year from the returned time.var_pair
-rank_shift$Year <- as.numeric(substr(rank_shift$year_pair, 6, 9))
-rank_shift$Site <- as.factor(rank_shift$Site)
-rank_shift$Habitat <- as.factor(rank_shift$Habitat)
-
-rank_shift_means <-
-  rank_shift %>%
-  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
-  dplyr::summarise(Rank_shift_mean = mean(MRS), 
-            Rank_shift_se = sd(MRS)/sqrt(n()))
-
-(rank_shift_plot <-
-  ggplot(data = rank_shift_means) +
-  geom_point(aes(x = Year, y = Rank_shift_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
-  geom_line(aes(x = Year, y = Rank_shift_mean, colour = Site, group = Site_habitat), size = 1.2) +
-  geom_errorbar(aes(x = Year, ymin = Rank_shift_mean - Rank_shift_se , ymax = Rank_shift_mean + Rank_shift_se, colour = Site,
-                    group = Site_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site_habitat, nrow = 6)+
-  labs(title = "a. rank_shift\n", y = "Total rank_shift\n") +
-  theme_bw() + 
-  theme(panel.grid.major = element_blank(), 
-        panel.grid.minor = element_blank(),
-        panel.background = element_blank(),
-        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
-        axis.line.y = element_line(colour = "black"),
-        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
-        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
-        axis.line.x = element_line(colour = "black"),
-        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
-        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
-        strip.text.x = element_text(size = 32),
-        legend.key.size = unit(2, "cm"), # increase legend size
-        legend.text = element_text(size = rel(2)), # increase legend text
-        legend.title = element_text(size = rel(2.25)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal',
-        legend.box.background = element_rect(colour = "black") # make legend horizontal
-        )
-    )
-
-# rate change
-rate_change_table<-
-  rate_change(
-  df = transect_level,
-  time.var = "Year",
-  species.var = "Taxa",
-  abundance.var = "Mean_cover",
-  replicate.var = "Stratum"
-)
-# kable(rate_change_table)
-
-rate_change<-
-  rate_change_interval(
-  df = transect_level,
-  time.var = "Year",
-  species.var = "Taxa",
-  abundance.var = "Mean_cover",
-  replicate.var = "Stratum"
-)
-
-rate_change<- 
-  rate_change %>%
-    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
-rate_change$Site_habitat <- paste(rate_change$Site, rate_change$Habitat)
-
-# A lot of extra spaces
-
-rate_change <- rate_change %>%
-  mutate(across(everything(), str_trim))
-
-str(rate_change) # everything became characters
-
-rate_change$distance <- as.numeric(rate_change$distance)
-rate_change$interval <- as.numeric(rate_change$interval)
-rate_change$Site <- as.factor(rate_change$Site)
-rate_change$Habitat <- as.factor(rate_change$Habitat)
-
-ggplot(rate_change, aes(interval, distance, colour = Site, shape = Habitat, group = Site_habitat, linetype = Habitat)) +
-  facet_wrap(~Site) + 
-  geom_point() +
-  scale_colour_manual(values = site_colours) +
-  scale_linetype_manual(values = habitat_linetypes) +
-  stat_smooth(method = "lm", se = T, size = 2) +
-  scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site)+
-#  labs(y = "Distance\n") +
-  theme_bw() + 
-  theme(panel.grid.major = element_blank(), 
-        panel.grid.minor = element_blank(),
-        panel.background = element_blank(),
-        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
-        axis.line.y = element_line(colour = "black"),
-        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
-        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
-        axis.line.x = element_line(colour = "black"),
-        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
-        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
-        strip.text.x = element_text(size = 32),
-        legend.key.size = unit(2, "cm"), # increase legend size
-        legend.text = element_text(size = rel(2)), # increase legend text
-        legend.title = element_text(size = rel(2.25)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal',
-        legend.box.background = element_rect(colour = "black") # make legend horizontal
-        ) # I don't think these are linear relationships so the smoother isn't terriibly helpful
-```

### 2024-07-16 noam_data_exploration_v3.Rmd
Added codyn measure of community stability, which is similar to how other papers (including Pete's) has defined portfolio effects. This is just 1/CV.

@@
 
-alpha_diversity$Cover <- rowSums(wide_transect_level[5:74])/100
+alpha_diversity$Cover <- rowSums(wide_transect_level[ , -which(names(wide_transect_level) %in% meta_cols)])/100
+alpha_diversity$Stratum <- paste(alpha_diversity$Site, "-", alpha_diversity$Habitat, "-", alpha_diversity$Transect)
 
 diversity_stability <- alpha_diversity %>%
-  dplyr::group_by(Site, Habitat, Site_habitat, Transect) %>%
+  dplyr::group_by(Site, Habitat, Site_habitat, Transect, Stratum) %>%
   dplyr::summarise(Richness_mean = mean(Richness), 
@@
             Cover_cv = CV(Cover))
+stability_transect <- community_stability(df = transect_level,
+                                          time.var = "Year",
+                                          abundance.var = "Mean_cover",
+                                          replicate.var = "Stratum") # using codyn measure of stability
+# NOTE: codyn community_stability() is just the inverse of CV
+diversity_stability <- merge(diversity_stability, stability_transect, by = "Stratum")
+library(ecofolio) # possible other method for examining portfolio effects
 

### 2024-07-17 noam_data_exploration_v3.Rmd
Added another possible glmm approach to analyzing adiv that's more analagous to what other studies have done

--- 2024-07-16 noam_data_exploration_v3.Rmd
@@
 plot(allEffects(richness_mod))
+
+# probably the closest to what others have done:
+richness_mod_2 <- glmmTMB(
+  Richness ~ Year_factor*Habitat + (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),
+  family = poisson,
+  data = alpha_diversity_quad,
+  control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000))
+) # overspecified but going with it for now
+summary(richness_mod_2) # looks like strong autocorrelation, which makes sense to me
+Anova(richness_mod_2)
+plot(allEffects(richness_mod_2))
+emmeans(richness_mod_2, pairwise ~ Habitat | Year_factor) # if we want to compare habitats year to year
 
@@
 
-alpha_diversity$Cover <- rowSums(wide_transect_level[5:74])/100
-
+alpha_diversity$Cover <- rowSums(wide_transect_level[ , -which(names(wide_transect_level) %in% meta_cols)])/100
+alpha_diversity$Stratum <- paste(alpha_diversity$Site, "-", alpha_diversity$Habitat, "-", alpha_diversity$Transect)
+# stability_transect <- community_stability(df = transect_level,
+#                                           time.var = "Year",
+#                                           abundance.var = "Mean_cover",
+#                                           replicate.var = "Stratum") # using codyn measure of stability
+# NOTE: codyn community_stability() is just the inverse of CV
 diversity_stability <- alpha_diversity %>%
-  dplyr::group_by(Site, Habitat, Site_habitat, Transect) %>%
+  dplyr::group_by(Site, Habitat, Site_habitat, Transect, Stratum) %>%
   dplyr::summarise(Richness_mean = mean(Richness), 
@@
             Shannon_cv = CV(Shannon),
+            Shannon_stability = 1/CV(Shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
             Cover_mean = mean(Cover),
             Cover_se = sd(Cover)/sqrt(n()),
-            Cover_cv = CV(Cover))
+            Cover_cv = CV(Cover),
+            Cover_stability = 1/CV(Cover)) # as above
+
+library(ecofolio) # possible other method for examining portfolio effects
 
@@
 ```
+
+```{r diversity stability quadrat level}
+
+```

### 2024-07-17 noam_data_exploration_v3.Rmd
Added Anderson et al. 2013 portfolio effects at transect level

--- 2024-07-16 noam_data_exploration_v3.Rmd
@@
 library(emmeans)
+library(ecofolio) # possible other method for examining portfolio effects 
 ```
@@
 
-meta_cols <- c(colnames(data[1:7] ), "Island_side", "Island_side_habitat")
+meta_cols <- c(colnames(data[1:7] ), "Island_side", "Island_side_habitat", "Stratum")
 species_cols <- colnames(data[8:77])
@@
 
+# probably the closest to what others have done:
+richness_mod_2 <- glmmTMB(
+  Richness ~ Year_factor*Habitat + (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),
+  family = poisson,
+  data = alpha_diversity_quad,
+  control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000))
+) # overspecified but going with it for now
+summary(richness_mod_2) # looks like strong autocorrelation, which makes sense to me
+Anova(richness_mod_2)
+plot(allEffects(richness_mod_2))
+emmeans(richness_mod_2, pairwise ~ Habitat | Year_factor) # if we want to compare habitats year to year
+
 ```
@@
 
-```{r turnover}
+```{r turnover, eval = FALSE}
 # the turnover function calculates turnover as the proportion or species that either appear or disappear between timepoints
@@
         ) # I don't think these are linear relationships so the smoother isn't terriibly helpful
+```
+
+```{r transect level portfolio effect}
+# using Anderson et al. 2013 definitions of portfolio effects - relevant fxns in ecofolio packages
+wide_transect_level$Stratum <- as.factor(paste(wide_transect_level$Site, "-", wide_transect_level$Habitat, "-", wide_transect_level$Transect))
+x <- subset(wide_transect_level, Stratum == "LTER 1 - Backreef - 1")
+
+y <- x[ , -which(names(x) %in% meta_cols)]
+y <- y[, which(colSums(y) != 0)]
+z <- pe_avg_cv(y, ci = TRUE, detrending = "loess_detrended")
+
+transect_portfolio_columns <- c("Stratum", "Portfolio_effect") # names of columns for dataframe
+transect_level_portfolio <- 
+  data.frame(matrix(nrow = length(unique(wide_transect_level$Stratum)), 
+                    ncol = length(transect_portfolio_columns))
+             )
+colnames(transect_level_portfolio) <- transect_portfolio_columns # change column 
+transect_level_portfolio$Stratum <- levels(wide_transect_level$Stratum) # populate
+
+for(i in 1:length(unique(unique(wide_transect_level$Stratum))) ){ # for each  transect
+  # create a new dataframe for current transect
+  newData <- subset(wide_transect_level, Stratum == levels(wide_transect_level$Stratum)[i])
+  current_matrix <- newData[ , -which(names(newData) %in% meta_cols)] # remove metadata columns
+  if(sum(current_matrix) == 0){ # if this was a 0 algae transect
+    transect_level_portfolio[i, 2] <- NA # set portfolio effects to 0
+  }else{ # otherwise
+    current_matrix <- current_matrix[, which(colSums(current_matrix) != 0)] # remove irrelevant algae
+
+    transect_level_portfolio[i, 2] <- pe_avg_cv(current_matrix) # calculate portfolio effects for current time series
+  }
+
+}
+
+```
+
+```{r CV, fig.height = 8, fig.width = 10}
+# coefficient of variation
+CV<-function(x){
+  return(sd(x,na.rm=T)/mean(x,na.rm=T))
+}
+
+alpha_diversity$Cover <- rowSums(wide_transect_level[ , -which(names(wide_transect_level) %in% meta_cols)])/100
+alpha_diversity$Stratum <- paste(alpha_diversity$Site, "-", alpha_diversity$Habitat, "-", alpha_diversity$Transect)
+lm_synchrony_transect <- codyn::synchrony(df = transect_level,
+                                   time.var = "Year",
+                                   species.var = "Taxa",
+                                   abundance.var = "Mean_cover",
+                                   metric = "Loreau",
+                                   replicate.var = "Stratum") # Loreau and Mazancourt synchrony
+colnames(lm_synchrony_transect) = c("Stratum", "Loreau_synchrony") # rename cols for merge later
+
+gross_synchrony_transect <- codyn::synchrony(df = transect_level,
+                                   time.var = "Year",
+                                   species.var = "Taxa",
+                                   abundance.var = "Mean_cover",
+                                   metric = "Gross",
+                                   replicate.var = "Stratum") # lots of warnings in this one - examine further later
+colnames(gross_synchrony_transect) = c("Stratum", "Gross_synchrony") # rename cols for merge later
+
+# stability_transect <- community_stability(df = transect_level,
+#                                           time.var = "Year",
+#                                           abundance.var = "Mean_cover",
+#                                           replicate.var = "Stratum") # using codyn measure of stability
+# NOTE: codyn community_stability() is just the inverse of CV
+diversity_stability <- alpha_diversity %>%
+  dplyr::group_by(Site, Habitat, Site_habitat, Transect, Stratum) %>%
+  dplyr::summarise(Richness_mean = mean(Richness), 
+            Richness_se = sd(Richness)/sqrt(n()),
+            Richness_cv = CV(Richness),
+            Shannon_mean = mean(Shannon), 
+            Shannon_se = sd(Shannon)/sqrt(n()),
+            Shannon_cv = CV(Shannon),
+            Shannon_stability = 1/CV(Shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
+            Cover_mean = mean(Cover),
+            Cover_se = sd(Cover)/sqrt(n()),
+            Cover_cv = CV(Cover),
+            Cover_stability = 1/CV(Cover)) # as above
+
+diversity_stability <- merge(diversity_stability, lm_synchrony_transect, by = "Stratum") # add in loreau and de Mazancourt synchrony
+diversity_stability <- merge(diversity_stability, gross_synchrony_transect, by = "Stratum")
+diversity_stability <- merge(diversity_stability, transect_level_portfolio, by = "Stratum") # add in portfolio effects
+pairs(~ Cover_mean + Cover_cv + Cover_stability + 
+        Shannon_mean + Shannon_cv + Shannon_stability + 
+        Loreau_synchrony + Gross_synchrony + Portfolio_effect, data = diversity_stability)
+
+diversity_mod_1 <- glmmTMB(Cover_cv ~ Richness_mean, family = Gamma(link = "log"), data = diversity_stability)
+summary(diversity_mod_1)
+plot(allEffects(diversity_mod_1)) # richness reduces cv (i.e. enhances stability!)
+
+diversity_mod_2 <- glmmTMB(Cover_cv ~ Shannon_mean, family = Gamma(link = "log"), data = diversity_stability)
+summary(diversity_mod_2) # similar results
+
+diversity_mod_3 <- glmmTMB(Cover_cv ~ Habitat*Site, family = Gamma(link = "log"), data = diversity_stability)
+summary(diversity_mod_3) # christ 
+Anova(diversity_mod_3) # all highly sig but what the fuck am I doing here
+
+diversity_mod_4 <- glmmTMB(Cover_cv ~ Habitat, family = Gamma(link = "log"), data = diversity_stability)
+summary(diversity_mod_4)
+Anova(diversity_mod_4)
+plot(allEffects(diversity_mod_4)) # greatest variance in the lagoon (makes sense given beta diversity trends above)
+
+diversity_mod_5 <- glmmTMB(Cover_cv ~ Site, family = Gamma(link = "log"), data = diversity_stability)
+summary(diversity_mod_5)
+Anova(diversity_mod_5)
+
+diversity_mod_6 <- glmmTMB(Cover_cv ~ Shannon_mean*Habitat + (1|Site), family = Gamma(link = "log"), data = diversity_stability)
+summary(diversity_mod_6) # negligible variance in random effects
+Anova(diversity_mod_6) # Diversity the strongest predictor of "stability" by far!
+
+diversity_mod_7 <- glmmTMB(Cover_cv ~ Shannon_mean*Site + (1|Site_habitat), family = Gamma(link = "log"), data = diversity_stability)
+summary(diversity_mod_7)
+Anova(diversity_mod_7)
+
+diversity_mod_universal <- glmmTMB(Cover_cv ~ Shannon_mean*Site*Habitat + (1|Site_habitat), family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
+diversity_dredge <- dredge(diversity_mod_universal)
+model.sel(diversity_dredge)
+
+# best model based on dredge:
+diversity_best <-  glmmTMB(Cover_cv ~ Shannon_mean*Site + Site*Habitat + (1|Site_habitat), family = Gamma(link = "log"), data = diversity_stability)
+summary(diversity_best) # very low variance in random effect, which is great
+Anova(diversity_best) # Interactions all significant
+plot(dp <- simulateResiduals(diversity_best)) # some outliers, but otherwise good model fit. I am fine with this. 
+r.squaredGLMM(diversity_best) # 0.77, 0.77
+
+(shannon_cv_plot = 
+  ggplot(data = diversity_stability) +
+  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 8) + 
+   scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+    stat_smooth(aes(x = Shannon_mean, y = Shannon_cv),method = "lm", se = T, size = 2) + 
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+(shannon_cv_plot_2 = 
+  ggplot(data = diversity_stability) +
+  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+# maybe some heteroskedasticity?
+# just eyeballing but it looks like "stability" is highest in LTER 2 Fringe, which also has low diversity
+# Functional form vs. portfolio effects?
+# Either way looks sick as shit. Great clustering by habitat within site, and diversity trend is strong in LTER 3 4 5
+# LTER2 the trend is reversed
+# LTER 1 is mostly driven by habitat
+# LTER 6 potentially a wash
+
+# GOALS: test portfolio effects, look at turnover and temporal beta diversity stats, ways to test beta dispersion
+
+```
+
+```{r variance vs mean diversity plot, fig.height = 8, fig.width = 10}
+(diversity_mean_cv_plot = 
+  ggplot(data = diversity_stability) +
+  geom_point(aes(x = Shannon_mean, y = Shannon_cv, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+# More diverse regions are also those that fluctuate the least in their diversity
+
+(diversity_mean_cv_plot_2 = 
+  ggplot(data = diversity_stability) +
+  geom_point(aes(x = Shannon_mean, y = Shannon_cv, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+
+#  facet_wrap(~Site) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         ) # awesome relationship
+```
+
+```{r general cover diversity, fig.height = 30, fig.width = 30}
+(diversity_cover_plot = 
+  ggplot(data = alpha_diversity) +
+  geom_point(aes(x = Cover, y = Shannon, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Year) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         ) # ABORT ABORT ABORT
+```
+
+```{r diversity stability quadrat level}
+
 ```
@@
 ```
-
-```{r CV, fig.height = 8, fig.width = 10}
-# coefficient of variation
-CV<-function(x){
-  return(sd(x,na.rm=T)/mean(x,na.rm=T))
-}
-
-alpha_diversity$Cover <- rowSums(wide_transect_level[5:74])/100
-
-diversity_stability <- alpha_diversity %>%
-  dplyr::group_by(Site, Habitat, Site_habitat, Transect) %>%
-  dplyr::summarise(Richness_mean = mean(Richness), 
-            Richness_se = sd(Richness)/sqrt(n()),
-            Richness_cv = CV(Richness),
-            Shannon_mean = mean(Shannon), 
-            Shannon_se = sd(Shannon)/sqrt(n()),
-            Shannon_cv = CV(Shannon),
-            Cover_mean = mean(Cover),
-            Cover_se = sd(Cover)/sqrt(n()),
-            Cover_cv = CV(Cover))
-
-diversity_mod_1 <- glmmTMB(Cover_cv ~ Richness_mean, family = Gamma(link = "log"), data = diversity_stability)
-summary(diversity_mod_1)
-plot(allEffects(diversity_mod_1)) # richness reduces cv (i.e. enhances stability!)
-
-diversity_mod_2 <- glmmTMB(Cover_cv ~ Shannon_mean, family = Gamma(link = "log"), data = diversity_stability)
-summary(diversity_mod_2) # similar results
-
-diversity_mod_3 <- glmmTMB(Cover_cv ~ Habitat*Site, family = Gamma(link = "log"), data = diversity_stability)
-summary(diversity_mod_3) # christ 
-Anova(diversity_mod_3) # all highly sig but what the fuck am I doing here
-
-diversity_mod_4 <- glmmTMB(Cover_cv ~ Habitat, family = Gamma(link = "log"), data = diversity_stability)
-summary(diversity_mod_4)
-Anova(diversity_mod_4)
-plot(allEffects(diversity_mod_4)) # greatest variance in the lagoon (makes sense given beta diversity trends above)
-
-diversity_mod_5 <- glmmTMB(Cover_cv ~ Site, family = Gamma(link = "log"), data = diversity_stability)
-summary(diversity_mod_5)
-Anova(diversity_mod_5)
-
-diversity_mod_6 <- glmmTMB(Cover_cv ~ Shannon_mean*Habitat + (1|Site), family = Gamma(link = "log"), data = diversity_stability)
-summary(diversity_mod_6) # negligible variance in random effects
-Anova(diversity_mod_6) # Diversity the strongest predictor of "stability" by far!
-
-diversity_mod_7 <- glmmTMB(Cover_cv ~ Shannon_mean*Site + (1|Site_habitat), family = Gamma(link = "log"), data = diversity_stability)
-summary(diversity_mod_7)
-Anova(diversity_mod_7)
-
-diversity_mod_universal <- glmmTMB(Cover_cv ~ Shannon_mean*Site*Habitat + (1|Site_habitat), family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
-diversity_dredge <- dredge(diversity_mod_universal)
-model.sel(diversity_dredge)
-
-# best model based on dredge:
-diversity_best <-  glmmTMB(Cover_cv ~ Shannon_mean*Site + Site*Habitat + (1|Site_habitat), family = Gamma(link = "log"), data = diversity_stability)
-summary(diversity_best) # very low variance in random effect, which is great
-Anova(diversity_best) # Interactions all significant
-plot(dp <- simulateResiduals(diversity_best)) # some outliers, but otherwise good model fit. I am fine with this. 
-r.squaredGLMM(diversity_best) # 0.77, 0.77
-
-(shannon_cv_plot = 
-  ggplot(data = diversity_stability) +
-  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 8) + 
-   scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-    stat_smooth(aes(x = Shannon_mean, y = Shannon_cv),method = "lm", se = T, size = 2) + 
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-(shannon_cv_plot_2 = 
-  ggplot(data = diversity_stability) +
-  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site) +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-# maybe some heteroskedasticity?
-# just eyeballing but it looks like "stability" is highest in LTER 2 Fringe, which also has low diversity
-# Functional form vs. portfolio effects?
-# Either way looks sick as shit. Great clustering by habitat within site, and diversity trend is strong in LTER 3 4 5
-# LTER2 the trend is reversed
-# LTER 1 is mostly driven by habitat
-# LTER 6 potentially a wash
-
-# GOALS: test portfolio effects, look at turnover and temporal beta diversity stats, ways to test beta dispersion
-
-```
-
-```{r variance vs mean diversity plot, fig.height = 8, fig.width = 10}
-(diversity_mean_cv_plot = 
-  ggplot(data = diversity_stability) +
-  geom_point(aes(x = Shannon_mean, y = Shannon_cv, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site) +
-
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-# More diverse regions are also those that fluctuate the least in their diversity
-
-(diversity_mean_cv_plot_2 = 
-  ggplot(data = diversity_stability) +
-  geom_point(aes(x = Shannon_mean, y = Shannon_cv, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-
-#  facet_wrap(~Site) +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         ) # awesome relationship
-```
-
-```{r general cover diversity, fig.height = 30, fig.width = 30}
-(diversity_cover_plot = 
-  ggplot(data = alpha_diversity) +
-  geom_point(aes(x = Cover, y = Shannon, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Year) +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         ) # ABORT ABORT ABORT
-```

### 2024-07-17 noam_data_exploration_v3.Rmd
Looked at transect level diversity, stability, synchrony stuff. Redid model selection at transect level. Need to sort issues for work at quadrat level.

@@
 wide_transect_level$Stratum <- as.factor(paste(wide_transect_level$Site, "-", wide_transect_level$Habitat, "-", wide_transect_level$Transect))
-x <- subset(wide_transect_level, Stratum == "LTER 1 - Backreef - 1")
-
-y <- x[ , -which(names(x) %in% meta_cols)]
-y <- y[, which(colSums(y) != 0)]
-z <- pe_avg_cv(y, ci = TRUE, detrending = "loess_detrended")
 
@@
         Loreau_synchrony + Gross_synchrony + Portfolio_effect, data = diversity_stability)
-
-diversity_mod_1 <- glmmTMB(Cover_cv ~ Richness_mean, family = Gamma(link = "log"), data = diversity_stability)
-summary(diversity_mod_1)
-plot(allEffects(diversity_mod_1)) # richness reduces cv (i.e. enhances stability!)
-
-diversity_mod_2 <- glmmTMB(Cover_cv ~ Shannon_mean, family = Gamma(link = "log"), data = diversity_stability)
-summary(diversity_mod_2) # similar results
-
-diversity_mod_3 <- glmmTMB(Cover_cv ~ Habitat*Site, family = Gamma(link = "log"), data = diversity_stability)
-summary(diversity_mod_3) # christ 
-Anova(diversity_mod_3) # all highly sig but what the fuck am I doing here
-
-diversity_mod_4 <- glmmTMB(Cover_cv ~ Habitat, family = Gamma(link = "log"), data = diversity_stability)
-summary(diversity_mod_4)
-Anova(diversity_mod_4)
-plot(allEffects(diversity_mod_4)) # greatest variance in the lagoon (makes sense given beta diversity trends above)
-
-diversity_mod_5 <- glmmTMB(Cover_cv ~ Site, family = Gamma(link = "log"), data = diversity_stability)
-summary(diversity_mod_5)
-Anova(diversity_mod_5)
-
-diversity_mod_6 <- glmmTMB(Cover_cv ~ Shannon_mean*Habitat + (1|Site), family = Gamma(link = "log"), data = diversity_stability)
-summary(diversity_mod_6) # negligible variance in random effects
-Anova(diversity_mod_6) # Diversity the strongest predictor of "stability" by far!
-
-diversity_mod_7 <- glmmTMB(Cover_cv ~ Shannon_mean*Site + (1|Site_habitat), family = Gamma(link = "log"), data = diversity_stability)
-summary(diversity_mod_7)
-Anova(diversity_mod_7)
-
-diversity_mod_universal <- glmmTMB(Cover_cv ~ Shannon_mean*Site*Habitat + (1|Site_habitat), family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
-diversity_dredge <- dredge(diversity_mod_universal)
-model.sel(diversity_dredge)
-
-# best model based on dredge:
-diversity_best <-  glmmTMB(Cover_cv ~ Shannon_mean*Site + Site*Habitat + (1|Site_habitat), family = Gamma(link = "log"), data = diversity_stability)
-summary(diversity_best) # very low variance in random effect, which is great
-Anova(diversity_best) # Interactions all significant
-plot(dp <- simulateResiduals(diversity_best)) # some outliers, but otherwise good model fit. I am fine with this. 
-r.squaredGLMM(diversity_best) # 0.77, 0.77
-
+```
+
+```{r basic diversity stability plots transect level, fig.height=8, fig.width=10}
 (shannon_cv_plot = 
   ggplot(data = diversity_stability) +
-  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 8) + 
-   scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-    stat_smooth(aes(x = Shannon_mean, y = Shannon_cv),method = "lm", se = T, size = 2) + 
+  geom_point(aes(x = Shannon_mean, y = Cover_stability, colour = Site, shape = Habitat), size = 8) + 
+   scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+#    stat_smooth(aes(x = Shannon_mean, y = Shannon_cv),method = "lm", se = T, size = 2) + 
   theme_bw() + 
@@
   ggplot(data = diversity_stability) +
-  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+  geom_point(aes(x = Shannon_mean, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
    scale_colour_manual(values = site_colours) +
@@
 
-# GOALS: test portfolio effects, look at turnover and temporal beta diversity stats, ways to test beta dispersion
-
 ```
@@
 
-```{r general cover diversity, fig.height = 30, fig.width = 30}
-(diversity_cover_plot = 
-  ggplot(data = alpha_diversity) +
-  geom_point(aes(x = Cover, y = Shannon, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Year) +
+```{r portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
+(loreau_stability_plot = 
+  ggplot(data = diversity_stability) +
+  geom_point(aes(x = Loreau_synchrony, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site) +
+
   theme_bw() + 
@@
         legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+(gross_stability_plot = 
+  ggplot(data = diversity_stability) +
+  geom_point(aes(x = Gross_synchrony, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+(portfolio_stability_plot = 
+  ggplot(data = diversity_stability) +
+  geom_point(aes(x = Portfolio_effect, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         ) # almost a perfect line, which makes sense given that CV is a component of portfolio effects
+```
+
+```{r diversity, portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
+(loreau_diversity_plot = 
+  ggplot(data = diversity_stability) +
+  geom_point(aes(x = Shannon_mean, y = Loreau_synchrony, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+(gross_diversity_plot = 
+  ggplot(data = diversity_stability) +
+  geom_point(aes(x = Shannon_mean, y = Gross_synchrony, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+(portfolio_diversity_plot = 
+  ggplot(data = diversity_stability) +
+  geom_point(aes(x = Shannon_mean, y = Portfolio_effect, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+```
+
+```{r model selection}
+diversity_mod_universal <- glmmTMB(Cover_stability ~ Loreau_synchrony + Shannon_mean*Site*Habitat, family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
+diversity_dredge <- dredge(diversity_mod_universal)
+model.sel(diversity_dredge)
+
+# top model is as follows (by a decent margin):
+diversity_top <- glmmTMB(Cover_stability ~ Loreau_synchrony + Shannon_mean*Site + Site*Habitat, family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
+summary(diversity_top)
+Anova(diversity_top)
+```
+
+```{r general cover diversity, fig.height = 30, fig.width = 30}
+(diversity_cover_plot = 
+  ggplot(data = alpha_diversity) +
+  geom_point(aes(x = Cover, y = Shannon, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Year) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
          ) # ABORT ABORT ABORT
@@
 
-```{r diversity stability quadrat level}
+```{r diversity stability quadrat level setup}
+data$Stratum <- as.factor(paste(data$Site, "-", data$Habitat, "-", data$Transect))
+data$Location <- as.factor(data$Location)
+
+quadrat_portfolio_columns <- c("Location", "Portfolio_effect") # names of columns for dataframe
+quadrat_level_portfolio <- 
+  data.frame(matrix(nrow = length(unique(data$Location)), 
+                    ncol = length(quadrat_portfolio_columns))
+             )
+colnames(quadrat_level_portfolio) <- quadrat_portfolio_columns # change column 
+quadrat_level_portfolio$Location <- unique(data$Location) # populate
+
+for(i in 1:length(unique(unique(data$Location))) ){ # for each  quadrat
+  # create a new dataframe for current quadrat
+  newData <- subset(data, Location == levels(data$Location)[i])
+  current_matrix <- newData[ , -which(names(newData) %in% meta_cols)] # remove metadata columns
+  if(sum(current_matrix) == 0){ # if this was a 0 algae quadrat
+    quadrat_level_portfolio[i, 2] <- NA # set portfolio effects to 0
+  }else{ # otherwise
+    current_matrix <- current_matrix[, which(colSums(current_matrix) != 0)] # remove irrelevant algae
+
+    quadrat_level_portfolio[i, 2] <- pe_avg_cv(current_matrix) # calculate portfolio effects for current time series
+  }
+
+}
 

### 2024-07-17 noam_data_exploration_v3.Rmd
Figured out how to make portfolio effects work at quadrat level, but the data are weird because of too many quadrats with extremely low richness

@@
 
-problem_children <- c("LTER 5 Fringing Reef Algae Transect 3 Quad 9", "LTER 5 Fringing Reef Algae Transect 3 Quad 8", "LTER 5 Fringing Reef Algae Transect 3 Quad 6", "LTER 5 Fringing Reef Algae Transect 3 Quad 3", "LTER 5 Fringing Reef Algae Transect 3 Quad 10")
+problem_children <- c("LTER 5 Fringing Reef Algae Transect 2 Quad 9", "LTER 5 Fringing Reef Algae Transect 3 Quad 9", "LTER 5 Fringing Reef Algae Transect 3 Quad 8", "LTER 5 Fringing Reef Algae Transect 3 Quad 6", "LTER 5 Fringing Reef Algae Transect 3 Quad 3", "LTER 5 Fringing Reef Algae Transect 3 Quad 10")
 
@@
   current_matrix <- newData[ , -which(names(newData) %in% meta_cols)] # remove metadata columns
-  if(sum(current_matrix) == 0){ # if this was a 0 algae quadrat
+  if(sum(current_matrix) == 0 | # if this was a 0 algae quadrat
+     # or if there was only 1 column after removing all irrelevant algae
+     is.null(ncol(current_matrix[, which(colSums(current_matrix) != 0)]))){ 
     quadrat_level_portfolio[i, 2] <- NA # set portfolio effects to 0
@@
 }
+
+sum(is.na(quadrat_level_portfolio$Portfolio_effect)) # 4 quadrats causing trouble
+problem_children <- droplevels(quadrat_level_portfolio[is.na(quadrat_level_portfolio$Portfolio_effect),]$Location)
+hist(quadrat_level_portfolio$Portfolio_effect) # some crazy outliers due to low richness
 

### 2024-07-17 noam_data_exploration_v3.Rmd
Rearranged a couple chunks

@@
 
+```
+```{r gam approach}
+# NOTE: Not sure where I'm going with this yet
+# Idea is to look at smoother estimates for each time series to get an idea for the "shape" of the time series
+# Not sure if worthwhile, might skip
+
+# The below takes 17 years to run in and of itself
+# gam_model <- gam(Richness ~ s(Year) + s(Year, by = Habitat) + 
+#                  s(Island_side, bs = "re") + s(Transect, bs = "re") + s(Location, bs = "re"),
+#                  data = alpha_diversity_quad, family = poisson())
+
+gam_columns <- c("Location", "Intercept", "edf") # names of columns for dataframe
+gam_coefficients <- # new dataframe name - same workflow as above
+  data.frame(matrix(nrow = length(unique(alpha_diversity_quad$Location)),
+                    ncol = length(gam_columns))
+             )
+colnames(gam_coefficients) <- gam_columns # change column names to what was specified
+gam_coefficients$Location <- levels(alpha_diversity_quad$Location) # populate Location column with each quad/time series
+
+# same workflow as *below (I did gls first originally)
+for(i in 1:length(unique(alpha_diversity_quad$Location)) ){ # for each  quadrat
+  # create a new dataframe for the quadrat of the current iteration
+  newData <- subset(alpha_diversity_quad, Location == levels(alpha_diversity_quad$Location)[i])
+
+  # fit gam for current time series
+  loop_mod <- gamm(Richness ~ s(Year), data = newData, correlation = corCAR1(form = ~ Year), method = "REML", family = poisson)
+  loop_sum <- summary(loop_mod$gam) # store summary data
+  # populate appropriate row and column with the relevant coeffcients from gls
+  gam_coefficients[i, 2] <- unname(loop_sum$p.coeff)
+  gam_coefficients[i, 3] <- unname(loop_sum$edf)
+
+}
+
+hist(gam_coefficients$Intercept)
+hist(gam_coefficients$edf) # generally around 1 except for an outlier
+# LTER 5 Fringing Reef Algae Transect 3 Quad 6 - a quadrat that has almost no algae the entire time series
+# issues with quadrat - smaller sample = more 0s, makes modeling trickier
+hist(subset(gam_coefficients, Location != "LTER 5 Fringing Reef Algae Transect 3 Quad 6")$edf)
+sum(gam_coefficients$edf > 1.01) # all but 18/1200 quadrats can be fit by a smoother with edf < 1.01
+# trends more straightforward than I anticipated?
+# argument for just looking at mean richness and slope?
 ```
@@
 
-```
-
-```{r gam approach}
-# NOTE: Not sure where I'm going with this yet
-# Idea is to look at smoother estimates for each time series to get an idea for the "shape" of the time series
-# Not sure if worthwhile, might skip
-
-# The below takes 17 years to run in and of itself
-# gam_model <- gam(Richness ~ s(Year) + s(Year, by = Habitat) + 
-#                  s(Island_side, bs = "re") + s(Transect, bs = "re") + s(Location, bs = "re"),
-#                  data = alpha_diversity_quad, family = poisson())
-
-gam_columns <- c("Location", "Intercept", "edf") # names of columns for dataframe
-gam_coefficients <- # new dataframe name - same workflow as above
-  data.frame(matrix(nrow = length(unique(alpha_diversity_quad$Location)),
-                    ncol = length(gam_columns))
-             )
-colnames(gam_coefficients) <- gam_columns # change column names to what was specified
-gam_coefficients$Location <- levels(alpha_diversity_quad$Location) # populate Location column with each quad/time series
-
-# same workflow as above
-for(i in 1:length(unique(alpha_diversity_quad$Location)) ){ # for each  quadrat
-  # create a new dataframe for the quadrat of the current iteration
-  newData <- subset(alpha_diversity_quad, Location == levels(alpha_diversity_quad$Location)[i])
-
-  # fit gam for current time series
-  loop_mod <- gamm(Richness ~ s(Year), data = newData, correlation = corCAR1(form = ~ Year), method = "REML", family = poisson)
-  loop_sum <- summary(loop_mod$gam) # store summary data
-  # populate appropriate row and column with the relevant coeffcients from gls
-  gam_coefficients[i, 2] <- unname(loop_sum$p.coeff)
-  gam_coefficients[i, 3] <- unname(loop_sum$edf)
-
-}
-
-hist(gam_coefficients$Intercept)
-hist(gam_coefficients$edf) # generally around 1 except for an outlier
-# LTER 5 Fringing Reef Algae Transect 3 Quad 6 - a quadrat that has almost no algae the entire time series
-# issues with quadrat - smaller sample = more 0s, makes modeling trickier
-hist(subset(gam_coefficients, Location != "LTER 5 Fringing Reef Algae Transect 3 Quad 6")$edf)
-sum(gam_coefficients$edf > 1.01) # all but 18/1200 quadrats can be fit by a smoother with edf < 1.01
-# trends more straightforward than I anticipated?
-# argument for just looking at mean richness and slope?
 ```

### 2024-07-18 noam_data_exploration_v3.Rmd
Subsetted quad level data according to what Lauren did (I messed it up somehow) and ran total turnover + some exploratory stats at quad level.

@@
 
-problem_children <- c("LTER 5 Fringing Reef Algae Transect 2 Quad 9", "LTER 5 Fringing Reef Algae Transect 3 Quad 9", "LTER 5 Fringing Reef Algae Transect 3 Quad 8", "LTER 5 Fringing Reef Algae Transect 3 Quad 6", "LTER 5 Fringing Reef Algae Transect 3 Quad 3", "LTER 5 Fringing Reef Algae Transect 3 Quad 10")
-
-long_data_sub <- long_data[long_data$Location != problem_children, ] # remove problem children
-# still get the same error when I try to run at quadrat level
-# I think it's because of the 0s
-# Error in names(output) <- `*vtmp*` : 
-#  'names' attribute [1] must be the same length as the vector [0]
-```
-
-```{r turnover, eval = FALSE}
+problem_children <- c(
+                      "LTER 5 Fringing Reef Algae Transect 3 Quad 9",
+                      "LTER 5 Fringing Reef Algae Transect 3 Quad 8",
+                      "LTER 5 Fringing Reef Algae Transect 3 Quad 6",
+                      "LTER 5 Fringing Reef Algae Transect 3 Quad 3",
+                      "LTER 5 Fringing Reef Algae Transect 3 Quad 10")
+
+long_data_sub <- long_data[!long_data$Location %in% problem_children, ] # remove problem children
+
+data_sub <- data[!data$Location %in% problem_children, ]
+
+```
+
+```{r total turnover, fig.height = 24, fig.width=16}
 # the turnover function calculates turnover as the proportion or species that either appear or disappear between timepoints
@@
 
-total_turnover<- 
+# get relevant metadata for turnover metrics
+data_sub <- data[paste(data$Year, data$Location) %in% paste(total_turnover$Year, total_turnover$Location), ]
+
+# check to make sure everything is the same length
+# nrow(data_sub) # 17829
+# nrow(total_turnover) # 17829
+
+# extract metadata
+turnover_meta <- data_sub[ ,which(names(data_sub) %in% meta_cols)]
+
+# merge with turnover df
+total_turnover <- merge(total_turnover, turnover_meta, by = c("Year", "Location"))
+total_turnover$Year_factor <- as.factor(total_turnover$Year)
+
+# turnover bounded between 0 and 1 --> beta regression
+# need to transform due to presence of 0s and 1s
+# Smithson verkuilen transformation for beta-regression
+sv_trans = function(prop, N, s = 0.000005){
+  (prop*(nrow(N) - 1) + s)/nrow(N)
+  # where prop is the proportional value you're transforming, 
+  # N is the sample size, which is specified by taking the number of rows/observations from a given dataframe,
+  # and s is a small offset 
+}
+total_turnover$Turnover_trans <- sv_trans(total_turnover$total, total_turnover)
+
+# model
+total_turnover_mod1 <- glmmTMB(Turnover_trans ~ Year_factor*Habitat*Site +
+                                 (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),
+                               family = beta_family(), data = total_turnover)
+summary(total_turnover_mod1) # low temporal autocorrelation... maybe don't need ar1 term?
+plot(total_turnover_res <- simulateResiduals(total_turnover_mod1)) # residuals look very weird
+Anova(total_turnover_mod1) # obviously everything sig
+
+total_turnover_mod2 <- glmmTMB(Turnover_trans ~ Year_factor*Habitat*Site + (1 | Site/Transect/Location),
+                               family = beta_family(), data = total_turnover)
+summary(total_turnover_mod2) # 
+plot(total_turnover_res2 <- simulateResiduals(total_turnover_mod2)) # residuals still look like shit
+Anova(total_turnover_mod2) 
+anova(total_turnover_mod2, total_turnover_mod1) # models are sig different from each other
+AICc(total_turnover_mod2, total_turnover_mod1) # first model is better
+# need to figure out residuals problem
+plot(allEffects(total_turnover_mod1))
+
+# plot
+total_turnover_means <-
   total_turnover %>%
-    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
-total_turnover$Site_habitat <- paste(total_turnover$Site, total_turnover$Habitat)
-# Add back in stratum column
-
-# A lot of extra spaces
-
-total_turnover <- total_turnover %>%
-  mutate(across(everything(), str_trim))
-
-str(total_turnover) # everything became characters
-
-total_turnover$total <- as.numeric(total_turnover$total)
-total_turnover$Year <- as.factor(total_turnover$Year)
-total_turnover$Site <- as.factor(total_turnover$Site)
-total_turnover$Habitat <- as.factor(total_turnover$Habitat)
-
-turnover_means <-
-  total_turnover %>%
-  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
+  dplyr::group_by(Year, Island_side, Site, Habitat, Site_habitat) %>%
   dplyr::summarise(Turnover_mean = mean(total), 
@@
 (turnover_plot <-
-  ggplot(data = turnover_means) +
+  ggplot(data = total_turnover_means) +
   geom_point(aes(x = Year, y = Turnover_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
@@
 #
-
+```
+
+```{r appearances}
 # examine appearance and disappearances independently 

### 2024-07-18 noam_data_exploration_v3.Rmd
Worked through turnover, rank shift, and rate change at quadrat level. Rate change is dependent on transformed vs untransformed data. Currently it's based on Euclidean distances, which I don't love. Would like to dig into base code of codyn to see if I can change to Jaccard or Bray-Curtis dissimilarity

@@
 
-```{r appearances}
+```{r appearances, fig.height = 24, fig.width=16}
 # examine appearance and disappearances independently 
@@
   turnover(
-  df = long_data,
+  df = long_data_sub,
   time.var = "Year",
@@
 
-appearance<- 
-  appearance %>%
-    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
-appearance$Site_habitat <- paste(appearance$Site, appearance$Habitat)
-# Add back in stratum column
-
-# A lot of extra spaces
-
-appearance <- appearance %>%
-  mutate(across(everything(), str_trim))
-
-str(appearance) # everything became characters
-
-appearance$appearance <- as.numeric(appearance$appearance)
-appearance$Year <- as.factor(appearance$Year)
-appearance$Site <- as.factor(appearance$Site)
-appearance$Habitat <- as.factor(appearance$Habitat)
+appearance <- merge(turnover_meta, appearance, by = c("Year", "Location"))
+appearance$Appearance_trans <- sv_trans(appearance$appearance, appearance)
+appearance$Year_factor <- as.factor(appearance$Year)
+
+# model
+appearance_turnover_mod1 <- glmmTMB(Appearance_trans ~ Year_factor*Habitat*Site +
+                                 (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),
+                               family = beta_family(), data = appearance)
+summary(appearance_turnover_mod1) # low temporal autocorrelation and it's negative
+plot(appearance_turnover_res <- simulateResiduals(appearance_turnover_mod1)) # similar residuals problem
+Anova(appearance_turnover_mod1) # obviously everything sig
+
+appearance_turnover_mod2 <- glmmTMB(Appearance_trans ~ Year_factor*Habitat*Site + (1 | Site/Transect/Location),
+                               family = beta_family(), data = appearance)
+summary(appearance_turnover_mod2) # 
+plot(appearance_turnover_res2 <- simulateResiduals(appearance_turnover_mod2)) #
+Anova(appearance_turnover_mod2) 
+anova(appearance_turnover_mod2, appearance_turnover_mod1) # models are sig different from each other
+AICc(appearance_turnover_mod2, appearance_turnover_mod1) # first model is better but not as drastic as total
+
+plot(allEffects(appearance_turnover_mod1))
 
@@
 #
-
+```
+
+```{r disappearances, fig.height = 24, fig.width=16}
 disappearance <-
   turnover(
-  df = long_data,
+  df = long_data_sub,
   time.var = "Year",
   species.var = "Taxa",
-  abundance.var = "Mean_cover",
-  replicate.var = "Stratum",
+  abundance.var = "Percent_cover",
+  replicate.var = "Location",
   metric = "disappearance"
@@
 
-disappearance<- 
-  disappearance %>%
-    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
-disappearance$Site_habitat <- paste(disappearance$Site, disappearance$Habitat)
-# Add back in stratum column
-
-# A lot of extra spaces
-
-disappearance <- disappearance %>%
-  mutate(across(everything(), str_trim))
-
-str(disappearance) # everything became characters
-
-disappearance$disappearance <- as.numeric(disappearance$disappearance)
-disappearance$Year <- as.factor(disappearance$Year)
-disappearance$Site <- as.factor(disappearance$Site)
-disappearance$Habitat <- as.factor(disappearance$Habitat)
+disappearance <- merge(turnover_meta, disappearance, by = c("Year", "Location"))
+disappearance$Disappearance_trans <- sv_trans(disappearance$disappearance, disappearance)
+disappearance$Year_factor <- as.factor(disappearance$Year)
+
+# model
+disappearance_turnover_mod1 <- glmmTMB(Disappearance_trans ~ Year_factor*Habitat*Site +
+                                 (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),
+                               family = beta_family(), data = disappearance)
+summary(disappearance_turnover_mod1) # strong autocorrelation here
+plot(disappearance_turnover_res <- simulateResiduals(disappearance_turnover_mod1)) # 
+Anova(disappearance_turnover_mod1) #
+
+disappearance_turnover_mod2 <- glmmTMB(Disappearance_trans ~ Year_factor*Habitat*Site + (1 | Site/Transect/Location),
+                               family = beta_family(), data = disappearance)
+summary(disappearance_turnover_mod2) # 
+plot(disappearance_turnover_res2 <- simulateResiduals(disappearance_turnover_mod2)) #
+Anova(disappearance_turnover_mod2) 
+anova(disappearance_turnover_mod2, disappearance_turnover_mod1) # only case where models are not sig different from each other
+AICc(disappearance_turnover_mod2, disappearance_turnover_mod1) # models almost equivalent
+
+plot(allEffects(disappearance_turnover_mod1))
 
@@
     )
-
+```
+
+```{r mean rank shift, fig.height = 24, fig.width=16}
 # rank shift
@@
   rank_shift(
-  df = long_data,
+  df = long_data_sub,
   time.var = "Year",
   species.var = "Taxa",
-  abundance.var = "Mean_cover",
-  replicate.var = "Stratum"
-)
-
-rank_shift<- 
-  rank_shift %>%
-    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
-rank_shift$Site_habitat <- paste(rank_shift$Site, rank_shift$Habitat)
-# Add back in stratum column
-
-# A lot of extra spaces
-
-rank_shift <- rank_shift %>%
-  mutate(across(everything(), str_trim))
-
-str(rank_shift) # everything became characters
-
-rank_shift$MRS <- as.numeric(rank_shift$MRS)
-# Create a column with the final year from the returned time.var_pair
-rank_shift$Year <- as.numeric(substr(rank_shift$year_pair, 6, 9))
-rank_shift$Site <- as.factor(rank_shift$Site)
-rank_shift$Habitat <- as.factor(rank_shift$Habitat)
+  abundance.var = "Percent_cover",
+  replicate.var = "Location"
+)
+
+rank_shift <- merge(turnover_meta, rank_shift, by = c("Location"))
 
@@
             Rank_shift_se = sd(MRS)/sqrt(n()))
-
+# Looks very weird - not sure what to do with this
 (rank_shift_plot <-
@@
     )
-
+```
+
+```{r rate of change, fig.height=15, fig.width=15}
 # rate change
-rate_change_table<-
+long_data_sub$Percent_cover_transformed <- wisconsin(long_data_sub$Percent_cover) # wisconsin transform abundance data
+
+# NOTE: the results here vary considerably if you used transformed vs untransformed data
+# Looking into way to change distance matrix from Euclidiean to something else
+rate_change_table_transformed<-
   rate_change(
-  df = long_data,
+  df = long_data_sub,
   time.var = "Year",
   species.var = "Taxa",
-  abundance.var = "Mean_cover",
-  replicate.var = "Stratum"
-)
-# kable(rate_change_table)
-
-rate_change<-
-  rate_change_interval(
-  df = long_data,
+  abundance.var = "Percent_cover_transformed",
+  replicate.var = "Location"
+) # summarizes rate of change for each quadrat over the course of the time series
+
+rate_change_table_transformed <- merge(turnover_meta, rate_change_table_transformed, by = "Location")
+hist(rate_change_table_transformed$rate_change)
+
+rate_change_mod1 <- glmmTMB(rate_change ~ Habitat*Site + (1|Habitat/Site/Transect), data = rate_change_table_transformed)
+summary(rate_change_mod1)
+plot(rc_res <- simulateResiduals(rate_change_mod1)) # some deviations, but sig likely due to large n
+Anova(rate_change_mod1)
+#                Chisq Df Pr(>Chisq)    
+# Habitat      187.988  3  < 2.2e-16 ***
+# Site          81.209  5  4.686e-16 ***
+# Habitat:Site 143.638 15  < 2.2e-16 ***
+plot(allEffects(rate_change_mod1))
+# Outer 17 has fewest differences
+# LTER 1 2 3 changing more than 4 5 6 at outer10
+# Backreef change highest at LTER 1 2
+# Fringing highest at LTER 1 2 6
+emmeans(rate_change_mod1, pairwise ~ Site | Habitat)
+
+# Untransformed
+rate_change_table_untransformed<-
+  rate_change(
+  df = long_data_sub,
   time.var = "Year",
   species.var = "Taxa",
-  abundance.var = "Mean_cover",
-  replicate.var = "Stratum"
-)
-
-rate_change<- 
-  rate_change %>%
-    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
-rate_change$Site_habitat <- paste(rate_change$Site, rate_change$Habitat)
-
-# A lot of extra spaces
-
-rate_change <- rate_change %>%
-  mutate(across(everything(), str_trim))
-
-str(rate_change) # everything became characters
-
-rate_change$distance <- as.numeric(rate_change$distance)
-rate_change$interval <- as.numeric(rate_change$interval)
-rate_change$Site <- as.factor(rate_change$Site)
-rate_change$Habitat <- as.factor(rate_change$Habitat)
-
-ggplot(rate_change, aes(interval, distance, colour = Site, shape = Habitat, group = Site_habitat, linetype = Habitat)) +
-  facet_wrap(~Site) + 
-  geom_point() +
-  scale_colour_manual(values = site_colours) +
-  scale_linetype_manual(values = habitat_linetypes) +
-  stat_smooth(method = "lm", se = T, size = 2) +
-  scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site)+
-#  labs(y = "Distance\n") +
+  abundance.var = "Percent_cover",
+  replicate.var = "Location"
+) # summarizes rate of change for each quadrat over the course of the time series
+
+rate_change_table_untransformed <- merge(turnover_meta, rate_change_table_untransformed, by = "Location")
+hist(rate_change_table_untransformed$rate_change)
+
+rate_change_mod2 <- glmmTMB(rate_change ~ Habitat*Site + (1|Habitat/Site/Transect), data = rate_change_table_untransformed)
+summary(rate_change_mod2)
+plot(rc_res <- simulateResiduals(rate_change_mod2)) # as above
+Anova(rate_change_mod2)
+#                Chisq Df Pr(>Chisq)    
+# Habitat       39.224  3  1.556e-08 ***
+# Site          17.882  5   0.003098 ** 
+# Habitat:Site 123.014 15  < 2.2e-16 ***
+plot(allEffects(rate_change_mod1))
+# eyeballing it
+# Outer 17 scarcely changing regardless of site
+# Rate of change similar in backreef between sites
+# big variance between sites in outer10 and fringing
+# Outer 10: As you go around the island, sites become increasing more "stable" over time (i.e. less shift toward dissimilarity)
+# Effectively the opposite trend in Fringing - more dissimilar over time as you go around the island
+emmeans(rate_change_mod1, pairwise ~ Site | Habitat)
+
+ggplot(data = rate_change_table_transformed) +
+  geom_boxplot(aes(x = Site, y = rate_change, colour = Site) ) +
+  geom_jitter(aes(x = Site, y = rate_change, colour = Site), width = 0.2) +
+  facet_wrap(~Habitat) +
+  scale_colour_manual(values = site_colours) +
   theme_bw() + 
@@
         legend.box.background = element_rect(colour = "black") # make legend horizontal
-        ) # I don't think these are linear relationships so the smoother isn't terriibly helpful
+        )
+
+ggplot(data = rate_change_table_untransformed) +
+  geom_boxplot(aes(x = Site, y = rate_change, colour = Site) ) +
+  geom_jitter(aes(x = Site, y = rate_change, colour = Site), width = 0.2) +
+  facet_wrap(~Habitat) +
+  scale_colour_manual(values = site_colours) +
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+        axis.line.y = element_line(colour = "black"),
+        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        axis.line.x = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(size = 32),
+        legend.key.size = unit(2, "cm"), # increase legend size
+        legend.text = element_text(size = rel(2)), # increase legend text
+        legend.title = element_text(size = rel(2.25)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal',
+        legend.box.background = element_rect(colour = "black") # make legend horizontal
+        )
+
+```
+
+```{r rate change vis method 2, fig.height=15, fig.width=15}
+rate_change_untransformed<-
+  rate_change_interval(
+  df = long_data_sub,
+  time.var = "Year",
+  species.var = "Taxa",
+  abundance.var = "Percent_cover",
+  replicate.var = "Location"
+)
+
+rate_change_untransformed_distances <- merge(rate_change_untransformed, rate_change_meta, by = "Location")
+
+rate_change_transformed<-
+  rate_change_interval(
+  df = long_data_sub,
+  time.var = "Year",
+  species.var = "Taxa",
+  abundance.var = "Percent_cover_transformed",
+  replicate.var = "Location"
+)
+
+rate_change_transformed_distances <- merge(rate_change_transformed, rate_change_meta, by = "Location")
+
+ ggplot(rate_change_untransformed_distances,
+        aes(x = interval, y = distance, colour = Site, shape = Habitat) )+
+   facet_wrap(~Habitat) + 
+   geom_point() +
+   scale_colour_manual(values = site_colours) +
+   scale_linetype_manual(values = habitat_linetypes) +
+   stat_smooth(method = "lm", se = T, size = 2) +
+   scale_shape_manual(values = habitat_shapes) +
+   labs(y = "Untransformed distance\n") +
+   theme_bw() + 
+   theme(panel.grid.major = element_blank(), 
+         panel.grid.minor = element_blank(),
+         panel.background = element_blank(),
+         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+         axis.line.y = element_line(colour = "black"),
+         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         axis.line.x = element_line(colour = "black"),
+         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         strip.text.x = element_text(size = 32),
+         legend.key.size = unit(2, "cm"), # increase legend size
+         legend.text = element_text(size = rel(2)), # increase legend text
+         legend.title = element_text(size = rel(2.25)),
+         legend.position = 'bottom', # move legend to bottom
+         legend.box = 'horizontal',
+         legend.box.background = element_rect(colour = "black") # make legend horizontal
+         )
+
+ ggplot(rate_change_transformed_distances,
+        aes(x = interval, y = distance, colour = Site, shape = Habitat) )+
+   facet_wrap(~Habitat) + 
+   geom_point() +
+   scale_colour_manual(values = site_colours) +
+   scale_linetype_manual(values = habitat_linetypes) +
+   stat_smooth(method = "lm", se = T, size = 2) +
+   scale_shape_manual(values = habitat_shapes) +
+   labs(y = "Transformed distance\n") +
+   theme_bw() + 
+   theme(panel.grid.major = element_blank(), 
+         panel.grid.minor = element_blank(),
+         panel.background = element_blank(),
+         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+         axis.line.y = element_line(colour = "black"),
+         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         axis.line.x = element_line(colour = "black"),
+         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         strip.text.x = element_text(size = 32),
+         legend.key.size = unit(2, "cm"), # increase legend size
+         legend.text = element_text(size = rel(2)), # increase legend text
+         legend.title = element_text(size = rel(2.25)),
+         legend.position = 'bottom', # move legend to bottom
+         legend.box = 'horizontal',
+         legend.box.background = element_rect(colour = "black") # make legend horizontal
+         ) 
 ```

### 2024-07-19 noam_data_exploration_v3.Rmd
Codyn has a built in function to do what I want for temporal beta diversity but I can't run it (takes 13+ hr). Need to specify questions before subsetting.

@@
          ) 
+```
+
+```{r multivariate change}
+## Ran this for 13 hours and it didn't finish (obviously)
+## Need to subset or specify questions or try remote access on more powerful computer
+# multi_change_untransformed <-
+#   multivariate_change(
+#    df = long_data_sub,
+#    time.var = "Year",
+#    species.var = "Taxa",
+#    abundance.var = "Percent_cover",
+#    replicate.var = "Location"
+#   )
+#
+
 ```

### 2024-07-19 noam_data_exploration_v3.Rmd
Added diversity stability at the quadrat level. looks great!

@@
 
-richness_mod <- glmmTMB(
-  Richness ~ Island_side*Habitat + (1 | Island_side/Site/Transect/Location) + ar1(Year_factor + 0 | Location),
+# richness_mod <- glmmTMB(
+#   Richness ~ Island_side*Habitat + (1 | Island_side/Site/Transect/Location) + ar1(Year_factor + 0 | Location),
+#   family = poisson,
+#   data = alpha_diversity_quad,
+#   control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000))
+# ) # overspecified but going with it for now
+# summary(richness_mod) # looks like strong autocorrelation, which makes sense to me
+# Anova(richness_mod)
+# plot(allEffects(richness_mod))
+
+# probably the closest to what others have done:
+richness_mod_2 <- glmmTMB(
+  Richness ~ Year_factor*Habitat*Site + (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),
   family = poisson,
@@
 ) # overspecified but going with it for now
-summary(richness_mod) # looks like strong autocorrelation, which makes sense to me
-Anova(richness_mod)
-plot(allEffects(richness_mod))
-
-# probably the closest to what others have done:
-richness_mod_2 <- glmmTMB(
-  Richness ~ Year_factor*Habitat + (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),
-  family = poisson,
-  data = alpha_diversity_quad,
-  control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000))
-) # overspecified but going with it for now
+
 summary(richness_mod_2) # looks like strong autocorrelation, which makes sense to me
@@
 plot(allEffects(richness_mod_2))
-emmeans(richness_mod_2, pairwise ~ Habitat | Year_factor) # if we want to compare habitats year to year
+# emmeans(richness_mod_2, pairwise ~ Habitat | Year_factor) # if we want to compare habitats year to year
 
@@
 
-rate_change_untransformed_distances <- merge(rate_change_untransformed, rate_change_meta, by = "Location")
+rate_change_untransformed_distances <- merge(rate_change_untransformed, turnover_meta, by = "Location")
 
@@
 
-rate_change_transformed_distances <- merge(rate_change_transformed, rate_change_meta, by = "Location")
+rate_change_transformed_distances <- merge(rate_change_transformed, turnover_meta, by = "Location")
 
@@
 
-```{r diversity stability quadrat level setup}
-data$Stratum <- as.factor(paste(data$Site, "-", data$Habitat, "-", data$Transect))
-data$Location <- as.factor(data$Location)
+```{r diversity stability quadrat level setup, fig.height = 20, fig.width=20}
+data_sub$Stratum <- as.factor(paste(data_sub$Site, "-", data_sub$Habitat, "-", data_sub$Transect))
+data_sub$Location <- as.factor(data_sub$Location)
 
@@
 quadrat_level_portfolio <- 
-  data.frame(matrix(nrow = length(unique(data$Location)), 
+  data.frame(matrix(nrow = length(unique(data_sub$Location)), 
                     ncol = length(quadrat_portfolio_columns))
@@
 colnames(quadrat_level_portfolio) <- quadrat_portfolio_columns # change column 
-quadrat_level_portfolio$Location <- unique(data$Location) # populate
-
-for(i in 1:length(unique(unique(data$Location))) ){ # for each  quadrat
+quadrat_level_portfolio$Location <- unique(data_sub$Location) # populate
+
+for(i in 1:length(unique(unique(data_sub$Location))) ){ # for each  quadrat
   # create a new dataframe for current quadrat
-  newData <- subset(data, Location == levels(data$Location)[i])
+  newData <- subset(data, Location == levels(data_sub$Location)[i])
   current_matrix <- newData[ , -which(names(newData) %in% meta_cols)] # remove metadata columns
@@
 
-sum(is.na(quadrat_level_portfolio$Portfolio_effect)) # 4 quadrats causing trouble
-problem_children <- droplevels(quadrat_level_portfolio[is.na(quadrat_level_portfolio$Portfolio_effect),]$Location)
+# sum(is.na(quadrat_level_portfolio$Portfolio_effect)) # 4 quadrats causing trouble
+
 hist(quadrat_level_portfolio$Portfolio_effect) # some crazy outliers due to low richness
 
+alpha_diversity_quad_sub <- alpha_diversity_quad[!alpha_diversity_quad$Location %in% problem_children, ]
+
+alpha_diversity_quad_sub$Cover <- rowSums(data_sub[ , -which(names(data_sub) %in% meta_cols)])/100
+alpha_diversity_quad_sub$Stratum <- 
+  paste(alpha_diversity_quad_sub$Site, "-", alpha_diversity_quad_sub$Habitat, "-", alpha_diversity_quad_sub$Transect)
+lm_synchrony_quadrat <- codyn::synchrony(df = long_data_sub,
+                                   time.var = "Year",
+                                   species.var = "Taxa",
+                                   abundance.var = "Percent_cover",
+                                   metric = "Loreau",
+                                   replicate.var = "Location") # Loreau and Mazancourt synchrony
+colnames(lm_synchrony_quadrat) = c("Location", "Loreau_synchrony") # rename cols for merge later
+
+gross_synchrony_quadrat <- codyn::synchrony(df = long_data_sub,
+                                   time.var = "Year",
+                                   species.var = "Taxa",
+                                   abundance.var = "Percent_cover",
+                                   metric = "Gross",
+                                   replicate.var = "Location") # lots of warnings in this one - examine further later
+colnames(gross_synchrony_transect) = c("Location", "Gross_synchrony") # rename cols for merge later
+
+diversity_stability_quadrat <- alpha_diversity_quad_sub %>%
+  dplyr::group_by(Location, Site, Habitat, Site_habitat, Transect, Stratum) %>%
+  dplyr::summarise(Richness_mean = mean(Richness), 
+            Richness_se = sd(Richness)/sqrt(n()),
+            Richness_cv = CV(Richness),
+            Shannon_mean = mean(Shannon), 
+            Shannon_se = sd(Shannon)/sqrt(n()),
+            Shannon_cv = CV(Shannon),
+            Shannon_stability = 1/CV(Shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
+            Cover_mean = mean(Cover),
+            Cover_se = sd(Cover)/sqrt(n()),
+            Cover_cv = CV(Cover),
+            Cover_stability = 1/CV(Cover)) # as above
+
+diversity_stability_quadrat <- 
+  merge(diversity_stability_quadrat, lm_synchrony_quadrat, by = "Location") # add in loreau and de Mazancourt synchrony
+diversity_stability_quadrat <- merge(diversity_stability_quadrat, gross_synchrony_quadrat, by = "Location")
+diversity_stability_quadrat <- merge(diversity_stability_quadrat, quadrat_level_portfolio, by = "Location") # add in portfolio effects
+pairs(~ Cover_mean + Cover_cv + Cover_stability + 
+        Shannon_mean + Shannon_cv + Shannon_stability + 
+        Loreau_synchrony + Gross_synchrony + Portfolio_effect, data = diversity_stability)
+
+```
+
+```{r basic diversity stability plots quadrat level, fig.height=8, fig.width=10}
+(shannon_cv_plot = 
+  ggplot(data = diversity_stability_quadrat) +
+  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 8) + 
+   scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+#    stat_smooth(aes(x = Shannon_mean, y = Shannon_cv),method = "lm", se = T, size = 2) + 
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+(shannon_cv_plot_2 = 
+  ggplot(data = diversity_stability_quadrat) +
+  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 4, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+# LTER 2 obscures a lot of the trends
+
+```
+
+```{r portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
+(loreau_stability_plot_quad = 
+  ggplot(data = diversity_stability_quadrat) +
+  geom_point(aes(x = Loreau_synchrony, y = Cover_cv, colour = Site, shape = Habitat), size = 4, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+# (gross_stability_plot_quad = 
+#   ggplot(data = diversity_stability_quadrat) +
+#   geom_point(aes(x = Gross_synchrony, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+#    scale_colour_manual(values = site_colours) +
+#    scale_shape_manual(values = habitat_shapes) +
+#   facet_wrap(~Site) +
+#    
+#   theme_bw() + 
+#   theme(axis.title.x = element_text(size=18), 
+#         axis.title.y = element_text(size=18), 
+#         panel.background = element_blank(), 
+#         panel.grid.major = element_blank(),  #remove major-grid labels
+#         panel.grid.minor = element_blank(),  #remove minor-grid labels
+#         plot.background = element_blank(),
+#         legend.text = element_text(size = rel(1)), # increase legend text
+#         legend.title = element_text(size = rel(1)),
+#         legend.position = 'bottom', # move legend to bottom
+#         legend.box = 'horizontal', # make legend horizontal
+#         legend.box.background = element_rect(colour = "black"))  # add box around legend
+#          )
+
+(portfolio_stability_plot_quad = 
+  ggplot(data = diversity_stability_quadrat) +
+  geom_point(aes(x = Portfolio_effect, y = Cover_cv, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         ) # almost a perfect line, which makes sense given that CV is a component of portfolio effects
+```
+
+```{r diversity portfolio effects and synchrony quadrat level,  fig.height = 8, fig.width = 10}
+(loreau_diversity_plot_quad = 
+  ggplot(data = diversity_stability_quadrat) +
+  geom_point(aes(x = Shannon_mean, y = Loreau_synchrony, colour = Site, shape = Habitat), size = 4, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+(portfolio_diversity_plot_quad = 
+  ggplot(data = diversity_stability_quadrat) +
+  geom_point(aes(x = Shannon_mean, y = Portfolio_effect, colour = Site, shape = Habitat), size = 4, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Site) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
 ```

### 2024-07-19 noam_data_exploration_v3.Rmd
Halfway through figuring out how I might want to model diversity/stability but my boy just defended so I'm pivoting.

@@
   ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 4, stroke = 2) + 
+  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 4, stroke = 1) + 
    scale_colour_manual(values = site_colours) +
@@
   ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Loreau_synchrony, y = Cover_cv, colour = Site, shape = Habitat), size = 4, stroke = 2) + 
+  geom_point(aes(x = Loreau_synchrony, y = Cover_cv, colour = Site, shape = Habitat), size = 4, stroke = 1) + 
    scale_colour_manual(values = site_colours) +
@@
   ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Shannon_mean, y = Loreau_synchrony, colour = Site, shape = Habitat), size = 4, stroke = 2) + 
+  geom_point(aes(x = Shannon_mean, y = Loreau_synchrony, colour = Site, shape = Habitat), size = 4, stroke = 1) + 
    scale_colour_manual(values = site_colours) +
@@
          )
+```
+
+```{r diversity models quad level}
+diversity_mod_universal <- glmmTMB(Cover_cv ~Shannon_mean + Site*Habitat, family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
+diversity_dredge <- dredge(diversity_mod_universal)
+model.sel(diversity_dredge)
+
+synchrony_mod_universal <- glmmTMB(Cover_cv ~Loreau_synchrony*Site*Habitat, family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
+synchrony_dredge <- dredge(synchrony_mod_universal)
+model.sel(synchrony_dredge)
+
 ```

### 2024-07-22 noam_data_exploration_v3.Rmd
Continuing to update/play around with models and visualization.

--- 2024-07-19 noam_data_exploration_v3.Rmd
@@
   ggplot(data = alpha_diversity_quad_means) +
-  geom_point(aes(x = Year, y = Richness_mean, colour = Island_side, shape = Habitat, group = Island_side_habitat), size = 8) +
-  geom_line(aes(x = Year, y = Richness_mean, colour = Island_side, group = Island_side_habitat), size = 1.2) +
-  geom_errorbar(aes(x = Year, ymin = Richness_mean - Richness_se , ymax = Richness_mean + Richness_se, colour = Island_side,
-                    group = Island_side_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = island_side_colours) +
+  geom_point(aes(x = Year, y = Richness_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8) +
+  geom_line(aes(x = Year, y = Richness_mean, colour = Site, group = Site_habitat), size = 1.2) +
+  geom_errorbar(aes(x = Year, ymin = Richness_mean - Richness_se , ymax = Richness_mean + Richness_se, colour = Site,
+                    group = Site_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
@@
 
-slope_mod <- glmmTMB(Slope ~ Habitat*Island_side + (1|Site/Transect), data = gls_coefficients_merged)
+slope_mod <- glmmTMB(Slope ~ Habitat*Site + (1|Site/Transect), data = gls_coefficients_merged)
+# model convergence problem
 summary(slope_mod) # pretty low variance in random effects
@@
 Anova(slope_mod)
+# Response: Slope
 #                 Chisq Df Pr(>Chisq)    
-# Island_side           16.075  2  0.0003232 ***
-# Habitat             1301.260  3  < 2.2e-16 ***
-# Island_side:Habitat  799.995  6  < 2.2e-16 ***
-r.squaredGLMM(slope_mod) # 0.16, 0.18
+# Habitat      1397.999  3  < 2.2e-16 ***
+# Site           39.095  5  2.273e-07 ***
+# Habitat:Site 1738.155 15  < 2.2e-16 ***
+r.squaredGLMM(slope_mod) # 0.22, 0.24
 plot(allEffects(slope_mod))
@@
 
-slope_emm <- emmeans(slope_mod, ~ Habitat*Island_side) # create emmeans object
+slope_emm <- emmeans(slope_mod, ~ Habitat*Site) # create emmeans object
 emmeans::contrast(slope_emm, "consec", simple = "each", combine = TRUE, adjust = "bonferroni")
 # Note: not sure I want to include this
+emmeans(slope_mod, pairwise ~ Site | Habitat)
 ```
@@
   geom_vline(xintercept = 0, size = 2, linetype = 2) +
-  facet_grid(Habitat ~ Island_side) +
+  facet_grid(Habitat ~ Site) +
 #  theme_classic() +
@@
   data = alpha_diversity_quad,
-  control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000))
+  control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000)), na.action = na.pass
 ) # overspecified but going with it for now
@@
 # emmeans(richness_mod_2, pairwise ~ Habitat | Year_factor) # if we want to compare habitats year to year
-
+# model.sel(dredge(richness_mod_2))
+
+top_rich_mod <- glmmTMB(Richness ~ Year_factor*Habitat + Year_factor*Site + (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),   family = poisson, data = alpha_diversity_quad, control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000)))
+summary(top_rich_mod)
+Anova(top_rich_mod)
 ```
@@
 
+alpha_diversity_quad$Cover <- rowSums(data[ , -which(names(data) %in% meta_cols)])/100
 alpha_diversity_quad_sub <- alpha_diversity_quad[!alpha_diversity_quad$Location %in% problem_children, ]
 
-alpha_diversity_quad_sub$Cover <- rowSums(data_sub[ , -which(names(data_sub) %in% meta_cols)])/100
+# alpha_diversity_quad_sub$Cover <- rowSums(data_sub[ , -which(names(data_sub) %in% meta_cols)])/100
 alpha_diversity_quad_sub$Stratum <- 
@@
   ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 4, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site) +
+  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 4, stroke = 1) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Habitat) +
   theme_bw() + 
@@
   ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Loreau_synchrony, y = Cover_cv, colour = Site, shape = Habitat), size = 4, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site) +
+  geom_point(aes(x = Loreau_synchrony, y = Cover_cv, colour = Site, shape = Habitat), size = 4, stroke = 1) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Habitat) +
 
@@
   ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Portfolio_effect, y = Cover_cv, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+  geom_point(aes(x = Portfolio_effect, y = Cover_cv, colour = Site, shape = Habitat), size = 4, stroke = 2) + 
    scale_colour_manual(values = site_colours) +
@@
   ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Shannon_mean, y = Loreau_synchrony, colour = Site, shape = Habitat), size = 4, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site) +
+  geom_point(aes(x = Shannon_mean, y = Loreau_synchrony, colour = Site, shape = Habitat), size = 4, stroke = 1) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~Habitat) +
 
@@
          )
+```
+
+```{r diversity models quad level}
+diversity_mod_universal <- glmmTMB(Cover_cv ~Shannon_mean*Site + Shannon_mean*Habitat + Site*Habitat + (1|Transect/Location), family = Gamma(link = "log"), data = diversity_stability_quadrat, na.action = na.pass)
+diversity_dredge <- dredge(diversity_mod_universal)
+model.sel(diversity_dredge) # best model is as specified
+summary(diversity_mod_universal) # low variance in random effects
+Anova(diversity_mod_universal) # LR Tests are kinda useless
+r.squaredGLMM(diversity_mod_universal) # 0.653, 0.655
+
+synchrony_mod_universal <- glmmTMB(Cover_cv ~Loreau_synchrony*Site + Loreau_synchrony*Habitat + Site*Habitat + (1|Transect/Location), family = Gamma(link = "log"), data = diversity_stability_quadrat, na.action = na.pass)
+synchrony_dredge <- dredge(synchrony_mod_universal)
+model.sel(synchrony_dredge) # best model is as specified
+summary(synchrony_mod_universal)
+r.squaredGLMM(synchrony_mod_universal) # 0.800, 0.802
+
+diversity_stability_quadrat$Loreau_synchrony_trans <- sv_trans(diversity_stability_quadrat$Loreau_synchrony, diversity_stability_quadrat) # for modelling on beta
+synchrony_diversity_universal <- glmmTMB(Loreau_synchrony_trans ~Shannon_mean*Site + Shannon_mean*Habitat + Site*Habitat + (1|Transect/Location), family = beta_family(), data = diversity_stability_quadrat, na.action = na.pass)
+model.sel(dredge(synchrony_diversity_universal)) # best model as specified
+summary(synchrony_diversity_universal)
+Anova(synchrony_diversity_universal)
+r.squaredGLMM(synchrony_diversity_universal) # Model's distribution-specific variance is negative. Results are not reliable.
+
 ```

### 2024-07-22 noam_data_exploration_v3.Rmd
Continuing to add more vis options.

@@
 
+variance_ratio_quadrat <- variance_ratio(df = long_data_sub,
+                                   time.var = "Year",
+                                   species.var = "Taxa",
+                                   abundance.var = "Percent_cover",
+                                   average.replicates = F,
+                                   bootnumber = 1,
+                                   replicate.var = "Location")
+
 diversity_stability_quadrat <- alpha_diversity_quad_sub %>%
@@
 diversity_stability_quadrat <- merge(diversity_stability_quadrat, quadrat_level_portfolio, by = "Location") # add in portfolio effects
+diversity_stability_quadrat <- merge(diversity_stability_quadrat, variance_ratio_quadrat, by = "Location")
 pairs(~ Cover_mean + Cover_cv + Cover_stability + 
         Shannon_mean + Shannon_cv + Shannon_stability + 
-        Loreau_synchrony + Gross_synchrony + Portfolio_effect, data = diversity_stability)
+        Loreau_synchrony +  Portfolio_effect + VR, data = diversity_stability_quadrat)
 
@@
   ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 8) + 
-   scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-#    stat_smooth(aes(x = Shannon_mean, y = Shannon_cv),method = "lm", se = T, size = 2) + 
+  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 4) + 
+   scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+   facet_wrap(~Site) +
+
   theme_bw() + 
@@
    scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Habitat) +
+  facet_wrap(~Site) +
 
@@
    scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site) +
+ # facet_wrap(~Site) +
 
@@
          ) # almost a perfect line, which makes sense given that CV is a component of portfolio effects
+
 ```

### 2024-07-22 noam_data_exploration_v3.Rmd
Continuing to add more vis options.

@@
 diversity_stability_quadrat <- merge(diversity_stability_quadrat, quadrat_level_portfolio, by = "Location") # add in portfolio effects
-diversity_stability_quadrat <- merge(diversity_stability_quadrat, variance_ratio_quadrat, by = "Location")
 pairs(~ Cover_mean + Cover_cv + Cover_stability + 
         Shannon_mean + Shannon_cv + Shannon_stability + 
-        Loreau_synchrony +  Portfolio_effect + VR, data = diversity_stability_quadrat)
+        Loreau_synchrony + Gross_synchrony + Portfolio_effect, data = diversity_stability)
 

### 2024-07-23 noam_data_exploration_v3.Rmd
Added a cover vs cv figure, but I don't like it. Played around with how to calculate portfolio effects some more.

--- 2024-07-22 noam_data_exploration_v3.Rmd
@@
 
-    transect_level_portfolio[i, 2] <- pe_avg_cv(current_matrix) # calculate portfolio effects for current time series
+    transect_level_portfolio[i, 2] <- pe_mv(current_matrix, "loess_detrended") # calculate portfolio effects for current time series
   }
@@
 
-gross_synchrony_transect <- codyn::synchrony(df = transect_level,
-                                   time.var = "Year",
-                                   species.var = "Taxa",
-                                   abundance.var = "Mean_cover",
-                                   metric = "Gross",
-                                   replicate.var = "Stratum") # lots of warnings in this one - examine further later
-colnames(gross_synchrony_transect) = c("Stratum", "Gross_synchrony") # rename cols for merge later
+# gross_synchrony_transect <- codyn::synchrony(df = transect_level,
+#                                    time.var = "Year",
+#                                    species.var = "Taxa",
+#                                    abundance.var = "Mean_cover",
+#                                    metric = "Gross",
+#                                    replicate.var = "Stratum") # lots of warnings in this one - examine further later
+# colnames(gross_synchrony_transect) = c("Stratum", "Gross_synchrony") # rename cols for merge later
 
@@
 diversity_stability <- merge(diversity_stability, lm_synchrony_transect, by = "Stratum") # add in loreau and de Mazancourt synchrony
-diversity_stability <- merge(diversity_stability, gross_synchrony_transect, by = "Stratum")
+# diversity_stability <- merge(diversity_stability, gross_synchrony_transect, by = "Stratum")
 diversity_stability <- merge(diversity_stability, transect_level_portfolio, by = "Stratum") # add in portfolio effects
@@
         Shannon_mean + Shannon_cv + Shannon_stability + 
-        Loreau_synchrony + Gross_synchrony + Portfolio_effect, data = diversity_stability)
+        Loreau_synchrony + Portfolio_effect, data = diversity_stability)
 ```
@@
 
-(gross_stability_plot = 
+# (gross_stability_plot = 
+#   ggplot(data = diversity_stability) +
+#   geom_point(aes(x = Gross_synchrony, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+#    scale_colour_manual(values = site_colours) +
+#    scale_shape_manual(values = habitat_shapes) +
+#   facet_wrap(~Site) +
+#    
+#   theme_bw() + 
+#   theme(axis.title.x = element_text(size=18), 
+#         axis.title.y = element_text(size=18), 
+#         panel.background = element_blank(), 
+#         panel.grid.major = element_blank(),  #remove major-grid labels
+#         panel.grid.minor = element_blank(),  #remove minor-grid labels
+#         plot.background = element_blank(),
+#         legend.text = element_text(size = rel(1)), # increase legend text
+#         legend.title = element_text(size = rel(1)),
+#         legend.position = 'bottom', # move legend to bottom
+#         legend.box = 'horizontal', # make legend horizontal
+#         legend.box.background = element_rect(colour = "black"))  # add box around legend
+#          )
+# 
+(portfolio_stability_plot = 
   ggplot(data = diversity_stability) +
-  geom_point(aes(x = Gross_synchrony, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+  geom_point(aes(x = Portfolio_effect, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
    scale_colour_manual(values = site_colours) +
@@
         legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-(portfolio_stability_plot = 
+         ) # almost a perfect line, which makes sense given that CV is a component of portfolio effects
+```
+
+```{r diversity, portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
+(loreau_diversity_plot = 
   ggplot(data = diversity_stability) +
-  geom_point(aes(x = Portfolio_effect, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+  geom_point(aes(x = Shannon_mean, y = Loreau_synchrony, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
    scale_colour_manual(values = site_colours) +
@@
         legend.box.background = element_rect(colour = "black"))  # add box around legend
-         ) # almost a perfect line, which makes sense given that CV is a component of portfolio effects
-```
-
-```{r diversity, portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
-(loreau_diversity_plot = 
-  ggplot(data = diversity_stability) +
-  geom_point(aes(x = Shannon_mean, y = Loreau_synchrony, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site) +
-
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-(gross_diversity_plot = 
-  ggplot(data = diversity_stability) +
-  geom_point(aes(x = Shannon_mean, y = Gross_synchrony, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site) +
-
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
+         )
+
+# (gross_diversity_plot = 
+#   ggplot(data = diversity_stability) +
+#   geom_point(aes(x = Shannon_mean, y = Gross_synchrony, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+#    scale_colour_manual(values = site_colours) +
+#    scale_shape_manual(values = habitat_shapes) +
+#   facet_wrap(~Site) +
+#    
+#   theme_bw() + 
+#   theme(axis.title.x = element_text(size=18), 
+#         axis.title.y = element_text(size=18), 
+#         panel.background = element_blank(), 
+#         panel.grid.major = element_blank(),  #remove major-grid labels
+#         panel.grid.minor = element_blank(),  #remove minor-grid labels
+#         plot.background = element_blank(),
+#         legend.text = element_text(size = rel(1)), # increase legend text
+#         legend.title = element_text(size = rel(1)),
+#         legend.position = 'bottom', # move legend to bottom
+#         legend.box = 'horizontal', # make legend horizontal
+#         legend.box.background = element_rect(colour = "black"))  # add box around legend
+#          )
 
@@
 
-    quadrat_level_portfolio[i, 2] <- pe_avg_cv(current_matrix) # calculate portfolio effects for current time series
+    quadrat_level_portfolio[i, 2] <- pe_mv(current_matrix, "loess_detrended") # calculate portfolio effects for current time series
   }
@@
 
-gross_synchrony_quadrat <- codyn::synchrony(df = long_data_sub,
-                                   time.var = "Year",
-                                   species.var = "Taxa",
-                                   abundance.var = "Percent_cover",
-                                   metric = "Gross",
-                                   replicate.var = "Location") # lots of warnings in this one - examine further later
-colnames(gross_synchrony_transect) = c("Location", "Gross_synchrony") # rename cols for merge later
+## gross_synchrony_quadrat <- codyn::synchrony(df = long_data_sub,
+##                                    time.var = "Year",
+##                                    species.var = "Taxa",
+##                                    abundance.var = "Percent_cover",
+##                                    metric = "Gross",
+##                                    replicate.var = "Location") # lots of warnings in this one - examine further later
+## colnames(gross_synchrony_transect) = c("Location", "Gross_synchrony") # rename cols for merge later
 
@@
         Loreau_synchrony +  Portfolio_effect + VR, data = diversity_stability_quadrat)
+
+```
+
+```{r cover and cv, fig.height = 8, fig.width = 10}
+(cover_cv_plot = 
+  ggplot(data = diversity_stability_quadrat) +
+  geom_point(aes(x = Cover_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 4) + 
+   scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+   facet_wrap(~Site) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+(cover_cv_plot = 
+  ggplot(data = diversity_stability_quadrat) +
+  geom_point(aes(x = Cover_mean, y = Shannon_mean, colour = Site, shape = Habitat), size = 4) + 
+   scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+   facet_wrap(~Site) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
 

### 2024-07-23 noam_data_exploration_v3.Rmd
Found slight differences in how codyn calculates stability, so using that for the overall diversity stability dataframe. I still prefer CV.

--- 2024-07-23 noam_data_exploration_v3.Rmd
@@
 
+community_stability_quadrat <- community_stability(df = long_data_sub,
+                                   time.var = "Year",
+                                   abundance.var = "Percent_cover",
+                                   replicate.var = "Location")
+colnames(community_stability_quadrat) = c("Location", "Cover_stability")
+
 diversity_stability_quadrat <- alpha_diversity_quad_sub %>%
@@
             Cover_se = sd(Cover)/sqrt(n()),
-            Cover_cv = CV(Cover),
-            Cover_stability = 1/CV(Cover)) # as above
+            Cover_cv = CV(Cover))
 
@@
 diversity_stability_quadrat <- merge(diversity_stability_quadrat, variance_ratio_quadrat, by = "Location")
+diversity_stability_quadrat <- merge(diversity_stability_quadrat, community_stability_quadrat, by = "Location")
 pairs(~ Cover_mean + Cover_cv + Cover_stability + 

### 2024-07-29 julianna_data_exploration.Rmd
CV

@@
 ```
+
+# CV: Noam's way
+
+First try looking at CV Noam's way.
+I think this is more looking at if diverse areas (all time) are more stable, rather than if stability from one year to the next is dictated by current diversity (I think we can do this also since quadrats are fixed--will try that next)
+
+## Calculate
+
+Make a coefficient of variation function:
+
+```{r cv function}
+# coefficient of variation
+CV <- function(x){
+  return(sd(x,na.rm=T)/mean(x,na.rm=T))
+}
+```
+
+Apply it when we summarize
+
+```{r mega quadrat df}
+# get mega df with just about anything we need
+coralAlgDiv %>% 
+  full_join(algSum) -> megaQuadDF
+```
+
+```{r calculate CVs}
+diversity_stability <- megaQuadDF %>%
+  dplyr::group_by(site, habitat, site_habitat, island_side) %>%
+  dplyr::summarise(
+    Coral_mean = mean(coral),
+    Richness_mean = mean(Richness), 
+    Richness_se = sd(Richness)/sqrt(n()),
+    Richness_cv = CV(Richness),
+    Shannon_mean = mean(Shannon), 
+    Shannon_se = sd(Shannon)/sqrt(n()),
+    Shannon_cv = CV(Shannon),
+    Cover_mean = mean(AlgSum),
+    Cover_se = sd(AlgSum)/sqrt(n()),
+    Cover_cv = CV(AlgSum))
+```
+
+## Plot
+
+Algae diversity vs. cover CV:
+
+```{r Shannon}
+# divided by habitat
+diversity_stability %>% 
+  ggplot(aes(x = Shannon_mean, y = Cover_cv, color = site)) +
+  geom_point() +
+  facet_wrap(~habitat) +
+  theme_bw()
+
+# or all together
+diversity_stability %>% 
+  ggplot(aes(x = Shannon_mean, y = Cover_cv, color = site, shape = habitat)) +
+  geom_point(size = 2) +
+  theme_bw()
+```
+
+```{r richness}
+# divided by habitat
+diversity_stability %>% 
+  ggplot(aes(x = Richness_mean, y = Cover_cv, color = site)) +
+  geom_point() +
+  facet_wrap(~habitat) +
+  theme_bw()
+
+# or all together
+diversity_stability %>% 
+  ggplot(aes(x = Richness_mean, y = Cover_cv, color = site, shape = habitat)) +
+  geom_point(size = 2) +
+  theme_bw()
+```
+
+Versus coral cover
+
+```{r vs coral cover}
+diversity_stability %>% 
+  ggplot(aes(x = Coral_mean, y = Cover_cv, color = site, shape = habitat)) +
+  geom_point(size = 2) +
+  theme_bw()
+```
+
+# Stability
+
+Is change in algal cover in a given quadrat from one year to the next influenced by diversity? Need to set up a lag/change to calculate this
+
+# Synchrony
+
+Maybe try this way: https://cran.r-project.org/web/packages/codyn/vignettes/Community_Stability_Metrics.html using the `codyn` package?
+
+```{r}
+
+```

### 2024-07-29 julianna_data_exploration.Rmd
Fixed potential "fig 1" to be by site instead of island side and include lines for each of the disturbances in Lauren's Excel

@@
   select(-year, -location, -site, -island_side, -habitat, -site_habitat, -transect, -quadrat) -> sppMat
+
+# get ranges to calculate over
+startAlg <- grep("acanthophora_spicifera", colnames(algae))
+endAlg <- grep("valonia_ventricosa", colnames(algae))
 
@@
   # relevel for graphing
-  mutate(island_side = factor(island_side, levels = c("north", "east", "west")),
-         habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) -> quadHabIS
+  mutate(island_side = str_to_sentence(island_side),
+         island_side = factor(island_side, levels = c("North", "East", "West")),
+         habitat = case_when(
+           str_detect(habitat, pattern = "_") ~ str_to_sentence(paste0(str_split(habitat, "_")[[1]][1], " ", str_split(habitat, "_")[[1]][2])),
+           TRUE ~ str_to_sentence(habitat)),
+         habitat = factor(habitat, levels = c("Fringing", "Backreef", "Outer 10", "Outer 17"))) -> quadHabIS
 
@@
     moma.colors("Klein", type = "discrete")[4] # turf
-  )) +
+  ),
+  labels = c("CCA", "Coral", "Macroalgae", "Turf algae")) +
   facet_grid(island_side ~ habitat) +
@@
   xlab("Year")
+```
+
+## Quarat-level habitat-site
+
+```{r get summary}
+
+quad4sum %>% 
+  # summarize and group quadrats by island/habitat
+  group_by(year, habitat, site, island_side, substrate) %>% 
+  summarise(mean_cov = mean(sum_cover),
+            sd_cov = sd(sum_cover, na.rm = TRUE),
+            n = n(),
+            CI95 = 1.96*(sd_cov)/sqrt(n)) %>% 
+  # relevel for graphing
+  mutate(island_side = str_to_sentence(island_side),
+         island_side = factor(island_side, levels = c("North", "East", "West")),
+         habitat = case_when(
+           str_detect(habitat, pattern = "_") ~ str_to_sentence(paste0(str_split(habitat, "_")[[1]][1], " ", str_split(habitat, "_")[[1]][2])),
+           TRUE ~ str_to_sentence(habitat)),
+         habitat = factor(habitat, levels = c("Fringing", "Backreef", "Outer 10", "Outer 17"))) %>% 
+  mutate(site = case_when(
+    site == "lter_1" ~ "LTER 1 (N)",
+    site == "lter_2" ~ "LTER 2 (N)",
+    site == "lter_3" ~ "LTER 3 (E)",
+    site == "lter_4" ~ "LTER 4 (E)",
+    site == "lter_5" ~ "LTER 5 (W)",
+    site == "lter_6" ~ "LTER 6 (W)"
+  )) -> quadHabSite
+
+```
+
+```{r plot cover through time}
+# pdf("Fig1_quadHabSiteNo20.pdf", height = 6.5, width = 12)
+
+quadHabSite %>% 
+  filter(year != 2020) %>% 
+  ggplot(aes(x = year, y = mean_cov, group = substrate, fill = substrate)) +
+  geom_ribbon(aes(ymin = mean_cov -CI95, 
+                    ymax = mean_cov + CI95),
+              alpha  = 0.7) +
+  geom_line() +
+  theme_bw() +
+  scale_fill_manual(values = c(
+    moma.colors("Klein", type = "discrete")[2], # CCA
+    moma.colors("Klein", type = "discrete")[1], # coral
+    moma.colors("Klein", type = "discrete")[5], # macro
+    moma.colors("Klein", type = "discrete")[4] # turf
+  ),
+  labels = c("CCA", "Coral", "Macroalgae", "Turf algae")) +
+  facet_grid(habitat ~ site) + # not sure which direction is best here
+  ylab("Mean cover (%)") +
+  xlab("Year") +
+  # add lines for disturbances
+  geom_vline(xintercept = 2006, linetype = "dotdash") + # start of COTS
+  geom_vline(xintercept = 2010,  color = "darkgray") + # cyclone Oli
+  geom_vline(xintercept = 2019, linetype = "dashed", color = "red")  # bleaching
+
+# dev.off()
 ```

### 2024-07-29 julianna_data_exploration.Rmd
Calculated the change in algal cover over the next year based on the richness/diversity in that year

@@
 
+- Would need to start in 2008 because no change to 2007
+
+## Wrangle
+
+Might just brute force this...
+
+```{r get the change in algal cover each year}
+megaQuadDF %>% 
+  select(year, site, habitat, site_habitat, transect, quadrat, island_side, AlgSum) %>% 
+  pivot_wider(names_from = year,
+              values_from = AlgSum) %>% 
+  # calculate changes FROM the previous year 
+  # this means that a change FROM 2007 to 2008 will be joined with diversity in 2007 
+  # that means that a positive correlation between richness in 2007 and "ChFrom2007" (etc) == more diverse communities are more likely to INCREASE in cover/less likely to decrease in cover (not necessarily stability)
+  mutate(AlgChFrom2007 = `2008` - `2007`,
+         AlgChFrom2008 = `2009` - `2008`,
+         AlgChFrom2009 = `2010` - `2009`,
+         AlgChFrom2010 = `2011` - `2010`,
+         AlgChFrom2011 = `2012` - `2011`,
+         AlgChFrom2012 = `2013` - `2012`,
+         AlgChFrom2013 = `2014` - `2013`,
+         AlgChFrom2014 = `2015` - `2014`,
+         AlgChFrom2015 = `2016` - `2015`,
+         AlgChFrom2016 = `2017` - `2016`,
+         AlgChFrom2017 = `2018` - `2017`,
+         AlgChFrom2018 = `2019` - `2018`,
+         #AlgChFrom2007 = `2021` - `2019`, # maybe skip because this isn't comparable?
+         AlgChFrom2021 = `2022` - `2021`, 
+         AlgChFrom2022 = `2023` - `2022`
+         ) -> algalCh
+```
+
+There are **NA's in 2021**--confirm those are just missing quadrats.
+
+Now get it back into long/joinable format (probably a more elegant way to do this):
+
+```{r get into long format and join}
+startChA <- grep("AlgChFrom2007", colnames(algalCh))
+endtChA <- grep("AlgChFrom2022", colnames(algalCh))
+
+algalCh %>% 
+  # get the columns we need
+  select(c(site, habitat, site_habitat, transect, quadrat, island_side, colnames(algalCh)[startChA:endtChA])) %>% 
+  pivot_longer(cols = colnames(algalCh)[startChA:endtChA],
+               names_to = "Change_year",
+               values_to = "Change_in_percent"
+               ) %>% 
+  # match with correct year
+  mutate(year = case_when(
+    Change_year == "AlgChFrom2007" ~ 2007,
+    Change_year == "AlgChFrom2008" ~ 2008,
+    Change_year == "AlgChFrom2009" ~ 2009,
+    Change_year == "AlgChFrom2010" ~ 2010,
+    Change_year == "AlgChFrom2011" ~ 2011,
+    Change_year == "AlgChFrom2012" ~ 2012,
+    Change_year == "AlgChFrom2013" ~ 2013,
+    Change_year == "AlgChFrom2014" ~ 2014,
+    Change_year == "AlgChFrom2015" ~ 2015,
+    Change_year == "AlgChFrom2016" ~ 2016,
+    Change_year == "AlgChFrom2017" ~ 2017,
+    Change_year == "AlgChFrom2018" ~ 2018,
+    Change_year == "AlgChFrom2019" ~ 2019,
+    Change_year == "AlgChFrom2021" ~ 2021,
+    Change_year == "AlgChFrom2022" ~ 2022,
+    TRUE ~ -9999
+  )) %>%
+  full_join(megaQuadDF) %>% 
+  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17")),
+         island_side = factor(island_side, levels = c("north", "east", "west"))) -> algChYr
+
+```
+
+## Plot
+
+```{r plot richness}
+algChYr %>% 
+  ggplot(aes(x = Richness, y = Change_in_percent, group = Richness)) +
+  geom_boxplot() +
+  theme_classic() +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
+  facet_grid(island_side~habitat)
+```
+
+```{r plot abs(change) ~ richness}
+algChYr %>% 
+  ggplot(aes(x = Richness, y = abs(Change_in_percent), group = Richness)) +
+  geom_boxplot() +
+  theme_classic() +
+  facet_grid(island_side~habitat)
+```
+
+```{r plot Shannon}
+algChYr %>% 
+  ggplot(aes(x = Shannon, y = Change_in_percent)) +
+  geom_point() +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
+  geom_smooth() +
+  theme_classic() +
+  facet_grid(island_side~habitat)
+```
+
+```{r plot Shannon ~ abs(change)}
+# pdf("change_in_cov_Shann.pdf", height = 7, width = 10)
+
+algChYr %>% 
+  ggplot(aes(x = Shannon, y = abs(Change_in_percent), color = year)) +
+  geom_point() +
+  geom_smooth(color = "coral") +
+  theme_classic() + # kind of negative
+  facet_grid(island_side~habitat)
+
+# dev.off()
+```
+
 # Synchrony
@@
 ```
+
+# Nic's plot of change through time
+
+Want to plot how each spp changes in abundance through time just to get an idea
+
+## Wrangling
+
+```{r}
+megaQuadDF
+```

### 2024-07-30 julianna_data_exploration.Rmd
Made graph Nic suggested that is kind of synchrony vibes

@@
 
+## Total cover (not change)
+
 Want to plot how each spp changes in abundance through time just to get an idea
 
-## Wrangling
-
-```{r}
-megaQuadDF
-```
+### Wrangling
+
+Need to pivot so species are a column
+
+```{r summarize and pivot}
+startAlg <- grep("acanthophora_spicifera", colnames(algae))
+endAlg <- grep("valonia_ventricosa", colnames(algae))
+
+algae %>% 
+  pivot_longer(cols = colnames(algae)[startAlg:endAlg],
+               names_to = "Taxa",
+               values_to = "Perc_cov") %>% 
+  group_by(year, site, island_side, habitat, Taxa) %>% 
+  summarize(Mean_cov = mean(Perc_cov),
+            Sd_cov = sd(Perc_cov)) -> taxa_long
+
+```
+
+### Plot
+
+Check for most abundant things
+
+```{r calculate most abundant throughout time}
+taxa_long %>% 
+  group_by(habitat, Taxa) %>% 
+  summarize(Mean_cov = mean(Mean_cov)) %>% view()
+```
+
+Plot all
+
+```{r plot}
+# pdf("abundance_time_noTurf.pdf", height = 10, width = 15)
+
+taxa_long %>% 
+  filter(Taxa != "algal_turf") %>% 
+  ggplot(aes(x = year, y = Mean_cov, group = Taxa, color = Taxa)) +
+  geom_line() +
+  facet_grid(habitat ~ site) +
+  theme_bw() +
+  theme(legend.position = "top")
+
+# dev.off()
+```
+
+Plot some interesting ones (skipping CCA/peysonnellia/turf; focusing on most abundant things)
+
+```{r plot most abundant}
+taxa_long %>% 
+  filter(Taxa == "damselfish_turf" |
+           Taxa == "lobophora_variegata" |
+           Taxa == "turbinaria_ornata" |
+           Taxa == "dictyota_bartayresiana" |
+           Taxa == "halimeda_minima" |
+           Taxa == "asparagopsis_taxiformis" |
+           Taxa == "amansia_rhodantha" |
+           Taxa == "cyanophyta" |
+           Taxa == "halimeda_discoidea" |
+           Taxa == "halimeda_opuntia" |
+           Taxa == "sargassum_pacificum" |
+           Taxa == "halimeda_taenicola" |
+           Taxa == "halimeda_sp" |
+           Taxa == "halimeda_distorta" 
+           ) %>% 
+  ggplot(aes(x = year, y = Mean_cov, group = Taxa, color = Taxa)) +
+  geom_line() +
+  facet_grid(site ~ habitat)
+
+```
+
+## Change each time (synchrony vibes)
+
+### Wrangle
+
+```{r calculate change per taxa per year}
+algae %>% 
+  pivot_longer(cols = colnames(algae)[startAlg:endAlg],
+               names_to = "Taxa",
+               values_to = "Perc_cov") %>% 
+  pivot_wider(names_from = year,
+              values_from = Perc_cov) %>% 
+  mutate(AlgChFrom2007 = `2008` - `2007`,
+       AlgChFrom2008 = `2009` - `2008`,
+       AlgChFrom2009 = `2010` - `2009`,
+       AlgChFrom2010 = `2011` - `2010`,
+       AlgChFrom2011 = `2012` - `2011`,
+       AlgChFrom2012 = `2013` - `2012`,
+       AlgChFrom2013 = `2014` - `2013`,
+       AlgChFrom2014 = `2015` - `2014`,
+       AlgChFrom2015 = `2016` - `2015`,
+       AlgChFrom2016 = `2017` - `2016`,
+       AlgChFrom2017 = `2018` - `2017`,
+       AlgChFrom2018 = `2019` - `2018`, 
+       # skip 2020
+       AlgChFrom2021 = `2022` - `2021`, 
+       AlgChFrom2022 = `2023` - `2022`) -> algChTaxa
+```
+
+```{r get longer with values per taxa}
+startChAT <- grep("AlgChFrom2007", colnames(algChTaxa))
+endtChAT <- grep("AlgChFrom2022", colnames(algChTaxa))
+
+algChTaxa %>% 
+  select(c(site, habitat, site_habitat, transect, quadrat, island_side, Taxa, colnames(algChTaxa)[startChAT:endtChAT])) %>% 
+  pivot_longer(cols = colnames(algChTaxa)[startChAT:endtChAT],
+               names_to = "Change_year",
+               values_to = "Change_in_percent"
+               ) %>% 
+  # join with the correct year
+  mutate(year = case_when(
+    Change_year == "AlgChFrom2007" ~ 2007,
+    Change_year == "AlgChFrom2008" ~ 2008,
+    Change_year == "AlgChFrom2009" ~ 2009,
+    Change_year == "AlgChFrom2010" ~ 2010,
+    Change_year == "AlgChFrom2011" ~ 2011,
+    Change_year == "AlgChFrom2012" ~ 2012,
+    Change_year == "AlgChFrom2013" ~ 2013,
+    Change_year == "AlgChFrom2014" ~ 2014,
+    Change_year == "AlgChFrom2015" ~ 2015,
+    Change_year == "AlgChFrom2016" ~ 2016,
+    Change_year == "AlgChFrom2017" ~ 2017,
+    Change_year == "AlgChFrom2018" ~ 2018,
+    Change_year == "AlgChFrom2019" ~ 2019,
+    Change_year == "AlgChFrom2021" ~ 2021,
+    Change_year == "AlgChFrom2022" ~ 2022,
+    TRUE ~ -9999)) -> longChTaxa
+```
+
+```{r get averages per site/hab}
+longChTaxa %>% 
+  group_by(site, habitat, year, island_side, Taxa) %>% 
+  summarize(Mean_ch_cov = mean(Change_in_percent, na.rm=TRUE),
+            Sd_ch_cov = sd(Change_in_percent, na.rm=TRUE)) %>% 
+  # if you take the mean of NAs ONLY, get NaNs--change those to NA's:
+  mutate(Mean_ch_cov = case_when(is.nan(Mean_ch_cov) ~ NA,
+                                 TRUE ~ Mean_ch_cov)) %>% 
+  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) -> meanChTaxSiteHab
+```
+
+Not sure how to calculate N because **THERE ARE NA'S**!!! I think these are quadrats that are missed
+
+### Plot
+
+```{r}
+# pdf("NicsPlot.pdf", height = 13, width = 10)
+
+meanChTaxSiteHab %>%
+  ggplot(aes(x = year, y = Mean_ch_cov, group = Taxa, color = Taxa)) +
+  geom_line() +
+  theme_bw() +
+  facet_grid(site~habitat) +
+  theme(legend.position = "top") +
+  ylab("Mean change in cover (%)")
+
+# dev.off()
+```
+
+Can we super-impose total change in cover on that?
+
+```{r}
+algChYr %>% 
+  group_by(site, habitat, year, island_side) %>% 
+  summarize(Mean_alg_ChAll = mean(Change_in_percent, na.rm = TRUE)) %>% 
+  mutate(Mean_alg_ChAll = case_when(is.nan(Mean_alg_ChAll) ~ NA,
+                                    TRUE ~ Mean_alg_ChAll)) %>%
+  full_join(meanChTaxSiteHab) -> synchDat
+
+synchDat %>% 
+  ggplot(aes(x = year)) +
+  geom_line(aes(x = year, y = Mean_ch_cov, group = Taxa, color = Taxa)) +
+  theme_bw() +
+  geom_line(aes(x = year, y = Mean_alg_ChAll+50), color = "black") +
+  facet_grid(site~habitat) +
+  theme(legend.position = "top") +
+  ylab("Mean change in cover (%)") +
+  ylim(-50, 150) # maybe this really needs to be done wrt change in stuff without turf?
+```

### 2024-07-30 julianna_algalMetric_wrangling.Rmd
Created a fresh Rmd for just the algal metric creation code

---
title: "Julianna's algal metrics wrangling code"
output: html_document
date: "2024-07-30"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clean up
rm(list=ls())

# load packages
require(here) # relative file paths
require(tidyverse) # data wrangling
require(vegan) # community/diversity analyses
```

This script generates the CSV of algal metrics to regress other variables against

# Bring in the data

```{r}
# algal data
algae <- read_csv(here("data", "algaeonly_cleaned_colnames.csv")) %>% 
  mutate(island_side = case_when(site == "lter_1" ~ "north",
                                 site == "lter_2" ~ "north",
                                 site == "lter_3" ~ "east",
                                 site == "lter_4" ~ "east",
                                 site == "lter_5" ~ "west",
                                 site == "lter_6" ~ "west",
                                 TRUE ~ "UdINGLEbUTT"))
```

### 2024-07-30 julianna_algalMetric_wrangling.Rmd
Updated wrangling/metric calculating code

@@
 ```{r}
-# algal data
+# algal cover 
 algae <- read_csv(here("data", "algaeonly_cleaned_colnames.csv")) %>% 
@@
                                  TRUE ~ "UdINGLEbUTT"))
+
+# sand cover
+sand <- read_csv(here("data", "allbenthicdata_cleaned_colnames.csv")) %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, sand, coral)
 ```
+
+# Calculate algal alpha diversity
+
+## Make matrices
+
+Make a spp matrix to use with `vegan`
+
+```{r make species matrix}
+# matrix of just species and their relative percent covers in the same order as the algae tibble
+algae %>% 
+  select(-year, -location, -site, -island_side, -habitat, -site_habitat, -transect, -quadrat) -> sppMat
+```
+
+```{r make a metadata matrix in the same order}
+# get ranges to calculate over (this is just for indexing purposes in the next bit of code)
+startAlg <- grep("acanthophora_spicifera", colnames(algae))
+endAlg <- grep("valonia_ventricosa", colnames(algae))
+
+# metadata matrix
+algae %>% 
+  # bring in sand data
+  full_join(sand) %>% 
+  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg]), # percent cover of algae
+         AlgCov_percHard = case_when(
+           sand == 0 ~ AlgCov_perc, # since we can't divide by a zero
+           TRUE ~ AlgCov_perc/sand 
+           )) %>% # percent cover of hard substrate
+   # make a year-habitat column for betadisper
+  mutate(hab_year = paste0(habitat, "_", year),
+         island_side_hab_year = paste0(island_side, "_", hab_year)) %>% 
+  select(year, location, site, habitat, island_side, site_habitat, 
+         hab_year, island_side_hab_year, transect, quadrat, AlgCov_perc, 
+         AlgCov_percHard, sand, coral)  -> sppMeta
+```
+
+## Calculate diversity metrics
+
+```{r get richness and Shannon diversity}
+sppMeta %>% 
+  # these functions use the vegan package
+  mutate(Richness = specnumber(sppMat),
+         Shannon = diversity(sppMat, "shannon")) -> algDiv
+```
+
+# Calculate algal beta diversity
+
+## Distance matrix: habitat-year
+
+We get an error if we run `vegdist()` with quadrats that have no algae. We can see there are 672 of these that we need to filter out:
+
+```{r look for zeros}
+sppMat %>% 
+  mutate(AlgCov_perc = rowSums(across(where(is.numeric)))) %>% 
+  filter(AlgCov_perc == 0) %>% dim() # 672 with no algae at all
+```
+
+Now we can exclude quadrats with no algae and make a species matrix again
+
+```{r get spp mat df}
+startAlg <- grep("acanthophora_spicifera", colnames(algae))
+endAlg <- grep("valonia_ventricosa", colnames(algae))
+
+# spp matrix where 0 observations are filtered out
+algae %>% 
+  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(AlgCov_perc != 0) %>% 
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgCov_perc) -> sppMat0
+```
+
+```{r metadata matix}
+# matching metadata matrix
+algae %>% 
+  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(AlgCov_perc != 0) %>%
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
+  # make a year-habitat column for betadisper
+  mutate(hab_year = paste0(habitat, "_", year),
+         island_side_hab_year = paste0(island_side, hab_year)) -> sppMeta0
+```
+
+Now we'll calculate this by habitat because we're compute limited
+
+### Calculate fringing
+
+```{r get fringe mat}
+startAlg <- grep("acanthophora_spicifera", colnames(algae))
+endAlg <- grep("valonia_ventricosa", colnames(algae))
+
+# spp matrix
+algae %>% 
+  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(AlgCov_perc != 0) %>% 
+  filter(habitat == "fringing") %>% 
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgCov_perc) -> sppMat0fring
+
+# metadata matrix
+algae %>% 
+  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(AlgCov_perc != 0) %>%
+   filter(habitat == "fringing") %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
+  # make a year-habitat column for betadisper
+  mutate(hab_year = paste0(habitat, "_", year),
+         island_side_hab_year = paste0(island_side, hab_year),
+         island_side_year = paste0(island_side, year)) -> sppMeta0fring
+```
+
+Calculate Bray Curtis distances
+
+```{r calculate distances}
+brayMatFring <- vegdist(sppMat0fring, "bray") # 1 is maximum dissimilarity
+```
+
+Calculate dispersion (this is using Noam's code)
+
+```{r calculate beta dispersion}
+yrBetaDisperFring <- vegan::betadisper(brayMatFring, sppMeta0fring$year) # warning: some squared distances are negative and changed to zero (ask Jamie about this??)
+```
+
+### Backreef
+
+```{r get backreef df}
+startAlg <- grep("acanthophora_spicifera", colnames(algae))
+endAlg <- grep("valonia_ventricosa", colnames(algae))
+
+# spp matrix
+algae %>% 
+  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(AlgCov_perc != 0) %>% 
+  filter(habitat == "backreef") %>% 
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgCov_perc) -> sppMat0back
+
+# metadata matrix
+algae %>% 
+  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(AlgCov_perc != 0) %>%
+   filter(habitat == "backreef") %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
+  # make a year-habitat column for betadisper
+  mutate(hab_year = paste0(habitat, "_", year),
+         island_side_hab_year = paste0(island_side, hab_year),
+         island_side_year = paste0(island_side, year)) -> sppMeta0back
+```
+
+Calculate Bray Curtis distances
+
+```{r calculate distance}
+brayMatBack <- vegdist(sppMat0back, "bray") # 1 is maximum dissimilarity
+```
+
+Calculate beta disperion
+
+```{r calculate beta dispersion}
+yrBetaDisperBack <- vegan::betadisper(brayMatBack, sppMeta0back$year) 
+```
+
+### Outer 10
+
+```{r get outer 10 df}
+startAlg <- grep("acanthophora_spicifera", colnames(algae))
+endAlg <- grep("valonia_ventricosa", colnames(algae))
+
+# spp matrix
+algae %>% 
+  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(AlgCov_perc != 0) %>% 
+  filter(habitat == "outer_10") %>% 
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgCov_perc) -> sppMat0outer10
+
+# metadata matrix
+algae %>% 
+  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(AlgCov_perc != 0) %>%
+   filter(habitat == "outer_10") %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
+  # make a year-habitat column for betadisper
+  mutate(hab_year = paste0(habitat, "_", year),
+         island_side_hab_year = paste0(island_side, hab_year),
+         island_side_year = paste0(island_side, year)) -> sppMeta0outer10
+```
+
+Calculate Bray Curtis distances
+
+```{r calculate distance}
+brayMatOuter10 <- vegdist(sppMat0outer10, "bray") # 1 is maximum dissimilarity
+```
+
+Calculate beta disperion
+
+```{r calculate beta dispersion}
+yrBetaDisperOuter10 <- vegan::betadisper(brayMatOuter10, sppMeta0outer10$year) # error: some squared distances are negative and changed to zero
+```
+
+### Outer 17
+
+```{r get outer 17 df}
+startAlg <- grep("acanthophora_spicifera", colnames(algae))
+endAlg <- grep("valonia_ventricosa", colnames(algae))
+
+# spp matrix
+algae %>% 
+  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(AlgCov_perc != 0) %>% 
+  filter(habitat == "outer_17") %>% 
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgCov_perc) -> sppMat0outer17
+
+# metadata matrix
+algae %>% 
+  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(AlgCov_perc != 0) %>%
+   filter(habitat == "outer_17") %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
+  # make a year-habitat column for betadisper
+  mutate(hab_year = paste0(habitat, "_", year),
+         island_side_hab_year = paste0(island_side, hab_year),
+         island_side_year = paste0(island_side, year)) -> sppMeta0outer17
+```
+
+Calculate Bray Curtis distances
+
+```{r calculate distance}
+brayMatOuter17 <- vegdist(sppMat0outer17, "bray") # 1 is maximum dissimilarity
+```
+
+Calculate beta disperion
+
+```{r calculate beta dispersion}
+yrBetaDisperOuter17 <- vegan::betadisper(brayMatOuter17, sppMeta0outer17$year) # error: some squared distances are negative and changed to zero
+```
+
+## Join together
+
+Attach back to metadata
+
+```{r join distances with metadata}
+sppMeta0fring %>% 
+  mutate(distance_centriod = yrBetaDisperFring$distances) -> fringeDist
+
+sppMeta0back %>% 
+  mutate(distance_centriod = yrBetaDisperBack$distances) -> backDist
+
+sppMeta0outer10 %>% 
+  mutate(distance_centriod = yrBetaDisperOuter10$distances) -> outer10Dist
+
+sppMeta0outer17 %>% 
+  mutate(distance_centriod = yrBetaDisperOuter17$distances) -> outer17Dist
+```
+
+```{r join together}
+# make sure this looks good
+colnames(fringeDist) == colnames(backDist) 
+colnames(backDist) == colnames(outer10Dist) 
+colnames(outer10Dist) == colnames(outer17Dist)
+
+# join together and tidy up
+betaDHabYr <- rbind(fringeDist, backDist, outer10Dist, outer17Dist) %>% 
+  mutate(hab_yr_distcentroid = distance_centriod) %>% 
+  select(year, location, site, habitat, transect, quadrat, hab_yr_distcentroid)
+```
+
+### Save dispersions
+
+```{r}
+# save(yrBetaDisperFring, file = here("data/fringingyrdispersion.Rdata"))
+# save(yrBetaDisperBack, file = here("data/backyrdispersion.Rdata"))
+# save(yrBetaDisperOuter10, file = here("data/outer10yrdispersion.Rdata"))
+# save(yrBetaDisperOuter17, file = here("data/outer17yrdispersion.Rdata"))
+```
+
+### Load dispersions
+
+```{r}
+load(here("data/fringingyrdispersion.Rdata"))
+load(here("data/backyrdispersion.Rdata"))
+load(here("data/outer10yrdispersion.Rdata"))
+load(here("data/outer17yrdispersion.Rdata"))
+```
+
+# Join all quadrat-level datasets together
+
+```{r join}
+algDiv %>% 
+  full_join(betaDHabYr) %>% 
+  # follow a cleaner naming convention
+  rename(algal_perc_cov = AlgCov_perc,
+         algal_perc_cov_hard = AlgCov_percHard,
+         algal_richness = Richness,
+         algal_Shann = Shannon) %>% 
+  # get just the columns we need
+  select(-site_habitat, -hab_year, -island_side_hab_year) -> quadData
+```
+
+```{r save}
+# write_csv(quadData, here("algal_metrics_quadrat.csv"))
+```
+
+# Site/habitat/year-level mean
+
+We'll also average these data at the site/habitat/year-level to be more compatible with other data types.
+This will be done at the *quadrat-level*, meaning we won't take the transect-level summary first, since they aren't meant to be factors in a statistical design
+
+```{r get averages}
+quadData %>% 
+  group_by(year, site, habitat, island_side) %>% 
+  summarize(mean_coral_perc_cov = mean(coral),
+            mean_alg_perc_cov = mean(algal_perc_cov),
+            mean_alg_perc_cov_hard = mean(algal_perc_cov_hard),
+            mean_alg_richness = mean(algal_richness),
+            mean_alg_Shann = mean(algal_Shann),
+            mean_dist_centroid_habyear = mean(hab_yr_distcentroid, na.rm = TRUE)
+            ) -> siteData
+```
+
+```{r save dataframe}
+# write_csv(siteData, here("algal_metrics_site.csv"))
+```
+
+# All time metrics
+
+## Calculate the coefficient of variation
+
+This is basically just copying/pasting Noam's code
+
+```{r cv function}
+# coefficient of variation
+CV <- function(x){
+  return(sd(x,na.rm=T)/mean(x,na.rm=T))
+}
+```
+
+Apply it when we summarize
+
+```{r calculate CV and summarize}
+quadData %>% 
+  # don't include year this time
+  group_by(site, habitat, island_side) %>%
+  summarize(
+    mean_coral_perc = mean(coral),
+    mean_alg_richness = mean(algal_richness), 
+    richness_se = sd(algal_richness)/sqrt(n()),
+    richness_cv = CV(algal_richness),
+    mean_Shann = mean(algal_Shann), 
+    Shann_se = sd(algal_Shann)/sqrt(n()),
+    Shann_cv = CV(algal_Shann),
+    # for total algal cover
+    mean_alg_cov = mean(algal_perc_cov),
+    alg_cov_se = sd(algal_perc_cov)/sqrt(n()),
+    cv_alg_cov = CV(algal_perc_cov),
+    # and for % of hard substrate
+    mean_alg_cov_hard = mean(algal_perc_cov_hard),
+    alg_cov_hard_se = sd(algal_perc_cov_hard)/sqrt(n()),
+    cv_alg_cov_hard = CV(algal_perc_cov_hard),
+    ) -> alltimeCV
+```
+
+```{r save}
+# write_csv(alltimeCV, here(here("algal_metrics_all_time.csv")))
+```
+
+- Noam: Add **synchrony** to this dataset + any other relevant metrics
+
+s

### 2024-07-30 julianna_algalMetric_wrangling.Rmd
Created the model markdown

@@
 ```{r save}
-# write_csv(quadData, here("algal_metrics_quadrat.csv"))
+# write_csv(quadData, here("data/algal_metrics_quadrat.csv"))
 ```
@@
 ```{r save dataframe}
-# write_csv(siteData, here("algal_metrics_site.csv"))
+# write_csv(siteData, here("data/algal_metrics_site.csv"))
 ```
@@
 ```{r save}
-# write_csv(alltimeCV, here(here("algal_metrics_all_time.csv")))
+# write_csv(alltimeCV, here(here("data/algal_metrics_all_time.csv")))
 ```

### 2024-07-30 julianna_richness_model.Rmd
Created the model markdown

---
title: "Richness model"
output: html_document
date: "2024-07-30"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clean up
rm(list=ls())

# load packages
require(here) # relative file paths
require(tidyverse) # data wrangling
require(lme4) # mixed models
```

Make a model for algal richness through time

# Bring in the data

```{r load data}
cleaned_rich <- read_csv(here("data/algal_metrics_quadrat.csv")) # generated using my (Julianna's) algal metric code
```

# Make a model

```{r}
mRich <- glmer(,
               family = Poisson,
               data = cleaned_rich)
```

### 2024-07-30 julianna_richness_model.Rmd
Worked on richness model--made a first pass--maybe actually want to add an interaction?

@@
 ```{r load data}
-cleaned_rich <- read_csv(here("data/algal_metrics_quadrat.csv")) # generated using my (Julianna's) algal metric code
+cleaned_rich <- read_csv(here("data/algal_metrics_quadrat.csv")) %>% 
+  filter(year != 2020) # fake data
+  # generated using my (Julianna's) algal metric code
 ```
@@
 
-```{r}
-mRich <- glmer(,
-               family = Poisson,
+```{r make Poisson model}
+mRich <- glmer(algal_richness ~ as.factor(year) + habitat + (1|site/transect),
+               family = poisson,
                data = cleaned_rich)
 ```
+
+# Examine output
+
+```{r make prediction grid}
+# create a grid to predict over
+newdata <- expand_grid(year = unique(cleaned_rich$year),
+                       habitat = unique(cleaned_rich$habitat),
+                       site = unique(cleaned_rich$site),
+                       transect = unique(cleaned_rich$transect))
+```
+
+```{r generate predictions}
+newdata$Preds <- predict(mRich, newdata = newdata)
+```
+
+```{r plot predictions}
+newdata %>% 
+  # get transect means
+  group_by(year, habitat, site) %>% 
+  summarize(Pred_mean_richness = mean(Preds),
+            SD_richness = sd(Preds)) %>% 
+  ggplot(aes(x = year, y = Pred_mean_richness, group = habitat, color = habitat)) +
+  geom_line() +
+  theme_bw() +
+  facet_wrap(~site)
+```
+
+- Should year be a factor?

### 2024-08-04 noam_data_exploration_v3.Rmd
Updated for use with the corrected colnames file

--- 2024-07-23 noam_data_exploration_v3.Rmd
@@
 # starting with all data now because of 0 algae transects/quadrats
-data <- read.csv(here::here("data", "allbenthicdata_cleaned.csv"), stringsAsFactors = F)
-
-not_algae <- c("Ascidian", "Bare.Space", "Coral", "Coral.Rubble", "Corallimorpharia", "Heteractis.sp.", "Millepora.platyphylla", "No.data", "Sand", "Sacrophyton.sp.", "Shell.Debris", "Soft.Coral", "Sponge", "Tridacna.sp.", "Zooxanthellae", "row_sums")
+data <- read.csv(here::here("data", "allbenthicdata_cleaned_colnames.csv"), stringsAsFactors = F)
+
+not_algae <- c("ascidian", "bare_space", "coral", "coral_rubble", "corallimorpharia", "heteractis_sp", "millepora_platyphylla", "no_data", "sand", "sacrophyton_sp", "shell_debris", "soft_coral", "sponge", "tridacna_sp", "zooxanthellae", "row_sums")
 
@@
 # adding island side - adapted from Julianna's code
-data <- data %>% mutate(Island_side = case_when(Site == "LTER 1" ~ "North",
-                                 Site == "LTER 2" ~ "North",
-                                 Site == "LTER 3" ~ "East",
-                                 Site == "LTER 4" ~ "East",
-                                 Site == "LTER 5" ~ "West",
-                                 Site == "LTER 6" ~ "West"))
-data$Island_side_habitat <- paste(data$Island_side, data$Habitat)
+data <- data %>% mutate(island_side = case_when(site == "lter_1" ~ "north",
+                                 site == "lter_2" ~ "north",
+                                 site == "lter_3" ~ "east",
+                                 site == "lter_4" ~ "east",
+                                 site == "lter_5" ~ "west",
+                                 site == "lter_6" ~ "west"))
+data$island_side_habitat <- paste(data$island_side, data$habitat)
 
@@
 
-meta_cols <- c(colnames(data[1:7] ), "Island_side", "Island_side_habitat", "Stratum")
-species_cols <- colnames(data[8:77])
+meta_cols <- c(colnames(data[1:7] ), "island_side", "island_side_habitat", "stratum")
+species_cols <- colnames(data)[!colnames(data) %in% meta_cols]
 
@@
   cols = 8:77,
-  names_to = "Taxa",
-  values_to = "Percent_cover")
+  names_to = "taxa",
+  values_to = "percent_cover")
 
  transect_level <- long_data %>%
-   dplyr::group_by(Year, Site, Habitat, Transect, Taxa) %>%
-   dplyr::summarise(Mean_cover = mean(Percent_cover), 
-             SE = sd(Percent_cover)/sqrt(n()))
-
-# site_level <- transect_level %>%
-#   dplyr::group_by(Year, Site, Habitat, Taxa) %>%
-#   dplyr::summarise(Mean_site_cover = mean(Mean_cover), 
-#             Site_se = sd(Mean_cover)/sqrt(n()))
+   dplyr::group_by(year, island_side, site, habitat, transect, taxa) %>%
+   dplyr::summarise(mean_cover = mean(percent_cover), 
+             SE = sd(percent_cover)/sqrt(n()))
+
+site_level <- transect_level %>%
+  dplyr::group_by(year, island_side, site, habitat, taxa) %>%
+  dplyr::summarise(mean_site_cover = mean(mean_cover), 
+            site_se = sd(mean_cover)/sqrt(n()))
 ```
@@
 wide_transect_level <- data %>%
-  dplyr::group_by(Year, Island_side, Island_side_habitat, Site, Habitat, Site_habitat, Transect) %>%
+  dplyr::group_by(year, island_side, island_side_habitat, site, habitat, site_habitat, transect) %>%
   dplyr::summarize(across(all_of(species_cols), mean, na.rm = TRUE))
@@
   data.frame(
-    Year = as.factor(wide_transect_level$Year),
-    Island_side = as.factor(wide_transect_level$Island_side),
-    Island_side_habitat = as.factor(wide_transect_level$Island_side_habitat),
-    Site = as.factor(wide_transect_level$Site),
-    Habitat = as.factor(wide_transect_level$Habitat),
-    Site_habitat = as.factor(wide_transect_level$Site_habitat),
-    Transect = as.factor(wide_transect_level$Transect),
-    Site_habitat_transect = as.factor(paste(wide_transect_level$Site_habitat, wide_transect_level$Transect)),
-    Shannon = diversity(wide_transect_level[ , -which(names(wide_transect_level) %in% meta_cols)], "shannon"),
-    Richness = rowSums(wide_transect_level[ , -which(names(wide_transect_level) %in% meta_cols)] > 0)
+    year = as.factor(wide_transect_level$year),
+    island_side = as.factor(wide_transect_level$island_side),
+    island_side_habitat = as.factor(wide_transect_level$island_side_habitat),
+    site = as.factor(wide_transect_level$site),
+    habitat = as.factor(wide_transect_level$habitat),
+    site_habitat = as.factor(wide_transect_level$site_habitat),
+    transect = as.factor(wide_transect_level$transect),
+    site_habitat_transect = as.factor(paste(wide_transect_level$site_habitat, wide_transect_level$transect)),
+    shannon = diversity(wide_transect_level[ , -which(names(wide_transect_level) %in% meta_cols)], "shannon"),
+    richness = rowSums(wide_transect_level[ , -which(names(wide_transect_level) %in% meta_cols)] > 0)
   )
@@
 alpha_diversity_means <- alpha_diversity %>%
-  dplyr::group_by(Year, Island_side, Island_side_habitat, Site, Habitat, Site_habitat) %>%
-  dplyr::summarise(Richness_mean = mean(Richness), 
-            Richness_se = sd(Richness)/sqrt(n()),
-            Shannon_mean = mean(Shannon), 
-            Shannon_se = sd(Shannon)/sqrt(n()))
-
-site_colours <- c("LTER 1" = "cornflowerblue",
-               "LTER 2" = "darkblue",
-               "LTER 3" = "yellowgreen",
-               "LTER 4" = "forestgreen",
-               "LTER 5" = "mediumorchid",
-               "LTER 6" = "orange")
-island_side_colours <- c("North" = "darkblue",
-                         "East" = "mediumorchid",
-                         "West" = "orange")
-habitat_linetypes <- c("Fringing" = "solid",
-                       "Backreef" = "longdash",
-                       "Outer 10" = "twodash",
-                       "Outer 17" = "dotted")
-habitat_shapes <- c("Fringing" = 16,
-                     "Backreef" = 17,
-                     "Outer 10" = 1,
-                     "Outer 17" = 2)
+  dplyr::group_by(year, island_side, island_side_habitat, site, habitat, site_habitat) %>%
+  dplyr::summarise(richness_mean = mean(richness), 
+            richness_se = sd(richness)/sqrt(n()),
+            shannon_mean = mean(shannon), 
+            shannon_se = sd(shannon)/sqrt(n()))
+
+#CALLIE ADDED
+#if you need to export this:
+#alpha_export <- left_join(alpha_diversity, alpha_diversity_means)
+#write_csv(alpha_export, here("data", "algae_alpha_diversity.csv"))
+
+site_colours <- c("lter_1" = "cornflowerblue",
+                  "lter_2" = "darkblue",
+                  "lter_3" = "yellowgreen",
+                  "lter_4" = "forestgreen",
+                  "lter_5" = "mediumorchid",
+                  "lter_6" = "orange")
+island_side_colours <- c("north" = "darkblue",
+                         "east" = "mediumorchid",
+                         "west" = "orange")
+habitat_linetypes <- c("fringing" = "solid",
+                       "backreef" = "longdash",
+                       "outer_10" = "twodash",
+                       "outer_17" = "dotted")
+habitat_shapes <- c("fringing" = 16,
+                     "backreef" = 17,
+                     "outer_10" = 1,
+                     "outer_17" = 2)
 
@@
   ggplot(data = alpha_diversity_means) +
-  geom_point(aes(x = Year, y = Richness_mean, colour = Island_side, shape = Habitat, group = Island_side_habitat), size = 8) +
-  geom_line(aes(x = Year, y = Richness_mean, colour = Island_side, group = Island_side_habitat), size = 1.2) +
-  geom_errorbar(aes(x = Year, ymin = Richness_mean - Richness_se , ymax = Richness_mean + Richness_se, colour = Island_side,
-                    group = Island_side_habitat), width = 0.5, size = 1.2) +
+  geom_point(aes(x = year, y = richness_mean, colour = island_side, shape = habitat, group = island_side_habitat), size = 8) +
+  geom_line(aes(x = year, y = richness_mean, colour = island_side, group = island_side_habitat), size = 1.2) +
+  geom_errorbar(aes(x = year, ymin = richness_mean - richness_se , ymax = richness_mean + richness_se, colour = island_side,
+                    group = island_side_habitat), width = 0.5, size = 1.2) +
   scale_colour_manual(values = island_side_colours) +
   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Habitat)+
-  labs(title = "a. Richness\n", y = "Richness\n") +
+  facet_wrap(~habitat)+
+  labs(title = "a. richness\n", y = "richness\n") +
   theme_bw() + 
@@
   ggplot(data = alpha_diversity_means) +
-  geom_point(aes(x = Year, y = Shannon_mean, colour = Island_side, shape = Habitat, group = Island_side_habitat), size = 8) +
-  geom_line(aes(x = Year, y = Shannon_mean, colour = Island_side, group = Island_side_habitat), size = 1.2) +
-  geom_errorbar(aes(x = Year, ymin = Shannon_mean - Shannon_se , ymax = Shannon_mean + Shannon_se, colour = Island_side,
-                    group = Island_side_habitat), width = 0.5, size = 1.2) +
+  geom_point(aes(x = year, y = shannon_mean, colour = island_side, shape = habitat, group = island_side_habitat), size = 8) +
+  geom_line(aes(x = year, y = shannon_mean, colour = island_side, group = island_side_habitat), size = 1.2) +
+  geom_errorbar(aes(x = year, ymin = shannon_mean - shannon_se , ymax = shannon_mean + shannon_se, colour = island_side,
+                    group = island_side_habitat), width = 0.5, size = 1.2) +
   scale_colour_manual(values = island_side_colours) +
   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Habitat)+
-  labs(title = "b. Shannon\n", y = "shannon\n") +
+  facet_wrap(~habitat)+
+  labs(title = "b. shannon\n", y = "shannon\n") +
   theme_bw() + 
@@
   data.frame(
-    Year = as.numeric(data$Year),
-    Location = as.factor(data$Location),
-    Island_side = as.factor(data$Island_side),
-    Island_side_habitat = as.factor(data$Island_side_habitat),
-    Site = as.factor(data$Site),
-    Habitat = as.factor(data$Habitat),
-    Site_habitat = as.factor(data$Site_habitat),
-    Transect = as.factor(data$Transect),
-    Site_habitat_transect = as.factor(paste(data$Site_habitat, data$Transect)),
-    Shannon = diversity(data[ , -which(names(data) %in% meta_cols)], "shannon"),
-    Richness = rowSums(data[ , -which(names(data) %in% meta_cols)] > 0)
+    year = as.numeric(data$year),
+    location = as.factor(data$location),
+    island_side = as.factor(data$island_side),
+    island_side_habitat = as.factor(data$island_side_habitat),
+    site = as.factor(data$site),
+    habitat = as.factor(data$habitat),
+    site_habitat = as.factor(data$site_habitat),
+    transect = as.factor(data$transect),
+    site_habitat_transect = as.factor(paste(data$site_habitat, data$transect)),
+    shannon = diversity(data[ , -which(names(data) %in% meta_cols)], "shannon"),
+    richness = rowSums(data[ , -which(names(data) %in% meta_cols)] > 0)
   )
@@
 alpha_diversity_quad_means <- alpha_diversity_quad %>%
-  dplyr::group_by(Year, Island_side, Island_side_habitat, Site, Habitat, Site_habitat) %>%
-  dplyr::summarise(Richness_mean = mean(Richness), 
-            Richness_se = sd(Richness)/sqrt(n()),
-            Shannon_mean = mean(Shannon), 
-            Shannon_se = sd(Shannon)/sqrt(n()))
+  dplyr::group_by(year, island_side, island_side_habitat, site, habitat, site_habitat) %>%
+  dplyr::summarise(richness_mean = mean(richness), 
+            richness_se = sd(richness)/sqrt(n()),
+            shannon_mean = mean(shannon), 
+            shannon_se = sd(shannon)/sqrt(n()))
 
@@
   ggplot(data = alpha_diversity_quad_means) +
-  geom_point(aes(x = Year, y = Richness_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8) +
-  geom_line(aes(x = Year, y = Richness_mean, colour = Site, group = Site_habitat), size = 1.2) +
-  geom_errorbar(aes(x = Year, ymin = Richness_mean - Richness_se , ymax = Richness_mean + Richness_se, colour = Site,
-                    group = Site_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Habitat)+
-  labs(title = "a. Richness quad level\n", y = "Richness\n") +
+  geom_point(aes(x = year, y = richness_mean, colour = site, shape = habitat, group = site_habitat), size = 8) +
+  geom_line(aes(x = year, y = richness_mean, colour = site, group = site_habitat), size = 1.2) +
+  geom_errorbar(aes(x = year, ymin = richness_mean - richness_se , ymax = richness_mean + richness_se, colour = site,
+                    group = site_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~habitat)+
+  labs(title = "a. richness quad level\n", y = "richness\n") +
   theme_bw() + 
@@
   ggplot(data = alpha_diversity_quad_means) +
-  geom_point(aes(x = Year, y = Shannon_mean, colour = Island_side, shape = Habitat, group = Island_side_habitat), size = 8) +
-  geom_line(aes(x = Year, y = Shannon_mean, colour = Island_side, group = Island_side_habitat), size = 1.2) +
-  geom_errorbar(aes(x = Year, ymin = Shannon_mean - Shannon_se , ymax = Shannon_mean + Shannon_se, colour = Island_side,
-                    group = Island_side_habitat), width = 0.5, size = 1.2) +
+  geom_point(aes(x = year, y = shannon_mean, colour = island_side, shape = habitat, group = island_side_habitat), size = 8) +
+  geom_line(aes(x = year, y = shannon_mean, colour = island_side, group = island_side_habitat), size = 1.2) +
+  geom_errorbar(aes(x = year, ymin = shannon_mean - shannon_se , ymax = shannon_mean + shannon_se, colour = island_side,
+                    group = island_side_habitat), width = 0.5, size = 1.2) +
   scale_colour_manual(values = island_side_colours) +
   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Habitat)+
-  labs(title = "b. Shannon quad level\n", y = "shannon\n") +
+  facet_wrap(~habitat)+
+  labs(title = "b. shannon quad level\n", y = "shannon\n") +
   theme_bw() + 
@@
 ```
+
 ```{r compare quad and transect level plots, fig.height = 30, fig.width= 40}
@@
 ```
+
 ```{r gam approach}
@@
 # The below takes 17 years to run in and of itself
-# gam_model <- gam(Richness ~ s(Year) + s(Year, by = Habitat) + 
-#                  s(Island_side, bs = "re") + s(Transect, bs = "re") + s(Location, bs = "re"),
+# gam_model <- gam(richness ~ s(year) + s(year, by = habitat) + 
+#                  s(island_side, bs = "re") + s(transect, bs = "re") + s(location, bs = "re"),
 #                  data = alpha_diversity_quad, family = poisson())
 
-gam_columns <- c("Location", "Intercept", "edf") # names of columns for dataframe
+gam_columns <- c("location", "Intercept", "edf") # names of columns for dataframe
 gam_coefficients <- # new dataframe name - same workflow as above
-  data.frame(matrix(nrow = length(unique(alpha_diversity_quad$Location)),
+  data.frame(matrix(nrow = length(unique(alpha_diversity_quad$location)),
                     ncol = length(gam_columns))
@@
 colnames(gam_coefficients) <- gam_columns # change column names to what was specified
-gam_coefficients$Location <- levels(alpha_diversity_quad$Location) # populate Location column with each quad/time series
+gam_coefficients$location <- levels(alpha_diversity_quad$location) # populate location column with each quad/time series
 
 # same workflow as *below (I did gls first originally)
-for(i in 1:length(unique(alpha_diversity_quad$Location)) ){ # for each  quadrat
+for(i in 1:length(unique(alpha_diversity_quad$location)) ){ # for each  quadrat
   # create a new dataframe for the quadrat of the current iteration
-  newData <- subset(alpha_diversity_quad, Location == levels(alpha_diversity_quad$Location)[i])
+  newData <- subset(alpha_diversity_quad, location == levels(alpha_diversity_quad$location)[i])
 
   # fit gam for current time series
-  loop_mod <- gamm(Richness ~ s(Year), data = newData, correlation = corCAR1(form = ~ Year), method = "REML", family = poisson)
+  loop_mod <- gamm(richness ~ s(year), data = newData, correlation = corCAR1(form = ~ year), method = "REML", family = poisson)
   loop_sum <- summary(loop_mod$gam) # store summary data
@@
 hist(gam_coefficients$edf) # generally around 1 except for an outlier
-# LTER 5 Fringing Reef Algae Transect 3 Quad 6 - a quadrat that has almost no algae the entire time series
+# lter_5 fringing Reef Algae transect 3 Quad 6 - a quadrat that has almost no algae the entire time series
 # issues with quadrat - smaller sample = more 0s, makes modeling trickier
-hist(subset(gam_coefficients, Location != "LTER 5 Fringing Reef Algae Transect 3 Quad 6")$edf)
+hist(subset(gam_coefficients, location != "lter_5 fringing Reef Algae transect 3 Quad 6")$edf)
 sum(gam_coefficients$edf > 1.01) # all but 18/1200 quadrats can be fit by a smoother with edf < 1.01
@@
 
-columns <- c("Location", "Intercept", "Slope") # names of columns for dataframe
+columns <- c("location", "Intercept", "Slope") # names of columns for dataframe
 gls_coefficients <- # new dataframe name
-  data.frame(matrix(nrow = length(unique(alpha_diversity_quad$Location)), # 1200 quadrats = 1200 different time series
+  data.frame(matrix(nrow = length(unique(alpha_diversity_quad$location)), # 1200 quadrats = 1200 different time series
                     ncol = length(columns)) # specified to make sure number of columns doesn't exceed what was specified
@@
 colnames(gls_coefficients) <- columns # change column names to what was specified
-gls_coefficients$Location <- levels(alpha_diversity_quad$Location) # populate Location column with each quad/time series
-
-for(i in 1:length(unique(alpha_diversity_quad$Location)) ){ # for each  quadrat
+gls_coefficients$location <- levels(alpha_diversity_quad$location) # populate location column with each quad/time series
+
+for(i in 1:length(unique(alpha_diversity_quad$location)) ){ # for each  quadrat
   # create a new dataframe for the quadrat of the current iteration
-  newData <- subset(alpha_diversity_quad, Location == levels(alpha_diversity_quad$Location)[i])
+  newData <- subset(alpha_diversity_quad, location == levels(alpha_diversity_quad$location)[i])
 
   # fit gls for current time series
-  loop_mod <- gls(Richness ~ Year, correlation =corAR1(form=~Year), data = newData)
+  loop_mod <- gls(richness ~ year, correlation =corAR1(form=~year), data = newData)
 
@@
 # Start adding back in relevant metadata - probably a more elegant way to do this!!!
-habitat_vector <- c(rep("Backreef", 50), rep("Fringing", 50), rep("Outer 10", 50), rep("Outer 17", 50))
+habitat_vector <- c(rep("backreef", 50), rep("fringing", 50), rep("outer_10", 50), rep("outer_17", 50))
 transect_sequence <- c(rep(1,10),
@@
 extra_meta <- 
-  data.frame(Site = as.factor(c(rep("LTER 1", 200),
-                                rep("LTER 2", 200),
-                                rep("LTER 3", 200),
-                                rep("LTER 4", 200),
-                                rep("LTER 5", 200),
-                                rep("LTER 6", 200)) ),
-             Island_side = as.factor(c(rep("North", 400),
-                                     rep("East", 400),
-                                     rep("West", 400))
+  data.frame(site = as.factor(c(rep("lter_1", 200),
+                                rep("lter_2", 200),
+                                rep("lter_3", 200),
+                                rep("lter_4", 200),
+                                rep("lter_5", 200),
+                                rep("lter_6", 200)) ),
+             island_side = as.factor(c(rep("north", 400),
+                                     rep("east", 400),
+                                     rep("west", 400))
                                      ),
-             Habitat = as.factor(rep(habitat_vector, 6 )),
-             Transect = as.factor(rep(transect_sequence, 240)),
-             Location = levels(alpha_diversity_quad$Location) )
-
-gls_coefficients_merged <- merge(gls_coefficients, extra_meta, by = "Location")
-
-slope_mod <- glmmTMB(Slope ~ Habitat*Site + (1|Site/Transect), data = gls_coefficients_merged)
+             habitat = as.factor(rep(habitat_vector, 6 )),
+             transect = as.factor(rep(transect_sequence, 240)),
+             location = levels(alpha_diversity_quad$location) )
+
+gls_coefficients_merged <- merge(gls_coefficients, extra_meta, by = "location")
+
+slope_mod <- glmmTMB(Slope ~ habitat*site + (1|site/transect), data = gls_coefficients_merged)
 # model convergence problem
@@
 #                 Chisq Df Pr(>Chisq)    
-# Habitat      1397.999  3  < 2.2e-16 ***
-# Site           39.095  5  2.273e-07 ***
-# Habitat:Site 1738.155 15  < 2.2e-16 ***
+# habitat      1397.999  3  < 2.2e-16 ***
+# site           39.095  5  2.273e-07 ***
+# habitat:site 1738.155 15  < 2.2e-16 ***
 r.squaredGLMM(slope_mod) # 0.22, 0.24
 plot(allEffects(slope_mod))
-# Backreef: richness increases on East and North shore, but relatively constant west side
+# backreef: richness increases on east and north shore, but relatively constant west side
 # Fringe: richness decreasing on east and west side, but relatvely constant on north shore
-# Outer 10: Richness neutral/decreasing on east side, but neutral/increasing on north and west side
-# Outer 17: Richness decreasing north/east side, but neutral west side
+# outer_10: richness neutral/decreasing on east side, but neutral/increasing on north and west side
+# outer_17: richness decreasing north/east side, but neutral west side
 # Worth noting trends are slight - arguable whether they are biologically meaningful
 
-slope_emm <- emmeans(slope_mod, ~ Habitat*Site) # create emmeans object
+slope_emm <- emmeans(slope_mod, ~ habitat*site) # create emmeans object
 emmeans::contrast(slope_emm, "consec", simple = "each", combine = TRUE, adjust = "bonferroni")
 # Note: not sure I want to include this
-emmeans(slope_mod, pairwise ~ Site | Habitat)
+emmeans(slope_mod, pairwise ~ site | habitat)
 ```
@@
   geom_vline(xintercept = 0, size = 2, linetype = 2) +
-  facet_grid(Habitat ~ Site) +
+  facet_grid(habitat ~ site) +
 #  theme_classic() +
@@
 ```{r gls intercept, fig.height=20, fig.width=20}
-intercept_mod <- glmmTMB(Intercept ~ Habitat*Island_side + (1|Site/Transect), data = gls_coefficients_merged)
+intercept_mod <- glmmTMB(Intercept ~ habitat*island_side + (1|site/transect), data = gls_coefficients_merged)
 summary(intercept_mod) # a lot of variance w/i transect, less so w/i site
@@
 #                        Chisq Df Pr(>Chisq)    
-# Habitat             1307.838  3  < 2.2e-16 ***
-# Island_side           26.358  2  1.889e-06 ***
-# Habitat:Island_side  789.627  6  < 2.2e-16 ***
+# habitat             1307.838  3  < 2.2e-16 ***
+# island_side           26.358  2  1.889e-06 ***
+# habitat:island_side  789.627  6  < 2.2e-16 ***
 r.squaredGLMM(intercept_mod) # 0.16, 0.18
@@
   geom_vline(xintercept = 0, size = 2, linetype = 2) +
-  facet_grid(Habitat ~ Island_side) +
+  facet_grid(habitat ~ island_side) +
 #  theme_classic() +
@@
 ```{r glmm approach}
-alpha_diversity_quad$Year_factor <- as.factor(alpha_diversity_quad$Year) # for glmmTMB ar1 structure
+alpha_diversity_quad$year_factor <- as.factor(alpha_diversity_quad$year) # for glmmTMB ar1 structure
 # Framing - is adiv generally higher in certain habitats/regions?
@@
 # richness_mod <- glmmTMB(
-#   Richness ~ Island_side*Habitat + (1 | Island_side/Site/Transect/Location) + ar1(Year_factor + 0 | Location),
+#   richness ~ island_side*habitat + (1 | island_side/site/transect/location) + ar1(year_factor + 0 | location),
 #   family = poisson,
@@
 richness_mod_2 <- glmmTMB(
-  Richness ~ Year_factor*Habitat*Site + (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),
+  richness ~ year_factor*habitat*site + (1 | site/transect/location) + ar1(year_factor + 0 | location),
   family = poisson,
@@
 plot(allEffects(richness_mod_2))
-# emmeans(richness_mod_2, pairwise ~ Habitat | Year_factor) # if we want to compare habitats year to year
+# emmeans(richness_mod_2, pairwise ~ habitat | year_factor) # if we want to compare habitats year to year
 # model.sel(dredge(richness_mod_2))
 
-top_rich_mod <- glmmTMB(Richness ~ Year_factor*Habitat + Year_factor*Site + (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),   family = poisson, data = alpha_diversity_quad, control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000)))
+top_rich_mod <- glmmTMB(richness ~ year_factor*habitat + year_factor*site + (1 | site/transect/location) + ar1(year_factor + 0 | location),   family = poisson, data = alpha_diversity_quad, control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000)))
 summary(top_rich_mod)
@@
 
-transect_level$Site_habitat <-  paste(transect_level$Site, transect_level$Habitat)
-transect_level$Stratum <- paste(transect_level$Site, "-", transect_level$Habitat, "-", transect_level$Transect)
-
-# long_data$Site_habitat <-  paste(long_data$Site, long_data$Habitat)
-# long_data$Stratum <- paste(long_data$Site, "-", long_data$Habitat, "-", long_data$Transect)
+transect_level$site_habitat <-  paste(transect_level$site, transect_level$habitat)
+transect_level$stratum <- paste(transect_level$site, "-", transect_level$habitat, "-", transect_level$transect)
+
+# long_data$site_habitat <-  paste(long_data$site, long_data$habitat)
+# long_data$stratum <- paste(long_data$site, "-", long_data$habitat, "-", long_data$transect)
 
@@
 total_cover_per_quadrat_year <- long_data %>%
-  group_by(Location, Year) %>%
-  summarize(total_cover = sum(Percent_cover), .groups = 'drop')
+  group_by(location, year) %>%
+  summarize(total_cover = sum(percent_cover), .groups = 'drop')
 
@@
 zero_cover_count <- zero_cover_quadrats %>%
-  group_by(Location) %>%
+  group_by(location) %>%
   summarize(years_with_zero_cover = n())
@@
 
-length(unique(quadrats_multiple_years_zero_cover$Location)) # 144/1200 quadrats have multiple years with no algal cover
+length(unique(quadrats_multiple_years_zero_cover$location)) # 144/1200 quadrats have multiple years with no algal cover
 
@@
 total_cover_per_transect_year <- transect_level %>%
-  group_by(Stratum, Year) %>%
-  summarize(total_cover = sum(Mean_cover), .groups = 'drop')
+  group_by(stratum, year) %>%
+  summarize(total_cover = sum(mean_cover), .groups = 'drop')
 
@@
 zero_cover_transect_count <- zero_cover_transects %>%
-  group_by(Stratum) %>%
+  group_by(stratum) %>%
   summarize(years_with_zero_cover = n())
@@
 # identify quadrats missing timepoints
-years <- unique(long_data$Year)
-quadrats <- unique(long_data$Location)
+years <- unique(long_data$year)
+quadrats <- unique(long_data$location)
 
 # Create a dataframe with all combinations of quadrats and years
-all_combinations <- expand.grid(Location = quadrats, Year = years)
+all_combinations <- expand.grid(location = quadrats, year = years)
 
 # Identify the missing combinations by performing an anti-join with your actual data
-missing_timepoints <- anti_join(all_combinations, long_data, by = c("Location", "Year"))
+missing_timepoints <- anti_join(all_combinations, long_data, by = c("location", "year"))
 
@@
 problem_children <- c(
-                      "LTER 5 Fringing Reef Algae Transect 3 Quad 9",
-                      "LTER 5 Fringing Reef Algae Transect 3 Quad 8",
-                      "LTER 5 Fringing Reef Algae Transect 3 Quad 6",
-                      "LTER 5 Fringing Reef Algae Transect 3 Quad 3",
-                      "LTER 5 Fringing Reef Algae Transect 3 Quad 10")
-
-long_data_sub <- long_data[!long_data$Location %in% problem_children, ] # remove problem children
-
-data_sub <- data[!data$Location %in% problem_children, ]
+                      "lter_5_fringing_reef_algae_transect_3_quad_9",
+                      "lter_5_fringing_reef_algae_transect_3_quad_8",
+                      "lter_5_fringing_reef_algae_transect_3_quad_6",
+                      "lter_5_fringing_reef_algae_transect_3_quad_3",
+                      "lter_5_fringing_reef_algae_transect_3_quad_10")
+
+long_data_sub <- long_data[!long_data$location %in% problem_children, ] # remove problem children
+
+data_sub <- data[!data$location %in% problem_children, ]
 
@@
   df = long_data_sub,
-  time.var = "Year",
-  species.var = "Taxa",
-  abundance.var = "Percent_cover",
-  replicate.var = "Location",
+  time.var = "year",
+  species.var = "taxa",
+  abundance.var = "percent_cover",
+  replicate.var = "location",
   metric = "total"
@@
 # get relevant metadata for turnover metrics
-data_sub <- data[paste(data$Year, data$Location) %in% paste(total_turnover$Year, total_turnover$Location), ]
+data_sub <- data[paste(data$year, data$location) %in% paste(total_turnover$year, total_turnover$location), ]
 
@@
 # merge with turnover df
-total_turnover <- merge(total_turnover, turnover_meta, by = c("Year", "Location"))
-total_turnover$Year_factor <- as.factor(total_turnover$Year)
+total_turnover <- merge(total_turnover, turnover_meta, by = c("year", "location"))
+total_turnover$year_factor <- as.factor(total_turnover$year)
 
@@
 # model
-total_turnover_mod1 <- glmmTMB(Turnover_trans ~ Year_factor*Habitat*Site +
-                                 (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),
+total_turnover_mod1 <- glmmTMB(Turnover_trans ~ year_factor*habitat*site +
+                                 (1 | site/transect/location) + ar1(year_factor + 0 | location),
                                family = beta_family(), data = total_turnover)
@@
 
-total_turnover_mod2 <- glmmTMB(Turnover_trans ~ Year_factor*Habitat*Site + (1 | Site/Transect/Location),
+total_turnover_mod2 <- glmmTMB(Turnover_trans ~ year_factor*habitat*site + (1 | site/transect/location),
                                family = beta_family(), data = total_turnover)
@@
   total_turnover %>%
-  dplyr::group_by(Year, Island_side, Site, Habitat, Site_habitat) %>%
+  dplyr::group_by(year, island_side, site, habitat, site_habitat) %>%
   dplyr::summarise(Turnover_mean = mean(total), 
@@
   ggplot(data = total_turnover_means) +
-  geom_point(aes(x = Year, y = Turnover_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
-  geom_line(aes(x = Year, y = Turnover_mean, colour = Site, group = Site_habitat), size = 1.2) +
-  geom_errorbar(aes(x = Year, ymin = Turnover_mean - Turnover_se , ymax = Turnover_mean + Turnover_se, colour = Site,
-                    group = Site_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site_habitat, nrow = 6)+
+  geom_point(aes(x = year, y = Turnover_mean, colour = site, shape = habitat, group = site_habitat), size = 8, stroke = 2) +
+  geom_line(aes(x = year, y = Turnover_mean, colour = site, group = site_habitat), size = 1.2) +
+  geom_errorbar(aes(x = year, ymin = Turnover_mean - Turnover_se , ymax = Turnover_mean + Turnover_se, colour = site,
+                    group = site_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~site_habitat, nrow = 6)+
   labs(title = "a. Turnover\n", y = "Total turnover\n") +
@@
   df = long_data_sub,
-  time.var = "Year",
-  species.var = "Taxa",
-  abundance.var = "Percent_cover",
-  replicate.var = "Location",
+  time.var = "year",
+  species.var = "taxa",
+  abundance.var = "percent_cover",
+  replicate.var = "location",
   metric = "appearance"
@@
 
-appearance <- merge(turnover_meta, appearance, by = c("Year", "Location"))
+appearance <- merge(turnover_meta, appearance, by = c("year", "location"))
 appearance$Appearance_trans <- sv_trans(appearance$appearance, appearance)
-appearance$Year_factor <- as.factor(appearance$Year)
+appearance$year_factor <- as.factor(appearance$year)
 
 # model
-appearance_turnover_mod1 <- glmmTMB(Appearance_trans ~ Year_factor*Habitat*Site +
-                                 (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),
+appearance_turnover_mod1 <- glmmTMB(Appearance_trans ~ year_factor*habitat*site +
+                                 (1 | site/transect/location) + ar1(year_factor + 0 | location),
                                family = beta_family(), data = appearance)
@@
 
-appearance_turnover_mod2 <- glmmTMB(Appearance_trans ~ Year_factor*Habitat*Site + (1 | Site/Transect/Location),
+appearance_turnover_mod2 <- glmmTMB(Appearance_trans ~ year_factor*habitat*site + (1 | site/transect/location),
                                family = beta_family(), data = appearance)
@@
   appearance %>%
-  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
+  dplyr::group_by(year, site, habitat, site_habitat) %>%
   dplyr::summarise(Appearance_mean = mean(appearance), 
@@
   ggplot(data = appearance_means) +
-  geom_point(aes(x = Year, y = Appearance_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
-  geom_line(aes(x = Year, y = Appearance_mean, colour = Site, group = Site_habitat), size = 1.2) +
-  geom_errorbar(aes(x = Year, ymin = Appearance_mean - Appearance_se , ymax = Appearance_mean + Appearance_se, colour = Site,
-                    group = Site_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site_habitat, nrow = 6)+
+  geom_point(aes(x = year, y = Appearance_mean, colour = site, shape = habitat, group = site_habitat), size = 8, stroke = 2) +
+  geom_line(aes(x = year, y = Appearance_mean, colour = site, group = site_habitat), size = 1.2) +
+  geom_errorbar(aes(x = year, ymin = Appearance_mean - Appearance_se , ymax = Appearance_mean + Appearance_se, colour = site,
+                    group = site_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~site_habitat, nrow = 6)+
   labs(title = "b. Appearances\n", y = "Total turnover\n") +
@@
   df = long_data_sub,
-  time.var = "Year",
-  species.var = "Taxa",
-  abundance.var = "Percent_cover",
-  replicate.var = "Location",
+  time.var = "year",
+  species.var = "taxa",
+  abundance.var = "percent_cover",
+  replicate.var = "location",
   metric = "disappearance"
@@
 
-disappearance <- merge(turnover_meta, disappearance, by = c("Year", "Location"))
+disappearance <- merge(turnover_meta, disappearance, by = c("year", "location"))
 disappearance$Disappearance_trans <- sv_trans(disappearance$disappearance, disappearance)
-disappearance$Year_factor <- as.factor(disappearance$Year)
+disappearance$year_factor <- as.factor(disappearance$year)
 
 # model
-disappearance_turnover_mod1 <- glmmTMB(Disappearance_trans ~ Year_factor*Habitat*Site +
-                                 (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),
+disappearance_turnover_mod1 <- glmmTMB(Disappearance_trans ~ year_factor*habitat*site +
+                                 (1 | site/transect/location) + ar1(year_factor + 0 | location),
                                family = beta_family(), data = disappearance)
@@
 
-disappearance_turnover_mod2 <- glmmTMB(Disappearance_trans ~ Year_factor*Habitat*Site + (1 | Site/Transect/Location),
+disappearance_turnover_mod2 <- glmmTMB(Disappearance_trans ~ year_factor*habitat*site + (1 | site/transect/location),
                                family = beta_family(), data = disappearance)
@@
   disappearance %>%
-  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
+  dplyr::group_by(year, site, habitat, site_habitat) %>%
   dplyr::summarise(Disappearance_mean = mean(disappearance), 
@@
   ggplot(data = disappearance_means) +
-  geom_point(aes(x = Year, y = Disappearance_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
-  geom_line(aes(x = Year, y = Disappearance_mean, colour = Site, group = Site_habitat), size = 1.2) +
-  geom_errorbar(aes(x = Year, ymin = Disappearance_mean - Disappearance_se , ymax = Disappearance_mean + Disappearance_se, colour = Site,
-                    group = Site_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site_habitat, nrow = 6)+
+  geom_point(aes(x = year, y = Disappearance_mean, colour = site, shape = habitat, group = site_habitat), size = 8, stroke = 2) +
+  geom_line(aes(x = year, y = Disappearance_mean, colour = site, group = site_habitat), size = 1.2) +
+  geom_errorbar(aes(x = year, ymin = Disappearance_mean - Disappearance_se , ymax = Disappearance_mean + Disappearance_se, colour = site,
+                    group = site_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~site_habitat, nrow = 6)+
   labs(title = "c. Disappearance\n", y = "Total turnover\n") +
@@
   df = long_data_sub,
-  time.var = "Year",
-  species.var = "Taxa",
-  abundance.var = "Percent_cover",
-  replicate.var = "Location"
-)
-
-rank_shift <- merge(turnover_meta, rank_shift, by = c("Location"))
+  time.var = "year",
+  species.var = "taxa",
+  abundance.var = "percent_cover",
+  replicate.var = "location"
+)
+
+rank_shift <- merge(turnover_meta, rank_shift, by = c("location"))
 
@@
   rank_shift %>%
-  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
+  dplyr::group_by(year, site, habitat, site_habitat) %>%
   dplyr::summarise(Rank_shift_mean = mean(MRS), 
@@
   ggplot(data = rank_shift_means) +
-  geom_point(aes(x = Year, y = Rank_shift_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
-  geom_line(aes(x = Year, y = Rank_shift_mean, colour = Site, group = Site_habitat), size = 1.2) +
-  geom_errorbar(aes(x = Year, ymin = Rank_shift_mean - Rank_shift_se , ymax = Rank_shift_mean + Rank_shift_se, colour = Site,
-                    group = Site_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site_habitat, nrow = 6)+
+  geom_point(aes(x = year, y = Rank_shift_mean, colour = site, shape = habitat, group = site_habitat), size = 8, stroke = 2) +
+  geom_line(aes(x = year, y = Rank_shift_mean, colour = site, group = site_habitat), size = 1.2) +
+  geom_errorbar(aes(x = year, ymin = Rank_shift_mean - Rank_shift_se , ymax = Rank_shift_mean + Rank_shift_se, colour = site,
+                    group = site_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~site_habitat, nrow = 6)+
   labs(title = "a. rank_shift\n", y = "Total rank_shift\n") +
@@
 # rate change
-long_data_sub$Percent_cover_transformed <- wisconsin(long_data_sub$Percent_cover) # wisconsin transform abundance data
+long_data_sub$percent_cover_transformed <- wisconsin(long_data_sub$percent_cover) # wisconsin transform abundance data
 
@@
   df = long_data_sub,
-  time.var = "Year",
-  species.var = "Taxa",
-  abundance.var = "Percent_cover_transformed",
-  replicate.var = "Location"
+  time.var = "year",
+  species.var = "taxa",
+  abundance.var = "percent_cover_transformed",
+  replicate.var = "location"
 ) # summarizes rate of change for each quadrat over the course of the time series
 
-rate_change_table_transformed <- merge(turnover_meta, rate_change_table_transformed, by = "Location")
+rate_change_table_transformed <- merge(turnover_meta, rate_change_table_transformed, by = "location")
 hist(rate_change_table_transformed$rate_change)
 
-rate_change_mod1 <- glmmTMB(rate_change ~ Habitat*Site + (1|Habitat/Site/Transect), data = rate_change_table_transformed)
+rate_change_mod1 <- glmmTMB(rate_change ~ habitat*site + (1|habitat/site/transect), data = rate_change_table_transformed)
 summary(rate_change_mod1)
@@
 #                Chisq Df Pr(>Chisq)    
-# Habitat      187.988  3  < 2.2e-16 ***
-# Site          81.209  5  4.686e-16 ***
-# Habitat:Site 143.638 15  < 2.2e-16 ***
+# habitat      187.988  3  < 2.2e-16 ***
+# site          81.209  5  4.686e-16 ***
+# habitat:site 143.638 15  < 2.2e-16 ***
 plot(allEffects(rate_change_mod1))
-# Outer 17 has fewest differences
-# LTER 1 2 3 changing more than 4 5 6 at outer10
-# Backreef change highest at LTER 1 2
-# Fringing highest at LTER 1 2 6
-emmeans(rate_change_mod1, pairwise ~ Site | Habitat)
+# outer_17 has fewest differences
+# lter_1 2 3 changing more than 4 5 6 at outer10
+# backreef change highest at lter_1 2
+# fringing highest at lter_1 2 6
+emmeans(rate_change_mod1, pairwise ~ site | habitat)
 
@@
   df = long_data_sub,
-  time.var = "Year",
-  species.var = "Taxa",
-  abundance.var = "Percent_cover",
-  replicate.var = "Location"
+  time.var = "year",
+  species.var = "taxa",
+  abundance.var = "percent_cover",
+  replicate.var = "location"
 ) # summarizes rate of change for each quadrat over the course of the time series
 
-rate_change_table_untransformed <- merge(turnover_meta, rate_change_table_untransformed, by = "Location")
+rate_change_table_untransformed <- merge(turnover_meta, rate_change_table_untransformed, by = "location")
 hist(rate_change_table_untransformed$rate_change)
 
-rate_change_mod2 <- glmmTMB(rate_change ~ Habitat*Site + (1|Habitat/Site/Transect), data = rate_change_table_untransformed)
+rate_change_mod2 <- glmmTMB(rate_change ~ habitat*site + (1|habitat/site/transect), data = rate_change_table_untransformed)
 summary(rate_change_mod2)
@@
 #                Chisq Df Pr(>Chisq)    
-# Habitat       39.224  3  1.556e-08 ***
-# Site          17.882  5   0.003098 ** 
-# Habitat:Site 123.014 15  < 2.2e-16 ***
+# habitat       39.224  3  1.556e-08 ***
+# site          17.882  5   0.003098 ** 
+# habitat:site 123.014 15  < 2.2e-16 ***
 plot(allEffects(rate_change_mod1))
 # eyeballing it
-# Outer 17 scarcely changing regardless of site
+# outer_17 scarcely changing regardless of site
 # Rate of change similar in backreef between sites
 # big variance between sites in outer10 and fringing
-# Outer 10: As you go around the island, sites become increasing more "stable" over time (i.e. less shift toward dissimilarity)
-# Effectively the opposite trend in Fringing - more dissimilar over time as you go around the island
-emmeans(rate_change_mod1, pairwise ~ Site | Habitat)
+# outer_10: As you go around the island, sites become increasing more "stable" over time (i.e. less shift toward dissimilarity)
+# Effectively the opposite trend in fringing - more dissimilar over time as you go around the island
+emmeans(rate_change_mod1, pairwise ~ site | habitat)
 
 ggplot(data = rate_change_table_transformed) +
-  geom_boxplot(aes(x = Site, y = rate_change, colour = Site) ) +
-  geom_jitter(aes(x = Site, y = rate_change, colour = Site), width = 0.2) +
-  facet_wrap(~Habitat) +
+  geom_boxplot(aes(x = site, y = rate_change, colour = site) ) +
+  geom_jitter(aes(x = site, y = rate_change, colour = site), width = 0.2) +
+  facet_wrap(~habitat) +
   scale_colour_manual(values = site_colours) +
@@
 ggplot(data = rate_change_table_untransformed) +
-  geom_boxplot(aes(x = Site, y = rate_change, colour = Site) ) +
-  geom_jitter(aes(x = Site, y = rate_change, colour = Site), width = 0.2) +
-  facet_wrap(~Habitat) +
+  geom_boxplot(aes(x = site, y = rate_change, colour = site) ) +
+  geom_jitter(aes(x = site, y = rate_change, colour = site), width = 0.2) +
+  facet_wrap(~habitat) +
   scale_colour_manual(values = site_colours) +
@@
   df = long_data_sub,
-  time.var = "Year",
-  species.var = "Taxa",
-  abundance.var = "Percent_cover",
-  replicate.var = "Location"
-)
-
-rate_change_untransformed_distances <- merge(rate_change_untransformed, turnover_meta, by = "Location")
+  time.var = "year",
+  species.var = "taxa",
+  abundance.var = "percent_cover",
+  replicate.var = "location"
+)
+
+rate_change_untransformed_distances <- merge(rate_change_untransformed, turnover_meta, by = "location")
 
@@
   df = long_data_sub,
-  time.var = "Year",
-  species.var = "Taxa",
-  abundance.var = "Percent_cover_transformed",
-  replicate.var = "Location"
-)
-
-rate_change_transformed_distances <- merge(rate_change_transformed, turnover_meta, by = "Location")
+  time.var = "year",
+  species.var = "taxa",
+  abundance.var = "percent_cover_transformed",
+  replicate.var = "location"
+)
+
+rate_change_transformed_distances <- merge(rate_change_transformed, turnover_meta, by = "location")
 
  ggplot(rate_change_untransformed_distances,
-        aes(x = interval, y = distance, colour = Site, shape = Habitat) )+
-   facet_wrap(~Habitat) + 
+        aes(x = interval, y = distance, colour = site, shape = habitat) )+
+   facet_wrap(~habitat) + 
    geom_point() +
@@
  ggplot(rate_change_transformed_distances,
-        aes(x = interval, y = distance, colour = Site, shape = Habitat) )+
-   facet_wrap(~Habitat) + 
+        aes(x = interval, y = distance, colour = site, shape = habitat) )+
+   facet_wrap(~habitat) + 
    geom_point() +
@@
 #    df = long_data_sub,
-#    time.var = "Year",
-#    species.var = "Taxa",
-#    abundance.var = "Percent_cover",
-#    replicate.var = "Location"
+#    time.var = "year",
+#    species.var = "taxa",
+#    abundance.var = "percent_cover",
+#    replicate.var = "location"
 #   )
@@
 # using Anderson et al. 2013 definitions of portfolio effects - relevant fxns in ecofolio packages
-wide_transect_level$Stratum <- as.factor(paste(wide_transect_level$Site, "-", wide_transect_level$Habitat, "-", wide_transect_level$Transect))
-
-transect_portfolio_columns <- c("Stratum", "Portfolio_effect") # names of columns for dataframe
+wide_transect_level$stratum <- as.factor(paste(wide_transect_level$site, "-", wide_transect_level$habitat, "-", wide_transect_level$transect))
+
+transect_portfolio_columns <- c("stratum", "Portfolio_effect") # names of columns for dataframe
 transect_level_portfolio <- 
-  data.frame(matrix(nrow = length(unique(wide_transect_level$Stratum)), 
+  data.frame(matrix(nrow = length(unique(wide_transect_level$stratum)), 
                     ncol = length(transect_portfolio_columns))
@@
 colnames(transect_level_portfolio) <- transect_portfolio_columns # change column 
-transect_level_portfolio$Stratum <- levels(wide_transect_level$Stratum) # populate
-
-for(i in 1:length(unique(unique(wide_transect_level$Stratum))) ){ # for each  transect
+transect_level_portfolio$stratum <- levels(wide_transect_level$stratum) # populate
+
+for(i in 1:length(unique(unique(wide_transect_level$stratum))) ){ # for each  transect
   # create a new dataframe for current transect
-  newData <- subset(wide_transect_level, Stratum == levels(wide_transect_level$Stratum)[i])
+  newData <- subset(wide_transect_level, stratum == levels(wide_transect_level$stratum)[i])
   current_matrix <- newData[ , -which(names(newData) %in% meta_cols)] # remove metadata columns
@@
 alpha_diversity$Cover <- rowSums(wide_transect_level[ , -which(names(wide_transect_level) %in% meta_cols)])/100
-alpha_diversity$Stratum <- paste(alpha_diversity$Site, "-", alpha_diversity$Habitat, "-", alpha_diversity$Transect)
+alpha_diversity$stratum <- paste(alpha_diversity$site, "-", alpha_diversity$habitat, "-", alpha_diversity$transect)
 lm_synchrony_transect <- codyn::synchrony(df = transect_level,
-                                   time.var = "Year",
-                                   species.var = "Taxa",
-                                   abundance.var = "Mean_cover",
+                                   time.var = "year",
+                                   species.var = "taxa",
+                                   abundance.var = "mean_cover",
                                    metric = "Loreau",
-                                   replicate.var = "Stratum") # Loreau and Mazancourt synchrony
-colnames(lm_synchrony_transect) = c("Stratum", "Loreau_synchrony") # rename cols for merge later
-
-gross_synchrony_transect <- codyn::synchrony(df = transect_level,
-                                   time.var = "Year",
-                                   species.var = "Taxa",
-                                   abundance.var = "Mean_cover",
-                                   metric = "Gross",
-                                   replicate.var = "Stratum") # lots of warnings in this one - examine further later
-colnames(gross_synchrony_transect) = c("Stratum", "Gross_synchrony") # rename cols for merge later
+                                   replicate.var = "stratum") # Loreau and Mazancourt synchrony
+colnames(lm_synchrony_transect) = c("stratum", "Loreau_synchrony") # rename cols for merge later
+
+# gross_synchrony_transect <- codyn::synchrony(df = transect_level,
+#                                    time.var = "year",
+#                                    species.var = "taxa",
+#                                    abundance.var = "mean_cover",
+#                                    metric = "Gross",
+#                                    replicate.var = "stratum") # lots of warnings in this one - examine further later
+# colnames(gross_synchrony_transect) = c("stratum", "Gross_synchrony") # rename cols for merge later
 
 # stability_transect <- community_stability(df = transect_level,
-#                                           time.var = "Year",
-#                                           abundance.var = "Mean_cover",
-#                                           replicate.var = "Stratum") # using codyn measure of stability
+#                                           time.var = "year",
+#                                           abundance.var = "mean_cover",
+#                                           replicate.var = "stratum") # using codyn measure of stability
 # NOTE: codyn community_stability() is just the inverse of CV
 diversity_stability <- alpha_diversity %>%
-  dplyr::group_by(Site, Habitat, Site_habitat, Transect, Stratum) %>%
-  dplyr::summarise(Richness_mean = mean(Richness), 
-            Richness_se = sd(Richness)/sqrt(n()),
-            Richness_cv = CV(Richness),
-            Shannon_mean = mean(Shannon), 
-            Shannon_se = sd(Shannon)/sqrt(n()),
-            Shannon_cv = CV(Shannon),
-            Shannon_stability = 1/CV(Shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
+  dplyr::group_by(site, habitat, site_habitat, transect, stratum) %>%
+  dplyr::summarise(richness_mean = mean(richness), 
+            richness_se = sd(richness)/sqrt(n()),
+            richness_cv = CV(richness),
+            shannon_mean = mean(shannon), 
+            shannon_se = sd(shannon)/sqrt(n()),
+            shannon_cv = CV(shannon),
+            shannon_stability = 1/CV(shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
             Cover_mean = mean(Cover),
@@
 
-diversity_stability <- merge(diversity_stability, lm_synchrony_transect, by = "Stratum") # add in loreau and de Mazancourt synchrony
-diversity_stability <- merge(diversity_stability, gross_synchrony_transect, by = "Stratum")
-diversity_stability <- merge(diversity_stability, transect_level_portfolio, by = "Stratum") # add in portfolio effects
+diversity_stability <- merge(diversity_stability, lm_synchrony_transect, by = "stratum") # add in loreau and de Mazancourt synchrony
+# diversity_stability <- merge(diversity_stability, gross_synchrony_transect, by = "stratum")
+diversity_stability <- merge(diversity_stability, transect_level_portfolio, by = "stratum") # add in portfolio effects
 pairs(~ Cover_mean + Cover_cv + Cover_stability + 
-        Shannon_mean + Shannon_cv + Shannon_stability + 
-        Loreau_synchrony + Gross_synchrony + Portfolio_effect, data = diversity_stability)
+        shannon_mean + shannon_cv + shannon_stability + 
+        Loreau_synchrony + Portfolio_effect, data = diversity_stability)
 ```
@@
   ggplot(data = diversity_stability) +
-  geom_point(aes(x = Shannon_mean, y = Cover_stability, colour = Site, shape = Habitat), size = 8) + 
-   scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-#    stat_smooth(aes(x = Shannon_mean, y = Shannon_cv),method = "lm", se = T, size = 2) + 
+  geom_point(aes(x = shannon_mean, y = Cover_stability, colour = site, shape = habitat), size = 8) + 
+   scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+#    stat_smooth(aes(x = shannon_mean, y = shannon_cv),method = "lm", se = T, size = 2) + 
   theme_bw() + 
@@
   ggplot(data = diversity_stability) +
-  geom_point(aes(x = Shannon_mean, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site) +
+  geom_point(aes(x = shannon_mean, y = Cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~site) +
   theme_bw() + 
@@
 # maybe some heteroskedasticity?
-# just eyeballing but it looks like "stability" is highest in LTER 2 Fringe, which also has low diversity
+# just eyeballing but it looks like "stability" is highest in lter_2 Fringe, which also has low diversity
 # Functional form vs. portfolio effects?
-# Either way looks sick as shit. Great clustering by habitat within site, and diversity trend is strong in LTER 3 4 5
+# Either way looks sick as shit. Great clustering by habitat within site, and diversity trend is strong in lter_3 4 5
 # LTER2 the trend is reversed
-# LTER 1 is mostly driven by habitat
-# LTER 6 potentially a wash
+# lter_1 is mostly driven by habitat
+# lter_6 potentially a wash
 
@@
   ggplot(data = diversity_stability) +
-  geom_point(aes(x = Shannon_mean, y = Shannon_cv, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site) +
+  geom_point(aes(x = shannon_mean, y = shannon_cv, colour = site, shape = habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~site) +
 
@@
   ggplot(data = diversity_stability) +
-  geom_point(aes(x = Shannon_mean, y = Shannon_cv, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-
-#  facet_wrap(~Site) +
+  geom_point(aes(x = shannon_mean, y = shannon_cv, colour = site, shape = habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+
+#  facet_wrap(~site) +
   theme_bw() + 
@@
   ggplot(data = diversity_stability) +
-  geom_point(aes(x = Loreau_synchrony, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site) +
+  geom_point(aes(x = Loreau_synchrony, y = Cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~site) +
 
@@
 
-(gross_stability_plot = 
-  ggplot(data = diversity_stability) +
-  geom_point(aes(x = Gross_synchrony, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site) +
-
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-(portfolio_stability_plot = 
-  ggplot(data = diversity_stability) +
-  geom_point(aes(x = Portfolio_effect, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site) +
-
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         ) # almost a perfect line, which makes sense given that CV is a component of portfolio effects
-```
-
-```{r diversity, portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
-(loreau_diversity_plot = 
-  ggplot(data = diversity_stability) +
-  geom_point(aes(x = Shannon_mean, y = Loreau_synchrony, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site) +
-
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-(gross_diversity_plot = 
-  ggplot(data = diversity_stability) +
-  geom_point(aes(x = Shannon_mean, y = Gross_synchrony, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site) +
-
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-(portfolio_diversity_plot = 
-  ggplot(data = diversity_stability) +
-  geom_point(aes(x = Shannon_mean, y = Portfolio_effect, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site) +
-
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-```
-
-```{r model selection}
-diversity_mod_universal <- glmmTMB(Cover_stability ~ Loreau_synchrony + Shannon_mean*Site*Habitat, family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
-diversity_dredge <- dredge(diversity_mod_universal)
-model.sel(diversity_dredge)
-
-# top model is as follows (by a decent margin):
-diversity_top <- glmmTMB(Cover_stability ~ Loreau_synchrony + Shannon_mean*Site + Site*Habitat, family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
-summary(diversity_top)
-Anova(diversity_top)
-```
-
-```{r general cover diversity, fig.height = 30, fig.width = 30}
-(diversity_cover_plot = 
-  ggplot(data = alpha_diversity) +
-  geom_point(aes(x = Cover, y = Shannon, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Year) +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         ) # ABORT ABORT ABORT
-```
-
-```{r diversity stability quadrat level setup, fig.height = 20, fig.width=20}
-data_sub$Stratum <- as.factor(paste(data_sub$Site, "-", data_sub$Habitat, "-", data_sub$Transect))
-data_sub$Location <- as.factor(data_sub$Location)
-
-quadrat_portfolio_columns <- c("Location", "Portfolio_effect") # names of columns for dataframe
-quadrat_level_portfolio <- 
-  data.frame(matrix(nrow = length(unique(data_sub$Location)), 
-                    ncol = length(quadrat_portfolio_columns))
-             )
-colnames(quadrat_level_portfolio) <- quadrat_portfolio_columns # change column 
-quadrat_level_portfolio$Location <- unique(data_sub$Location) # populate
-
-for(i in 1:length(unique(unique(data_sub$Location))) ){ # for each  quadrat
-  # create a new dataframe for current quadrat
-  newData <- subset(data, Location == levels(data_sub$Location)[i])
-  current_matrix <- newData[ , -which(names(newData) %in% meta_cols)] # remove metadata columns
-  if(sum(current_matrix) == 0 | # if this was a 0 algae quadrat
-     # or if there was only 1 column after removing all irrelevant algae
-     is.null(ncol(current_matrix[, which(colSums(current_matrix) != 0)]))){ 
-    quadrat_level_portfolio[i, 2] <- NA # set portfolio effects to 0
-  }else{ # otherwise
-    current_matrix <- current_matrix[, which(colSums(current_matrix) != 0)] # remove irrelevant algae
-
-    quadrat_level_portfolio[i, 2] <- pe_mv(current_matrix, "loess_detrended") # calculate portfolio effects for current time series
-  }
-
-}
-
-# sum(is.na(quadrat_level_portfolio$Portfolio_effect)) # 4 quadrats causing trouble
-
-hist(quadrat_level_portfolio$Portfolio_effect) # some crazy outliers due to low richness
-
-alpha_diversity_quad$Cover <- rowSums(data[ , -which(names(data) %in% meta_cols)])/100
-alpha_diversity_quad_sub <- alpha_diversity_quad[!alpha_diversity_quad$Location %in% problem_children, ]
-
-# alpha_diversity_quad_sub$Cover <- rowSums(data_sub[ , -which(names(data_sub) %in% meta_cols)])/100
-alpha_diversity_quad_sub$Stratum <- 
-  paste(alpha_diversity_quad_sub$Site, "-", alpha_diversity_quad_sub$Habitat, "-", alpha_diversity_quad_sub$Transect)
-lm_synchrony_quadrat <- codyn::synchrony(df = long_data_sub,
-                                   time.var = "Year",
-                                   species.var = "Taxa",
-                                   abundance.var = "Percent_cover",
-                                   metric = "Loreau",
-                                   replicate.var = "Location") # Loreau and Mazancourt synchrony
-colnames(lm_synchrony_quadrat) = c("Location", "Loreau_synchrony") # rename cols for merge later
-
-## gross_synchrony_quadrat <- codyn::synchrony(df = long_data_sub,
-##                                    time.var = "Year",
-##                                    species.var = "Taxa",
-##                                    abundance.var = "Percent_cover",
-##                                    metric = "Gross",
-##                                    replicate.var = "Location") # lots of warnings in this one - examine further later
-## colnames(gross_synchrony_transect) = c("Location", "Gross_synchrony") # rename cols for merge later
-
-variance_ratio_quadrat <- variance_ratio(df = long_data_sub,
-                                   time.var = "Year",
-                                   species.var = "Taxa",
-                                   abundance.var = "Percent_cover",
-                                   average.replicates = F,
-                                   bootnumber = 1,
-                                   replicate.var = "Location")
-
-diversity_stability_quadrat <- alpha_diversity_quad_sub %>%
-  dplyr::group_by(Location, Site, Habitat, Site_habitat, Transect, Stratum) %>%
-  dplyr::summarise(Richness_mean = mean(Richness), 
-            Richness_se = sd(Richness)/sqrt(n()),
-            Richness_cv = CV(Richness),
-            Shannon_mean = mean(Shannon), 
-            Shannon_se = sd(Shannon)/sqrt(n()),
-            Shannon_cv = CV(Shannon),
-            Shannon_stability = 1/CV(Shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
-            Cover_mean = mean(Cover),
-            Cover_se = sd(Cover)/sqrt(n()),
-            Cover_cv = CV(Cover),
-            Cover_stability = 1/CV(Cover)) # as above
-
-diversity_stability_quadrat <- 
-  merge(diversity_stability_quadrat, lm_synchrony_quadrat, by = "Location") # add in loreau and de Mazancourt synchrony
-diversity_stability_quadrat <- merge(diversity_stability_quadrat, gross_synchrony_quadrat, by = "Location")
-diversity_stability_quadrat <- merge(diversity_stability_quadrat, quadrat_level_portfolio, by = "Location") # add in portfolio effects
-diversity_stability_quadrat <- merge(diversity_stability_quadrat, variance_ratio_quadrat, by = "Location")
-pairs(~ Cover_mean + Cover_cv + Cover_stability + 
-        Shannon_mean + Shannon_cv + Shannon_stability + 
-        Loreau_synchrony +  Portfolio_effect + VR, data = diversity_stability_quadrat)
-
-```
-
-```{r cover and cv, fig.height = 8, fig.width = 10}
-(cover_cv_plot = 
-  ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Cover_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 4) + 
-   scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-   facet_wrap(~Site) +
-
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-(cover_cv_plot = 
-  ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Cover_mean, y = Shannon_mean, colour = Site, shape = Habitat), size = 4) + 
-   scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-   facet_wrap(~Site) +
-
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-```
-
-```{r basic diversity stability plots quadrat level, fig.height=8, fig.width=10}
-(shannon_cv_plot = 
-  ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 4) + 
-   scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-   facet_wrap(~Site) +
-
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-(shannon_cv_plot_2 = 
-  ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 4, stroke = 1) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Habitat) +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-# LTER 2 obscures a lot of the trends
-
-```
-
-```{r portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
-(loreau_stability_plot_quad = 
-  ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Loreau_synchrony, y = Cover_cv, colour = Site, shape = Habitat), size = 4, stroke = 1) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site) +
-
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-# (gross_stability_plot_quad = 
-#   ggplot(data = diversity_stability_quadrat) +
-#   geom_point(aes(x = Gross_synchrony, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
+# (gross_stability_plot = 
+#   ggplot(data = diversity_stability) +
+#   geom_point(aes(x = Gross_synchrony, y = Cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
 #    scale_colour_manual(values = site_colours) +
 #    scale_shape_manual(values = habitat_shapes) +
-#   facet_wrap(~Site) +
+#   facet_wrap(~site) +
 #    
@@
 #          )
-
-(portfolio_stability_plot_quad = 
-  ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Portfolio_effect, y = Cover_cv, colour = Site, shape = Habitat), size = 4, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
- # facet_wrap(~Site) +
+# 
+(portfolio_stability_plot = 
+  ggplot(data = diversity_stability) +
+  geom_point(aes(x = Portfolio_effect, y = Cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~site) +
 
@@
          ) # almost a perfect line, which makes sense given that CV is a component of portfolio effects
-
-```
-
-```{r diversity portfolio effects and synchrony quadrat level,  fig.height = 8, fig.width = 10}
-(loreau_diversity_plot_quad = 
-  ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Shannon_mean, y = Loreau_synchrony, colour = Site, shape = Habitat), size = 4, stroke = 1) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site) +
+```
+
+```{r diversity, portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
+(loreau_diversity_plot = 
+  ggplot(data = diversity_stability) +
+  geom_point(aes(x = shannon_mean, y = Loreau_synchrony, colour = site, shape = habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~site) +
 
@@
 
-(portfolio_diversity_plot_quad = 
-  ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Shannon_mean, y = Portfolio_effect, colour = Site, shape = Habitat), size = 4, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~Site) +
+# (gross_diversity_plot = 
+#   ggplot(data = diversity_stability) +
+#   geom_point(aes(x = shannon_mean, y = Gross_synchrony, colour = site, shape = habitat), size = 8, stroke = 2) + 
+#    scale_colour_manual(values = site_colours) +
+#    scale_shape_manual(values = habitat_shapes) +
+#   facet_wrap(~site) +
+#    
+#   theme_bw() + 
+#   theme(axis.title.x = element_text(size=18), 
+#         axis.title.y = element_text(size=18), 
+#         panel.background = element_blank(), 
+#         panel.grid.major = element_blank(),  #remove major-grid labels
+#         panel.grid.minor = element_blank(),  #remove minor-grid labels
+#         plot.background = element_blank(),
+#         legend.text = element_text(size = rel(1)), # increase legend text
+#         legend.title = element_text(size = rel(1)),
+#         legend.position = 'bottom', # move legend to bottom
+#         legend.box = 'horizontal', # make legend horizontal
+#         legend.box.background = element_rect(colour = "black"))  # add box around legend
+#          )
+
+(portfolio_diversity_plot = 
+  ggplot(data = diversity_stability) +
+  geom_point(aes(x = shannon_mean, y = Portfolio_effect, colour = site, shape = habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~site) +
 
@@
 
+```{r model selection}
+diversity_mod_universal <- glmmTMB(Cover_stability ~ Loreau_synchrony + shannon_mean*site*habitat, family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
+diversity_dredge <- dredge(diversity_mod_universal)
+model.sel(diversity_dredge)
+
+# top model is as follows (by a decent margin):
+diversity_top <- glmmTMB(Cover_stability ~ Loreau_synchrony + shannon_mean*site + site*habitat, family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
+summary(diversity_top)
+Anova(diversity_top)
+```
+
+```{r general cover diversity, fig.height = 30, fig.width = 30}
+(diversity_cover_plot = 
+  ggplot(data = alpha_diversity) +
+  geom_point(aes(x = Cover, y = shannon, colour = site, shape = habitat), size = 8, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~year) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         ) # ABORT ABORT ABORT
+```
+
+```{r diversity stability quadrat level setup, fig.height = 20, fig.width=20}
+data_sub$stratum <- as.factor(paste(data_sub$site, "-", data_sub$habitat, "-", data_sub$transect))
+data_sub$location <- as.factor(data_sub$location)
+
+quadrat_portfolio_columns <- c("location", "Portfolio_effect") # names of columns for dataframe
+quadrat_level_portfolio <- 
+  data.frame(matrix(nrow = length(unique(data_sub$location)), 
+                    ncol = length(quadrat_portfolio_columns))
+             )
+colnames(quadrat_level_portfolio) <- quadrat_portfolio_columns # change column 
+quadrat_level_portfolio$location <- unique(data_sub$location) # populate
+
+for(i in 1:length(unique(unique(data_sub$location))) ){ # for each  quadrat
+  # create a new dataframe for current quadrat
+  newData <- subset(data, location == levels(data_sub$location)[i])
+  current_matrix <- newData[ , -which(names(newData) %in% meta_cols)] # remove metadata columns
+  if(sum(current_matrix) == 0 | # if this was a 0 algae quadrat
+     # or if there was only 1 column after removing all irrelevant algae
+     is.null(ncol(current_matrix[, which(colSums(current_matrix) != 0)]))){ 
+    quadrat_level_portfolio[i, 2] <- NA # set portfolio effects to 0
+  }else{ # otherwise
+    current_matrix <- current_matrix[, which(colSums(current_matrix) != 0)] # remove irrelevant algae
+
+    quadrat_level_portfolio[i, 2] <- pe_mv(current_matrix, "loess_detrended") # calculate portfolio effects for current time series
+  }
+
+}
+
+# sum(is.na(quadrat_level_portfolio$Portfolio_effect)) # 4 quadrats causing trouble
+
+hist(quadrat_level_portfolio$Portfolio_effect) # some crazy outliers due to low richness
+
+alpha_diversity_quad$Cover <- rowSums(data[ , -which(names(data) %in% meta_cols)])/100
+alpha_diversity_quad_sub <- alpha_diversity_quad[!alpha_diversity_quad$location %in% problem_children, ]
+
+# alpha_diversity_quad_sub$Cover <- rowSums(data_sub[ , -which(names(data_sub) %in% meta_cols)])/100
+alpha_diversity_quad_sub$stratum <- 
+  paste(alpha_diversity_quad_sub$site, "-", alpha_diversity_quad_sub$habitat, "-", alpha_diversity_quad_sub$transect)
+lm_synchrony_quadrat <- codyn::synchrony(df = long_data_sub,
+                                   time.var = "year",
+                                   species.var = "taxa",
+                                   abundance.var = "percent_cover",
+                                   metric = "Loreau",
+                                   replicate.var = "location") # Loreau and Mazancourt synchrony
+colnames(lm_synchrony_quadrat) = c("location", "Loreau_synchrony") # rename cols for merge later
+
+## gross_synchrony_quadrat <- codyn::synchrony(df = long_data_sub,
+##                                    time.var = "year",
+##                                    species.var = "taxa",
+##                                    abundance.var = "percent_cover",
+##                                    metric = "Gross",
+##                                    replicate.var = "location") # lots of warnings in this one - examine further later
+## colnames(gross_synchrony_transect) = c("location", "Gross_synchrony") # rename cols for merge later
+
+variance_ratio_quadrat <- variance_ratio(df = long_data_sub,
+                                   time.var = "year",
+                                   species.var = "taxa",
+                                   abundance.var = "percent_cover",
+                                   average.replicates = F,
+                                   bootnumber = 1,
+                                   replicate.var = "location")
+
+community_stability_quadrat <- community_stability(df = long_data_sub,
+                                   time.var = "year",
+                                   abundance.var = "percent_cover",
+                                   replicate.var = "location")
+colnames(community_stability_quadrat) = c("location", "Cover_stability")
+
+diversity_stability_quadrat <- alpha_diversity_quad_sub %>%
+  dplyr::group_by(location, site, habitat, site_habitat, transect, stratum) %>%
+  dplyr::summarise(richness_mean = mean(richness), 
+            richness_se = sd(richness)/sqrt(n()),
+            richness_cv = CV(richness),
+            shannon_mean = mean(shannon), 
+            shannon_se = sd(shannon)/sqrt(n()),
+            shannon_cv = CV(shannon),
+            shannon_stability = 1/CV(shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
+            Cover_mean = mean(Cover),
+            Cover_se = sd(Cover)/sqrt(n()),
+            Cover_cv = CV(Cover))
+
+diversity_stability_quadrat <- 
+  merge(diversity_stability_quadrat, lm_synchrony_quadrat, by = "location") # add in loreau and de Mazancourt synchrony
+# diversity_stability_quadrat <- merge(diversity_stability_quadrat, gross_synchrony_quadrat, by = "location")
+diversity_stability_quadrat <- merge(diversity_stability_quadrat, quadrat_level_portfolio, by = "location") # add in portfolio effects
+diversity_stability_quadrat <- merge(diversity_stability_quadrat, variance_ratio_quadrat, by = "location")
+diversity_stability_quadrat <- merge(diversity_stability_quadrat, community_stability_quadrat, by = "location")
+pairs(~ Cover_mean + Cover_cv + Cover_stability + 
+        shannon_mean + shannon_cv + shannon_stability + 
+        Loreau_synchrony +  Portfolio_effect + VR, data = diversity_stability_quadrat)
+
+```
+
+```{r cover and cv, fig.height = 8, fig.width = 10}
+(cover_cv_plot = 
+  ggplot(data = diversity_stability_quadrat) +
+  geom_point(aes(x = Cover_mean, y = Cover_cv, colour = site, shape = habitat), size = 4) + 
+   scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+   facet_wrap(~site) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+(cover_cv_plot = 
+  ggplot(data = diversity_stability_quadrat) +
+  geom_point(aes(x = Cover_mean, y = shannon_mean, colour = site, shape = habitat), size = 4) + 
+   scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+   facet_wrap(~site) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+```
+
+```{r basic diversity stability plots quadrat level, fig.height=8, fig.width=10}
+(shannon_cv_plot = 
+  ggplot(data = diversity_stability_quadrat) +
+  geom_point(aes(x = shannon_mean, y = Cover_cv, colour = site, shape = habitat), size = 4) + 
+   scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+   facet_wrap(~site) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+(shannon_cv_plot_2 = 
+  ggplot(data = diversity_stability_quadrat) +
+  geom_point(aes(x = shannon_mean, y = Cover_cv, colour = site, shape = habitat), size = 4, stroke = 1) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~habitat) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+# lter_2 obscures a lot of the trends
+
+```
+
+```{r portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
+(loreau_stability_plot_quad = 
+  ggplot(data = diversity_stability_quadrat) +
+  geom_point(aes(x = Loreau_synchrony, y = Cover_cv, colour = site, shape = habitat), size = 4, stroke = 1) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~site) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+# (gross_stability_plot_quad = 
+#   ggplot(data = diversity_stability_quadrat) +
+#   geom_point(aes(x = Gross_synchrony, y = Cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
+#    scale_colour_manual(values = site_colours) +
+#    scale_shape_manual(values = habitat_shapes) +
+#   facet_wrap(~site) +
+#    
+#   theme_bw() + 
+#   theme(axis.title.x = element_text(size=18), 
+#         axis.title.y = element_text(size=18), 
+#         panel.background = element_blank(), 
+#         panel.grid.major = element_blank(),  #remove major-grid labels
+#         panel.grid.minor = element_blank(),  #remove minor-grid labels
+#         plot.background = element_blank(),
+#         legend.text = element_text(size = rel(1)), # increase legend text
+#         legend.title = element_text(size = rel(1)),
+#         legend.position = 'bottom', # move legend to bottom
+#         legend.box = 'horizontal', # make legend horizontal
+#         legend.box.background = element_rect(colour = "black"))  # add box around legend
+#          )
+
+(portfolio_stability_plot_quad = 
+  ggplot(data = diversity_stability_quadrat) +
+  geom_point(aes(x = Portfolio_effect, y = Cover_cv, colour = site, shape = habitat), size = 4, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+ # facet_wrap(~site) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         ) # almost a perfect line, which makes sense given that CV is a component of portfolio effects
+
+```
+
+```{r diversity portfolio effects and synchrony quadrat level,  fig.height = 8, fig.width = 10}
+(loreau_diversity_plot_quad = 
+  ggplot(data = diversity_stability_quadrat) +
+  geom_point(aes(x = shannon_mean, y = Loreau_synchrony, colour = site, shape = habitat), size = 4, stroke = 1) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~site) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+(portfolio_diversity_plot_quad = 
+  ggplot(data = diversity_stability_quadrat) +
+  geom_point(aes(x = shannon_mean, y = Portfolio_effect, colour = site, shape = habitat), size = 4, stroke = 2) + 
+   scale_colour_manual(values = site_colours) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~site) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+```
+
 ```{r diversity models quad level}
-diversity_mod_universal <- glmmTMB(Cover_cv ~Shannon_mean*Site + Shannon_mean*Habitat + Site*Habitat + (1|Transect/Location), family = Gamma(link = "log"), data = diversity_stability_quadrat, na.action = na.pass)
+diversity_mod_universal <- glmmTMB(Cover_cv ~shannon_mean*site + shannon_mean*habitat + site*habitat + (1|transect/location), family = Gamma(link = "log"), data = diversity_stability_quadrat, na.action = na.pass)
 diversity_dredge <- dredge(diversity_mod_universal)
@@
 
-synchrony_mod_universal <- glmmTMB(Cover_cv ~Loreau_synchrony*Site + Loreau_synchrony*Habitat + Site*Habitat + (1|Transect/Location), family = Gamma(link = "log"), data = diversity_stability_quadrat, na.action = na.pass)
+synchrony_mod_universal <- glmmTMB(Cover_cv ~Loreau_synchrony*site + Loreau_synchrony*habitat + site*habitat + (1|transect/location), family = Gamma(link = "log"), data = diversity_stability_quadrat, na.action = na.pass)
 synchrony_dredge <- dredge(synchrony_mod_universal)
@@
 diversity_stability_quadrat$Loreau_synchrony_trans <- sv_trans(diversity_stability_quadrat$Loreau_synchrony, diversity_stability_quadrat) # for modelling on beta
-synchrony_diversity_universal <- glmmTMB(Loreau_synchrony_trans ~Shannon_mean*Site + Shannon_mean*Habitat + Site*Habitat + (1|Transect/Location), family = beta_family(), data = diversity_stability_quadrat, na.action = na.pass)
+synchrony_diversity_universal <- glmmTMB(Loreau_synchrony_trans ~shannon_mean*site + shannon_mean*habitat + site*habitat + (1|transect/location), family = beta_family(), data = diversity_stability_quadrat, na.action = na.pass)
 model.sel(dredge(synchrony_diversity_universal)) # best model as specified
@@
 species_matrix <- species_matrix/100 # convert from percent to proportion
-rownames(species_matrix) <- paste(data$Year, data$Site_habitat, data$Transect)
+rownames(species_matrix) <- paste(data$year, data$site_habitat, data$transect)
 rows_to_remove <- rowSums(species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
@@
 data_subset <- data[!rows_to_remove,] # make sure rows are also removed from dataframe
-data_subset$Stratum <- paste(data_subset$Year, "-", data_subset$Site, "-", data_subset$Habitat )
+data_subset$stratum <- paste(data_subset$year, "-", data_subset$site, "-", data_subset$habitat )
 
@@
 # add back in metadata
-site.scores$Year <- as.factor(data_subset$Year)
-site.scores$Site <- as.factor(data_subset$Site)
-site.scores$Habitat <- as.factor(data_subset$Habitat)
-site.scores$Site_habitat <- as.factor(data_subset$Site_habitat)
-site.scores$Transect <- as.factor(data_subset$Transect)
-site.scores$Stratum <- as.factor(data_subset$Stratum)
-
-cent <- aggregate(cbind(NMDS1, NMDS2) ~ Stratum, data = site.scores, FUN = mean) # compute centroids for NMDS1 and NMDS2
+site.scores$year <- as.factor(data_subset$year)
+site.scores$site <- as.factor(data_subset$site)
+site.scores$habitat <- as.factor(data_subset$habitat)
+site.scores$site_habitat <- as.factor(data_subset$site_habitat)
+site.scores$transect <- as.factor(data_subset$transect)
+site.scores$stratum <- as.factor(data_subset$stratum)
+
+cent <- aggregate(cbind(NMDS1, NMDS2) ~ stratum, data = site.scores, FUN = mean) # compute centroids for NMDS1 and NMDS2
 
@@
   cent %>%
-    separate(Stratum, into = c("Year", "Site", "Habitat"), sep = "-")
-cent$Site_habitat <- paste(cent$Site, cent$Habitat)
+    separate(stratum, into = c("year", "site", "habitat"), sep = "-")
+cent$site_habitat <- paste(cent$site, cent$habitat)
 # Add back in stratum column
-cent$Stratum <-  aggregate(cbind(NMDS1, NMDS2) ~ Stratum, data = site.scores, FUN = mean)$Stratum
+cent$stratum <-  aggregate(cbind(NMDS1, NMDS2) ~ stratum, data = site.scores, FUN = mean)$stratum
 # A lot of extra spaces
@@
   mutate(across(everything(), str_trim))
-cent$Year <- as.factor(cent$Year)
-cent$Habitat <- as.factor(cent$Habitat)
-cent$Site <- as.factor(cent$Site)
-cent$Site_habitat <- as.factor(cent$Site_habitat)
+cent$year <- as.factor(cent$year)
+cent$habitat <- as.factor(cent$habitat)
+cent$site <- as.factor(cent$site)
+cent$site_habitat <- as.factor(cent$site_habitat)
 cent$NMDS1 <- as.numeric(cent$NMDS1) # read in as characters for some reason
@@
 ggplot() + 
-  geom_point(data= site.scores, aes(x=NMDS1,y=NMDS2, shape = Habitat, colour = Site),
+  geom_point(data= site.scores, aes(x=NMDS1,y=NMDS2, shape = habitat, colour = site),
              alpha = 0.25, size=2) + # add the point markers
@@
 
-  geom_path(data = cent, aes(x = NMDS1, y = NMDS2, group = Site_habitat, colour = Site),
+  geom_path(data = cent, aes(x = NMDS1, y = NMDS2, group = site_habitat, colour = site),
             arrow = arrow(length=unit(0.5,"cm"), ends="last", type = "closed"), size = 1) +
-  facet_wrap(~factor(Habitat, levels=c('Fringing','Backreef','Outer 10','Outer 17'))) +
+  facet_wrap(~factor(habitat, levels=c('fringing','backreef','outer_10','outer_17'))) +
   scale_linetype_manual(values = habitat_linetypes) +
@@
 
-# looks like stochasticity increases from outer 17 --> Fringing
-# Even in lagoon though it looks like communities return to initial state... except for North Shore LTER sites
+# looks like stochasticity increases from outer_17 --> fringing
+# Even in lagoon though it looks like communities return to initial state... except for north Shore lter_sites
 # ways to test stochasticity vs. determinism?
@@
  bray_matrix <- vegdist(species_matrix, method = "bray")
-# year_beta_disper <- vegan::betadisper(bray_matrix, data_subset$Year)
+# year_beta_disper <- vegan::betadisper(bray_matrix, data_subset$year)
 # year_beta_disper #
@@
 # plot(year_beta_disper, sub = "") # lol
-# boxplot(year_beta_disper, main = "Year", xlab = "") 
+# boxplot(year_beta_disper, main = "year", xlab = "") 
 # 
-# site_beta_disper <- vegan::betadisper(bray_matrix, data_subset$Site)
+# site_beta_disper <- vegan::betadisper(bray_matrix, data_subset$site)
 # plot(site_beta_disper, sub = "") # site five stands out
-# boxplot(site_beta_disper, main = "Site", xlab = "") 
+# boxplot(site_beta_disper, main = "site", xlab = "") 
 # 
 # 
-# habitat_beta_disper <- vegan::betadisper(bray_matrix, data_subset$Habitat)
+# habitat_beta_disper <- vegan::betadisper(bray_matrix, data_subset$habitat)
 # plot(habitat_beta_disper, sub = "") 
-# boxplot(habitat_beta_disper, main = "Habitat", xlab = "") # greatest variance in fringing it seems (as expected)
+# boxplot(habitat_beta_disper, main = "habitat", xlab = "") # greatest variance in fringing it seems (as expected)
 # 
-# site_habitat_beta_disper <- vegan::betadisper(bray_matrix, data_subset$Site_habitat)
+# site_habitat_beta_disper <- vegan::betadisper(bray_matrix, data_subset$site_habitat)
 # plot(site_habitat_beta_disper, sub = "") 
-# boxplot(site_habitat_beta_disper, main = "Site_habitat", xlab = "") 
+# boxplot(site_habitat_beta_disper, main = "site_habitat", xlab = "") 
 # 
-stratum_beta_disper <- vegan::betadisper(bray_matrix, data_subset$Stratum)
+stratum_beta_disper <- vegan::betadisper(bray_matrix, data_subset$stratum)
 # plot(site_habitat_beta_disper, sub = "") 
-boxplot(stratum_beta_disper, main = "Site_habitat", xlab = "")
+boxplot(stratum_beta_disper, main = "site_habitat", xlab = "")
 stratum_distances <- as.data.frame(stratum_beta_disper$distances)
 colnames(stratum_distances)[1] <- "Distance"
-stratum_distances$Stratum <- data_subset$Stratum
-stratum_distances$Transect <- data_subset$Transect
-stratum_distances$Year <- data_subset$Year
-stratum_distances$Site <- data_subset$Site
-stratum_distances$Habitat <- data_subset$Habitat
-stratum_distances$Site_habitat <- data_subset$Site_habitat
+stratum_distances$stratum <- data_subset$stratum
+stratum_distances$transect <- data_subset$transect
+stratum_distances$year <- data_subset$year
+stratum_distances$site <- data_subset$site
+stratum_distances$habitat <- data_subset$habitat
+stratum_distances$site_habitat <- data_subset$site_habitat
 
 beta_disper_means <- stratum_distances %>%
-  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
+  dplyr::group_by(year, site, habitat, site_habitat) %>%
   dplyr::summarise(Distance_mean = mean(Distance), 
@@
   ggplot(data = beta_disper_means) +
-  geom_point(aes(x = Year, y = Distance_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
-  geom_line(aes(x = Year, y = Distance_mean, colour = Site, group = Site_habitat), size = 1.2) +
-  geom_errorbar(aes(x = Year, ymin = Distance_mean - Distance_se , ymax = Distance_mean + Distance_se, colour = Site,
-                    group = Site_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~factor(Habitat, levels=c('Fringing','Backreef','Outer 10','Outer 17')))+
+  geom_point(aes(x = year, y = Distance_mean, colour = site, shape = habitat, group = site_habitat), size = 8, stroke = 2) +
+  geom_line(aes(x = year, y = Distance_mean, colour = site, group = site_habitat), size = 1.2) +
+  geom_errorbar(aes(x = year, ymin = Distance_mean - Distance_se , ymax = Distance_mean + Distance_se, colour = site,
+                    group = site_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~factor(habitat, levels=c('fringing','backreef','outer_10','outer_17')))+
   labs(title = "Distance to centroid\n", y = "Distance\n") +

### 2024-08-05 noam_data_exploration_v3.Rmd
in the middle of creating site level csvs to see if they match J but need to run again.

@@
 
-transect_portfolio_columns <- c("stratum", "Portfolio_effect") # names of columns for dataframe
+transect_portfolio_columns <- c("stratum", "portfolio_effect") # names of columns for dataframe
 transect_level_portfolio <- 
@@
 
-alpha_diversity$Cover <- rowSums(wide_transect_level[ , -which(names(wide_transect_level) %in% meta_cols)])/100
+alpha_diversity$cover <- rowSums(wide_transect_level[ , -which(names(wide_transect_level) %in% meta_cols)])/100
 alpha_diversity$stratum <- paste(alpha_diversity$site, "-", alpha_diversity$habitat, "-", alpha_diversity$transect)
@@
                                    replicate.var = "stratum") # Loreau and Mazancourt synchrony
-colnames(lm_synchrony_transect) = c("stratum", "Loreau_synchrony") # rename cols for merge later
+colnames(lm_synchrony_transect) = c("stratum", "loreau_synchrony") # rename cols for merge later
 
@@
             shannon_stability = 1/CV(shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
-            Cover_mean = mean(Cover),
-            Cover_se = sd(Cover)/sqrt(n()),
-            Cover_cv = CV(Cover),
-            Cover_stability = 1/CV(Cover)) # as above
+            cover_mean = mean(cover),
+            cover_se = sd(cover)/sqrt(n()),
+            cover_cv = CV(cover),
+            cover_stability = 1/CV(cover)) # as above
 
@@
 diversity_stability <- merge(diversity_stability, transect_level_portfolio, by = "stratum") # add in portfolio effects
-pairs(~ Cover_mean + Cover_cv + Cover_stability + 
+pairs(~ cover_mean + cover_cv + cover_stability + 
         shannon_mean + shannon_cv + shannon_stability + 
-        Loreau_synchrony + Portfolio_effect, data = diversity_stability)
+        loreau_synchrony + portfolio_effect, data = diversity_stability)
 ```
@@
   ggplot(data = diversity_stability) +
-  geom_point(aes(x = shannon_mean, y = Cover_stability, colour = site, shape = habitat), size = 8) + 
+  geom_point(aes(x = shannon_mean, y = cover_stability, colour = site, shape = habitat), size = 8) + 
    scale_colour_manual(values = site_colours) +
@@
   ggplot(data = diversity_stability) +
-  geom_point(aes(x = shannon_mean, y = Cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
+  geom_point(aes(x = shannon_mean, y = cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
    scale_colour_manual(values = site_colours) +
@@
   ggplot(data = diversity_stability) +
-  geom_point(aes(x = Loreau_synchrony, y = Cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
+  geom_point(aes(x = loreau_synchrony, y = cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
    scale_colour_manual(values = site_colours) +
@@
 #   ggplot(data = diversity_stability) +
-#   geom_point(aes(x = Gross_synchrony, y = Cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
+#   geom_point(aes(x = Gross_synchrony, y = cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
 #    scale_colour_manual(values = site_colours) +
@@
   ggplot(data = diversity_stability) +
-  geom_point(aes(x = Portfolio_effect, y = Cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
+  geom_point(aes(x = portfolio_effect, y = cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
    scale_colour_manual(values = site_colours) +
@@
   ggplot(data = diversity_stability) +
-  geom_point(aes(x = shannon_mean, y = Loreau_synchrony, colour = site, shape = habitat), size = 8, stroke = 2) + 
+  geom_point(aes(x = shannon_mean, y = loreau_synchrony, colour = site, shape = habitat), size = 8, stroke = 2) + 
    scale_colour_manual(values = site_colours) +
@@
   ggplot(data = diversity_stability) +
-  geom_point(aes(x = shannon_mean, y = Portfolio_effect, colour = site, shape = habitat), size = 8, stroke = 2) + 
+  geom_point(aes(x = shannon_mean, y = portfolio_effect, colour = site, shape = habitat), size = 8, stroke = 2) + 
    scale_colour_manual(values = site_colours) +
@@
 ```{r model selection}
-diversity_mod_universal <- glmmTMB(Cover_stability ~ Loreau_synchrony + shannon_mean*site*habitat, family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
+diversity_mod_universal <- glmmTMB(cover_stability ~ loreau_synchrony + shannon_mean*site*habitat, family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
 diversity_dredge <- dredge(diversity_mod_universal)
@@
 # top model is as follows (by a decent margin):
-diversity_top <- glmmTMB(Cover_stability ~ Loreau_synchrony + shannon_mean*site + site*habitat, family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
+diversity_top <- glmmTMB(cover_stability ~ loreau_synchrony + shannon_mean*site + site*habitat, family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
 summary(diversity_top)
@@
   ggplot(data = alpha_diversity) +
-  geom_point(aes(x = Cover, y = shannon, colour = site, shape = habitat), size = 8, stroke = 2) + 
+  geom_point(aes(x = cover, y = shannon, colour = site, shape = habitat), size = 8, stroke = 2) + 
    scale_colour_manual(values = site_colours) +
@@
 
-quadrat_portfolio_columns <- c("location", "Portfolio_effect") # names of columns for dataframe
+quadrat_portfolio_columns <- c("location", "portfolio_effect") # names of columns for dataframe
 quadrat_level_portfolio <- 
@@
 
-# sum(is.na(quadrat_level_portfolio$Portfolio_effect)) # 4 quadrats causing trouble
-
-hist(quadrat_level_portfolio$Portfolio_effect) # some crazy outliers due to low richness
-
-alpha_diversity_quad$Cover <- rowSums(data[ , -which(names(data) %in% meta_cols)])/100
+# sum(is.na(quadrat_level_portfolio$portfolio_effect)) # 4 quadrats causing trouble
+
+hist(quadrat_level_portfolio$portfolio_effect) # some crazy outliers due to low richness
+
+alpha_diversity_quad$cover <- rowSums(data[ , -which(names(data) %in% meta_cols)])/100
 alpha_diversity_quad_sub <- alpha_diversity_quad[!alpha_diversity_quad$location %in% problem_children, ]
 
-# alpha_diversity_quad_sub$Cover <- rowSums(data_sub[ , -which(names(data_sub) %in% meta_cols)])/100
+# alpha_diversity_quad_sub$cover <- rowSums(data_sub[ , -which(names(data_sub) %in% meta_cols)])/100
 alpha_diversity_quad_sub$stratum <- 
@@
                                    replicate.var = "location") # Loreau and Mazancourt synchrony
-colnames(lm_synchrony_quadrat) = c("location", "Loreau_synchrony") # rename cols for merge later
+colnames(lm_synchrony_quadrat) = c("location", "loreau_synchrony") # rename cols for merge later
 
@@
                                    replicate.var = "location")
-colnames(community_stability_quadrat) = c("location", "Cover_stability")
+colnames(community_stability_quadrat) = c("location", "cover_stability")
 
 diversity_stability_quadrat <- alpha_diversity_quad_sub %>%
-  dplyr::group_by(location, site, habitat, site_habitat, transect, stratum) %>%
+  dplyr::group_by(location, site, habitat, island_side, site_habitat, transect, stratum) %>%
   dplyr::summarise(richness_mean = mean(richness), 
@@
             shannon_stability = 1/CV(shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
-            Cover_mean = mean(Cover),
-            Cover_se = sd(Cover)/sqrt(n()),
-            Cover_cv = CV(Cover))
+            cover_mean = mean(cover),
+            cover_se = sd(cover)/sqrt(n()),
+            cover_cv = CV(cover))
 
@@
 diversity_stability_quadrat <- merge(diversity_stability_quadrat, community_stability_quadrat, by = "location")
-pairs(~ Cover_mean + Cover_cv + Cover_stability + 
+pairs(~ cover_mean + cover_cv + cover_stability + 
         shannon_mean + shannon_cv + shannon_stability + 
-        Loreau_synchrony +  Portfolio_effect + VR, data = diversity_stability_quadrat)
+        loreau_synchrony +  portfolio_effect + VR, data = diversity_stability_quadrat)
 
@@
   ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Cover_mean, y = Cover_cv, colour = site, shape = habitat), size = 4) + 
+  geom_point(aes(x = cover_mean, y = cover_cv, colour = site, shape = habitat), size = 4) + 
    scale_colour_manual(values = site_colours) +
@@
   ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Cover_mean, y = shannon_mean, colour = site, shape = habitat), size = 4) + 
+  geom_point(aes(x = cover_mean, y = shannon_mean, colour = site, shape = habitat), size = 4) + 
    scale_colour_manual(values = site_colours) +
@@
   ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = shannon_mean, y = Cover_cv, colour = site, shape = habitat), size = 4) + 
+  geom_point(aes(x = shannon_mean, y = cover_cv, colour = site, shape = habitat), size = 4) + 
    scale_colour_manual(values = site_colours) +
@@
   ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = shannon_mean, y = Cover_cv, colour = site, shape = habitat), size = 4, stroke = 1) + 
+  geom_point(aes(x = shannon_mean, y = cover_cv, colour = site, shape = habitat), size = 4, stroke = 1) + 
    scale_colour_manual(values = site_colours) +
@@
   ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Loreau_synchrony, y = Cover_cv, colour = site, shape = habitat), size = 4, stroke = 1) + 
+  geom_point(aes(x = loreau_synchrony, y = cover_cv, colour = site, shape = habitat), size = 4, stroke = 1) + 
    scale_colour_manual(values = site_colours) +
@@
 #   ggplot(data = diversity_stability_quadrat) +
-#   geom_point(aes(x = Gross_synchrony, y = Cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
+#   geom_point(aes(x = Gross_synchrony, y = cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
 #    scale_colour_manual(values = site_colours) +
@@
   ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = Portfolio_effect, y = Cover_cv, colour = site, shape = habitat), size = 4, stroke = 2) + 
+  geom_point(aes(x = portfolio_effect, y = cover_cv, colour = site, shape = habitat), size = 4, stroke = 2) + 
    scale_colour_manual(values = site_colours) +
@@
   ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = shannon_mean, y = Loreau_synchrony, colour = site, shape = habitat), size = 4, stroke = 1) + 
+  geom_point(aes(x = shannon_mean, y = loreau_synchrony, colour = site, shape = habitat), size = 4, stroke = 1) + 
    scale_colour_manual(values = site_colours) +
@@
   ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = shannon_mean, y = Portfolio_effect, colour = site, shape = habitat), size = 4, stroke = 2) + 
+  geom_point(aes(x = shannon_mean, y = portfolio_effect, colour = site, shape = habitat), size = 4, stroke = 2) + 
    scale_colour_manual(values = site_colours) +
@@
 ```{r diversity models quad level}
-diversity_mod_universal <- glmmTMB(Cover_cv ~shannon_mean*site + shannon_mean*habitat + site*habitat + (1|transect/location), family = Gamma(link = "log"), data = diversity_stability_quadrat, na.action = na.pass)
+diversity_mod_universal <- glmmTMB(cover_cv ~shannon_mean*site + shannon_mean*habitat + site*habitat + (1|transect/location), family = Gamma(link = "log"), data = diversity_stability_quadrat, na.action = na.pass)
 diversity_dredge <- dredge(diversity_mod_universal)
@@
 
-synchrony_mod_universal <- glmmTMB(Cover_cv ~Loreau_synchrony*site + Loreau_synchrony*habitat + site*habitat + (1|transect/location), family = Gamma(link = "log"), data = diversity_stability_quadrat, na.action = na.pass)
+synchrony_mod_universal <- glmmTMB(cover_cv ~loreau_synchrony*site + loreau_synchrony*habitat + site*habitat + (1|transect/location), family = Gamma(link = "log"), data = diversity_stability_quadrat, na.action = na.pass)
 synchrony_dredge <- dredge(synchrony_mod_universal)
@@
 
-diversity_stability_quadrat$Loreau_synchrony_trans <- sv_trans(diversity_stability_quadrat$Loreau_synchrony, diversity_stability_quadrat) # for modelling on beta
-synchrony_diversity_universal <- glmmTMB(Loreau_synchrony_trans ~shannon_mean*site + shannon_mean*habitat + site*habitat + (1|transect/location), family = beta_family(), data = diversity_stability_quadrat, na.action = na.pass)
+diversity_stability_quadrat$loreau_synchrony_trans <- sv_trans(diversity_stability_quadrat$loreau_synchrony, diversity_stability_quadrat) # for modelling on beta
+synchrony_diversity_universal <- glmmTMB(loreau_synchrony_trans ~shannon_mean*site + shannon_mean*habitat + site*habitat + (1|transect/location), family = beta_family(), data = diversity_stability_quadrat, na.action = na.pass)
 model.sel(dredge(synchrony_diversity_universal)) # best model as specified
@@
 r.squaredGLMM(synchrony_diversity_universal) # Model's distribution-specific variance is negative. Results are not reliable.
+
+```
+
+```{r new csvs for joint analyses}
+# write_csv(alpha_diversity_quad, here("data", "algae_diversity_quadrat.csv"))
+# write_csv(diversity_stability_quadrat, here("data", "algae_diversity_stability_metrics_quadrat.csv"))
+
+diversity_stability_quadrat <- alpha_diversity_quad_sub %>%
+  dplyr::group_by(location, site, habitat, site_habitat, transect, stratum) %>%
+  dplyr::summarise(richness_mean = mean(richness), 
+            richness_se = sd(richness)/sqrt(n()),
+            richness_cv = CV(richness),
+            shannon_mean = mean(shannon), 
+            shannon_se = sd(shannon)/sqrt(n()),
+            shannon_cv = CV(shannon),
+            shannon_stability = 1/CV(shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
+            cover_mean = mean(cover),
+            cover_se = sd(cover)/sqrt(n()),
+            cover_cv = CV(cover)) 
+# create csv for site level
+new_csv_meta <- c("year", "island_side", "site", "habitat", "site_habitat") # relevant metadata
+cols_to_remove <- meta_cols[!meta_cols %in% new_csv_meta] # columns that don't overlap between metadata vectors
+reduced_diversity_stability <- diversity_stability_quadrat[ , -which(names(diversity_stability_quadrat) %in% cols_to_remove)]
+
+## HAD TO RUN TO CITY OFFICE FOR FAMILY SHIT AND COULDN'T FINISH THIS YET - NOAM AUG 08
+# diversity_stability_site <- reduced_diversity_stability %>%
+#   dplyr::group_by(island_side, site, habitat, site_habitat) %>%
+#   dplyr::summarise(richness_mean = mean(richness_mean), 
+#             richness_se = sd(richness_mean)/sqrt(n()),
+#             richness_cv = mean(richness_cv),
+#             shannon_mean = mean(shannon), 
+#             shannon_se = sd(shannon)/sqrt(n()),
+#             shannon_cv = CV(shannon),
+#             shannon_stability = 1/CV(shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
+#             cover_mean = mean(cover),
+#             cover_se = sd(cover)/sqrt(n()),
+#             cover_cv = CV(cover),
+#             loreau_mean = mean(loreau_synchrony),
+#             loreau_se = se(loreau_synchrony),
+#             portfolio_mean = mean(portfolio_effect),
+#             portfolio_se = se(portfolio_effect))
+#
 

### 2024-08-10 comparing_datasets.Rmd
Markdown comparing datasets generated by J and N independently

---
title: "comparing noam and j datasets"
author: "Noam Altman-Kurosaki"
date: "2024-08-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clean up
rm(list=ls())

library(here) 
library(tidyverse)

```

This is a document for comparing the datasets N and J prepared

```{r quad level}
meta_cols <- c("year", "site", "island_side", "habitat")

j_quad <- read_csv(here("data", "j_algal_metrics_quadrat.csv"))
n_quad <- read_csv(here("data", "algae_diversity_quadrat.csv"))

quad_comp <- merge(j_quad, n_quad, by = meta_cols)

# cover - should be effectively 0 differnces
sum(quad_comp$cover - quad_comp$algal_perc_cov/100) # 7.297496e-13 --> some rounding issues, but fine

# richness
sum(quad_comp$richness - quad_comp$algal_richness) # 0

# shannon
sum(quad_comp$shannon - quad_comp$algal_Shann) # -2.218337e-12 --> rounding issues
```

```{r site level}

j_site <- read_csv(here("data", "j_algal_metrics_site.csv"))
n_site <- read_csv(here("data", "algae_diversity_site.csv"))

site_comp <- merge(j_site, n_site, by = meta_cols)

# cover - should be effectively 0 differnces
sum(site_comp$algae_cover_mean - site_comp$mean_alg_perc_cov/100) # 1.11e-15 --> some rounding issues, but fine

# richness
sum(site_comp$mean_alg_richness - site_comp$algae_richness_mean) # sameish here

# shannon
sum(site_comp$algae_shannon_mean - site_comp$mean_alg_Shann) # 0

```

```{r diversity stability}
j_stab <- read_csv(here("data", "j_algal_metrics_all_time.csv"))
n_stab <- read_csv(here("data", "algae_diversity_stability_metrics_site.csv"))

stab_comp <- merge(j_stab, n_stab, by = c("site", "island_side", "habitat"))

# cover 
sum(stab_comp$algae_cover_mean - stab_comp$mean_alg_cov/100) # -3.191891e-15--> rounding issues

# richness
sum(stab_comp$algae_richness_mean - stab_comp$mean_alg_richness) # 5.107026e-15

# shannon
sum(stab_comp$algae_shannon_mean - stab_comp$mean_Shann) # 0

# CV
sum(stab_comp$algae_cover_cv - stab_comp$alg_cov_cv) # -9.15934e-15
```

### 2024-08-15 julianna_algalMetric_wrangling.Rmd
Cleaned algal metric wrangling to have the same names as the datasets shared for folks to use with their ancillary datasets

@@
 
+### Save dispersions
+
+```{r}
+# save(yrBetaDisperFring, file = here("data/fringingyrdispersion.Rdata"))
+# save(yrBetaDisperBack, file = here("data/backyrdispersion.Rdata"))
+# save(yrBetaDisperOuter10, file = here("data/outer10yrdispersion.Rdata"))
+# save(yrBetaDisperOuter17, file = here("data/outer17yrdispersion.Rdata"))
+```
+
+### Load dispersions
+
+This is for if you're not running betadisper() for the first time (loading old runs)
+
+```{r}
+load(here("data/fringingyrdispersion.Rdata"))
+load(here("data/backyrdispersion.Rdata"))
+load(here("data/outer10yrdispersion.Rdata"))
+load(here("data/outer17yrdispersion.Rdata"))
+```
+
 ## Join together
@@
 
-### Save dispersions
-
-```{r}
-# save(yrBetaDisperFring, file = here("data/fringingyrdispersion.Rdata"))
-# save(yrBetaDisperBack, file = here("data/backyrdispersion.Rdata"))
-# save(yrBetaDisperOuter10, file = here("data/outer10yrdispersion.Rdata"))
-# save(yrBetaDisperOuter17, file = here("data/outer17yrdispersion.Rdata"))
-```
-
-### Load dispersions
-
-```{r}
-load(here("data/fringingyrdispersion.Rdata"))
-load(here("data/backyrdispersion.Rdata"))
-load(here("data/outer10yrdispersion.Rdata"))
-load(here("data/outer17yrdispersion.Rdata"))
-```
-
 # Join all quadrat-level datasets together
@@
   full_join(betaDHabYr) %>% 
-  # follow a cleaner naming convention
-  rename(algal_perc_cov = AlgCov_perc,
-         algal_perc_cov_hard = AlgCov_percHard,
-         algal_richness = Richness,
-         algal_Shann = Shannon) %>% 
+  # follow a cleaner naming convention to match Noam's
+  rename(algae_cover = AlgCov_perc,
+         algae_cover_hard = AlgCov_percHard,
+         algae_richness = Richness,
+         algae_shannon = Shannon) %>% 
   # get just the columns we need
   select(-site_habitat, -hab_year, -island_side_hab_year) -> quadData
+
 ```
@@
   group_by(year, site, habitat, island_side) %>% 
-  summarize(mean_coral_perc_cov = mean(coral),
-            mean_alg_perc_cov = mean(algal_perc_cov),
-            mean_alg_perc_cov_hard = mean(algal_perc_cov_hard),
-            mean_alg_richness = mean(algal_richness),
-            mean_alg_Shann = mean(algal_Shann),
-            mean_dist_centroid_habyear = mean(hab_yr_distcentroid, na.rm = TRUE)
+  summarize(coral_cover_mean = mean(coral),
+            N = n(),
+            coral_cover_se = sd(coral)/sqrt(N),
+            algae_cover_mean = mean(algae_cover),
+            algae_cover_se = sd(algae_cover)/sqrt(N),
+            algae_cover_hard_mean = mean(algae_cover_hard),
+            algae_cover_hard_se = sd(algae_cover_hard)/sqrt(N),
+            algae_richness_mean = mean(algae_richness),
+            algae_richness_se = sd(algae_richness)/sqrt(N),
+            algae_shannon_mean = mean(algae_shannon),
+            algae_shannon_se = sd(algae_shannon)/sqrt(N),
+            centroid_dist_habyear_mean = mean(hab_yr_distcentroid, na.rm = TRUE)
             ) -> siteData
+
 ```
@@
   summarize(
-    mean_coral_perc = mean(coral),
-    mean_alg_richness = mean(algal_richness), 
-    richness_se = sd(algal_richness)/sqrt(n()),
-    richness_cv = CV(algal_richness),
-    mean_Shann = mean(algal_Shann), 
-    Shann_se = sd(algal_Shann)/sqrt(n()),
-    Shann_cv = CV(algal_Shann),
+    coral_cover_mean = mean(coral),
+    algae_richness_mean = mean(algae_richness), 
+    algae_richness_se = sd(algae_richness)/sqrt(n()),
+    algae_richness_cv = CV(algae_richness),
+    algae_shannon_mean = mean(algae_shannon), 
+    algae_shannon_se = sd(algae_shannon)/sqrt(n()),
+    algae_shannon_cv = CV(algae_shannon),
     # for total algal cover
-    mean_alg_cov = mean(algal_perc_cov),
-    alg_cov_se = sd(algal_perc_cov)/sqrt(n()),
-    cv_alg_cov = CV(algal_perc_cov),
+    algae_cover_mean = mean(algae_cover),
+    algae_cover_se = sd(algae_cover)/sqrt(n()),
+    algae_cover_cv = CV(algae_cover),
     # and for % of hard substrate
-    mean_alg_cov_hard = mean(algal_perc_cov_hard),
-    alg_cov_hard_se = sd(algal_perc_cov_hard)/sqrt(n()),
-    cv_alg_cov_hard = CV(algal_perc_cov_hard),
+    algae_cover_hard_mean = mean(algae_cover_hard),
+    algae_cover_hard_se = sd(algae_cover_hard)/sqrt(n()),
+    algae_cover_hard_cv = CV(algae_cover_hard),
     ) -> alltimeCV
@@
 
-```{r save}
-# write_csv(alltimeCV, here(here("data/algal_metrics_all_time.csv")))
+Combine with Noam's synchrony/portfolio estimates:
+
+```{r}
+noamallTime <- read_csv(here("data/algae_diversity_stability_metrics_alltime.csv"))
 ```
 
-- Noam: Add **synchrony** to this dataset + any other relevant metrics
+```{r}
+noamallTime %>% 
+  select(site, habitat, island_side, algae_shannon_stability, algae_loreau_mean, algae_loreau_se, algae_portfolio_mean, algae_portfolio_se) %>% 
+  full_join(alltimeCV) -> alltimeJoin
+```
 
-s
+```{r save}
+write_csv(alltimeJoin, here(here("data/algal_metrics_all_time.csv")))
+```

### 2024-08-15 noam_data_exploration_v3.Rmd

@@
 
+# algal diversity site level
+diversity_site <- alpha_diversity_quad %>%
+  dplyr::group_by(year, island_side, site, habitat, site_habitat) %>%
+  dplyr::summarise(algae_cover_mean = mean(cover),
+                   algae_cover_se = sd(cover)/sqrt(n()),
+                   algae_richness_mean = mean(richness),
+                   algae_shannon_mean = mean(shannon), 
+                   algae_shannon_se = sd(shannon)/sqrt(n())
+  )
+# write_csv(diversity_site, here("data", "algae_diversity_site.csv"))
+
 diversity_stability_quadrat <- alpha_diversity_quad_sub %>%
@@
 
-## HAD TO RUN TO CITY OFFICE FOR FAMILY SHIT AND COULDN'T FINISH THIS YET - NOAM AUG 08
-# diversity_stability_site <- reduced_diversity_stability %>%
-#   dplyr::group_by(island_side, site, habitat, site_habitat) %>%
-#   dplyr::summarise(richness_mean = mean(richness_mean), 
-#             richness_se = sd(richness_mean)/sqrt(n()),
-#             richness_cv = mean(richness_cv),
-#             shannon_mean = mean(shannon), 
-#             shannon_se = sd(shannon)/sqrt(n()),
-#             shannon_cv = CV(shannon),
-#             shannon_stability = 1/CV(shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
-#             cover_mean = mean(cover),
-#             cover_se = sd(cover)/sqrt(n()),
-#             cover_cv = CV(cover),
-#             loreau_mean = mean(loreau_synchrony),
-#             loreau_se = se(loreau_synchrony),
-#             portfolio_mean = mean(portfolio_effect),
-#             portfolio_se = se(portfolio_effect))
-#
+# averaging from the original here to avoid taking average of average, cv of cv, etc.
+diversity_cv_site <- alpha_diversity_quad %>%
+  dplyr::group_by(site, habitat, island_side, site_habitat) %>%
+  dplyr::summarise(algae_richness_mean = mean(richness), 
+            algae_richness_se = sd(richness)/sqrt(n()),
+            algae_richness_cv = CV(richness),
+            algae_shannon_mean = mean(shannon), 
+            algae_shannon_se = sd(shannon)/sqrt(n()),
+            algae_shannon_cv = CV(shannon),
+            algae_shannon_stability = 1/CV(shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
+            algae_cover_mean = mean(cover),
+            algae_cover_se = sd(cover)/sqrt(n()),
+            algae_cover_cv = CV(cover))
+
+# portfolio and synchrony need to be calculated at quad level, so taking the mean of the quad level for these
+synchrony_portfolio_site <- diversity_stability_quadrat %>%
+  dplyr::group_by(site, habitat, island_side, site_habitat) %>%
+  dplyr::summarise(
+            algae_loreau_mean = mean(loreau_synchrony),
+            algae_loreau_se = sd(loreau_synchrony)/sqrt(n()),
+            algae_portfolio_mean = mean(portfolio_effect),
+            algae_portfolio_se = sd(portfolio_effect)/sqrt(n()))
+
+# merge together
+diversity_stability_site <- merge(diversity_cv_site, synchrony_portfolio_site, by = c("site", "habitat", "island_side", "site_habitat"))
+# write_csv(diversity_stability_site, here("data", "algae_diversity_stability_metrics_site.csv"))
 

### 2024-08-15 julianna_coral_macroalg_wrangling.Rmd
Start a cleaned script with just the coral/macroalgae wrangling

--- 2024-07-30 julianna_algalMetric_wrangling.Rmd
@@
 ---
-title: "Julianna's algal metrics wrangling code"
+title: "Wrangling coral and macroalgae data to be joinable"
 output: html_document
-date: "2024-07-30"
-editor_options: 
-  chunk_output_type: console
+date: "2024-08-15"
 ---
@@
 
-This script generates the CSV of algal metrics to regress other variables against
-
-# Bring in the data
-
-```{r}
-# algal cover 
-algae <- read_csv(here("data", "algaeonly_cleaned_colnames.csv")) %>% 
-  mutate(island_side = case_when(site == "lter_1" ~ "north",
-                                 site == "lter_2" ~ "north",
-                                 site == "lter_3" ~ "east",
-                                 site == "lter_4" ~ "east",
-                                 site == "lter_5" ~ "west",
-                                 site == "lter_6" ~ "west",
-                                 TRUE ~ "UdINGLEbUTT"))
-
-# sand cover
-sand <- read_csv(here("data", "allbenthicdata_cleaned_colnames.csv")) %>% 
-  select(year, location, site, habitat, site_habitat, transect, quadrat, sand, coral)
-```
-
-# Calculate algal alpha diversity
-
-## Make matrices
-
-Make a spp matrix to use with `vegan`
-
-```{r make species matrix}
-# matrix of just species and their relative percent covers in the same order as the algae tibble
-algae %>% 
-  select(-year, -location, -site, -island_side, -habitat, -site_habitat, -transect, -quadrat) -> sppMat
-```
-
-```{r make a metadata matrix in the same order}
-# get ranges to calculate over (this is just for indexing purposes in the next bit of code)
-startAlg <- grep("acanthophora_spicifera", colnames(algae))
-endAlg <- grep("valonia_ventricosa", colnames(algae))
-
-# metadata matrix
-algae %>% 
-  # bring in sand data
-  full_join(sand) %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg]), # percent cover of algae
-         AlgCov_percHard = case_when(
-           sand == 0 ~ AlgCov_perc, # since we can't divide by a zero
-           TRUE ~ AlgCov_perc/sand 
-           )) %>% # percent cover of hard substrate
-   # make a year-habitat column for betadisper
-  mutate(hab_year = paste0(habitat, "_", year),
-         island_side_hab_year = paste0(island_side, "_", hab_year)) %>% 
-  select(year, location, site, habitat, island_side, site_habitat, 
-         hab_year, island_side_hab_year, transect, quadrat, AlgCov_perc, 
-         AlgCov_percHard, sand, coral)  -> sppMeta
-```
-
-## Calculate diversity metrics
-
-```{r get richness and Shannon diversity}
-sppMeta %>% 
-  # these functions use the vegan package
-  mutate(Richness = specnumber(sppMat),
-         Shannon = diversity(sppMat, "shannon")) -> algDiv
-```
-
-# Calculate algal beta diversity
-
-## Distance matrix: habitat-year
-
-We get an error if we run `vegdist()` with quadrats that have no algae. We can see there are 672 of these that we need to filter out:
-
-```{r look for zeros}
-sppMat %>% 
-  mutate(AlgCov_perc = rowSums(across(where(is.numeric)))) %>% 
-  filter(AlgCov_perc == 0) %>% dim() # 672 with no algae at all
-```
-
-Now we can exclude quadrats with no algae and make a species matrix again
-
-```{r get spp mat df}
-startAlg <- grep("acanthophora_spicifera", colnames(algae))
-endAlg <- grep("valonia_ventricosa", colnames(algae))
-
-# spp matrix where 0 observations are filtered out
-algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>% 
-  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgCov_perc) -> sppMat0
-```
-
-```{r metadata matix}
-# matching metadata matrix
-algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>%
-  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
-  # make a year-habitat column for betadisper
-  mutate(hab_year = paste0(habitat, "_", year),
-         island_side_hab_year = paste0(island_side, hab_year)) -> sppMeta0
-```
-
-Now we'll calculate this by habitat because we're compute limited
-
-### Calculate fringing
-
-```{r get fringe mat}
-startAlg <- grep("acanthophora_spicifera", colnames(algae))
-endAlg <- grep("valonia_ventricosa", colnames(algae))
-
-# spp matrix
-algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>% 
-  filter(habitat == "fringing") %>% 
-  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgCov_perc) -> sppMat0fring
-
-# metadata matrix
-algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>%
-   filter(habitat == "fringing") %>% 
-  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
-  # make a year-habitat column for betadisper
-  mutate(hab_year = paste0(habitat, "_", year),
-         island_side_hab_year = paste0(island_side, hab_year),
-         island_side_year = paste0(island_side, year)) -> sppMeta0fring
-```
-
-Calculate Bray Curtis distances
-
-```{r calculate distances}
-brayMatFring <- vegdist(sppMat0fring, "bray") # 1 is maximum dissimilarity
-```
-
-Calculate dispersion (this is using Noam's code)
-
-```{r calculate beta dispersion}
-yrBetaDisperFring <- vegan::betadisper(brayMatFring, sppMeta0fring$year) # warning: some squared distances are negative and changed to zero (ask Jamie about this??)
-```
-
-### Backreef
-
-```{r get backreef df}
-startAlg <- grep("acanthophora_spicifera", colnames(algae))
-endAlg <- grep("valonia_ventricosa", colnames(algae))
-
-# spp matrix
-algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>% 
-  filter(habitat == "backreef") %>% 
-  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgCov_perc) -> sppMat0back
-
-# metadata matrix
-algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>%
-   filter(habitat == "backreef") %>% 
-  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
-  # make a year-habitat column for betadisper
-  mutate(hab_year = paste0(habitat, "_", year),
-         island_side_hab_year = paste0(island_side, hab_year),
-         island_side_year = paste0(island_side, year)) -> sppMeta0back
-```
-
-Calculate Bray Curtis distances
-
-```{r calculate distance}
-brayMatBack <- vegdist(sppMat0back, "bray") # 1 is maximum dissimilarity
-```
-
-Calculate beta disperion
-
-```{r calculate beta dispersion}
-yrBetaDisperBack <- vegan::betadisper(brayMatBack, sppMeta0back$year) 
-```
-
-### Outer 10
-
-```{r get outer 10 df}
-startAlg <- grep("acanthophora_spicifera", colnames(algae))
-endAlg <- grep("valonia_ventricosa", colnames(algae))
-
-# spp matrix
-algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>% 
-  filter(habitat == "outer_10") %>% 
-  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgCov_perc) -> sppMat0outer10
-
-# metadata matrix
-algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>%
-   filter(habitat == "outer_10") %>% 
-  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
-  # make a year-habitat column for betadisper
-  mutate(hab_year = paste0(habitat, "_", year),
-         island_side_hab_year = paste0(island_side, hab_year),
-         island_side_year = paste0(island_side, year)) -> sppMeta0outer10
-```
-
-Calculate Bray Curtis distances
-
-```{r calculate distance}
-brayMatOuter10 <- vegdist(sppMat0outer10, "bray") # 1 is maximum dissimilarity
-```
-
-Calculate beta disperion
-
-```{r calculate beta dispersion}
-yrBetaDisperOuter10 <- vegan::betadisper(brayMatOuter10, sppMeta0outer10$year) # error: some squared distances are negative and changed to zero
-```
-
-### Outer 17
-
-```{r get outer 17 df}
-startAlg <- grep("acanthophora_spicifera", colnames(algae))
-endAlg <- grep("valonia_ventricosa", colnames(algae))
-
-# spp matrix
-algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>% 
-  filter(habitat == "outer_17") %>% 
-  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgCov_perc) -> sppMat0outer17
-
-# metadata matrix
-algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>%
-   filter(habitat == "outer_17") %>% 
-  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
-  # make a year-habitat column for betadisper
-  mutate(hab_year = paste0(habitat, "_", year),
-         island_side_hab_year = paste0(island_side, hab_year),
-         island_side_year = paste0(island_side, year)) -> sppMeta0outer17
-```
-
-Calculate Bray Curtis distances
-
-```{r calculate distance}
-brayMatOuter17 <- vegdist(sppMat0outer17, "bray") # 1 is maximum dissimilarity
-```
-
-Calculate beta disperion
-
-```{r calculate beta dispersion}
-yrBetaDisperOuter17 <- vegan::betadisper(brayMatOuter17, sppMeta0outer17$year) # error: some squared distances are negative and changed to zero
-```
-
-## Join together
-
-Attach back to metadata
-
-```{r join distances with metadata}
-sppMeta0fring %>% 
-  mutate(distance_centriod = yrBetaDisperFring$distances) -> fringeDist
-
-sppMeta0back %>% 
-  mutate(distance_centriod = yrBetaDisperBack$distances) -> backDist
-
-sppMeta0outer10 %>% 
-  mutate(distance_centriod = yrBetaDisperOuter10$distances) -> outer10Dist
-
-sppMeta0outer17 %>% 
-  mutate(distance_centriod = yrBetaDisperOuter17$distances) -> outer17Dist
-```
-
-```{r join together}
-# make sure this looks good
-colnames(fringeDist) == colnames(backDist) 
-colnames(backDist) == colnames(outer10Dist) 
-colnames(outer10Dist) == colnames(outer17Dist)
-
-# join together and tidy up
-betaDHabYr <- rbind(fringeDist, backDist, outer10Dist, outer17Dist) %>% 
-  mutate(hab_yr_distcentroid = distance_centriod) %>% 
-  select(year, location, site, habitat, transect, quadrat, hab_yr_distcentroid)
-```
-
-### Save dispersions
-
-```{r}
-# save(yrBetaDisperFring, file = here("data/fringingyrdispersion.Rdata"))
-# save(yrBetaDisperBack, file = here("data/backyrdispersion.Rdata"))
-# save(yrBetaDisperOuter10, file = here("data/outer10yrdispersion.Rdata"))
-# save(yrBetaDisperOuter17, file = here("data/outer17yrdispersion.Rdata"))
-```
-
-### Load dispersions
-
-```{r}
-load(here("data/fringingyrdispersion.Rdata"))
-load(here("data/backyrdispersion.Rdata"))
-load(here("data/outer10yrdispersion.Rdata"))
-load(here("data/outer17yrdispersion.Rdata"))
-```
-
-# Join all quadrat-level datasets together
-
-```{r join}
-algDiv %>% 
-  full_join(betaDHabYr) %>% 
-  # follow a cleaner naming convention
-  rename(algal_perc_cov = AlgCov_perc,
-         algal_perc_cov_hard = AlgCov_percHard,
-         algal_richness = Richness,
-         algal_Shann = Shannon) %>% 
-  # get just the columns we need
-  select(-site_habitat, -hab_year, -island_side_hab_year) -> quadData
-```
-
-```{r save}
-# write_csv(quadData, here("algal_metrics_quadrat.csv"))
-```
-
-# Site/habitat/year-level mean
-
-We'll also average these data at the site/habitat/year-level to be more compatible with other data types.
-This will be done at the *quadrat-level*, meaning we won't take the transect-level summary first, since they aren't meant to be factors in a statistical design
-
-```{r get averages}
-quadData %>% 
-  group_by(year, site, habitat, island_side) %>% 
-  summarize(mean_coral_perc_cov = mean(coral),
-            mean_alg_perc_cov = mean(algal_perc_cov),
-            mean_alg_perc_cov_hard = mean(algal_perc_cov_hard),
-            mean_alg_richness = mean(algal_richness),
-            mean_alg_Shann = mean(algal_Shann),
-            mean_dist_centroid_habyear = mean(hab_yr_distcentroid, na.rm = TRUE)
-            ) -> siteData
-```
-
-```{r save dataframe}
-# write_csv(siteData, here("algal_metrics_site.csv"))
-```
-
-# All time metrics
-
-## Calculate the coefficient of variation
-
-This is basically just copying/pasting Noam's code
-
-```{r cv function}
-# coefficient of variation
-CV <- function(x){
-  return(sd(x,na.rm=T)/mean(x,na.rm=T))
-}
-```
-
-Apply it when we summarize
-
-```{r calculate CV and summarize}
-quadData %>% 
-  # don't include year this time
-  group_by(site, habitat, island_side) %>%
-  summarize(
-    mean_coral_perc = mean(coral),
-    mean_alg_richness = mean(algal_richness), 
-    richness_se = sd(algal_richness)/sqrt(n()),
-    richness_cv = CV(algal_richness),
-    mean_Shann = mean(algal_Shann), 
-    Shann_se = sd(algal_Shann)/sqrt(n()),
-    Shann_cv = CV(algal_Shann),
-    # for total algal cover
-    mean_alg_cov = mean(algal_perc_cov),
-    alg_cov_se = sd(algal_perc_cov)/sqrt(n()),
-    cv_alg_cov = CV(algal_perc_cov),
-    # and for % of hard substrate
-    mean_alg_cov_hard = mean(algal_perc_cov_hard),
-    alg_cov_hard_se = sd(algal_perc_cov_hard)/sqrt(n()),
-    cv_alg_cov_hard = CV(algal_perc_cov_hard),
-    ) -> alltimeCV
-```
-
-```{r save}
-# write_csv(alltimeCV, here(here("algal_metrics_all_time.csv")))
-```
-
-- Noam: Add **synchrony** to this dataset + any other relevant metrics
-
-s
+This script summarizes the coral and macroalgae subsets to be joinable with the full algae dataset

### 2024-08-15 julianna_algalMetric_wrangling.Rmd
Updated wrangling code to fix sand bug for algal cover of hard substrate

@@
   full_join(sand) %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg]), # percent cover of algae
-         AlgCov_percHard = case_when(
-           sand == 0 ~ AlgCov_perc, # since we can't divide by a zero
-           TRUE ~ AlgCov_perc/sand 
+  mutate(algae_cover = rowSums(algae[startAlg:endAlg]), # percent cover of algae
+         algae_cover_hard = case_when(
+           sand == 100 ~ NA, # since we can't divide by a zero
+           TRUE ~ (algae_cover/(100-sand))*100 
            )) %>% # percent cover of hard substrate
@@
   select(year, location, site, habitat, island_side, site_habitat, 
-         hab_year, island_side_hab_year, transect, quadrat, AlgCov_perc, 
-         AlgCov_percHard, sand, coral)  -> sppMeta
+         hab_year, island_side_hab_year, transect, quadrat, algae_cover, 
+         algae_cover_hard, sand, coral)  -> sppMeta
 ```
@@
 sppMat %>% 
-  mutate(AlgCov_perc = rowSums(across(where(is.numeric)))) %>% 
-  filter(AlgCov_perc == 0) %>% dim() # 672 with no algae at all
+  mutate(algae_cover = rowSums(across(where(is.numeric)))) %>% 
+  filter(algae_cover == 0) %>% dim() # 672 with no algae at all
 ```
@@
 algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>% 
-  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgCov_perc) -> sppMat0
+  mutate(algae_cover = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(algae_cover != 0) %>% 
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -algae_cover) -> sppMat0
 ```
@@
 algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>%
+  mutate(algae_cover = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(algae_cover != 0) %>%
   select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
@@
 algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>% 
+  mutate(algae_cover = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(algae_cover != 0) %>% 
   filter(habitat == "fringing") %>% 
-  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgCov_perc) -> sppMat0fring
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -algae_cover) -> sppMat0fring
 
@@
 algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>%
+  mutate(algae_cover = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(algae_cover != 0) %>%
    filter(habitat == "fringing") %>% 
@@
 algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>% 
+  mutate(algae_cover = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(algae_cover != 0) %>% 
   filter(habitat == "backreef") %>% 
-  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgCov_perc) -> sppMat0back
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -algae_cover) -> sppMat0back
 
@@
 algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>%
+  mutate(algae_cover = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(algae_cover != 0) %>%
    filter(habitat == "backreef") %>% 
@@
 algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>% 
+  mutate(algae_cover = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(algae_cover != 0) %>% 
   filter(habitat == "outer_10") %>% 
-  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgCov_perc) -> sppMat0outer10
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -algae_cover) -> sppMat0outer10
 
@@
 algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>%
+  mutate(algae_cover = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(algae_cover != 0) %>%
    filter(habitat == "outer_10") %>% 
@@
 algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>% 
+  mutate(algae_cover = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(algae_cover != 0) %>% 
   filter(habitat == "outer_17") %>% 
-  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgCov_perc) -> sppMat0outer17
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -algae_cover) -> sppMat0outer17
 
@@
 algae %>% 
-  mutate(AlgCov_perc = rowSums(algae[startAlg:endAlg])) %>% 
-  filter(AlgCov_perc != 0) %>%
+  mutate(algae_cover = rowSums(algae[startAlg:endAlg])) %>% 
+  filter(algae_cover != 0) %>%
    filter(habitat == "outer_17") %>% 
@@
   # follow a cleaner naming convention to match Noam's
-  rename(algae_cover = AlgCov_perc,
-         algae_cover_hard = AlgCov_percHard,
-         algae_richness = Richness,
+  rename(algae_richness = Richness,
          algae_shannon = Shannon) %>% 
@@
             algae_cover_se = sd(algae_cover)/sqrt(N),
-            algae_cover_hard_mean = mean(algae_cover_hard),
-            algae_cover_hard_se = sd(algae_cover_hard)/sqrt(N),
+            algae_cover_hard_mean = mean(algae_cover_hard, na.rm = TRUE),
+            algae_cover_hard_se = sd(algae_cover_hard, na.rm = TRUE)/sqrt(N),
             algae_richness_mean = mean(algae_richness),
@@
 
-Apply it when we summarize
-
 ```{r calculate CV and summarize}
@@
     # and for % of hard substrate
-    algae_cover_hard_mean = mean(algae_cover_hard),
-    algae_cover_hard_se = sd(algae_cover_hard)/sqrt(n()),
+    algae_cover_hard_mean = mean(algae_cover_hard, na.rm = TRUE),
+    algae_cover_hard_se = sd(algae_cover_hard, na.rm = TRUE)/sqrt(n()),
     algae_cover_hard_cv = CV(algae_cover_hard),

### 2024-08-15 julianna_coral_macroalg_wrangling.Rmd
Updated wrangling code to fix sand bug for algal cover of hard substrate

@@
 This script summarizes the coral and macroalgae subsets to be joinable with the full algae dataset
+
+# Bring in the data
+
+```{r}
+# algal cover 
+macroalgae <- read_csv(here("data", "algaeonly_cleaned_colnames.csv")) %>% 
+  mutate(island_side = case_when(site == "lter_1" ~ "north",
+                                 site == "lter_2" ~ "north",
+                                 site == "lter_3" ~ "east",
+                                 site == "lter_4" ~ "east",
+                                 site == "lter_5" ~ "west",
+                                 site == "lter_6" ~ "west",
+                                 TRUE ~ "UdINGLEbUTT")) %>% 
+  select(-algal_turf, -crustose_corallines, -cyanophyta, -damselfish_turf, -lithophyllum_kotschyanum,
+         -neogoniolithon_brassica_florida, -sporolithon_sp) # got this from the cleaning code--should Peysonellia be included?
+
+# sand cover and coral cover
+sand_coral <- read_csv(here("data", "allbenthicdata_cleaned_colnames.csv")) %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, sand, coral)
+```
+
+# Calculate macroalgal alpha diversity
+
+## Make matrices
+
+Make a spp matrix to use with `vegan`
+
+```{r make species matrix}
+# matrix of just species and their relative percent covers in the same order as the algae tibble
+macroalgae %>% 
+  select(-year, -location, -site, -island_side, -habitat, -site_habitat, -transect, -quadrat) -> sppMat
+```

### 2024-08-15 julianna_data_exploration.Rmd
Updated wrangling code to fix sand bug for algal cover of hard substrate

@@
 
+require(librarian)
+
 # load packages
-require(here) # relative file paths
-require(tidyverse) # data wrangling
-require(vegan) # community/diversity analyses
-require(lme4) # glmer()
-require(emmeans) # emmeans()
-require(viridis) # color palette
-require(MoMAColors) # another color palette
-require(patchwork) # putting figures together
-require(BiodiversityR)
-require(ggsci) # scientific jounral and sci-fi themed color palette
+librarian::shelf(
+  here, # relative file paths
+  tidyverse,  # data wrangling
+  vegan, # data wrangling
+  lme4, # glmer()
+  emmeans, # emmeans()
+  viridis, # color palette
+  MoMAColors, # another color palette: devtools::install_github("BlakeRMills/MoMAColors")
+  patchwork, # putting figures together
+  BiodiversityR,
+  ggsci # scientific jounral and sci-fi themed color palette
+)
+
 ```

### 2024-08-15 julianna_algalMetric_wrangling.Rmd
Made independent macroalgae and coral wrangling Rmkdown to produce coral/macroalgal summaries

@@
 sand <- read_csv(here("data", "allbenthicdata_cleaned_colnames.csv")) %>% 
-  select(year, location, site, habitat, site_habitat, transect, quadrat, sand, coral)
+  select(year, location, site, habitat, site_habitat, transect, quadrat, sand)
 ```
@@
          hab_year, island_side_hab_year, transect, quadrat, algae_cover, 
-         algae_cover_hard, sand, coral)  -> sppMeta
+         algae_cover_hard)  -> sppMeta
 ```
@@
   group_by(year, site, habitat, island_side) %>% 
-  summarize(coral_cover_mean = mean(coral),
-            N = n(),
-            coral_cover_se = sd(coral)/sqrt(N),
+  summarize(N = n(),
             algae_cover_mean = mean(algae_cover),
@@
   summarize(
-    coral_cover_mean = mean(coral),
     algae_richness_mean = mean(algae_richness), 

### 2024-08-15 julianna_coral_macroalg_wrangling.Rmd
Made independent macroalgae and coral wrangling Rmkdown to produce coral/macroalgal summaries

--- 2024-08-15 julianna_coral_macroalg_wrangling.Rmd
@@
 This script summarizes the coral and macroalgae subsets to be joinable with the full algae dataset
+
+# Bring in the data
+
+```{r}
+# algal cover 
+macroalgae <- read_csv(here("data", "algaeonly_cleaned_colnames.csv")) %>% 
+  mutate(island_side = case_when(site == "lter_1" ~ "north",
+                                 site == "lter_2" ~ "north",
+                                 site == "lter_3" ~ "east",
+                                 site == "lter_4" ~ "east",
+                                 site == "lter_5" ~ "west",
+                                 site == "lter_6" ~ "west",
+                                 TRUE ~ "UdINGLEbUTT")) %>% 
+  select(-algal_turf, -crustose_corallines, -cyanophyta, -damselfish_turf, -lithophyllum_kotschyanum,
+         -neogoniolithon_brassica_florida, -sporolithon_sp) # on Aug 15 Noam and I decided to keep Peysonellia, etc. in
+
+# sand cover and coral cover
+sand_coral <- read_csv(here("data", "allbenthicdata_cleaned_colnames.csv")) %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, sand, coral)
+```
+
+# Calculate macroalgal alpha diversity
+
+## Make matrices
+
+Make a macroalgal spp matrix to use with `vegan`
+
+```{r make species matrix}
+# matrix of just species and their relative percent covers in the same order as the algae tibble
+macroalgae %>% 
+  select(-year, -location, -site, -island_side, -habitat, -site_habitat, -transect, -quadrat) -> sppMat
+```
+
+```{r make a metadata matrix in the same order}
+# get ranges to calculate over (this is just for indexing purposes in the next bit of code)
+startAlg <- grep("acanthophora_spicifera", colnames(macroalgae))
+endAlg <- grep("valonia_ventricosa", colnames(macroalgae))
+
+# metadata matrix
+macroalgae %>% 
+  # bring in sand data
+  full_join(sand_coral) %>% 
+  rename(coral_cover = coral,
+         sand_cover = sand) %>% 
+  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg]), # percent cover of algae
+         macroalgae_cover_hard = case_when(
+           sand_cover == 100 ~ NA, # since we can't divide by a zero
+           TRUE ~ (macroalgae_cover/(100-sand_cover))*100 
+           ),
+         # also for corals
+         coral_cover_hard = case_when(
+           sand_cover == 100 ~ NA, # since we can't divide by a zero
+           TRUE ~ (coral_cover/(100-sand_cover))*100 
+           )
+         ) %>% # percent cover of hard substrate
+   # make a year-habitat column for betadisper
+  mutate(hab_year = paste0(habitat, "_", year),
+         island_side_hab_year = paste0(island_side, "_", hab_year)) %>% 
+  select(year, location, site, habitat, island_side, site_habitat, 
+         hab_year, island_side_hab_year, transect, quadrat, macroalgae_cover, 
+         macroalgae_cover_hard, sand_cover, coral_cover, coral_cover_hard)  -> sppMeta
+```
+
+## Calculate diversity metrics
+
+```{r get richness and Shannon diversity}
+sppMeta %>% 
+  # these functions use the vegan package
+  mutate(macroalgae_richness = specnumber(sppMat),
+         macroalgae_shannon = diversity(sppMat, "shannon")) -> macroAlgDiv
+```
+
+# Calculate algal beta diversity
+
+## Distance matrix: habitat-year
+
+We get an error if we run `vegdist()` with quadrats that have no algae. We can see there are 672 of these that we need to filter out:
+
+```{r look for zeros}
+sppMat %>% 
+  mutate(macroalgae_cover = rowSums(across(where(is.numeric)))) %>% 
+  filter(macroalgae_cover == 0) %>% dim() # 7208 with no algae at all (that's like over 1/3!!)
+```
+
+Now we can exclude quadrats with no algae and make a species matrix again
+
+```{r get spp mat df}
+startAlg <- grep("acanthophora_spicifera", colnames(macroalgae))
+endAlg <- grep("valonia_ventricosa", colnames(macroalgae))
+
+# spp matrix where 0 observations are filtered out
+macroalgae %>% 
+  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
+  filter(macroalgae_cover != 0) %>% 
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -macroalgae_cover) -> sppMat0
+```
+
+```{r metadata matix}
+# matching metadata matrix
+macroalgae %>% 
+  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
+  filter(macroalgae_cover != 0) %>%
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
+  # make a year-habitat column for betadisper
+  mutate(hab_year = paste0(habitat, "_", year),
+         island_side_hab_year = paste0(island_side, hab_year)) -> sppMeta0
+```
+
+Now we'll calculate this by habitat because we're compute limited
+
+### Calculate fringing
+
+```{r get fringe mat}
+startAlg <- grep("acanthophora_spicifera", colnames(macroalgae))
+endAlg <- grep("valonia_ventricosa", colnames(macroalgae))
+
+# spp matrix
+macroalgae %>% 
+  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
+  filter(macroalgae_cover != 0) %>% 
+  filter(habitat == "fringing") %>% 
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -macroalgae_cover) -> sppMat0fring
+
+# metadata matrix
+macroalgae %>% 
+  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
+  filter(macroalgae_cover != 0) %>%
+   filter(habitat == "fringing") %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
+  # make a year-habitat column for betadisper
+  mutate(hab_year = paste0(habitat, "_", year),
+         island_side_hab_year = paste0(island_side, hab_year),
+         island_side_year = paste0(island_side, year)) -> sppMeta0fring
+```
+
+Calculate Bray Curtis distances
+
+```{r calculate distances}
+brayMatFringMacro <- vegdist(sppMat0fring, "bray") # 1 is maximum dissimilarity
+```
+
+Calculate dispersion (this is using Noam's code)
+
+```{r calculate beta dispersion}
+yrBetaDisperFringMacro <- vegan::betadisper(brayMatFringMacro, sppMeta0fring$year) # warning: some squared distances are negative and changed to zero (ask Jamie about this??)
+```
+
+### Backreef
+
+```{r get backreef df}
+startAlg <- grep("acanthophora_spicifera", colnames(macroalgae))
+endAlg <- grep("valonia_ventricosa", colnames(macroalgae))
+
+# spp matrix
+macroalgae %>% 
+  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
+  filter(macroalgae_cover != 0) %>% 
+  filter(habitat == "backreef") %>% 
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -macroalgae_cover) -> sppMat0back
+
+# metadata matrix
+macroalgae %>% 
+  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
+  filter(macroalgae_cover != 0) %>%
+   filter(habitat == "backreef") %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
+  # make a year-habitat column for betadisper
+  mutate(hab_year = paste0(habitat, "_", year),
+         island_side_hab_year = paste0(island_side, hab_year),
+         island_side_year = paste0(island_side, year)) -> sppMeta0back
+```
+
+Calculate Bray Curtis distances
+
+```{r calculate distance}
+brayMatBackMacro <- vegdist(sppMat0back, "bray") # 1 is maximum dissimilarity
+```
+
+Calculate beta disperion
+
+```{r calculate beta dispersion}
+yrBetaDisperBackMacro <- vegan::betadisper(brayMatBackMacro, sppMeta0back$year) 
+```
+
+### Outer 10
+
+```{r get outer 10 df}
+startAlg <- grep("acanthophora_spicifera", colnames(macroalgae))
+endAlg <- grep("valonia_ventricosa", colnames(macroalgae))
+
+# spp matrix
+macroalgae %>% 
+  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
+  filter(macroalgae_cover != 0) %>% 
+  filter(habitat == "outer_10") %>% 
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -macroalgae_cover) -> sppMat0outer10
+
+# metadata matrix
+macroalgae %>% 
+  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
+  filter(macroalgae_cover != 0) %>%
+   filter(habitat == "outer_10") %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
+  # make a year-habitat column for betadisper
+  mutate(hab_year = paste0(habitat, "_", year),
+         island_side_hab_year = paste0(island_side, hab_year),
+         island_side_year = paste0(island_side, year)) -> sppMeta0outer10
+```
+
+Calculate Bray Curtis distances
+
+```{r calculate distance}
+brayMatOuter10Macro <- vegdist(sppMat0outer10, "bray") # 1 is maximum dissimilarity
+```
+
+Calculate beta disperion
+
+```{r calculate beta dispersion}
+yrBetaDisperOuter10Macro <- vegan::betadisper(brayMatOuter10Macro, sppMeta0outer10$year) # error: some squared distances are negative and changed to zero
+```
+
+### Outer 17
+
+```{r get outer 17 df}
+startAlg <- grep("acanthophora_spicifera", colnames(macroalgae))
+endAlg <- grep("valonia_ventricosa", colnames(macroalgae))
+
+# spp matrix
+macroalgae %>% 
+  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
+  filter(macroalgae_cover != 0) %>% 
+  filter(habitat == "outer_17") %>% 
+  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -macroalgae_cover) -> sppMat0outer17
+
+# metadata matrix
+macroalgae %>% 
+  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
+  filter(macroalgae_cover != 0) %>%
+   filter(habitat == "outer_17") %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
+  # make a year-habitat column for betadisper
+  mutate(hab_year = paste0(habitat, "_", year),
+         island_side_hab_year = paste0(island_side, hab_year),
+         island_side_year = paste0(island_side, year)) -> sppMeta0outer17
+```
+
+Calculate Bray Curtis distances
+
+```{r calculate distance}
+brayMatOuter17Macro <- vegdist(sppMat0outer17, "bray") # 1 is maximum dissimilarity
+```
+
+Calculate beta disperion
+
+```{r calculate beta dispersion}
+yrBetaDisperOuter17Macro <- vegan::betadisper(brayMatOuter17Macro, sppMeta0outer17$year) # error: some squared distances are negative and changed to zero
+```
+
+### Save dispersions
+
+```{r}
+# save(yrBetaDisperFringMacro, file = here("data/fringingyrdispersionMacro.Rdata"))
+# save(yrBetaDisperBackMacro, file = here("data/backyrdispersionMacro.Rdata"))
+# save(yrBetaDisperOuter10Macro, file = here("data/outer10yrdispersionMacro.Rdata"))
+# save(yrBetaDisperOuter17Macro, file = here("data/outer17yrdispersionMacro.Rdata"))
+```
+
+### Load dispersions
+
+This is for if you're not running betadisper() for the first time (loading old runs)
+
+```{r}
+load(here("data/fringingyrdispersionMacro.Rdata"))
+load(here("data/backyrdispersionMacro.Rdata"))
+load(here("data/outer10yrdispersionMacro.Rdata"))
+load(here("data/outer17yrdispersionMacro.Rdata"))
+```
+
+## Join together
+
+Attach back to metadata
+
+```{r join distances with metadata}
+sppMeta0fring %>% 
+  mutate(distance_centriod = yrBetaDisperFringMacro$distances) -> fringeDist
+
+sppMeta0back %>% 
+  mutate(distance_centriod = yrBetaDisperBackMacro$distances) -> backDist
+
+sppMeta0outer10 %>% 
+  mutate(distance_centriod = yrBetaDisperOuter10Macro$distances) -> outer10Dist
+
+sppMeta0outer17 %>% 
+  mutate(distance_centriod = yrBetaDisperOuter17Macro$distances) -> outer17Dist
+```
+
+```{r join together}
+# make sure this looks good
+colnames(fringeDist) == colnames(backDist) 
+colnames(backDist) == colnames(outer10Dist) 
+colnames(outer10Dist) == colnames(outer17Dist)
+
+# join together and tidy up
+betaDHabYr <- rbind(fringeDist, backDist, outer10Dist, outer17Dist) %>% 
+  mutate(hab_yr_distcentroid = distance_centriod) %>% 
+  select(year, location, site, habitat, transect, quadrat, hab_yr_distcentroid)
+```
+
+# Join all quadrat-level datasets together
+
+```{r join}
+macroAlgDiv %>% 
+  full_join(betaDHabYr) %>% 
+  # get just the columns we need
+  select(-site_habitat, -hab_year, -island_side_hab_year) -> quadData
+
+```
+
+```{r save}
+# write_csv(quadData, here("data/macroalgal_metrics_quadrat.csv"))
+```
+
+# Site/habitat/year-level mean
+
+We'll also average these data at the site/habitat/year-level to be more compatible with other data types.
+This will be done at the *quadrat-level*, meaning we won't take the transect-level summary first, since they aren't meant to be factors in a statistical design
+
+```{r get averages}
+quadData %>% 
+  group_by(year, site, habitat, island_side) %>% 
+  summarize(N = n(),
+            coral_cover_mean = mean(coral_cover),
+            coral_cover_se = sd(coral_cover)/sqrt(N),
+            coral_cover_hard_mean = mean(coral_cover_hard, na.rm = TRUE),
+            coral_cover_hard_se = sd(coral_cover_hard, na.rm = TRUE)/sqrt(N),
+            macroalgae_cover_mean = mean(macroalgae_cover),
+            macroalgae_cover_se = sd(macroalgae_cover)/sqrt(N),
+            macroalgae_cover_hard_mean = mean(macroalgae_cover_hard, na.rm = TRUE),
+            macroalgae_cover_hard_se = sd(macroalgae_cover_hard, na.rm = TRUE)/sqrt(N),
+            macroalgae_richness_mean = mean(macroalgae_richness),
+            macroalgae_richness_se = sd(macroalgae_richness)/sqrt(N),
+            macroalgae_shannon_mean = mean(macroalgae_shannon),
+            macroalgae_shannon_se = sd(macroalgae_shannon)/sqrt(N),
+            centroid_dist_habyear_mean = mean(hab_yr_distcentroid, na.rm = TRUE)
+            ) -> siteData
+
+```
+
+```{r save dataframe}
+# write_csv(siteData, here("data/macroalgal_metrics_site.csv"))
+```
+
+# All time metrics
+
+## Calculate the coefficient of variation
+
+This is basically just copying/pasting Noam's code
+
+```{r cv function}
+# coefficient of variation
+CV <- function(x){
+  return(sd(x,na.rm=T)/mean(x,na.rm=T))
+}
+```
+
+```{r calculate CV and summarize}
+quadData %>% 
+  # don't include year this time
+  group_by(site, habitat, island_side) %>%
+  summarize(N = n(),
+            coral_cover_mean = mean(coral_cover),
+            coral_cover_se = sd(coral_cover)/sqrt(N),
+            coral_cover_hard_mean = mean(coral_cover_hard, na.rm = TRUE),
+            coral_cover_hard_se = sd(coral_cover_hard, na.rm = TRUE)/sqrt(N),
+            macroalgae_cover_mean = mean(macroalgae_cover),
+            macroalgae_cover_se = sd(macroalgae_cover)/sqrt(N),
+            macroalgae_cover_hard_mean = mean(macroalgae_cover_hard, na.rm = TRUE),
+            macroalgae_cover_hard_se = sd(macroalgae_cover_hard, na.rm = TRUE)/sqrt(N),
+            macroalgae_richness_mean = mean(macroalgae_richness),
+            macroalgae_richness_se = sd(macroalgae_richness)/sqrt(N),
+            macroalgae_shannon_mean = mean(macroalgae_shannon),
+            macroalgae_shannon_se = sd(macroalgae_shannon)/sqrt(N),
+            centroid_dist_habyear_mean = mean(hab_yr_distcentroid, na.rm = TRUE)
+            )  -> alltimeCV
+```
+
+```{r save}
+write_csv(alltimeCV, here(here("data/macroalgal_metrics_all_time.csv")))
+```

### 2024-08-15 julianna_algalMetric_wrangling.Rmd
Fixed CSV names to match Noam's for max reproducibility and took coral/sand out of algal metric datasheets

@@
 ```{r calculate beta dispersion}
-yrBetaDisperFring <- vegan::betadisper(brayMatFring, sppMeta0fring$year) # warning: some squared distances are negative and changed to zero (ask Jamie about this??)
+# yrBetaDisperFring <- vegan::betadisper(brayMatFring, sppMeta0fring$year) # warning: some squared distances are negative and changed to zero (ask Jamie about this??)
 ```
@@
 ```{r calculate beta dispersion}
-yrBetaDisperBack <- vegan::betadisper(brayMatBack, sppMeta0back$year) 
+# yrBetaDisperBack <- vegan::betadisper(brayMatBack, sppMeta0back$year) 
 ```
@@
 ```{r calculate beta dispersion}
-yrBetaDisperOuter10 <- vegan::betadisper(brayMatOuter10, sppMeta0outer10$year) # error: some squared distances are negative and changed to zero
+# yrBetaDisperOuter10 <- vegan::betadisper(brayMatOuter10, sppMeta0outer10$year) # error: some squared distances are negative and changed to zero
 ```
@@
 ```{r calculate beta dispersion}
-yrBetaDisperOuter17 <- vegan::betadisper(brayMatOuter17, sppMeta0outer17$year) # error: some squared distances are negative and changed to zero
+# yrBetaDisperOuter17 <- vegan::betadisper(brayMatOuter17, sppMeta0outer17$year) # error: some squared distances are negative and changed to zero
 ```
@@
 ```{r save}
-# write_csv(quadData, here("data/algal_metrics_quadrat.csv"))
+write_csv(quadData, here("data/algae_diversity_quadrat.csv"))
 ```
@@
 ```{r save dataframe}
-# write_csv(siteData, here("data/algal_metrics_site.csv"))
+write_csv(siteData, here("data/algae_diversity_site.csv"))
 ```
@@
 ```{r save}
-write_csv(alltimeJoin, here(here("data/algal_metrics_all_time.csv")))
+write_csv(alltimeJoin, here(here("data/algae_diversity_stability_metrics_alltime.csv")))
 ```

### 2024-08-15 julianna_coral_macroalg_wrangling.Rmd
Fixed CSV names to match Noam's for max reproducibility and took coral/sand out of algal metric datasheets

@@
 ```{r calculate beta dispersion}
-yrBetaDisperFringMacro <- vegan::betadisper(brayMatFringMacro, sppMeta0fring$year) # warning: some squared distances are negative and changed to zero (ask Jamie about this??)
+# yrBetaDisperFringMacro <- vegan::betadisper(brayMatFringMacro, sppMeta0fring$year) # warning: some squared distances are negative and changed to zero (ask Jamie about this??)
 ```
@@
 ```{r calculate beta dispersion}
-yrBetaDisperBackMacro <- vegan::betadisper(brayMatBackMacro, sppMeta0back$year) 
+# yrBetaDisperBackMacro <- vegan::betadisper(brayMatBackMacro, sppMeta0back$year) 
 ```
@@
 ```{r calculate beta dispersion}
-yrBetaDisperOuter10Macro <- vegan::betadisper(brayMatOuter10Macro, sppMeta0outer10$year) # error: some squared distances are negative and changed to zero
+# yrBetaDisperOuter10Macro <- vegan::betadisper(brayMatOuter10Macro, sppMeta0outer10$year) # error: some squared distances are negative and changed to zero
 ```
@@
 ```{r calculate beta dispersion}
-yrBetaDisperOuter17Macro <- vegan::betadisper(brayMatOuter17Macro, sppMeta0outer17$year) # error: some squared distances are negative and changed to zero
+# yrBetaDisperOuter17Macro <- vegan::betadisper(brayMatOuter17Macro, sppMeta0outer17$year) # error: some squared distances are negative and changed to zero
 ```
@@
 ```{r save}
-# write_csv(quadData, here("data/macroalgal_metrics_quadrat.csv"))
+ write_csv(quadData, here("data/macroalgae_diversity_quadrat.csv"))
 ```
@@
 ```{r save dataframe}
-# write_csv(siteData, here("data/macroalgal_metrics_site.csv"))
+ write_csv(siteData, here("data/macroalgae_diversity_site.csv"))
 ```
@@
 ```{r save}
-write_csv(alltimeCV, here(here("data/macroalgal_metrics_all_time.csv")))
+write_csv(alltimeCV, here(here("data/macroalgae_stability_metrics_all_time.csv")))
 ```

### 2024-08-16 noam_data_exploration_v3.Rmd
Fixed site level synchrony and portfolio effects

@@
 
-site_level <- transect_level %>%
-  dplyr::group_by(year, island_side, site, habitat, taxa) %>%
-  dplyr::summarise(mean_site_cover = mean(mean_cover), 
-            site_se = sd(mean_cover)/sqrt(n()))
+site_level <- long_data %>%
+  dplyr::group_by(year, island_side, site, habitat, site_habitat, taxa) %>%
+  dplyr::summarise(mean_site_cover = mean(percent_cover), 
+            site_se = sd(percent_cover)/sqrt(n()))
+
 ```
@@
 # portfolio and synchrony need to be calculated at quad level, so taking the mean of the quad level for these
-synchrony_portfolio_site <- diversity_stability_quadrat %>%
-  dplyr::group_by(site, habitat, island_side, site_habitat) %>%
-  dplyr::summarise(
-            algae_loreau_mean = mean(loreau_synchrony),
-            algae_loreau_se = sd(loreau_synchrony)/sqrt(n()),
-            algae_portfolio_mean = mean(portfolio_effect),
-            algae_portfolio_se = sd(portfolio_effect)/sqrt(n()))
+# synchrony_portfolio_site <- diversity_stability_quadrat %>%
+#   dplyr::group_by(site, habitat, island_side, site_habitat) %>%
+#   dplyr::summarise(
+#             algae_loreau_mean = mean(loreau_synchrony),
+#             algae_loreau_se = sd(loreau_synchrony)/sqrt(n()),
+#             algae_portfolio_mean = mean(portfolio_effect),
+#             algae_portfolio_se = sd(portfolio_effect)/sqrt(n()))
+# NOT DOING IT LIKE THIS ANYMORE! DOING IT BY RECALCULATING SYNCHRONY WITH SITE AS REPLICATE
+lm_synchrony_site <- codyn::synchrony(df = site_level,
+                                   time.var = "year",
+                                   species.var = "taxa",
+                                   abundance.var = "mean_site_cover",
+                                   metric = "Loreau",
+                                   replicate.var = "site_habitat") # Loreau and Mazancourt synchrony
+colnames(lm_synchrony_site) = c("site_habitat", "loreau_synchrony") # rename cols for merge later
+
+wide_site_level <- data %>%
+  dplyr::group_by(year, island_side, island_side_habitat, site, habitat, site_habitat) %>%
+  dplyr::summarize(across(all_of(species_cols), mean, na.rm = TRUE))
+wide_site_level$site_habitat <- as.factor(wide_site_level$site_habitat)
+
+site_portfolio_columns <- c("site_habitat", "portfolio_effect") # names of columns for dataframe
+site_level_portfolio <- 
+  data.frame(matrix(nrow = length(unique(wide_site_level$site_habitat)), 
+                    ncol = length(site_portfolio_columns))
+             )
+colnames(site_level_portfolio) <- site_portfolio_columns # change column 
+site_level_portfolio$site_habitat <- unique(wide_site_level$site_habitat) # populate
+
+for(i in 1:length(unique(wide_site_level$site_habitat)) ){ # for each  site
+  # create a new dataframe for current site
+  newData <- subset(wide_site_level, site_habitat == levels(wide_site_level$site_habitat)[i])
+  current_matrix <- newData[ , -which(names(newData) %in% meta_cols)] # remove metadata columns
+  if(sum(current_matrix) == 0){ # if this was a 0 algae site
+    site_level_portfolio[i, 2] <- NA # set portfolio effects to 0
+  }else{ # otherwise
+    current_matrix <- current_matrix[, which(colSums(current_matrix) != 0)] # remove irrelevant algae
+
+    site_level_portfolio[i, 2] <- pe_mv(current_matrix, "loess_detrended") # calculate portfolio effects for current time series
+  }
+
+}
 
 # merge together
-diversity_stability_site <- merge(diversity_cv_site, synchrony_portfolio_site, by = c("site", "habitat", "island_side", "site_habitat"))
+diversity_stability_site <- merge(diversity_cv_site, lm_synchrony_site, by = "site_habitat")
+diversity_stability_site <- merge(diversity_stability_site, site_level_portfolio, by = "site_habitat")
+
 # write_csv(diversity_stability_site, here("data", "algae_diversity_stability_metrics_site.csv"))

### 2024-08-16 julianna_algalMetric_wrangling.Rmd
Making a clean script for Aug 29 figgies that only uses standardized dataset

@@
 ```{r}
-noamallTime <- read_csv(here("data/algae_diversity_stability_metrics_alltime.csv"))
+noamallTime <- read_csv(here("data/NOAMalgae_diversity_stability_metrics_alltime.csv"))
 ```
@@
 noamallTime %>% 
-  select(site, habitat, island_side, algae_shannon_stability, algae_loreau_mean, algae_loreau_se, algae_portfolio_mean, algae_portfolio_se) %>% 
+  select(site, habitat, island_side, algae_shannon_stability, loreau_synchrony, portfolio_effect) %>% 
   full_join(alltimeCV) -> alltimeJoin

### 2024-08-16 julianna_figs_Aug29.Rmd
Making a clean script for Aug 29 figgies that only uses standardized dataset

---
title: "Juilanna's figures for August 29"
output: html_document
date: "2024-08-16"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clean up
rm(list=ls())

# load librarian (for repeatable package loading/installation)
if (!require("librarian")) install.packages("librarian")

# load packages
librarian::shelf(
  here, # relative file paths
  tidyverse, # data wrangling
  vegan # community/diversity analyses
                 )
```

# Read in data

Want data at the quadrat level, which is our level of replication

```{r load data}
quads <- read_csv(here("data"))
```

### 2024-08-16 julianna_figs_Aug29.Rmd
Finished exploratory plots/code for Aug 29 meeting (now on the slide deck)

@@
   tidyverse, # data wrangling
-  vegan # community/diversity analyses
+  vegan, # community/diversity analyses
+  MoMAColors # color palette: can install using devtools::install_github("BlakeRMills/MoMAColors")
                  )
@@
 
-Want data at the quadrat level, which is our level of replication
+Want data at the quadrat-level data for the trends through time plot.
 
 ```{r load data}
-quads <- read_csv(here("data"))
+# raw algae information (e.g., for grabbing CCA for Fig 1.)
+algae <- read_csv(here("data/algaeonly_cleaned_colnames.csv"))
+
+# algae standardized dataset with diversity information at the quadrat-level (generated using julianna_algalMetric_wrangling.Rmd)
+algaeQuad <- read_csv(here("data/algae_diversity_quadrat.csv"))
+
+# macroalgae & coral dataset in the ancillary data Drive (generated using julianna_coral_macroalg_wrangling.Rmd)
+macroalgQuad <- read_csv(here("data/macroalgae_coral_diversity_quadrat.csv"))
+
+# all time metrics for stability
+algaeAT <- read_csv(here("data/algae_diversity_stability_metrics_alltime.csv")) 
+macroalgAT <- read_csv(here("data/macroalgae_coral_diversity_stability_metrics_alltime.csv")) 
 ```
+
+# Patterns of 4 key spaceholders through time
+
+Deron suggested we make this figure: macroalgae, CCA, turf, and coral through time
+
+```{r get a dataset with the main four spaceholders}
+macroalgQuad %>% 
+  select(year, location, site, habitat, transect, quadrat, island_side, macroalgae_cover, macroalgae_cover_hard, coral_cover, coral_cover_hard, sand_cover) %>% 
+  full_join(algae) %>%
+  rowwise() %>% 
+  # get turf and CCA cover
+  mutate(turf_cover = sum(algal_turf, damselfish_turf),
+         turf_cover_hard = case_when(
+           sand_cover == 100 ~ NA, # since we can't divide by a zero
+           TRUE ~ (turf_cover/(100-sand_cover))*100 
+           ),
+         CCA_cover = sum(lithophyllum_kotschyanum, crustose_corallines, neogoniolithon_brassica_florida, sporolithon_sp),
+         CCA_cover_hard = case_when(
+           sand_cover == 100 ~ NA, # since we can't divide by a zero
+           TRUE ~ (CCA_cover/(100-sand_cover))*100 
+           )
+         ) %>% 
+  select(year, location, site, habitat, transect, quadrat, island_side, 
+         macroalgae_cover, macroalgae_cover_hard, coral_cover, coral_cover_hard, sand_cover, 
+         turf_cover, turf_cover_hard, CCA_cover, CCA_cover_hard
+         ) %>% 
+  # relevel for graphing
+  mutate(island_side = str_to_sentence(island_side),
+         island_side = factor(island_side, levels = c("North", "East", "West")),
+         habitat = case_when(
+           str_detect(habitat, pattern = "_") ~ str_to_sentence(paste0(str_split(habitat, "_")[[1]][1], " ", str_split(habitat, "_")[[1]][2])),
+           TRUE ~ str_to_sentence(habitat)),
+         habitat = factor(habitat, levels = c("Fringing", "Backreef", "Outer 10", "Outer 17"))) %>% 
+  mutate(site = case_when(
+    site == "lter_1" ~ "LTER 1 (N)",
+    site == "lter_2" ~ "LTER 2 (N)",
+    site == "lter_3" ~ "LTER 3 (E)",
+    site == "lter_4" ~ "LTER 4 (E)",
+    site == "lter_5" ~ "LTER 5 (W)",
+    site == "lter_6" ~ "LTER 6 (W)"
+  )) -> main4quad
+```
+
+```{r plot for absolute cover}
+main4quad %>% 
+  select(year, location, site, habitat, transect, quadrat, island_side, 
+         macroalgae_cover, turf_cover, CCA_cover, coral_cover) %>% 
+  # pivot key columns longer
+  pivot_longer(cols = c(macroalgae_cover, turf_cover, CCA_cover, coral_cover),
+               names_to = "Substrate",
+               values_to = "cover") %>% 
+  # summarize and group quadrats by island/habitat
+  group_by(year, habitat, site, island_side, Substrate) %>% 
+  summarise(mean_cov = mean(cover),
+            sd_cov = sd(cover, na.rm = TRUE),
+            n = n(),
+            CI95 = 1.96*(sd_cov)/sqrt(n)) %>% 
+  # plot
+  ggplot(aes(x = year, y = mean_cov, group = Substrate, fill = Substrate)) +
+  geom_ribbon(aes(ymin = mean_cov -CI95, 
+                    ymax = mean_cov + CI95),
+              alpha  = 0.7) +
+  geom_line() +
+  theme_bw() +
+   scale_fill_manual(values = c(
+    moma.colors("Klein", type = "discrete")[2], # CCA
+    moma.colors("Klein", type = "discrete")[1], # coral
+    moma.colors("Klein", type = "discrete")[5], # macro
+    moma.colors("Klein", type = "discrete")[4] # turf
+  ),
+  labels = c("CCA", "Coral", "Macroalgae", "Turf algae")) +
+  facet_grid(site ~ habitat) + # not sure which direction is best here
+  ylab("Mean absolute cover (%)") +
+  xlab("Year") +
+  # add lines for disturbances
+  geom_vline(xintercept = 2006, linetype = "dotdash") + # start of COTS
+  geom_vline(xintercept = 2010,  color = "darkgray") + # cyclone Oli
+  geom_vline(xintercept = 2019, linetype = "dashed", color = "red") 
+```
+
+```{r plot for percent of hard cover}
+main4quad %>% 
+  select(year, location, site, habitat, transect, quadrat, island_side, 
+         macroalgae_cover_hard, turf_cover_hard, CCA_cover_hard, coral_cover_hard) %>% 
+  # pivot key columns longer
+  pivot_longer(cols = c(macroalgae_cover_hard, turf_cover_hard, CCA_cover_hard, coral_cover_hard),
+               names_to = "Substrate",
+               values_to = "cover") %>% 
+  # summarize and group quadrats by island/habitat
+  group_by(year, habitat, site, island_side, Substrate) %>% 
+  summarise(mean_cov = mean(cover, na.rm = TRUE),
+            sd_cov = sd(cover, na.rm = TRUE),
+            n = n(),
+            CI95 = 1.96*(sd_cov)/sqrt(n)) %>% 
+  # plot
+  ggplot(aes(x = year, y = mean_cov, group = Substrate, fill = Substrate)) +
+  geom_ribbon(aes(ymin = mean_cov -CI95, 
+                    ymax = mean_cov + CI95),
+              alpha  = 0.7) +
+  geom_line() +
+  theme_bw() +
+   scale_fill_manual(values = c(
+    moma.colors("Klein", type = "discrete")[2], # CCA
+    moma.colors("Klein", type = "discrete")[1], # coral
+    moma.colors("Klein", type = "discrete")[5], # macro
+    moma.colors("Klein", type = "discrete")[4] # turf
+  ),
+  labels = c("CCA", "Coral", "Macroalgae", "Turf algae")) +
+  facet_grid(site ~ habitat) + # not sure which direction is best here
+  ylab("Mean cover of hard substrate(%)") +
+  xlab("Year") +
+  # add lines for disturbances
+  geom_vline(xintercept = 2006, linetype = "dotdash") + # start of COTS
+  geom_vline(xintercept = 2010,  color = "darkgray") + # cyclone Oli
+  geom_vline(xintercept = 2019, linetype = "dashed", color = "red") 
+```
+
+# Coral cover vs. algal metrics
+
+```{r join coral and algae}
+algaeQuad %>% 
+  full_join(macroalgQuad) -> coralAlgaeQuad
+```
+
+```{r plot algal-richness vs. coral}
+coralAlgaeQuad %>% 
+  ggplot(aes(x = algae_richness, y = coral_cover_hard, color = island_side, group = algae_richness)) +
+  geom_boxplot() + 
+  facet_grid(habitat~site) +
+  scale_color_manual(values = c("darkblue", "darkred", "darkgreen")) +
+  theme_classic()
+```
+
+```{r plot macroalgal-richness vs. coral}
+coralAlgaeQuad %>% 
+  ggplot(aes(x = macroalgae_richness, y = coral_cover_hard, color = island_side, group = macroalgae_richness)) +
+  geom_boxplot() + 
+  facet_grid(habitat~site) +
+  scale_color_manual(values = c("darkblue", "darkred", "darkgreen")) +
+  theme_classic()
+```
+
+```{r plot macroalgal-richness vs. algal richness}
+coralAlgaeQuad %>% 
+  ggplot(aes(x = algae_richness, y = macroalgae_richness, color = island_side, group = algae_richness)) +
+  geom_point() + 
+  facet_grid(habitat~site) +
+  scale_color_manual(values = c("darkblue", "darkred", "darkgreen")) +
+  theme_classic()
+```
+
+# Beta diversity
+
+```{r beta diversity with disturbances}
+coralAlgaeQuad %>% 
+  # first get average coral cover per year
+  group_by(year, habitat, site) %>% 
+  summarize(coral_cover_hard_mean = mean(coral_cover_hard, na.rm = TRUE)) %>% 
+  full_join(coralAlgaeQuad) %>% 
+  ggplot(aes(x = year, y = hab_yr_distcentroid, group = year)) +
+  geom_boxplot() +
+  ylab("Distance to centroid") +
+  xlab("Year") + 
+  theme_bw() +
+  facet_grid(site~habitat) +
+  # add lines for disturbances
+  geom_vline(xintercept = 2006, linetype = "dotdash") + # start of COTS
+  geom_vline(xintercept = 2010,  color = "darkgray") + # cyclone Oli
+  geom_vline(xintercept = 2019, linetype = "dashed", color = "red")  + 
+  # Custom the Y scales:
+  scale_y_continuous(
+    # Features of the first axis
+    name = "Algal beta diversity",
+    # Add a second axis and specify its features
+    sec.axis = sec_axis( trans=~.*100, name="Coral (% hard)")
+  ) +
+  geom_line(aes(group = 1, x = year, y = coral_cover_hard_mean/100), size = 1, color = "coral")
+```
+
+# Algal stability vs. coral cover
+
+Join both data frames
+
+```{r join all time data frames}
+macroalgAT %>% 
+  full_join(algaeAT) -> alltime
+```
+
+Plot
+
+```{r cover CV vs. coral cover}
+alltime %>% 
+  ggplot(aes(x = coral_cover_hard_mean, y = algae_cover_hard_cv)) +
+  geom_point(aes(color = island_side)) +
+  facet_wrap(~habitat) +
+  geom_smooth(method = "lm") +
+  theme_bw()
+```
+
+```{r cover CV vs. coral cover CV}
+alltime %>% 
+  ggplot(aes(x = coral_cover_hard_cv, y = algae_cover_hard_cv)) +
+  geom_point(aes(color = island_side)) +
+  facet_wrap(~habitat) +
+  geom_smooth(method = "lm") +
+  theme_bw()
+```
+
+```{r richness CV vs. coral cover CV}
+alltime %>% 
+  ggplot(aes(x = coral_cover_hard_cv, y = algae_richness_cv)) +
+  geom_point(aes(color = island_side)) +
+  facet_wrap(~habitat) +
+  geom_smooth(method = "lm") +
+  theme_bw()
+```
+
+```{r synchrony vs. coral cover}
+alltime %>% 
+  #filter(habitat == "outer_10" | habitat ==  "outer_17") %>% 
+  ggplot(aes(x = coral_cover_hard_mean, y = loreau_synchrony)) +
+  geom_point(aes(color = island_side)) +
+  #facet_wrap(~habitat) +
+  geom_smooth(method = "lm") +
+  theme_bw()
+```
+
+```{r synchrony vs. coral cover CV}
+alltime %>% 
+  ggplot(aes(x = coral_cover_hard_cv, y = loreau_synchrony)) +
+  geom_point(aes(color = island_side)) +
+  facet_wrap(~habitat) +
+  geom_smooth(method = "lm") +
+  theme_bw()
+```
+
+```{r portfolio effects vs. coral cover CV}
+alltime %>% 
+  ggplot(aes(x = coral_cover_hard_cv, y = portfolio_effect)) +
+  geom_point(aes(color = island_side)) +
+  facet_wrap(~habitat) +
+  geom_smooth(method = "lm") +
+  theme_bw()
+```

### 2024-08-16 noam_data_exploration_v3.Rmd
re-ran models at island side level. Began figuring out bray curtis/beta diversity stuff at quad level. Still WIP.

@@
   ggplot(data = alpha_diversity_quad_means) +
-  geom_point(aes(x = year, y = richness_mean, colour = site, shape = habitat, group = site_habitat), size = 8) +
-  geom_line(aes(x = year, y = richness_mean, colour = site, group = site_habitat), size = 1.2) +
-  geom_errorbar(aes(x = year, ymin = richness_mean - richness_se , ymax = richness_mean + richness_se, colour = site,
-                    group = site_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = site_colours) +
+  geom_point(aes(x = year, y = richness_mean, colour = island_side, shape = habitat, group = island_side_habitat), size = 8) +
+  geom_line(aes(x = year, y = richness_mean, colour = island_side, group = island_side_habitat), size = 1.2) +
+  geom_errorbar(aes(x = year, ymin = richness_mean - richness_se , ymax = richness_mean + richness_se, colour = island_side,
+                    group = island_side_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = island_side_colours) +
   scale_shape_manual(values = habitat_shapes) +
@@
 
-columns <- c("location", "Intercept", "Slope") # names of columns for dataframe
+columns <- c("location", "Intercept", "Slope", "Phi") # names of columns for dataframe
 gls_coefficients <- # new dataframe name
@@
   gls_coefficients[i, 3] <- unname(coef(loop_mod)[2])
+  gls_coefficients[i, 4] <- loop_mod[["modelStruct"]][["corStruct"]]
 
@@
 hist(gls_coefficients$Intercept) # same with intercept - although it probably shouldn't be
+hist(gls_coefficients$Phi)
 
@@
 emmeans(slope_mod, pairwise ~ site | habitat)
-```
-
-```{r gls plot, fig.height=20, fig.width=20}
+
+## AT COAST-LEVEL
+slope_mod <- glmmTMB(Slope ~ habitat*island_side + (1|island_side/site/transect), data = gls_coefficients_merged)
+summary(slope_mod) # pretty low variance in random effects
+plot(gls_res <- simulateResiduals(slope_mod))
+
+hist(residuals(slope_mod))
+plot(predict(slope_mod) ~ residuals(slope_mod))
+Anova(slope_mod)
+#                        Chisq Df Pr(>Chisq)    
+# habitat             1301.246  3  < 2.2e-16 ***
+# island_side           16.075  2  0.0003231 ***
+# habitat:island_side  799.999  6  < 2.2e-16 ***
+r.squaredGLMM(slope_mod) # 0.16, 0.18
+plot(allEffects(slope_mod))
+# backreef: richness increases on east and north shore, but relatively constant west side
+# Fringe: richness decreasing on east and west side, but relatvely constant on north shore
+# outer_10: richness neutral/decreasing on east side, but neutral/increasing on north and west side
+# outer_17: richness decreasing north/east side, but neutral west side
+# Worth noting trends are slight - arguable whether they are biologically meaningful
+
+slope_emm <- emmeans(slope_mod, ~ habitat*island_side) # create emmeans object
+emmeans::contrast(slope_emm, "consec", simple = "each", combine = TRUE, adjust = "bonferroni")
+# Note: not sure I want to include this
+emmeans(slope_mod, pairwise ~ island_side | habitat)
+# north and east coast not different from each other in backreef
+# east and west not different from each other in fringing
+# north and west not different from each other outer 10
+# no sites different from each other outer 17
+slope_est <- as.data.frame(slope_emm) # create df of emmeans estimates
+
+```
+
+```{r gls plot, fig.height=20, fig.width=25}
 # not sure how to visualize here... thinking something like Dornelas et al. 2014 Fig 2a
@@
 
-ggplot(data = gls_coefficients_merged) +
-  geom_histogram(aes(x = Slope), bins = 30) +
+ggplot(data = gls_coefficients_merged, aes(x = Slope)) +
+  geom_histogram(bins = 40, aes(y = after_stat(ndensity))) +
+  geom_density(alpha=0,aes(y = after_stat(ndensity)), size = 1.5 ) +
   geom_vline(xintercept = 0, size = 2, linetype = 2) +
-  facet_grid(habitat ~ site) +
-#  theme_classic() +
+  geom_ribbon(data = merge(gls_coefficients_merged, slope_est), aes(xmin = lower.CL, xmax = upper.CL, ymin = 0, ymax = Inf),
+              fill = "red", alpha = 0.1)+
+  geom_vline(data = slope_est, aes(xintercept = emmean), size = 2, linetype = 1, colour = "red") +
+  facet_grid(island_side ~ habitat) +
+  theme_classic() +
+  labs(y = "Density\n") +
   theme(panel.grid.major = element_blank(), 
@@
 
-```{r gls intercept, fig.height=20, fig.width=20}
-intercept_mod <- glmmTMB(Intercept ~ habitat*island_side + (1|site/transect), data = gls_coefficients_merged)
+```{r gls intercept, fig.height=20, fig.width=25}
+intercept_mod <- glmmTMB(Intercept ~ habitat*island_side + (1|island_side/site/transect), data = gls_coefficients_merged)
 summary(intercept_mod) # a lot of variance w/i transect, less so w/i site
@@
 #                        Chisq Df Pr(>Chisq)    
-# habitat             1307.838  3  < 2.2e-16 ***
-# island_side           26.358  2  1.889e-06 ***
-# habitat:island_side  789.627  6  < 2.2e-16 ***
-r.squaredGLMM(intercept_mod) # 0.16, 0.18
+# habitat             1285.762  3  < 2.2e-16 ***
+# island_side           15.644  2  0.0004009 ***
+# habitat:island_side  776.312  6  < 2.2e-16 ***
+r.squaredGLMM(intercept_mod) # 0.16, 0.17
 plot(allEffects(intercept_mod))
 
-ggplot(data = gls_coefficients_merged) +
-  geom_histogram(aes(x = Intercept), bins = 30) +
+int_emm <- emmeans(intercept_mod, ~ habitat*island_side) # create emmeans object
+#emmeans::contrast(int_emm, "consec", simple = "each", combine = TRUE, adjust = "bonferroni")
+
+emmeans(intercept_mod, pairwise ~ island_side | habitat)
+# same trends as with slope
+
+int_est <- as.data.frame(int_emm) # create df of emmeans estimates
+
+ggplot(data = gls_coefficients_merged, aes(x = Intercept)) +
+  geom_histogram(bins = 40, aes(y = after_stat(ndensity))) +
+  geom_density(alpha=0,aes(y = after_stat(ndensity)), size = 1.5 ) +
   geom_vline(xintercept = 0, size = 2, linetype = 2) +
-  facet_grid(habitat ~ island_side) +
-#  theme_classic() +
+  geom_ribbon(data = merge(gls_coefficients_merged, int_est), aes(xmin = lower.CL, xmax = upper.CL, ymin = 0, ymax = Inf),
+              fill = "red", alpha = 0.1)+
+  geom_vline(data = int_est, aes(xintercept = emmean), size = 2, linetype = 1, colour = "red") +
+  facet_grid(island_side ~ habitat) +
+  theme_classic() +
+  labs(y = "Density\n") +
   theme(panel.grid.major = element_blank(), 
@@
 richness_mod_2 <- glmmTMB(
-  richness ~ year_factor*habitat*site + (1 | site/transect/location) + ar1(year_factor + 0 | location),
+  richness ~ year_factor*habitat*island_side + (1 | island_side/transect/location) + ar1(year_factor + 0 | location),
   family = poisson,
@@
 
-top_rich_mod <- glmmTMB(richness ~ year_factor*habitat + year_factor*site + (1 | site/transect/location) + ar1(year_factor + 0 | location),   family = poisson, data = alpha_diversity_quad, control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000)))
-summary(top_rich_mod)
-Anova(top_rich_mod)
-```
+# top_rich_mod <- glmmTMB(richness ~ year_factor*habitat + year_factor*island_side + (1 | island_side/transect/location) + # ar1(year_factor + 0 | location),   family = poisson, data = alpha_diversity_quad, control = glmmTMBControl(optCtrl = list(iter.max = # 2000, eval.max = 2000)))
+# summary(top_rich_mod)
+# Anova(top_rich_mod)
+
+# Julianna's model structure
+j_rich_mod <- glmmTMB(richness ~ year + habitat + (1|site/transect),
+               family = poisson,
+               data = alpha_diversity_quad)
+summary(j_rich_mod)
+plot(allEffects(j_rich_mod))
+
+grid <- expand_grid(year = unique(alpha_diversity_quad$year),
+                       habitat = unique(alpha_diversity_quad$habitat),
+                       site = unique(alpha_diversity_quad$site),
+                       island_side = unique(alpha_diversity_quad$island_side),
+                       transect = unique(alpha_diversity_quad$transect))
+grid$Preds <- predict(j_rich_mod, newdata = grid)
+
+grid %>% 
+  # get transect means
+  group_by(year, habitat, island_side) %>% 
+  summarize(Pred_mean_richness = mean(Preds),
+            SD_richness = sd(Preds)) %>% 
+  ggplot(aes(x = year, y = Pred_mean_richness, group = habitat, color = habitat)) +
+  geom_line() +
+  theme_bw() +
+  facet_wrap(~island_side)
+
+# Given that this is largely looking at slopes for temporal trends in adiv, I prefer the gls
+```
+
 ```{r turnover troubleshooting, fig.height = 30, fig.width=25}
@@
 # model
-total_turnover_mod1 <- glmmTMB(Turnover_trans ~ year_factor*habitat*site +
-                                 (1 | site/transect/location) + ar1(year_factor + 0 | location),
+total_turnover_mod1 <- glmmTMB(Turnover_trans ~ year_factor*habitat*island_side +
+                                 (1 | island_side/site/transect) + ar1(year_factor + 0 | location),
                                family = beta_family(), data = total_turnover)
@@
 
-total_turnover_mod2 <- glmmTMB(Turnover_trans ~ year_factor*habitat*site + (1 | site/transect/location),
-                               family = beta_family(), data = total_turnover)
-summary(total_turnover_mod2) # 
-plot(total_turnover_res2 <- simulateResiduals(total_turnover_mod2)) # residuals still look like shit
-Anova(total_turnover_mod2) 
-anova(total_turnover_mod2, total_turnover_mod1) # models are sig different from each other
-AICc(total_turnover_mod2, total_turnover_mod1) # first model is better
-# need to figure out residuals problem
-plot(allEffects(total_turnover_mod1))
+# total_turnover_mod2 <- glmmTMB(Turnover_trans ~ year_factor*habitat*site + (1 | site/transect/location),
+#                                family = beta_family(), data = total_turnover)
+# summary(total_turnover_mod2) # 
+# plot(total_turnover_res2 <- simulateResiduals(total_turnover_mod2)) # residuals still look like shit
+# Anova(total_turnover_mod2) 
+# anova(total_turnover_mod2, total_turnover_mod1) # models are sig different from each other
+# AICc(total_turnover_mod2, total_turnover_mod1) # first model is better
+# # need to figure out residuals problem
+# plot(allEffects(total_turnover_mod1))
+#
 
@@
   ggplot(data = total_turnover_means) +
-  geom_point(aes(x = year, y = Turnover_mean, colour = site, shape = habitat, group = site_habitat), size = 8, stroke = 2) +
-  geom_line(aes(x = year, y = Turnover_mean, colour = site, group = site_habitat), size = 1.2) +
-  geom_errorbar(aes(x = year, ymin = Turnover_mean - Turnover_se , ymax = Turnover_mean + Turnover_se, colour = site,
-                    group = site_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~site_habitat, nrow = 6)+
+  geom_point(aes(x = year, y = Turnover_mean, colour = island_side, shape = habitat, group = island_side_habitat), size = 8, stroke = 2) +
+  geom_line(aes(x = year, y = Turnover_mean, colour = island_side, group = island_side_habitat), size = 1.2) +
+  geom_errorbar(aes(x = year, ymin = Turnover_mean - Turnover_se , ymax = Turnover_mean + Turnover_se, colour = island_side,
+                    group = island_side_habitat), width = 0.5, size = 1.2) +
+  scale_colour_manual(values = island_side_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~island_side_habitat, nrow = 6)+
   labs(title = "a. Turnover\n", y = "Total turnover\n") +
@@
 
-```{r nmds, fig.height = 15, fig.width = 15}
+```{r bray curtis set up}
+species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
+species_matrix <- species_matrix/100 # convert from percent to proportion
+rownames(species_matrix) <- data$location
+rows_to_remove <- rowSums(species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
+species_matrix <- species_matrix[!rows_to_remove,] 
+data_subset <- data[!rows_to_remove,] # make sure rows are also removed from dataframe
+# data_subset$stratum <- paste(data_subset$year, "-", data_subset$site, "-", data_subset$habitat )
+brayfull <- vegdist(species_matrix, "bray")
+# View(brayfull)
+
+# playing around with stuff here - ignore for now
+x <- subset(data_subset, location == "lter_1_backreef_algae_transect_1_quad_1")
+year_combinations <- expand.grid(Year1 = x$year, Year2 = x$year)
+year_combinations <- year_combinations[year_combinations$Year1 != year_combinations$Year2, ] # filter out year 1 == year 2
+year_combinations$Year_Difference <- abs(year_combinations$Year1 - year_combinations$Year2) # find difference in years
+year_combinations$location <- rep("lter_1_backreef_algae_transect_1_quad_1", nrow(year_combinations))
+year_combinations$bray <- NA
+for(i in 1:nrow(year_combinations)){
+  subset(x)
+}
+
+```
+
+```{r old nmds code, fig.height = 15, fig.width = 15, eval = FALSE}
 species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
@@
 data_subset$stratum <- paste(data_subset$year, "-", data_subset$site, "-", data_subset$habitat )
-
 # identifying species w/ > 10% cover for plotting later

### 2024-08-16 julianna_coral_macroalg_wrangling.Rmd

@@
   full_join(betaDHabYr) %>% 
+  rename(macroalgae_centroid_dist_habyear_mean = hab_yr_distcentroid) %>% 
   # get just the columns we need
@@
 ```{r save}
- write_csv(quadData, here("data/macroalgae_diversity_quadrat.csv"))
+ write_csv(quadData, here("data/macroalgae_coral_diversity_quadrat.csv"))
 ```
@@
             macroalgae_shannon_se = sd(macroalgae_shannon)/sqrt(N),
-            centroid_dist_habyear_mean = mean(hab_yr_distcentroid, na.rm = TRUE)
+            macroalgae_centroid_dist_habyear_mean = mean(macroalgae_centroid_dist_habyear_mean, na.rm = TRUE)
             ) -> siteData
@@
 ```{r save dataframe}
- write_csv(siteData, here("data/macroalgae_diversity_site.csv"))
+ write_csv(siteData, here("data/macroalgae_coral_diversity_site.csv"))
 ```
@@
   summarize(N = n(),
+
+            # coral
             coral_cover_mean = mean(coral_cover),
@@
             coral_cover_hard_se = sd(coral_cover_hard, na.rm = TRUE)/sqrt(N),
+            coral_cover_cv = CV(coral_cover),
+            coral_cover_hard_cv = CV(coral_cover_hard),
+
+            # macroalgae cover
             macroalgae_cover_mean = mean(macroalgae_cover),
@@
             macroalgae_cover_hard_se = sd(macroalgae_cover_hard, na.rm = TRUE)/sqrt(N),
+            macroalgae_cover_cv = CV(macroalgae_cover),
+            macroalgae_cover_hard_cv = CV(macroalgae_cover_hard),
+
+            # macroalgae richness
             macroalgae_richness_mean = mean(macroalgae_richness),
@@
             macroalgae_shannon_se = sd(macroalgae_shannon)/sqrt(N),
-            centroid_dist_habyear_mean = mean(hab_yr_distcentroid, na.rm = TRUE)
+            macroalgae_centroid_dist_habyear_mean = mean(macroalgae_centroid_dist_habyear_mean, na.rm = TRUE)
             )  -> alltimeCV
@@
 ```{r save}
-write_csv(alltimeCV, here(here("data/macroalgae_stability_metrics_all_time.csv")))
+write_csv(alltimeCV, here(here("data/macroalgae_coral_diversity_stability_metrics_alltime.csv")))
 ```

### 2024-08-16 julianna_richness_model.Rmd

--- 2024-07-30 julianna_richness_model.Rmd
@@
 require(lme4) # mixed models
+
 ```
@@
 ```{r load data}
-cleaned_rich <- read_csv(here("data/algal_metrics_quadrat.csv")) # generated using my (Julianna's) algal metric code
+cleaned_rich <- read_csv(here("data/algae_diversity_quadrat.csv")) %>% # generated using my (Julianna's) algal metric wrangling code
+  filter(year != 2020) %>% # fake data
+  # make a unique quadrat ID
+  mutate(quad_id = paste0(transect, "_", quadrat))
+
 ```
 
-# Make a model
+# Full model no site
 
-```{r}
-mRich <- glmer(,
-               family = Poisson,
+WARNING: This is reeeeeally slow code to run
+
+```{r make Poisson model}
+mRich <- glmer(algae_richness ~ as.factor(year)*habitat + (1|site/quad_id),
+               family = poisson,
                data = cleaned_rich)
 ```
+
+# Examine output
+
+```{r make prediction grid}
+# create a grid to predict over
+newdata <- expand_grid(year = unique(cleaned_rich$year),
+                       habitat = unique(cleaned_rich$habitat),
+                       site = unique(cleaned_rich$site),
+                       quad_id = unique(cleaned_rich$quad_id))
+```
+
+```{r generate predictions}
+newdata$Preds <- predict(mRich, newdata = newdata)
+```
+
+**IMPORTANT**: I don't think this is the correct way to get the error bars
+
+```{r plot predictions}
+newdata %>% 
+  # get transect means
+  group_by(year, site, habitat) %>% 
+  summarize(Pred_mean_richness = mean(Preds),
+            SD_richness = sd(Preds),
+            N = n(),
+            SE_richness = SD_richness/sqrt(N)
+            ) %>% 
+  ggplot(aes(x = year, y = Pred_mean_richness, group = habitat, color = habitat)) +
+  geom_line() +
+  theme_bw() +
+  # add lines for disturbances
+  geom_vline(xintercept = 2006, linetype = "dotdash") + # start of COTS
+  geom_vline(xintercept = 2010,  color = "darkgray") + # cyclone Oli
+  geom_vline(xintercept = 2019, linetype = "dashed", color = "red")  # bleaching
+
+```
+
+# Full model with site
+
+```{r make Poisson model}
+mRichSite <- glmer(algae_richness ~ as.factor(year)*habitat*site + (1|quad_id),
+               family = poisson,
+               data = cleaned_rich,
+               control = glmerControl(optimizer = "bobyqa")) # this taken from : https://stackoverflow.com/questions/69929896/glmer-taking-days-to-compute-and-never-ending)
+```
+
+# Examine output
+
+```{r make prediction grid}
+# create a grid to predict over
+newdataSite <- expand_grid(year = unique(cleaned_rich$year),
+                       habitat = unique(cleaned_rich$habitat),
+                       site = unique(cleaned_rich$site),
+                       quad_id = unique(cleaned_rich$quad_id))
+```
+
+```{r generate predictions}
+newdataSite$Preds <- predict(mRichSite, newdata = newdataSite)
+```
+
+```{r plot predictions}
+newdataSite %>% 
+  # get transect means
+  group_by(year, habitat, site) %>% 
+  summarize(Pred_mean_richness = mean(Preds),
+            SD_richness = sd(Preds),
+            N = n(),
+            SE_richness = SD_richness/sqrt(N)
+            ) %>% 
+  ggplot(aes(x = year, y = Pred_mean_richness, group = habitat, color = habitat)) +
+  geom_line() +
+  theme_bw() +
+  facet_wrap(~site) +
+  geom_ribbon(ymin = Pred_mean_richness-(2*SE_richness),
+              ymax = Pred_mean_richness+(2*SE_richness),
+              alpha = 0.5) +
+  # add lines for disturbances
+  geom_vline(xintercept = 2006, linetype = "dotdash") + # start of COTS
+  geom_vline(xintercept = 2010,  color = "darkgray") + # cyclone Oli
+  geom_vline(xintercept = 2019, linetype = "dashed", color = "red")  # bleaching
+
+```

### 2024-08-16 julianna_richness_model.Rmd
Made some bad richness model/predictions--still need to figure out this rank deficient thing

--- 2024-07-30 julianna_richness_model.Rmd
@@
 ```{r load data}
-cleaned_rich <- read_csv(here("data/algal_metrics_quadrat.csv")) # generated using my (Julianna's) algal metric code
+cleaned_rich <- read_csv(here("data/algae_diversity_quadrat.csv")) %>% # generated using my (Julianna's) algal metric wrangling code
+  filter(year != 2020) %>% # fake data
+  # make a unique quadrat ID
+  mutate(quad_id = paste0(transect, "_", quadrat))
+
 ```
 
-# Make a model
+# Full model no site
 
-```{r}
-mRich <- glmer(,
-               family = Poisson,
+WARNING: This is reeeeeally slow code to run
+
+```{r make Poisson model}
+mRich <- glmer(algae_richness ~ as.factor(year)*habitat + (1|site/quad_id),
+               family = poisson,
                data = cleaned_rich)
 ```
+
+# Examine output
+
+```{r make prediction grid}
+# create a grid to predict over
+newdata <- expand_grid(year = unique(cleaned_rich$year),
+                       habitat = unique(cleaned_rich$habitat),
+                       site = unique(cleaned_rich$site),
+                       quad_id = unique(cleaned_rich$quad_id))
+```
+
+```{r generate predictions}
+newdata$Preds <- predict(mRich, newdata = newdata, "response")
+```
+
+**IMPORTANT**: I don't think this is the correct way to get the error bars
+
+```{r plot predictions}
+newdata %>% 
+  # get transect means
+  group_by(year, habitat) %>% 
+  summarize(Pred_mean_richness = mean(Preds),
+            SD_richness = sd(Preds),
+            N = n(),
+            SE_richness = SD_richness/sqrt(N)
+            ) %>% 
+  ggplot(aes(x = year, y = Pred_mean_richness, group = habitat, color = habitat)) +
+  geom_line() +
+  theme_bw() +
+  ylab("Predicted mean richness") +
+  xlab("Year") +
+  # add lines for disturbances
+  geom_vline(xintercept = 2006, linetype = "dotdash") + # start of COTS
+  geom_vline(xintercept = 2010,  color = "darkgray") + # cyclone Oli
+  geom_vline(xintercept = 2019, linetype = "dashed", color = "red")  # bleaching
+
+```
+
+# Full model with site
+
+```{r make Poisson model}
+mRichSite <- glm(algae_richness ~ as.factor(year)*habitat*site,
+               family = poisson,
+               data = cleaned_rich) # optimizer when using glmer() was taken from : https://stackoverflow.com/questions/69929896/glmer-taking-days-to-compute-and-never-ending)
+# not converging I'm not sure what's wrong? Maybe that's just like literally specifying everything? Took 1|quad_id out
+```
+
+# Examine output
+
+```{r make prediction grid}
+# create a grid to predict over
+newdataSite <- expand_grid(year = unique(cleaned_rich$year),
+                       habitat = unique(cleaned_rich$habitat),
+                       site = unique(cleaned_rich$site),
+                       quad_id = unique(cleaned_rich$quad_id))
+```
+
+```{r generate predictions}
+newdataSite$Preds <- predict(mRichSite, newdata = newdataSite, "response")
+```
+
+```{r plot predictions}
+newdataSite %>% 
+  # get transect means
+  group_by(year, habitat, site) %>% 
+  summarize(Pred_mean_richness = mean(Preds),
+            SD_richness = sd(Preds),
+            N = n(),
+            SE_richness = SD_richness/sqrt(N)
+            ) %>% 
+  ggplot(aes(x = year, y = Pred_mean_richness, group = habitat)) +
+  geom_line(aes(color = site, group = site)) +
+  theme_bw() +
+  facet_wrap(~habitat) +
+  # add lines for disturbances
+  geom_vline(xintercept = 2006, linetype = "dotdash") + # start of COTS
+  geom_vline(xintercept = 2010,  color = "darkgray") + # cyclone Oli
+  geom_vline(xintercept = 2019, linetype = "dashed", color = "red")  # bleaching
+
+```

### 2024-08-16 noam_data_exploration_v3.Rmd

@@
 for(i in 1:nrow(year_combinations)){
-  subset(x)
+  subset(x, year == year_combinations[i,1] | year_combinations[i,2])
 }

### 2024-08-19 noam_data_exploration_v3.Rmd
Moved bray curtis/beta diversity stuff to new markdown due to memory issues

--- 2024-08-16 noam_data_exploration_v3.Rmd
@@
 ```
-
-```{r bray curtis set up}
-species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
-species_matrix <- species_matrix/100 # convert from percent to proportion
-rownames(species_matrix) <- data$location
-rows_to_remove <- rowSums(species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
-species_matrix <- species_matrix[!rows_to_remove,] 
-data_subset <- data[!rows_to_remove,] # make sure rows are also removed from dataframe
-# data_subset$stratum <- paste(data_subset$year, "-", data_subset$site, "-", data_subset$habitat )
-brayfull <- vegdist(species_matrix, "bray")
-# View(brayfull)
-
-# playing around with stuff here - ignore for now
-x <- subset(data_subset, location == "lter_1_backreef_algae_transect_1_quad_1")
-year_combinations <- expand.grid(Year1 = x$year, Year2 = x$year)
-year_combinations <- year_combinations[year_combinations$Year1 != year_combinations$Year2, ] # filter out year 1 == year 2
-year_combinations$Year_Difference <- abs(year_combinations$Year1 - year_combinations$Year2) # find difference in years
-year_combinations$location <- rep("lter_1_backreef_algae_transect_1_quad_1", nrow(year_combinations))
-year_combinations$bray <- NA
-for(i in 1:nrow(year_combinations)){
-  subset(x)
-}
-
-```
-
-```{r old nmds code, fig.height = 15, fig.width = 15, eval = FALSE}
-species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
-species_matrix <- species_matrix/100 # convert from percent to proportion
-rownames(species_matrix) <- paste(data$year, data$site_habitat, data$transect)
-rows_to_remove <- rowSums(species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
-species_matrix <- species_matrix[!rows_to_remove,] 
-data_subset <- data[!rows_to_remove,] # make sure rows are also removed from dataframe
-data_subset$stratum <- paste(data_subset$year, "-", data_subset$site, "-", data_subset$habitat )
-# identifying species w/ > 10% cover for plotting later
-# NOTE: Used ChatGPT to generate these next few lines of code but it looked good so I went with it
-columns_to_keep <- apply(species_matrix, 2, function(x) any(x >= 0.1))
-abundant_species_matrix <- as.data.frame(species_matrix)[ , c(TRUE, columns_to_keep)]
-abundant_species_names <- colnames(abundant_species_matrix)
-
-# My computer can't run this at the quadrat level
-set.seed(1032) 
- ord2 <- metaMDS(species_matrix, dist	= "bray",	trymax	= 100,	k	= 2) 
-# ord2 # stress = 0.22
-# stressplot(ord2)
-
-set.seed(1032)
-ord3 <- metaMDS(species_matrix, dist	= "bray",	trymax	= 100,	k	= 3)
-ord3 # 0.166
-stressplot(ord3)
-
-# pro <- procrustes(ord2, ord3) # visually compare w procrustes rotation
-
-# plot(pro, cex=1.5)
-
-site.scores <- as.data.frame(vegan::scores(ord3)$sites)  # extract the site scores and convert to a data.frame
-# add back in metadata
-site.scores$year <- as.factor(data_subset$year)
-site.scores$site <- as.factor(data_subset$site)
-site.scores$habitat <- as.factor(data_subset$habitat)
-site.scores$site_habitat <- as.factor(data_subset$site_habitat)
-site.scores$transect <- as.factor(data_subset$transect)
-site.scores$stratum <- as.factor(data_subset$stratum)
-
-cent <- aggregate(cbind(NMDS1, NMDS2) ~ stratum, data = site.scores, FUN = mean) # compute centroids for NMDS1 and NMDS2
-
-# Need to split stratum up for plotting later
-cent<- 
-  cent %>%
-    separate(stratum, into = c("year", "site", "habitat"), sep = "-")
-cent$site_habitat <- paste(cent$site, cent$habitat)
-# Add back in stratum column
-cent$stratum <-  aggregate(cbind(NMDS1, NMDS2) ~ stratum, data = site.scores, FUN = mean)$stratum
-# A lot of extra spaces
-
-cent <- cent %>%
-  mutate(across(everything(), str_trim))
-cent$year <- as.factor(cent$year)
-cent$habitat <- as.factor(cent$habitat)
-cent$site <- as.factor(cent$site)
-cent$site_habitat <- as.factor(cent$site_habitat)
-cent$NMDS1 <- as.numeric(cent$NMDS1) # read in as characters for some reason
-cent$NMDS2 <- as.numeric(cent$NMDS2)
-species.scores <- as.data.frame(scores(ord3)$species)  # extract the species scores and convert to a data.frame
-species.scores$Species <- rownames(species.scores)
-
-# subset only the species w/ > 10% cover
-abundant.species.scores <- species.scores[species.scores$Species %in% abundant_species_names,]
-
-ggplot() + 
-  geom_point(data= site.scores, aes(x=NMDS1,y=NMDS2, shape = habitat, colour = site),
-             alpha = 0.25, size=2) + # add the point markers
-  geom_text(data= abundant.species.scores,aes(x=NMDS1,y=NMDS2, label = Species), size = 4) +  # add the species labels
-
-  geom_path(data = cent, aes(x = NMDS1, y = NMDS2, group = site_habitat, colour = site),
-            arrow = arrow(length=unit(0.5,"cm"), ends="last", type = "closed"), size = 1) +
-  facet_wrap(~factor(habitat, levels=c('fringing','backreef','outer_10','outer_17'))) +
-  scale_linetype_manual(values = habitat_linetypes) +
-  scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        strip.text.x = element_text(size = 25),
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.key.size = unit(2, "cm"), # increase legend size
-        legend.text = element_text(size = rel(1.5)), # increase legend text
-        legend.title = element_text(size = rel(1.5)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black")) + # add box around legend
-  guides(colour=guide_legend(title.position="top", # make color legend title go above legend
-                                     title.hjust =0.5, nrow = 2),
-         shape=guide_legend(title.position="top",  # make shape legend title go above legend
-                                     title.hjust =0.5,
-                            nrow = 2),
-         path = guide_legend(title.position="top",  # make shape legend title go above legend
-                                     title.hjust =0.5)
-         )
-
-# looks like stochasticity increases from outer_17 --> fringing
-# Even in lagoon though it looks like communities return to initial state... except for north Shore lter_sites
-# ways to test stochasticity vs. determinism?
-# seems like deterministic trajectories, although we're not starting from a clean slate
-# accounting for founder effects?
-
-```
-
-```{r beta dispersion, fig.height = 15, fig.width=20}
-# Want to plot mean distance to centroid, but not sure how to do that with so many factors/interactions...
- bray_matrix <- vegdist(species_matrix, method = "bray")
-# year_beta_disper <- vegan::betadisper(bray_matrix, data_subset$year)
-# year_beta_disper #
-
-# plot of dispersion test
-# plot(year_beta_disper, sub = "") # lol
-# boxplot(year_beta_disper, main = "year", xlab = "") 
-# 
-# site_beta_disper <- vegan::betadisper(bray_matrix, data_subset$site)
-# plot(site_beta_disper, sub = "") # site five stands out
-# boxplot(site_beta_disper, main = "site", xlab = "") 
-# 
-# 
-# habitat_beta_disper <- vegan::betadisper(bray_matrix, data_subset$habitat)
-# plot(habitat_beta_disper, sub = "") 
-# boxplot(habitat_beta_disper, main = "habitat", xlab = "") # greatest variance in fringing it seems (as expected)
-# 
-# site_habitat_beta_disper <- vegan::betadisper(bray_matrix, data_subset$site_habitat)
-# plot(site_habitat_beta_disper, sub = "") 
-# boxplot(site_habitat_beta_disper, main = "site_habitat", xlab = "") 
-# 
-stratum_beta_disper <- vegan::betadisper(bray_matrix, data_subset$stratum)
-# plot(site_habitat_beta_disper, sub = "") 
-boxplot(stratum_beta_disper, main = "site_habitat", xlab = "")
-stratum_distances <- as.data.frame(stratum_beta_disper$distances)
-colnames(stratum_distances)[1] <- "Distance"
-stratum_distances$stratum <- data_subset$stratum
-stratum_distances$transect <- data_subset$transect
-stratum_distances$year <- data_subset$year
-stratum_distances$site <- data_subset$site
-stratum_distances$habitat <- data_subset$habitat
-stratum_distances$site_habitat <- data_subset$site_habitat
-
-beta_disper_means <- stratum_distances %>%
-  dplyr::group_by(year, site, habitat, site_habitat) %>%
-  dplyr::summarise(Distance_mean = mean(Distance), 
-            Distance_se = sd(Distance)/sqrt(n()))
-
-(Distance_plot <-
-  ggplot(data = beta_disper_means) +
-  geom_point(aes(x = year, y = Distance_mean, colour = site, shape = habitat, group = site_habitat), size = 8, stroke = 2) +
-  geom_line(aes(x = year, y = Distance_mean, colour = site, group = site_habitat), size = 1.2) +
-  geom_errorbar(aes(x = year, ymin = Distance_mean - Distance_se , ymax = Distance_mean + Distance_se, colour = site,
-                    group = site_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~factor(habitat, levels=c('fringing','backreef','outer_10','outer_17')))+
-  labs(title = "Distance to centroid\n", y = "Distance\n") +
-  theme_bw() + 
-  theme(panel.grid.major = element_blank(), 
-        panel.grid.minor = element_blank(),
-        panel.background = element_blank(),
-        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
-        axis.line.y = element_line(colour = "black"),
-        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
-        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
-        axis.line.x = element_line(colour = "black"),
-        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
-        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
-        strip.text.x = element_text(size = 32),
-        legend.key.size = unit(2, "cm"), # increase legend size
-        legend.text = element_text(size = rel(2)), # increase legend text
-        legend.title = element_text(size = rel(2.25)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal',
-        legend.box.background = element_rect(colour = "black") # make legend horizontal
-        )
-    )
-# beta dispersion increases as we move from forereef to lagoon, lagoon to shore
-
-```

### 2024-08-20 R/Fish_ord.Rmd
committing fishy things :fish:

---
title: "Fish_Stuff"
author: "Callie Stephenson"
date: "2024-03-28"
output: html_document
---
```{r packages, message=FALSE, echo=FALSE}
library(dplyr)
library(readr)
library(here)
library(fishualize)
library(ggplot2)
library(vegan)
library(mclust)
library(gridExtra)
library(glue)
library(patchwork) 
library(scales)
```

```{r data, message=FALSE, echo=FALSE}
#algae <- read.csv(here("data", "algaeonly_cleaned.csv"))
fish_fun <- read.csv(here("data", "biomass_functional_summary.csv"))
fish <- read.csv(here("data", "biomass_herb_summary.csv"))
alpha <- read.csv(here("data", "algae_alpha_diversity.csv")) %>% 
  filter(Habitat != "Outer 17")
```

Noam Algae Code:
```{r}
data <- read.csv(here::here("data", "allbenthicdata_cleaned_colnames.csv"), stringsAsFactors = F)

not_algae <- c("ascidian", "bare_space", "coral", "coral_rubble", "corallimorpharia", "heteractis_sp", "millepora_platyphylla", "no_data", "sand", "sacrophyton_sp", "shell_debris", "soft_coral", "sponge", "tridacna_sp", "zooxanthellae", "row_sums")

algae <- data[ , -which(names(data) %in% not_algae)]
```

```{r}
data <- read.csv(here::here("data", "algaeonly_cleaned.csv"), stringsAsFactors = F) # replacements don't work when strings are factors
meta_cols <- colnames(data[1:7]) 
species_cols <- colnames(data[8:77])

wide_site_level <- data %>%
  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
  summarize(across(all_of(species_cols), mean, na.rm = TRUE))

wide_site_level_merge <- wide_site_level[wide_site_level$Habitat != "Outer 17", ]
```

```{r plot colors from Noam}
#copied from Noam
site_colours <- c("LTER 1" = "cornflowerblue",
               "LTER 2" = "darkblue",
               "LTER 3" = "yellowgreen",
               "LTER 4" = "forestgreen",
               "LTER 5" = "mediumorchid",
               "LTER 6" = "orange")
island_side_colours <- c("North" = "darkblue",
                         "East" = "mediumorchid",
                         "West" = "orange")
habitat_linetypes <- c("Fringing" = "solid",
                       "Backreef" = "longdash",
                       "Outer 10" = "twodash",
                       "Outer 17" = "dotted")
habitat_shapes <- c("Fringing" = 16,
                     "Backreef" = 17,
                     "Outer 10" = 1,
                     "Outer 17" = 2)
```

### Ordination of Algal Communities with Fish Community as Drivers:

Transform the data and remove rows that are 100% coral cover

```{r}
species.data.transform = wisconsin(sqrt(wide_site_level_merge[c(5:74)]))
non_zero_rows <- rowSums(species.data.transform == 0) != ncol(species.data.transform)
species.data.use <- species.data.transform[non_zero_rows, ]
```

```{r}
pcoa = capscale(species.data.use ~ 1, dist = "bray")
pcoa
summary <- summary(pcoa)
summary$cont
```

```{r}
par(mfrow = c(1,2))
plot(pcoa, type = 'n', xlim = c(-1.5, 1.5))
text(pcoa, "species", col = 'blue', cex = 0.5)
plot(pcoa, type = 'n', xlim = c(-1.5, 1.5))
#text(pcoa, "location", col = 'blue', cex = 0.5)
points(pcoa, col = 'black', bg = 'grey', pch = 21)
```

```{r}
# Step 1: Ensure the 5th column is a vector
colors <- as.factor(wide_site_level_merge[non_zero_rows, 4][[1]])

# Step 2: Create the color palette based on the factor levels
color_palette <- rainbow(length(levels(colors)))

# Step 3: Plot the PCoA with colored points
par(mfrow = c(1,3))

# First plot: species plot (same as before)
plot(pcoa, type = 'n', xlim = c(-1.5, 1.5))
text(pcoa, "species", col = 'blue', cex = 1)

# Second plot: location plot with colored points
plot(pcoa, type = 'n', xlim = c(-1.5, 1.5))
points(pcoa, bg = color_palette[colors], pch = 21)

# Optionally, add a legend
plot(1, type = 'n', xlab = '', ylab = '', xaxt = 'n', yaxt = 'n')
legend("center", legend = levels(colors), col = color_palette, pch = 21, pt.bg = color_palette)
```

```{r}
fish_fun1 <- fish_fun
fish_fun1[is.na(fish_fun1)] <- 0
fish_fun1 <- fish_fun1 %>% 
  mutate(Site = paste0("LTER ",Site))
merged <- left_join(wide_site_level_merge, fish_fun1, by = join_by(Year, Site, Habitat))
merged_clean <- merged[complete.cases(merged[c("Browser", "Brusher", "Cropper", "Concealed.Cropper", "Excavator", "Scraper")]), ]

species.data.transform = wisconsin(sqrt(merged_clean[c(6:75)]))
non_zero_rows <- rowSums(species.data.transform == 0) != ncol(species.data.transform)
species.data.use <- species.data.transform[non_zero_rows, ]
merged_clean_filtered <- merged_clean[non_zero_rows, ]
```

```{r}
cont = capscale(species.data.use ~ Browser + Brusher + Cropper + Concealed.Cropper + Excavator + Scraper, dist = "bray", data = merged_clean_filtered)
```

```{r}
plot(cont)
```

```{r}
summary(cont)$cont
```

```{r}
par(mfrow = c(1, 3))  # 1 row, 3 columns

# Plot 1: Species driving the community composition
plot(cont, display = "species", main = "Species Driving Community Composition", type = "n")
text(cont, display = "species", col = "blue", cex = 1)

# Plot 2: Communities (sites) plotted
plot(cont, display = "sites", main = "Communities", type = "n")
points(cont, display = "sites", col = color_palette[colors], pch = 21)  # Sites colored by the same scheme

# Plot 3: Constrained ordination axes (environmental variables)
plot(cont, display = "bp", main = "Constrained Ordination Axes", cex = 1)
```

```{r}
anova(cont)
anova(cont, by = "margin")
```

# Mantel 
I have no idea what I'm doing here
```{r}
library(vegan)
dist.algae <- dist(scale(x = merged_clean_filtered[,c(5:74)], scale = T, center = T), method = "euclidean")
dist.fish <- dist(scale(x = merged_clean_filtered[,c(75:88)], scale = T, center = T), method = "euclidean")
response_by_explan <- mantel(dist.algae, dist.fish, method = "spearman", permutations = 900, na.rm = TRUE)
response_by_explan
```

### Part 3: Model-Based Clustering

### 2024-08-20 noam_beta_diversity.Rmd
New beta diversity markdown for ordinations and dissimilarity. NOT FINISHED YET.

---
title: "noam_beta_diversity"
author: "Noam Altman-Kurosaki"
date: "2024-08-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(here)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(vegan)
library(glmmTMB)
library(car)
library(effects)
library(DHARMa)
library(stringr)
library(MuMIn)
library(codyn)
library(nlme)
library(mgcv)
library(emmeans)
# library(ecofolio) # possible other method for examining portfolio effects 
```

```{r import and format data}
# starting with all data now because of 0 algae transects/quadrats
data <- read.csv(here::here("data", "allbenthicdata_cleaned_colnames.csv"), stringsAsFactors = F)

not_algae <- c("ascidian", "bare_space", "coral", "coral_rubble", "corallimorpharia", "heteractis_sp", "millepora_platyphylla", "no_data", "sand", "sacrophyton_sp", "shell_debris", "soft_coral", "sponge", "tridacna_sp", "zooxanthellae", "row_sums")

data <- data[ , -which(names(data) %in% not_algae)]

# adding island side - adapted from Julianna's code
data <- data %>% mutate(island_side = case_when(site == "lter_1" ~ "north",
                                 site == "lter_2" ~ "north",
                                 site == "lter_3" ~ "east",
                                 site == "lter_4" ~ "east",
                                 site == "lter_5" ~ "west",
                                 site == "lter_6" ~ "west"))
data$island_side_habitat <- paste(data$island_side, data$habitat)

str(data)

meta_cols <- c(colnames(data[1:7] ), "island_side", "island_side_habitat", "stratum")
species_cols <- colnames(data)[!colnames(data) %in% meta_cols]

meta_df <- data[ , which(names(data) %in% meta_cols)]

site_colours <- c("lter_1" = "cornflowerblue",
                  "lter_2" = "darkblue",
                  "lter_3" = "yellowgreen",
                  "lter_4" = "forestgreen",
                  "lter_5" = "mediumorchid",
                  "lter_6" = "orange")
island_side_colours <- c("north" = "darkblue",
                         "east" = "mediumorchid",
                         "west" = "orange")
habitat_linetypes <- c("fringing" = "solid",
                       "backreef" = "longdash",
                       "outer_10" = "twodash",
                       "outer_17" = "dotted")
habitat_shapes <- c("fringing" = 16,
                     "backreef" = 17,
                     "outer_10" = 1,
                     "outer_17" = 2)

#pivot long 
long_data <- pivot_longer(
  data = data,
  cols = 8:77,
  names_to = "taxa",
  values_to = "percent_cover")

 transect_level <- long_data %>%
   dplyr::group_by(year, island_side, site, habitat, transect, taxa) %>%
   dplyr::summarise(mean_cover = mean(percent_cover), 
             SE = sd(percent_cover)/sqrt(n()))

site_level <- long_data %>%
  dplyr::group_by(year, island_side, site, habitat, site_habitat, taxa) %>%
  dplyr::summarise(mean_site_cover = mean(percent_cover), 
            site_se = sd(percent_cover)/sqrt(n()))

# subset based on habitat
fringing <- subset(data, habitat == "fringing")
backreef <- subset(data, habitat == "backreef")
outer10 <- subset(data, habitat == "outer_10")
outer17 <- subset(data, habitat == "outer_17")

```

```{r fringing bray curtis}
fringing_species_matrix <- as.matrix(fringing[ , -which(names(fringing) %in% meta_cols)])
fringing_species_matrix <- fringing_species_matrix/100 # convert from percent to proportion
rownames(fringing_species_matrix) <- paste(fringing$year,"-", fringing$location, sep = "")
fringing_rows_to_remove <- rowSums(fringing_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
fringing_species_matrix <- fringing_species_matrix[!fringing_rows_to_remove,] 
fringing_species_matrix <- asin(sqrt(fringing_species_matrix)) # arcsin squareroot transform
fringing_subset <- fringing[!fringing_rows_to_remove,] # make sure rows are also removed from dataframe
fringing_subset$stratum <- paste(fringing_subset$year,"-", fringing_subset$location, sep = "") # for merging with bray matrix later
brayfringing <- vegdist(fringing_species_matrix, "bray") # create bray dissimilarity matrix
# View(brayfringing)
fringing_braymat <- as.matrix(brayfringing) # conver to matrix object
# convert to pairwise comparison
fringing_comp <- fringing_braymat[lower.tri(fringing_braymat)] # extract lower triangle
fringing_row_indices <- row(fringing_braymat)[lower.tri(fringing_braymat)] # extract row indices for first quad in comparison
fringing_col_indices <- col(fringing_braymat)[lower.tri(fringing_braymat)] # extract col indices = second quad in comp
fringing_loc1 <- rownames(fringing_braymat)[fringing_row_indices] # populate appropriate vector
fringing_loc2 <- colnames(fringing_braymat)[fringing_col_indices] # populate appropriate vector
fringing_pairwise_df <- data.frame(loc1 = fringing_loc1, loc2 = fringing_loc2, dissimilarity = fringing_comp)
fringing_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", fringing_pairwise_df$loc1)) # extract timepoint from first quad in comp
fringing_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", fringing_pairwise_df$loc2)) # repeat for second quad
fringing_pairwise_df$time <- abs(fringing_pairwise_df$year2 - fringing_pairwise_df$year1) # extract length of differences
fringing_pairwise_df$location_1 <- sub(".*-(.*)", "\\1", fringing_pairwise_df$loc1) # extract quadrat id from first quad in comp
fringing_pairwise_df$location_2 <- sub(".*-(.*)", "\\1", fringing_pairwise_df$loc2) # repeat for second quad

```

```{r backreef bray curtis}
backreef_species_matrix <- as.matrix(backreef[ , -which(names(backreef) %in% meta_cols)])
backreef_species_matrix <- backreef_species_matrix/100 # convert from percent to proportion
rownames(backreef_species_matrix) <- paste(backreef$year,"-", backreef$location, sep = "")
backreef_rows_to_remove <- rowSums(backreef_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
backreef_species_matrix <- backreef_species_matrix[!backreef_rows_to_remove,] 
backreef_species_matrix <- asin(sqrt(backreef_species_matrix)) # arcsin squareroot transform
backreef_subset <- backreef[!backreef_rows_to_remove,] # make sure rows are also removed from dataframe
backreef_subset$stratum <- paste(backreef_subset$year,"-", backreef_subset$location, sep = "") # for merging with bray matrix later
braybackreef <- vegdist(backreef_species_matrix, "bray") # create bray dissimilarity matrix
# View(braybackreef)
backreef_braymat <- as.matrix(braybackreef) # conver to matrix object
# convert to pairwise comparison
backreef_comp <- backreef_braymat[lower.tri(backreef_braymat)] # extract lower triangle
backreef_row_indices <- row(backreef_braymat)[lower.tri(backreef_braymat)] # extract row indices for first quad in comparison
backreef_col_indices <- col(backreef_braymat)[lower.tri(backreef_braymat)] # extract col indices = second quad in comp
backreef_loc1 <- rownames(backreef_braymat)[backreef_row_indices] # populate appropriate vector
backreef_loc2 <- colnames(backreef_braymat)[backreef_col_indices] # populate appropriate vector
backreef_pairwise_df <- data.frame(loc1 = backreef_loc1, loc2 = backreef_loc2, dissimilarity = backreef_comp)
backreef_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", backreef_pairwise_df$loc1)) # extract timepoint from first quad in comp
backreef_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", backreef_pairwise_df$loc2)) # repeat for second quad
backreef_pairwise_df$time <- abs(backreef_pairwise_df$year2 - backreef_pairwise_df$year1) # extract length of differences
backreef_pairwise_df$location_1 <- sub(".*-(.*)", "\\1", backreef_pairwise_df$loc1) # extract quadrat id from first quad in comp
backreef_pairwise_df$location_2 <- sub(".*-(.*)", "\\1", backreef_pairwise_df$loc2) # repeat for second quad

```

```{r outer10 bray curtis}
outer10_species_matrix <- as.matrix(outer10[ , -which(names(outer10) %in% meta_cols)])
outer10_species_matrix <- outer10_species_matrix/100 # convert from percent to proportion
rownames(outer10_species_matrix) <- paste(outer10$year,"-", outer10$location, sep = "")
outer10_rows_to_remove <- rowSums(outer10_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
outer10_species_matrix <- outer10_species_matrix[!outer10_rows_to_remove,] 
outer10_species_matrix <- asin(sqrt(outer10_species_matrix)) # arcsin squareroot transform
outer10_subset <- outer10[!outer10_rows_to_remove,] # make sure rows are also removed from dataframe
outer10_subset$stratum <- paste(outer10_subset$year,"-", outer10_subset$location, sep = "") # for merging with bray matrix later
brayouter10 <- vegdist(outer10_species_matrix, "bray") # create bray dissimilarity matrix
# View(brayouter10)
outer10_braymat <- as.matrix(brayouter10) # conver to matrix object
# convert to pairwise comparison
outer10_comp <- outer10_braymat[lower.tri(outer10_braymat)] # extract lower triangle
outer10_row_indices <- row(outer10_braymat)[lower.tri(outer10_braymat)] # extract row indices for first quad in comparison
outer10_col_indices <- col(outer10_braymat)[lower.tri(outer10_braymat)] # extract col indices = second quad in comp
outer10_loc1 <- rownames(outer10_braymat)[outer10_row_indices] # populate appropriate vector
outer10_loc2 <- colnames(outer10_braymat)[outer10_col_indices] # populate appropriate vector
outer10_pairwise_df <- data.frame(loc1 = outer10_loc1, loc2 = outer10_loc2, dissimilarity = outer10_comp)
outer10_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer10_pairwise_df$loc1)) # extract timepoint from first quad in comp
outer10_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer10_pairwise_df$loc2)) # repeat for second quad
outer10_pairwise_df$time <- abs(outer10_pairwise_df$year2 - outer10_pairwise_df$year1) # extract length of differences
outer10_pairwise_df$location_1 <- sub(".*-(.*)", "\\1", outer10_pairwise_df$loc1) # extract quadrat id from first quad in comp
outer10_pairwise_df$location_2 <- sub(".*-(.*)", "\\1", outer10_pairwise_df$loc2) # repeat for second quad

```

```{r outer17 bray curtis}
outer17_species_matrix <- as.matrix(outer17[ , -which(names(outer17) %in% meta_cols)])
outer17_species_matrix <- outer17_species_matrix/100 # convert from percent to proportion
rownames(outer17_species_matrix) <- paste(outer17$year,"-", outer17$location, sep = "")
outer17_rows_to_remove <- rowSums(outer17_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
outer17_species_matrix <- outer17_species_matrix[!outer17_rows_to_remove,] 
outer17_species_matrix <- asin(sqrt(outer17_species_matrix)) # arcsin squareroot transform
outer17_subset <- outer17[!outer17_rows_to_remove,] # make sure rows are also removed from dataframe
outer17_subset$stratum <- paste(outer17_subset$year,"-", outer17_subset$location, sep = "") # for merging with bray matrix later
brayouter17 <- vegdist(outer17_species_matrix, "bray") # create bray dissimilarity matrix
# View(brayouter17)
outer17_braymat <- as.matrix(brayouter17) # conver to matrix object
# convert to pairwise comparison
outer17_comp <- outer17_braymat[lower.tri(outer17_braymat)] # extract lower triangle
outer17_row_indices <- row(outer17_braymat)[lower.tri(outer17_braymat)] # extract row indices for first quad in comparison
outer17_col_indices <- col(outer17_braymat)[lower.tri(outer17_braymat)] # extract col indices = second quad in comp
outer17_loc1 <- rownames(outer17_braymat)[outer17_row_indices] # populate appropriate vector
outer17_loc2 <- colnames(outer17_braymat)[outer17_col_indices] # populate appropriate vector
outer17_pairwise_df <- data.frame(loc1 = outer17_loc1, loc2 = outer17_loc2, dissimilarity = outer17_comp)
outer17_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer17_pairwise_df$loc1)) # extract timepoint from first quad in comp
outer17_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer17_pairwise_df$loc2)) # repeat for second quad
outer17_pairwise_df$time <- abs(outer17_pairwise_df$year2 - outer17_pairwise_df$year1) # extract length of differences
outer17_pairwise_df$location_1 <- sub(".*-(.*)", "\\1", outer17_pairwise_df$loc1) # extract quadrat id from first quad in comp
outer17_pairwise_df$location_2 <- sub(".*-(.*)", "\\1", outer17_pairwise_df$loc2) # repeat for second quad

```

```{r bray dissimilarity over time, fig.height=15, fig.width=15}
# for the purposes of this first set of analyses, I'm interested in dissimilarity of a given quad over time
# not necessarily interested in how quads compare to each other over time/at each timepoint
fringing_pairwise_subset <- fringing_pairwise_df[fringing_pairwise_df$location_1 == fringing_pairwise_df$location_2,]
backreef_pairwise_subset <- backreef_pairwise_df[backreef_pairwise_df$location_1 == backreef_pairwise_df$location_2,]
outer10_pairwise_subset <- outer10_pairwise_df[outer10_pairwise_df$location_1 == outer10_pairwise_df$location_2,]
outer17_pairwise_subset <- outer17_pairwise_df[outer17_pairwise_df$location_1 == outer17_pairwise_df$location_2,]

comb_subset <- rbind(fringing_pairwise_subset,
                     backreef_pairwise_subset,
                     outer10_pairwise_subset,
                     outer17_pairwise_subset)
# location_1 and location_2 columns are all the same now
# create new col just called "location" for merge with other metadata
comb_subset$location <- comb_subset$location_1
bray_time_series <- merge(meta_df, comb_subset, by = "location")

# model

# some quadrats have dissimilarity = 0 or 1

# Smithson verkuilen transformation for beta-regression
sv_trans = function(prop, N, s = 0.000005){
  (prop*(nrow(N) - 1) + s)/nrow(N)
  # where prop is the proportional value you're transforming, 
  # N is the sample size, which is specified by taking the number of rows/observations from a given dataframe,
  # and s is a small offset 
}

# transformed dissimilarities
bray_time_series$dissimilarity_trans <- sv_trans(bray_time_series$dissimilarity, bray_time_series)

# bray_mod_universal <- glmmTMB(dissimilarity_trans ~ time*habitat*island_side + (1 | island_side/site/transect/location),
#                               na.action = na.pass, family = beta_family(link = "logit"), data = bray_time_series) #
# summary(bray_mod_universal) # a fair bit of variance w/i quads/transects
# 
# bray_mod_reduced <- glmmTMB(dissimilarity_trans ~ time*habitat + time*island_side + (1 | island_side/site/transect/location),
#                               na.action = na.pass, family = beta_family(link = "logit"), data = bray_time_series)

```

```{r dissimilarity over time plot, fig.height=15, fig.width=15}
 ggplot(bray_time_series,
        aes(x = time, y = dissimilarity, colour = island_side) )+
   facet_wrap(~habitat) + 
   geom_point() +
   scale_colour_manual(values = island_side_colours) +
   stat_smooth(se = T, linewidth = 2) +
   labs(y = "Dissimilarity\n") +
   theme_bw() + 
   theme(panel.grid.major = element_blank(), 
         panel.grid.minor = element_blank(),
         panel.background = element_blank(),
         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
         axis.line.y = element_line(colour = "black"),
         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         axis.line.x = element_line(colour = "black"),
         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         strip.text.x = element_text(size = 32),
         legend.key.size = unit(2, "cm"), # increase legend size
         legend.text = element_text(size = rel(2)), # increase legend text
         legend.title = element_text(size = rel(2.25)),
         legend.position = 'bottom', # move legend to bottom
         legend.box = 'horizontal',
         legend.box.background = element_rect(colour = "black") # make legend horizontal
         )

```

```{r fringing nmds set up}
# identifying species w/ > 10% cover for plotting later

#### FRINGING
# NOTE: Used ChatGPT to generate these next few lines of code but it looked good so I went with it
# fringing_columns_to_keep <- apply(fringing_species_matrix, 2, function(x) any(x >= 0.1))
# fringing_abundant_species_matrix <- as.data.frame(fringing_species_matrix)[ , fringing_columns_to_keep]
# fringing_abundant_species_names <- colnames(fringing_abundant_species_matrix)
# 
# fringing_subset_meta <- fringing_subset[ , which(names(fringing_subset) %in% meta_cols)]
# 
# # for mean centroid:
# fringing_subset_meta$island_side_year <- paste(fringing_subset_meta$year, "-", fringing_subset_meta$island_side_habitat, sep = "")
# 
# # takes several hours to run - do so at your own peril
# set.seed(1032) 
# # ord3 <- metaMDS(fringing_braymat, dist	= "bray", k = 3, trymax = 100) 
# fringing_ord3 # stress = 0.018
# 
# 
# sppscores(fringing_ord3) <- fringing_species_matrix # add back in species scores
# fringing.species.scores <- as.data.frame(scores(fringing_ord3, "species")) # some NaN for absent species
# fringing.species.scores <- na.omit(fringing.species.scores)
# fringing.species.scores$species <- rownames(fringing.species.scores) 
# 
# 
# fringing.site.scores <- as.data.frame(vegan::scores(fringing_ord3, "sites"))  # extract the site scores and convert to a data.frame
# # add back in metadata
# fringing.site.scores$stratum <- rownames(fringing.site.scores)
# fringing.site.scores <- merge(fringing_subset_meta, fringing.site.scores, by = "stratum")
# 
# 
# 
# 
# write_csv(site.scores, here("data", "fringing_site_scores.csv"))
# write_csv(species.scores, here("data", "fringing_species_scores.csv"))

fringing.species.scores <- read.csv(here::here("data", "fringing_species_scores.csv"), stringsAsFactors = F) 
fringing.site.scores <- read.csv(here::here("data", "fringing_site_scores.csv"), stringsAsFactors = F) 
# fringing.cent <- read.csv(here::here("data", "fringing_mean_centroid.csv"), stringsAsFactors = F)

fringing.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = fringing.site.scores, FUN = mean) 
# fringing.cent.meta <- fringing_subset_meta %>%
#    dplyr::group_by(year, island_side, island_side_habitat, island_side_year, site, habitat, transect, quadrat) %>%
#    dplyr::summarise(year = unique(year), 
#                     island_side = unique(island_side),
#                     island_side_habitat = unique(island_side_habitat),
#                     island_side_year = unique(island_side_year),
#                     site = unique(site),
#                     habitat = unique(habitat),
#                     transect = unique(transect),
#                     quadrat = unique(quadrat))
# fringing.cent <- merge(fringing.cent, fringing_subset_meta[,-which(names(fringing_subset_meta) %in% c("stratum"))],  by = "island_side_year")

# subset only the species w/ > 10% cover
fringing.abundant.species.scores <- fringing.species.scores[fringing.species.scores$species %in% fringing_abundant_species_names,]

#############
# BACKREEF #
############
backreef_columns_to_keep <- apply(backreef_species_matrix, 2, function(x) any(x >= 0.1))
backreef_abundant_species_matrix <- as.data.frame(backreef_species_matrix)[ , backreef_columns_to_keep]
backreef_abundant_species_names <- colnames(backreef_abundant_species_matrix)

backreef_subset_meta <- backreef_subset[ , which(names(backreef_subset) %in% meta_cols)]

# for mean centroid:
backreef_subset_meta$island_side_year <- paste(backreef_subset_meta$year, "-", backreef_subset_meta$island_side_habitat, sep = "")

# takes several hours to run - do so at your own peril
set.seed(1032) 
# ord3 <- metaMDS(backreef_braymat, dist	= "bray", k = 3, trymax = 100) 
backreef_ord3 # stress = 0.018

sppscores(backreef_ord3) <- backreef_species_matrix # add back in species scores
backreef.species.scores <- as.data.frame(scores(backreef_ord3, "species")) # some NaN for absent species
backreef.species.scores <- na.omit(backreef.species.scores)
backreef.species.scores$species <- rownames(backreef.species.scores)

backreef.site.scores <- as.data.frame(vegan::scores(backreef_ord3, "sites"))  # extract the site scores and convert to a data.frame
# add back in metadata
backreef.site.scores$stratum <- rownames(backreef.site.scores)
backreef.site.scores <- merge(backreef_subset_meta, backreef.site.scores, by = "stratum")

write_csv(site.scores, here("data", "backreef_site_scores.csv"))
write_csv(species.scores, here("data", "backreef_species_scores.csv"))

backreef.species.scores <- read.csv(here::here("data", "backreef_species_scores.csv"), stringsAsFactors = F) 
backreef.site.scores <- read.csv(here::here("data", "backreef_site_scores.csv"), stringsAsFactors = F) 
# backreef.cent <- read.csv(here::here("data", "backreef_mean_centroid.csv"), stringsAsFactors = F)

backreef.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = backreef.site.scores, FUN = mean) 
# backreef.cent.meta <- backreef_subset_meta %>%
#    dplyr::group_by(year, island_side, island_side_habitat, island_side_year, site, habitat, transect, quadrat) %>%
#    dplyr::summarise(year = unique(year), 
#                     island_side = unique(island_side),
#                     island_side_habitat = unique(island_side_habitat),
#                     island_side_year = unique(island_side_year),
#                     site = unique(site),
#                     habitat = unique(habitat),
#                     transect = unique(transect),
#                     quadrat = unique(quadrat))
# backreef.cent <- merge(backreef.cent, backreef_subset_meta[,-which(names(backreef_subset_meta) %in% c("stratum"))],  by = "island_side_year")

# subset only the species w/ > 10% cover
backreef.abundant.species.scores <- backreef.species.scores[backreef.species.scores$species %in% backreef_abundant_species_names,]

#######
# OUTER10
#######

outer10_columns_to_keep <- apply(outer10_species_matrix, 2, function(x) any(x >= 0.1))
outer10_abundant_species_matrix <- as.data.frame(outer10_species_matrix)[ , outer10_columns_to_keep]
outer10_abundant_species_names <- colnames(outer10_abundant_species_matrix)

outer10_subset_meta <- outer10_subset[ , which(names(outer10_subset) %in% meta_cols)]

# for mean centroid:
outer10_subset_meta$island_side_year <- paste(outer10_subset_meta$year, "-", outer10_subset_meta$island_side_habitat, sep = "")

# takes several hours to run - do so at your own peril
set.seed(1032) 
# ord3 <- metaMDS(outer10_braymat, dist	= "bray", k = 3, trymax = 100) 
outer10_ord3 # stress = 0.018

sppscores(outer10_ord3) <- outer10_species_matrix # add back in species scores
outer10.species.scores <- as.data.frame(scores(outer10_ord3, "species")) # some NaN for absent species
outer10.species.scores <- na.omit(outer10.species.scores)
outer10.species.scores$species <- rownames(outer10.species.scores)

outer10.site.scores <- as.data.frame(vegan::scores(outer10_ord3, "sites"))  # extract the site scores and convert to a data.frame
# add back in metadata
outer10.site.scores$stratum <- rownames(outer10.site.scores)
outer10.site.scores <- merge(outer10_subset_meta, outer10.site.scores, by = "stratum")

write_csv(site.scores, here("data", "outer10_site_scores.csv"))
write_csv(species.scores, here("data", "outer10_species_scores.csv"))

outer10.species.scores <- read.csv(here::here("data", "outer10_species_scores.csv"), stringsAsFactors = F) 
outer10.site.scores <- read.csv(here::here("data", "outer10_site_scores.csv"), stringsAsFactors = F) 
# outer10.cent <- read.csv(here::here("data", "outer10_mean_centroid.csv"), stringsAsFactors = F)

outer10.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = outer10.site.scores, FUN = mean) 
# outer10.cent.meta <- outer10_subset_meta %>%
#    dplyr::group_by(year, island_side, island_side_habitat, island_side_year, site, habitat, transect, quadrat) %>%
#    dplyr::summarise(year = unique(year), 
#                     island_side = unique(island_side),
#                     island_side_habitat = unique(island_side_habitat),
#                     island_side_year = unique(island_side_year),
#                     site = unique(site),
#                     habitat = unique(habitat),
#                     transect = unique(transect),
#                     quadrat = unique(quadrat))
# outer10.cent <- merge(outer10.cent, outer10_subset_meta[,-which(names(outer10_subset_meta) %in% c("stratum"))],  by = "island_side_year")

# subset only the species w/ > 10% cover
outer10.abundant.species.scores <- outer10.species.scores[outer10.species.scores$species %in% outer10_abundant_species_names,]

#############
# OUTER 17 ##
#############

outer17_columns_to_keep <- apply(outer17_species_matrix, 2, function(x) any(x >= 0.1))
outer17_abundant_species_matrix <- as.data.frame(outer17_species_matrix)[ , outer17_columns_to_keep]
outer17_abundant_species_names <- colnames(outer17_abundant_species_matrix)

outer17_subset_meta <- outer17_subset[ , which(names(outer17_subset) %in% meta_cols)]

# for mean centroid:
outer17_subset_meta$island_side_year <- paste(outer17_subset_meta$year, "-", outer17_subset_meta$island_side_habitat, sep = "")

# takes several hours to run - do so at your own peril
set.seed(1032) 
# ord3 <- metaMDS(outer17_braymat, dist	= "bray", k = 3, trymax = 100) 
outer17_ord3 # stress = 0.018

sppscores(outer17_ord3) <- outer17_species_matrix # add back in species scores
outer17.species.scores <- as.data.frame(scores(outer17_ord3, "species")) # some NaN for absent species
outer17.species.scores <- na.omit(outer17.species.scores)
outer17.species.scores$species <- rownames(outer17.species.scores)

outer17.site.scores <- as.data.frame(vegan::scores(outer17_ord3, "sites"))  # extract the site scores and convert to a data.frame
# add back in metadata
outer17.site.scores$stratum <- rownames(outer17.site.scores)
outer17.site.scores <- merge(outer17_subset_meta, outer17.site.scores, by = "stratum")

write_csv(site.scores, here("data", "outer17_site_scores.csv"))
write_csv(species.scores, here("data", "outer17_species_scores.csv"))

outer17.species.scores <- read.csv(here::here("data", "outer17_species_scores.csv"), stringsAsFactors = F) 
outer17.site.scores <- read.csv(here::here("data", "outer17_site_scores.csv"), stringsAsFactors = F) 
# outer17.cent <- read.csv(here::here("data", "outer17_mean_centroid.csv"), stringsAsFactors = F)

outer17.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = outer17.site.scores, FUN = mean) 
# outer17.cent.meta <- outer17_subset_meta %>%
#    dplyr::group_by(year, island_side, island_side_habitat, island_side_year, site, habitat, transect, quadrat) %>%
#    dplyr::summarise(year = unique(year), 
#                     island_side = unique(island_side),
#                     island_side_habitat = unique(island_side_habitat),
#                     island_side_year = unique(island_side_year),
#                     site = unique(site),
#                     habitat = unique(habitat),
#                     transect = unique(transect),
#                     quadrat = unique(quadrat))
# outer17.cent <- merge(outer17.cent, outer17_subset_meta[,-which(names(outer17_subset_meta) %in% c("stratum"))],  by = "island_side_year")

# subset only the species w/ > 10% cover
outer17.abundant.species.scores <- outer17.species.scores[outer17.species.scores$species %in% outer17_abundant_species_names,]
```

```{r plot fringing}
ggplot() + 
  geom_point(data= site.scores, aes(x=NMDS1,y=NMDS2, colour = island_side),
             alpha = 0.25, size=2) + # add the point markers
  geom_text(data= abundant.species.scores,aes(x=NMDS1,y=NMDS2, label = species), size = 4) +  # add the species labels

  geom_path(data = cent, aes(x = NMDS1, y = NMDS2, group = island_side_habitat, colour = island_side),
            arrow = arrow(length=unit(0.5,"cm"), ends="last", type = "closed"), size = 1) +
#  facet_wrap(~factor(habitat, levels=c('fringing','backreef','outer_10','outer_17'))) +
  scale_linetype_manual(values = habitat_linetypes) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        strip.text.x = element_text(size = 25),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(1.5)), # increase legend text
        legend.title = element_text(size = rel(1.5)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) + # add box around legend
  guides(colour=guide_legend(title.position="top", # make color legend title go above legend
                                     title.hjust =0.5, nrow = 2),
         shape=guide_legend(title.position="top",  # make shape legend title go above legend
                                     title.hjust =0.5,
                            nrow = 2),
         path = guide_legend(title.position="top",  # make shape legend title go above legend
                                     title.hjust =0.5)
         )

```

```{r full bray curtis set up, eval = FALSE}
### I DON'T THINK ANY COMPUTER CAN RUN THIS
species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
species_matrix <- species_matrix/100 # convert from percent to proportion
rownames(species_matrix) <- paste(data$year,"-", data$location, sep = "")
rows_to_remove <- rowSums(species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
species_matrix <- species_matrix[!rows_to_remove,] 
data_subset <- data[!rows_to_remove,] # make sure rows are also removed from dataframe
data_subset$stratum <- paste(data_subset$year,"-", data_subset$location, sep = "") # for merging with bray matrix later
brayfull <- vegdist(species_matrix, "bray") # create bray dissimilarity matrix
# View(brayfull)
braymat <- as.matrix(brayfull) # conver to matrix object
# convert to pairwise comparison
comp <- braymat[lower.tri(braymat)] # extract lower triangle
row_indices <- row(braymat)[lower.tri(braymat)] # extract row indices for first quad in comparison
col_indices <- col(braymat)[lower.tri(braymat)] # 
loc1 <- rownames(braymat)[row_indices]
loc2 <- colnames(braymat)[col_indices]
pairwise_df <- data.frame(loc1 = loc1, loc2 = loc2, dissimilarity = comp)
# pairwise_df$year1 <- sub("(.*)-.*", "\\1", pairwise_df$loc1)
year1 <- strsplit(pairwise_df$loc1, "-")[[1]]

```

```{r old nmds code, fig.height = 15, fig.width = 15, eval = FALSE}
species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
species_matrix <- species_matrix/100 # convert from percent to proportion
rownames(species_matrix) <- paste(data$year, data$site_habitat, data$transect)
rows_to_remove <- rowSums(species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
species_matrix <- species_matrix[!rows_to_remove,] 
data_subset <- data[!rows_to_remove,] # make sure rows are also removed from dataframe
data_subset$stratum <- paste(data_subset$year, "-", data_subset$site, "-", data_subset$habitat )
# identifying species w/ > 10% cover for plotting later
# NOTE: Used ChatGPT to generate these next few lines of code but it looked good so I went with it
columns_to_keep <- apply(species_matrix, 2, function(x) any(x >= 0.1))
abundant_species_matrix <- as.data.frame(species_matrix)[ , c(TRUE, columns_to_keep)]
abundant_species_names <- colnames(abundant_species_matrix)

# My computer can't run this at the quadrat level

# ord2 # stress = 0.22
# stressplot(ord2)

set.seed(1032)
ord3 <- metaMDS(species_matrix, dist	= "bray",	trymax	= 100,	k	= 3)
ord3 # 0.166
stressplot(ord3)

# pro <- procrustes(ord2, ord3) # visually compare w procrustes rotation

# plot(pro, cex=1.5)

site.scores <- as.data.frame(vegan::scores(ord3)$sites)  # extract the site scores and convert to a data.frame
# add back in metadata
site.scores$year <- as.factor(data_subset$year)
site.scores$site <- as.factor(data_subset$site)
site.scores$habitat <- as.factor(data_subset$habitat)
site.scores$site_habitat <- as.factor(data_subset$site_habitat)
site.scores$transect <- as.factor(data_subset$transect)
site.scores$stratum <- as.factor(data_subset$stratum)

cent <- aggregate(cbind(NMDS1, NMDS2) ~ stratum, data = site.scores, FUN = mean) # compute centroids for NMDS1 and NMDS2

# Need to split stratum up for plotting later
cent<- 
  cent %>%
    separate(stratum, into = c("year", "site", "habitat"), sep = "-")
cent$site_habitat <- paste(cent$site, cent$habitat)
# Add back in stratum column
cent$stratum <-  aggregate(cbind(NMDS1, NMDS2) ~ stratum, data = site.scores, FUN = mean)$stratum
# A lot of extra spaces

cent <- cent %>%
  mutate(across(everything(), str_trim))
cent$year <- as.factor(cent$year)
cent$habitat <- as.factor(cent$habitat)
cent$site <- as.factor(cent$site)
cent$site_habitat <- as.factor(cent$site_habitat)
cent$NMDS1 <- as.numeric(cent$NMDS1) # read in as characters for some reason
cent$NMDS2 <- as.numeric(cent$NMDS2)
species.scores <- as.data.frame(scores(ord3)$species)  # extract the species scores and convert to a data.frame
species.scores$Species <- rownames(species.scores)

# subset only the species w/ > 10% cover
abundant.species.scores <- species.scores[species.scores$Species %in% abundant_species_names,]

ggplot() + 
  geom_point(data= site.scores, aes(x=NMDS1,y=NMDS2, shape = habitat, colour = site),
             alpha = 0.25, size=2) + # add the point markers
  geom_text(data= abundant.species.scores,aes(x=NMDS1,y=NMDS2, label = Species), size = 4) +  # add the species labels

  geom_path(data = cent, aes(x = NMDS1, y = NMDS2, group = site_habitat, colour = site),
            arrow = arrow(length=unit(0.5,"cm"), ends="last", type = "closed"), size = 1) +
  facet_wrap(~factor(habitat, levels=c('fringing','backreef','outer_10','outer_17'))) +
  scale_linetype_manual(values = habitat_linetypes) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        strip.text.x = element_text(size = 25),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(1.5)), # increase legend text
        legend.title = element_text(size = rel(1.5)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) + # add box around legend
  guides(colour=guide_legend(title.position="top", # make color legend title go above legend
                                     title.hjust =0.5, nrow = 2),
         shape=guide_legend(title.position="top",  # make shape legend title go above legend
                                     title.hjust =0.5,
                            nrow = 2),
         path = guide_legend(title.position="top",  # make shape legend title go above legend
                                     title.hjust =0.5)
         )

# looks like stochasticity increases from outer_17 --> fringing
# Even in lagoon though it looks like communities return to initial state... except for north Shore lter_sites
# ways to test stochasticity vs. determinism?
# seems like deterministic trajectories, although we're not starting from a clean slate
# accounting for founder effects?

```

```{r beta dispersion, fig.height = 15, fig.width=20}
# Want to plot mean distance to centroid, but not sure how to do that with so many factors/interactions...
 bray_matrix <- vegdist(species_matrix, method = "bray")
# year_beta_disper <- vegan::betadisper(bray_matrix, data_subset$year)
# year_beta_disper #

# plot of dispersion test
# plot(year_beta_disper, sub = "") # lol
# boxplot(year_beta_disper, main = "year", xlab = "") 
# 
# site_beta_disper <- vegan::betadisper(bray_matrix, data_subset$site)
# plot(site_beta_disper, sub = "") # site five stands out
# boxplot(site_beta_disper, main = "site", xlab = "") 
# 
# 
# habitat_beta_disper <- vegan::betadisper(bray_matrix, data_subset$habitat)
# plot(habitat_beta_disper, sub = "") 
# boxplot(habitat_beta_disper, main = "habitat", xlab = "") # greatest variance in fringing it seems (as expected)
# 
# site_habitat_beta_disper <- vegan::betadisper(bray_matrix, data_subset$site_habitat)
# plot(site_habitat_beta_disper, sub = "") 
# boxplot(site_habitat_beta_disper, main = "site_habitat", xlab = "") 
# 
stratum_beta_disper <- vegan::betadisper(bray_matrix, data_subset$stratum)
# plot(site_habitat_beta_disper, sub = "") 
boxplot(stratum_beta_disper, main = "site_habitat", xlab = "")
stratum_distances <- as.data.frame(stratum_beta_disper$distances)
colnames(stratum_distances)[1] <- "Distance"
stratum_distances$stratum <- data_subset$stratum
stratum_distances$transect <- data_subset$transect
stratum_distances$year <- data_subset$year
stratum_distances$site <- data_subset$site
stratum_distances$habitat <- data_subset$habitat
stratum_distances$site_habitat <- data_subset$site_habitat

beta_disper_means <- stratum_distances %>%
  dplyr::group_by(year, site, habitat, site_habitat) %>%
  dplyr::summarise(Distance_mean = mean(Distance), 
            Distance_se = sd(Distance)/sqrt(n()))

(Distance_plot <-
  ggplot(data = beta_disper_means) +
  geom_point(aes(x = year, y = Distance_mean, colour = site, shape = habitat, group = site_habitat), size = 8, stroke = 2) +
  geom_line(aes(x = year, y = Distance_mean, colour = site, group = site_habitat), size = 1.2) +
  geom_errorbar(aes(x = year, ymin = Distance_mean - Distance_se , ymax = Distance_mean + Distance_se, colour = site,
                    group = site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~factor(habitat, levels=c('fringing','backreef','outer_10','outer_17')))+
  labs(title = "Distance to centroid\n", y = "Distance\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )
# beta dispersion increases as we move from forereef to lagoon, lagoon to shore

```

### 2024-08-20 R/Fish_ord.Rmd
ordinated with overall herb biomass and functional biomass but nothing explains much :fish:

@@
 fish_fun <- read.csv(here("data", "biomass_functional_summary.csv"))
-fish <- read.csv(here("data", "biomass_herb_summary.csv"))
+herb <- read.csv(here("data", "biomass_herb_summary.csv"))
 alpha <- read.csv(here("data", "algae_alpha_diversity.csv")) %>% 
@@
 
-### Ordination of Algal Communities with Fish Community as Drivers:
+# Ordination of Algal Communities with Fish Community as Drivers:
 
@@
 
+## Just Herbivore Biomass:
+```{r}
+herb_merge <- herb %>% 
+  mutate(Site = paste0("LTER ",Site))
+herb_algal_merge <- left_join(wide_site_level_merge, herb_merge[,c("Year", "Site", "Habitat", "Herbivore.Biomass")], 
+                              by = join_by(Year, Site, Habitat)) %>% 
+  filter(Year != "2023")
+
+species.data.transform = wisconsin(sqrt(herb_algal_merge[c(5:74)]))
+non_zero_rows <- rowSums(species.data.transform == 0) != ncol(species.data.transform)
+species.data.use <- species.data.transform[non_zero_rows, ]
+merged_clean_filtered <- herb_algal_merge[non_zero_rows, ]
+```
+
+```{r}
+herb_algal_constrained_ord = capscale(species.data.use ~ Herbivore.Biomass, dist = "bray", data = merged_clean_filtered)
+```
+
+```{r}
+herb_algal_constrained_ord
+```
+
+```{r}
+par(mfrow = c(1, 3))  # 1 row, 3 columns
+
+# Plot 1: Species driving the community composition
+plot(herb_algal_constrained_ord, display = "species", main = "Species Driving Community Composition", type = "n")
+text(herb_algal_constrained_ord, display = "species", col = "blue", cex = 1)
+
+# Plot 2: Communities (sites) plotted
+plot(herb_algal_constrained_ord, display = "sites", main = "Communities", type = "n")
+points(herb_algal_constrained_ord, display = "sites", col = color_palette[colors], pch = 21)  # Sites colored by the same scheme
+
+# Plot 3: Constrained ordination axes (environmental variables)
+plot(herb_algal_constrained_ord, display = "sites", type = 'n', main = "Relationship with Herbivore Biomass")
+points(herb_algal_constrained_ord, display = "sites", col = 'grey')
+text(herb_algal_constrained_ord, display = 'bp', col = 'blue')
+```
+
+## The whole fish community:
 ```{r}

### 2024-08-20 noam_data_exploration_v3.Rmd
started adding in effects size plots from emmeans

@@
 
-```{r glmm approach}
+```{r glmm approach, fig.height = 15, fig.width=20}
 alpha_diversity_quad$year_factor <- as.factor(alpha_diversity_quad$year) # for glmmTMB ar1 structure
@@
 plot(allEffects(richness_mod_2))
-# emmeans(richness_mod_2, pairwise ~ habitat | year_factor) # if we want to compare habitats year to year
+mod_emm <- emmeans(richness_mod_2, ~ habitat * island_side * year_factor) # if we want to compare habitats year to year
+plot(mod_emm) +
+  geom_point(aes(x = emmean, y = year_factor, colour = habitat))
+  facet_wrap(island_side~habitat)
 # model.sel(dredge(richness_mod_2))
+
+emmip(richness_mod_2, island_side ~ year_factor | habitat, CIs = TRUE) +
+  scale_colour_manual(values = island_side_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  scale_y_continuous(limits = c(0,2)) +
+#  facet_wrap(~habitat)+
+#  labs(title = "b. shannon quad level\n", y = "shannon\n") +
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+        axis.line.y = element_line(colour = "black"),
+        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        axis.line.x = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(size = 32),
+        legend.key.size = unit(2, "cm"), # increase legend size
+        legend.text = element_text(size = rel(2)), # increase legend text
+        legend.title = element_text(size = rel(2.25)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal',
+        legend.box.background = element_rect(colour = "black") # make legend horizontal
+        )
 

### 2024-08-20 noam_data_exploration_v3.Rmd
finished creating effects size plots for richness and slope models

--- 2024-08-19 noam_data_exploration_v3.Rmd
@@
 
-```{r gls approach}
+```{r gls approach, fig.height=10, fig.width=14}
 
@@
 emmeans::contrast(slope_emm, "consec", simple = "each", combine = TRUE, adjust = "bonferroni")
+slope_plot_info <- plot(slope_emm, plotit = FALSE)
+(emmeans_slope_graph <- 
+  ggplot(data =slope_plot_info) +
+  geom_point(aes(y = the.emmean, x = habitat, colour = island_side), size = 8, position = position_dodge(0.3)) +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL, x = habitat, colour = island_side), width = 0.2, size = 1.5,
+                position = position_dodge(0.3) )+
+  geom_hline(yintercept = 0) +
+  theme_classic() +
+  labs(y = "Estimated maringal mean slope\n", x = "Habitat") +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+#        panel.background = element_blank(),
+        legend.text = element_text(size=20),
+        legend.title = element_text(size=20),
+        axis.line = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
+        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain") ) )
+
 # Note: not sure I want to include this
@@
 
-```{r glmm approach}
+```{r glmm approach, fig.height = 20, fig.width=15}
 alpha_diversity_quad$year_factor <- as.factor(alpha_diversity_quad$year) # for glmmTMB ar1 structure
@@
 richness_mod_2 <- glmmTMB(
-  richness ~ year_factor*habitat*island_side + (1 | island_side/transect/location) + ar1(year_factor + 0 | location),
+  richness ~ year_factor*habitat*island_side + (1 | site/transect/location) + ar1(year_factor + 0 | location),
   family = poisson,
@@
 plot(allEffects(richness_mod_2))
-# emmeans(richness_mod_2, pairwise ~ habitat | year_factor) # if we want to compare habitats year to year
-# model.sel(dredge(richness_mod_2))
+mod_emm <- emmeans(richness_mod_2, ~ habitat * island_side * year_factor) # if we want to compare habitats year to year
+richness_plot_info <- plot(mod_emm, plotit = FALSE)
+(emmeans_graph <- 
+  ggplot(data =richness_plot_info) +
+  geom_point(aes(x = exp(the.emmean), y = year_factor, colour = island_side), size = 8) +
+  geom_errorbar(aes(xmin = exp(asymp.LCL), xmax = exp(asymp.UCL), y = year_factor, colour = island_side), width = 0.2, size = 1.5) +
+  facet_wrap(~habitat) +
+  theme_classic() +
+  labs(y = "Year\n", x = "Estimated marginal mean richness") +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+#        panel.background = element_blank(),
+        legend.text = element_text(size=20),
+        legend.title = element_text(size=20),
+        axis.line = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
+        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain") ) )
 

### 2024-08-20 noam_data_exploration_v3.Rmd
added log richness as a factor for gls modeling but hit an error - will look into tomorrow.

@@
 # FRAMING: What is the general trend in algal alpha diversity over the course of the time series? Increasing, decreasing, neutral?
-
+alpha_diversity_quad$logRich <- log(alpha_diversity_quad$richness + 1)
 columns <- c("location", "Intercept", "Slope", "Phi") # names of columns for dataframe

### 2024-08-21 noam_data_exploration_v3.Rmd
Continued troubleshooting issues with log transformed slope. May drop this issue in general. Continued updating models and figures to be at island_side instead of site level, added more effects-style plots.

@@
   # fit gls for current time series
-  loop_mod <- gls(richness ~ year, correlation =corAR1(form=~year), data = newData)
+  loop_mod <- gls(logRich ~ year, correlation =corAR1(form=~year), 
+                  control=glsControl(msVerbose=T), data = newData)
 
@@
 
-slope_mod <- glmmTMB(Slope ~ habitat*site + (1|site/transect), data = gls_coefficients_merged)
-# model convergence problem
-summary(slope_mod) # pretty low variance in random effects
-plot(gls_res <- simulateResiduals(slope_mod)) # there are issues, but I can't figure out what they are...
-# I think this is due to 1200 obs in model
-hist(residuals(slope_mod))
-plot(predict(slope_mod) ~ residuals(slope_mod))
-Anova(slope_mod)
-# Response: Slope
-#                 Chisq Df Pr(>Chisq)    
-# habitat      1397.999  3  < 2.2e-16 ***
-# site           39.095  5  2.273e-07 ***
-# habitat:site 1738.155 15  < 2.2e-16 ***
-r.squaredGLMM(slope_mod) # 0.22, 0.24
-plot(allEffects(slope_mod))
-# backreef: richness increases on east and north shore, but relatively constant west side
-# Fringe: richness decreasing on east and west side, but relatvely constant on north shore
-# outer_10: richness neutral/decreasing on east side, but neutral/increasing on north and west side
-# outer_17: richness decreasing north/east side, but neutral west side
-# Worth noting trends are slight - arguable whether they are biologically meaningful
-
-slope_emm <- emmeans(slope_mod, ~ habitat*site) # create emmeans object
-emmeans::contrast(slope_emm, "consec", simple = "each", combine = TRUE, adjust = "bonferroni")
-# Note: not sure I want to include this
-emmeans(slope_mod, pairwise ~ site | habitat)
+# slope_mod <- glmmTMB(Slope ~ habitat*site + (1|site/transect), data = gls_coefficients_merged)
+# # model convergence problem
+# summary(slope_mod) # pretty low variance in random effects
+# plot(gls_res <- simulateResiduals(slope_mod)) # there are issues, but I can't figure out what they are...
+# # I think this is due to 1200 obs in model
+# hist(residuals(slope_mod))
+# plot(predict(slope_mod) ~ residuals(slope_mod))
+# Anova(slope_mod)
+# # Response: Slope
+# #                 Chisq Df Pr(>Chisq)    
+# # habitat      1397.999  3  < 2.2e-16 ***
+# # site           39.095  5  2.273e-07 ***
+# # habitat:site 1738.155 15  < 2.2e-16 ***
+# r.squaredGLMM(slope_mod) # 0.22, 0.24
+# plot(allEffects(slope_mod))
+# # backreef: richness increases on east and north shore, but relatively constant west side
+# # Fringe: richness decreasing on east and west side, but relatvely constant on north shore
+# # outer_10: richness neutral/decreasing on east side, but neutral/increasing on north and west side
+# # outer_17: richness decreasing north/east side, but neutral west side
+# # Worth noting trends are slight - arguable whether they are biologically meaningful
+# 
+# 
+# 
+# 
+# slope_emm <- emmeans(slope_mod, ~ habitat*site) # create emmeans object
+# emmeans::contrast(slope_emm, "consec", simple = "each", combine = TRUE, adjust = "bonferroni")
+# # Note: not sure I want to include this
+# emmeans(slope_mod, pairwise ~ site | habitat)
+#
 
@@
 
-```{r glmm approach, fig.height = 20, fig.width=15}
+```{r glmm approach, fig.height = 15, fig.width=25}
 alpha_diversity_quad$year_factor <- as.factor(alpha_diversity_quad$year) # for glmmTMB ar1 structure
@@
 richness_plot_info <- plot(mod_emm, plotit = FALSE)
+richness_plot_info$year_numeric <- as.numeric(levels(richness_plot_info$year_factor))
 (emmeans_graph <- 
   ggplot(data =richness_plot_info) +
-  geom_point(aes(x = exp(the.emmean), y = year_factor, colour = island_side), size = 8) +
-  geom_errorbar(aes(xmin = exp(asymp.LCL), xmax = exp(asymp.UCL), y = year_factor, colour = island_side), width = 0.2, size = 1.5) +
+  geom_point(aes(y = exp(the.emmean), x = year_numeric, colour = island_side), size = 8) +
+  geom_errorbar(aes(ymin = exp(asymp.LCL), ymax = exp(asymp.UCL), x = year_numeric, colour = island_side), 
+                width = 0.2, size = 1.5) +
+  geom_segment(aes(y = 5.2, yend = 5.2, x = 2006, xend = 2010), linewidth = 4, alpha = 0.2, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 2) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 2) +
   facet_wrap(~habitat) +
   theme_classic() +
-  labs(y = "Year\n", x = "Estimated marginal mean richness") +
+  labs(x = "Year", y = "Estimated marginal mean richness\n") +
   theme(panel.grid.major = element_blank(), 
@@
 
-```{r total turnover, fig.height = 24, fig.width=16}
+```{r total turnover, fig.height = 16, fig.width=24}
 # the turnover function calculates turnover as the proportion or species that either appear or disappear between timepoints
@@
 }
+
+# for backtransforming logit transformed data in graph
+inv_logit <- function(x){
+  exp(x)/(1+exp(x))
+}
+
 total_turnover$Turnover_trans <- sv_trans(total_turnover$total, total_turnover)
@@
 total_turnover_mod1 <- glmmTMB(Turnover_trans ~ year_factor*habitat*island_side +
-                                 (1 | island_side/site/transect) + ar1(year_factor + 0 | location),
-                               family = beta_family(), data = total_turnover)
+                                 (1 | site/transect/location) + ar1(year_factor + 0 | location),
+                               family = beta_family(), data = total_turnover) # convergence problem, probably overspecified
 summary(total_turnover_mod1) # low temporal autocorrelation... maybe don't need ar1 term?
@@
 Anova(total_turnover_mod1) # obviously everything sig
-
-# total_turnover_mod2 <- glmmTMB(Turnover_trans ~ year_factor*habitat*site + (1 | site/transect/location),
-#                                family = beta_family(), data = total_turnover)
-# summary(total_turnover_mod2) # 
-# plot(total_turnover_res2 <- simulateResiduals(total_turnover_mod2)) # residuals still look like shit
-# Anova(total_turnover_mod2) 
-# anova(total_turnover_mod2, total_turnover_mod1) # models are sig different from each other
-# AICc(total_turnover_mod2, total_turnover_mod1) # first model is better
-# # need to figure out residuals problem
-# plot(allEffects(total_turnover_mod1))
-#
-
-# plot
-total_turnover_means <-
-  total_turnover %>%
-  dplyr::group_by(year, island_side, site, habitat, site_habitat) %>%
-  dplyr::summarise(Turnover_mean = mean(total), 
-            Turnover_se = sd(total)/sqrt(n()))
-
-(turnover_plot <-
-  ggplot(data = total_turnover_means) +
-  geom_point(aes(x = year, y = Turnover_mean, colour = island_side, shape = habitat, group = island_side_habitat), size = 8, stroke = 2) +
-  geom_line(aes(x = year, y = Turnover_mean, colour = island_side, group = island_side_habitat), size = 1.2) +
-  geom_errorbar(aes(x = year, ymin = Turnover_mean - Turnover_se , ymax = Turnover_mean + Turnover_se, colour = island_side,
-                    group = island_side_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = island_side_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~island_side_habitat, nrow = 6)+
-  labs(title = "a. Turnover\n", y = "Total turnover\n") +
-  theme_bw() + 
+total_turnover_emm <- emmeans(total_turnover_mod1, ~ year_factor*habitat*island_side)
+total_turnover_plot_info <- plot(total_turnover_emm, plotit = F)
+total_turnover_plot_info$year_numeric <- as.numeric(levels(total_turnover_plot_info$year_factor))
+
+(total_turnover_emmeans_graph <- 
+  ggplot(data =total_turnover_plot_info) +
+  geom_point(aes(y = inv_logit(the.emmean), x = year_numeric, colour = island_side), size = 8) +
+  geom_errorbar(aes(ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), x = year_numeric, colour = island_side), 
+                width = 0.2, size = 1.5) +
+  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 4, alpha = 0.2, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 2) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 2) +
+  facet_wrap(~habitat) +
+  theme_classic() +
+  labs(x = "Year", y = "Estimated marginal mean total turnover\n") +
   theme(panel.grid.major = element_blank(), 
         panel.grid.minor = element_blank(),
-        panel.background = element_blank(),
-        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
-        axis.line.y = element_line(colour = "black"),
-        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
-        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
-        axis.line.x = element_line(colour = "black"),
-        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
-        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
-        strip.text.x = element_text(size = 32),
-        legend.key.size = unit(2, "cm"), # increase legend size
-        legend.text = element_text(size = rel(2)), # increase legend text
-        legend.title = element_text(size = rel(2.25)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal',
-        legend.box.background = element_rect(colour = "black") # make legend horizontal
-        )
-    )
+#        panel.background = element_blank(),
+        legend.text = element_text(size=20),
+        legend.title = element_text(size=20),
+        axis.line = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
+        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain") ) )
+
+# plot
+#total_turnover_means <-
+#  total_turnover %>%
+#  dplyr::group_by(year, island_side, site, habitat, site_habitat) %>%
+#  dplyr::summarise(Turnover_mean = mean(total), 
+#            Turnover_se = sd(total)/sqrt(n()))
 #
-```
-
-```{r appearances, fig.height = 24, fig.width=16}
+#(turnover_plot <-
+#  ggplot(data = total_turnover_means) +
+#  geom_point(aes(x = year, y = Turnover_mean, colour = island_side, shape = habitat, group = island_side_habitat), size = 8, stroke = #2) +
+#  geom_line(aes(x = year, y = Turnover_mean, colour = island_side, group = island_side_habitat), size = 1.2) +
+#  geom_errorbar(aes(x = year, ymin = Turnover_mean - Turnover_se , ymax = Turnover_mean + Turnover_se, colour = island_side,
+#                    group = island_side_habitat), width = 0.5, size = 1.2) +
+#  scale_colour_manual(values = island_side_colours) +
+#  scale_shape_manual(values = habitat_shapes) +
+#  facet_wrap(~island_side_habitat, nrow = 6)+
+#  labs(title = "a. Turnover\n", y = "Total turnover\n") +
+#  theme_bw() + 
+#  theme(panel.grid.major = element_blank(), 
+#        panel.grid.minor = element_blank(),
+#        panel.background = element_blank(),
+#        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+#        axis.line.y = element_line(colour = "black"),
+#        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+#        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+#        axis.line.x = element_line(colour = "black"),
+#        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+#        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+#        strip.text.x = element_text(size = 32),
+#        legend.key.size = unit(2, "cm"), # increase legend size
+#        legend.text = element_text(size = rel(2)), # increase legend text
+#        legend.title = element_text(size = rel(2.25)),
+#        legend.position = 'bottom', # move legend to bottom
+#        legend.box = 'horizontal',
+#        legend.box.background = element_rect(colour = "black") # make legend horizontal
+#        )
+#    )
+#
+```
+
+```{r appearances, fig.height = 16, fig.width=24}
 # examine appearance and disappearances independently 
@@
 # model
-appearance_turnover_mod1 <- glmmTMB(Appearance_trans ~ year_factor*habitat*site +
+appearance_turnover_mod1 <- glmmTMB(Appearance_trans ~ year_factor*habitat*island_side +
                                  (1 | site/transect/location) + ar1(year_factor + 0 | location),
-                               family = beta_family(), data = appearance)
+                               family = beta_family(), data = appearance) # overspecified
 summary(appearance_turnover_mod1) # low temporal autocorrelation and it's negative
@@
 
-appearance_turnover_mod2 <- glmmTMB(Appearance_trans ~ year_factor*habitat*site + (1 | site/transect/location),
-                               family = beta_family(), data = appearance)
-summary(appearance_turnover_mod2) # 
-plot(appearance_turnover_res2 <- simulateResiduals(appearance_turnover_mod2)) #
-Anova(appearance_turnover_mod2) 
-anova(appearance_turnover_mod2, appearance_turnover_mod1) # models are sig different from each other
-AICc(appearance_turnover_mod2, appearance_turnover_mod1) # first model is better but not as drastic as total
-
-plot(allEffects(appearance_turnover_mod1))
-
-appearance_means <-
-  appearance %>%
-  dplyr::group_by(year, site, habitat, site_habitat) %>%
-  dplyr::summarise(Appearance_mean = mean(appearance), 
-            Appearance_se = sd(appearance)/sqrt(n()))
-
-(appearance_plot <-
-  ggplot(data = appearance_means) +
-  geom_point(aes(x = year, y = Appearance_mean, colour = site, shape = habitat, group = site_habitat), size = 8, stroke = 2) +
-  geom_line(aes(x = year, y = Appearance_mean, colour = site, group = site_habitat), size = 1.2) +
-  geom_errorbar(aes(x = year, ymin = Appearance_mean - Appearance_se , ymax = Appearance_mean + Appearance_se, colour = site,
-                    group = site_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~site_habitat, nrow = 6)+
-  labs(title = "b. Appearances\n", y = "Total turnover\n") +
-  theme_bw() + 
+# appearance_turnover_mod2 <- glmmTMB(Appearance_trans ~ year_factor*habitat*site + (1 | site/transect/location),
+#                                family = beta_family(), data = appearance)
+# summary(appearance_turnover_mod2) # 
+# plot(appearance_turnover_res2 <- simulateResiduals(appearance_turnover_mod2)) #
+# Anova(appearance_turnover_mod2) 
+# anova(appearance_turnover_mod2, appearance_turnover_mod1) # models are sig different from each other
+# AICc(appearance_turnover_mod2, appearance_turnover_mod1) # first model is better but not as drastic as total
+
+appearance_turnover_emm <- emmeans(appearance_turnover_mod1, ~ year_factor*habitat*island_side)
+appearance_turnover_plot_info <- plot(appearance_turnover_emm, plotit = F)
+appearance_turnover_plot_info$year_numeric <- as.numeric(levels(appearance_turnover_plot_info$year_factor))
+
+(appearance_turnover_emmeans_graph <- 
+  ggplot(data =appearance_turnover_plot_info) +
+  geom_point(aes(y = inv_logit(the.emmean), x = year_numeric, colour = island_side), size = 8) +
+  geom_errorbar(aes(ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), x = year_numeric, colour = island_side), 
+                width = 0.2, size = 1.5) +
+  geom_segment(aes(y = 0.5, yend = 0.5, x = 2006, xend = 2010), linewidth = 4, alpha = 0.2, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 2) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 2) +
+  facet_wrap(~habitat) +
+  theme_classic() +
+  labs(x = "Year", y = "Estimated marginal mean appearance turnover\n") +
   theme(panel.grid.major = element_blank(), 
         panel.grid.minor = element_blank(),
-        panel.background = element_blank(),
-        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
-        axis.line.y = element_line(colour = "black"),
-        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
-        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
-        axis.line.x = element_line(colour = "black"),
-        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
-        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
-        strip.text.x = element_text(size = 32),
-        legend.key.size = unit(2, "cm"), # increase legend size
-        legend.text = element_text(size = rel(2)), # increase legend text
-        legend.title = element_text(size = rel(2.25)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal',
-        legend.box.background = element_rect(colour = "black") # make legend horizontal
-        )
-    )
+#        panel.background = element_blank(),
+        legend.text = element_text(size=20),
+        legend.title = element_text(size=20),
+        axis.line = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
+        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain") ) )
+
+plot(allEffects(appearance_turnover_mod1))
+
+# appearance_means <-
+#   appearance %>%
+#   dplyr::group_by(year, site, habitat, site_habitat) %>%
+#   dplyr::summarise(Appearance_mean = mean(appearance), 
+#             Appearance_se = sd(appearance)/sqrt(n()))
 # 
-```
-
-```{r disappearances, fig.height = 24, fig.width=16}
+# (appearance_plot <-
+#   ggplot(data = appearance_means) +
+#   geom_point(aes(x = year, y = Appearance_mean, colour = site, shape = habitat, group = site_habitat), size = 8, stroke = 2) +
+#   geom_line(aes(x = year, y = Appearance_mean, colour = site, group = site_habitat), size = 1.2) +
+#   geom_errorbar(aes(x = year, ymin = Appearance_mean - Appearance_se , ymax = Appearance_mean + Appearance_se, colour = site,
+#                     group = site_habitat), width = 0.5, size = 1.2) +
+#   scale_colour_manual(values = site_colours) +
+#   scale_shape_manual(values = habitat_shapes) +
+#   facet_wrap(~site_habitat, nrow = 6)+
+#   labs(title = "b. Appearances\n", y = "Total turnover\n") +
+#   theme_bw() + 
+#   theme(panel.grid.major = element_blank(), 
+#         panel.grid.minor = element_blank(),
+#         panel.background = element_blank(),
+#         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+#         axis.line.y = element_line(colour = "black"),
+#         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+#         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+#         axis.line.x = element_line(colour = "black"),
+#         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+#         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+#         strip.text.x = element_text(size = 32),
+#         legend.key.size = unit(2, "cm"), # increase legend size
+#         legend.text = element_text(size = rel(2)), # increase legend text
+#         legend.title = element_text(size = rel(2.25)),
+#         legend.position = 'bottom', # move legend to bottom
+#         legend.box = 'horizontal',
+#         legend.box.background = element_rect(colour = "black") # make legend horizontal
+#         )
+#     )
+# #
+```
+
+```{r disappearances, fig.width = 24, fig.height=16}
 disappearance <-
@@
 disappearance <- merge(turnover_meta, disappearance, by = c("year", "location"))
-disappearance$Disappearance_trans <- sv_trans(disappearance$disappearance, disappearance)
+disappearance$disappearance_trans <- sv_trans(disappearance$disappearance, disappearance)
 disappearance$year_factor <- as.factor(disappearance$year)
@@
 # model
-disappearance_turnover_mod1 <- glmmTMB(Disappearance_trans ~ year_factor*habitat*site +
+disappearance_turnover_mod1 <- glmmTMB(disappearance_trans ~ year_factor*habitat*island_side +
                                  (1 | site/transect/location) + ar1(year_factor + 0 | location),
-                               family = beta_family(), data = disappearance)
-summary(disappearance_turnover_mod1) # strong autocorrelation here
-plot(disappearance_turnover_res <- simulateResiduals(disappearance_turnover_mod1)) # 
-Anova(disappearance_turnover_mod1) #
-
-disappearance_turnover_mod2 <- glmmTMB(Disappearance_trans ~ year_factor*habitat*site + (1 | site/transect/location),
-                               family = beta_family(), data = disappearance)
-summary(disappearance_turnover_mod2) # 
-plot(disappearance_turnover_res2 <- simulateResiduals(disappearance_turnover_mod2)) #
-Anova(disappearance_turnover_mod2) 
-anova(disappearance_turnover_mod2, disappearance_turnover_mod1) # only case where models are not sig different from each other
-AICc(disappearance_turnover_mod2, disappearance_turnover_mod1) # models almost equivalent
-
-plot(allEffects(disappearance_turnover_mod1))
-
-disappearance_means <-
-  disappearance %>%
-  dplyr::group_by(year, site, habitat, site_habitat) %>%
-  dplyr::summarise(Disappearance_mean = mean(disappearance), 
-            Disappearance_se = sd(disappearance)/sqrt(n()))
-
-(disappearance_plot <-
-  ggplot(data = disappearance_means) +
-  geom_point(aes(x = year, y = Disappearance_mean, colour = site, shape = habitat, group = site_habitat), size = 8, stroke = 2) +
-  geom_line(aes(x = year, y = Disappearance_mean, colour = site, group = site_habitat), size = 1.2) +
-  geom_errorbar(aes(x = year, ymin = Disappearance_mean - Disappearance_se , ymax = Disappearance_mean + Disappearance_se, colour = site,
-                    group = site_habitat), width = 0.5, size = 1.2) +
-  scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~site_habitat, nrow = 6)+
-  labs(title = "c. Disappearance\n", y = "Total turnover\n") +
-  theme_bw() + 
+                               family = beta_family(), data = disappearance) # overspecified
+summary(disappearance_turnover_mod1) # low temporal autocorrelation and it's negative
+plot(disappearance_turnover_res <- simulateResiduals(disappearance_turnover_mod1)) # similar residuals problem
+Anova(disappearance_turnover_mod1) # obviously everything sig
+
+# disappearance_turnover_mod2 <- glmmTMB(disappearance_trans ~ year_factor*habitat*site + (1 | site/transect/location),
+#                                family = beta_family(), data = disappearance)
+# summary(disappearance_turnover_mod2) # 
+# plot(disappearance_turnover_res2 <- simulateResiduals(disappearance_turnover_mod2)) #
+# Anova(disappearance_turnover_mod2) 
+# anova(disappearance_turnover_mod2, disappearance_turnover_mod1) # models are sig different from each other
+# AICc(disappearance_turnover_mod2, disappearance_turnover_mod1) # first model is better but not as drastic as total
+
+disappearance_turnover_emm <- emmeans(disappearance_turnover_mod1, ~ year_factor*habitat*island_side)
+disappearance_turnover_plot_info <- plot(disappearance_turnover_emm, plotit = F)
+disappearance_turnover_plot_info$year_numeric <- as.numeric(levels(disappearance_turnover_plot_info$year_factor))
+
+(disappearance_turnover_emmeans_graph <- 
+  ggplot(data =disappearance_turnover_plot_info) +
+  geom_point(aes(y = inv_logit(the.emmean), x = year_numeric, colour = island_side), size = 8) +
+  geom_errorbar(aes(ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), x = year_numeric, colour = island_side), 
+                width = 0.2, size = 1.5) +
+  geom_segment(aes(y = 0.5, yend = 0.5, x = 2006, xend = 2010), linewidth = 4, alpha = 0.2, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 2) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 2) +
+  facet_wrap(~habitat) +
+  theme_classic() +
+  labs(x = "Year", y = "Estimated marginal mean disappearance turnover\n") +
   theme(panel.grid.major = element_blank(), 
         panel.grid.minor = element_blank(),
-        panel.background = element_blank(),
-        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
-        axis.line.y = element_line(colour = "black"),
-        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
-        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
-        axis.line.x = element_line(colour = "black"),
-        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
-        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
-        strip.text.x = element_text(size = 32),
-        legend.key.size = unit(2, "cm"), # increase legend size
-        legend.text = element_text(size = rel(2)), # increase legend text
-        legend.title = element_text(size = rel(2.25)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal',
-        legend.box.background = element_rect(colour = "black") # make legend horizontal
-        )
-    )
+#        panel.background = element_blank(),
+        legend.text = element_text(size=20),
+        legend.title = element_text(size=20),
+        axis.line = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
+        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain") ) )
+
+plot(allEffects(disappearance_turnover_mod1))
+
+# disappearance_means <-
+#   disappearance %>%
+#   dplyr::group_by(year, site, habitat, site_habitat) %>%
+#   dplyr::summarise(Disappearance_mean = mean(disappearance), 
+#             Disappearance_se = sd(disappearance)/sqrt(n()))
+# 
+# (disappearance_plot <-
+#   ggplot(data = disappearance_means) +
+#   geom_point(aes(x = year, y = Disappearance_mean, colour = site, shape = habitat, group = site_habitat), size = 8, stroke = 2) +
+#   geom_line(aes(x = year, y = Disappearance_mean, colour = site, group = site_habitat), size = 1.2) +
+#   geom_errorbar(aes(x = year, ymin = Disappearance_mean - Disappearance_se , ymax = Disappearance_mean + Disappearance_se, colour = # site,
+#                     group = site_habitat), width = 0.5, size = 1.2) +
+#   scale_colour_manual(values = site_colours) +
+#   scale_shape_manual(values = habitat_shapes) +
+#   facet_wrap(~site_habitat, nrow = 6)+
+#   labs(title = "c. Disappearance\n", y = "Total turnover\n") +
+#   theme_bw() + 
+#   theme(panel.grid.major = element_blank(), 
+#         panel.grid.minor = element_blank(),
+#         panel.background = element_blank(),
+#         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+#         axis.line.y = element_line(colour = "black"),
+#         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+#         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+#         axis.line.x = element_line(colour = "black"),
+#         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+#         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+#         strip.text.x = element_text(size = 32),
+#         legend.key.size = unit(2, "cm"), # increase legend size
+#         legend.text = element_text(size = rel(2)), # increase legend text
+#         legend.title = element_text(size = rel(2.25)),
+#         legend.position = 'bottom', # move legend to bottom
+#         legend.box = 'horizontal',
+#         legend.box.background = element_rect(colour = "black") # make legend horizontal
+#         )
+#     )
 ```
@@
 ```{r basic diversity stability plots quadrat level, fig.height=8, fig.width=10}
-(shannon_cv_plot = 
-  ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = shannon_mean, y = cover_cv, colour = site, shape = habitat), size = 4) + 
-   scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-   facet_wrap(~site) +
+
+richness_cv_mod <- glmmTMB(cover_cv ~ richness_mean*habitat*island_side + (1|site/transect), family = Gamma(link = "log"),
+                          data = diversity_stability_quadrat, na.action = na.pass)
+# model.sel(dredge(richness_cv_mod)) # top model is as specified, second best (delta = 10) doesn't include rich:hab:coast
+r.squaredGLMM(richness_cv_mod) # 0.61, 0.64
+summary(richness_cv_mod)
+emmeans(richness_cv_mod, pairwise ~ island_side | habitat, var = "richness_mean")
+richness_cv_emm <- emmip(richness_cv_mod,
+                        habitat ~ richness_mean | island_side,
+                        at = list(richness_mean = seq(0,5,0.1)), plotit = F, CIs = T)
+
+(richness_cv_plot = 
+  ggplot() +
+  geom_point(data = diversity_stability_quadrat,
+             aes(x = richness_mean, y = cover_cv, colour = island_side, shape = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = richness_cv_emm, aes(x = richness_mean, y = exp(yvar), colour = island_side), size = 2) +
+  geom_ribbon(data = richness_cv_emm, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = island_side), alpha = 0.3) +
+#   scale_colour_manual(values = site_colours) +
+  scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~habitat) +
 
@@
 ```{r portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
+synchrony_cv_mod <- glmmTMB(cover_cv ~ loreau_synchrony*habitat*island_side + (1|site/transect), 
+                             family = Gamma(link = "log"),
+                          data = diversity_stability_quadrat, na.action = na.pass)
+# model.sel(dredge(synchrony_cv_mod)) # top model by far is as specified
+r.squaredGLMM(synchrony_cv_mod) # 0.73, 0.74
+summary(synchrony_cv_mod)
+emmeans(synchrony_cv_mod, pairwise ~ island_side | habitat, var = "loreau_synchrony", at = list(loreau_synchrony = seq(0,1,0.05)))
+synchrony_cv_emm <- emmip(synchrony_cv_mod,
+                        habitat ~ loreau_synchrony | island_side,
+                        at = list(loreau_synchrony = seq(0,1,0.05)), plotit = F, CIs = T)
+
 (loreau_stability_plot_quad = 
   ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = loreau_synchrony, y = cover_cv, colour = site, shape = habitat), size = 4, stroke = 1) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~site) +
+  geom_point(aes(x = loreau_synchrony, y = cover_cv, colour = island_side, shape = habitat), size = 2, alpha = 0.5) + 
+ # geom_line(data = synchrony_cv_emm, aes(x = loreau_synchrony, y = exp(yvar), colour = island_side), size = 1) +
+ # geom_ribbon(data = synchrony_cv_emm, aes(x = loreau_synchrony, ymin = LCL, ymax = UCL, fill = island_side),
+ #             alpha = 0.3) +
+   scale_shape_manual(values = habitat_shapes) +
+#  facet_wrap(~habitat) +
 

### 2024-08-21 noam_data_exploration_v3.Rmd
Continued troubleshooting issues with log transformed slope. May drop this issue in general. Continued updating models and figures to be at island_side instead of site level, added more effects-style plots.

@@
 ```{r portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
-synchrony_cv_mod <- glmmTMB(cover_cv ~ loreau_synchrony*habitat*island_side + (1|site/transect), 
+synchrony_cv_mod <- glmmTMB(cover_cv ~ loreau_synchrony*habitat*island_side + (1|island_side/site/transect), 
                              family = Gamma(link = "log"),
@@
 # model.sel(dredge(synchrony_cv_mod)) # top model by far is as specified
-r.squaredGLMM(synchrony_cv_mod) # 0.73, 0.74
+r.squaredGLMM(synchrony_cv_mod) # 0.78, 0.79
 summary(synchrony_cv_mod)

### 2024-08-22 noam_data_exploration_v3.Rmd
added shannon diversity for adiv models/effects plots. Adjusted figs to log scale y axis for parity with model.

--- 2024-08-21 noam_data_exploration_v3.Rmd
@@
 
+```{r gls shannon, fig.height=10, fig.width=14}
+alpha_diversity_quad$logShann <- log(alpha_diversity_quad$shannon + 1)
+columns <- c("location", "Intercept", "Slope", "Phi") # names of columns for dataframe
+gls_shann_coefficients <- # new dataframe name
+  data.frame(matrix(nrow = length(unique(alpha_diversity_quad$location)), # 1200 quadrats = 1200 different time series
+                    ncol = length(columns)) # specified to make sure number of columns doesn't exceed what was specified
+             )
+colnames(gls_shann_coefficients) <- columns # change column names to what was specified
+gls_shann_coefficients$location <- levels(alpha_diversity_quad$location) # populate location column with each quad/time series
+
+for(i in 1:length(unique(alpha_diversity_quad$location)) ){ # for each  quadrat
+  # create a new dataframe for the quadrat of the current iteration
+  newData <- subset(alpha_diversity_quad, location == levels(alpha_diversity_quad$location)[i])
+
+  # fit gls for current time series
+  loop_mod <- gls(shannon ~ year, correlation =corAR1(form=~year), control = list(singular.ok = TRUE), data = newData)
+
+  # populate appropriate row and column with the relevant coeffcients from gls
+  gls_shann_coefficients[i, 2] <- unname(coef(loop_mod)[1])
+  gls_shann_coefficients[i, 3] <- unname(coef(loop_mod)[2])
+  gls_shann_coefficients[i, 4] <- loop_mod[["modelStruct"]][["corStruct"]]
+
+}
+
+hist(gls_shann_coefficients$Slope) # looks basically normal, bounded -Inf, Inf
+
+hist(gls_shann_coefficients$Intercept) # same with intercept - although it probably shouldn't be
+hist(gls_shann_coefficients$Phi)
+
+gls_shann_coefficients_merged <- merge(gls_shann_coefficients, extra_meta, by = "location")
+
+slope_shann_mod <- glmmTMB(Slope ~ habitat*island_side + (1|island_side/site/transect), data = gls_shann_coefficients_merged)
+summary(slope_shann_mod) # pretty low variance in random effects
+plot(gls_res <- simulateResiduals(slope_shann_mod))
+
+hist(residuals(slope_shann_mod))
+plot(predict(slope_shann_mod) ~ residuals(slope_shann_mod))
+Anova(slope_shann_mod)
+
+r.squaredGLMM(slope_shann_mod) # 0.19, 0.22
+plot(allEffects(slope_shann_mod))
+
+slope_emm <- emmeans(slope_shann_mod, ~ habitat*island_side) # create emmeans object
+slope_plot_info <- plot(slope_emm, plotit = FALSE)
+(emmeans_slope_graph <- 
+  ggplot(data =slope_plot_info) +
+  geom_point(aes(y = the.emmean, x = habitat, colour = island_side), size = 8, position = position_dodge(0.3)) +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL, x = habitat, colour = island_side), width = 0.2, size = 1.5,
+                position = position_dodge(0.3) )+
+  geom_hline(yintercept = 0) +
+  theme_classic() +
+  labs(y = "Estimated maringal mean slope\n", x = "Habitat") +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+#        panel.background = element_blank(),
+        legend.text = element_text(size=20),
+        legend.title = element_text(size=20),
+        axis.line = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
+        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain") ) )
+
+# Note: not sure I want to include this
+emmeans(slope_shann_mod, pairwise ~ island_side | habitat)
+# north and east coast not different from each other in backreef
+# east and west not different from each other in fringing
+# north and west not different from each other outer 10
+# no sites different from each other outer 17
+slope_est <- as.data.frame(slope_emm) # create df of emmeans estimates
+
+```
+
 ```{r glmm approach, fig.height = 15, fig.width=25}
@@
   geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 2) +
+  scale_y_log10() +
   facet_wrap(~habitat) +
@@
 
+```{r glmm shannon, fig.height = 15, fig.width=25}
+shannon_mod <- glmmTMB(
+  log(shannon+1) ~ year_factor*habitat*island_side + (1 | site/transect/location) + ar1(year_factor + 0 | location),
+
+  data = alpha_diversity_quad,
+  control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000)), na.action = na.pass
+) # overspecified but going with it for now
+
+summary(shannon_mod) # looks like strong autocorrelation, which makes sense to me
+Anova(shannon_mod)
+plot(allEffects(shannon_mod_2))
+mod_emm <- emmeans(shannon_mod, ~ habitat * island_side * year_factor) # if we want to compare habitats year to year
+shannon_plot_info <- plot(mod_emm, plotit = FALSE)
+shannon_plot_info$year_numeric <- as.numeric(levels(shannon_plot_info$year_factor))
+(emmeans_graph <- 
+  ggplot(data =shannon_plot_info) +
+  geom_point(aes(y = exp(the.emmean), x = year_numeric, colour = island_side), size = 8) +
+  geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL), x = year_numeric, colour = island_side), 
+                width = 0.2, size = 1.5) +
+  geom_segment(aes(y = 2.5, yend = 2.5, x = 2006, xend = 2010), linewidth = 4, alpha = 0.2, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 2) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 2) +
+  scale_y_log10() +
+  facet_wrap(~habitat) +
+  theme_classic() +
+  labs(x = "Year", y = "Shannon-Weiner Index (H')\n") +
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+#        panel.background = element_blank(),
+        legend.text = element_text(size=20),
+        legend.title = element_text(size=20),
+        axis.line = element_line(colour = "black"),
+        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
+        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
+        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain") ) )
+
+```
+
 ```{r turnover troubleshooting, fig.height = 30, fig.width=25}
@@
   theme_classic() +
-  labs(x = "Year", y = "Estimated marginal mean total turnover\n") +
+  labs(x = "Year", y = "Total turnover\n") +
   theme(panel.grid.major = element_blank(), 
@@
   theme_classic() +
-  labs(x = "Year", y = "Estimated marginal mean appearance turnover\n") +
+  labs(x = "Year", y = "Appearances\n") +
   theme(panel.grid.major = element_blank(), 
@@
   theme_classic() +
-  labs(x = "Year", y = "Estimated marginal mean disappearance turnover\n") +
+  labs(x = "Year", y = "Disappearances\n") +
   theme(panel.grid.major = element_blank(), 
@@
   facet_wrap(~habitat) +
-
+  scale_y_log10() +
   theme_bw() + 
@@
 # model.sel(dredge(synchrony_cv_mod)) # top model by far is as specified
-r.squaredGLMM(synchrony_cv_mod) # 0.73, 0.74
+r.squaredGLMM(synchrony_cv_mod) # 0.78, 0.79
 summary(synchrony_cv_mod)
-emmeans(synchrony_cv_mod, pairwise ~ island_side | habitat, var = "loreau_synchrony", at = list(loreau_synchrony = seq(0,1,0.05)))
+synchrony_cv_effects <- emmeans(synchrony_cv_mod, pairwise ~ island_side | habitat, var = "loreau_synchrony")
 synchrony_cv_emm <- emmip(synchrony_cv_mod,
@@
   geom_point(aes(x = loreau_synchrony, y = cover_cv, colour = island_side, shape = habitat), size = 2, alpha = 0.5) + 
- # geom_line(data = synchrony_cv_emm, aes(x = loreau_synchrony, y = exp(yvar), colour = island_side), size = 1) +
- # geom_ribbon(data = synchrony_cv_emm, aes(x = loreau_synchrony, ymin = LCL, ymax = UCL, fill = island_side),
- #             alpha = 0.3) +
-   scale_shape_manual(values = habitat_shapes) +
-#  facet_wrap(~habitat) +
-
+  geom_line(data = synchrony_cv_emm, aes(x = loreau_synchrony, y = exp(yvar), colour = island_side), size = 1) +
+  geom_ribbon(data = synchrony_cv_emm, aes(x = loreau_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = island_side),
+             alpha = 0.3) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~habitat) +
+   scale_y_log10() +
   theme_bw() + 

### 2024-08-22 noam_data_exploration_v3.Rmd
Filtered model effects for diversity-stability model to plot only over the ranges of the data

@@
 
+## FOR FILTERING MODEL EFFECTS LATER
+diversity_ranges <- diversity_stability_quadrat %>%
+  group_by(island_side, habitat) %>%
+  summarise(min_richness = min(richness_mean),
+            max_richness = max(richness_mean),
+            min_shannon = min(shannon_mean),
+            max_shannon = max(shannon_mean),
+            min_loreau = min(loreau_synchrony),
+            max_loreau = max(loreau_synchrony))
+
 ```
@@
 summary(richness_cv_mod)
-emmeans(richness_cv_mod, pairwise ~ island_side | habitat, var = "richness_mean")
+emtrends(richness_cv_mod, pairwise ~ island_side | habitat, var = "richness_mean")
 richness_cv_emm <- emmip(richness_cv_mod,
@@
                         at = list(richness_mean = seq(0,5,0.1)), plotit = F, CIs = T)
+
+filtered_richness_effects <- richness_cv_emm %>%
+  left_join(diversity_ranges, by = c("island_side", "habitat")) %>%
+  filter(richness_mean >= min_richness & richness_mean <= max_richness)
 
@@
              aes(x = richness_mean, y = cover_cv, colour = island_side, shape = habitat), size = 2, alpha = 0.5) + 
-  geom_line(data = richness_cv_emm, aes(x = richness_mean, y = exp(yvar), colour = island_side), size = 2) +
-  geom_ribbon(data = richness_cv_emm, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = island_side), alpha = 0.3) +
+  geom_line(data = filtered_richness_effects, aes(x = richness_mean, y = exp(yvar), colour = island_side), size = 2) +
+  geom_ribbon(data = filtered_richness_effects, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = island_side), alpha = 0.3) +
 #   scale_colour_manual(values = site_colours) +
@@
 
-(shannon_cv_plot_2 = 
-  ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = shannon_mean, y = cover_cv, colour = site, shape = habitat), size = 4, stroke = 1) + 
-   scale_colour_manual(values = site_colours) +
+shannon_cv_mod <- glmmTMB(cover_cv ~ shannon_mean*habitat*island_side + (1|site/transect), family = Gamma(link = "log"),
+                          data = diversity_stability_quadrat, na.action = na.pass)
+# model.sel(dredge(shannon_cv_mod)) # top model is as specified, second best (delta = 10) doesn't include shann:hab:coast
+r.squaredGLMM(shannon_cv_mod) # 0.55, 0.59
+summary(shannon_cv_mod)
+emtrends(shannon_cv_mod, pairwise ~ island_side | habitat, var = "shannon_mean")
+shannon_cv_emm <- emmip(shannon_cv_mod,
+                        habitat ~ shannon_mean | island_side,
+                        at = list(shannon_mean = seq(0,1.25,0.05)), plotit = F, CIs = T)
+
+filtered_shannon_effects <- shannon_cv_emm %>%
+  left_join(diversity_ranges, by = c("island_side", "habitat")) %>%
+  filter(shannon_mean >= min_shannon & shannon_mean <= max_shannon)
+
+(shannon_cv_plot = 
+  ggplot() +
+  geom_point(data = diversity_stability_quadrat,
+             aes(x = shannon_mean, y = cover_cv, colour = island_side, shape = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = filtered_shannon_effects, aes(x = shannon_mean, y = exp(yvar), colour = island_side), size = 2) +
+  geom_ribbon(data = filtered_shannon_effects, aes(x = shannon_mean, ymin = exp(LCL), ymax = exp(UCL), fill = island_side), alpha = 0.3) +
+#   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
   facet_wrap(~habitat) +
+  scale_y_log10() +
   theme_bw() + 
@@
          )
-
-# lter_2 obscures a lot of the trends
 
@@
 
+filtered_synchrony_effects <- synchrony_cv_emm %>%
+  left_join(diversity_ranges, by = c("island_side", "habitat")) %>%
+  filter(loreau_synchrony >= min_loreau & loreau_synchrony <= max_loreau)
+
 (loreau_stability_plot_quad = 
@@
   geom_point(aes(x = loreau_synchrony, y = cover_cv, colour = island_side, shape = habitat), size = 2, alpha = 0.5) + 
-  geom_line(data = synchrony_cv_emm, aes(x = loreau_synchrony, y = exp(yvar), colour = island_side), size = 1) +
-  geom_ribbon(data = synchrony_cv_emm, aes(x = loreau_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = island_side),
+  geom_line(data = filtered_synchrony_effects, aes(x = loreau_synchrony, y = exp(yvar), colour = island_side), size = 1) +
+  geom_ribbon(data = filtered_synchrony_effects, aes(x = loreau_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = island_side),
              alpha = 0.3) +

### 2024-08-23 noam_data_exploration_v3.Rmd
Added figs and models for diversity synchrony stability relationships

@@
 ```{r diversity portfolio effects and synchrony quadrat level,  fig.height = 8, fig.width = 10}
-(loreau_diversity_plot_quad = 
-  ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = shannon_mean, y = loreau_synchrony, colour = site, shape = habitat), size = 4, stroke = 1) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~site) +
+diversity_stability_quadrat$loreau_synchrony_trans <- sv_trans(diversity_stability_quadrat$loreau_synchrony, diversity_stability_quadrat) # for modelling on beta
+synchrony_shann_mod <- glmmTMB(loreau_synchrony_trans ~ shannon_mean*habitat*island_side + (1|island_side/site/transect), 
+                             family = beta_family(),
+                          data = diversity_stability_quadrat, na.action = na.pass)
+# model.sel(dredge(synchrony_shann_mod)) # top model is as specified
+summary(synchrony_shann_mod)
+synchrony_shann_effects <- emtrends(synchrony_shann_mod, pairwise ~ island_side | habitat, var = "shannon_mean")
+synchrony_shann_emm <- emmip(synchrony_shann_mod,
+                        habitat ~ shannon_mean | island_side,
+                        at = list(shannon_mean = seq(0,1.25,0.05)), plotit = F, CIs = T)
+
+filtered_synchrony_div_effects <- synchrony_shann_emm %>%
+  left_join(diversity_ranges, by = c("island_side", "habitat")) %>%
+  filter(shannon_mean >= min_shannon & shannon_mean <= max_shannon)
+
+(loreau_diversity_plot_quad <-
+  ggplot() +
+  geom_point(data = diversity_stability_quadrat,
+             aes(x = shannon_mean, y = loreau_synchrony, colour = island_side, shape = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = filtered_synchrony_div_effects, aes(x = shannon_mean, y = inv_logit(yvar), colour = island_side), size = 1) +
+  geom_ribbon(data = filtered_synchrony_div_effects, aes(x = shannon_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL),
+                                                         fill = island_side),
+             alpha = 0.3) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~habitat) +
 
@@
 
-(portfolio_diversity_plot_quad = 
-  ggplot(data = diversity_stability_quadrat) +
-  geom_point(aes(x = shannon_mean, y = portfolio_effect, colour = site, shape = habitat), size = 4, stroke = 2) + 
-   scale_colour_manual(values = site_colours) +
-   scale_shape_manual(values = habitat_shapes) +
-  facet_wrap(~site) +
+synchrony_rich_mod <- glmmTMB(loreau_synchrony_trans ~ richness_mean*habitat*island_side + (1|island_side/site/transect), 
+                             family = beta_family(),
+                          data = diversity_stability_quadrat, na.action = na.pass)
+# model.sel(dredge(synchrony_rich_mod)) # top model is as specified
+summary(synchrony_rich_mod)
+synchrony_rich_effects <- emtrends(synchrony_rich_mod, pairwise ~ island_side | habitat, var = "richness_mean")
+synchrony_rich_emm <- emmip(synchrony_rich_mod,
+                        habitat ~ richness_mean | island_side,
+                        at = list(richness_mean = seq(0,5,0.1)), plotit = F, CIs = T)
+
+filtered_synchrony_rich_effects <- synchrony_rich_emm %>%
+  left_join(diversity_ranges, by = c("island_side", "habitat")) %>%
+  filter(richness_mean >= min_richness & richness_mean <= max_richness)
+
+(loreau_richness_plot_quad <-
+  ggplot() +
+  geom_point(data = diversity_stability_quadrat,
+             aes(x = richness_mean, y = loreau_synchrony, colour = island_side, shape = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = filtered_synchrony_rich_effects, aes(x = richness_mean, y = inv_logit(yvar), colour = island_side), size = 1) +
+  geom_ribbon(data = filtered_synchrony_rich_effects, aes(x = richness_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL),
+                                                         fill = island_side),
+             alpha = 0.3) +
+   scale_shape_manual(values = habitat_shapes) +
+  facet_wrap(~habitat) +
 
@@
          )
+
 ```
@@
 
-diversity_stability_quadrat$loreau_synchrony_trans <- sv_trans(diversity_stability_quadrat$loreau_synchrony, diversity_stability_quadrat) # for modelling on beta
 synchrony_diversity_universal <- glmmTMB(loreau_synchrony_trans ~shannon_mean*site + shannon_mean*habitat + site*habitat + (1|transect/location), family = beta_family(), data = diversity_stability_quadrat, na.action = na.pass)

### 2024-08-26 noam_beta_diversity.Rmd
Ran metaMDS on bray curtis matrices at habitat level and on full data. Created dissimilarity over time (interval) model and graph.

@@
 #pivot long 
-long_data <- pivot_longer(
-  data = data,
-  cols = 8:77,
-  names_to = "taxa",
-  values_to = "percent_cover")
-
- transect_level <- long_data %>%
-   dplyr::group_by(year, island_side, site, habitat, transect, taxa) %>%
-   dplyr::summarise(mean_cover = mean(percent_cover), 
-             SE = sd(percent_cover)/sqrt(n()))
-
-site_level <- long_data %>%
-  dplyr::group_by(year, island_side, site, habitat, site_habitat, taxa) %>%
-  dplyr::summarise(mean_site_cover = mean(percent_cover), 
-            site_se = sd(percent_cover)/sqrt(n()))
-
+# long_data <- pivot_longer(
+#   data = data,
+#   cols = 8:77,
+#   names_to = "taxa",
+#   values_to = "percent_cover") 
+# 
+# 
+#  transect_level <- long_data %>%
+#    dplyr::group_by(year, island_side, site, habitat, transect, taxa) %>%
+#    dplyr::summarise(mean_cover = mean(percent_cover), 
+#              SE = sd(percent_cover)/sqrt(n()))
+#  
+# site_level <- long_data %>%
+#   dplyr::group_by(year, island_side, site, habitat, site_habitat, taxa) %>%
+#   dplyr::summarise(mean_site_cover = mean(percent_cover), 
+#             site_se = sd(percent_cover)/sqrt(n()))
+# 
 # subset based on habitat
@@
 
-# bray_mod_universal <- glmmTMB(dissimilarity_trans ~ time*habitat*island_side + (1 | island_side/site/transect/location),
-#                               na.action = na.pass, family = beta_family(link = "logit"), data = bray_time_series) #
-# summary(bray_mod_universal) # a fair bit of variance w/i quads/transects
-# 
-# bray_mod_reduced <- glmmTMB(dissimilarity_trans ~ time*habitat + time*island_side + (1 | island_side/site/transect/location),
-#                               na.action = na.pass, family = beta_family(link = "logit"), data = bray_time_series)
+bray_mod_universal <- glmmTMB(dissimilarity_trans ~ time*habitat*island_side + (1 | island_side/site/transect/location),
+                              na.action = na.pass, family = beta_family(link = "logit"), data = bray_time_series) #
+# model.sel(dredge(bray_mod_universal)) # top model as specified
+summary(bray_mod_universal) # a fair bit of variance w/i quads/transects
+hist(residuals(bray_mod_universal))
+plot(predict(bray_mod_universal) ~ residuals(bray_mod_universal)) # looks like reflection of logit curve
 
+emtrends(bray_mod_universal, pairwise ~ habitat |island_side, "time")
+bray_trends_emm <- emmip(bray_mod_universal, habitat ~ time | island_side, at = list(time = seq(1,16,1)), plotit = F, CIs = T)
+inv_logit <- function(x) {
+  exp(x)/(1+exp(x))
+}
 ```
@@
 ```{r dissimilarity over time plot, fig.height=15, fig.width=15}
- ggplot(bray_time_series,
-        aes(x = time, y = dissimilarity, colour = island_side) )+
+ ggplot()+
+   geom_point(data = bray_time_series, aes(x = time, y = dissimilarity, colour = island_side), alpha = 0.01 ) +
+   geom_line(data = bray_trends_emm, aes(x = time, y = inv_logit(yvar), colour = island_side), linewidth = 2) +
+   geom_ribbon(data = bray_trends_emm, aes(x = time, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = island_side), alpha = 0.3) +
    facet_wrap(~habitat) +
-   geom_point() +
-   scale_colour_manual(values = island_side_colours) +
+
+ #  scale_colour_manual(values = island_side_colours) +
    stat_smooth(se = T, linewidth = 2) +
@@
 
-```{r fringing nmds set up}
-# identifying species w/ > 10% cover for plotting later
+```{r hab level nmds set up}
+
+#### WARNING: THIS CHUNK TAKES ~2 DAYS TO RUN! SUGGESTEDF TO USE PROVIDED CSV FILES WITH ORDINATION DATA ######
 
 #### FRINGING
-# NOTE: Used ChatGPT to generate these next few lines of code but it looked good so I went with it
+
 # fringing_columns_to_keep <- apply(fringing_species_matrix, 2, function(x) any(x >= 0.1))
@@
 #
-# # takes several hours to run - do so at your own peril
+
 # set.seed(1032) 
-# # ord3 <- metaMDS(fringing_braymat, dist	= "bray", k = 3, trymax = 100) 
+# # fringing_ord3 <- metaMDS(fringing_braymat, dist	= "bray", k = 3, trymax = 100) 
 # fringing_ord3 # stress = 0.018
@@
 
-fringing.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = fringing.site.scores, FUN = mean) 
+# fringing.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = fringing.site.scores, FUN = mean) 
 # fringing.cent.meta <- fringing_subset_meta %>%
@@
 # subset only the species w/ > 10% cover
-fringing.abundant.species.scores <- fringing.species.scores[fringing.species.scores$species %in% fringing_abundant_species_names,]
+# fringing.abundant.species.scores <- fringing.species.scores[fringing.species.scores$species %in% fringing_abundant_species_names,]
 
@@
 ############
-backreef_columns_to_keep <- apply(backreef_species_matrix, 2, function(x) any(x >= 0.1))
-backreef_abundant_species_matrix <- as.data.frame(backreef_species_matrix)[ , backreef_columns_to_keep]
-backreef_abundant_species_names <- colnames(backreef_abundant_species_matrix)
-
-backreef_subset_meta <- backreef_subset[ , which(names(backreef_subset) %in% meta_cols)]
-
-# for mean centroid:
-backreef_subset_meta$island_side_year <- paste(backreef_subset_meta$year, "-", backreef_subset_meta$island_side_habitat, sep = "")
-
-# takes several hours to run - do so at your own peril
-set.seed(1032) 
-# ord3 <- metaMDS(backreef_braymat, dist	= "bray", k = 3, trymax = 100) 
-backreef_ord3 # stress = 0.018
-
-sppscores(backreef_ord3) <- backreef_species_matrix # add back in species scores
-backreef.species.scores <- as.data.frame(scores(backreef_ord3, "species")) # some NaN for absent species
-backreef.species.scores <- na.omit(backreef.species.scores)
-backreef.species.scores$species <- rownames(backreef.species.scores)
-
-backreef.site.scores <- as.data.frame(vegan::scores(backreef_ord3, "sites"))  # extract the site scores and convert to a data.frame
-# add back in metadata
-backreef.site.scores$stratum <- rownames(backreef.site.scores)
-backreef.site.scores <- merge(backreef_subset_meta, backreef.site.scores, by = "stratum")
-
-write_csv(site.scores, here("data", "backreef_site_scores.csv"))
-write_csv(species.scores, here("data", "backreef_species_scores.csv"))
+# backreef_columns_to_keep <- apply(backreef_species_matrix, 2, function(x) any(x >= 0.1))
+# backreef_abundant_species_matrix <- as.data.frame(backreef_species_matrix)[ , backreef_columns_to_keep]
+# backreef_abundant_species_names <- colnames(backreef_abundant_species_matrix)
+# 
+# backreef_subset_meta <- backreef_subset[ , which(names(backreef_subset) %in% meta_cols)]
+# 
+# # for mean centroid:
+# backreef_subset_meta$island_side_year <- paste(backreef_subset_meta$year, "-", backreef_subset_meta$island_side_habitat, sep = "")
+# 
+# # takes several hours to run - do so at your own peril
+# set.seed(1032) 
+# backreef_ord3 <- metaMDS(backreef_braymat, dist	= "bray", k = 3, trymax = 100) 
+# backreef_ord3 # 
+# 
+# 
+# sppscores(backreef_ord3) <- backreef_species_matrix # add back in species scores
+# backreef.species.scores <- as.data.frame(scores(backreef_ord3, "species")) # some NaN for absent species
+# backreef.species.scores <- na.omit(backreef.species.scores)
+# backreef.species.scores$species <- rownames(backreef.species.scores) 
+# 
+# 
+# backreef.site.scores <- as.data.frame(vegan::scores(backreef_ord3, "sites"))  # extract the site scores and convert to a data.frame
+# # add back in metadata
+# backreef.site.scores$stratum <- rownames(backreef.site.scores)
+# backreef.site.scores <- merge(backreef_subset_meta, backreef.site.scores, by = "stratum")
+# 
+# 
+# 
+# 
+# write_csv(backreef.site.scores, here("data", "backreef_site_scores.csv"))
+# write_csv(backreef.species.scores, here("data", "backreef_species_scores.csv"))
 
@@
 
-outer10_columns_to_keep <- apply(outer10_species_matrix, 2, function(x) any(x >= 0.1))
-outer10_abundant_species_matrix <- as.data.frame(outer10_species_matrix)[ , outer10_columns_to_keep]
-outer10_abundant_species_names <- colnames(outer10_abundant_species_matrix)
+# outer10_columns_to_keep <- apply(outer10_species_matrix, 2, function(x) any(x >= 0.1))
+# outer10_abundant_species_matrix <- as.data.frame(outer10_species_matrix)[ , outer10_columns_to_keep]
+# outer10_abundant_species_names <- colnames(outer10_abundant_species_matrix)
+# 
+# outer10_subset_meta <- outer10_subset[ , which(names(outer10_subset) %in% meta_cols)]
+# 
+# # for mean centroid:
+# outer10_subset_meta$island_side_year <- paste(outer10_subset_meta$year, "-", outer10_subset_meta$island_side_habitat, sep = "")
+#
 
-outer10_subset_meta <- outer10_subset[ , which(names(outer10_subset) %in% meta_cols)]
-
-# for mean centroid:
-outer10_subset_meta$island_side_year <- paste(outer10_subset_meta$year, "-", outer10_subset_meta$island_side_habitat, sep = "")
-
-# takes several hours to run - do so at your own peril
-set.seed(1032) 
-# ord3 <- metaMDS(outer10_braymat, dist	= "bray", k = 3, trymax = 100) 
-outer10_ord3 # stress = 0.018
-
-sppscores(outer10_ord3) <- outer10_species_matrix # add back in species scores
-outer10.species.scores <- as.data.frame(scores(outer10_ord3, "species")) # some NaN for absent species
-outer10.species.scores <- na.omit(outer10.species.scores)
-outer10.species.scores$species <- rownames(outer10.species.scores)
-
-outer10.site.scores <- as.data.frame(vegan::scores(outer10_ord3, "sites"))  # extract the site scores and convert to a data.frame
-# add back in metadata
-outer10.site.scores$stratum <- rownames(outer10.site.scores)
-outer10.site.scores <- merge(outer10_subset_meta, outer10.site.scores, by = "stratum")
-
-write_csv(site.scores, here("data", "outer10_site_scores.csv"))
-write_csv(species.scores, here("data", "outer10_species_scores.csv"))
+# set.seed(1032) 
+# outer10_ord3 <- metaMDS(outer10_braymat, dist	= "bray", k = 3, trymax = 100) 
+# outer10_ord3 # 
+# 
+# 
+# sppscores(outer10_ord3) <- outer10_species_matrix # add back in species scores
+# outer10.species.scores <- as.data.frame(scores(outer10_ord3, "species")) # some NaN for absent species
+# outer10.species.scores <- na.omit(outer10.species.scores)
+# outer10.species.scores$species <- rownames(outer10.species.scores) 
+# 
+# 
+# outer10.site.scores <- as.data.frame(vegan::scores(outer10_ord3, "sites"))  # extract the site scores and convert to a data.frame
+# # add back in metadata
+# outer10.site.scores$stratum <- rownames(outer10.site.scores)
+# outer10.site.scores <- merge(outer10_subset_meta, outer10.site.scores, by = "stratum")
+# 
+# 
+# 
+# 
+# write_csv(outer10.site.scores, here("data", "outer10_site_scores.csv"))
+# write_csv(outer10.species.scores, here("data", "outer10_species_scores.csv"))
 
@@
 
-outer17_columns_to_keep <- apply(outer17_species_matrix, 2, function(x) any(x >= 0.1))
-outer17_abundant_species_matrix <- as.data.frame(outer17_species_matrix)[ , outer17_columns_to_keep]
-outer17_abundant_species_names <- colnames(outer17_abundant_species_matrix)
+## outer17_columns_to_keep <- apply(outer17_species_matrix, 2, function(x) any(x >= 0.1))
+## outer17_abundant_species_matrix <- as.data.frame(outer17_species_matrix)[ , outer17_columns_to_keep]
+## outer17_abundant_species_names <- colnames(outer17_abundant_species_matrix)
+## 
+## outer17_subset_meta <- outer17_subset[ , which(names(outer17_subset) %in% meta_cols)]
+## 
+## # for mean centroid:
+## outer17_subset_meta$island_side_year <- paste(outer17_subset_meta$year, "-", outer17_subset_meta$island_side_habitat, sep = "")
+##
 
-outer17_subset_meta <- outer17_subset[ , which(names(outer17_subset) %in% meta_cols)]
-
-# for mean centroid:
-outer17_subset_meta$island_side_year <- paste(outer17_subset_meta$year, "-", outer17_subset_meta$island_side_habitat, sep = "")
-
-# takes several hours to run - do so at your own peril
-set.seed(1032) 
-# ord3 <- metaMDS(outer17_braymat, dist	= "bray", k = 3, trymax = 100) 
-outer17_ord3 # stress = 0.018
-
-sppscores(outer17_ord3) <- outer17_species_matrix # add back in species scores
-outer17.species.scores <- as.data.frame(scores(outer17_ord3, "species")) # some NaN for absent species
-outer17.species.scores <- na.omit(outer17.species.scores)
-outer17.species.scores$species <- rownames(outer17.species.scores)
-
-outer17.site.scores <- as.data.frame(vegan::scores(outer17_ord3, "sites"))  # extract the site scores and convert to a data.frame
-# add back in metadata
-outer17.site.scores$stratum <- rownames(outer17.site.scores)
-outer17.site.scores <- merge(outer17_subset_meta, outer17.site.scores, by = "stratum")
-
-write_csv(site.scores, here("data", "outer17_site_scores.csv"))
-write_csv(species.scores, here("data", "outer17_species_scores.csv"))
+## set.seed(1032) 
+## outer17_ord3 <- metaMDS(outer17_braymat, dist	= "bray", k = 3, trymax = 100) 
+## outer17_ord3 # 
+## 
+## 
+## sppscores(outer17_ord3) <- outer17_species_matrix # add back in species scores
+## outer17.species.scores <- as.data.frame(scores(outer17_ord3, "species")) # some NaN for absent species
+## outer17.species.scores <- na.omit(outer17.species.scores)
+## outer17.species.scores$species <- rownames(outer17.species.scores) 
+## 
+## 
+## outer17.site.scores <- as.data.frame(vegan::scores(outer17_ord3, "sites"))  # extract the site scores and convert to a data.frame
+## # add back in metadata
+## outer17.site.scores$stratum <- rownames(outer17.site.scores)
+## outer17.site.scores <- merge(outer17_subset_meta, outer17.site.scores, by = "stratum")
+## 
+## 
+## 
+## 
+## write_csv(outer17.site.scores, here("data", "outer17_site_scores.csv"))
+## write_csv(outer17.species.scores, here("data", "outer17_species_scores.csv"))
 
@@
 
-```{r full bray curtis set up, eval = FALSE}
-### I DON'T THINK ANY COMPUTER CAN RUN THIS
-species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
-species_matrix <- species_matrix/100 # convert from percent to proportion
-rownames(species_matrix) <- paste(data$year,"-", data$location, sep = "")
-rows_to_remove <- rowSums(species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
-species_matrix <- species_matrix[!rows_to_remove,] 
-data_subset <- data[!rows_to_remove,] # make sure rows are also removed from dataframe
-data_subset$stratum <- paste(data_subset$year,"-", data_subset$location, sep = "") # for merging with bray matrix later
-brayfull <- vegdist(species_matrix, "bray") # create bray dissimilarity matrix
-# View(brayfull)
-braymat <- as.matrix(brayfull) # conver to matrix object
-# convert to pairwise comparison
-comp <- braymat[lower.tri(braymat)] # extract lower triangle
-row_indices <- row(braymat)[lower.tri(braymat)] # extract row indices for first quad in comparison
-col_indices <- col(braymat)[lower.tri(braymat)] # 
-loc1 <- rownames(braymat)[row_indices]
-loc2 <- colnames(braymat)[col_indices]
-pairwise_df <- data.frame(loc1 = loc1, loc2 = loc2, dissimilarity = comp)
-# pairwise_df$year1 <- sub("(.*)-.*", "\\1", pairwise_df$loc1)
-year1 <- strsplit(pairwise_df$loc1, "-")[[1]]
+```{r full ordination}
+### TAKES ~1 wk to run 
+# full_species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
+# full_species_matrix <- full_species_matrix/100 # convert from percent to proportion
+# rownames(full_species_matrix) <- paste(data$year,"-", data$location, sep = "")
+# full_rows_to_remove <- rowSums(full_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
+# full_species_matrix <- full_species_matrix[!full_rows_to_remove,] 
+# full_species_matrix <- asin(sqrt(full_species_matrix)) # arcsin squareroot transform
+# full_subset <- data[!full_rows_to_remove,] # make sure rows are also removed from dataframe
+# full_subset$stratum <- paste(full_subset$year,"-", full_subset$location, sep = "") # for merging with bray matrix later
+# brayfull <- vegdist(full_species_matrix, "bray") # create bray dissimilarity matrix
+# # View(brayfull)
+# full_braymat <- as.matrix(brayfull) # conver to matrix object
+# 
+# full_columns_to_keep <- apply(full_species_matrix, 2, function(x) any(x >= 0.1))
+# full_abundant_species_matrix <- as.data.frame(full_species_matrix)[ , full_columns_to_keep]
+# full_abundant_species_names <- colnames(full_abundant_species_matrix)
+# 
+# full_subset_meta <- full_subset[ , which(names(full_subset) %in% meta_cols)]
+# 
+# # for mean centroid:
+# full_subset_meta$island_side_year <- paste(full_subset_meta$year, "-", full_subset_meta$island_side_habitat, sep = "")
+# 
+# # takes several hours to run - do so at your own peril
+# set.seed(1032) 
+# full_ord3 <- metaMDS(full_braymat, dist	= "bray", k = 3, trymax = 100) 
+# full_ord3 # 
+#
+
+# sppscores(full_ord3) <- full_species_matrix # add back in species scores
+# full.species.scores <- as.data.frame(scores(full_ord3, "species")) # some NaN for absent species
+# full.species.scores <- na.omit(full.species.scores)
+# full.species.scores$species <- rownames(full.species.scores) 
+# 
+# 
+# full.site.scores <- as.data.frame(vegan::scores(full_ord3, "sites"))  # extract the site scores and convert to a data.frame
+# # add back in metadata
+# full.site.scores$stratum <- rownames(full.site.scores)
+# full.site.scores <- merge(full_subset_meta, full.site.scores, by = "stratum")
+# 
+# 
+# 
+# 
+# write_csv(full.site.scores, here("data", "full_site_scores.csv"))
+# write_csv(full.species.scores, here("data", "full_species_scores.csv"))
 

### 2024-08-29 noam_beta_diversity.Rmd

@@
 
-not_algae <- c("ascidian", "bare_space", "coral", "coral_rubble", "corallimorpharia", "heteractis_sp", "millepora_platyphylla", "no_data", "sand", "sacrophyton_sp", "shell_debris", "soft_coral", "sponge", "tridacna_sp", "zooxanthellae", "row_sums")
+not_algae <- c("ascidian", "bare_space", "coral", "coral_rubble", "corallimorpharia", "heteractis_sp", "millepora_platyphylla", "no_data", "sand", "sarcophyton_sp", "shell_debris", "soft_coral", "sponge", "tridacna_sp", "zooxanthellae", "row_sums")
 
@@
 rownames(fringing_species_matrix) <- paste(fringing$year,"-", fringing$location, sep = "")
+
+# filter rare species
+fringing_species_presence <- colSums(fringing_species_matrix > 0) # Presence absence matrix
+fringing_species_percentage <- (fringing_species_presence/nrow(fringing_species_matrix)) # proportion of rows where a species is present
+fringing_common_species <- names(fringing_species_percentage[fringing_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
+fringing_species_matrix <- fringing_species_matrix[,colnames(fringing_species_matrix) %in% fringing_common_species] # filter original species matrix
+
 fringing_rows_to_remove <- rowSums(fringing_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
 fringing_species_matrix <- fringing_species_matrix[!fringing_rows_to_remove,]
-fringing_species_matrix <- asin(sqrt(fringing_species_matrix)) # arcsin squareroot transform
+
+fringing_species_matrix_untransformed <- fringing_species_matrix
+fringing_species_matrix_transformed <- asin(sqrt(fringing_species_matrix)) # arcsin squareroot transform
 fringing_subset <- fringing[!fringing_rows_to_remove,] # make sure rows are also removed from dataframe
 fringing_subset$stratum <- paste(fringing_subset$year,"-", fringing_subset$location, sep = "") # for merging with bray matrix later
-brayfringing <- vegdist(fringing_species_matrix, "bray") # create bray dissimilarity matrix
+
+brayfringing <- vegdist(fringing_species_matrix_transformed, "bray") # create bray dissimilarity matrix
 # View(brayfringing)
+brayfringing_untransformed <- vegdist(fringing_species_matrix_untransformed, "bray")
 fringing_braymat <- as.matrix(brayfringing) # conver to matrix object
@@
 
+fringing_untransformed_braymat <- as.matrix(brayfringing_untransformed) # conver to matrix object
+# convert to pairwise comparison
+fringing_untransformed_comp <- fringing_untransformed_braymat[lower.tri(fringing_untransformed_braymat)] # extract lower triangle
+fringing_untransformed_row_indices <- row(fringing_untransformed_braymat)[lower.tri(fringing_untransformed_braymat)] # extract row indices for first quad in comparison
+fringing_untransformed_col_indices <- col(fringing_untransformed_braymat)[lower.tri(fringing_untransformed_braymat)] # extract col indices = second quad in comp
+fringing_untransformed_loc1 <- rownames(fringing_untransformed_braymat)[fringing_untransformed_row_indices] # populate appropriate vector
+fringing_untransformed_loc2 <- colnames(fringing_untransformed_braymat)[fringing_untransformed_col_indices] # populate appropriate vector
+fringing_untransformed_pairwise_df <- data.frame(loc1 = fringing_untransformed_loc1, loc2 = fringing_untransformed_loc2, dissimilarity = fringing_untransformed_comp)
+fringing_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", fringing_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
+fringing_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", fringing_untransformed_pairwise_df$loc2)) # repeat for second quad
+fringing_untransformed_pairwise_df$time <- abs(fringing_untransformed_pairwise_df$year2 - fringing_untransformed_pairwise_df$year1) # extract length of differences
+fringing_untransformed_pairwise_df$location_1 <- sub(".*-(.*)", "\\1", fringing_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
+fringing_untransformed_pairwise_df$location_2 <- sub(".*-(.*)", "\\1", fringing_untransformed_pairwise_df$loc2) # repeat for second quad
+
 ```
@@
 rownames(backreef_species_matrix) <- paste(backreef$year,"-", backreef$location, sep = "")
+
+# filter rare species
+backreef_species_presence <- colSums(backreef_species_matrix > 0) # Presence absence matrix
+backreef_species_percentage <- (backreef_species_presence/nrow(backreef_species_matrix)) # proportion of rows where a species is present
+backreef_common_species <- names(backreef_species_percentage[backreef_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
+backreef_species_matrix <- backreef_species_matrix[,colnames(backreef_species_matrix) %in% backreef_common_species] # filter original species matrix
+
 backreef_rows_to_remove <- rowSums(backreef_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
 backreef_species_matrix <- backreef_species_matrix[!backreef_rows_to_remove,]
-backreef_species_matrix <- asin(sqrt(backreef_species_matrix)) # arcsin squareroot transform
+
+backreef_species_matrix_untransformed <- backreef_species_matrix
+backreef_species_matrix_transformed <- asin(sqrt(backreef_species_matrix)) # arcsin squareroot transform
 backreef_subset <- backreef[!backreef_rows_to_remove,] # make sure rows are also removed from dataframe
 backreef_subset$stratum <- paste(backreef_subset$year,"-", backreef_subset$location, sep = "") # for merging with bray matrix later
-braybackreef <- vegdist(backreef_species_matrix, "bray") # create bray dissimilarity matrix
+braybackreef <- vegdist(backreef_species_matrix_transformed, "bray") # create bray dissimilarity matrix
 # View(braybackreef)
+braybackreef_untransformed <- vegdist(backreef_species_matrix_untransformed, "bray")
 backreef_braymat <- as.matrix(braybackreef) # conver to matrix object
@@
 
+backreef_untransformed_braymat <- as.matrix(braybackreef_untransformed) # conver to matrix object
+# convert to pairwise comparison
+backreef_untransformed_comp <- backreef_untransformed_braymat[lower.tri(backreef_untransformed_braymat)] # extract lower triangle
+backreef_untransformed_row_indices <- row(backreef_untransformed_braymat)[lower.tri(backreef_untransformed_braymat)] # extract row indices for first quad in comparison
+backreef_untransformed_col_indices <- col(backreef_untransformed_braymat)[lower.tri(backreef_untransformed_braymat)] # extract col indices = second quad in comp
+backreef_untransformed_loc1 <- rownames(backreef_untransformed_braymat)[backreef_untransformed_row_indices] # populate appropriate vector
+backreef_untransformed_loc2 <- colnames(backreef_untransformed_braymat)[backreef_untransformed_col_indices] # populate appropriate vector
+backreef_untransformed_pairwise_df <- data.frame(loc1 = backreef_untransformed_loc1, loc2 = backreef_untransformed_loc2, dissimilarity = backreef_untransformed_comp)
+backreef_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", backreef_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
+backreef_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", backreef_untransformed_pairwise_df$loc2)) # repeat for second quad
+backreef_untransformed_pairwise_df$time <- abs(backreef_untransformed_pairwise_df$year2 - backreef_untransformed_pairwise_df$year1) # extract length of differences
+backreef_untransformed_pairwise_df$location_1 <- sub(".*-(.*)", "\\1", backreef_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
+backreef_untransformed_pairwise_df$location_2 <- sub(".*-(.*)", "\\1", backreef_untransformed_pairwise_df$loc2) # repeat for second quad
+
 ```
@@
 rownames(outer10_species_matrix) <- paste(outer10$year,"-", outer10$location, sep = "")
+
+# filter rare species
+outer10_species_presence <- colSums(outer10_species_matrix > 0) # Presence absence matrix
+outer10_species_percentage <- (outer10_species_presence/nrow(outer10_species_matrix)) # proportion of rows where a species is present
+outer10_common_species <- names(outer10_species_percentage[outer10_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
+outer10_species_matrix <- outer10_species_matrix[,colnames(outer10_species_matrix) %in% outer10_common_species] # filter original species matrix
+
 outer10_rows_to_remove <- rowSums(outer10_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
 outer10_species_matrix <- outer10_species_matrix[!outer10_rows_to_remove,]
-outer10_species_matrix <- asin(sqrt(outer10_species_matrix)) # arcsin squareroot transform
+
+outer10_species_matrix_untransformed <- outer10_species_matrix
+outer10_species_matrix_transformed <- asin(sqrt(outer10_species_matrix)) # arcsin squareroot transform
 outer10_subset <- outer10[!outer10_rows_to_remove,] # make sure rows are also removed from dataframe
 outer10_subset$stratum <- paste(outer10_subset$year,"-", outer10_subset$location, sep = "") # for merging with bray matrix later
-brayouter10 <- vegdist(outer10_species_matrix, "bray") # create bray dissimilarity matrix
+brayouter10 <- vegdist(outer10_species_matrix_transformed, "bray") # create bray dissimilarity matrix
 # View(brayouter10)
+brayouter10_untransformed <- vegdist(outer10_species_matrix_untransformed, "bray")
 outer10_braymat <- as.matrix(brayouter10) # conver to matrix object
@@
 
+outer10_untransformed_braymat <- as.matrix(brayouter10_untransformed) # conver to matrix object
+# convert to pairwise comparison
+outer10_untransformed_comp <- outer10_untransformed_braymat[lower.tri(outer10_untransformed_braymat)] # extract lower triangle
+outer10_untransformed_row_indices <- row(outer10_untransformed_braymat)[lower.tri(outer10_untransformed_braymat)] # extract row indices for first quad in comparison
+outer10_untransformed_col_indices <- col(outer10_untransformed_braymat)[lower.tri(outer10_untransformed_braymat)] # extract col indices = second quad in comp
+outer10_untransformed_loc1 <- rownames(outer10_untransformed_braymat)[outer10_untransformed_row_indices] # populate appropriate vector
+outer10_untransformed_loc2 <- colnames(outer10_untransformed_braymat)[outer10_untransformed_col_indices] # populate appropriate vector
+outer10_untransformed_pairwise_df <- data.frame(loc1 = outer10_untransformed_loc1, loc2 = outer10_untransformed_loc2, dissimilarity = outer10_untransformed_comp)
+outer10_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer10_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
+outer10_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer10_untransformed_pairwise_df$loc2)) # repeat for second quad
+outer10_untransformed_pairwise_df$time <- abs(outer10_untransformed_pairwise_df$year2 - outer10_untransformed_pairwise_df$year1) # extract length of differences
+outer10_untransformed_pairwise_df$location_1 <- sub(".*-(.*)", "\\1", outer10_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
+outer10_untransformed_pairwise_df$location_2 <- sub(".*-(.*)", "\\1", outer10_untransformed_pairwise_df$loc2) # repeat for second quad
+
 ```
@@
 rownames(outer17_species_matrix) <- paste(outer17$year,"-", outer17$location, sep = "")
+
+# filter rare species
+outer17_species_presence <- colSums(outer17_species_matrix > 0) # Presence absence matrix
+outer17_species_percentage <- (outer17_species_presence/nrow(outer17_species_matrix)) # proportion of rows where a species is present
+outer17_common_species <- names(outer17_species_percentage[outer17_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
+outer17_species_matrix <- outer17_species_matrix[,colnames(outer17_species_matrix) %in% outer17_common_species] # filter original species matrix
+
 outer17_rows_to_remove <- rowSums(outer17_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
 outer17_species_matrix <- outer17_species_matrix[!outer17_rows_to_remove,]
-outer17_species_matrix <- asin(sqrt(outer17_species_matrix)) # arcsin squareroot transform
+
+outer17_species_matrix_untransformed <- outer17_species_matrix
+outer17_species_matrix_transformed <- asin(sqrt(outer17_species_matrix)) # arcsin squareroot transform
 outer17_subset <- outer17[!outer17_rows_to_remove,] # make sure rows are also removed from dataframe
 outer17_subset$stratum <- paste(outer17_subset$year,"-", outer17_subset$location, sep = "") # for merging with bray matrix later
-brayouter17 <- vegdist(outer17_species_matrix, "bray") # create bray dissimilarity matrix
+brayouter17 <- vegdist(outer17_species_matrix_transformed, "bray") # create bray dissimilarity matrix
 # View(brayouter17)
+brayouter17_untransformed <- vegdist(outer17_species_matrix_untransformed, "bray")
 outer17_braymat <- as.matrix(brayouter17) # conver to matrix object
@@
 outer17_pairwise_df$location_2 <- sub(".*-(.*)", "\\1", outer17_pairwise_df$loc2) # repeat for second quad
+
+outer17_untransformed_braymat <- as.matrix(brayouter17_untransformed) # conver to matrix object
+# convert to pairwise comparison
+outer17_untransformed_comp <- outer17_untransformed_braymat[lower.tri(outer17_untransformed_braymat)] # extract lower triangle
+outer17_untransformed_row_indices <- row(outer17_untransformed_braymat)[lower.tri(outer17_untransformed_braymat)] # extract row indices for first quad in comparison
+outer17_untransformed_col_indices <- col(outer17_untransformed_braymat)[lower.tri(outer17_untransformed_braymat)] # extract col indices = second quad in comp
+outer17_untransformed_loc1 <- rownames(outer17_untransformed_braymat)[outer17_untransformed_row_indices] # populate appropriate vector
+outer17_untransformed_loc2 <- colnames(outer17_untransformed_braymat)[outer17_untransformed_col_indices] # populate appropriate vector
+outer17_untransformed_pairwise_df <- data.frame(loc1 = outer17_untransformed_loc1, loc2 = outer17_untransformed_loc2, dissimilarity = outer17_untransformed_comp)
+outer17_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer17_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
+outer17_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer17_untransformed_pairwise_df$loc2)) # repeat for second quad
+outer17_untransformed_pairwise_df$time <- abs(outer17_untransformed_pairwise_df$year2 - outer17_untransformed_pairwise_df$year1) # extract length of differences
+outer17_untransformed_pairwise_df$location_1 <- sub(".*-(.*)", "\\1", outer17_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
+outer17_untransformed_pairwise_df$location_2 <- sub(".*-(.*)", "\\1", outer17_untransformed_pairwise_df$loc2) # repeat for second quad
 
@@
 
+fringing_untransformed_pairwise_subset <- fringing_untransformed_pairwise_df[fringing_untransformed_pairwise_df$location_1 == fringing_untransformed_pairwise_df$location_2,]
+backreef_untransformed_pairwise_subset <- backreef_untransformed_pairwise_df[backreef_untransformed_pairwise_df$location_1 == backreef_untransformed_pairwise_df$location_2,]
+outer10_untransformed_pairwise_subset <- outer10_untransformed_pairwise_df[outer10_untransformed_pairwise_df$location_1 == outer10_untransformed_pairwise_df$location_2,]
+outer17_untransformed_pairwise_subset <- outer17_untransformed_pairwise_df[outer17_untransformed_pairwise_df$location_1 == outer17_untransformed_pairwise_df$location_2,]
+
+comb_untransformed_subset <- rbind(fringing_untransformed_pairwise_subset,
+                     backreef_untransformed_pairwise_subset,
+                     outer10_untransformed_pairwise_subset,
+                     outer17_untransformed_pairwise_subset)
+# location_1 and location_2 columns are all the same now
+# create new col just called "location" for merge with other metadata
+comb_untransformed_subset$location <- comb_untransformed_subset$location_1
+bray_untransformed_time_series <- merge(meta_df, comb_untransformed_subset, by = "location")
+
 # model
@@
 # Smithson verkuilen transformation for beta-regression
-sv_trans = function(prop, N, s = 0.000005){
+sv_trans <- function(prop, N, s = 0.000005){
   (prop*(nrow(N) - 1) + s)/nrow(N)
@@
 
+inv_logit <- function(x) {
+  exp(x)/(1+exp(x))
+}
+
 # transformed dissimilarities
 bray_time_series$dissimilarity_trans <- sv_trans(bray_time_series$dissimilarity, bray_time_series)
+bray_untransformed_time_series$dissimilarity_trans <- sv_trans(bray_untransformed_time_series$dissimilarity, bray_time_series)
 
@@
 bray_trends_emm <- emmip(bray_mod_universal, habitat ~ time | island_side, at = list(time = seq(1,16,1)), plotit = F, CIs = T)
-inv_logit <- function(x) {
-  exp(x)/(1+exp(x))
-}
+
+bray_untransformed_mod_universal <- glmmTMB(dissimilarity_trans ~ time*habitat*island_side + (1 | island_side/site/transect/location),
+                              na.action = na.pass, family = beta_family(link = "logit"), data = bray_untransformed_time_series) #
+# model.sel(dredge(bray_untransformed_mod_universal)) # top model as specified
+summary(bray_untransformed_mod_universal) # a fair bit of variance w/i quads/transects
+hist(residuals(bray_untransformed_mod_universal))
+plot(predict(bray_untransformed_mod_universal) ~ residuals(bray_untransformed_mod_universal)) # looks like reflection of logit curve
+
+emtrends(bray_untransformed_mod_universal, pairwise ~ habitat |island_side, "time")
+bray_untransformed_trends_emm <- emmip(bray_untransformed_mod_universal, habitat ~ time | island_side, at = list(time = seq(1,16,1)), plotit = F, CIs = T)
+
 ```
@@
 
+ggplot()+
+   geom_point(data = bray_untransformed_time_series, aes(x = time, y = dissimilarity, colour = island_side), alpha = 0.01 ) +
+   geom_line(data = bray_untransformed_trends_emm, aes(x = time, y = inv_logit(yvar), colour = island_side), linewidth = 2) +
+   geom_ribbon(data = bray_untransformed_trends_emm, aes(x = time, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = island_side), alpha = 0.3) +
+   facet_wrap(~habitat) +
+
+ #  scale_colour_manual(values = island_side_colours) +
+   stat_smooth(se = T, linewidth = 2) +
+   labs(y = "Dissimilarity\n") +
+   theme_bw() + 
+   theme(panel.grid.major = element_blank(), 
+         panel.grid.minor = element_blank(),
+         panel.background = element_blank(),
+         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+         axis.line.y = element_line(colour = "black"),
+         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         axis.line.x = element_line(colour = "black"),
+         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         strip.text.x = element_text(size = 32),
+         legend.key.size = unit(2, "cm"), # increase legend size
+         legend.text = element_text(size = rel(2)), # increase legend text
+         legend.title = element_text(size = rel(2.25)),
+         legend.position = 'bottom', # move legend to bottom
+         legend.box = 'horizontal',
+         legend.box.background = element_rect(colour = "black") # make legend horizontal
+         )
+
 ```
@@
 
-#### WARNING: THIS CHUNK TAKES ~2 DAYS TO RUN! SUGGESTEDF TO USE PROVIDED CSV FILES WITH ORDINATION DATA ######
+#### WARNING: THIS CHUNK TAKES ~2 DAYS TO RUN! SUGGESTED TO USE PROVIDED CSV FILES WITH ORDINATION DATA ######
 
@@
 # 
-# fringing_subset_meta <- fringing_subset[ , which(names(fringing_subset) %in% meta_cols)]
+fringing_subset_meta <- fringing_subset[ , which(names(fringing_subset) %in% meta_cols)]
 # 
 # # for mean centroid:
-# fringing_subset_meta$island_side_year <- paste(fringing_subset_meta$year, "-", fringing_subset_meta$island_side_habitat, sep = "")
+fringing_subset_meta$island_side_year <- paste(fringing_subset_meta$year, "-", fringing_subset_meta$island_side_habitat, sep = "")
 #
 
-# set.seed(1032) 
-# # fringing_ord3 <- metaMDS(fringing_braymat, dist	= "bray", k = 3, trymax = 100) 
-# fringing_ord3 # stress = 0.018
-# 
-# 
-# sppscores(fringing_ord3) <- fringing_species_matrix # add back in species scores
-# fringing.species.scores <- as.data.frame(scores(fringing_ord3, "species")) # some NaN for absent species
-# fringing.species.scores <- na.omit(fringing.species.scores)
-# fringing.species.scores$species <- rownames(fringing.species.scores) 
-# 
-# 
-# fringing.site.scores <- as.data.frame(vegan::scores(fringing_ord3, "sites"))  # extract the site scores and convert to a data.frame
-# # add back in metadata
-# fringing.site.scores$stratum <- rownames(fringing.site.scores)
-# fringing.site.scores <- merge(fringing_subset_meta, fringing.site.scores, by = "stratum")
-# 
-# 
-# 
-# 
-# write_csv(site.scores, here("data", "fringing_site_scores.csv"))
-# write_csv(species.scores, here("data", "fringing_species_scores.csv"))
+set.seed(1032)
+fringing_untransformed_ord3 <- metaMDS(fringing_species_matrix_untransformed, distance = "bray", k = 3)
+fringing_untransformed_ord3 # 0.075
+fringing_untransformed_species_scores <- as.data.frame(scores(fringing_untransformed_ord3, "species"))
+fringing_untransformed_species_scores <- na.omit(fringing_untransformed_species_scores)
+fringing_untransformed_species_scores$species <- rownames(fringing_untransformed_species_scores)
 
-fringing.species.scores <- read.csv(here::here("data", "fringing_species_scores.csv"), stringsAsFactors = F) 
-fringing.site.scores <- read.csv(here::here("data", "fringing_site_scores.csv"), stringsAsFactors = F) 
-# fringing.cent <- read.csv(here::here("data", "fringing_mean_centroid.csv"), stringsAsFactors = F)
+fringing_untransformed_site_scores <- as.data.frame(scores(fringing_untransformed_ord3, "site"))
+fringing_untransformed_site_scores$stratum <- rownames(fringing_untransformed_site_scores)
+fringing_untransformed_site_scores <- merge(fringing_subset_meta, fringing_untransformed_site_scores, by = "stratum")
 
-# fringing.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = fringing.site.scores, FUN = mean) 
-# fringing.cent.meta <- fringing_subset_meta %>%
-#    dplyr::group_by(year, island_side, island_side_habitat, island_side_year, site, habitat, transect, quadrat) %>%
-#    dplyr::summarise(year = unique(year), 
-#                     island_side = unique(island_side),
-#                     island_side_habitat = unique(island_side_habitat),
-#                     island_side_year = unique(island_side_year),
-#                     site = unique(site),
-#                     habitat = unique(habitat),
-#                     transect = unique(transect),
-#                     quadrat = unique(quadrat))
-# fringing.cent <- merge(fringing.cent, fringing_subset_meta[,-which(names(fringing_subset_meta) %in% c("stratum"))],  by = "island_side_year")
+set.seed(1032)
+fringing_transformed_ord3 <- metaMDS(fringing_species_matrix_transformed, distance = "bray", k = 3)
+fringing_transformed_ord3 # 0.101
 
-# subset only the species w/ > 10% cover
-# fringing.abundant.species.scores <- fringing.species.scores[fringing.species.scores$species %in% fringing_abundant_species_names,]
+fringing_transformed_species_scores <- as.data.frame(scores(fringing_transformed_ord3, "species"))
+fringing_transformed_species_scores <- na.omit(fringing_transformed_species_scores)
+fringing_transformed_species_scores$species <- rownames(fringing_transformed_species_scores)
+
+fringing_transformed_site_scores <- as.data.frame(scores(fringing_transformed_ord3, "site"))
+fringing_transformed_site_scores$stratum <- rownames(fringing_transformed_site_scores)
+fringing_transformed_site_scores <- merge(fringing_subset_meta, fringing_transformed_site_scores, by = "stratum")
+plot(fringing_untransformed_ord3) # outlier driven by quadrat dominated by asparagopsis taxiformis 
+plot(fringing_transformed_ord3)
+
+write_csv(fringing_untransformed_species_scores, here("data", "fringing_untransformed_species_rare_removed.csv"))
+write_csv(fringing_untransformed_site_scores, here("data", "fringing_untransformed_site_rare_removed.csv"))
+
+write_csv(fringing_transformed_species_scores, here("data", "fringing_transformed_species_rare_removed.csv"))
+write_csv(fringing_transformed_site_scores, here("data", "fringing_transformed_site_rare_removed.csv"))
+
+# fringing.species.scores <- read.csv(here::here("data", "fringing_species_rare_removed.csv"), stringsAsFactors = F) 
+# fringing.site.scores <- read.csv(here::here("data", "fringing_site_rare_removed.csv"), stringsAsFactors = F)
 
@@
 # 
-# backreef_subset_meta <- backreef_subset[ , which(names(backreef_subset) %in% meta_cols)]
+backreef_subset_meta <- backreef_subset[ , which(names(backreef_subset) %in% meta_cols)]
 # 
 # # for mean centroid:
-# backreef_subset_meta$island_side_year <- paste(backreef_subset_meta$year, "-", backreef_subset_meta$island_side_habitat, sep = "")
+backreef_subset_meta$island_side_year <- paste(backreef_subset_meta$year, "-", backreef_subset_meta$island_side_habitat, sep = "")
 #
-# # takes several hours to run - do so at your own peril
-# set.seed(1032) 
-# backreef_ord3 <- metaMDS(backreef_braymat, dist	= "bray", k = 3, trymax = 100) 
-# backreef_ord3 # 
-# 
-# 
-# sppscores(backreef_ord3) <- backreef_species_matrix # add back in species scores
-# backreef.species.scores <- as.data.frame(scores(backreef_ord3, "species")) # some NaN for absent species
-# backreef.species.scores <- na.omit(backreef.species.scores)
-# backreef.species.scores$species <- rownames(backreef.species.scores) 
-# 
-# 
-# backreef.site.scores <- as.data.frame(vegan::scores(backreef_ord3, "sites"))  # extract the site scores and convert to a data.frame
-# # add back in metadata
-# backreef.site.scores$stratum <- rownames(backreef.site.scores)
-# backreef.site.scores <- merge(backreef_subset_meta, backreef.site.scores, by = "stratum")
-# 
-# 
-# 
-# 
-# write_csv(backreef.site.scores, here("data", "backreef_site_scores.csv"))
-# write_csv(backreef.species.scores, here("data", "backreef_species_scores.csv"))
 
-backreef.species.scores <- read.csv(here::here("data", "backreef_species_scores.csv"), stringsAsFactors = F) 
-backreef.site.scores <- read.csv(here::here("data", "backreef_site_scores.csv"), stringsAsFactors = F) 
-# backreef.cent <- read.csv(here::here("data", "backreef_mean_centroid.csv"), stringsAsFactors = F)
+set.seed(1032)
+backreef_untransformed_ord3 <- metaMDS(backreef_species_matrix_untransformed, distance = "bray", k = 3)
+backreef_untransformed_ord3 # 0.088
+backreef_untransformed_species_scores <- as.data.frame(scores(backreef_untransformed_ord3, "species"))
+backreef_untransformed_species_scores <- na.omit(backreef_untransformed_species_scores)
+backreef_untransformed_species_scores$species <- rownames(backreef_untransformed_species_scores)
 
-backreef.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = backreef.site.scores, FUN = mean) 
-# backreef.cent.meta <- backreef_subset_meta %>%
-#    dplyr::group_by(year, island_side, island_side_habitat, island_side_year, site, habitat, transect, quadrat) %>%
-#    dplyr::summarise(year = unique(year), 
-#                     island_side = unique(island_side),
-#                     island_side_habitat = unique(island_side_habitat),
-#                     island_side_year = unique(island_side_year),
-#                     site = unique(site),
-#                     habitat = unique(habitat),
-#                     transect = unique(transect),
-#                     quadrat = unique(quadrat))
-# backreef.cent <- merge(backreef.cent, backreef_subset_meta[,-which(names(backreef_subset_meta) %in% c("stratum"))],  by = "island_side_year")
+backreef_untransformed_site_scores <- as.data.frame(scores(backreef_untransformed_ord3, "site"))
+backreef_untransformed_site_scores$stratum <- rownames(backreef_untransformed_site_scores)
+backreef_untransformed_site_scores <- merge(backreef_subset_meta, backreef_untransformed_site_scores, by = "stratum")
 
-# subset only the species w/ > 10% cover
-backreef.abundant.species.scores <- backreef.species.scores[backreef.species.scores$species %in% backreef_abundant_species_names,]
+set.seed(1032)
+backreef_transformed_ord3 <- metaMDS(backreef_species_matrix_transformed, distance = "bray", k = 3)
+backreef_transformed_ord3 # 0.121
+
+backreef_transformed_species_scores <- as.data.frame(scores(backreef_transformed_ord3, "species"))
+backreef_transformed_species_scores <- na.omit(backreef_transformed_species_scores)
+backreef_transformed_species_scores$species <- rownames(backreef_transformed_species_scores)
+
+backreef_transformed_site_scores <- as.data.frame(scores(backreef_transformed_ord3, "site"))
+backreef_transformed_site_scores$stratum <- rownames(backreef_transformed_site_scores)
+backreef_transformed_site_scores <- merge(backreef_subset_meta, backreef_transformed_site_scores, by = "stratum")
+plot(backreef_untransformed_ord3) # outlier driven by quadrat dominated by dichotomaria marginata, microdictyon, okamurae peysonnelia, and amphiroa also causing issues
+plot(backreef_transformed_ord3)
+
+write_csv(backreef_untransformed_species_scores, here("data", "backreef_untransformed_species_rare_removed.csv"))
+write_csv(backreef_untransformed_site_scores, here("data", "backreef_untransformed_site_rare_removed.csv"))
+
+write_csv(backreef_transformed_species_scores, here("data", "backreef_transformed_species_rare_removed.csv"))
+write_csv(backreef_transformed_site_scores, here("data", "backreef_transformed_site_rare_removed.csv"))
 
@@
 # 
-# outer10_subset_meta <- outer10_subset[ , which(names(outer10_subset) %in% meta_cols)]
+outer10_subset_meta <- outer10_subset[ , which(names(outer10_subset) %in% meta_cols)]
 # 
 # # for mean centroid:
-# outer10_subset_meta$island_side_year <- paste(outer10_subset_meta$year, "-", outer10_subset_meta$island_side_habitat, sep = "")
+outer10_subset_meta$island_side_year <- paste(outer10_subset_meta$year, "-", outer10_subset_meta$island_side_habitat, sep = "")
 #
 
+set.seed(1032)
+outer10_untransformed_ord3 <- metaMDS(outer10_species_matrix_untransformed, distance = "bray", k = 3)
+outer10_untransformed_ord3 # 0.122 - more normal
+outer10_untransformed_species_scores <- as.data.frame(scores(outer10_untransformed_ord3, "species"))
+outer10_untransformed_species_scores <- na.omit(outer10_untransformed_species_scores)
+outer10_untransformed_species_scores$species <- rownames(outer10_untransformed_species_scores)
+
+outer10_untransformed_site_scores <- as.data.frame(scores(outer10_untransformed_ord3, "site"))
+outer10_untransformed_site_scores$stratum <- rownames(outer10_untransformed_site_scores)
+outer10_untransformed_site_scores <- merge(outer10_subset_meta, outer10_untransformed_site_scores, by = "stratum")
+
+set.seed(1032)
+outer10_transformed_ord3 <- metaMDS(outer10_species_matrix_transformed, distance = "bray", k = 3)
+outer10_transformed_ord3 # 0.174
+
+outer10_transformed_species_scores <- as.data.frame(scores(outer10_transformed_ord3, "species"))
+outer10_transformed_species_scores <- na.omit(outer10_transformed_species_scores)
+outer10_transformed_species_scores$species <- rownames(outer10_transformed_species_scores)
+
+outer10_transformed_site_scores <- as.data.frame(scores(outer10_transformed_ord3, "site"))
+outer10_transformed_site_scores$stratum <- rownames(outer10_transformed_site_scores)
+outer10_transformed_site_scores <- merge(outer10_subset_meta, outer10_transformed_site_scores, by = "stratum")
+plot(outer10_untransformed_ord3) 
+plot(outer10_transformed_ord3) # transformation emphasizes outlier, removes legibility of patterns (driven by sudden dominance of halimeda distorta)
+
+write_csv(outer10_untransformed_species_scores, here("data", "outer10_untransformed_species_rare_removed.csv"))
+write_csv(outer10_untransformed_site_scores, here("data", "outer10_untransformed_site_rare_removed.csv"))
+
+write_csv(outer10_transformed_species_scores, here("data", "outer10_transformed_species_rare_removed.csv"))
+write_csv(outer10_transformed_site_scores, here("data", "outer10_transformed_site_rare_removed.csv"))
+
+#############
+# OUTER 17 ##
+#############
+# outer17_columns_to_keep <- apply(outer17_species_matrix, 2, function(x) any(x >= 0.1))
+# outer17_abundant_species_matrix <- as.data.frame(outer17_species_matrix)[ , outer17_columns_to_keep]
+# outer17_abundant_species_names <- colnames(outer17_abundant_species_matrix)
+# 
+outer17_subset_meta <- outer17_subset[ , which(names(outer17_subset) %in% meta_cols)]
+# 
+# # for mean centroid:
+outer17_subset_meta$island_side_year <- paste(outer17_subset_meta$year, "-", outer17_subset_meta$island_side_habitat, sep = "")
+#
+
+set.seed(1032)
+outer17_untransformed_ord3 <- metaMDS(outer17_species_matrix_untransformed, distance = "bray", k = 3)
+outer17_untransformed_ord3 # 0.109
+outer17_untransformed_species_scores <- as.data.frame(scores(outer17_untransformed_ord3, "species"))
+outer17_untransformed_species_scores <- na.omit(outer17_untransformed_species_scores)
+outer17_untransformed_species_scores$species <- rownames(outer17_untransformed_species_scores)
+
+outer17_untransformed_site_scores <- as.data.frame(scores(outer17_untransformed_ord3, "site"))
+outer17_untransformed_site_scores$stratum <- rownames(outer17_untransformed_site_scores)
+outer17_untransformed_site_scores <- merge(outer17_subset_meta, outer17_untransformed_site_scores, by = "stratum")
+
+set.seed(1032)
+outer17_transformed_ord3 <- metaMDS(outer17_species_matrix_transformed, distance = "bray", k = 3)
+outer17_transformed_ord3 # 0.148 - more stress in transformed matrix?
+
+outer17_transformed_species_scores <- as.data.frame(scores(outer17_transformed_ord3, "species"))
+outer17_transformed_species_scores <- na.omit(outer17_transformed_species_scores)
+outer17_transformed_species_scores$species <- rownames(outer17_transformed_species_scores)
+
+outer17_transformed_site_scores <- as.data.frame(scores(outer17_transformed_ord3, "site"))
+outer17_transformed_site_scores$stratum <- rownames(outer17_transformed_site_scores)
+outer17_transformed_site_scores <- merge(outer17_subset_meta, outer17_transformed_site_scores, by = "stratum")
+plot(outer17_untransformed_ord3) 
+plot(outer17_transformed_ord3) # trasnsformation removes arch pattern - probably good
+
+write_csv(outer17_untransformed_species_scores, here("data", "outer17_untransformed_species_rare_removed.csv"))
+write_csv(outer17_untransformed_site_scores, here("data", "outer17_untransformed_site_rare_removed.csv"))
+
+write_csv(outer17_transformed_species_scores, here("data", "outer17_transformed_species_rare_removed.csv"))
+write_csv(outer17_transformed_site_scores, here("data", "outer17_transformed_site_rare_removed.csv"))
+
+```
+
+```{r plotting individually}
+fringing.site.scores <- read_csv(here("data", "fringing_transformed_site_rare_removed.csv")) 
+backreef.site.scores <- read_csv(here("data", "backreef_transformed_site_rare_removed.csv")) 
+outer10.site.scores <- read_csv(here("data", "outer10_transformed_site_rare_removed.csv")) 
+outer17.site.scores <- read_csv(here("data", "outer17_transformed_site_rare_removed.csv"))
+
+fringing.species.scores <- read_csv(here("data", "fringing_transformed_species_rare_removed.csv")) 
+backreef.species.scores <- read_csv(here("data", "backreef_transformed_species_rare_removed.csv")) 
+outer10.species.scores <- read_csv(here("data", "outer10_transformed_species_rare_removed.csv")) 
+outer17.species.scores <- read_csv(here("data", "outer17_transformed_species_rare_removed.csv"))
+
+fringing_subset_meta_condensed <- fringing_subset_meta %>%
+  group_by(year, habitat, island_side, island_side_habitat, island_side_year) %>%
+  summarize() %>%
+  ungroup()
+fringing.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = fringing.site.scores, FUN = mean) 
+fringing.cent <- merge(fringing.cent, fringing_subset_meta_condensed, by = "island_side_year")
+
+# read in disturbances
+disturbances <- read.csv(here("data", "disturbance_list.csv"))
+fringing.cent <- merge(fringing.cent, disturbances, by = "year") # add disturbances to path
+
+backreef_subset_meta_condensed <- backreef_subset_meta %>%
+  group_by(year, habitat, island_side, island_side_habitat, island_side_year) %>%
+  summarize() %>%
+  ungroup()
+backreef.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = backreef.site.scores, FUN = mean) 
+backreef.cent <- merge(backreef.cent, backreef_subset_meta_condensed, by = "island_side_year")
+backreef.cent <- merge(backreef.cent, disturbances, by = "year") # add disturbances to path
+
+outer10_subset_meta_condensed <- outer10_subset_meta %>%
+  group_by(year, habitat, island_side, island_side_habitat, island_side_year) %>%
+  summarize() %>%
+  ungroup()
+outer10.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = outer10.site.scores, FUN = mean) 
+outer10.cent <- merge(outer10.cent, outer10_subset_meta_condensed, by = "island_side_year")
+outer10.cent <- merge(outer10.cent, disturbances, by = "year") # add disturbances to path
+
+outer17_subset_meta_condensed <- outer17_subset_meta %>%
+  group_by(year, habitat, island_side, island_side_habitat, island_side_year) %>%
+  summarize() %>%
+  ungroup()
+outer17.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = outer17.site.scores, FUN = mean) 
+outer17.cent <- merge(outer17.cent, outer17_subset_meta_condensed, by = "island_side_year")
+outer17.cent <- merge(outer17.cent, disturbances, by = "year") # add disturbances to path
+
+library(ggrepel)
+(fringing_nmds_plot <-
+  ggplot() + 
+  geom_point(data= fringing.site.scores, aes(x=NMDS1,y=NMDS2, colour = island_side), alpha = 0.05) + # add the point markers
+  geom_text_repel(data= fringing.species.scores,aes(x=NMDS1,y=NMDS2, label = species), size = 4, alpha = 0.5) +  # add the species labels
+#  
+  geom_path(data = fringing.cent, aes(x = NMDS1, y = NMDS2, group = island_side_habitat, colour = island_side),
+            arrow = arrow(length=unit(0.25,"cm"), ends="last", type = "closed"), size = 1) +
+  geom_point(data = fringing.cent, aes(x = NMDS1, y = NMDS2, group = island_side_habitat, colour = island_side, shape = disturbance), size = 3, stroke = 1.4) +
+  scale_shape_manual(values = c(
+    "cots" = 8,
+    "cyclone" = 13,
+    "bleaching" = 4
+  )) +
+  labs(title = "a. Fringing") + 
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        strip.text.x = element_text(size = 25),
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.key.size = unit(2, "cm"), # increase legend size
+
+        legend.text = element_text(size = rel(1.5)), # increase legend text
+        legend.title = element_text(size = rel(1.5)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) + # add box around legend
+  guides(colour=guide_legend(title.position="top", # make color legend title go above legend
+                                     title.hjust =0.5, nrow = 2),
+         shape=guide_legend(title.position="top",  # make shape legend title go above legend
+                                     title.hjust =0.5,
+                            nrow = 2),
+         path = guide_legend(title.position="top",  # make shape legend title go above legend
+                                     title.hjust =0.5)
+         )  )
+
+(backreef_nmds_plot <-
+  ggplot() + 
+  geom_point(data= backreef.site.scores, aes(x=NMDS1,y=NMDS2, colour = island_side), alpha = 0.05) + # add the point markers
+  geom_text_repel(data= backreef.species.scores,aes(x=NMDS1,y=NMDS2, label = species), size = 4, alpha = 0.5) +  # add the species labels
+#  
+  geom_path(data = backreef.cent, aes(x = NMDS1, y = NMDS2, group = island_side_habitat, colour = island_side),
+            arrow = arrow(length=unit(0.25,"cm"), ends="last", type = "closed"), size = 1) +
+  geom_point(data = backreef.cent, aes(x = NMDS1, y = NMDS2, group = island_side_habitat, colour = island_side, shape = disturbance), size = 3, stroke = 1.4) +
+  scale_shape_manual(values = c(
+    "cots" = 8,
+    "cyclone" = 13,
+    "bleaching" = 4
+  )) +
+  labs(title = "b. backreef") + 
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        strip.text.x = element_text(size = 25),
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.key.size = unit(2, "cm"), # increase legend size
+
+        legend.text = element_text(size = rel(1.5)), # increase legend text
+        legend.title = element_text(size = rel(1.5)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) + # add box around legend
+  guides(colour=guide_legend(title.position="top", # make color legend title go above legend
+                                     title.hjust =0.5, nrow = 2),
+         shape=guide_legend(title.position="top",  # make shape legend title go above legend
+                                     title.hjust =0.5,
+                            nrow = 2),
+         path = guide_legend(title.position="top",  # make shape legend title go above legend
+                                     title.hjust =0.5)
+         )  )
+
+(outer10_nmds_plot <-
+  ggplot() + 
+  geom_point(data= outer10.site.scores, aes(x=NMDS1,y=NMDS2, colour = island_side), alpha = 0.05) + # add the point markers
+  geom_text_repel(data= outer10.species.scores,aes(x=NMDS1,y=NMDS2, label = species), size = 4, alpha = 0.5) +  # add the species labels
+#  
+  geom_path(data = outer10.cent, aes(x = NMDS1, y = NMDS2, group = island_side_habitat, colour = island_side),
+            arrow = arrow(length=unit(0.25,"cm"), ends="last", type = "closed"), size = 1) +
+  geom_point(data = outer10.cent, aes(x = NMDS1, y = NMDS2, group = island_side_habitat, colour = island_side, shape = disturbance), size = 3, stroke = 1.4) +
+  scale_shape_manual(values = c(
+    "cots" = 8,
+    "cyclone" = 13,
+    "bleaching" = 4
+  )) +
+  labs(title = "c. outer10") + 
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        strip.text.x = element_text(size = 25),
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.key.size = unit(2, "cm"), # increase legend size
+
+        legend.text = element_text(size = rel(1.5)), # increase legend text
+        legend.title = element_text(size = rel(1.5)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) + # add box around legend
+  guides(colour=guide_legend(title.position="top", # make color legend title go above legend
+                                     title.hjust =0.5, nrow = 2),
+         shape=guide_legend(title.position="top",  # make shape legend title go above legend
+                                     title.hjust =0.5,
+                            nrow = 2),
+         path = guide_legend(title.position="top",  # make shape legend title go above legend
+                                     title.hjust =0.5)
+         )  )
+
+(outer17_nmds_plot <-
+  ggplot() + 
+  geom_point(data= outer17.site.scores, aes(x=NMDS1,y=NMDS2, colour = island_side), alpha = 0.05) + # add the point markers
+  geom_text_repel(data= outer17.species.scores,aes(x=NMDS1,y=NMDS2, label = species), size = 4, alpha = 0.5) +  # add the species labels
+#  
+  geom_path(data = outer17.cent, aes(x = NMDS1, y = NMDS2, group = island_side_habitat, colour = island_side),
+            arrow = arrow(length=unit(0.25,"cm"), ends="last", type = "closed"), size = 1) +
+  geom_point(data = outer17.cent, aes(x = NMDS1, y = NMDS2, group = island_side_habitat, colour = island_side, shape = disturbance), size = 3, stroke = 1.4) +
+  scale_shape_manual(values = c(
+    "cots" = 8,
+    "cyclone" = 13,
+    "bleaching" = 4
+  )) +
+  labs(title = "d. outer17") + 
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        strip.text.x = element_text(size = 25),
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.key.size = unit(2, "cm"), # increase legend size
+        legend.text = element_text(size = rel(1.5)), # increase legend text
+        legend.title = element_text(size = rel(1.5)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) + # add box around legend
+  guides(colour=guide_legend(title.position="top", # make color legend title go above legend
+                                     title.hjust =0.5, nrow = 2),
+         shape=guide_legend(title.position="top",  # make shape legend title go above legend
+                                     title.hjust =0.5,
+                            nrow = 2),
+         path = guide_legend(title.position="top",  # make shape legend title go above legend
+                                     title.hjust =0.5)
+         )  )
+```
+```{r panel graph, fig.height= 16, fig.width= 15}
+ggarrange(fringing_nmds_plot, backreef_nmds_plot, outer10_nmds_plot, outer17_nmds_plot, nrow = 2, ncol = 2, common.legend = T)
+
+```
+
+```{r full ordination setup}
+
+# get names of quads that remained after removing rare species
+relevant_quads <- c(rownames(fringing_species_matrix_untransformed), rownames(backreef_species_matrix_untransformed),
+                    rownames(outer10_species_matrix_untransformed), rownames(outer17_species_matrix_untransformed))
+
+full_species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
+full_species_matrix <- full_species_matrix/100 # convert from percent to proportion
+rownames(full_species_matrix) <- paste(data$year,"-", data$location, sep = "")
+
+full_species_matrix <- full_species_matrix[rownames(full_species_matrix) %in% relevant_quads,] # subset so only quads that didn't contain rare species are represented
+full_species_matrix <- asin(sqrt(full_species_matrix)) # arcsin squareroot transform
+data$stratum <- paste(data$year,"-", data$location, sep = "") # for merging with bray matrix later
+full_subset <- data[data$stratum %in% relevant_quads,] # make sure rows are also removed from dataframe
+
+brayfull <- vegdist(full_species_matrix, "bray") # create bray dissimilarity matrix
+# View(brayfull)
+full_braymat <- as.matrix(brayfull) # conver to matrix object
+
+full_columns_to_keep <- apply(full_species_matrix, 2, function(x) any(x >= asin(sqrt(0.1) )))
+full_abundant_species_matrix <- as.data.frame(full_species_matrix)[ , full_columns_to_keep]
+full_abundant_species_names <- colnames(full_abundant_species_matrix)
+
+full_subset_meta <- full_subset[ , which(names(full_subset) %in% meta_cols)]
+
+# for mean centroid:
+full_subset_meta$island_side_year <- paste(full_subset_meta$year, "-", full_subset_meta$island_side_habitat, sep = "")
+
+full_subset_meta_condensed <- full_subset_meta %>%
+  group_by(year, habitat, island_side, island_side_habitat, island_side_year) %>%
+  summarize() %>%
+  ungroup()
+
+### TAKES ~1 wk to run 
 # set.seed(1032) 
-# outer10_ord3 <- metaMDS(outer10_braymat, dist	= "bray", k = 3, trymax = 100) 
-# outer10_ord3 # 
+# full_ord3 <- metaMDS(full_braymat, dist	= "bray", k = 3, trymax = 100) 
+# full_ord3 # 0.03 seems suspect
+#
+
+# sppscores(full_ord3) <- full_species_matrix # add back in species scores
+# full.species.scores <- as.data.frame(scores(full_ord3, "species")) # some NaN for absent species
+# full.species.scores <- na.omit(full.species.scores)
+# full.species.scores$species <- rownames(full.species.scores) 
 # 
 # 
-# sppscores(outer10_ord3) <- outer10_species_matrix # add back in species scores
-# outer10.species.scores <- as.data.frame(scores(outer10_ord3, "species")) # some NaN for absent species
-# outer10.species.scores <- na.omit(outer10.species.scores)
-# outer10.species.scores$species <- rownames(outer10.species.scores) 
-# 
-# 
-# outer10.site.scores <- as.data.frame(vegan::scores(outer10_ord3, "sites"))  # extract the site scores and convert to a data.frame
+# full.site.scores <- as.data.frame(vegan::scores(full_ord3, "sites"))  # extract the site scores and convert to a data.frame
 # # add back in metadata
-# outer10.site.scores$stratum <- rownames(outer10.site.scores)
-# outer10.site.scores <- merge(outer10_subset_meta, outer10.site.scores, by = "stratum")
+# full.site.scores$stratum <- rownames(full.site.scores)
+# full.site.scores <- merge(full_subset_meta, full.site.scores, by = "stratum")
 # 
@@
 # 
-# write_csv(outer10.site.scores, here("data", "outer10_site_scores.csv"))
-# write_csv(outer10.species.scores, here("data", "outer10_species_scores.csv"))
+# write_csv(full.site.scores, here("data", "full_site_scores.csv"))
+# write_csv(full.species.scores, here("data", "full_species_scores.csv"))
 
-outer10.species.scores <- read.csv(here::here("data", "outer10_species_scores.csv"), stringsAsFactors = F) 
-outer10.site.scores <- read.csv(here::here("data", "outer10_site_scores.csv"), stringsAsFactors = F) 
-# outer10.cent <- read.csv(here::here("data", "outer10_mean_centroid.csv"), stringsAsFactors = F)
-
-outer10.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = outer10.site.scores, FUN = mean) 
-# outer10.cent.meta <- outer10_subset_meta %>%
-#    dplyr::group_by(year, island_side, island_side_habitat, island_side_year, site, habitat, transect, quadrat) %>%
-#    dplyr::summarise(year = unique(year), 
-#                     island_side = unique(island_side),
-#                     island_side_habitat = unique(island_side_habitat),
-#                     island_side_year = unique(island_side_year),
-#                     site = unique(site),
-#                     habitat = unique(habitat),
-#                     transect = unique(transect),
-#                     quadrat = unique(quadrat))
-# outer10.cent <- merge(outer10.cent, outer10_subset_meta[,-which(names(outer10_subset_meta) %in% c("stratum"))],  by = "island_side_year")
-
-# subset only the species w/ > 10% cover
-outer10.abundant.species.scores <- outer10.species.scores[outer10.species.scores$species %in% outer10_abundant_species_names,]
-
-#############
-# OUTER 17 ##
-#############
-
-## outer17_columns_to_keep <- apply(outer17_species_matrix, 2, function(x) any(x >= 0.1))
-## outer17_abundant_species_matrix <- as.data.frame(outer17_species_matrix)[ , outer17_columns_to_keep]
-## outer17_abundant_species_names <- colnames(outer17_abundant_species_matrix)
-## 
-## outer17_subset_meta <- outer17_subset[ , which(names(outer17_subset) %in% meta_cols)]
-## 
-## # for mean centroid:
-## outer17_subset_meta$island_side_year <- paste(outer17_subset_meta$year, "-", outer17_subset_meta$island_side_habitat, sep = "")
-##
-
-## set.seed(1032) 
-## outer17_ord3 <- metaMDS(outer17_braymat, dist	= "bray", k = 3, trymax = 100) 
-## outer17_ord3 # 
-## 
-## 
-## sppscores(outer17_ord3) <- outer17_species_matrix # add back in species scores
-## outer17.species.scores <- as.data.frame(scores(outer17_ord3, "species")) # some NaN for absent species
-## outer17.species.scores <- na.omit(outer17.species.scores)
-## outer17.species.scores$species <- rownames(outer17.species.scores) 
-## 
-## 
-## outer17.site.scores <- as.data.frame(vegan::scores(outer17_ord3, "sites"))  # extract the site scores and convert to a data.frame
-## # add back in metadata
-## outer17.site.scores$stratum <- rownames(outer17.site.scores)
-## outer17.site.scores <- merge(outer17_subset_meta, outer17.site.scores, by = "stratum")
-## 
-## 
-## 
-## 
-## write_csv(outer17.site.scores, here("data", "outer17_site_scores.csv"))
-## write_csv(outer17.species.scores, here("data", "outer17_species_scores.csv"))
-
-outer17.species.scores <- read.csv(here::here("data", "outer17_species_scores.csv"), stringsAsFactors = F) 
-outer17.site.scores <- read.csv(here::here("data", "outer17_site_scores.csv"), stringsAsFactors = F) 
-# outer17.cent <- read.csv(here::here("data", "outer17_mean_centroid.csv"), stringsAsFactors = F)
-
-outer17.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = outer17.site.scores, FUN = mean) 
-# outer17.cent.meta <- outer17_subset_meta %>%
-#    dplyr::group_by(year, island_side, island_side_habitat, island_side_year, site, habitat, transect, quadrat) %>%
-#    dplyr::summarise(year = unique(year), 
-#                     island_side = unique(island_side),
-#                     island_side_habitat = unique(island_side_habitat),
-#                     island_side_year = unique(island_side_year),
-#                     site = unique(site),
-#                     habitat = unique(habitat),
-#                     transect = unique(transect),
-#                     quadrat = unique(quadrat))
-# outer17.cent <- merge(outer17.cent, outer17_subset_meta[,-which(names(outer17_subset_meta) %in% c("stratum"))],  by = "island_side_year")
-
-# subset only the species w/ > 10% cover
-outer17.abundant.species.scores <- outer17.species.scores[outer17.species.scores$species %in% outer17_abundant_species_names,]
 ```
 
-```{r plot fringing}
+```{r full ordination, fig.height = 15, fig.width=15}
+
+full.site.scores <- read_csv(here("data", "full_site_scores.csv"))
+
+hist(subset(full.site.scores, habitat == "fringing")$NMDS1)
+
+# option 1 for removing outliers - remove 1.5*IQR identified by boxplot
+full.site.scores.outliers.nmds1 <- boxplot(full.site.scores$NMDS1, plot = F)$out
+full.site.scores.outliers.nmds2 <- boxplot(full.site.scores$NMDS2, plot = F)$out
+full.site.scores.outliers.removed.method1 <- full.site.scores[!full.site.scores$NMDS1 %in% full.site.scores.outliers.nmds1 &
+                                                      !full.site.scores$NMDS2 %in% full.site.scores.outliers.nmds2, ]
+
+# option 2 - triple IQR
+nmds1.iqr <- IQR(full.site.scores$NMDS1)
+nmds1.lq <-unname(quantile(full.site.scores$NMDS1)[2])
+nmds1.uq <-unname(quantile(full.site.scores$NMDS1)[4])
+full.site.scores.outliers.nmds1.method2 <- full.site.scores[3*nmds1.iqr + nmds1.uq < full.site.scores$NMDS1 & 
+                                                             full.site.scores$NMDS1 > nmds1.lq - 3*nmds1.iqr, ]
+
+nmds2.iqr <- IQR(full.site.scores$NMDS2)
+nmds2.lq <-unname(quantile(full.site.scores$NMDS2)[2])
+nmds2.uq <-unname(quantile(full.site.scores$NMDS2)[4])
+full.site.scores.outliers.nmds2.method2 <- full.site.scores[3*nmds2.iqr + nmds2.uq < full.site.scores$NMDS2 & 
+                                                             full.site.scores$NMDS2 > nmds2.lq - 3*nmds2.iqr, ]
+
+full.site.scores.outliers.removed.method2 <- anti_join(full.site.scores, rbind(full.site.scores.outliers.nmds1.method2, full.site.scores.outliers.nmds2.method2))
+problem.quads <- rbind(full.site.scores.outliers.nmds1.method2, full.site.scores.outliers.nmds2.method2)$stratum
+
+# option 3 - go at habitat level and manually tweak based on boxplots, histograms, and ordinations
+# Fringing
+boxplot(subset(full.site.scores, habitat == "fringing")$NMDS2)
+fringing.outliers.nmds1 <- subset(full.site.scores, habitat == "fringing")[subset(full.site.scores, habitat == "fringing")$NMDS1 > 0.001 | 
+                                                                            subset(full.site.scores, habitat == "fringing")$NMDS1 < -0.001, ]
+fringing.outliers.nmds2 <- subset(full.site.scores, habitat == "fringing")[subset(full.site.scores, habitat == "fringing")$NMDS2 < -3e-04 |
+                                                                             subset(full.site.scores, habitat == "fringing")$NMDS2 > 3e-04,]
+
+# Backreef
+boxplot(subset(full.site.scores, habitat == "backreef")$NMDS1) # anything above 0.1 or below -0.1 should go
+boxplot(subset(full.site.scores, habitat == "backreef")$NMDS2) # anything below -0.01 or above 0.02
+
+backreef.outliers.nmds1 <- subset(full.site.scores, habitat == "backreef")[subset(full.site.scores, habitat == "backreef")$NMDS1 > 0.1 | 
+                                                                            subset(full.site.scores, habitat == "backreef")$NMDS1 < -0.1, ]
+backreef.outliers.nmds2 <- subset(full.site.scores, habitat == "backreef")[subset(full.site.scores, habitat == "backreef")$NMDS2 > 0.02 | 
+                                                                            subset(full.site.scores, habitat == "backreef")$NMDS2 < -0.0005, ]
+
+# Outer10
+boxplot(subset(full.site.scores, habitat == "outer_10")$NMDS1) # anything above 0.001 
+boxplot(subset(full.site.scores, habitat == "outer_10")$NMDS2) # anything above 0.005
+
+outer10.outliers.nmds1 <- subset(full.site.scores, habitat == "outer_10")[subset(full.site.scores, habitat == "outer_10")$NMDS1 > 0.001,]
+outer10.outliers.nmds2 <- subset(full.site.scores, habitat == "outer_10")[subset(full.site.scores, habitat == "outer_10")$NMDS2 > 0.001,]
+
+# Outer17
+boxplot(subset(full.site.scores, habitat == "outer_17")$NMDS1) # potentially above 0.00015 but these all look fine to me 
+boxplot(subset(full.site.scores, habitat == "outer_17")$NMDS2) # nothing too troublesome
+
+full.site.outliers <- rbind(fringing.outliers.nmds1, fringing.outliers.nmds2, backreef.outliers.nmds1, backreef.outliers.nmds2, outer10.outliers.nmds1, outer10.outliers.nmds2)
+problem.quads <- full.site.outliers$stratum
+full.site.scores.outliers.removed.method3 <- full.site.scores[!full.site.scores$stratum %in% problem.quads, ]
+
+## Option 4 - only sites that did not contain rare species
+full.site.scores.outliers.removed.method4 <- full.site.scores[full.site.scores$stratum %in% relevant_quads,]
+
+full.species.scores <- read_csv(here("data", "full_species_scores.csv"))
+
+full.abundant.species.scores <- full.species.scores[full.species.scores$species %in% full_abundant_species_names,]
+
+full.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = full.site.scores.outliers.removed.method4, FUN = mean) 
+full.cent <- merge(full.cent, full_subset_meta_condensed, by = "island_side_year")
+
+# read in disturbances
+disturbances <- read.csv(here("data", "disturbance_list.csv"))
+full.cent <- merge(full.cent, disturbances, by = "year") # add disturbances to path
+library(ggrepel)
 ggplot() + 
-  geom_point(data= site.scores, aes(x=NMDS1,y=NMDS2, colour = island_side),
-             alpha = 0.25, size=2) + # add the point markers
-  geom_text(data= abundant.species.scores,aes(x=NMDS1,y=NMDS2, label = species), size = 4) +  # add the species labels
+  geom_point(data= full.site.scores.outliers.removed.method4, aes(x=NMDS1,y=NMDS2, colour = island_side), alpha = 0.05) + # add the point markers
+  geom_text_repel(data= full.abundant.species.scores,aes(x=NMDS1,y=NMDS2, label = species), size = 4, alpha = 0.5) +  # add the species labels
 
-  geom_path(data = cent, aes(x = NMDS1, y = NMDS2, group = island_side_habitat, colour = island_side),
-            arrow = arrow(length=unit(0.5,"cm"), ends="last", type = "closed"), size = 1) +
-#  facet_wrap(~factor(habitat, levels=c('fringing','backreef','outer_10','outer_17'))) +
-  scale_linetype_manual(values = habitat_linetypes) +
-  scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
+  geom_path(data = full.cent, aes(x = NMDS1, y = NMDS2, group = island_side_habitat, colour = island_side),
+            arrow = arrow(length=unit(0.25,"cm"), ends="last", type = "closed"), size = 1) +
+  geom_point(data = full.cent, aes(x = NMDS1, y = NMDS2, group = island_side_habitat, colour = island_side, shape = disturbance), size = 3, stroke = 1.4) +
+  scale_shape_manual(values = c(
+    "cots" = 8,
+    "cyclone" = 13,
+    "bleaching" = 4
+  )) +
+  scale_y_continuous(limits = c(-2.5e-04,2.5e-04)) +
+  scale_x_continuous(limits = c(-1.75e-04, 1.75e-04)) +
+  facet_wrap(~factor(habitat, levels=c('fringing','backreef','outer_10','outer_17'))) +
+# scale_linetype_manual(values = habitat_linetypes) +
+#  scale_colour_manual(values = site_colours) +
+#  scale_shape_manual(values = habitat_shapes) +
   theme_bw() + 
@@
          )
-
-```
-
-```{r full ordination}
-### TAKES ~1 wk to run 
-# full_species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
-# full_species_matrix <- full_species_matrix/100 # convert from percent to proportion
-# rownames(full_species_matrix) <- paste(data$year,"-", data$location, sep = "")
-# full_rows_to_remove <- rowSums(full_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
-# full_species_matrix <- full_species_matrix[!full_rows_to_remove,] 
-# full_species_matrix <- asin(sqrt(full_species_matrix)) # arcsin squareroot transform
-# full_subset <- data[!full_rows_to_remove,] # make sure rows are also removed from dataframe
-# full_subset$stratum <- paste(full_subset$year,"-", full_subset$location, sep = "") # for merging with bray matrix later
-# brayfull <- vegdist(full_species_matrix, "bray") # create bray dissimilarity matrix
-# # View(brayfull)
-# full_braymat <- as.matrix(brayfull) # conver to matrix object
-# 
-# full_columns_to_keep <- apply(full_species_matrix, 2, function(x) any(x >= 0.1))
-# full_abundant_species_matrix <- as.data.frame(full_species_matrix)[ , full_columns_to_keep]
-# full_abundant_species_names <- colnames(full_abundant_species_matrix)
-# 
-# full_subset_meta <- full_subset[ , which(names(full_subset) %in% meta_cols)]
-# 
-# # for mean centroid:
-# full_subset_meta$island_side_year <- paste(full_subset_meta$year, "-", full_subset_meta$island_side_habitat, sep = "")
-# 
-# # takes several hours to run - do so at your own peril
-# set.seed(1032) 
-# full_ord3 <- metaMDS(full_braymat, dist	= "bray", k = 3, trymax = 100) 
-# full_ord3 # 
-#
-
-# sppscores(full_ord3) <- full_species_matrix # add back in species scores
-# full.species.scores <- as.data.frame(scores(full_ord3, "species")) # some NaN for absent species
-# full.species.scores <- na.omit(full.species.scores)
-# full.species.scores$species <- rownames(full.species.scores) 
-# 
-# 
-# full.site.scores <- as.data.frame(vegan::scores(full_ord3, "sites"))  # extract the site scores and convert to a data.frame
-# # add back in metadata
-# full.site.scores$stratum <- rownames(full.site.scores)
-# full.site.scores <- merge(full_subset_meta, full.site.scores, by = "stratum")
-# 
-# 
-# 
-# 
-# write_csv(full.site.scores, here("data", "full_site_scores.csv"))
-# write_csv(full.species.scores, here("data", "full_species_scores.csv"))
-
-```
-
-```{r old nmds code, fig.height = 15, fig.width = 15, eval = FALSE}
-species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
-species_matrix <- species_matrix/100 # convert from percent to proportion
-rownames(species_matrix) <- paste(data$year, data$site_habitat, data$transect)
-rows_to_remove <- rowSums(species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
-species_matrix <- species_matrix[!rows_to_remove,] 
-data_subset <- data[!rows_to_remove,] # make sure rows are also removed from dataframe
-data_subset$stratum <- paste(data_subset$year, "-", data_subset$site, "-", data_subset$habitat )
-# identifying species w/ > 10% cover for plotting later
-# NOTE: Used ChatGPT to generate these next few lines of code but it looked good so I went with it
-columns_to_keep <- apply(species_matrix, 2, function(x) any(x >= 0.1))
-abundant_species_matrix <- as.data.frame(species_matrix)[ , c(TRUE, columns_to_keep)]
-abundant_species_names <- colnames(abundant_species_matrix)
-
-# My computer can't run this at the quadrat level
-
-# ord2 # stress = 0.22
-# stressplot(ord2)
-
-set.seed(1032)
-ord3 <- metaMDS(species_matrix, dist	= "bray",	trymax	= 100,	k	= 3)
-ord3 # 0.166
-stressplot(ord3)
-
-# pro <- procrustes(ord2, ord3) # visually compare w procrustes rotation
-
-# plot(pro, cex=1.5)
-
-site.scores <- as.data.frame(vegan::scores(ord3)$sites)  # extract the site scores and convert to a data.frame
-# add back in metadata
-site.scores$year <- as.factor(data_subset$year)
-site.scores$site <- as.factor(data_subset$site)
-site.scores$habitat <- as.factor(data_subset$habitat)
-site.scores$site_habitat <- as.factor(data_subset$site_habitat)
-site.scores$transect <- as.factor(data_subset$transect)
-site.scores$stratum <- as.factor(data_subset$stratum)
-
-cent <- aggregate(cbind(NMDS1, NMDS2) ~ stratum, data = site.scores, FUN = mean) # compute centroids for NMDS1 and NMDS2
-
-# Need to split stratum up for plotting later
-cent<- 
-  cent %>%
-    separate(stratum, into = c("year", "site", "habitat"), sep = "-")
-cent$site_habitat <- paste(cent$site, cent$habitat)
-# Add back in stratum column
-cent$stratum <-  aggregate(cbind(NMDS1, NMDS2) ~ stratum, data = site.scores, FUN = mean)$stratum
-# A lot of extra spaces
-
-cent <- cent %>%
-  mutate(across(everything(), str_trim))
-cent$year <- as.factor(cent$year)
-cent$habitat <- as.factor(cent$habitat)
-cent$site <- as.factor(cent$site)
-cent$site_habitat <- as.factor(cent$site_habitat)
-cent$NMDS1 <- as.numeric(cent$NMDS1) # read in as characters for some reason
-cent$NMDS2 <- as.numeric(cent$NMDS2)
-species.scores <- as.data.frame(scores(ord3)$species)  # extract the species scores and convert to a data.frame
-species.scores$Species <- rownames(species.scores)
-
-# subset only the species w/ > 10% cover
-abundant.species.scores <- species.scores[species.scores$Species %in% abundant_species_names,]
-
-ggplot() + 
-  geom_point(data= site.scores, aes(x=NMDS1,y=NMDS2, shape = habitat, colour = site),
-             alpha = 0.25, size=2) + # add the point markers
-  geom_text(data= abundant.species.scores,aes(x=NMDS1,y=NMDS2, label = Species), size = 4) +  # add the species labels
-
-  geom_path(data = cent, aes(x = NMDS1, y = NMDS2, group = site_habitat, colour = site),
-            arrow = arrow(length=unit(0.5,"cm"), ends="last", type = "closed"), size = 1) +
-  facet_wrap(~factor(habitat, levels=c('fringing','backreef','outer_10','outer_17'))) +
-  scale_linetype_manual(values = habitat_linetypes) +
-  scale_colour_manual(values = site_colours) +
-  scale_shape_manual(values = habitat_shapes) +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        strip.text.x = element_text(size = 25),
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.key.size = unit(2, "cm"), # increase legend size
-        legend.text = element_text(size = rel(1.5)), # increase legend text
-        legend.title = element_text(size = rel(1.5)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black")) + # add box around legend
-  guides(colour=guide_legend(title.position="top", # make color legend title go above legend
-                                     title.hjust =0.5, nrow = 2),
-         shape=guide_legend(title.position="top",  # make shape legend title go above legend
-                                     title.hjust =0.5,
-                            nrow = 2),
-         path = guide_legend(title.position="top",  # make shape legend title go above legend
-                                     title.hjust =0.5)
-         )
-
-# looks like stochasticity increases from outer_17 --> fringing
-# Even in lagoon though it looks like communities return to initial state... except for north Shore lter_sites
-# ways to test stochasticity vs. determinism?
-# seems like deterministic trajectories, although we're not starting from a clean slate
-# accounting for founder effects?
 

### 2024-08-30 julianna_algalMetric_wrangling.Rmd
Updated wrangling code to create CSVs with hard cover estimates + richness/Shannon normalized to cover

@@
            TRUE ~ (algae_cover/(100-sand))*100 
-           )) %>% # percent cover of hard substrate
+           ),
+         hard_cover = 100-sand) %>% # percent cover of hard substrate
    # make a year-habitat column for betadisper
@@
          hab_year, island_side_hab_year, transect, quadrat, algae_cover, 
-         algae_cover_hard)  -> sppMeta
+         algae_cover_hard, hard_cover, sand)  -> sppMeta
 ```
@@
          algae_shannon = Shannon) %>% 
+  # get algal metrics relative to hard
+  mutate(algae_richness_hard = algae_richness/hard_cover,
+         algae_shannon_hard = algae_shannon/hard_cover) %>% 
   # get just the columns we need
@@
             algae_shannon_se = sd(algae_shannon)/sqrt(N),
-            centroid_dist_habyear_mean = mean(hab_yr_distcentroid, na.rm = TRUE)
+            centroid_dist_habyear_mean = mean(hab_yr_distcentroid, na.rm = TRUE),
+            # add hard metrics
+            algae_richness_hard_mean = mean(algae_richness_hard, na.rm = TRUE),
+            algae_richness_hard_se = sd(algae_richness_hard, na.rm = TRUE)/sqrt(N),
+            algae_shannon_hard_mean = mean(algae_shannon_hard, na.rm = TRUE),
+            algae_shannon_se = sd(algae_shannon_hard, na.rm = TRUE)/sqrt(N),
+            hard_cover_mean = mean(hard_cover, na.rm = TRUE),
+            hard_cover_se = sd(hard_cover, na.rm = TRUE)/sqrt(N)
             ) -> siteData
@@
     algae_cover_hard_cv = CV(algae_cover_hard),
+    # including diversity
+    algae_richness_hard_mean = mean(algae_richness_hard, na.rm = TRUE),
+    algae_richness_hard_se = sd(algae_richness_hard, na.rm = TRUE)/sqrt(n()),
+    algae_shannon_hard_mean = mean(algae_shannon_hard, na.rm = TRUE),
+    algae_shannon_se = sd(algae_shannon_hard, na.rm = TRUE)/sqrt(n()),
+    hard_cover_mean = mean(hard_cover, na.rm = TRUE),
+    hard_cover_se = sd(hard_cover, na.rm = TRUE)/sqrt(n())
     ) -> alltimeCV

### 2024-08-30 julianna_coral_macroalg_wrangling.Rmd
Updated wrangling code to create CSVs with hard cover estimates + richness/Shannon normalized to cover

@@
          sand_cover = sand) %>% 
-  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg]), # percent cover of algae
+  mutate(hard_cover = 100-sand_cover) %>% 
+  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg]), # percent cover of macroalgae
          macroalgae_cover_hard = case_when(
            sand_cover == 100 ~ NA, # since we can't divide by a zero
-           TRUE ~ (macroalgae_cover/(100-sand_cover))*100 
+           TRUE ~ (macroalgae_cover/hard_cover)*100 
            ),
@@
            sand_cover == 100 ~ NA, # since we can't divide by a zero
-           TRUE ~ (coral_cover/(100-sand_cover))*100 
+           TRUE ~ (coral_cover/hard_cover)*100 
            )
@@
          hab_year, island_side_hab_year, transect, quadrat, macroalgae_cover, 
-         macroalgae_cover_hard, sand_cover, coral_cover, coral_cover_hard)  -> sppMeta
+         macroalgae_cover_hard, sand_cover, hard_cover, coral_cover, coral_cover_hard)  -> sppMeta
 ```
@@
   mutate(macroalgae_richness = specnumber(sppMat),
-         macroalgae_shannon = diversity(sppMat, "shannon")) -> macroAlgDiv
+         macroalgae_shannon = diversity(sppMat, "shannon"),
+         macroalgae_richness_hard = macroalgae_richness/hard_cover,
+         macroalgae_shannon_hard = macroalgae_shannon/hard_cover
+         ) -> macroAlgDiv
 ```
@@
             coral_cover_hard_se = sd(coral_cover_hard, na.rm = TRUE)/sqrt(N),
+            hard_cover_mean = mean(hard_cover),
+            hard_cover_se = sd(hard_cover)/sqrt(N),
+
             macroalgae_cover_mean = mean(macroalgae_cover),
@@
             macroalgae_shannon_se = sd(macroalgae_shannon)/sqrt(N),
+            # of hard cover
+            macroalgae_richness_hard_mean = mean(macroalgae_richness_hard, na.rm = TRUE),
+            macroalgae_richness_hard_se = sd(macroalgae_richness_hard, na.rm = TRUE)/sqrt(N),
+            macroalgae_shannon_hard_mean = mean(macroalgae_shannon_hard, na.rm = TRUE),
+            macroalgae_shannon_hard_se = sd(macroalgae_shannon_hard, na.rm = TRUE)/sqrt(N),
             macroalgae_centroid_dist_habyear_mean = mean(macroalgae_centroid_dist_habyear_mean, na.rm = TRUE)
@@
             coral_cover_hard_cv = CV(coral_cover_hard),
+            hard_cover_mean = mean(hard_cover),
+            hard_cover_se = sd(hard_cover)/sqrt(N),
+            hard_cover_cv = CV(hard_cover),
 
@@
             macroalgae_shannon_se = sd(macroalgae_shannon)/sqrt(N),
+
+            # of hard
+            macroalgae_richness_hard_mean = mean(macroalgae_richness_hard, na.rm = TRUE),
+            macroalgae_richness_hard_se = sd(macroalgae_richness_hard, na.rm = TRUE)/sqrt(N),
+            macroalgae_shannon_hard_mean = mean(macroalgae_shannon_hard, na.rm = TRUE),
+            macroalgae_shannon_hard_se = sd(macroalgae_shannon_hard, na.rm = TRUE)/sqrt(N),
+
             macroalgae_centroid_dist_habyear_mean = mean(macroalgae_centroid_dist_habyear_mean, na.rm = TRUE)

### 2024-08-30 R/Fish_ord.Rmd

@@
 
+### PERMANOVA
+
+```{r}
+herb_perm <- adonis2(species.data.use ~ Herbivore.Biomass + Year + Site + Habitat, data = merged_clean_filtered, dist = 'bray', by = 'margin')
+herb_perm
+```
+
+##### stratified? 
+```{r}
+library(permute)
+merged_clean_filtered$Site <- as.factor(merged_clean_filtered$Site)
+merged_clean_filtered$Habitat <- as.factor(merged_clean_filtered$Habitat)
+
+# Set up the permutation design
+perm_design <- how(
+  blocks = merged_clean_filtered$Year,  # stratify by Year
+  plots = Plots(strata = merged_clean_filtered$Site),  # stratify by Site within Year
+  within = Within(type = "free"),  # free permutations within each Site
+  nperm = 999  # number of permutations
+)
+
+# Apply the permutation design in adonis2
+herb_perm <- adonis2(
+  species.data.use ~ Herbivore.Biomass + Year + Site + Habitat, 
+  data = merged_clean_filtered, 
+  dist = 'bray', 
+  by = 'margin', 
+  permutations = perm_design
+)
+
+herb_perm
+```
+
 ## The whole fish community:
@@
 ### Part 3: Model-Based Clustering
+
+#BECAUSE I LIKE WORKING IN MARKDOWN AND PRINTING IN SCRIPTS:
+
+```{r}
+herb_beta <- ggplot(merged_div, aes(x = year)) +
+  geom_line(aes(y = centroid_dist_habyear_mean, color = "Centroid Distance")) +
+  geom_line(aes(y = Herbivore.Biomass /103388.3, color = "Herbivore Biomass")) + # Rescale Herbivore.Biomass
+  facet_grid(site ~ habitat) +
+  scale_y_continuous(
+    name = "Mean Centroid Distance",
+    sec.axis = sec_axis(~ . * 103388.3, name = "Herbivore Biomass") # Rescale back for labeling
+  ) +
+  scale_color_manual(values = c("Centroid Distance" = "darkgreen", "Herbivore Biomass" = "darkblue")) +
+  labs(x = "Year", color = "Legend", title = "Trends in Algal Beta Dispersion and Herbivore Biomass" # Add the plot title here
+  ) +
+  theme_bw() +
+  theme(legend.position = "bottom")
+```

### 2024-09-03 noam_final_analyses.Rmd
start of final analysis and data viz

---
title: "noam_final_analyses"
author: "Noam Altman-Kurosaki"
date: "2024-08-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(here)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(vegan)
library(glmmTMB)
library(car)
library(effects)
library(DHARMa)
library(stringr)
library(MuMIn)
library(codyn)
library(nlme)
library(mgcv)
library(emmeans)
library(ecofolio) # possible other method for examining portfolio effects 
```

```{r import and format data}
# coefficient of variation
CV<-function(x){
  return(sd(x,na.rm=T)/mean(x,na.rm=T))
}

# for figures
island_side_colours <- c("north" = "darkblue",
                         "east" = "mediumorchid",
                         "west" = "orange")
habitat_colours <- c("fringing" = "#D81B60",
                     "backreef" = "#FFC107",
                     "outer_10" = "#1E88E5",
                     "outer_17" = "#004D40")

# starting with all data now because of 0 algae transects/quadrats
data <- read.csv(here::here("data", "allbenthicdata_cleaned_colnames.csv"), stringsAsFactors = F)

not_algae <- c("ascidian", "bare_space", "coral", "coral_rubble", "corallimorpharia", "heteractis_sp", "millepora_platyphylla", "no_data", "sand", "sarcophyton_sp", "shell_debris", "soft_coral", "sponge", "tridacna_sp", "zooxanthellae", "row_sums")

data <- data[ , -which(names(data) %in% not_algae)]

# adding island side - adapted from Julianna's code
data <- data %>% mutate(island_side = case_when(site == "lter_1" ~ "north",
                                 site == "lter_2" ~ "north",
                                 site == "lter_3" ~ "east",
                                 site == "lter_4" ~ "east",
                                 site == "lter_5" ~ "west",
                                 site == "lter_6" ~ "west"))
data$island_side_habitat <- paste(data$island_side, data$habitat)

macroalgae <- data %>% 
  select(-algal_turf, -crustose_corallines, -cyanophyta, -damselfish_turf, -lithophyllum_kotschyanum,
         -neogoniolithon_brassica_florida, -sporolithon_sp)

sand_coral <- read_csv(here("data", "allbenthicdata_cleaned_colnames.csv")) %>% 
  select(year, location, site, habitat, site_habitat, transect, quadrat, sand, coral)

# get ranges to calculate over (this is just for indexing purposes in the next bit of code)
startAlg <- grep("acanthophora_spicifera", colnames(macroalgae))
endAlg <- grep("valonia_ventricosa", colnames(macroalgae))

macroalgae %>% 
  # bring in sand data
  full_join(sand_coral) %>% 
  rename(coral_cover = coral,
         sand_cover = sand) %>% 
  mutate(hard_cover = 100-sand_cover) %>% 
  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg]), # percent cover of macroalgae

         macroalgae_cover_hard = case_when(
           sand_cover == 100 ~ NA, # since we can't divide by a zero
           TRUE ~ (macroalgae_cover/hard_cover)*100 
           ),
         # also for corals
         coral_cover_hard = case_when(
           sand_cover == 100 ~ NA, # since we can't divide by a zero
           TRUE ~ (coral_cover/hard_cover)*100 
           )
         ) %>% # percent cover of hard substrate
   # make a year-habitat column for betadisper
  mutate(hab_year = paste0(habitat, "_", year),
         site_hab_year = paste0(site, "_", hab_year)) %>% 
  select(year, location, site, habitat, island_side, site_habitat, island_side_habitat,
         hab_year, site_hab_year, transect, quadrat, macroalgae_cover, 
         macroalgae_cover_hard, sand_cover, hard_cover, coral_cover, coral_cover_hard)  -> meta_df

meta_df$all_algae_cover <- rowSums(data[startAlg:endAlg]) # percent cover of all algae
meta_df$all_algae_cover_hard <- case_when(
           meta_df$sand_cover == 100 ~ NA, # since we can't divide by a zero
           TRUE ~ (meta_df$all_algae_cover/meta_df$hard_cover)*100   
         )

data <- merge(meta_df, data )
macroalgae <- merge(meta_df, macroalgae)

meta_cols <- colnames(meta_df)

full_species_cols <- colnames(data)[!colnames(data) %in% meta_cols]
macro_species_cols <- colnames(macroalgae)[!colnames(macroalgae) %in% meta_cols]

# summarize at site level
site_full <- data %>%
  dplyr::group_by(year, island_side, site, habitat, site_habitat, hab_year, site_hab_year) %>%
  dplyr::summarize(across(all_of(
    c("macroalgae_cover", "macroalgae_cover_hard", "sand_cover", "hard_cover", "coral_cover",
      "coral_cover_hard","all_algae_cover", "all_algae_cover_hard", full_species_cols)), mean, na.rm = TRUE))

# summarize at species level
site_macro <- macroalgae %>%
  dplyr::group_by(year, island_side, site, habitat, site_habitat, hab_year, site_hab_year) %>%
  dplyr::summarize(across(all_of(
    c("macroalgae_cover", "macroalgae_cover_hard", "sand_cover", "hard_cover", "coral_cover",
      "coral_cover_hard","all_algae_cover", "all_algae_cover_hard", macro_species_cols)), mean, na.rm = TRUE))

# extract site level meta_df for future merges
site_meta_df <- site_full[,which(names(site_full) %in% meta_cols)]

#pivot long 
long_data <- pivot_longer(
  data = data,
  cols = grep("acanthophora_spicifera", colnames(data)):grep("valonia_ventricosa", colnames(data)),
  names_to = "taxa",
  values_to = "percent_cover")

macro_long_data <- pivot_longer(
  data = macroalgae,
  cols = grep("acanthophora_spicifera", colnames(macroalgae)):grep("valonia_ventricosa", colnames(macroalgae)),
  names_to = "taxa",
  values_to = "percent_cover")

site_long_data <- pivot_longer(
  data = site_full,
  cols = grep("acanthophora_spicifera", colnames(site_full)):grep("valonia_ventricosa", colnames(site_full)),
  names_to = "taxa",
  values_to = "percent_cover")

site_long_macro <- pivot_longer(
  data = site_macro,
  cols = grep("acanthophora_spicifera", colnames(site_macro)):grep("valonia_ventricosa", colnames(site_macro)),
  names_to = "taxa",
  values_to = "percent_cover")

```

```{r read and merge secondary datasets}
wave <- read.csv(here::here("data", "algae_wave_dynamics.csv"), stringsAsFactors = F)
nutrients <- read.csv(here::here("data", "lter_nuts.csv"), stringsAsFactors = F)
temp <- read.csv(here::here("data", "temp_sum.csv"), stringsAsFactors = F)

```

```{r all algae alpha diversity quadrat level}
alpha_diversity_quad <-
  data.frame(
    year = data$year,
    location = data$location,
    shannon = diversity(data[ , -which(names(data) %in% meta_cols)], "shannon"),
    richness = rowSums(data[ , -which(names(data) %in% meta_cols)] > 0)
  )

alpha_diversity_quad <- merge(meta_df, alpha_diversity_quad, by = c("location", "year") )

full_richness_mod <- glmmTMB(richness ~ year*habitat + (1|site/location) + offset(log(hard_cover)), family = poisson,
                             data = alpha_diversity_quad[!is.na(alpha_diversity_quad$macroalgae_cover_hard),]) 
summary(full_richness_mod)
hist(residuals(full_richness_mod)) # slightly skewed but not terrible

# striping/patterining expected because of discrete response var. otherwise fine
plot(residuals(full_richness_mod) ~ fitted(full_richness_mod))

emtrends(full_richness_mod, pairwise ~  habitat, var = "year")
full_richness_emm <- emmip(full_richness_mod,habitat ~ year,
                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)

(base_richness_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_quad,
             aes(x = year, y = richness/(hard_cover/100), colour = habitat), size = 2, alpha = 0.05) + 
  geom_line(data = full_richness_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = full_richness_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10() +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

```

```{r macroalgae diversity quadrat level}
alpha_diversity_quad_macro <-
  data.frame(
    year = macroalgae$year,
    location = macroalgae$location,
    shannon = diversity(macroalgae[ , -which(names(macroalgae) %in% meta_cols)], "shannon"),
    richness = rowSums(macroalgae[ , -which(names(macroalgae) %in% meta_cols)] > 0)
  )

alpha_diversity_quad_macro <- merge(meta_df, alpha_diversity_quad_macro, by = c("location", "year"))

```

```{r all algae alpha diversity site level}
alpha_diversity_site <-
  data.frame(
    year = site_full$year,
    site_hab_year = site_full$site_hab_year,
    shannon = diversity(site_full[ , -which(names(site_full) %in% meta_cols)], "shannon"),
    richness = rowSums(site_full[ , -which(names(site_full) %in% meta_cols)] > 0)
  )

alpha_diversity_site <- merge(site_meta_df, alpha_diversity_site, by= c("year", "site_hab_year"))

site_full_richness_mod <- glmmTMB(richness ~ year*habitat + (1|site) + offset(log(hard_cover)), family = poisson,
                             data = alpha_diversity_site) 
# Warning in (function (start, objective, gradient = NULL, hessian = NULL,  : NA/NaN function evaluation
# This warning occurs when the optimizer visits a region of parameter space that is invalid. It is not a problem as long as the optimizer has left that region of parameter space upon convergence, which is indicated by an absence of the model convergence warnings described above.
summary(site_full_richness_mod)
hist(residuals(site_full_richness_mod)) # similar slight skew to before

# striping/patterining expected because of discrete response var. otherwise fine
plot(residuals(site_full_richness_mod) ~ fitted(site_full_richness_mod)) # slightly heteroskedastic

emtrends(site_full_richness_mod, pairwise ~  habitat, var = "year") # habitat trends don't differ from each other
site_full_richness_emm <- emmip(site_full_richness_mod,habitat ~ year,
                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)

(site_base_richness_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_site,
             aes(x = year, y = richness/(hard_cover/100), colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_full_richness_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = site_full_richness_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10() +
    labs(x = "Year", y = "Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

site_full_shannon_mod <- glmmTMB(log(shannon/(hard_cover/100) +1 ) ~ year*habitat + (1|site),
                             data = alpha_diversity_site) 
# sites with shannon div = 0 so log+1 transformed
summary(site_full_shannon_mod)
hist(residuals(site_full_shannon_mod)) # looks good

# slight patterning but generally fine
plot(residuals(site_full_shannon_mod) ~ fitted(site_full_shannon_mod))

emtrends(site_full_shannon_mod, pairwise ~  habitat, var = "year") # backreef and fringing differ
site_full_shannon_emm <- emmip(site_full_shannon_mod,habitat ~ year,
                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)

(site_base_shannon_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_site,
             aes(x = year, y = shannon/(hard_cover/100), colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_full_shannon_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = site_full_shannon_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10() +
    labs(x = "Year", y = "shannon") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

```

```{r macroalgae alpha diversity site level}
alpha_diversity_site_macro <-
  data.frame(
    year = site_macro$year,
    site_hab_year = site_macro$site_hab_year,
    shannon = diversity(site_macro[ , -which(names(site_macro) %in% meta_cols)], "shannon"),
    richness = rowSums(site_macro[ , -which(names(site_macro) %in% meta_cols)] > 0)
  )

alpha_diversity_site_macro <- merge(site_meta_df, alpha_diversity_site_macro, by= c("year", "site_hab_year"))

site_macro_richness_mod <- glmmTMB(richness ~ year*habitat + (1|site) + offset(log(hard_cover)), family = poisson,
                             data = alpha_diversity_site_macro) 
# Warning in (function (start, objective, gradient = NULL, hessian = NULL,  : NA/NaN function evaluation
# This warning occurs when the optimizer visits a region of parameter space that is invalid. It is not a problem as long as the optimizer has left that region of parameter space upon convergence, which is indicated by an absence of the model convergence warnings described above.
summary(site_macro_richness_mod)
hist(residuals(site_macro_richness_mod)) # slightly more skew

# striping/patterining expected because of discrete response var. otherwise fine
plot(residuals(site_macro_richness_mod) ~ fitted(site_macro_richness_mod))

emtrends(site_macro_richness_mod, pairwise ~  habitat, var = "year") # habitat trends don't differ from each other
site_macro_richness_emm <- emmip(site_macro_richness_mod,habitat ~ year,
                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)

(site_base_richness_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_site_macro,
             aes(x = year, y = richness/(hard_cover/100), colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_macro_richness_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = site_macro_richness_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10() +
    labs(x = "Year", y = "Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

site_macro_shannon_mod <- glmmTMB(log(shannon/(hard_cover/100) +1 ) ~ year*habitat + (1|site),
                             data = alpha_diversity_site) 
# sites with shannon div = 0 so log+1 transformed
summary(site_macro_shannon_mod)
hist(residuals(site_macro_shannon_mod)) # looks good

# slight patterning but generally fine
plot(residuals(site_macro_shannon_mod) ~ fitted(site_macro_shannon_mod))

emtrends(site_macro_shannon_mod, pairwise ~  habitat, var = "year") # backreef and fringing differ
site_macro_shannon_emm <- emmip(site_macro_shannon_mod,habitat ~ year,
                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)

(site_base_shannon_plot <- 
  ggplot() +
 # geom_point(data = alpha_diversity_site,
 #            aes(x = year, y = shannon/(hard_cover/100), colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_macro_shannon_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = site_macro_shannon_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
#  scale_y_log10() +
    labs(x = "Year", y = "shannon") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

```

### 2024-09-03 noam_final_analyses.Rmd
more setup

@@
   geom_point(data = alpha_diversity_quad,
-             aes(x = year, y = richness/(hard_cover/100), colour = habitat), size = 2, alpha = 0.05) + 
+             aes(x = year, y = richness, colour = habitat), size = 2, alpha = 0.05) + 
   geom_line(data = full_richness_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
@@
   scale_fill_manual(values = habitat_colours) +
+  labs(y = "Richness") +
   scale_y_log10() +
@@
   scale_fill_manual(values = habitat_colours) +
-  scale_y_log10() +
+  scale_y_log10(lim = c(0.1,20)) +
     labs(x = "Year", y = "Richness") +

### 2024-09-03 noam_beta_diversity.Rmd
Compleed site level dissimilarity over time and site level ordination for all algae. working on macro now

@@
 ```{r import and format data}
+# for figures
+island_side_colours <- c("north" = "darkblue",
+                         "east" = "mediumorchid",
+                         "west" = "orange")
+habitat_colours <- c("fringing" = "#D81B60",
+                     "backreef" = "#FFC107",
+                     "outer_10" = "#1E88E5",
+                     "outer_17" = "#004D40")
+
+# some quadrats have dissimilarity = 0 or 1
+# Smithson verkuilen transformation for beta-regression
+sv_trans <- function(prop, N, s = 0.000005){
+  (prop*(nrow(N) - 1) + s)/nrow(N)
+  # where prop is the proportional value you're transforming, 
+  # N is the sample size, which is specified by taking the number of rows/observations from a given dataframe,
+  # and s is a small offset 
+}
+
+inv_logit <- function(x) {
+  exp(x)/(1+exp(x))
+}
+
 # starting with all data now because of 0 algae transects/quadrats
@@
 
-str(data)
-
-meta_cols <- c(colnames(data[1:7] ), "island_side", "island_side_habitat", "stratum")
-species_cols <- colnames(data)[!colnames(data) %in% meta_cols]
-
-meta_df <- data[ , which(names(data) %in% meta_cols)]
-
-site_colours <- c("lter_1" = "cornflowerblue",
-                  "lter_2" = "darkblue",
-                  "lter_3" = "yellowgreen",
-                  "lter_4" = "forestgreen",
-                  "lter_5" = "mediumorchid",
-                  "lter_6" = "orange")
-island_side_colours <- c("north" = "darkblue",
-                         "east" = "mediumorchid",
-                         "west" = "orange")
-habitat_linetypes <- c("fringing" = "solid",
-                       "backreef" = "longdash",
-                       "outer_10" = "twodash",
-                       "outer_17" = "dotted")
-habitat_shapes <- c("fringing" = 16,
-                     "backreef" = 17,
-                     "outer_10" = 1,
-                     "outer_17" = 2)
-
-#pivot long 
-# long_data <- pivot_longer(
-#   data = data,
-#   cols = 8:77,
-#   names_to = "taxa",
-#   values_to = "percent_cover") 
-# 
-# 
-#  transect_level <- long_data %>%
-#    dplyr::group_by(year, island_side, site, habitat, transect, taxa) %>%
-#    dplyr::summarise(mean_cover = mean(percent_cover), 
-#              SE = sd(percent_cover)/sqrt(n()))
-#  
-# site_level <- long_data %>%
-#   dplyr::group_by(year, island_side, site, habitat, site_habitat, taxa) %>%
-#   dplyr::summarise(mean_site_cover = mean(percent_cover), 
-#             site_se = sd(percent_cover)/sqrt(n()))
-# 
-# subset based on habitat
+macroalgae <- data %>% 
+  select(-algal_turf, -crustose_corallines, -cyanophyta, -damselfish_turf, -lithophyllum_kotschyanum,
+         -neogoniolithon_brassica_florida, -sporolithon_sp)
+
+sand_coral <- read_csv(here("data", "allbenthicdata_cleaned_colnames.csv")) %>% 
+  select(year, location, site, habitat, site_habitat, transect, quadrat, sand, coral)
+
+# get ranges to calculate over (this is just for indexing purposes in the next bit of code)
+startAlg <- grep("acanthophora_spicifera", colnames(macroalgae))
+endAlg <- grep("valonia_ventricosa", colnames(macroalgae))
+
+macroalgae %>% 
+  # bring in sand data
+  full_join(sand_coral) %>% 
+  rename(coral_cover = coral,
+         sand_cover = sand) %>% 
+  mutate(hard_cover = 100-sand_cover) %>% 
+  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg]), # percent cover of macroalgae
+
+         macroalgae_cover_hard = case_when(
+           sand_cover == 100 ~ NA, # since we can't divide by a zero
+           TRUE ~ (macroalgae_cover/hard_cover)*100 
+           ),
+         # also for corals
+         coral_cover_hard = case_when(
+           sand_cover == 100 ~ NA, # since we can't divide by a zero
+           TRUE ~ (coral_cover/hard_cover)*100 
+           )
+         ) %>% # percent cover of hard substrate
+   # make a year-habitat column for betadisper
+  mutate(hab_year = paste0(habitat, "_", year),
+         site_hab_year = paste0(site, "_", hab_year)) %>% 
+  select(year, location, site, habitat, island_side, site_habitat, island_side_habitat,
+         hab_year, site_hab_year, transect, quadrat, macroalgae_cover, 
+         macroalgae_cover_hard, sand_cover, hard_cover, coral_cover, coral_cover_hard)  -> meta_df
+
+meta_df$all_algae_cover <- rowSums(data[startAlg:endAlg]) # percent cover of all algae
+meta_df$all_algae_cover_hard <- case_when(
+           meta_df$sand_cover == 100 ~ NA, # since we can't divide by a zero
+           TRUE ~ (meta_df$all_algae_cover/meta_df$hard_cover)*100   
+         )
+
+data <- merge(meta_df, data )
+macroalgae <- merge(meta_df, macroalgae)
+
+meta_cols <- colnames(meta_df)
+
+full_species_cols <- colnames(data)[!colnames(data) %in% meta_cols]
+macro_species_cols <- colnames(macroalgae)[!colnames(macroalgae) %in% meta_cols]
+
+# summarize at site level
+site_full <- data %>%
+  dplyr::group_by(year, island_side, site, habitat, site_habitat, hab_year, site_hab_year) %>%
+  dplyr::summarize(across(all_of(
+    c("macroalgae_cover", "macroalgae_cover_hard", "sand_cover", "hard_cover", "coral_cover",
+      "coral_cover_hard","all_algae_cover", "all_algae_cover_hard", full_species_cols)), mean, na.rm = TRUE))
+
+# summarize at species level
+site_macro <- macroalgae %>%
+  dplyr::group_by(year, island_side, site, habitat, site_habitat, hab_year, site_hab_year) %>%
+  dplyr::summarize(across(all_of(
+    c("macroalgae_cover", "macroalgae_cover_hard", "sand_cover", "hard_cover", "coral_cover",
+      "coral_cover_hard","all_algae_cover", "all_algae_cover_hard", macro_species_cols)), mean, na.rm = TRUE))
+
+# extract site level meta_df for future merges
+site_meta_df <- site_full[,which(names(site_full) %in% meta_cols)]
+
 fringing <- subset(data, habitat == "fringing")
@@
 
-```
-
+fringing_macro <- subset(macroalgae, habitat == "fringing")
+backreef_macro <- subset(macroalgae, habitat == "backreef")
+outer10_macro <- subset(macroalgae, habitat == "outer_10")
+outer17_macro <- subset(macroalgae, habitat == "outer_17")
+
+fringing_site <- subset(site_full, habitat == "fringing")
+backreef_site <- subset(site_full, habitat == "backreef")
+outer10_site <- subset(site_full, habitat == "outer_10")
+outer17_site <- subset(site_full, habitat == "outer_17")
+
+fringing_site_macro <- subset(site_macro, habitat == "fringing")
+backreef_site_macro <- subset(site_macro, habitat == "backreef")
+outer10_site_macro <- subset(site_macro, habitat == "outer_10")
+outer17_site_macro <- subset(site_macro, habitat == "outer_17")
+```
+
+Moving all site level analyses to before quadrat for beta diversity
+
+```{r site level bray curtis setup}
+fringing_site_species_matrix <- as.matrix(fringing_site[ , -which(names(fringing_site) %in% meta_cols)])
+fringing_site_species_matrix <- fringing_site_species_matrix/fringing_site$hard_cover # convert from percent to proportion of hard cover
+rownames(fringing_site_species_matrix) <- paste(fringing_site$year,"-", fringing_site$site_habitat, sep = "")
+
+# filter rare species
+# fringing_site_species_presence <- colSums(fringing_site_species_matrix > 0) # Presence absence matrix
+# fringing_site_species_percentage <- (fringing_site_species_presence/nrow(fringing_site_species_matrix)) # proportion of rows where a species is present
+# fringing_site_common_species <- names(fringing_site_species_percentage[fringing_site_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
+# fringing_site_species_matrix <- fringing_site_species_matrix[,colnames(fringing_site_species_matrix) %in% fringing_site_common_species] # filter original species matrix
+
+fringing_site_rows_to_remove <- rowSums(fringing_site_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
+fringing_site_species_matrix <- fringing_site_species_matrix[!fringing_site_rows_to_remove,]
+
+fringing_site_species_matrix_untransformed <- fringing_site_species_matrix
+fringing_site_species_matrix_transformed <- asin(sqrt(fringing_site_species_matrix)) # arcsin squareroot transform
+fringing_site_subset <- fringing_site[!fringing_site_rows_to_remove,] # make sure rows are also removed from dataframe
+fringing_site_subset$stratum <- paste(fringing_site_subset$year,"-", fringing_site_subset$site_habitat, sep = "") # for merging with bray matrix later
+
+brayfringing_site <- vegdist(fringing_site_species_matrix_transformed, "bray") # create bray dissimilarity matrix
+# View(brayfringing_site)
+brayfringing_site_untransformed <- vegdist(fringing_site_species_matrix_untransformed, "bray")
+fringing_site_braymat <- as.matrix(brayfringing_site) # conver to matrix object
+# convert to pairwise comparison
+fringing_site_comp <- fringing_site_braymat[lower.tri(fringing_site_braymat)] # extract lower triangle
+fringing_site_row_indices <- row(fringing_site_braymat)[lower.tri(fringing_site_braymat)] # extract row indices for first quad in comparison
+fringing_site_col_indices <- col(fringing_site_braymat)[lower.tri(fringing_site_braymat)] # extract col indices = second quad in comp
+fringing_site_loc1 <- rownames(fringing_site_braymat)[fringing_site_row_indices] # populate appropriate vector
+fringing_site_loc2 <- colnames(fringing_site_braymat)[fringing_site_col_indices] # populate appropriate vector
+fringing_site_pairwise_df <- data.frame(loc1 = fringing_site_loc1, loc2 = fringing_site_loc2, dissimilarity = fringing_site_comp)
+fringing_site_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", fringing_site_pairwise_df$loc1)) # extract timepoint from first quad in comp
+fringing_site_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", fringing_site_pairwise_df$loc2)) # repeat for second quad
+fringing_site_pairwise_df$time <- abs(fringing_site_pairwise_df$year2 - fringing_site_pairwise_df$year1) # extract length of differences
+fringing_site_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", fringing_site_pairwise_df$loc1) # extract quadrat id from first quad in comp
+fringing_site_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", fringing_site_pairwise_df$loc2) # repeat for second quad
+
+fringing_site_untransformed_braymat <- as.matrix(brayfringing_site_untransformed) # conver to matrix object
+# convert to pairwise comparison
+fringing_site_untransformed_comp <- fringing_site_untransformed_braymat[lower.tri(fringing_site_untransformed_braymat)] # extract lower triangle
+fringing_site_untransformed_row_indices <- row(fringing_site_untransformed_braymat)[lower.tri(fringing_site_untransformed_braymat)] # extract row indices for first quad in comparison
+fringing_site_untransformed_col_indices <- col(fringing_site_untransformed_braymat)[lower.tri(fringing_site_untransformed_braymat)] # extract col indices = second quad in comp
+fringing_site_untransformed_loc1 <- rownames(fringing_site_untransformed_braymat)[fringing_site_untransformed_row_indices] # populate appropriate vector
+fringing_site_untransformed_loc2 <- colnames(fringing_site_untransformed_braymat)[fringing_site_untransformed_col_indices] # populate appropriate vector
+fringing_site_untransformed_pairwise_df <- data.frame(loc1 = fringing_site_untransformed_loc1, loc2 = fringing_site_untransformed_loc2, dissimilarity = fringing_site_untransformed_comp)
+fringing_site_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", fringing_site_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
+fringing_site_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", fringing_site_untransformed_pairwise_df$loc2)) # repeat for second quad
+fringing_site_untransformed_pairwise_df$time <- abs(fringing_site_untransformed_pairwise_df$year2 - fringing_site_untransformed_pairwise_df$year1) # extract length of differences
+fringing_site_untransformed_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", fringing_site_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
+fringing_site_untransformed_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", fringing_site_untransformed_pairwise_df$loc2) # repeat for second quad
+
+# backreef
+
+backreef_site_species_matrix <- as.matrix(backreef_site[ , -which(names(backreef_site) %in% meta_cols)])
+backreef_site_species_matrix <- backreef_site_species_matrix/backreef_site$hard_cover # convert from percent to proportion of hard cover
+rownames(backreef_site_species_matrix) <- paste(backreef_site$year,"-", backreef_site$site_habitat, sep = "")
+
+# filter rare species
+# backreef_site_species_presence <- colSums(backreef_site_species_matrix > 0) # Presence absence matrix
+# backreef_site_species_percentage <- (backreef_site_species_presence/nrow(backreef_site_species_matrix)) # proportion of rows where a species is present
+# backreef_site_common_species <- names(backreef_site_species_percentage[backreef_site_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
+# backreef_site_species_matrix <- backreef_site_species_matrix[,colnames(backreef_site_species_matrix) %in% backreef_site_common_species] # filter original species matrix
+
+backreef_site_rows_to_remove <- rowSums(backreef_site_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
+backreef_site_species_matrix <- backreef_site_species_matrix[!backreef_site_rows_to_remove,]
+
+backreef_site_species_matrix_untransformed <- backreef_site_species_matrix
+backreef_site_species_matrix_transformed <- asin(sqrt(backreef_site_species_matrix)) # arcsin squareroot transform
+backreef_site_subset <- backreef_site[!backreef_site_rows_to_remove,] # make sure rows are also removed from dataframe
+backreef_site_subset$stratum <- paste(backreef_site_subset$year,"-", backreef_site_subset$site_habitat, sep = "") # for merging with bray matrix later
+
+braybackreef_site <- vegdist(backreef_site_species_matrix_transformed, "bray") # create bray dissimilarity matrix
+# View(braybackreef_site)
+braybackreef_site_untransformed <- vegdist(backreef_site_species_matrix_untransformed, "bray")
+backreef_site_braymat <- as.matrix(braybackreef_site) # conver to matrix object
+# convert to pairwise comparison
+backreef_site_comp <- backreef_site_braymat[lower.tri(backreef_site_braymat)] # extract lower triangle
+backreef_site_row_indices <- row(backreef_site_braymat)[lower.tri(backreef_site_braymat)] # extract row indices for first quad in comparison
+backreef_site_col_indices <- col(backreef_site_braymat)[lower.tri(backreef_site_braymat)] # extract col indices = second quad in comp
+backreef_site_loc1 <- rownames(backreef_site_braymat)[backreef_site_row_indices] # populate appropriate vector
+backreef_site_loc2 <- colnames(backreef_site_braymat)[backreef_site_col_indices] # populate appropriate vector
+backreef_site_pairwise_df <- data.frame(loc1 = backreef_site_loc1, loc2 = backreef_site_loc2, dissimilarity = backreef_site_comp)
+backreef_site_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", backreef_site_pairwise_df$loc1)) # extract timepoint from first quad in comp
+backreef_site_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", backreef_site_pairwise_df$loc2)) # repeat for second quad
+backreef_site_pairwise_df$time <- abs(backreef_site_pairwise_df$year2 - backreef_site_pairwise_df$year1) # extract length of differences
+backreef_site_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", backreef_site_pairwise_df$loc1) # extract quadrat id from first quad in comp
+backreef_site_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", backreef_site_pairwise_df$loc2) # repeat for second quad
+
+backreef_site_untransformed_braymat <- as.matrix(braybackreef_site_untransformed) # conver to matrix object
+# convert to pairwise comparison
+backreef_site_untransformed_comp <- backreef_site_untransformed_braymat[lower.tri(backreef_site_untransformed_braymat)] # extract lower triangle
+backreef_site_untransformed_row_indices <- row(backreef_site_untransformed_braymat)[lower.tri(backreef_site_untransformed_braymat)] # extract row indices for first quad in comparison
+backreef_site_untransformed_col_indices <- col(backreef_site_untransformed_braymat)[lower.tri(backreef_site_untransformed_braymat)] # extract col indices = second quad in comp
+backreef_site_untransformed_loc1 <- rownames(backreef_site_untransformed_braymat)[backreef_site_untransformed_row_indices] # populate appropriate vector
+backreef_site_untransformed_loc2 <- colnames(backreef_site_untransformed_braymat)[backreef_site_untransformed_col_indices] # populate appropriate vector
+backreef_site_untransformed_pairwise_df <- data.frame(loc1 = backreef_site_untransformed_loc1, loc2 = backreef_site_untransformed_loc2, dissimilarity = backreef_site_untransformed_comp)
+backreef_site_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", backreef_site_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
+backreef_site_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", backreef_site_untransformed_pairwise_df$loc2)) # repeat for second quad
+backreef_site_untransformed_pairwise_df$time <- abs(backreef_site_untransformed_pairwise_df$year2 - backreef_site_untransformed_pairwise_df$year1) # extract length of differences
+backreef_site_untransformed_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", backreef_site_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
+backreef_site_untransformed_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", backreef_site_untransformed_pairwise_df$loc2) # repeat for second quad
+
+# outer10
+
+outer10_site_species_matrix <- as.matrix(outer10_site[ , -which(names(outer10_site) %in% meta_cols)])
+outer10_site_species_matrix <- outer10_site_species_matrix/outer10_site$hard_cover # convert from percent to proportion of hard cover
+rownames(outer10_site_species_matrix) <- paste(outer10_site$year,"-", outer10_site$site_habitat, sep = "")
+
+# filter rare species
+# outer10_site_species_presence <- colSums(outer10_site_species_matrix > 0) # Presence absence matrix
+# outer10_site_species_percentage <- (outer10_site_species_presence/nrow(outer10_site_species_matrix)) # proportion of rows where a species is present
+# outer10_site_common_species <- names(outer10_site_species_percentage[outer10_site_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
+# outer10_site_species_matrix <- outer10_site_species_matrix[,colnames(outer10_site_species_matrix) %in% outer10_site_common_species] # filter original species matrix
+
+outer10_site_rows_to_remove <- rowSums(outer10_site_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
+outer10_site_species_matrix <- outer10_site_species_matrix[!outer10_site_rows_to_remove,]
+
+outer10_site_species_matrix_untransformed <- outer10_site_species_matrix
+outer10_site_species_matrix_transformed <- asin(sqrt(outer10_site_species_matrix)) # arcsin squareroot transform
+outer10_site_subset <- outer10_site[!outer10_site_rows_to_remove,] # make sure rows are also removed from dataframe
+outer10_site_subset$stratum <- paste(outer10_site_subset$year,"-", outer10_site_subset$site_habitat, sep = "") # for merging with bray matrix later
+
+brayouter10_site <- vegdist(outer10_site_species_matrix_transformed, "bray") # create bray dissimilarity matrix
+# View(brayouter10_site)
+brayouter10_site_untransformed <- vegdist(outer10_site_species_matrix_untransformed, "bray")
+outer10_site_braymat <- as.matrix(brayouter10_site) # conver to matrix object
+# convert to pairwise comparison
+outer10_site_comp <- outer10_site_braymat[lower.tri(outer10_site_braymat)] # extract lower triangle
+outer10_site_row_indices <- row(outer10_site_braymat)[lower.tri(outer10_site_braymat)] # extract row indices for first quad in comparison
+outer10_site_col_indices <- col(outer10_site_braymat)[lower.tri(outer10_site_braymat)] # extract col indices = second quad in comp
+outer10_site_loc1 <- rownames(outer10_site_braymat)[outer10_site_row_indices] # populate appropriate vector
+outer10_site_loc2 <- colnames(outer10_site_braymat)[outer10_site_col_indices] # populate appropriate vector
+outer10_site_pairwise_df <- data.frame(loc1 = outer10_site_loc1, loc2 = outer10_site_loc2, dissimilarity = outer10_site_comp)
+outer10_site_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer10_site_pairwise_df$loc1)) # extract timepoint from first quad in comp
+outer10_site_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer10_site_pairwise_df$loc2)) # repeat for second quad
+outer10_site_pairwise_df$time <- abs(outer10_site_pairwise_df$year2 - outer10_site_pairwise_df$year1) # extract length of differences
+outer10_site_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", outer10_site_pairwise_df$loc1) # extract quadrat id from first quad in comp
+outer10_site_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", outer10_site_pairwise_df$loc2) # repeat for second quad
+
+outer10_site_untransformed_braymat <- as.matrix(brayouter10_site_untransformed) # conver to matrix object
+# convert to pairwise comparison
+outer10_site_untransformed_comp <- outer10_site_untransformed_braymat[lower.tri(outer10_site_untransformed_braymat)] # extract lower triangle
+outer10_site_untransformed_row_indices <- row(outer10_site_untransformed_braymat)[lower.tri(outer10_site_untransformed_braymat)] # extract row indices for first quad in comparison
+outer10_site_untransformed_col_indices <- col(outer10_site_untransformed_braymat)[lower.tri(outer10_site_untransformed_braymat)] # extract col indices = second quad in comp
+outer10_site_untransformed_loc1 <- rownames(outer10_site_untransformed_braymat)[outer10_site_untransformed_row_indices] # populate appropriate vector
+outer10_site_untransformed_loc2 <- colnames(outer10_site_untransformed_braymat)[outer10_site_untransformed_col_indices] # populate appropriate vector
+outer10_site_untransformed_pairwise_df <- data.frame(loc1 = outer10_site_untransformed_loc1, loc2 = outer10_site_untransformed_loc2, dissimilarity = outer10_site_untransformed_comp)
+outer10_site_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer10_site_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
+outer10_site_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer10_site_untransformed_pairwise_df$loc2)) # repeat for second quad
+outer10_site_untransformed_pairwise_df$time <- abs(outer10_site_untransformed_pairwise_df$year2 - outer10_site_untransformed_pairwise_df$year1) # extract length of differences
+outer10_site_untransformed_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", outer10_site_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
+outer10_site_untransformed_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", outer10_site_untransformed_pairwise_df$loc2) # repeat for second quad
+
+# outer 17
+
+outer17_site_species_matrix <- as.matrix(outer17_site[ , -which(names(outer17_site) %in% meta_cols)])
+outer17_site_species_matrix <- outer17_site_species_matrix/outer17_site$hard_cover # convert from percent to proportion of hard cover
+rownames(outer17_site_species_matrix) <- paste(outer17_site$year,"-", outer17_site$site_habitat, sep = "")
+
+# filter rare species
+# outer17_site_species_presence <- colSums(outer17_site_species_matrix > 0) # Presence absence matrix
+# outer17_site_species_percentage <- (outer17_site_species_presence/nrow(outer17_site_species_matrix)) # proportion of rows where a species is present
+# outer17_site_common_species <- names(outer17_site_species_percentage[outer17_site_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
+# outer17_site_species_matrix <- outer17_site_species_matrix[,colnames(outer17_site_species_matrix) %in% outer17_site_common_species] # filter original species matrix
+
+outer17_site_rows_to_remove <- rowSums(outer17_site_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
+outer17_site_species_matrix <- outer17_site_species_matrix[!outer17_site_rows_to_remove,]
+
+outer17_site_species_matrix_untransformed <- outer17_site_species_matrix
+outer17_site_species_matrix_transformed <- asin(sqrt(outer17_site_species_matrix)) # arcsin squareroot transform
+outer17_site_subset <- outer17_site[!outer17_site_rows_to_remove,] # make sure rows are also removed from dataframe
+outer17_site_subset$stratum <- paste(outer17_site_subset$year,"-", outer17_site_subset$site_habitat, sep = "") # for merging with bray matrix later
+
+brayouter17_site <- vegdist(outer17_site_species_matrix_transformed, "bray") # create bray dissimilarity matrix
+# View(brayouter17_site)
+brayouter17_site_untransformed <- vegdist(outer17_site_species_matrix_untransformed, "bray")
+outer17_site_braymat <- as.matrix(brayouter17_site) # conver to matrix object
+# convert to pairwise comparison
+outer17_site_comp <- outer17_site_braymat[lower.tri(outer17_site_braymat)] # extract lower triangle
+outer17_site_row_indices <- row(outer17_site_braymat)[lower.tri(outer17_site_braymat)] # extract row indices for first quad in comparison
+outer17_site_col_indices <- col(outer17_site_braymat)[lower.tri(outer17_site_braymat)] # extract col indices = second quad in comp
+outer17_site_loc1 <- rownames(outer17_site_braymat)[outer17_site_row_indices] # populate appropriate vector
+outer17_site_loc2 <- colnames(outer17_site_braymat)[outer17_site_col_indices] # populate appropriate vector
+outer17_site_pairwise_df <- data.frame(loc1 = outer17_site_loc1, loc2 = outer17_site_loc2, dissimilarity = outer17_site_comp)
+outer17_site_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer17_site_pairwise_df$loc1)) # extract timepoint from first quad in comp
+outer17_site_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer17_site_pairwise_df$loc2)) # repeat for second quad
+outer17_site_pairwise_df$time <- abs(outer17_site_pairwise_df$year2 - outer17_site_pairwise_df$year1) # extract length of differences
+outer17_site_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", outer17_site_pairwise_df$loc1) # extract quadrat id from first quad in comp
+outer17_site_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", outer17_site_pairwise_df$loc2) # repeat for second quad
+
+outer17_site_untransformed_braymat <- as.matrix(brayouter17_site_untransformed) # conver to matrix object
+# convert to pairwise comparison
+outer17_site_untransformed_comp <- outer17_site_untransformed_braymat[lower.tri(outer17_site_untransformed_braymat)] # extract lower triangle
+outer17_site_untransformed_row_indices <- row(outer17_site_untransformed_braymat)[lower.tri(outer17_site_untransformed_braymat)] # extract row indices for first quad in comparison
+outer17_site_untransformed_col_indices <- col(outer17_site_untransformed_braymat)[lower.tri(outer17_site_untransformed_braymat)] # extract col indices = second quad in comp
+outer17_site_untransformed_loc1 <- rownames(outer17_site_untransformed_braymat)[outer17_site_untransformed_row_indices] # populate appropriate vector
+outer17_site_untransformed_loc2 <- colnames(outer17_site_untransformed_braymat)[outer17_site_untransformed_col_indices] # populate appropriate vector
+outer17_site_untransformed_pairwise_df <- data.frame(loc1 = outer17_site_untransformed_loc1, loc2 = outer17_site_untransformed_loc2, dissimilarity = outer17_site_untransformed_comp)
+outer17_site_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer17_site_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
+outer17_site_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer17_site_untransformed_pairwise_df$loc2)) # repeat for second quad
+outer17_site_untransformed_pairwise_df$time <- abs(outer17_site_untransformed_pairwise_df$year2 - outer17_site_untransformed_pairwise_df$year1) # extract length of differences
+outer17_site_untransformed_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", outer17_site_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
+outer17_site_untransformed_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", outer17_site_untransformed_pairwise_df$loc2) # repeat for second quad
+
+```
+
+```{r site bray dissimilarity over time, fig.height=15, fig.width=15}
+# for the purposes of this first set of analyses, I'm interested in dissimilarity of a given quad over time
+# not necessarily interested in how quads compare to each other over time/at each timepoint
+fringing_site_pairwise_subset <- fringing_site_pairwise_df[fringing_site_pairwise_df$site_1 == fringing_site_pairwise_df$site_2,]
+backreef_site_pairwise_subset <- backreef_site_pairwise_df[backreef_site_pairwise_df$site_1 == backreef_site_pairwise_df$site_2,]
+outer10_site_pairwise_subset <- outer10_site_pairwise_df[outer10_site_pairwise_df$site_1 == outer10_site_pairwise_df$site_2,]
+outer17_site_pairwise_subset <- outer17_site_pairwise_df[outer17_site_pairwise_df$site_1 == outer17_site_pairwise_df$site_2,]
+
+comb_site_subset <- rbind(fringing_site_pairwise_subset,
+                     backreef_site_pairwise_subset,
+                     outer10_site_pairwise_subset,
+                     outer17_site_pairwise_subset)
+# site_1 and site_2 columns are all the same now
+# create new col just called for merge with other metadata
+comb_site_subset$site_habitat <- comb_site_subset$site_1
+
+bray_time_series_site <- left_join(unique(site_meta_df[ , c("site", "site_habitat", "habitat", "island_side")]), comb_site_subset, by = "site_habitat")
+
+# model
+
+# transformed dissimilarities
+bray_time_series_site$dissimilarity_trans <- sv_trans(bray_time_series_site$dissimilarity, bray_time_series_site)
+
+bray_mod_site <- glmmTMB(dissimilarity_trans ~ time*habitat + (1 | site),
+                              na.action = na.pass, family = beta_family(link = "logit"), data = bray_time_series_site) #
+summary(bray_mod_site) 
+hist(residuals(bray_mod_site)) # slight skew but not awful
+plot(residuals(bray_mod_site) ~ fitted(bray_mod_site)) # some heteroskedasticity but not bad
+
+emtrends(bray_mod_site, pairwise ~ habitat , "time") # no sig differences
+bray_trends_emm_site <- emmip(bray_mod_site, habitat ~ time , at = list(time = seq(1,16,1)), plotit = F, CIs = T)
+
+ggplot()+
+   geom_point(data = bray_time_series_site, aes(x = time, y = dissimilarity, colour = habitat), alpha = 0.1 ) +
+   geom_line(data = bray_trends_emm_site, aes(x = time, y = inv_logit(yvar), colour = habitat), linewidth = 2) +
+   geom_ribbon(data = bray_trends_emm_site, aes(x = time, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+   scale_fill_manual(values = habitat_colours) +
+
+   labs(y = "Dissimilarity\n") +
+   theme_bw() + 
+   theme(panel.grid.major = element_blank(), 
+         panel.grid.minor = element_blank(),
+         panel.background = element_blank(),
+         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+         axis.line.y = element_line(colour = "black"),
+         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         axis.line.x = element_line(colour = "black"),
+         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         strip.text.x = element_text(size = 32),
+         legend.key.size = unit(2, "cm"), # increase legend size
+         legend.text = element_text(size = rel(2)), # increase legend text
+         legend.title = element_text(size = rel(2.25)),
+         legend.position = 'bottom', # move legend to bottom
+         legend.box = 'horizontal',
+         legend.box.background = element_rect(colour = "black") )# make legend horizontal
+```
+
+```{r site macroalgae level bray curtis setup}
+fringing_site_macro_species_matrix <- as.matrix(fringing_site_macro[ , -which(names(fringing_site_macro) %in% meta_cols)])
+fringing_site_macro_species_matrix <- fringing_site_macro_species_matrix/fringing_site_macro$hard_cover # convert from percent to proportion of hard cover
+rownames(fringing_site_macro_species_matrix) <- paste(fringing_site_macro$year,"-", fringing_site_macro$site_habitat, sep = "")
+
+# filter rare species
+# fringing_site_macro_species_presence <- colSums(fringing_site_macro_species_matrix > 0) # Presence absence matrix
+# fringing_site_macro_species_percentage <- (fringing_site_macro_species_presence/nrow(fringing_site_macro_species_matrix)) # proportion of rows where a species is present
+# fringing_site_macro_common_species <- names(fringing_site_macro_species_percentage[fringing_site_macro_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
+# fringing_site_macro_species_matrix <- fringing_site_macro_species_matrix[,colnames(fringing_site_macro_species_matrix) %in% fringing_site_macro_common_species] # filter original species matrix
+
+fringing_site_macro_rows_to_remove <- rowSums(fringing_site_macro_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
+fringing_site_macro_species_matrix <- fringing_site_macro_species_matrix[!fringing_site_macro_rows_to_remove,]
+
+fringing_site_macro_species_matrix_untransformed <- fringing_site_macro_species_matrix
+fringing_site_macro_species_matrix_transformed <- asin(sqrt(fringing_site_macro_species_matrix)) # arcsin squareroot transform
+fringing_site_macro_subset <- fringing_site_macro[!fringing_site_macro_rows_to_remove,] # make sure rows are also removed from dataframe
+fringing_site_macro_subset$stratum <- paste(fringing_site_macro_subset$year,"-", fringing_site_macro_subset$site_habitat, sep = "") # for merging with bray matrix later
+
+brayfringing_site_macro <- vegdist(fringing_site_macro_species_matrix_transformed, "bray") # create bray dissimilarity matrix
+# View(brayfringing_site_macro)
+brayfringing_site_macro_untransformed <- vegdist(fringing_site_macro_species_matrix_untransformed, "bray")
+fringing_site_macro_braymat <- as.matrix(brayfringing_site_macro) # conver to matrix object
+# convert to pairwise comparison
+fringing_site_macro_comp <- fringing_site_macro_braymat[lower.tri(fringing_site_macro_braymat)] # extract lower triangle
+fringing_site_macro_row_indices <- row(fringing_site_macro_braymat)[lower.tri(fringing_site_macro_braymat)] # extract row indices for first quad in comparison
+fringing_site_macro_col_indices <- col(fringing_site_macro_braymat)[lower.tri(fringing_site_macro_braymat)] # extract col indices = second quad in comp
+fringing_site_macro_loc1 <- rownames(fringing_site_macro_braymat)[fringing_site_macro_row_indices] # populate appropriate vector
+fringing_site_macro_loc2 <- colnames(fringing_site_macro_braymat)[fringing_site_macro_col_indices] # populate appropriate vector
+fringing_site_macro_pairwise_df <- data.frame(loc1 = fringing_site_macro_loc1, loc2 = fringing_site_macro_loc2, dissimilarity = fringing_site_macro_comp)
+fringing_site_macro_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", fringing_site_macro_pairwise_df$loc1)) # extract timepoint from first quad in comp
+fringing_site_macro_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", fringing_site_macro_pairwise_df$loc2)) # repeat for second quad
+fringing_site_macro_pairwise_df$time <- abs(fringing_site_macro_pairwise_df$year2 - fringing_site_macro_pairwise_df$year1) # extract length of differences
+fringing_site_macro_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", fringing_site_macro_pairwise_df$loc1) # extract quadrat id from first quad in comp
+fringing_site_macro_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", fringing_site_macro_pairwise_df$loc2) # repeat for second quad
+
+fringing_site_macro_untransformed_braymat <- as.matrix(brayfringing_site_macro_untransformed) # conver to matrix object
+# convert to pairwise comparison
+fringing_site_macro_untransformed_comp <- fringing_site_macro_untransformed_braymat[lower.tri(fringing_site_macro_untransformed_braymat)] # extract lower triangle
+fringing_site_macro_untransformed_row_indices <- row(fringing_site_macro_untransformed_braymat)[lower.tri(fringing_site_macro_untransformed_braymat)] # extract row indices for first quad in comparison
+fringing_site_macro_untransformed_col_indices <- col(fringing_site_macro_untransformed_braymat)[lower.tri(fringing_site_macro_untransformed_braymat)] # extract col indices = second quad in comp
+fringing_site_macro_untransformed_loc1 <- rownames(fringing_site_macro_untransformed_braymat)[fringing_site_macro_untransformed_row_indices] # populate appropriate vector
+fringing_site_macro_untransformed_loc2 <- colnames(fringing_site_macro_untransformed_braymat)[fringing_site_macro_untransformed_col_indices] # populate appropriate vector
+fringing_site_macro_untransformed_pairwise_df <- data.frame(loc1 = fringing_site_macro_untransformed_loc1, loc2 = fringing_site_macro_untransformed_loc2, dissimilarity = fringing_site_macro_untransformed_comp)
+fringing_site_macro_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", fringing_site_macro_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
+fringing_site_macro_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", fringing_site_macro_untransformed_pairwise_df$loc2)) # repeat for second quad
+fringing_site_macro_untransformed_pairwise_df$time <- abs(fringing_site_macro_untransformed_pairwise_df$year2 - fringing_site_macro_untransformed_pairwise_df$year1) # extract length of differences
+fringing_site_macro_untransformed_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", fringing_site_macro_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
+fringing_site_macro_untransformed_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", fringing_site_macro_untransformed_pairwise_df$loc2) # repeat for second quad
+
+# backreef
+
+backreef_site_macro_species_matrix <- as.matrix(backreef_site_macro[ , -which(names(backreef_site_macro) %in% meta_cols)])
+backreef_site_macro_species_matrix <- backreef_site_macro_species_matrix/backreef_site_macro$hard_cover # convert from percent to proportion of hard cover
+rownames(backreef_site_macro_species_matrix) <- paste(backreef_site_macro$year,"-", backreef_site_macro$site_habitat, sep = "")
+
+# filter rare species
+# backreef_site_macro_species_presence <- colSums(backreef_site_macro_species_matrix > 0) # Presence absence matrix
+# backreef_site_macro_species_percentage <- (backreef_site_macro_species_presence/nrow(backreef_site_macro_species_matrix)) # proportion of rows where a species is present
+# backreef_site_macro_common_species <- names(backreef_site_macro_species_percentage[backreef_site_macro_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
+# backreef_site_macro_species_matrix <- backreef_site_macro_species_matrix[,colnames(backreef_site_macro_species_matrix) %in% backreef_site_macro_common_species] # filter original species matrix
+
+backreef_site_macro_rows_to_remove <- rowSums(backreef_site_macro_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
+backreef_site_macro_species_matrix <- backreef_site_macro_species_matrix[!backreef_site_macro_rows_to_remove,]
+
+backreef_site_macro_species_matrix_untransformed <- backreef_site_macro_species_matrix
+backreef_site_macro_species_matrix_transformed <- asin(sqrt(backreef_site_macro_species_matrix)) # arcsin squareroot transform
+backreef_site_macro_subset <- backreef_site_macro[!backreef_site_macro_rows_to_remove,] # make sure rows are also removed from dataframe
+backreef_site_macro_subset$stratum <- paste(backreef_site_macro_subset$year,"-", backreef_site_macro_subset$site_habitat, sep = "") # for merging with bray matrix later
+
+braybackreef_site_macro <- vegdist(backreef_site_macro_species_matrix_transformed, "bray") # create bray dissimilarity matrix
+# View(braybackreef_site_macro)
+braybackreef_site_macro_untransformed <- vegdist(backreef_site_macro_species_matrix_untransformed, "bray")
+backreef_site_macro_braymat <- as.matrix(braybackreef_site_macro) # conver to matrix object
+# convert to pairwise comparison
+backreef_site_macro_comp <- backreef_site_macro_braymat[lower.tri(backreef_site_macro_braymat)] # extract lower triangle
+backreef_site_macro_row_indices <- row(backreef_site_macro_braymat)[lower.tri(backreef_site_macro_braymat)] # extract row indices for first quad in comparison
+backreef_site_macro_col_indices <- col(backreef_site_macro_braymat)[lower.tri(backreef_site_macro_braymat)] # extract col indices = second quad in comp
+backreef_site_macro_loc1 <- rownames(backreef_site_macro_braymat)[backreef_site_macro_row_indices] # populate appropriate vector
+backreef_site_macro_loc2 <- colnames(backreef_site_macro_braymat)[backreef_site_macro_col_indices] # populate appropriate vector
+backreef_site_macro_pairwise_df <- data.frame(loc1 = backreef_site_macro_loc1, loc2 = backreef_site_macro_loc2, dissimilarity = backreef_site_macro_comp)
+backreef_site_macro_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", backreef_site_macro_pairwise_df$loc1)) # extract timepoint from first quad in comp
+backreef_site_macro_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", backreef_site_macro_pairwise_df$loc2)) # repeat for second quad
+backreef_site_macro_pairwise_df$time <- abs(backreef_site_macro_pairwise_df$year2 - backreef_site_macro_pairwise_df$year1) # extract length of differences
+backreef_site_macro_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", backreef_site_macro_pairwise_df$loc1) # extract quadrat id from first quad in comp
+backreef_site_macro_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", backreef_site_macro_pairwise_df$loc2) # repeat for second quad
+
+backreef_site_macro_untransformed_braymat <- as.matrix(braybackreef_site_macro_untransformed) # conver to matrix object
+# convert to pairwise comparison
+backreef_site_macro_untransformed_comp <- backreef_site_macro_untransformed_braymat[lower.tri(backreef_site_macro_untransformed_braymat)] # extract lower triangle
+backreef_site_macro_untransformed_row_indices <- row(backreef_site_macro_untransformed_braymat)[lower.tri(backreef_site_macro_untransformed_braymat)] # extract row indices for first quad in comparison
+backreef_site_macro_untransformed_col_indices <- col(backreef_site_macro_untransformed_braymat)[lower.tri(backreef_site_macro_untransformed_braymat)] # extract col indices = second quad in comp
+backreef_site_macro_untransformed_loc1 <- rownames(backreef_site_macro_untransformed_braymat)[backreef_site_macro_untransformed_row_indices] # populate appropriate vector
+backreef_site_macro_untransformed_loc2 <- colnames(backreef_site_macro_untransformed_braymat)[backreef_site_macro_untransformed_col_indices] # populate appropriate vector
+backreef_site_macro_untransformed_pairwise_df <- data.frame(loc1 = backreef_site_macro_untransformed_loc1, loc2 = backreef_site_macro_untransformed_loc2, dissimilarity = backreef_site_macro_untransformed_comp)
+backreef_site_macro_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", backreef_site_macro_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
+backreef_site_macro_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", backreef_site_macro_untransformed_pairwise_df$loc2)) # repeat for second quad
+backreef_site_macro_untransformed_pairwise_df$time <- abs(backreef_site_macro_untransformed_pairwise_df$year2 - backreef_site_macro_untransformed_pairwise_df$year1) # extract length of differences
+backreef_site_macro_untransformed_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", backreef_site_macro_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
+backreef_site_macro_untransformed_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", backreef_site_macro_untransformed_pairwise_df$loc2) # repeat for second quad
+
+# outer10
+
+outer10_site_macro_species_matrix <- as.matrix(outer10_site_macro[ , -which(names(outer10_site_macro) %in% meta_cols)])
+outer10_site_macro_species_matrix <- outer10_site_macro_species_matrix/outer10_site_macro$hard_cover # convert from percent to proportion of hard cover
+rownames(outer10_site_macro_species_matrix) <- paste(outer10_site_macro$year,"-", outer10_site_macro$site_habitat, sep = "")
+
+# filter rare species
+# outer10_site_macro_species_presence <- colSums(outer10_site_macro_species_matrix > 0) # Presence absence matrix
+# outer10_site_macro_species_percentage <- (outer10_site_macro_species_presence/nrow(outer10_site_macro_species_matrix)) # proportion of rows where a species is present
+# outer10_site_macro_common_species <- names(outer10_site_macro_species_percentage[outer10_site_macro_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
+# outer10_site_macro_species_matrix <- outer10_site_macro_species_matrix[,colnames(outer10_site_macro_species_matrix) %in% outer10_site_macro_common_species] # filter original species matrix
+
+outer10_site_macro_rows_to_remove <- rowSums(outer10_site_macro_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
+outer10_site_macro_species_matrix <- outer10_site_macro_species_matrix[!outer10_site_macro_rows_to_remove,]
+
+outer10_site_macro_species_matrix_untransformed <- outer10_site_macro_species_matrix
+outer10_site_macro_species_matrix_transformed <- asin(sqrt(outer10_site_macro_species_matrix)) # arcsin squareroot transform
+outer10_site_macro_subset <- outer10_site_macro[!outer10_site_macro_rows_to_remove,] # make sure rows are also removed from dataframe
+outer10_site_macro_subset$stratum <- paste(outer10_site_macro_subset$year,"-", outer10_site_macro_subset$site_habitat, sep = "") # for merging with bray matrix later
+
+brayouter10_site_macro <- vegdist(outer10_site_macro_species_matrix_transformed, "bray") # create bray dissimilarity matrix
+# View(brayouter10_site_macro)
+brayouter10_site_macro_untransformed <- vegdist(outer10_site_macro_species_matrix_untransformed, "bray")
+outer10_site_macro_braymat <- as.matrix(brayouter10_site_macro) # conver to matrix object
+# convert to pairwise comparison
+outer10_site_macro_comp <- outer10_site_macro_braymat[lower.tri(outer10_site_macro_braymat)] # extract lower triangle
+outer10_site_macro_row_indices <- row(outer10_site_macro_braymat)[lower.tri(outer10_site_macro_braymat)] # extract row indices for first quad in comparison
+outer10_site_macro_col_indices <- col(outer10_site_macro_braymat)[lower.tri(outer10_site_macro_braymat)] # extract col indices = second quad in comp
+outer10_site_macro_loc1 <- rownames(outer10_site_macro_braymat)[outer10_site_macro_row_indices] # populate appropriate vector
+outer10_site_macro_loc2 <- colnames(outer10_site_macro_braymat)[outer10_site_macro_col_indices] # populate appropriate vector
+outer10_site_macro_pairwise_df <- data.frame(loc1 = outer10_site_macro_loc1, loc2 = outer10_site_macro_loc2, dissimilarity = outer10_site_macro_comp)
+outer10_site_macro_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer10_site_macro_pairwise_df$loc1)) # extract timepoint from first quad in comp
+outer10_site_macro_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer10_site_macro_pairwise_df$loc2)) # repeat for second quad
+outer10_site_macro_pairwise_df$time <- abs(outer10_site_macro_pairwise_df$year2 - outer10_site_macro_pairwise_df$year1) # extract length of differences
+outer10_site_macro_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", outer10_site_macro_pairwise_df$loc1) # extract quadrat id from first quad in comp
+outer10_site_macro_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", outer10_site_macro_pairwise_df$loc2) # repeat for second quad
+
+outer10_site_macro_untransformed_braymat <- as.matrix(brayouter10_site_macro_untransformed) # conver to matrix object
+# convert to pairwise comparison
+outer10_site_macro_untransformed_comp <- outer10_site_macro_untransformed_braymat[lower.tri(outer10_site_macro_untransformed_braymat)] # extract lower triangle
+outer10_site_macro_untransformed_row_indices <- row(outer10_site_macro_untransformed_braymat)[lower.tri(outer10_site_macro_untransformed_braymat)] # extract row indices for first quad in comparison
+outer10_site_macro_untransformed_col_indices <- col(outer10_site_macro_untransformed_braymat)[lower.tri(outer10_site_macro_untransformed_braymat)] # extract col indices = second quad in comp
+outer10_site_macro_untransformed_loc1 <- rownames(outer10_site_macro_untransformed_braymat)[outer10_site_macro_untransformed_row_indices] # populate appropriate vector
+outer10_site_macro_untransformed_loc2 <- colnames(outer10_site_macro_untransformed_braymat)[outer10_site_macro_untransformed_col_indices] # populate appropriate vector
+outer10_site_macro_untransformed_pairwise_df <- data.frame(loc1 = outer10_site_macro_untransformed_loc1, loc2 = outer10_site_macro_untransformed_loc2, dissimilarity = outer10_site_macro_untransformed_comp)
+outer10_site_macro_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer10_site_macro_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
+outer10_site_macro_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer10_site_macro_untransformed_pairwise_df$loc2)) # repeat for second quad
+outer10_site_macro_untransformed_pairwise_df$time <- abs(outer10_site_macro_untransformed_pairwise_df$year2 - outer10_site_macro_untransformed_pairwise_df$year1) # extract length of differences
+outer10_site_macro_untransformed_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", outer10_site_macro_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
+outer10_site_macro_untransformed_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", outer10_site_macro_untransformed_pairwise_df$loc2) # repeat for second quad
+
+# outer 17
+
+outer17_site_macro_species_matrix <- as.matrix(outer17_site_macro[ , -which(names(outer17_site_macro) %in% meta_cols)])
+outer17_site_macro_species_matrix <- outer17_site_macro_species_matrix/outer17_site_macro$hard_cover # convert from percent to proportion of hard cover
+rownames(outer17_site_macro_species_matrix) <- paste(outer17_site_macro$year,"-", outer17_site_macro$site_habitat, sep = "")
+
+# filter rare species
+# outer17_site_macro_species_presence <- colSums(outer17_site_macro_species_matrix > 0) # Presence absence matrix
+# outer17_site_macro_species_percentage <- (outer17_site_macro_species_presence/nrow(outer17_site_macro_species_matrix)) # proportion of rows where a species is present
+# outer17_site_macro_common_species <- names(outer17_site_macro_species_percentage[outer17_site_macro_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
+# outer17_site_macro_species_matrix <- outer17_site_macro_species_matrix[,colnames(outer17_site_macro_species_matrix) %in% outer17_site_macro_common_species] # filter original species matrix
+
+outer17_site_macro_rows_to_remove <- rowSums(outer17_site_macro_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
+outer17_site_macro_species_matrix <- outer17_site_macro_species_matrix[!outer17_site_macro_rows_to_remove,]
+
+outer17_site_macro_species_matrix_untransformed <- outer17_site_macro_species_matrix
+outer17_site_macro_species_matrix_transformed <- asin(sqrt(outer17_site_macro_species_matrix)) # arcsin squareroot transform
+outer17_site_macro_subset <- outer17_site_macro[!outer17_site_macro_rows_to_remove,] # make sure rows are also removed from dataframe
+outer17_site_macro_subset$stratum <- paste(outer17_site_macro_subset$year,"-", outer17_site_macro_subset$site_habitat, sep = "") # for merging with bray matrix later
+
+brayouter17_site_macro <- vegdist(outer17_site_macro_species_matrix_transformed, "bray") # create bray dissimilarity matrix
+# View(brayouter17_site_macro)
+brayouter17_site_macro_untransformed <- vegdist(outer17_site_macro_species_matrix_untransformed, "bray")
+outer17_site_macro_braymat <- as.matrix(brayouter17_site_macro) # conver to matrix object
+# convert to pairwise comparison
+outer17_site_macro_comp <- outer17_site_macro_braymat[lower.tri(outer17_site_macro_braymat)] # extract lower triangle
+outer17_site_macro_row_indices <- row(outer17_site_macro_braymat)[lower.tri(outer17_site_macro_braymat)] # extract row indices for first quad in comparison
+outer17_site_macro_col_indices <- col(outer17_site_macro_braymat)[lower.tri(outer17_site_macro_braymat)] # extract col indices = second quad in comp
+outer17_site_macro_loc1 <- rownames(outer17_site_macro_braymat)[outer17_site_macro_row_indices] # populate appropriate vector
+outer17_site_macro_loc2 <- colnames(outer17_site_macro_braymat)[outer17_site_macro_col_indices] # populate appropriate vector
+outer17_site_macro_pairwise_df <- data.frame(loc1 = outer17_site_macro_loc1, loc2 = outer17_site_macro_loc2, dissimilarity = outer17_site_macro_comp)
+outer17_site_macro_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer17_site_macro_pairwise_df$loc1)) # extract timepoint from first quad in comp
+outer17_site_macro_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer17_site_macro_pairwise_df$loc2)) # repeat for second quad
+outer17_site_macro_pairwise_df$time <- abs(outer17_site_macro_pairwise_df$year2 - outer17_site_macro_pairwise_df$year1) # extract length of differences
+outer17_site_macro_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", outer17_site_macro_pairwise_df$loc1) # extract quadrat id from first quad in comp
+outer17_site_macro_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", outer17_site_macro_pairwise_df$loc2) # repeat for second quad
+
+outer17_site_macro_untransformed_braymat <- as.matrix(brayouter17_site_macro_untransformed) # conver to matrix object
+# convert to pairwise comparison
+outer17_site_macro_untransformed_comp <- outer17_site_macro_untransformed_braymat[lower.tri(outer17_site_macro_untransformed_braymat)] # extract lower triangle
+outer17_site_macro_untransformed_row_indices <- row(outer17_site_macro_untransformed_braymat)[lower.tri(outer17_site_macro_untransformed_braymat)] # extract row indices for first quad in comparison
+outer17_site_macro_untransformed_col_indices <- col(outer17_site_macro_untransformed_braymat)[lower.tri(outer17_site_macro_untransformed_braymat)] # extract col indices = second quad in comp
+outer17_site_macro_untransformed_loc1 <- rownames(outer17_site_macro_untransformed_braymat)[outer17_site_macro_untransformed_row_indices] # populate appropriate vector
+outer17_site_macro_untransformed_loc2 <- colnames(outer17_site_macro_untransformed_braymat)[outer17_site_macro_untransformed_col_indices] # populate appropriate vector
+outer17_site_macro_untransformed_pairwise_df <- data.frame(loc1 = outer17_site_macro_untransformed_loc1, loc2 = outer17_site_macro_untransformed_loc2, dissimilarity = outer17_site_macro_untransformed_comp)
+outer17_site_macro_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer17_site_macro_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
+outer17_site_macro_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer17_site_macro_untransformed_pairwise_df$loc2)) # repeat for second quad
+outer17_site_macro_untransformed_pairwise_df$time <- abs(outer17_site_macro_untransformed_pairwise_df$year2 - outer17_site_macro_untransformed_pairwise_df$year1) # extract length of differences
+outer17_site_macro_untransformed_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", outer17_site_macro_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
+outer17_site_macro_untransformed_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", outer17_site_macro_untransformed_pairwise_df$loc2) # repeat for second quad
+
+```
+
+```{r site macroalgae bray dissimilarity over time, fig.height=15, fig.width=15}
+# for the purposes of this first set of analyses, I'm interested in dissimilarity of a given quad over time
+# not necessarily interested in how quads compare to each other over time/at each timepoint
+fringing_site_macro_pairwise_subset <- fringing_site_macro_pairwise_df[fringing_site_macro_pairwise_df$site_1 == fringing_site_macro_pairwise_df$site_2,]
+backreef_site_macro_pairwise_subset <- backreef_site_macro_pairwise_df[backreef_site_macro_pairwise_df$site_1 == backreef_site_macro_pairwise_df$site_2,]
+outer10_site_macro_pairwise_subset <- outer10_site_macro_pairwise_df[outer10_site_macro_pairwise_df$site_1 == outer10_site_macro_pairwise_df$site_2,]
+outer17_site_macro_pairwise_subset <- outer17_site_macro_pairwise_df[outer17_site_macro_pairwise_df$site_1 == outer17_site_macro_pairwise_df$site_2,]
+
+comb_site_macro_subset <- rbind(fringing_site_macro_pairwise_subset,
+                     backreef_site_macro_pairwise_subset,
+                     outer10_site_macro_pairwise_subset,
+                     outer17_site_macro_pairwise_subset)
+# site_1 and site_2 columns are all the same now
+# create new col just called for merge with other metadata
+comb_site_macro_subset$site_habitat <- comb_site_macro_subset$site_1
+
+bray_time_series_site_macro <- left_join(unique(site_meta_df[ , c("site", "site_habitat", "habitat", "island_side")]), comb_site_macro_subset, by = "site_habitat")
+
+# model
+
+# transformed dissimilarities
+bray_time_series_site_macro$dissimilarity_trans <- sv_trans(bray_time_series_site_macro$dissimilarity, bray_time_series_site_macro)
+
+bray_mod_site_macro <- glmmTMB(dissimilarity_trans ~ time*habitat + (1 | site),
+                              na.action = na.pass, family = beta_family(link = "logit"), data = bray_time_series_site_macro) #
+summary(bray_mod_site_macro) 
+hist(residuals(bray_mod_site_macro)) # slight skew but not awful
+plot(residuals(bray_mod_site_macro) ~ fitted(bray_mod_site_macro)) # stronger patterning here - negative slope
+
+emtrends(bray_mod_site_macro, pairwise ~ habitat , "time") # fringing slope significantly different from all others
+bray_trends_emm_site_macro <- emmip(bray_mod_site_macro, habitat ~ time , at = list(time = seq(1,16,1)), plotit = F, CIs = T)
+
+ggplot()+
+   geom_point(data = bray_time_series_site_macro, aes(x = time, y = dissimilarity, colour = habitat), alpha = 0.1 ) +
+   geom_line(data = bray_trends_emm_site_macro, aes(x = time, y = inv_logit(yvar), colour = habitat), linewidth = 2) +
+   geom_ribbon(data = bray_trends_emm_site_macro, aes(x = time, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+   scale_fill_manual(values = habitat_colours) +
+
+   labs(y = "Dissimilarity\n") +
+   theme_bw() + 
+   theme(panel.grid.major = element_blank(), 
+         panel.grid.minor = element_blank(),
+         panel.background = element_blank(),
+         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+         axis.line.y = element_line(colour = "black"),
+         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         axis.line.x = element_line(colour = "black"),
+         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         strip.text.x = element_text(size = 32),
+         legend.key.size = unit(2, "cm"), # increase legend size
+         legend.text = element_text(size = rel(2)), # increase legend text
+         legend.title = element_text(size = rel(2.25)),
+         legend.position = 'bottom', # move legend to bottom
+         legend.box = 'horizontal',
+         legend.box.background = element_rect(colour = "black") )# make legend horizontal
+```
+
+```{r site level ordination all algae, fig.height=15, fig.width=15}
+
+# get names of quads that remained after removing rare species
+relevant_sites <- c(rownames(fringing_site_species_matrix_transformed), rownames(backreef_site_species_matrix_transformed),
+                    rownames(outer10_site_species_matrix_transformed), rownames(outer17_site_species_matrix_transformed))
+
+full_site_species_matrix <- as.matrix(site_full[ , -which(names(site_full) %in% meta_cols)])
+full_site_species_matrix <- full_site_species_matrix/site_full$hard_cover # convert from percent to proportion of hard cover
+rownames(full_site_species_matrix) <- paste(site_full$year,"-", site_full$site_habitat, sep = "")
+
+full_site_species_matrix <- full_site_species_matrix[rownames(full_site_species_matrix) %in% relevant_sites,] # subset so only quads that didn't contain rare species are represented
+full_site_species_matrix <- asin(sqrt(full_site_species_matrix)) # arcsin squareroot transform
+site_full$stratum <- paste(site_full$year,"-", site_full$location, sep = "") # for merging with bray matrix later
+full_subset <- site_full[site_full$stratum %in% relevant_sites,] # make sure rows are also removed from dataframe
+
+full_columns_to_keep <- apply(full_site_species_matrix, 2, function(x) any(x >= asin(sqrt(0.1) )))
+full_abundant_site_species_matrix <- as.data.frame(full_site_species_matrix)[ , full_columns_to_keep]
+full_abundant_site_species_names <- colnames(full_abundant_site_species_matrix)
+
+# for mean centroid:
+
+full_site_meta_condensed <- site_meta_df %>%
+ group_by(year, habitat, hab_year) %>%
+ summarize() %>%
+ ungroup()
+
+set.seed(1032) 
+full_ord3_site <- metaMDS(full_site_species_matrix, dist	= "bray", k = 3, trymax = 100) 
+full_ord3_site # 0.149
+
+full.species.scores_site <- as.data.frame(scores(full_ord3_site, "species")) # some NaN for absent species
+full.species.scores_site <- na.omit(full.species.scores_site)
+full.species.scores_site$species <- rownames(full.species.scores_site) 
+full.abundant.species.scores_site <- full.species.scores_site[full.species.scores_site$species %in% full_abundant_site_species_names,]
+
+full.site.scores_site <- as.data.frame(vegan::scores(full_ord3_site, "sites"))  # extract the site scores and convert to a data.frame
+# add back in metadata
+full.site.scores_site$stratum <- rownames(full.site.scores_site)
+full.site.scores_site <- merge(site_meta_df, full.site.scores_site, by = "stratum")
+
+# write_csv(full.site.scores_site, here("data", "full_site_level_site_scores.csv"))
+# write_csv(full.species.scores_site, here("data", "full_site_level_species_scores.csv"))
+
+full.cent_site <- aggregate(cbind(NMDS1, NMDS2) ~ hab_year, data = full.site.scores_site, FUN = mean) 
+full.cent_site <- merge(full.cent_site, full_site_meta_condensed, by = "hab_year")
+
+# read in disturbances
+disturbances <- read.csv(here("data", "disturbance_list.csv"))
+full.cent_site <- merge(full.cent_site, disturbances, by = "year") # add disturbances to path
+library(ggrepel)
+ggplot() + 
+  geom_point(data= full.site.scores_site, aes(x=NMDS1,y=NMDS2, colour = habitat), alpha = 0.05) + # add the point markers
+  geom_text_repel(data= full.abundant.species.scores_site,aes(x=NMDS1,y=NMDS2, label = species), size = 4, alpha = 0.5) +  # add the species labels
+
+  geom_path(data = full.cent_site, aes(x = NMDS1, y = NMDS2, group = habitat, colour = habitat),
+            arrow = arrow(length=unit(0.25,"cm"), ends="last", type = "closed"), size = 1) +
+  geom_point(data = full.cent_site, aes(x = NMDS1, y = NMDS2, group = habitat, colour = habitat, shape = disturbance), size = 3, stroke = 1.4) +
+  scale_shape_manual(values = c(
+    "cots" = 8,
+    "cyclone" = 13,
+    "bleaching" = 4
+  )) +
+
+  scale_colour_manual(values = habitat_colours) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        strip.text.x = element_text(size = 25),
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.key.size = unit(2, "cm"), # increase legend size
+        legend.text = element_text(size = rel(1.5)), # increase legend text
+        legend.title = element_text(size = rel(1.5)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) + # add box around legend
+  guides(colour=guide_legend(title.position="top", # make color legend title go above legend
+                                     title.hjust =0.5, nrow = 2),
+         shape=guide_legend(title.position="top",  # make shape legend title go above legend
+                                     title.hjust =0.5,
+                            nrow = 2),
+         path = guide_legend(title.position="top",  # make shape legend title go above legend
+                                     title.hjust =0.5)
+         )
+
+```
+
+```{r site level ordination set up macroalgae, fig.height=15, fig.width=15}
+
+# get names of quads that remained after removing rare species
+relevant_site_macros <- c(rownames(fringing_site_macro_species_matrix_transformed), rownames(backreef_site_macro_species_matrix_transformed),
+                    rownames(outer10_site_macro_species_matrix_transformed), rownames(outer17_site_macro_species_matrix_transformed))
+
+full_site_macro_species_matrix <- as.matrix(site_macro[ , -which(names(site_macro) %in% meta_cols)])
+full_site_macro_species_matrix <- full_site_macro_species_matrix/site_macro$hard_cover # convert from percent to proportion of hard cover
+rownames(full_site_macro_species_matrix) <- paste(site_macro$year,"-", site_macro$site_habitat, sep = "")
+
+full_site_macro_species_matrix <- full_site_macro_species_matrix[rownames(full_site_macro_species_matrix) %in% relevant_site_macros,] # subset so only quads that didn't contain rare species are represented
+full_site_macro_species_matrix <- asin(sqrt(full_site_macro_species_matrix)) # arcsin squareroot transform
+site_macro$stratum <- paste(site_macro$year,"-", site_macro$location, sep = "") # for merging with bray matrix later
+full_subset_site_macro <- site_macro[site_macro$stratum %in% relevant_site_macros,] # make sure rows are also removed from dataframe
+
+full_columns_to_keep <- apply(full_site_macro_species_matrix, 2, function(x) any(x >= asin(sqrt(0.1) )))
+full_abundant_site_macro_species_matrix <- as.data.frame(full_site_macro_species_matrix)[ , full_columns_to_keep]
+full_abundant_site_macro_species_names <- colnames(full_abundant_site_macro_species_matrix)
+
+# for mean centroid:
+
+full_site_macro_meta_condensed <- site_meta_df %>%
+ group_by(year, habitat, hab_year) %>%
+ summarize() %>%
+ ungroup()
+
+set.seed(1032) 
+full_ord3_site_macro <- metaMDS(full_site_macro_species_matrix, dist	= "bray", k = 3, trymax = 100) 
+full_ord3_site_macro # 0.149
+
+full.species.scores_site_macro <- as.data.frame(scores(full_ord3_site_macro, "species")) # some NaN for absent species
+full.species.scores_site_macro <- na.omit(full.species.scores_site_macro)
+full.species.scores_site_macro$species <- rownames(full.species.scores_site_macro) 
+full.abundant.species.scores_site_macro <- full.species.scores_site_macro[full.species.scores_site_macro$species %in% full_abundant_site_macro_species_names,]
+
+full.site.scores_site_macro <- as.data.frame(vegan::scores(full_ord3_site_macro, "sites"))  # extract the site scores and convert to a data.frame
+# add back in metadata
+full.site.scores_site_macro$stratum <- rownames(full.site.scores_site_macro)
+full.site.scores_site_macro <- merge(site_meta_df, full.site.scores_site_macro, by = "stratum")
+
+# write_csv(full.site.scores_site_macro, here("data", "full_site_macro_level_site_macro_scores.csv"))
+# write_csv(full.species.scores_site_macro, here("data", "full_site_macro_level_species_scores.csv"))
+
+full.cent_site_macro <- aggregate(cbind(NMDS1, NMDS2) ~ hab_year, data = full.site.scores_site_macro, FUN = mean) 
+full.cent_site_macro <- merge(full.cent_site_macro, full_site_macro_meta_condensed, by = "hab_year")
+
+# read in disturbances
+disturbances <- read.csv(here("data", "disturbance_list.csv"))
+full.cent_site_macro <- merge(full.cent_site_macro, disturbances, by = "year") # add disturbances to path
+
+ggplot() + 
+  geom_point(data= full.site.scores_site_macro, aes(x=NMDS1,y=NMDS2, colour = habitat), alpha = 0.05) + # add the point markers
+  geom_text_repel(data= full.abundant.species.scores_site_macro,aes(x=NMDS1,y=NMDS2, label = species), size = 4, alpha = 0.5) +  # add the species labels
+
+  geom_path(data = full.cent_site_macro, aes(x = NMDS1, y = NMDS2, group = habitat, colour = habitat),
+            arrow = arrow(length=unit(0.25,"cm"), ends="last", type = "closed"), size = 1) +
+  geom_point(data = full.cent_site_macro, aes(x = NMDS1, y = NMDS2, group = habitat, colour = habitat, shape = disturbance), size = 3, stroke = 1.4) +
+  scale_shape_manual(values = c(
+    "cots" = 8,
+    "cyclone" = 13,
+    "bleaching" = 4
+  )) +
+
+  scale_colour_manual(values = habitat_colours) +
+
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        strip.text.x = element_text(size = 25),
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.key.size = unit(2, "cm"), # increase legend size
+        legend.text = element_text(size = rel(1.5)), # increase legend text
+        legend.title = element_text(size = rel(1.5)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) + # add box around legend
+  guides(colour=guide_legend(title.position="top", # make color legend title go above legend
+                                     title.hjust =0.5, nrow = 2),
+         shape=guide_legend(title.position="top",  # make shape legend title go above legend
+                                     title.hjust =0.5,
+                            nrow = 2),
+         path = guide_legend(title.position="top",  # make shape legend title go above legend
+                                     title.hjust =0.5)
+         )
+
+```
+
+----------------- QUADRAT BELOW
 ```{r fringing bray curtis}

### 2024-09-04 noam_beta_diversity.Rmd
finished nmds on macroalgae, did permanova, modeled beta dispersion

@@
 
-```{r site macroalgae level bray curtis setup}
-fringing_site_macro_species_matrix <- as.matrix(fringing_site_macro[ , -which(names(fringing_site_macro) %in% meta_cols)])
-fringing_site_macro_species_matrix <- fringing_site_macro_species_matrix/fringing_site_macro$hard_cover # convert from percent to proportion of hard cover
-rownames(fringing_site_macro_species_matrix) <- paste(fringing_site_macro$year,"-", fringing_site_macro$site_habitat, sep = "")
-
-# filter rare species
-# fringing_site_macro_species_presence <- colSums(fringing_site_macro_species_matrix > 0) # Presence absence matrix
-# fringing_site_macro_species_percentage <- (fringing_site_macro_species_presence/nrow(fringing_site_macro_species_matrix)) # proportion of rows where a species is present
-# fringing_site_macro_common_species <- names(fringing_site_macro_species_percentage[fringing_site_macro_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
-# fringing_site_macro_species_matrix <- fringing_site_macro_species_matrix[,colnames(fringing_site_macro_species_matrix) %in% fringing_site_macro_common_species] # filter original species matrix
-
-fringing_site_macro_rows_to_remove <- rowSums(fringing_site_macro_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
-fringing_site_macro_species_matrix <- fringing_site_macro_species_matrix[!fringing_site_macro_rows_to_remove,]
-
-fringing_site_macro_species_matrix_untransformed <- fringing_site_macro_species_matrix
-fringing_site_macro_species_matrix_transformed <- asin(sqrt(fringing_site_macro_species_matrix)) # arcsin squareroot transform
-fringing_site_macro_subset <- fringing_site_macro[!fringing_site_macro_rows_to_remove,] # make sure rows are also removed from dataframe
-fringing_site_macro_subset$stratum <- paste(fringing_site_macro_subset$year,"-", fringing_site_macro_subset$site_habitat, sep = "") # for merging with bray matrix later
-
-brayfringing_site_macro <- vegdist(fringing_site_macro_species_matrix_transformed, "bray") # create bray dissimilarity matrix
-# View(brayfringing_site_macro)
-brayfringing_site_macro_untransformed <- vegdist(fringing_site_macro_species_matrix_untransformed, "bray")
-fringing_site_macro_braymat <- as.matrix(brayfringing_site_macro) # conver to matrix object
-# convert to pairwise comparison
-fringing_site_macro_comp <- fringing_site_macro_braymat[lower.tri(fringing_site_macro_braymat)] # extract lower triangle
-fringing_site_macro_row_indices <- row(fringing_site_macro_braymat)[lower.tri(fringing_site_macro_braymat)] # extract row indices for first quad in comparison
-fringing_site_macro_col_indices <- col(fringing_site_macro_braymat)[lower.tri(fringing_site_macro_braymat)] # extract col indices = second quad in comp
-fringing_site_macro_loc1 <- rownames(fringing_site_macro_braymat)[fringing_site_macro_row_indices] # populate appropriate vector
-fringing_site_macro_loc2 <- colnames(fringing_site_macro_braymat)[fringing_site_macro_col_indices] # populate appropriate vector
-fringing_site_macro_pairwise_df <- data.frame(loc1 = fringing_site_macro_loc1, loc2 = fringing_site_macro_loc2, dissimilarity = fringing_site_macro_comp)
-fringing_site_macro_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", fringing_site_macro_pairwise_df$loc1)) # extract timepoint from first quad in comp
-fringing_site_macro_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", fringing_site_macro_pairwise_df$loc2)) # repeat for second quad
-fringing_site_macro_pairwise_df$time <- abs(fringing_site_macro_pairwise_df$year2 - fringing_site_macro_pairwise_df$year1) # extract length of differences
-fringing_site_macro_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", fringing_site_macro_pairwise_df$loc1) # extract quadrat id from first quad in comp
-fringing_site_macro_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", fringing_site_macro_pairwise_df$loc2) # repeat for second quad
-
-fringing_site_macro_untransformed_braymat <- as.matrix(brayfringing_site_macro_untransformed) # conver to matrix object
-# convert to pairwise comparison
-fringing_site_macro_untransformed_comp <- fringing_site_macro_untransformed_braymat[lower.tri(fringing_site_macro_untransformed_braymat)] # extract lower triangle
-fringing_site_macro_untransformed_row_indices <- row(fringing_site_macro_untransformed_braymat)[lower.tri(fringing_site_macro_untransformed_braymat)] # extract row indices for first quad in comparison
-fringing_site_macro_untransformed_col_indices <- col(fringing_site_macro_untransformed_braymat)[lower.tri(fringing_site_macro_untransformed_braymat)] # extract col indices = second quad in comp
-fringing_site_macro_untransformed_loc1 <- rownames(fringing_site_macro_untransformed_braymat)[fringing_site_macro_untransformed_row_indices] # populate appropriate vector
-fringing_site_macro_untransformed_loc2 <- colnames(fringing_site_macro_untransformed_braymat)[fringing_site_macro_untransformed_col_indices] # populate appropriate vector
-fringing_site_macro_untransformed_pairwise_df <- data.frame(loc1 = fringing_site_macro_untransformed_loc1, loc2 = fringing_site_macro_untransformed_loc2, dissimilarity = fringing_site_macro_untransformed_comp)
-fringing_site_macro_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", fringing_site_macro_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
-fringing_site_macro_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", fringing_site_macro_untransformed_pairwise_df$loc2)) # repeat for second quad
-fringing_site_macro_untransformed_pairwise_df$time <- abs(fringing_site_macro_untransformed_pairwise_df$year2 - fringing_site_macro_untransformed_pairwise_df$year1) # extract length of differences
-fringing_site_macro_untransformed_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", fringing_site_macro_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
-fringing_site_macro_untransformed_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", fringing_site_macro_untransformed_pairwise_df$loc2) # repeat for second quad
-
-# backreef
-
-backreef_site_macro_species_matrix <- as.matrix(backreef_site_macro[ , -which(names(backreef_site_macro) %in% meta_cols)])
-backreef_site_macro_species_matrix <- backreef_site_macro_species_matrix/backreef_site_macro$hard_cover # convert from percent to proportion of hard cover
-rownames(backreef_site_macro_species_matrix) <- paste(backreef_site_macro$year,"-", backreef_site_macro$site_habitat, sep = "")
-
-# filter rare species
-# backreef_site_macro_species_presence <- colSums(backreef_site_macro_species_matrix > 0) # Presence absence matrix
-# backreef_site_macro_species_percentage <- (backreef_site_macro_species_presence/nrow(backreef_site_macro_species_matrix)) # proportion of rows where a species is present
-# backreef_site_macro_common_species <- names(backreef_site_macro_species_percentage[backreef_site_macro_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
-# backreef_site_macro_species_matrix <- backreef_site_macro_species_matrix[,colnames(backreef_site_macro_species_matrix) %in% backreef_site_macro_common_species] # filter original species matrix
-
-backreef_site_macro_rows_to_remove <- rowSums(backreef_site_macro_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
-backreef_site_macro_species_matrix <- backreef_site_macro_species_matrix[!backreef_site_macro_rows_to_remove,]
-
-backreef_site_macro_species_matrix_untransformed <- backreef_site_macro_species_matrix
-backreef_site_macro_species_matrix_transformed <- asin(sqrt(backreef_site_macro_species_matrix)) # arcsin squareroot transform
-backreef_site_macro_subset <- backreef_site_macro[!backreef_site_macro_rows_to_remove,] # make sure rows are also removed from dataframe
-backreef_site_macro_subset$stratum <- paste(backreef_site_macro_subset$year,"-", backreef_site_macro_subset$site_habitat, sep = "") # for merging with bray matrix later
-
-braybackreef_site_macro <- vegdist(backreef_site_macro_species_matrix_transformed, "bray") # create bray dissimilarity matrix
-# View(braybackreef_site_macro)
-braybackreef_site_macro_untransformed <- vegdist(backreef_site_macro_species_matrix_untransformed, "bray")
-backreef_site_macro_braymat <- as.matrix(braybackreef_site_macro) # conver to matrix object
-# convert to pairwise comparison
-backreef_site_macro_comp <- backreef_site_macro_braymat[lower.tri(backreef_site_macro_braymat)] # extract lower triangle
-backreef_site_macro_row_indices <- row(backreef_site_macro_braymat)[lower.tri(backreef_site_macro_braymat)] # extract row indices for first quad in comparison
-backreef_site_macro_col_indices <- col(backreef_site_macro_braymat)[lower.tri(backreef_site_macro_braymat)] # extract col indices = second quad in comp
-backreef_site_macro_loc1 <- rownames(backreef_site_macro_braymat)[backreef_site_macro_row_indices] # populate appropriate vector
-backreef_site_macro_loc2 <- colnames(backreef_site_macro_braymat)[backreef_site_macro_col_indices] # populate appropriate vector
-backreef_site_macro_pairwise_df <- data.frame(loc1 = backreef_site_macro_loc1, loc2 = backreef_site_macro_loc2, dissimilarity = backreef_site_macro_comp)
-backreef_site_macro_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", backreef_site_macro_pairwise_df$loc1)) # extract timepoint from first quad in comp
-backreef_site_macro_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", backreef_site_macro_pairwise_df$loc2)) # repeat for second quad
-backreef_site_macro_pairwise_df$time <- abs(backreef_site_macro_pairwise_df$year2 - backreef_site_macro_pairwise_df$year1) # extract length of differences
-backreef_site_macro_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", backreef_site_macro_pairwise_df$loc1) # extract quadrat id from first quad in comp
-backreef_site_macro_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", backreef_site_macro_pairwise_df$loc2) # repeat for second quad
-
-backreef_site_macro_untransformed_braymat <- as.matrix(braybackreef_site_macro_untransformed) # conver to matrix object
-# convert to pairwise comparison
-backreef_site_macro_untransformed_comp <- backreef_site_macro_untransformed_braymat[lower.tri(backreef_site_macro_untransformed_braymat)] # extract lower triangle
-backreef_site_macro_untransformed_row_indices <- row(backreef_site_macro_untransformed_braymat)[lower.tri(backreef_site_macro_untransformed_braymat)] # extract row indices for first quad in comparison
-backreef_site_macro_untransformed_col_indices <- col(backreef_site_macro_untransformed_braymat)[lower.tri(backreef_site_macro_untransformed_braymat)] # extract col indices = second quad in comp
-backreef_site_macro_untransformed_loc1 <- rownames(backreef_site_macro_untransformed_braymat)[backreef_site_macro_untransformed_row_indices] # populate appropriate vector
-backreef_site_macro_untransformed_loc2 <- colnames(backreef_site_macro_untransformed_braymat)[backreef_site_macro_untransformed_col_indices] # populate appropriate vector
-backreef_site_macro_untransformed_pairwise_df <- data.frame(loc1 = backreef_site_macro_untransformed_loc1, loc2 = backreef_site_macro_untransformed_loc2, dissimilarity = backreef_site_macro_untransformed_comp)
-backreef_site_macro_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", backreef_site_macro_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
-backreef_site_macro_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", backreef_site_macro_untransformed_pairwise_df$loc2)) # repeat for second quad
-backreef_site_macro_untransformed_pairwise_df$time <- abs(backreef_site_macro_untransformed_pairwise_df$year2 - backreef_site_macro_untransformed_pairwise_df$year1) # extract length of differences
-backreef_site_macro_untransformed_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", backreef_site_macro_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
-backreef_site_macro_untransformed_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", backreef_site_macro_untransformed_pairwise_df$loc2) # repeat for second quad
-
-# outer10
-
-outer10_site_macro_species_matrix <- as.matrix(outer10_site_macro[ , -which(names(outer10_site_macro) %in% meta_cols)])
-outer10_site_macro_species_matrix <- outer10_site_macro_species_matrix/outer10_site_macro$hard_cover # convert from percent to proportion of hard cover
-rownames(outer10_site_macro_species_matrix) <- paste(outer10_site_macro$year,"-", outer10_site_macro$site_habitat, sep = "")
-
-# filter rare species
-# outer10_site_macro_species_presence <- colSums(outer10_site_macro_species_matrix > 0) # Presence absence matrix
-# outer10_site_macro_species_percentage <- (outer10_site_macro_species_presence/nrow(outer10_site_macro_species_matrix)) # proportion of rows where a species is present
-# outer10_site_macro_common_species <- names(outer10_site_macro_species_percentage[outer10_site_macro_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
-# outer10_site_macro_species_matrix <- outer10_site_macro_species_matrix[,colnames(outer10_site_macro_species_matrix) %in% outer10_site_macro_common_species] # filter original species matrix
-
-outer10_site_macro_rows_to_remove <- rowSums(outer10_site_macro_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
-outer10_site_macro_species_matrix <- outer10_site_macro_species_matrix[!outer10_site_macro_rows_to_remove,]
-
-outer10_site_macro_species_matrix_untransformed <- outer10_site_macro_species_matrix
-outer10_site_macro_species_matrix_transformed <- asin(sqrt(outer10_site_macro_species_matrix)) # arcsin squareroot transform
-outer10_site_macro_subset <- outer10_site_macro[!outer10_site_macro_rows_to_remove,] # make sure rows are also removed from dataframe
-outer10_site_macro_subset$stratum <- paste(outer10_site_macro_subset$year,"-", outer10_site_macro_subset$site_habitat, sep = "") # for merging with bray matrix later
-
-brayouter10_site_macro <- vegdist(outer10_site_macro_species_matrix_transformed, "bray") # create bray dissimilarity matrix
-# View(brayouter10_site_macro)
-brayouter10_site_macro_untransformed <- vegdist(outer10_site_macro_species_matrix_untransformed, "bray")
-outer10_site_macro_braymat <- as.matrix(brayouter10_site_macro) # conver to matrix object
-# convert to pairwise comparison
-outer10_site_macro_comp <- outer10_site_macro_braymat[lower.tri(outer10_site_macro_braymat)] # extract lower triangle
-outer10_site_macro_row_indices <- row(outer10_site_macro_braymat)[lower.tri(outer10_site_macro_braymat)] # extract row indices for first quad in comparison
-outer10_site_macro_col_indices <- col(outer10_site_macro_braymat)[lower.tri(outer10_site_macro_braymat)] # extract col indices = second quad in comp
-outer10_site_macro_loc1 <- rownames(outer10_site_macro_braymat)[outer10_site_macro_row_indices] # populate appropriate vector
-outer10_site_macro_loc2 <- colnames(outer10_site_macro_braymat)[outer10_site_macro_col_indices] # populate appropriate vector
-outer10_site_macro_pairwise_df <- data.frame(loc1 = outer10_site_macro_loc1, loc2 = outer10_site_macro_loc2, dissimilarity = outer10_site_macro_comp)
-outer10_site_macro_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer10_site_macro_pairwise_df$loc1)) # extract timepoint from first quad in comp
-outer10_site_macro_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer10_site_macro_pairwise_df$loc2)) # repeat for second quad
-outer10_site_macro_pairwise_df$time <- abs(outer10_site_macro_pairwise_df$year2 - outer10_site_macro_pairwise_df$year1) # extract length of differences
-outer10_site_macro_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", outer10_site_macro_pairwise_df$loc1) # extract quadrat id from first quad in comp
-outer10_site_macro_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", outer10_site_macro_pairwise_df$loc2) # repeat for second quad
-
-outer10_site_macro_untransformed_braymat <- as.matrix(brayouter10_site_macro_untransformed) # conver to matrix object
-# convert to pairwise comparison
-outer10_site_macro_untransformed_comp <- outer10_site_macro_untransformed_braymat[lower.tri(outer10_site_macro_untransformed_braymat)] # extract lower triangle
-outer10_site_macro_untransformed_row_indices <- row(outer10_site_macro_untransformed_braymat)[lower.tri(outer10_site_macro_untransformed_braymat)] # extract row indices for first quad in comparison
-outer10_site_macro_untransformed_col_indices <- col(outer10_site_macro_untransformed_braymat)[lower.tri(outer10_site_macro_untransformed_braymat)] # extract col indices = second quad in comp
-outer10_site_macro_untransformed_loc1 <- rownames(outer10_site_macro_untransformed_braymat)[outer10_site_macro_untransformed_row_indices] # populate appropriate vector
-outer10_site_macro_untransformed_loc2 <- colnames(outer10_site_macro_untransformed_braymat)[outer10_site_macro_untransformed_col_indices] # populate appropriate vector
-outer10_site_macro_untransformed_pairwise_df <- data.frame(loc1 = outer10_site_macro_untransformed_loc1, loc2 = outer10_site_macro_untransformed_loc2, dissimilarity = outer10_site_macro_untransformed_comp)
-outer10_site_macro_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer10_site_macro_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
-outer10_site_macro_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer10_site_macro_untransformed_pairwise_df$loc2)) # repeat for second quad
-outer10_site_macro_untransformed_pairwise_df$time <- abs(outer10_site_macro_untransformed_pairwise_df$year2 - outer10_site_macro_untransformed_pairwise_df$year1) # extract length of differences
-outer10_site_macro_untransformed_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", outer10_site_macro_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
-outer10_site_macro_untransformed_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", outer10_site_macro_untransformed_pairwise_df$loc2) # repeat for second quad
-
-# outer 17
-
-outer17_site_macro_species_matrix <- as.matrix(outer17_site_macro[ , -which(names(outer17_site_macro) %in% meta_cols)])
-outer17_site_macro_species_matrix <- outer17_site_macro_species_matrix/outer17_site_macro$hard_cover # convert from percent to proportion of hard cover
-rownames(outer17_site_macro_species_matrix) <- paste(outer17_site_macro$year,"-", outer17_site_macro$site_habitat, sep = "")
-
-# filter rare species
-# outer17_site_macro_species_presence <- colSums(outer17_site_macro_species_matrix > 0) # Presence absence matrix
-# outer17_site_macro_species_percentage <- (outer17_site_macro_species_presence/nrow(outer17_site_macro_species_matrix)) # proportion of rows where a species is present
-# outer17_site_macro_common_species <- names(outer17_site_macro_species_percentage[outer17_site_macro_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
-# outer17_site_macro_species_matrix <- outer17_site_macro_species_matrix[,colnames(outer17_site_macro_species_matrix) %in% outer17_site_macro_common_species] # filter original species matrix
-
-outer17_site_macro_rows_to_remove <- rowSums(outer17_site_macro_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
-outer17_site_macro_species_matrix <- outer17_site_macro_species_matrix[!outer17_site_macro_rows_to_remove,]
-
-outer17_site_macro_species_matrix_untransformed <- outer17_site_macro_species_matrix
-outer17_site_macro_species_matrix_transformed <- asin(sqrt(outer17_site_macro_species_matrix)) # arcsin squareroot transform
-outer17_site_macro_subset <- outer17_site_macro[!outer17_site_macro_rows_to_remove,] # make sure rows are also removed from dataframe
-outer17_site_macro_subset$stratum <- paste(outer17_site_macro_subset$year,"-", outer17_site_macro_subset$site_habitat, sep = "") # for merging with bray matrix later
-
-brayouter17_site_macro <- vegdist(outer17_site_macro_species_matrix_transformed, "bray") # create bray dissimilarity matrix
-# View(brayouter17_site_macro)
-brayouter17_site_macro_untransformed <- vegdist(outer17_site_macro_species_matrix_untransformed, "bray")
-outer17_site_macro_braymat <- as.matrix(brayouter17_site_macro) # conver to matrix object
-# convert to pairwise comparison
-outer17_site_macro_comp <- outer17_site_macro_braymat[lower.tri(outer17_site_macro_braymat)] # extract lower triangle
-outer17_site_macro_row_indices <- row(outer17_site_macro_braymat)[lower.tri(outer17_site_macro_braymat)] # extract row indices for first quad in comparison
-outer17_site_macro_col_indices <- col(outer17_site_macro_braymat)[lower.tri(outer17_site_macro_braymat)] # extract col indices = second quad in comp
-outer17_site_macro_loc1 <- rownames(outer17_site_macro_braymat)[outer17_site_macro_row_indices] # populate appropriate vector
-outer17_site_macro_loc2 <- colnames(outer17_site_macro_braymat)[outer17_site_macro_col_indices] # populate appropriate vector
-outer17_site_macro_pairwise_df <- data.frame(loc1 = outer17_site_macro_loc1, loc2 = outer17_site_macro_loc2, dissimilarity = outer17_site_macro_comp)
-outer17_site_macro_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer17_site_macro_pairwise_df$loc1)) # extract timepoint from first quad in comp
-outer17_site_macro_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer17_site_macro_pairwise_df$loc2)) # repeat for second quad
-outer17_site_macro_pairwise_df$time <- abs(outer17_site_macro_pairwise_df$year2 - outer17_site_macro_pairwise_df$year1) # extract length of differences
-outer17_site_macro_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", outer17_site_macro_pairwise_df$loc1) # extract quadrat id from first quad in comp
-outer17_site_macro_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", outer17_site_macro_pairwise_df$loc2) # repeat for second quad
-
-outer17_site_macro_untransformed_braymat <- as.matrix(brayouter17_site_macro_untransformed) # conver to matrix object
-# convert to pairwise comparison
-outer17_site_macro_untransformed_comp <- outer17_site_macro_untransformed_braymat[lower.tri(outer17_site_macro_untransformed_braymat)] # extract lower triangle
-outer17_site_macro_untransformed_row_indices <- row(outer17_site_macro_untransformed_braymat)[lower.tri(outer17_site_macro_untransformed_braymat)] # extract row indices for first quad in comparison
-outer17_site_macro_untransformed_col_indices <- col(outer17_site_macro_untransformed_braymat)[lower.tri(outer17_site_macro_untransformed_braymat)] # extract col indices = second quad in comp
-outer17_site_macro_untransformed_loc1 <- rownames(outer17_site_macro_untransformed_braymat)[outer17_site_macro_untransformed_row_indices] # populate appropriate vector
-outer17_site_macro_untransformed_loc2 <- colnames(outer17_site_macro_untransformed_braymat)[outer17_site_macro_untransformed_col_indices] # populate appropriate vector
-outer17_site_macro_untransformed_pairwise_df <- data.frame(loc1 = outer17_site_macro_untransformed_loc1, loc2 = outer17_site_macro_untransformed_loc2, dissimilarity = outer17_site_macro_untransformed_comp)
-outer17_site_macro_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer17_site_macro_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
-outer17_site_macro_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer17_site_macro_untransformed_pairwise_df$loc2)) # repeat for second quad
-outer17_site_macro_untransformed_pairwise_df$time <- abs(outer17_site_macro_untransformed_pairwise_df$year2 - outer17_site_macro_untransformed_pairwise_df$year1) # extract length of differences
-outer17_site_macro_untransformed_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", outer17_site_macro_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
-outer17_site_macro_untransformed_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", outer17_site_macro_untransformed_pairwise_df$loc2) # repeat for second quad
-
-```
-
-```{r site macroalgae bray dissimilarity over time, fig.height=15, fig.width=15}
-# for the purposes of this first set of analyses, I'm interested in dissimilarity of a given quad over time
-# not necessarily interested in how quads compare to each other over time/at each timepoint
-fringing_site_macro_pairwise_subset <- fringing_site_macro_pairwise_df[fringing_site_macro_pairwise_df$site_1 == fringing_site_macro_pairwise_df$site_2,]
-backreef_site_macro_pairwise_subset <- backreef_site_macro_pairwise_df[backreef_site_macro_pairwise_df$site_1 == backreef_site_macro_pairwise_df$site_2,]
-outer10_site_macro_pairwise_subset <- outer10_site_macro_pairwise_df[outer10_site_macro_pairwise_df$site_1 == outer10_site_macro_pairwise_df$site_2,]
-outer17_site_macro_pairwise_subset <- outer17_site_macro_pairwise_df[outer17_site_macro_pairwise_df$site_1 == outer17_site_macro_pairwise_df$site_2,]
-
-comb_site_macro_subset <- rbind(fringing_site_macro_pairwise_subset,
-                     backreef_site_macro_pairwise_subset,
-                     outer10_site_macro_pairwise_subset,
-                     outer17_site_macro_pairwise_subset)
-# site_1 and site_2 columns are all the same now
-# create new col just called for merge with other metadata
-comb_site_macro_subset$site_habitat <- comb_site_macro_subset$site_1
-
-bray_time_series_site_macro <- left_join(unique(site_meta_df[ , c("site", "site_habitat", "habitat", "island_side")]), comb_site_macro_subset, by = "site_habitat")
-
-# model
-
-# transformed dissimilarities
-bray_time_series_site_macro$dissimilarity_trans <- sv_trans(bray_time_series_site_macro$dissimilarity, bray_time_series_site_macro)
-
-bray_mod_site_macro <- glmmTMB(dissimilarity_trans ~ time*habitat + (1 | site),
-                              na.action = na.pass, family = beta_family(link = "logit"), data = bray_time_series_site_macro) #
-summary(bray_mod_site_macro) 
-hist(residuals(bray_mod_site_macro)) # slight skew but not awful
-plot(residuals(bray_mod_site_macro) ~ fitted(bray_mod_site_macro)) # stronger patterning here - negative slope
-
-emtrends(bray_mod_site_macro, pairwise ~ habitat , "time") # fringing slope significantly different from all others
-bray_trends_emm_site_macro <- emmip(bray_mod_site_macro, habitat ~ time , at = list(time = seq(1,16,1)), plotit = F, CIs = T)
+```{r dissimilarity compared to baseline all, fig.height=15, fig.width = 15}
+bray_time_series_site_baseline <- subset(bray_time_series_site, year2 == 2007)
+bray_mod_site_baseline <- glmmTMB(dissimilarity_trans ~ time*habitat + (1 | site),
+                              na.action = na.pass, family = beta_family(link = "logit"), data = bray_time_series_site_baseline)
+hist(residuals(bray_mod_site_baseline)) # long tails but not bad
+plot(residuals(bray_mod_site_baseline) ~ fitted(bray_mod_site_baseline) ) # heteroskedastic
+
+emtrends(bray_mod_site_baseline, pairwise ~ habitat , "time") # fringing slope significantly different from all others
+bray_trends_emm_site_baseline <- emmip(bray_mod_site_baseline, habitat ~ time , at = list(time = seq(1,16,1)), plotit = F, CIs = T)
 
 ggplot()+
-   geom_point(data = bray_time_series_site_macro, aes(x = time, y = dissimilarity, colour = habitat), alpha = 0.1 ) +
-   geom_line(data = bray_trends_emm_site_macro, aes(x = time, y = inv_logit(yvar), colour = habitat), linewidth = 2) +
-   geom_ribbon(data = bray_trends_emm_site_macro, aes(x = time, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
+   geom_point(data = bray_time_series_site_baseline, aes(x = time, y = dissimilarity, colour = habitat), alpha = 0.1 ) +
+   geom_line(data = bray_trends_emm_site_baseline, aes(x = time, y = inv_logit(yvar), colour = habitat), linewidth = 2) +
+   geom_ribbon(data = bray_trends_emm_site_baseline, aes(x = time, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
    scale_colour_manual(values = habitat_colours) +
@@
 
+```{r site macroalgae level bray curtis setup}
+fringing_site_macro_species_matrix <- as.matrix(fringing_site_macro[ , -which(names(fringing_site_macro) %in% meta_cols)])
+fringing_site_macro_species_matrix <- fringing_site_macro_species_matrix/fringing_site_macro$hard_cover # convert from percent to proportion of hard cover
+rownames(fringing_site_macro_species_matrix) <- paste(fringing_site_macro$year,"-", fringing_site_macro$site_habitat, sep = "")
+
+# filter rare species
+# fringing_site_macro_species_presence <- colSums(fringing_site_macro_species_matrix > 0) # Presence absence matrix
+# fringing_site_macro_species_percentage <- (fringing_site_macro_species_presence/nrow(fringing_site_macro_species_matrix)) # proportion of rows where a species is present
+# fringing_site_macro_common_species <- names(fringing_site_macro_species_percentage[fringing_site_macro_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
+# fringing_site_macro_species_matrix <- fringing_site_macro_species_matrix[,colnames(fringing_site_macro_species_matrix) %in% fringing_site_macro_common_species] # filter original species matrix
+
+fringing_site_macro_rows_to_remove <- rowSums(fringing_site_macro_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
+fringing_site_macro_species_matrix <- fringing_site_macro_species_matrix[!fringing_site_macro_rows_to_remove,]
+
+fringing_site_macro_species_matrix_untransformed <- fringing_site_macro_species_matrix
+fringing_site_macro_species_matrix_transformed <- asin(sqrt(fringing_site_macro_species_matrix)) # arcsin squareroot transform
+fringing_site_macro_subset <- fringing_site_macro[!fringing_site_macro_rows_to_remove,] # make sure rows are also removed from dataframe
+fringing_site_macro_subset$stratum <- paste(fringing_site_macro_subset$year,"-", fringing_site_macro_subset$site_habitat, sep = "") # for merging with bray matrix later
+
+brayfringing_site_macro <- vegdist(fringing_site_macro_species_matrix_transformed, "bray") # create bray dissimilarity matrix
+# View(brayfringing_site_macro)
+brayfringing_site_macro_untransformed <- vegdist(fringing_site_macro_species_matrix_untransformed, "bray")
+fringing_site_macro_braymat <- as.matrix(brayfringing_site_macro) # conver to matrix object
+# convert to pairwise comparison
+fringing_site_macro_comp <- fringing_site_macro_braymat[lower.tri(fringing_site_macro_braymat)] # extract lower triangle
+fringing_site_macro_row_indices <- row(fringing_site_macro_braymat)[lower.tri(fringing_site_macro_braymat)] # extract row indices for first quad in comparison
+fringing_site_macro_col_indices <- col(fringing_site_macro_braymat)[lower.tri(fringing_site_macro_braymat)] # extract col indices = second quad in comp
+fringing_site_macro_loc1 <- rownames(fringing_site_macro_braymat)[fringing_site_macro_row_indices] # populate appropriate vector
+fringing_site_macro_loc2 <- colnames(fringing_site_macro_braymat)[fringing_site_macro_col_indices] # populate appropriate vector
+fringing_site_macro_pairwise_df <- data.frame(loc1 = fringing_site_macro_loc1, loc2 = fringing_site_macro_loc2, dissimilarity = fringing_site_macro_comp)
+fringing_site_macro_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", fringing_site_macro_pairwise_df$loc1)) # extract timepoint from first quad in comp
+fringing_site_macro_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", fringing_site_macro_pairwise_df$loc2)) # repeat for second quad
+fringing_site_macro_pairwise_df$time <- abs(fringing_site_macro_pairwise_df$year2 - fringing_site_macro_pairwise_df$year1) # extract length of differences
+fringing_site_macro_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", fringing_site_macro_pairwise_df$loc1) # extract quadrat id from first quad in comp
+fringing_site_macro_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", fringing_site_macro_pairwise_df$loc2) # repeat for second quad
+
+fringing_site_macro_untransformed_braymat <- as.matrix(brayfringing_site_macro_untransformed) # conver to matrix object
+# convert to pairwise comparison
+fringing_site_macro_untransformed_comp <- fringing_site_macro_untransformed_braymat[lower.tri(fringing_site_macro_untransformed_braymat)] # extract lower triangle
+fringing_site_macro_untransformed_row_indices <- row(fringing_site_macro_untransformed_braymat)[lower.tri(fringing_site_macro_untransformed_braymat)] # extract row indices for first quad in comparison
+fringing_site_macro_untransformed_col_indices <- col(fringing_site_macro_untransformed_braymat)[lower.tri(fringing_site_macro_untransformed_braymat)] # extract col indices = second quad in comp
+fringing_site_macro_untransformed_loc1 <- rownames(fringing_site_macro_untransformed_braymat)[fringing_site_macro_untransformed_row_indices] # populate appropriate vector
+fringing_site_macro_untransformed_loc2 <- colnames(fringing_site_macro_untransformed_braymat)[fringing_site_macro_untransformed_col_indices] # populate appropriate vector
+fringing_site_macro_untransformed_pairwise_df <- data.frame(loc1 = fringing_site_macro_untransformed_loc1, loc2 = fringing_site_macro_untransformed_loc2, dissimilarity = fringing_site_macro_untransformed_comp)
+fringing_site_macro_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", fringing_site_macro_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
+fringing_site_macro_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", fringing_site_macro_untransformed_pairwise_df$loc2)) # repeat for second quad
+fringing_site_macro_untransformed_pairwise_df$time <- abs(fringing_site_macro_untransformed_pairwise_df$year2 - fringing_site_macro_untransformed_pairwise_df$year1) # extract length of differences
+fringing_site_macro_untransformed_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", fringing_site_macro_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
+fringing_site_macro_untransformed_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", fringing_site_macro_untransformed_pairwise_df$loc2) # repeat for second quad
+
+# backreef
+
+backreef_site_macro_species_matrix <- as.matrix(backreef_site_macro[ , -which(names(backreef_site_macro) %in% meta_cols)])
+backreef_site_macro_species_matrix <- backreef_site_macro_species_matrix/backreef_site_macro$hard_cover # convert from percent to proportion of hard cover
+rownames(backreef_site_macro_species_matrix) <- paste(backreef_site_macro$year,"-", backreef_site_macro$site_habitat, sep = "")
+
+# filter rare species
+# backreef_site_macro_species_presence <- colSums(backreef_site_macro_species_matrix > 0) # Presence absence matrix
+# backreef_site_macro_species_percentage <- (backreef_site_macro_species_presence/nrow(backreef_site_macro_species_matrix)) # proportion of rows where a species is present
+# backreef_site_macro_common_species <- names(backreef_site_macro_species_percentage[backreef_site_macro_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
+# backreef_site_macro_species_matrix <- backreef_site_macro_species_matrix[,colnames(backreef_site_macro_species_matrix) %in% backreef_site_macro_common_species] # filter original species matrix
+
+backreef_site_macro_rows_to_remove <- rowSums(backreef_site_macro_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
+backreef_site_macro_species_matrix <- backreef_site_macro_species_matrix[!backreef_site_macro_rows_to_remove,]
+
+backreef_site_macro_species_matrix_untransformed <- backreef_site_macro_species_matrix
+backreef_site_macro_species_matrix_transformed <- asin(sqrt(backreef_site_macro_species_matrix)) # arcsin squareroot transform
+backreef_site_macro_subset <- backreef_site_macro[!backreef_site_macro_rows_to_remove,] # make sure rows are also removed from dataframe
+backreef_site_macro_subset$stratum <- paste(backreef_site_macro_subset$year,"-", backreef_site_macro_subset$site_habitat, sep = "") # for merging with bray matrix later
+
+braybackreef_site_macro <- vegdist(backreef_site_macro_species_matrix_transformed, "bray") # create bray dissimilarity matrix
+# View(braybackreef_site_macro)
+braybackreef_site_macro_untransformed <- vegdist(backreef_site_macro_species_matrix_untransformed, "bray")
+backreef_site_macro_braymat <- as.matrix(braybackreef_site_macro) # conver to matrix object
+# convert to pairwise comparison
+backreef_site_macro_comp <- backreef_site_macro_braymat[lower.tri(backreef_site_macro_braymat)] # extract lower triangle
+backreef_site_macro_row_indices <- row(backreef_site_macro_braymat)[lower.tri(backreef_site_macro_braymat)] # extract row indices for first quad in comparison
+backreef_site_macro_col_indices <- col(backreef_site_macro_braymat)[lower.tri(backreef_site_macro_braymat)] # extract col indices = second quad in comp
+backreef_site_macro_loc1 <- rownames(backreef_site_macro_braymat)[backreef_site_macro_row_indices] # populate appropriate vector
+backreef_site_macro_loc2 <- colnames(backreef_site_macro_braymat)[backreef_site_macro_col_indices] # populate appropriate vector
+backreef_site_macro_pairwise_df <- data.frame(loc1 = backreef_site_macro_loc1, loc2 = backreef_site_macro_loc2, dissimilarity = backreef_site_macro_comp)
+backreef_site_macro_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", backreef_site_macro_pairwise_df$loc1)) # extract timepoint from first quad in comp
+backreef_site_macro_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", backreef_site_macro_pairwise_df$loc2)) # repeat for second quad
+backreef_site_macro_pairwise_df$time <- abs(backreef_site_macro_pairwise_df$year2 - backreef_site_macro_pairwise_df$year1) # extract length of differences
+backreef_site_macro_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", backreef_site_macro_pairwise_df$loc1) # extract quadrat id from first quad in comp
+backreef_site_macro_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", backreef_site_macro_pairwise_df$loc2) # repeat for second quad
+
+backreef_site_macro_untransformed_braymat <- as.matrix(braybackreef_site_macro_untransformed) # conver to matrix object
+# convert to pairwise comparison
+backreef_site_macro_untransformed_comp <- backreef_site_macro_untransformed_braymat[lower.tri(backreef_site_macro_untransformed_braymat)] # extract lower triangle
+backreef_site_macro_untransformed_row_indices <- row(backreef_site_macro_untransformed_braymat)[lower.tri(backreef_site_macro_untransformed_braymat)] # extract row indices for first quad in comparison
+backreef_site_macro_untransformed_col_indices <- col(backreef_site_macro_untransformed_braymat)[lower.tri(backreef_site_macro_untransformed_braymat)] # extract col indices = second quad in comp
+backreef_site_macro_untransformed_loc1 <- rownames(backreef_site_macro_untransformed_braymat)[backreef_site_macro_untransformed_row_indices] # populate appropriate vector
+backreef_site_macro_untransformed_loc2 <- colnames(backreef_site_macro_untransformed_braymat)[backreef_site_macro_untransformed_col_indices] # populate appropriate vector
+backreef_site_macro_untransformed_pairwise_df <- data.frame(loc1 = backreef_site_macro_untransformed_loc1, loc2 = backreef_site_macro_untransformed_loc2, dissimilarity = backreef_site_macro_untransformed_comp)
+backreef_site_macro_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", backreef_site_macro_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
+backreef_site_macro_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", backreef_site_macro_untransformed_pairwise_df$loc2)) # repeat for second quad
+backreef_site_macro_untransformed_pairwise_df$time <- abs(backreef_site_macro_untransformed_pairwise_df$year2 - backreef_site_macro_untransformed_pairwise_df$year1) # extract length of differences
+backreef_site_macro_untransformed_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", backreef_site_macro_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
+backreef_site_macro_untransformed_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", backreef_site_macro_untransformed_pairwise_df$loc2) # repeat for second quad
+
+# outer10
+
+outer10_site_macro_species_matrix <- as.matrix(outer10_site_macro[ , -which(names(outer10_site_macro) %in% meta_cols)])
+outer10_site_macro_species_matrix <- outer10_site_macro_species_matrix/outer10_site_macro$hard_cover # convert from percent to proportion of hard cover
+rownames(outer10_site_macro_species_matrix) <- paste(outer10_site_macro$year,"-", outer10_site_macro$site_habitat, sep = "")
+
+# filter rare species
+# outer10_site_macro_species_presence <- colSums(outer10_site_macro_species_matrix > 0) # Presence absence matrix
+# outer10_site_macro_species_percentage <- (outer10_site_macro_species_presence/nrow(outer10_site_macro_species_matrix)) # proportion of rows where a species is present
+# outer10_site_macro_common_species <- names(outer10_site_macro_species_percentage[outer10_site_macro_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
+# outer10_site_macro_species_matrix <- outer10_site_macro_species_matrix[,colnames(outer10_site_macro_species_matrix) %in% outer10_site_macro_common_species] # filter original species matrix
+
+outer10_site_macro_rows_to_remove <- rowSums(outer10_site_macro_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
+outer10_site_macro_species_matrix <- outer10_site_macro_species_matrix[!outer10_site_macro_rows_to_remove,]
+
+outer10_site_macro_species_matrix_untransformed <- outer10_site_macro_species_matrix
+outer10_site_macro_species_matrix_transformed <- asin(sqrt(outer10_site_macro_species_matrix)) # arcsin squareroot transform
+outer10_site_macro_subset <- outer10_site_macro[!outer10_site_macro_rows_to_remove,] # make sure rows are also removed from dataframe
+outer10_site_macro_subset$stratum <- paste(outer10_site_macro_subset$year,"-", outer10_site_macro_subset$site_habitat, sep = "") # for merging with bray matrix later
+
+brayouter10_site_macro <- vegdist(outer10_site_macro_species_matrix_transformed, "bray") # create bray dissimilarity matrix
+# View(brayouter10_site_macro)
+brayouter10_site_macro_untransformed <- vegdist(outer10_site_macro_species_matrix_untransformed, "bray")
+outer10_site_macro_braymat <- as.matrix(brayouter10_site_macro) # conver to matrix object
+# convert to pairwise comparison
+outer10_site_macro_comp <- outer10_site_macro_braymat[lower.tri(outer10_site_macro_braymat)] # extract lower triangle
+outer10_site_macro_row_indices <- row(outer10_site_macro_braymat)[lower.tri(outer10_site_macro_braymat)] # extract row indices for first quad in comparison
+outer10_site_macro_col_indices <- col(outer10_site_macro_braymat)[lower.tri(outer10_site_macro_braymat)] # extract col indices = second quad in comp
+outer10_site_macro_loc1 <- rownames(outer10_site_macro_braymat)[outer10_site_macro_row_indices] # populate appropriate vector
+outer10_site_macro_loc2 <- colnames(outer10_site_macro_braymat)[outer10_site_macro_col_indices] # populate appropriate vector
+outer10_site_macro_pairwise_df <- data.frame(loc1 = outer10_site_macro_loc1, loc2 = outer10_site_macro_loc2, dissimilarity = outer10_site_macro_comp)
+outer10_site_macro_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer10_site_macro_pairwise_df$loc1)) # extract timepoint from first quad in comp
+outer10_site_macro_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer10_site_macro_pairwise_df$loc2)) # repeat for second quad
+outer10_site_macro_pairwise_df$time <- abs(outer10_site_macro_pairwise_df$year2 - outer10_site_macro_pairwise_df$year1) # extract length of differences
+outer10_site_macro_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", outer10_site_macro_pairwise_df$loc1) # extract quadrat id from first quad in comp
+outer10_site_macro_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", outer10_site_macro_pairwise_df$loc2) # repeat for second quad
+
+outer10_site_macro_untransformed_braymat <- as.matrix(brayouter10_site_macro_untransformed) # conver to matrix object
+# convert to pairwise comparison
+outer10_site_macro_untransformed_comp <- outer10_site_macro_untransformed_braymat[lower.tri(outer10_site_macro_untransformed_braymat)] # extract lower triangle
+outer10_site_macro_untransformed_row_indices <- row(outer10_site_macro_untransformed_braymat)[lower.tri(outer10_site_macro_untransformed_braymat)] # extract row indices for first quad in comparison
+outer10_site_macro_untransformed_col_indices <- col(outer10_site_macro_untransformed_braymat)[lower.tri(outer10_site_macro_untransformed_braymat)] # extract col indices = second quad in comp
+outer10_site_macro_untransformed_loc1 <- rownames(outer10_site_macro_untransformed_braymat)[outer10_site_macro_untransformed_row_indices] # populate appropriate vector
+outer10_site_macro_untransformed_loc2 <- colnames(outer10_site_macro_untransformed_braymat)[outer10_site_macro_untransformed_col_indices] # populate appropriate vector
+outer10_site_macro_untransformed_pairwise_df <- data.frame(loc1 = outer10_site_macro_untransformed_loc1, loc2 = outer10_site_macro_untransformed_loc2, dissimilarity = outer10_site_macro_untransformed_comp)
+outer10_site_macro_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer10_site_macro_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
+outer10_site_macro_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer10_site_macro_untransformed_pairwise_df$loc2)) # repeat for second quad
+outer10_site_macro_untransformed_pairwise_df$time <- abs(outer10_site_macro_untransformed_pairwise_df$year2 - outer10_site_macro_untransformed_pairwise_df$year1) # extract length of differences
+outer10_site_macro_untransformed_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", outer10_site_macro_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
+outer10_site_macro_untransformed_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", outer10_site_macro_untransformed_pairwise_df$loc2) # repeat for second quad
+
+# outer 17
+
+outer17_site_macro_species_matrix <- as.matrix(outer17_site_macro[ , -which(names(outer17_site_macro) %in% meta_cols)])
+outer17_site_macro_species_matrix <- outer17_site_macro_species_matrix/outer17_site_macro$hard_cover # convert from percent to proportion of hard cover
+rownames(outer17_site_macro_species_matrix) <- paste(outer17_site_macro$year,"-", outer17_site_macro$site_habitat, sep = "")
+
+# filter rare species
+# outer17_site_macro_species_presence <- colSums(outer17_site_macro_species_matrix > 0) # Presence absence matrix
+# outer17_site_macro_species_percentage <- (outer17_site_macro_species_presence/nrow(outer17_site_macro_species_matrix)) # proportion of rows where a species is present
+# outer17_site_macro_common_species <- names(outer17_site_macro_species_percentage[outer17_site_macro_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
+# outer17_site_macro_species_matrix <- outer17_site_macro_species_matrix[,colnames(outer17_site_macro_species_matrix) %in% outer17_site_macro_common_species] # filter original species matrix
+
+outer17_site_macro_rows_to_remove <- rowSums(outer17_site_macro_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
+outer17_site_macro_species_matrix <- outer17_site_macro_species_matrix[!outer17_site_macro_rows_to_remove,]
+
+outer17_site_macro_species_matrix_untransformed <- outer17_site_macro_species_matrix
+outer17_site_macro_species_matrix_transformed <- asin(sqrt(outer17_site_macro_species_matrix)) # arcsin squareroot transform
+outer17_site_macro_subset <- outer17_site_macro[!outer17_site_macro_rows_to_remove,] # make sure rows are also removed from dataframe
+outer17_site_macro_subset$stratum <- paste(outer17_site_macro_subset$year,"-", outer17_site_macro_subset$site_habitat, sep = "") # for merging with bray matrix later
+
+brayouter17_site_macro <- vegdist(outer17_site_macro_species_matrix_transformed, "bray") # create bray dissimilarity matrix
+# View(brayouter17_site_macro)
+brayouter17_site_macro_untransformed <- vegdist(outer17_site_macro_species_matrix_untransformed, "bray")
+outer17_site_macro_braymat <- as.matrix(brayouter17_site_macro) # conver to matrix object
+# convert to pairwise comparison
+outer17_site_macro_comp <- outer17_site_macro_braymat[lower.tri(outer17_site_macro_braymat)] # extract lower triangle
+outer17_site_macro_row_indices <- row(outer17_site_macro_braymat)[lower.tri(outer17_site_macro_braymat)] # extract row indices for first quad in comparison
+outer17_site_macro_col_indices <- col(outer17_site_macro_braymat)[lower.tri(outer17_site_macro_braymat)] # extract col indices = second quad in comp
+outer17_site_macro_loc1 <- rownames(outer17_site_macro_braymat)[outer17_site_macro_row_indices] # populate appropriate vector
+outer17_site_macro_loc2 <- colnames(outer17_site_macro_braymat)[outer17_site_macro_col_indices] # populate appropriate vector
+outer17_site_macro_pairwise_df <- data.frame(loc1 = outer17_site_macro_loc1, loc2 = outer17_site_macro_loc2, dissimilarity = outer17_site_macro_comp)
+outer17_site_macro_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer17_site_macro_pairwise_df$loc1)) # extract timepoint from first quad in comp
+outer17_site_macro_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer17_site_macro_pairwise_df$loc2)) # repeat for second quad
+outer17_site_macro_pairwise_df$time <- abs(outer17_site_macro_pairwise_df$year2 - outer17_site_macro_pairwise_df$year1) # extract length of differences
+outer17_site_macro_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", outer17_site_macro_pairwise_df$loc1) # extract quadrat id from first quad in comp
+outer17_site_macro_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", outer17_site_macro_pairwise_df$loc2) # repeat for second quad
+
+outer17_site_macro_untransformed_braymat <- as.matrix(brayouter17_site_macro_untransformed) # conver to matrix object
+# convert to pairwise comparison
+outer17_site_macro_untransformed_comp <- outer17_site_macro_untransformed_braymat[lower.tri(outer17_site_macro_untransformed_braymat)] # extract lower triangle
+outer17_site_macro_untransformed_row_indices <- row(outer17_site_macro_untransformed_braymat)[lower.tri(outer17_site_macro_untransformed_braymat)] # extract row indices for first quad in comparison
+outer17_site_macro_untransformed_col_indices <- col(outer17_site_macro_untransformed_braymat)[lower.tri(outer17_site_macro_untransformed_braymat)] # extract col indices = second quad in comp
+outer17_site_macro_untransformed_loc1 <- rownames(outer17_site_macro_untransformed_braymat)[outer17_site_macro_untransformed_row_indices] # populate appropriate vector
+outer17_site_macro_untransformed_loc2 <- colnames(outer17_site_macro_untransformed_braymat)[outer17_site_macro_untransformed_col_indices] # populate appropriate vector
+outer17_site_macro_untransformed_pairwise_df <- data.frame(loc1 = outer17_site_macro_untransformed_loc1, loc2 = outer17_site_macro_untransformed_loc2, dissimilarity = outer17_site_macro_untransformed_comp)
+outer17_site_macro_untransformed_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer17_site_macro_untransformed_pairwise_df$loc1)) # extract timepoint from first quad in comp
+outer17_site_macro_untransformed_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer17_site_macro_untransformed_pairwise_df$loc2)) # repeat for second quad
+outer17_site_macro_untransformed_pairwise_df$time <- abs(outer17_site_macro_untransformed_pairwise_df$year2 - outer17_site_macro_untransformed_pairwise_df$year1) # extract length of differences
+outer17_site_macro_untransformed_pairwise_df$site_1 <- sub(".*-(.*)", "\\1", outer17_site_macro_untransformed_pairwise_df$loc1) # extract quadrat id from first quad in comp
+outer17_site_macro_untransformed_pairwise_df$site_2 <- sub(".*-(.*)", "\\1", outer17_site_macro_untransformed_pairwise_df$loc2) # repeat for second quad
+
+```
+
+```{r site macroalgae bray dissimilarity over time, fig.height=15, fig.width=15}
+# for the purposes of this first set of analyses, I'm interested in dissimilarity of a given quad over time
+# not necessarily interested in how quads compare to each other over time/at each timepoint
+fringing_site_macro_pairwise_subset <- fringing_site_macro_pairwise_df[fringing_site_macro_pairwise_df$site_1 == fringing_site_macro_pairwise_df$site_2,]
+backreef_site_macro_pairwise_subset <- backreef_site_macro_pairwise_df[backreef_site_macro_pairwise_df$site_1 == backreef_site_macro_pairwise_df$site_2,]
+outer10_site_macro_pairwise_subset <- outer10_site_macro_pairwise_df[outer10_site_macro_pairwise_df$site_1 == outer10_site_macro_pairwise_df$site_2,]
+outer17_site_macro_pairwise_subset <- outer17_site_macro_pairwise_df[outer17_site_macro_pairwise_df$site_1 == outer17_site_macro_pairwise_df$site_2,]
+
+comb_site_macro_subset <- rbind(fringing_site_macro_pairwise_subset,
+                     backreef_site_macro_pairwise_subset,
+                     outer10_site_macro_pairwise_subset,
+                     outer17_site_macro_pairwise_subset)
+# site_1 and site_2 columns are all the same now
+# create new col just called for merge with other metadata
+comb_site_macro_subset$site_habitat <- comb_site_macro_subset$site_1
+
+bray_time_series_site_macro <- left_join(unique(site_meta_df[ , c("site", "site_habitat", "habitat", "island_side")]), comb_site_macro_subset, by = "site_habitat")
+
+# model
+# transformed dissimilarities
+bray_time_series_site_macro$dissimilarity_trans <- sv_trans(bray_time_series_site_macro$dissimilarity, bray_time_series_site_macro)
+
+bray_mod_site_macro <- glmmTMB(dissimilarity_trans ~ time*habitat + (1 | site),
+                              na.action = na.pass, family = beta_family(link = "logit"), data = bray_time_series_site_macro) #
+summary(bray_mod_site_macro) 
+hist(residuals(bray_mod_site_macro)) # slight skew but not awful
+plot(residuals(bray_mod_site_macro) ~ fitted(bray_mod_site_macro)) # stronger patterning here - negative slope
+
+emtrends(bray_mod_site_macro, pairwise ~ habitat , "time") # fringing slope significantly different from all others
+bray_trends_emm_site_macro <- emmip(bray_mod_site_macro, habitat ~ time , at = list(time = seq(1,16,1)), plotit = F, CIs = T)
+
+ggplot()+
+   geom_point(data = bray_time_series_site_macro, aes(x = time, y = dissimilarity, colour = habitat), alpha = 0.1 ) +
+   geom_line(data = bray_trends_emm_site_macro, aes(x = time, y = inv_logit(yvar), colour = habitat), linewidth = 2) +
+   geom_ribbon(data = bray_trends_emm_site_macro, aes(x = time, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+   scale_fill_manual(values = habitat_colours) +
+
+   labs(y = "Dissimilarity\n") +
+   theme_bw() + 
+   theme(panel.grid.major = element_blank(), 
+         panel.grid.minor = element_blank(),
+         panel.background = element_blank(),
+         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+         axis.line.y = element_line(colour = "black"),
+         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         axis.line.x = element_line(colour = "black"),
+         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         strip.text.x = element_text(size = 32),
+         legend.key.size = unit(2, "cm"), # increase legend size
+         legend.text = element_text(size = rel(2)), # increase legend text
+         legend.title = element_text(size = rel(2.25)),
+         legend.position = 'bottom', # move legend to bottom
+         legend.box = 'horizontal',
+         legend.box.background = element_rect(colour = "black") )# make legend horizontal
+```
+
+```{r dissimilarity compared to baseline macro, fig.height=15, fig.width = 15}
+bray_time_series_site_macro_baseline <- subset(bray_time_series_site_macro, year2 == 2007)
+bray_mod_site_macro_baseline <- glmmTMB(dissimilarity_trans ~ time*habitat + (1 | site),
+                              na.action = na.pass, family = beta_family(link = "logit"), data = bray_time_series_site_macro_baseline)
+hist(residuals(bray_mod_site_macro_baseline)) # slight negative skew
+plot(residuals(bray_mod_site_macro_baseline) ~ fitted(bray_mod_site_macro_baseline) ) # strong negative trend
+
+emtrends(bray_mod_site_macro_baseline, pairwise ~ habitat , "time") # fringing slope significantly different from all others
+bray_trends_emm_site_macro_baseline <- emmip(bray_mod_site_macro_baseline, habitat ~ time , at = list(time = seq(1,16,1)), plotit = F, CIs = T)
+
+ggplot()+
+   geom_point(data = bray_time_series_site_macro_baseline, aes(x = time, y = dissimilarity, colour = habitat), alpha = 0.1 ) +
+   geom_line(data = bray_trends_emm_site_macro_baseline, aes(x = time, y = inv_logit(yvar), colour = habitat), linewidth = 2) +
+   geom_ribbon(data = bray_trends_emm_site_macro_baseline, aes(x = time, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+   scale_fill_manual(values = habitat_colours) +
+
+   labs(y = "Dissimilarity\n") +
+   theme_bw() + 
+   theme(panel.grid.major = element_blank(), 
+         panel.grid.minor = element_blank(),
+         panel.background = element_blank(),
+         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+         axis.line.y = element_line(colour = "black"),
+         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         axis.line.x = element_line(colour = "black"),
+         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         strip.text.x = element_text(size = 32),
+         legend.key.size = unit(2, "cm"), # increase legend size
+         legend.text = element_text(size = rel(2)), # increase legend text
+         legend.title = element_text(size = rel(2.25)),
+         legend.position = 'bottom', # move legend to bottom
+         legend.box = 'horizontal',
+         legend.box.background = element_rect(colour = "black") )# make legend horizontal
+```
+
 ```{r site level ordination all algae, fig.height=15, fig.width=15}
@@
 full_site_species_matrix <- asin(sqrt(full_site_species_matrix)) # arcsin squareroot transform
-site_full$stratum <- paste(site_full$year,"-", site_full$location, sep = "") # for merging with bray matrix later
+site_full$stratum <- paste(site_full$year,"-", site_full$site_habitat, sep = "") # for merging with bray matrix later
 full_subset <- site_full[site_full$stratum %in% relevant_sites,] # make sure rows are also removed from dataframe
@@
 
-full_site_macro_species_matrix <- as.matrix(site_macro[ , -which(names(site_macro) %in% meta_cols)])
+full_site_macro_species_matrix <- as.matrix(site_macro[ , -which(names(site_macro) %in% colnames(site_meta_df))])
 full_site_macro_species_matrix <- full_site_macro_species_matrix/site_macro$hard_cover # convert from percent to proportion of hard cover
@@
 full_site_macro_species_matrix <- asin(sqrt(full_site_macro_species_matrix)) # arcsin squareroot transform
-site_macro$stratum <- paste(site_macro$year,"-", site_macro$location, sep = "") # for merging with bray matrix later
+site_macro$stratum <- paste(site_macro$year,"-", site_macro$site_habitat, sep = "") # for merging with bray matrix later
 full_subset_site_macro <- site_macro[site_macro$stratum %in% relevant_site_macros,] # make sure rows are also removed from dataframe
@@
 # read in disturbances
-disturbances <- read.csv(here("data", "disturbance_list.csv"))
 full.cent_site_macro <- merge(full.cent_site_macro, disturbances, by = "year") # add disturbances to path
@@
 ggplot() + 
-  geom_point(data= full.site.scores_site_macro, aes(x=NMDS1,y=NMDS2, colour = habitat), alpha = 0.05) + # add the point markers
-  geom_text_repel(data= full.abundant.species.scores_site_macro,aes(x=NMDS1,y=NMDS2, label = species), size = 4, alpha = 0.5) +  # add the species labels
+  geom_point(data= full.site.scores_site_macro, aes(x=NMDS1,y=NMDS2, colour = habitat), alpha = 0.2) + # add the point markers
+  geom_text_repel(data= full.abundant.species.scores_site_macro,aes(x=NMDS1,y=NMDS2, label = species), size = 6, alpha = 0.7) +  # add the species labels
 
   geom_path(data = full.cent_site_macro, aes(x = NMDS1, y = NMDS2, group = habitat, colour = habitat),
-            arrow = arrow(length=unit(0.25,"cm"), ends="last", type = "closed"), size = 1) +
-  geom_point(data = full.cent_site_macro, aes(x = NMDS1, y = NMDS2, group = habitat, colour = habitat, shape = disturbance), size = 3, stroke = 1.4) +
+            arrow = arrow(length=unit(0.4,"cm"), ends="last", type = "closed"), size = 1) +
+  geom_point(data = full.cent_site_macro, aes(x = NMDS1, y = NMDS2, group = habitat, colour = habitat, shape = disturbance), size = 5, stroke = 1.4) +
   scale_shape_manual(values = c(
@@
          )
+
+```
+
+```{r full site permanova}
+full_site_bray <- vegdist(full_site_species_matrix, "bray") 
+full_site_macro_bray <- vegdist(full_site_macro_species_matrix, "bray")
+
+# adonis2(full_site_bray ~ year*habitat, data = site_full)
+#               Df SumOfSqs      R2       F Pr(>F)    
+# year           1    1.293 0.03257 18.4422  0.001 ***
+# habitat        3    9.952 0.25065 47.3113  0.001 ***
+# year:habitat   3    1.255 0.03161  5.9659  0.001 ***
+# Residual     388   27.204 0.68518                   
+# Total        395   39.704 1.00000                   
+# habitat accounts for most of the variation in algal communities
+
+# adonis2(full_site_macro_bray ~ year*habitat, data = full_subset_site_macro)
+#               Df SumOfSqs      R2       F Pr(>F)    
+# year           1    6.209 0.04528 24.8046  0.001 ***
+# habitat        3   28.410 0.20719 37.8323  0.001 ***
+# year:habitat   3    5.882 0.04290  7.8332  0.001 ***
+# Residual     386   96.622 0.70464                   
+# Total        393  137.124 1.00000               
+# same when just looking at macroalgal communities, but still greater residual variance
+
+```
+
+```{r beta dispersion all algae, fig.height=15, fig.width=15}
+full_site_betadisper <- vegan::betadisper(full_site_bray, full.site.scores_site$hab_year)
+site_full$betadispersion <- full_site_betadisper$distances[match(site_full$stratum, names(full_site_betadisper$distances))]
+site_full$betadispersion_trans <- sv_trans(site_full$betadispersion, site_full)
+
+betadisper_site_mod <- glmmTMB(betadispersion_trans ~ year*habitat + (1|site), family = beta_family(link = "logit"), data = site_full)
+summary(betadisper_site_mod) # fair bit of within site variance - inv_logit(0.04) = 0.5
+hist(residuals(betadisper_site_mod)) # pretty good
+plot(residuals(betadisper_site_mod) ~ fitted(betadisper_site_mod)) # slight heteroskedasticity
+Anova(betadisper_site_mod)
+#                Chisq Df Pr(>Chisq)    
+# year         12.9561  1  0.0003189 ***
+# habitat      27.2191  3  5.296e-06 ***
+# year:habitat  6.8472  3  0.0769311 .
+
+emtrends(betadisper_site_mod, pairwise ~ habitat , "year") # fringing slope significantly different from outer17
+betadisper_site_emm <- emmip(betadisper_site_mod, habitat ~ year , at = list(year = seq(2007,2023,1)), plotit = F, CIs = T)
+
+ggplot()+
+   geom_point(data = site_full, aes(x = year, y = betadispersion_trans, colour = habitat), alpha = 0.1 ) +
+   geom_line(data = betadisper_site_emm, aes(x = year, y = inv_logit(yvar), colour = habitat), linewidth = 2) +
+   geom_ribbon(data = betadisper_site_emm, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+   scale_fill_manual(values = habitat_colours) +
+
+   labs(y = "Beta dispersion\n", x = "Year") +
+   theme_bw() + 
+   theme(panel.grid.major = element_blank(), 
+         panel.grid.minor = element_blank(),
+         panel.background = element_blank(),
+         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+         axis.line.y = element_line(colour = "black"),
+         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         axis.line.x = element_line(colour = "black"),
+         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         strip.text.x = element_text(size = 32),
+         legend.key.size = unit(2, "cm"), # increase legend size
+         legend.text = element_text(size = rel(2)), # increase legend text
+         legend.title = element_text(size = rel(2.25)),
+         legend.position = 'bottom', # move legend to bottom
+         legend.box = 'horizontal',
+         legend.box.background = element_rect(colour = "black") )# make legend horizontal
+
+```
+
+```{r beta dispersion macoralgae, fig.height=15, fig.width=15}
+full_site_macro_betadisper <- vegan::betadisper(full_site_macro_bray, full.site.scores_site_macro$hab_year)
+full_subset_site_macro$betadispersion <- full_site_macro_betadisper$distances[match(full_subset_site_macro$stratum, names(full_site_macro_betadisper$distances))]
+full_subset_site_macro$betadispersion_trans <- sv_trans(full_subset_site_macro$betadispersion, full_subset_site_macro)
 

### 2024-09-04 noam_final_analyses.Rmd

@@
 
+data[,full_species_cols] <- data[,full_species_cols]/data$hard_cover # transform cover to proportion of hardcover in quadrat
+
+macroalgae[,macro_species_cols] <- macroalgae[,macro_species_cols]/macroalgae$hard_cover # repeat for macro dataframe
+
 # summarize at site level
@@
 
-```{r read and merge secondary datasets}
-wave <- read.csv(here::here("data", "algae_wave_dynamics.csv"), stringsAsFactors = F)
-nutrients <- read.csv(here::here("data", "lter_nuts.csv"), stringsAsFactors = F)
-temp <- read.csv(here::here("data", "temp_sum.csv"), stringsAsFactors = F)
-
-```
-
-```{r all algae alpha diversity quadrat level}
+```{r create adiv dataframes}
 alpha_diversity_quad <-
@@
 alpha_diversity_quad <- merge(meta_df, alpha_diversity_quad, by = c("location", "year") )
+
+alpha_diversity_quad_macro <-
+  data.frame(
+    year = macroalgae$year,
+    location = macroalgae$location,
+    shannon = diversity(macroalgae[ , -which(names(macroalgae) %in% meta_cols)], "shannon"),
+    richness = rowSums(macroalgae[ , -which(names(macroalgae) %in% meta_cols)] > 0)
+  )
+
+alpha_diversity_quad_macro <- merge(meta_df, alpha_diversity_quad_macro, by = c("location", "year"))
+
+alpha_diversity_site <-
+  data.frame(
+    year = site_full$year,
+    site_hab_year = site_full$site_hab_year,
+    shannon = diversity(site_full[ , -which(names(site_full) %in% meta_cols)], "shannon"),
+    richness = rowSums(site_full[ , -which(names(site_full) %in% meta_cols)] > 0)
+  )
+
+alpha_diversity_site <- merge(site_meta_df, alpha_diversity_site, by= c("year", "site_hab_year"))
+
+alpha_diversity_site_macro <-
+  data.frame(
+    year = site_macro$year,
+    site_hab_year = site_macro$site_hab_year,
+    shannon = diversity(site_macro[ , -which(names(site_macro) %in% meta_cols)], "shannon"),
+    richness = rowSums(site_macro[ , -which(names(site_macro) %in% meta_cols)] > 0)
+  )
+
+alpha_diversity_site_macro <- merge(site_meta_df, alpha_diversity_site_macro, by= c("year", "site_hab_year"))
+
+```
+
+```{r read and merge secondary datasets, fig.height=15, fig.width=15}
+wave <- read.csv(here::here("data", "algae_wave_dynamics.csv"), stringsAsFactors = F)
+wave <- wave[,-which(names(wave) %in% "X")] # remove rownum columns
+wave$site_hab_year <- paste0(wave$site,"_", wave$habitat,"_", wave$year) # for merge
+
+nutrients <- read.csv(here::here("data", "lter_nuts.csv"), stringsAsFactors = F)
+nutrients$site_hab_year <- paste0(nutrients$site,"_", nutrients$habitat,"_", nutrients$year)
+
+temp <- read.csv(here::here("data", "temp_sum.csv"), stringsAsFactors = F)
+temp$site_hab_year <- paste0(temp$site,"_", temp$habitat,"_", temp$year)
+
+herb <- read.csv(here::here("data", "herbivore_biomass_and_diversity_site.csv"), stringsAsFactors = F)
+# create column for total herbivore biomass
+herb$herbivore_biomass <- rowSums(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))])
+herb<-herb %>%  # rename columns for merge
+  rename(
+    year = Year,
+    site = Site,
+    habitat = Habitat,
+    herbivore_richness = richness,
+    herbivore_shannon = shannon,
+    herbivore_evenness = evenness
+    )
+# fix site names and habitat names
+herb <-
+  herb %>% mutate(site = case_when(site == 1 ~ "lter_1",
+                              site == 2 ~ "lter_2",
+                              site == 3 ~ "lter_3",
+                              site == 4 ~ "lter_4",
+                              site == 5 ~ "lter_5",
+                              site == 6 ~ "lter_6"))
+herb <-
+  herb %>% mutate(habitat = case_when(habitat == "Fringing" ~ "fringing",
+                                      habitat == "Backreef" ~ "backreef",
+                                      habitat == "Outer 10" ~ "outer_10")) # there is no outer 17 for herb data
+herb$site_hab_year <- paste0(herb$site,"_", herb$habitat,"_", herb$year)
+herb <- subset(herb, year != 2006)
+
+### Merge with alpha diversity datasets
+
+alpha_diversity_site <- full_join(alpha_diversity_site, nutrients, by = c("year", "site_hab_year", "site", "habitat"))
+alpha_diversity_site <- full_join(alpha_diversity_site, temp, by = c("year", "site_hab_year", "site", "habitat"))
+
+alpha_diversity_site <- full_join(alpha_diversity_site,herb[,-which(names(herb) %in% colnames(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))]) )], by = c("year", "site_hab_year", "site", "habitat"))
+
+alpha_diversity_site <- full_join(alpha_diversity_site, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
+pairs(data = alpha_diversity_site, ~ all_algae_cover_hard + macroalgae_cover_hard + richness + shannon + coral_cover_hard + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)
+
+alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro, nutrients, by = c("year", "site_hab_year", "site", "habitat"))
+alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro, temp, by = c("year", "site_hab_year", "site", "habitat"))
+
+alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro,herb[,-which(names(herb) %in% colnames(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))]) )], by = c("year", "site_hab_year", "site", "habitat"))
+
+alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
+pairs(data = alpha_diversity_site_macro, ~ richness + shannon + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)
+
+alpha_diversity_quad <- full_join(alpha_diversity_quad, nutrients, by = c("year", "site_hab_year", "site", "habitat"))
+alpha_diversity_quad <- full_join(alpha_diversity_quad, temp, by = c("year", "site_hab_year", "site", "habitat"))
+
+alpha_diversity_quad <- full_join(alpha_diversity_quad,herb[,-which(names(herb) %in% colnames(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))]) )], by = c("year", "site_hab_year", "site", "habitat"))
+
+alpha_diversity_quad <- full_join(alpha_diversity_quad, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
+pairs(data = alpha_diversity_quad, ~ richness + shannon + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)
+
+alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro, nutrients, by = c("year", "site_hab_year", "site", "habitat"))
+alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro, temp, by = c("year", "site_hab_year", "site", "habitat"))
+
+alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro,herb[,-which(names(herb) %in% colnames(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))]) )], by = c("year", "site_hab_year", "site", "habitat"))
+
+alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
+pairs(data = alpha_diversity_quad_macro, ~ richness + shannon + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)
+
+```
+
+```{r all algae alpha diversity site level}
+
+site_full_richness_mod <- glmmTMB(richness ~ year*habitat + (1|site) + offset(log(hard_cover)), family = poisson,
+                             data = alpha_diversity_site) # need to include offset because of the way richness is calculated
+# Warning in (function (start, objective, gradient = NULL, hessian = NULL,  : NA/NaN function evaluation
+# This warning occurs when the optimizer visits a region of parameter space that is invalid.
+It is not a problem as long as the optimizer has left that region of parameter space upon convergence, which is indicated by an absence of the model convergence warnings described above.
+summary(site_full_richness_mod) # low variance in random effect
+Anova(site_full_richness_mod) 
+#                 Chisq Df Pr(>Chisq)    
+# year           1.5505  1     0.2131    
+# habitat      106.8578  3     <2e-16 ***
+# year:habitat   5.3674  3     0.1468    
+# habitat more important than time when time is numeric
+# check_mod
+hist(residuals(site_full_richness_mod)) # slight skew
+
+# striping/patterining expected because of discrete response var.
+plot(residuals(site_full_richness_mod) ~ fitted(site_full_richness_mod)) # slightly heteroskedastic
+
+emtrends(site_full_richness_mod, pairwise ~  habitat, var = "year") # habitat trends don't differ from each other
+site_full_richness_emm <- emmip(site_full_richness_mod,habitat ~ year,
+                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)
+
+(site_base_richness_plot <- 
+  ggplot() +
+  geom_point(data = alpha_diversity_site,
+             aes(x = year, y = richness, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = site_full_richness_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = site_full_richness_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
+  geom_segment(aes(y = 18, yend = 18, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  scale_y_log10() +
+    labs(x = "Year", y = "Richness") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) ) # add box around legend
+
+site_full_shannon_mod <- glmmTMB(log(shannon + 1) ~ year*habitat + (1|site), 
+                             data = alpha_diversity_site) 
+# sites with shannon div = 0 so log+1 transformed. No offset because calculation of shannon already incorporated differences in %hard
+summary(site_full_shannon_mod) # low variance again
+Anova(site_full_shannon_mod)
+# Response: log(shannon + 1)
+#                Chisq Df Pr(>Chisq)    
+# year          0.0133  1   0.908348    
+# habitat      97.6274  3  < 2.2e-16 ***
+# year:habitat 14.0548  3   0.002831 **
+
+hist(residuals(site_full_shannon_mod)) # looks good
+plot(residuals(site_full_shannon_mod) ~ fitted(site_full_shannon_mod)) # looks good here too
+
+emtrends(site_full_shannon_mod, pairwise ~  habitat, var = "year") # backreef and fringing differ, backreef different fro outer17
+site_full_shannon_emm <- emmip(site_full_shannon_mod,habitat ~ year,
+                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)
+
+(site_base_shannon_plot <- 
+  ggplot() +
+  geom_point(data = alpha_diversity_site,
+             aes(x = year, y = shannon, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = site_full_shannon_emm, aes(x = year, y = yvar, colour = habitat), size = 2) +
+  geom_ribbon(data = site_full_shannon_emm, aes(x = year, ymin = LCL, ymax = UCL, fill = habitat), alpha = 0.3) +
+  geom_segment(aes(y = 3, yend = 3, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  scale_y_log10() +
+    labs(x = "Year", y = "Shannon Index (H')") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) ) # add box around legend
+
+```
+
+```{r macroalgae alpha diversity site level}
+
+site_macro_richness_mod <- glmmTMB(richness ~ year*habitat + (1|site) + offset(log(hard_cover)), family = poisson,
+                             data = alpha_diversity_site_macro) 
+# Warning in (function (start, objective, gradient = NULL, hessian = NULL,  : NA/NaN function evaluation
+summary(site_macro_richness_mod) # low variance again
+Anova(site_macro_richness_mod)
+#                Chisq Df Pr(>Chisq)    
+# year          1.1856  1    0.27621    
+# habitat      87.7765  3    < 2e-16 ***
+# year:habitat  7.0855  3    0.06922 .  
+hist(residuals(site_macro_richness_mod)) # slightly more skew
+plot(residuals(site_macro_richness_mod) ~ fitted(site_macro_richness_mod)) # similar patterns
+
+emtrends(site_macro_richness_mod, pairwise ~  habitat, var = "year") # habitat trends don't differ from each other
+site_macro_richness_emm <- emmip(site_macro_richness_mod,habitat ~ year,
+                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)
+
+(site_base_richness_plot <- 
+  ggplot() +
+  geom_point(data = alpha_diversity_site_macro,
+             aes(x = year, y = richness, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = site_macro_richness_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = site_macro_richness_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  geom_segment(aes(y = 18, yend = 18, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_y_log10() +
+    labs(x = "Year", y = "Richness") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) ) # add box around legend
+
+site_macro_shannon_mod <- glmmTMB(log(shannon +1 ) ~ year*habitat + (1|site),
+                             data = alpha_diversity_site) 
+# sites with shannon div = 0 so log+1 transformed
+summary(site_macro_shannon_mod)
+Anova(site_macro_shannon_mod)
+#                Chisq Df Pr(>Chisq)    
+# year          0.0133  1   0.908348    
+# habitat      97.6274  3  < 2.2e-16 ***
+# year:habitat 14.0548  3   0.002831 ** 
+hist(residuals(site_macro_shannon_mod)) # looks good
+plot(residuals(site_macro_shannon_mod) ~ fitted(site_macro_shannon_mod)) # generally fine
+
+emtrends(site_macro_shannon_mod, pairwise ~  habitat, var = "year") # backreef and fringing differ
+site_macro_shannon_emm <- emmip(site_macro_shannon_mod,habitat ~ year,
+                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)
+
+(site_base_shannon_plot <- 
+  ggplot() +
+ geom_point(data = alpha_diversity_site,
+            aes(x = year, y = shannon, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = site_macro_shannon_emm, aes(x = year, y = yvar, colour = habitat), size = 2) +
+  geom_ribbon(data = site_macro_shannon_emm, aes(x = year, ymin = LCL, ymax = UCL, fill = habitat), alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  geom_segment(aes(y = 2.5, yend = 2.5, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_fill_manual(values = habitat_colours) +
+  scale_y_log10() +
+    labs(x = "Year", y = "Shannon Index (H')") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) ) # add box around legend
+
+```
+```{r stepwise models site level}
+
+## FRINGING - exclude wave power
+fringing_all_site <- na.omit(subset(alpha_diversity_site, habitat == "fringing")) # only use rows that contain all factors
+
+fringing_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + mean_n + mean_temp_c +  herbivore_evenness + offset(log(hard_cover)) + (1|site), family = poisson, data = fringing_all_site, na.action = na.pass) # universal model has convergence issues
+fringing_all_richness_dredge <- dredge(fringing_all_richness_mod_universal) # most models have convergence issues
+model.sel(fringing_all_richness_dredge) # top model just includes year
+fringing_all_richness_top_mod <- glmmTMB(richness ~ year + offset(log(hard_cover)) + (1|site), family = poisson, data = fringing_all_site, na.action = na.pass)
+summary(fringing_all_richness_top_mod)
+hist(residuals(fringing_all_richness_top_mod)) # slight skew but not bad
+plot(residuals(fringing_all_richness_top_mod) ~ fitted(fringing_all_richness_top_mod)) # looks fine
+
+fringing_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = fringing_all_site, na.action = na.pass)
+fringing_all_shannon_dredge <- dredge(fringing_all_shannon_mod_universal) 
+model.sel(fringing_all_shannon_dredge) # tie for the top model - best includes year and coral cover or just nitrogen
+fringing_all_shannon_top_mod1 <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + (1|site), data = fringing_all_site, na.action = na.pass)
+fringing_all_shannon_top_mod2 <- glmmTMB(log(shannon + 1) ~ mean_n + mean_temp_c + (1|site), data = fringing_all_site, na.action = na.pass)
+hist(residuals(fringing_all_shannon_top_mod1)) # slightly bimodal
+hist(residuals(fringing_all_shannon_top_mod2)) # also uniform - neither are great
+plot(residuals(fringing_all_shannon_top_mod1) ~ fitted(fringing_all_shannon_top_mod1)) # slight patterning
+plot(residuals(fringing_all_shannon_top_mod2) ~ fitted(fringing_all_shannon_top_mod2)) # similar here
+
+## Backreef - includes all 
+backreef_all_site <- na.omit(subset(alpha_diversity_site, habitat == "backreef")) # only use rows that contain all factors
+
+backreef_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + offset(log(hard_cover)) + (1|site), family = poisson, data = backreef_all_site, na.action = na.pass) # universal model has convergence issues
+backreef_all_richness_dredge <- dredge(backreef_all_richness_mod_universal) # most models have convergence issues
+model.sel(backreef_all_richness_dredge) # top model includes year and coral 
+backreef_all_richness_top_mod <- glmmTMB(richness ~ year + coral_cover_hard + offset(log(hard_cover)) + (1|site), family = poisson, data = backreef_all_site, na.action = na.pass)
+summary(backreef_all_richness_top_mod)
+hist(residuals(backreef_all_richness_top_mod)) # slight skew but not bad
+plot(residuals(backreef_all_richness_top_mod) ~ fitted(backreef_all_richness_top_mod)) # looks fine
+
+backreef_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = backreef_all_site, na.action = na.pass)
+backreef_all_shannon_dredge <- dredge(backreef_all_shannon_mod_universal) 
+model.sel(backreef_all_shannon_dredge) #  best includes just coral cover 
+backreef_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  coral_cover_hard + (1|site), data = backreef_all_site, na.action = na.pass)
+hist(residuals(backreef_all_shannon_top_mod)) # pretty good
+plot(residuals(backreef_all_shannon_top_mod) ~ fitted(backreef_all_shannon_top_mod)) # looks fine
+
+## outer10 - no nitrogen 
+outer10_all_site <- subset(alpha_diversity_site, habitat == "outer_10") # only use rows that contain all factors
+# remove nitrogent columns
+outer10_all_site <- outer10_all_site[,-which(names(outer10_all_site) %in% c("mean_n", "n_cv"))]
+# remove other na rows
+outer10_all_site <- na.omit(outer10_all_site)
+
+outer10_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c +  herbivore_evenness + offset(log(hard_cover)) + (1|site), family = poisson, data = outer10_all_site, na.action = na.pass) # universal model has convergence issues
+outer10_all_richness_dredge <- dredge(outer10_all_richness_mod_universal) # most models have convergence issues
+model.sel(outer10_all_richness_dredge) # top model includes just intercept... next best is coral 
+outer10_all_richness_top_mod <- glmmTMB(richness ~ coral_cover_hard + offset(log(hard_cover)) + (1|site), family = poisson, data = outer10_all_site, na.action = na.pass)
+summary(outer10_all_richness_top_mod)
+hist(residuals(outer10_all_richness_top_mod)) #  not bad
+plot(residuals(outer10_all_richness_top_mod) ~ fitted(outer10_all_richness_top_mod)) # looks fine
+
+outer10_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c +  herbivore_evenness + (1|site), data = outer10_all_site, na.action = na.pass)
+outer10_all_shannon_dredge <- dredge(outer10_all_shannon_mod_universal) # some convergence issues
+model.sel(outer10_all_shannon_dredge) #  best includes just herbivore evenness - first time there's been a decent gap in AICc
+outer10_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  herbivore_evenness + (1|site), data = outer10_all_site, na.action = na.pass)
+hist(residuals(outer10_all_shannon_top_mod)) # pretty good
+plot(residuals(outer10_all_shannon_top_mod) ~ fitted(outer10_all_shannon_top_mod)) # looks fine
+
+## outer17 - no nitrogen or herbivores
+outer17_all_site <- subset(alpha_diversity_site, habitat == "outer_17") # only use rows that contain all factors
+# remove nitrogent columns
+outer17_all_site <- outer17_all_site[,-which(names(outer17_all_site) %in% c("mean_n", "n_cv", "herbivore_biomass", "herbivore_evenness", "herbivore_richness", "herbivore_shannon"))]
+# remove other na rows
+outer17_all_site <- na.omit(outer17_all_site)
+
+outer17_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c +  offset(log(hard_cover)) + (1|site), family = poisson, data = outer17_all_site, na.action = na.pass) # universal model has convergence issues
+outer17_all_richness_dredge <- dredge(outer17_all_richness_mod_universal) # most models have convergence issues
+model.sel(outer17_all_richness_dredge) # top model includes just intercept... next best is coral 
+outer17_all_richness_top_mod <- glmmTMB(richness ~ coral_cover_hard + offset(log(hard_cover)) + (1|site), family = poisson, data = outer17_all_site, na.action = na.pass)
+summary(outer17_all_richness_top_mod)
+hist(residuals(outer17_all_richness_top_mod)) #  not bad
+plot(residuals(outer17_all_richness_top_mod) ~ fitted(outer17_all_richness_top_mod)) # looks fine
+
+outer17_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c + (1|site), data = outer17_all_site, na.action = na.pass)
+outer17_all_shannon_dredge <- dredge(outer17_all_shannon_mod_universal) # some convergence issues
+model.sel(outer17_all_shannon_dredge) #  best includes just year
+outer17_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  year + (1|site), data = outer17_all_site, na.action = na.pass)
+hist(residuals(outer17_all_shannon_top_mod)) # pretty good
+plot(residuals(outer17_all_shannon_top_mod) ~ fitted(outer17_all_shannon_top_mod)) # looks fine
+```
+
+```{r all algae alpha diversity quadrat level}
 
@@
 ```{r macroalgae diversity quadrat level}
-alpha_diversity_quad_macro <-
-  data.frame(
-    year = macroalgae$year,
-    location = macroalgae$location,
-    shannon = diversity(macroalgae[ , -which(names(macroalgae) %in% meta_cols)], "shannon"),
-    richness = rowSums(macroalgae[ , -which(names(macroalgae) %in% meta_cols)] > 0)
-  )
-
-alpha_diversity_quad_macro <- merge(meta_df, alpha_diversity_quad_macro, by = c("location", "year"))
 
 ```
-
-```{r all algae alpha diversity site level}
-alpha_diversity_site <-
-  data.frame(
-    year = site_full$year,
-    site_hab_year = site_full$site_hab_year,
-    shannon = diversity(site_full[ , -which(names(site_full) %in% meta_cols)], "shannon"),
-    richness = rowSums(site_full[ , -which(names(site_full) %in% meta_cols)] > 0)
-  )
-
-alpha_diversity_site <- merge(site_meta_df, alpha_diversity_site, by= c("year", "site_hab_year"))
-
-site_full_richness_mod <- glmmTMB(richness ~ year*habitat + (1|site) + offset(log(hard_cover)), family = poisson,
-                             data = alpha_diversity_site) 
-# Warning in (function (start, objective, gradient = NULL, hessian = NULL,  : NA/NaN function evaluation
-# This warning occurs when the optimizer visits a region of parameter space that is invalid.
-It is not a problem as long as the optimizer has left that region of parameter space upon convergence, which is indicated by an absence of the model convergence warnings described above.
-summary(site_full_richness_mod)
-hist(residuals(site_full_richness_mod)) # similar slight skew to before
-
-# striping/patterining expected because of discrete response var. otherwise fine
-plot(residuals(site_full_richness_mod) ~ fitted(site_full_richness_mod)) # slightly heteroskedastic
-
-emtrends(site_full_richness_mod, pairwise ~  habitat, var = "year") # habitat trends don't differ from each other
-site_full_richness_emm <- emmip(site_full_richness_mod,habitat ~ year,
-                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)
-
-(site_base_richness_plot <- 
-  ggplot() +
-  geom_point(data = alpha_diversity_site,
-             aes(x = year, y = richness/(hard_cover/100), colour = habitat), size = 2, alpha = 0.5) + 
-  geom_line(data = site_full_richness_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
-  geom_ribbon(data = site_full_richness_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
-  scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
-  scale_y_log10() +
-    labs(x = "Year", y = "Richness") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black")) ) # add box around legend
-
-site_full_shannon_mod <- glmmTMB(log(shannon/(hard_cover/100) +1 ) ~ year*habitat + (1|site),
-                             data = alpha_diversity_site) 
-# sites with shannon div = 0 so log+1 transformed
-summary(site_full_shannon_mod)
-hist(residuals(site_full_shannon_mod)) # looks good
-
-# slight patterning but generally fine
-plot(residuals(site_full_shannon_mod) ~ fitted(site_full_shannon_mod))
-
-emtrends(site_full_shannon_mod, pairwise ~  habitat, var = "year") # backreef and fringing differ
-site_full_shannon_emm <- emmip(site_full_shannon_mod,habitat ~ year,
-                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)
-
-(site_base_shannon_plot <- 
-  ggplot() +
-  geom_point(data = alpha_diversity_site,
-             aes(x = year, y = shannon/(hard_cover/100), colour = habitat), size = 2, alpha = 0.5) + 
-  geom_line(data = site_full_shannon_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
-  geom_ribbon(data = site_full_shannon_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
-  scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
-  scale_y_log10() +
-    labs(x = "Year", y = "shannon") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black")) ) # add box around legend
-
-```
-
-```{r macroalgae alpha diversity site level}
-alpha_diversity_site_macro <-
-  data.frame(
-    year = site_macro$year,
-    site_hab_year = site_macro$site_hab_year,
-    shannon = diversity(site_macro[ , -which(names(site_macro) %in% meta_cols)], "shannon"),
-    richness = rowSums(site_macro[ , -which(names(site_macro) %in% meta_cols)] > 0)
-  )
-
-alpha_diversity_site_macro <- merge(site_meta_df, alpha_diversity_site_macro, by= c("year", "site_hab_year"))
-
-site_macro_richness_mod <- glmmTMB(richness ~ year*habitat + (1|site) + offset(log(hard_cover)), family = poisson,
-                             data = alpha_diversity_site_macro) 
-# Warning in (function (start, objective, gradient = NULL, hessian = NULL,  : NA/NaN function evaluation
-# This warning occurs when the optimizer visits a region of parameter space that is invalid.
-It is not a problem as long as the optimizer has left that region of parameter space upon convergence, which is indicated by an absence of the model convergence warnings described above.
-summary(site_macro_richness_mod)
-hist(residuals(site_macro_richness_mod)) # slightly more skew
-
-# striping/patterining expected because of discrete response var. otherwise fine
-plot(residuals(site_macro_richness_mod) ~ fitted(site_macro_richness_mod))
-
-emtrends(site_macro_richness_mod, pairwise ~  habitat, var = "year") # habitat trends don't differ from each other
-site_macro_richness_emm <- emmip(site_macro_richness_mod,habitat ~ year,
-                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)
-
-(site_base_richness_plot <- 
-  ggplot() +
-  geom_point(data = alpha_diversity_site_macro,
-             aes(x = year, y = richness/(hard_cover/100), colour = habitat), size = 2, alpha = 0.5) + 
-  geom_line(data = site_macro_richness_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
-  geom_ribbon(data = site_macro_richness_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
-  scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
-  scale_y_log10(lim = c(0.1,20)) +
-    labs(x = "Year", y = "Richness") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black")) ) # add box around legend
-
-site_macro_shannon_mod <- glmmTMB(log(shannon/(hard_cover/100) +1 ) ~ year*habitat + (1|site),
-                             data = alpha_diversity_site) 
-# sites with shannon div = 0 so log+1 transformed
-summary(site_macro_shannon_mod)
-hist(residuals(site_macro_shannon_mod)) # looks good
-
-# slight patterning but generally fine
-plot(residuals(site_macro_shannon_mod) ~ fitted(site_macro_shannon_mod))
-
-emtrends(site_macro_shannon_mod, pairwise ~  habitat, var = "year") # backreef and fringing differ
-site_macro_shannon_emm <- emmip(site_macro_shannon_mod,habitat ~ year,
-                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)
-
-(site_base_shannon_plot <- 
-  ggplot() +
- # geom_point(data = alpha_diversity_site,
- #            aes(x = year, y = shannon/(hard_cover/100), colour = habitat), size = 2, alpha = 0.5) + 
-  geom_line(data = site_macro_shannon_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
-  geom_ribbon(data = site_macro_shannon_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
-  scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
-#  scale_y_log10() +
-    labs(x = "Year", y = "shannon") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black")) ) # add box around legend
-
-```

### 2024-09-04 noam_beta_diversity.Rmd
more beta dispersion work

@@
    scale_fill_manual(values = habitat_colours) +
-
+   geom_segment(aes(y = 0.6, yend = 0.6, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+   geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+   geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
    labs(y = "Beta dispersion\n", x = "Year") +
@@
 full_subset_site_macro$betadispersion_trans <- sv_trans(full_subset_site_macro$betadispersion, full_subset_site_macro)
+
+betadisper_site_macro_mod <- glmmTMB(betadispersion_trans ~ year*habitat + (1|site), family = beta_family(link = "logit"), data = full_subset_site_macro)
+summary(betadisper_site_macro_mod) # similar site variance - inv_logit(0.01) = 0.5
+hist(residuals(betadisper_site_macro_mod)) # pretty good
+plot(residuals(betadisper_site_macro_mod) ~ fitted(betadisper_site_macro_mod)) # slight heteroskedasticity and negative trend
+Anova(betadisper_site_macro_mod)
+#               Chisq Df Pr(>Chisq)    
+# year         36.591  1  1.457e-09 ***
+# habitat      46.029  3  5.590e-10 ***
+# year:habitat 31.232  3  7.597e-07 ***
+
+emtrends(betadisper_site_macro_mod, pairwise ~ habitat , "year") # backreef distinct from all other habitats
+betadisper_site_macro_emm <- emmip(betadisper_site_macro_mod, habitat ~ year , at = list(year = seq(2007,2023,1)), plotit = F, CIs = T)
+
+ggplot()+
+   geom_point(data = site_full, aes(x = year, y = betadispersion_trans, colour = habitat), alpha = 0.1 ) +
+   geom_line(data = betadisper_site_macro_emm, aes(x = year, y = inv_logit(yvar), colour = habitat), linewidth = 2) +
+   geom_ribbon(data = betadisper_site_macro_emm, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+   scale_fill_manual(values = habitat_colours) +
+   geom_segment(aes(y = 0.75, yend = 0.75, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+   geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+   geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+   labs(y = "Beta dispersion\n", x = "Year") +
+   theme_bw() + 
+   theme(panel.grid.major = element_blank(), 
+         panel.grid.minor = element_blank(),
+         panel.background = element_blank(),
+         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+         axis.line.y = element_line(colour = "black"),
+         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         axis.line.x = element_line(colour = "black"),
+         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         strip.text.x = element_text(size = 32),
+         legend.key.size = unit(2, "cm"), # increase legend size
+         legend.text = element_text(size = rel(2)), # increase legend text
+         legend.title = element_text(size = rel(2.25)),
+         legend.position = 'bottom', # move legend to bottom
+         legend.box = 'horizontal',
+         legend.box.background = element_rect(colour = "black") )# make legend horizontal
 

### 2024-09-05 noam_final_analyses.Rmd
added synchrony diversity stability models at the site level for all algae and just macroalgae. Added figures as well.

@@
 
+```{r synchrony diversity stability setup site level, fig.height = 8, fig.width= 8}
+
+lm_synchrony_site <- codyn::synchrony(df = site_long_data,
+                                   time.var = "year",
+                                   species.var = "taxa",
+                                   abundance.var = "percent_cover",
+                                   metric = "Loreau",
+                                   replicate.var = "site_habitat") 
+colnames(lm_synchrony_site) = c("site_habitat", "loreau_synchrony") # rename cols for merge later
+
+lm_synchrony_site_macro <- codyn::synchrony(df = site_long_macro,
+                                   time.var = "year",
+                                   species.var = "taxa",
+                                   abundance.var = "percent_cover",
+                                   metric = "Loreau",
+                                   replicate.var = "site_habitat") 
+colnames(lm_synchrony_site_macro) = c("site_habitat", "loreau_synchrony") # rename cols for merge later
+
+diversity_cv_site <- alpha_diversity_site %>%
+  dplyr::group_by(site, habitat, island_side, site_habitat) %>%
+  dplyr::summarise(
+            algae_richness_mean = mean(richness), 
+            algae_richness_cv = CV(richness),
+            algae_richness_stability = 1/CV(richness),
+            algae_shannon_mean = mean(shannon), 
+            algae_shannon_cv = CV(shannon),
+            algae_shannon_stability = 1/CV(shannon), 
+            algae_cover_mean = mean(all_algae_cover_hard),
+            algae_cover_cv = CV(all_algae_cover_hard),
+            algae_cover_stability = 1/CV(all_algae_cover_hard)) 
+# an extra outer_10 was introduced for every site for some reason that is all NAs
+# remove NAs and merge with synchrony
+diversity_stability_site <- merge(lm_synchrony_site, na.omit(diversity_cv_site), by = "site_habitat")
+
+diversity_cv_site_macro <- alpha_diversity_site_macro %>%
+  dplyr::group_by(site, habitat, island_side, site_habitat) %>%
+  dplyr::summarise(
+            algae_richness_mean = mean(richness), 
+            algae_richness_cv = CV(richness),
+            algae_richness_stability = 1/CV(richness),
+            algae_shannon_mean = mean(shannon), 
+            algae_shannon_cv = CV(shannon),
+            algae_shannon_stability = 1/CV(shannon), 
+            algae_cover_mean = mean(macroalgae_cover_hard),
+            algae_cover_cv = CV(macroalgae_cover_hard),
+            algae_cover_stability = 1/CV(macroalgae_cover_hard)) 
+# an extra outer_10 was introduced for every site for some reason that is all NAs
+# remove NAs and merge with synchrony
+diversity_stability_site_macro <- merge(lm_synchrony_site_macro, na.omit(diversity_cv_site_macro), by = "site_habitat")
+
+# for plotting later
+diversity_ranges_site <- diversity_stability_site %>%
+  group_by(habitat) %>%
+  summarise(min_richness = min(algae_richness_mean),
+            max_richness = max(algae_richness_mean),
+            min_shannon = min(algae_shannon_mean),
+            max_shannon = max(algae_shannon_mean),
+            min_loreau = min(loreau_synchrony),
+            max_loreau = max(loreau_synchrony))
+
+diversity_ranges_site_macro <- diversity_stability_site_macro %>%
+  group_by(habitat) %>%
+  summarise(min_richness = min(algae_richness_mean),
+            max_richness = max(algae_richness_mean),
+            min_shannon = min(algae_shannon_mean),
+            max_shannon = max(algae_shannon_mean),
+            min_loreau = min(loreau_synchrony),
+            max_loreau = max(loreau_synchrony))
+
+```
+
+```{r site all diversity stability synchrony, fig.height = 8, fig.width=8}
+richness_cv_site <- glmmTMB(algae_cover_cv ~ algae_richness_mean*habitat, family = Gamma(link = "log"),
+                          data = diversity_stability_site)
+r.squaredGLMM(richness_cv_site) # 0.49
+summary(richness_cv_site)
+hist(residuals(richness_cv_site)) # symmetrical
+plot(residuals(richness_cv_site) ~ fitted(richness_cv_site)) # slightly heteroskedastic
+Anova(richness_cv_site)
+#                               Chisq Df Pr(>Chisq)   
+# algae_richness_mean          5.9078  1   0.015074 * 
+# habitat                     14.5560  3   0.002238 **
+# algae_richness_mean:habitat  3.8220  3   0.281336
+
+emtrends(richness_cv_site, pairwise ~ habitat, var = "algae_richness_mean") # no significant differences
+richness_cv_emm_site <- emmip(richness_cv_site,
+                        habitat ~ algae_richness_mean,
+                        at = list(algae_richness_mean = seq(0,10,0.1)), plotit = F, CIs = T)
+
+filtered_richness_effects_site <- richness_cv_emm_site %>%
+  left_join(diversity_ranges_site, by = c("habitat")) %>%
+  filter(algae_richness_mean >= min_richness & algae_richness_mean <= max_richness)
+
+(richness_cv_plot = 
+  ggplot() +
+  geom_point(data = diversity_stability_site,
+             aes(x = algae_richness_mean, y = algae_cover_cv, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = filtered_richness_effects_site, aes(x = algae_richness_mean, y = exp(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = filtered_richness_effects_site, aes(x = algae_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+## REPEAT WITH SHANNON
+shannon_cv_site <- glmmTMB(algae_cover_cv ~ algae_shannon_mean*habitat, family = Gamma(link = "log"),
+                          data = diversity_stability_site)
+r.squaredGLMM(shannon_cv_site) # 0.42
+summary(shannon_cv_site)
+hist(residuals(shannon_cv_site)) # symmetrical
+plot(residuals(shannon_cv_site) ~ fitted(shannon_cv_site)) # slightly heteroskedastic
+Anova(shannon_cv_site)
+#                              Chisq Df Pr(>Chisq)   
+# algae_shannon_mean          1.5291  1   0.216254   
+# habitat                    12.1017  3   0.007043 **
+# algae_shannon_mean:habitat  2.8675  3   0.412512
+
+emtrends(shannon_cv_site, pairwise ~ habitat, var = "algae_shannon_mean") # no significant differences
+shannon_cv_emm_site <- emmip(shannon_cv_site,
+                        habitat ~ algae_shannon_mean,
+                        at = list(algae_shannon_mean = seq(0,10,0.1)), plotit = F, CIs = T)
+
+filtered_shannon_effects_site <- shannon_cv_emm_site %>%
+  left_join(diversity_ranges_site, by = c("habitat")) %>%
+  filter(algae_shannon_mean >= min_shannon & algae_shannon_mean <= max_shannon)
+
+(shannon_cv_plot = 
+  ggplot() +
+  geom_point(data = diversity_stability_site,
+             aes(x = algae_shannon_mean, y = algae_cover_cv, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = filtered_shannon_effects_site, aes(x = algae_shannon_mean, y = exp(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = filtered_shannon_effects_site, aes(x = algae_shannon_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+synchrony_cv_site <- glmmTMB(algae_cover_cv ~ loreau_synchrony*habitat, family = Gamma(link = "log"),
+                          data = diversity_stability_site)
+r.squaredGLMM(synchrony_cv_site) # 0.76
+summary(synchrony_cv_site)
+hist(residuals(synchrony_cv_site)) # symmetrical
+plot(residuals(synchrony_cv_site) ~ fitted(synchrony_cv_site)) # slightly heteroskedastic
+Anova(synchrony_cv_site)
+#                           Chisq Df Pr(>Chisq)    
+# loreau_synchrony         45.857  1  1.272e-11 ***
+# habitat                  14.018  3    0.00288 ** 
+# loreau_synchrony:habitat  2.456  3    0.48330
+
+emtrends(synchrony_cv_site, pairwise ~ habitat, var = "loreau_synchrony") # no differences between habitats
+synchrony_cv_emm_site <- emmip(synchrony_cv_site,
+                        habitat ~ loreau_synchrony,
+                        at = list(loreau_synchrony = seq(0,1,0.01)), plotit = F, CIs = T)
+
+filtered_synchrony_cv_effects_site <- synchrony_cv_emm_site %>%
+  left_join(diversity_ranges_site, by = c("habitat")) %>%
+  filter(loreau_synchrony >= min_loreau & loreau_synchrony <= max_loreau)
+
+(synchrony_cv_plot = 
+  ggplot() +
+  geom_point(data = diversity_stability_site,
+             aes(x = loreau_synchrony, y = algae_cover_cv, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = filtered_synchrony_cv_effects_site, aes(x = loreau_synchrony, y = exp(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = filtered_synchrony_cv_effects_site, aes(x = loreau_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+synchrony_richness_site <- glmmTMB(loreau_synchrony  ~ algae_richness_mean*habitat, family = beta_family(link = "logit"),
+                          data = diversity_stability_site)
+# r.squaredLR(synchrony_richness_site)
+summary(synchrony_richness_site)
+hist(residuals(synchrony_richness_site)) # symmetrical
+plot(residuals(synchrony_richness_site) ~ fitted(synchrony_richness_site)) # slightly heteroskedastic
+Anova(synchrony_richness_site)
+# 
+#                               Chisq Df Pr(>Chisq)    
+# algae_richness_mean         12.6670  1  0.0003722 ***
+# habitat                      5.0218  3  0.1702104    
+# algae_richness_mean:habitat  6.0890  3  0.1073579
+
+emtrends(synchrony_richness_site, pairwise ~ habitat, var = "algae_richness_mean") # no differences between habitats
+synchrony_richness_emm_site <- emmip(synchrony_richness_site,
+                        habitat ~ algae_richness_mean,
+                        at = list(algae_richness_mean = seq(0,10,0.1)), plotit = F, CIs = T)
+
+filtered_synchrony_richness_effects_site <- synchrony_richness_emm_site %>%
+  left_join(diversity_ranges_site, by = c("habitat")) %>%
+  filter(algae_richness_mean >= min_richness & algae_richness_mean <= max_richness)
+
+(synchrony_richness_plot = 
+  ggplot() +
+  geom_point(data = diversity_stability_site,
+             aes(x = algae_richness_mean, y = loreau_synchrony, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = filtered_synchrony_richness_effects_site, aes(x = algae_richness_mean, y = exp(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = filtered_synchrony_richness_effects_site, aes(x = algae_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+synchrony_shannon_site <- glmmTMB(loreau_synchrony  ~ algae_shannon_mean*habitat, family = beta_family(link = "logit"),
+                          data = diversity_stability_site)
+# r.squaredLR(synchrony_shannon_site)
+summary(synchrony_shannon_site)
+hist(residuals(synchrony_shannon_site)) # symmetrical
+plot(residuals(synchrony_shannon_site) ~ fitted(synchrony_shannon_site)) # slightly heteroskedastic
+Anova(synchrony_shannon_site)
+# 
+#                              Chisq Df Pr(>Chisq)   
+# algae_shannon_mean         10.5506  1   0.001161 **
+# habitat                     4.9410  3   0.176171   
+# algae_shannon_mean:habitat  6.1573  3   0.104205
+
+emtrends(synchrony_shannon_site, pairwise ~ habitat, var = "algae_shannon_mean") # no differences between habitats
+synchrony_shannon_emm_site <- emmip(synchrony_shannon_site,
+                        habitat ~ algae_shannon_mean,
+                        at = list(algae_shannon_mean = seq(0,1.5,0.01)), plotit = F, CIs = T)
+
+filtered_synchrony_shannon_effects_site <- synchrony_shannon_emm_site %>%
+  left_join(diversity_ranges_site, by = c("habitat")) %>%
+  filter(algae_shannon_mean >= min_shannon & algae_shannon_mean <= max_shannon)
+
+(synchrony_shannon_plot = 
+  ggplot() +
+  geom_point(data = diversity_stability_site,
+             aes(x = algae_shannon_mean, y = loreau_synchrony, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = filtered_synchrony_shannon_effects_site, aes(x = algae_shannon_mean, y = exp(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = filtered_synchrony_shannon_effects_site, aes(x = algae_shannon_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+```
+
+```{r site macroalgae diversity stability synchrony, fig.height = 8, fig.width=8}
+richness_cv_site_macro <- glmmTMB(algae_cover_cv ~ algae_richness_mean*habitat, family = Gamma(link = "log"),
+                          data = diversity_stability_site_macro)
+r.squaredGLMM(richness_cv_site_macro) # 0.65
+summary(richness_cv_site_macro)
+hist(residuals(richness_cv_site_macro)) # symmetrical
+plot(residuals(richness_cv_site_macro) ~ fitted(richness_cv_site_macro)) # not bad
+Anova(richness_cv_site_macro)
+#                               Chisq Df Pr(>Chisq)   
+# algae_richness_mean         15.772  1  7.144e-05 ***
+# habitat                     21.689  3  7.572e-05 ***
+# algae_richness_mean:habitat 12.244  3   0.006591 **
+
+emtrends(richness_cv_site_macro, pairwise ~ habitat, var = "algae_richness_mean") # backreef differs from two forereef sites
+richness_cv_emm_site_macro <- emmip(richness_cv_site_macro,
+                        habitat ~ algae_richness_mean,
+                        at = list(algae_richness_mean = seq(0,10,0.1)), plotit = F, CIs = T)
+
+filtered_richness_effects_site_macro <- richness_cv_emm_site_macro %>%
+  left_join(diversity_ranges_site_macro, by = c("habitat")) %>%
+  filter(algae_richness_mean >= min_richness & algae_richness_mean <= max_richness)
+
+(richness_cv_plot = 
+  ggplot() +
+  geom_point(data = diversity_stability_site_macro,
+             aes(x = algae_richness_mean, y = algae_cover_cv, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = filtered_richness_effects_site_macro, aes(x = algae_richness_mean, y = exp(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = filtered_richness_effects_site_macro, aes(x = algae_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+## REPEAT WITH SHANNON
+shannon_cv_site_macro <- glmmTMB(algae_cover_cv ~ algae_shannon_mean*habitat, family = Gamma(link = "log"),
+                          data = diversity_stability_site_macro)
+r.squaredGLMM(shannon_cv_site_macro) # 0.37
+summary(shannon_cv_site_macro)
+hist(residuals(shannon_cv_site_macro)) # a little chunky
+plot(residuals(shannon_cv_site_macro) ~ fitted(shannon_cv_site_macro)) # pretty good
+Anova(shannon_cv_site_macro)
+#                             Chisq Df Pr(>Chisq)  
+# algae_shannon_mean         1.1533  1    0.28287  
+# habitat                    7.9691  3    0.04666 *
+# algae_shannon_mean:habitat 4.1867  3    0.24199
+
+emtrends(shannon_cv_site_macro, pairwise ~ habitat, var = "algae_shannon_mean") # no significant differences
+shannon_cv_emm_site_macro <- emmip(shannon_cv_site_macro,
+                        habitat ~ algae_shannon_mean,
+                        at = list(algae_shannon_mean = seq(0,2,0.01)), plotit = F, CIs = T)
+
+filtered_shannon_effects_site_macro <- shannon_cv_emm_site_macro %>%
+  left_join(diversity_ranges_site_macro, by = c("habitat")) %>%
+  filter(algae_shannon_mean >= min_shannon & algae_shannon_mean <= max_shannon)
+
+(shannon_cv_plot = 
+  ggplot() +
+  geom_point(data = diversity_stability_site_macro,
+             aes(x = algae_shannon_mean, y = algae_cover_cv, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = filtered_shannon_effects_site_macro, aes(x = algae_shannon_mean, y = exp(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = filtered_shannon_effects_site_macro, aes(x = algae_shannon_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+synchrony_cv_site_macro <- glmmTMB(algae_cover_cv ~ loreau_synchrony*habitat, family = Gamma(link = "log"),
+                          data = diversity_stability_site_macro)
+r.squaredGLMM(synchrony_cv_site_macro) # 0.71
+summary(synchrony_cv_site_macro)
+hist(residuals(synchrony_cv_site_macro)) # symmetrical
+plot(residuals(synchrony_cv_site_macro) ~ fitted(synchrony_cv_site_macro)) # slightly heteroskedastic
+Anova(synchrony_cv_site_macro)
+#                           Chisq Df Pr(>Chisq)    
+# loreau_synchrony         22.575  1  2.021e-06 ***
+# habitat                  11.611  3   0.008841 ** 
+# loreau_synchrony:habitat 11.389  3   0.009799 **
+
+emtrends(synchrony_cv_site_macro, pairwise ~ habitat, var = "loreau_synchrony") # backreef different from outer17
+synchrony_cv_emm_site_macro <- emmip(synchrony_cv_site_macro,
+                        habitat ~ loreau_synchrony,
+                        at = list(loreau_synchrony = seq(0,1,0.01)), plotit = F, CIs = T)
+
+filtered_synchrony_cv_effects_site_macro <- synchrony_cv_emm_site_macro %>%
+  left_join(diversity_ranges_site_macro, by = c("habitat")) %>%
+  filter(loreau_synchrony >= min_loreau & loreau_synchrony <= max_loreau)
+
+(synchrony_cv_plot = 
+  ggplot() +
+  geom_point(data = diversity_stability_site_macro,
+             aes(x = loreau_synchrony, y = algae_cover_cv, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = filtered_synchrony_cv_effects_site_macro, aes(x = loreau_synchrony, y = exp(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = filtered_synchrony_cv_effects_site_macro, aes(x = loreau_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+synchrony_richness_site_macro <- glmmTMB(loreau_synchrony  ~ algae_richness_mean*habitat, family = beta_family(link = "logit"),
+                          data = diversity_stability_site_macro)
+# r.squaredGLMM(synchrony_richness_site_macro)
+summary(synchrony_richness_site_macro)
+hist(residuals(synchrony_richness_site_macro)) # kind of chunky
+plot(residuals(synchrony_richness_site_macro) ~ fitted(synchrony_richness_site_macro)) # slightly heteroskedastic
+Anova(synchrony_richness_site_macro)
+# 
+#                               Chisq Df Pr(>Chisq)    
+# algae_richness_mean         0.2299  1    0.63161  
+# habitat                     5.4261  3    0.14312  
+# algae_richness_mean:habitat 9.1650  3    0.02718 *
+
+emtrends(synchrony_richness_site_macro, pairwise ~ habitat, var = "algae_richness_mean") # backreef almost diff from fing and outer17
+synchrony_richness_emm_site_macro <- emmip(synchrony_richness_site_macro,
+                        habitat ~ algae_richness_mean,
+                        at = list(algae_richness_mean = seq(0,10,0.1)), plotit = F, CIs = T)
+
+filtered_synchrony_richness_effects_site_macro <- synchrony_richness_emm_site_macro %>%
+  left_join(diversity_ranges_site_macro, by = c("habitat")) %>%
+  filter(algae_richness_mean >= min_richness & algae_richness_mean <= max_richness)
+
+(synchrony_richness_plot = 
+  ggplot() +
+  geom_point(data = diversity_stability_site_macro,
+             aes(x = algae_richness_mean, y = loreau_synchrony, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = filtered_synchrony_richness_effects_site_macro, aes(x = algae_richness_mean, y = exp(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = filtered_synchrony_richness_effects_site_macro, aes(x = algae_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+synchrony_shannon_site_macro <- glmmTMB(loreau_synchrony  ~ algae_shannon_mean*habitat, family = beta_family(link = "logit"),
+                          data = diversity_stability_site_macro)
+summary(synchrony_shannon_site_macro)
+hist(residuals(synchrony_shannon_site_macro)) # a little chunky
+plot(residuals(synchrony_shannon_site_macro) ~ fitted(synchrony_shannon_site_macro)) # slightly heteroskedastic
+Anova(synchrony_shannon_site_macro)
+# 
+#                              Chisq Df Pr(>Chisq)   
+# algae_richness_mean         0.2299  1    0.63161  
+# habitat                     5.4261  3    0.14312  
+# algae_richness_mean:habitat 9.1650  3    0.02718 *
+
+emtrends(synchrony_shannon_site_macro, pairwise ~ habitat, var = "algae_shannon_mean") # no differences between habitats
+synchrony_shannon_emm_site_macro <- emmip(synchrony_shannon_site_macro,
+                        habitat ~ algae_shannon_mean,
+                        at = list(algae_shannon_mean = seq(0,1.5,0.01)), plotit = F, CIs = T)
+
+filtered_synchrony_shannon_effects_site_macro <- synchrony_shannon_emm_site_macro %>%
+  left_join(diversity_ranges_site_macro, by = c("habitat")) %>%
+  filter(algae_shannon_mean >= min_shannon & algae_shannon_mean <= max_shannon)
+
+(synchrony_shannon_plot = 
+  ggplot() +
+  geom_point(data = diversity_stability_site_macro,
+             aes(x = algae_shannon_mean, y = loreau_synchrony, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = filtered_synchrony_shannon_effects_site_macro, aes(x = algae_shannon_mean, y = exp(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = filtered_synchrony_shannon_effects_site_macro, aes(x = algae_shannon_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+```
+
 ```{r all algae alpha diversity quadrat level}

### 2024-09-05 noam_final_analyses.Rmd
Added turnover models for site level

@@
 
+inv_logit <- function(x){
+  exp(x)/(1+exp(x))
+}
+
+sv_trans <- function(prop, s = 0.000005){
+  (prop*(length(prop) - 1) + s)/length(prop)
+  # where prop is the vector of the proportional value you're transforming, 
+  # N is the sample size, which is specified by taking the number of rows/observations from a given dataframe,
+  # and s is a small offset 
+}
+
 # for figures
@@
                         habitat ~ algae_shannon_mean,
-                        at = list(algae_shannon_mean = seq(0,10,0.1)), plotit = F, CIs = T)
+                        at = list(algae_shannon_mean = seq(0,10,0.01)), plotit = F, CIs = T)
 
@@
              aes(x = algae_richness_mean, y = loreau_synchrony, colour = habitat), size = 2, alpha = 0.5) + 
-  geom_line(data = filtered_synchrony_richness_effects_site, aes(x = algae_richness_mean, y = exp(yvar), colour = habitat), size = 2) +
-  geom_ribbon(data = filtered_synchrony_richness_effects_site, aes(x = algae_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+  geom_line(data = filtered_synchrony_richness_effects_site, aes(x = algae_richness_mean, y = inv_logit(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = filtered_synchrony_richness_effects_site, aes(x = algae_richness_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
               alpha = 0.3) +
@@
              aes(x = algae_shannon_mean, y = loreau_synchrony, colour = habitat), size = 2, alpha = 0.5) + 
-  geom_line(data = filtered_synchrony_shannon_effects_site, aes(x = algae_shannon_mean, y = exp(yvar), colour = habitat), size = 2) +
-  geom_ribbon(data = filtered_synchrony_shannon_effects_site, aes(x = algae_shannon_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+  geom_line(data = filtered_synchrony_shannon_effects_site, aes(x = algae_shannon_mean, y = inv_logit(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = filtered_synchrony_shannon_effects_site, aes(x = algae_shannon_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
               alpha = 0.3) +
@@
 
-emtrends(synchrony_cv_site_macro, pairwise ~ habitat, var = "loreau_synchrony") # backreef different from outer17
+emtrends(synchrony_cv_site_macro, pairwise ~ habitat, var = "loreau_synchrony") # backreef and fringe different from outer17
 synchrony_cv_emm_site_macro <- emmip(synchrony_cv_site_macro,
@@
              aes(x = algae_richness_mean, y = loreau_synchrony, colour = habitat), size = 2, alpha = 0.5) + 
-  geom_line(data = filtered_synchrony_richness_effects_site_macro, aes(x = algae_richness_mean, y = exp(yvar), colour = habitat), size = 2) +
-  geom_ribbon(data = filtered_synchrony_richness_effects_site_macro, aes(x = algae_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+  geom_line(data = filtered_synchrony_richness_effects_site_macro, aes(x = algae_richness_mean, y = inv_logit(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = filtered_synchrony_richness_effects_site_macro, aes(x = algae_richness_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
               alpha = 0.3) +
@@
 #                              Chisq Df Pr(>Chisq)   
-# algae_richness_mean         0.2299  1    0.63161  
-# habitat                     5.4261  3    0.14312  
-# algae_richness_mean:habitat 9.1650  3    0.02718 *
+# algae_shannon_mean         0.6238  1     0.4296
+# habitat                    4.8196  3     0.1855
+# algae_shannon_mean:habitat 2.2517  3     0.5218
 
@@
              aes(x = algae_shannon_mean, y = loreau_synchrony, colour = habitat), size = 2, alpha = 0.5) + 
-  geom_line(data = filtered_synchrony_shannon_effects_site_macro, aes(x = algae_shannon_mean, y = exp(yvar), colour = habitat), size = 2) +
-  geom_ribbon(data = filtered_synchrony_shannon_effects_site_macro, aes(x = algae_shannon_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+  geom_line(data = filtered_synchrony_shannon_effects_site_macro, aes(x = algae_shannon_mean, y = inv_logit(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = filtered_synchrony_shannon_effects_site_macro, aes(x = algae_shannon_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
               alpha = 0.3) +
@@
          )
+
+```
+
+```{r site all turnover}
+total_turnover_site<-
+  turnover(
+  df = site_long_data,
+  time.var = "year",
+  species.var = "taxa",
+  abundance.var = "percent_cover",
+  replicate.var = "site_habitat",
+  metric = "total"
+)
+total_turnover_site$site_hab_year <- paste0(total_turnover_site$site_habitat, "_", total_turnover_site$year) # for merge
+total_turnover_site <- merge(total_turnover_site, site_meta_df, by = c("year", "site_habitat", "site_hab_year") )
+
+total_turnover_site$turnover_trans <- sv_trans(total_turnover_site$total)
+
+total_turnover_mod_site <- glmmTMB(turnover_trans ~ year*habitat + (1 | site), family = beta_family(), data = total_turnover_site)
+summary(total_turnover_mod_site)
+hist(residuals(total_turnover_mod_site)) # a little stumpy nut not bad
+plot(residuals(total_turnover_mod_site) ~ fitted(total_turnover_mod_site)) # negative trend
+Anova(total_turnover_mod_site)
+emtrends(total_turnover_mod_site, pairwise ~ habitat, var = "year") # differences between backreef and fringing
+total_turnover_emm_site <- emmip(total_turnover_mod_site,
+                        habitat ~ year,
+                        at = list(year = seq(2008,2023,1)), plotit = F, CIs = T)
+
+(turnover_site_plot <- 
+  ggplot() +
+  geom_point(data = total_turnover_site,
+             aes(x = year, y = turnover_trans, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = total_turnover_emm_site, aes(x = year, y = inv_logit(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = total_turnover_emm_site, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL),
+                                                  fill = habitat), alpha = 0.3) +
+  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+    labs(x = "Year", y = "Total turnover\n") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) )
+
+```
+```{r site macro turnover}
+total_turnover_site_macro<-
+  turnover(
+  df = site_long_macro,
+  time.var = "year",
+  species.var = "taxa",
+  abundance.var = "percent_cover",
+  replicate.var = "site_habitat",
+  metric = "total"
+)
+total_turnover_site_macro$site_hab_year <- paste0(total_turnover_site_macro$site_habitat, "_", total_turnover_site_macro$year) # for merge
+total_turnover_site_macro <- merge(total_turnover_site_macro, site_meta_df, by = c("year", "site_habitat", "site_hab_year") )
+
+total_turnover_site_macro$turnover_trans <- sv_trans(total_turnover_site_macro$total)
+
+total_turnover_mod_site_macro <- glmmTMB(turnover_trans ~ year*habitat + (1 | site), family = beta_family(), data = total_turnover_site_macro)
+summary(total_turnover_mod_site_macro)
+hist(residuals(total_turnover_mod_site_macro)) # slight negative skew
+plot(residuals(total_turnover_mod_site_macro) ~ fitted(total_turnover_mod_site_macro)) # negative trend
+Anova(total_turnover_mod_site_macro)
+#               Chisq Df Pr(>Chisq)    
+# year         73.936  1  < 2.2e-16 ***
+# habitat      19.201  3  0.0002485 ***
+# year:habitat 12.099  3  0.0070512 ** 
+emtrends(total_turnover_mod_site_macro, pairwise ~ habitat, var = "year") # differences between forereef sites and fringing
+total_turnover_emm_site_macro <- emmip(total_turnover_mod_site_macro,
+                        habitat ~ year,
+                        at = list(year = seq(2008,2023,1)), plotit = F, CIs = T)
+
+(turnover_site_macro_plot <- 
+  ggplot() +
+  geom_point(data = total_turnover_site_macro,
+             aes(x = year, y = turnover_trans, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = total_turnover_emm_site_macro, aes(x = year, y = inv_logit(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = total_turnover_emm_site_macro, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL),
+                                                  fill = habitat), alpha = 0.3) +
+  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+    labs(x = "Year", y = "Total turnover\n") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) )
+
+```
+
+```{r site all appearance}
+appearance_site<-
+  turnover(
+  df = site_long_data,
+  time.var = "year",
+  species.var = "taxa",
+  abundance.var = "percent_cover",
+  replicate.var = "site_habitat",
+  metric = "appearance"
+)
+appearance_site$site_hab_year <- paste0(appearance_site$site_habitat, "_", appearance_site$year) # for merge
+appearance_site <- merge(appearance_site, site_meta_df, by = c("year", "site_habitat", "site_hab_year") )
+
+appearance_site$turnover_trans <- sv_trans(appearance_site$appearance)
+
+appearance_mod_site <- glmmTMB(turnover_trans ~ year*habitat , family = beta_family(), data = appearance_site)
+# convergence issue from inclusion of random effect
+summary(appearance_mod_site)
+hist(residuals(appearance_mod_site)) # a little stumpy nut not bad
+plot(residuals(appearance_mod_site) ~ fitted(appearance_mod_site)) # negative trend
+Anova(appearance_mod_site)
+#                Chisq Df Pr(>Chisq)    
+# year         58.4417  1  2.094e-14 ***
+# habitat       9.5344  3  0.0229685 *  
+# year:habitat 18.4862  3  0.0003491 ***
+emtrends(appearance_mod_site, pairwise ~ habitat, var = "year") # lagoons both diff from forereefs but not each other
+appearance_emm_site <- emmip(appearance_mod_site,
+                        habitat ~ year,
+                        at = list(year = seq(2008,2023,1)), plotit = F, CIs = T)
+
+(appearance_site_plot <- 
+  ggplot() +
+  geom_point(data = appearance_site,
+             aes(x = year, y = turnover_trans, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = appearance_emm_site, aes(x = year, y = inv_logit(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = appearance_emm_site, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL),
+                                                  fill = habitat), alpha = 0.3) +
+  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+    labs(x = "Year", y = "Appearances\n") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) )
+
+```
+```{r site macro disappearance}
+appearance_site_macro<-
+  turnover(
+  df = site_long_macro,
+  time.var = "year",
+  species.var = "taxa",
+  abundance.var = "percent_cover",
+  replicate.var = "site_habitat",
+  metric = "appearance"
+)
+appearance_site_macro$site_hab_year <- paste0(appearance_site_macro$site_habitat, "_", appearance_site_macro$year) # for merge
+appearance_site_macro <- merge(appearance_site_macro, site_meta_df, by = c("year", "site_habitat", "site_hab_year") )
+
+appearance_site_macro$turnover_trans <- sv_trans(appearance_site_macro$appearance)
+
+appearance_mod_site_macro <- glmmTMB(turnover_trans ~ year*habitat, family = beta_family(), data = appearance_site_macro) # same convergence issue
+summary(appearance_mod_site_macro)
+hist(residuals(appearance_mod_site_macro)) # stumpy
+plot(residuals(appearance_mod_site_macro) ~ fitted(appearance_mod_site_macro)) # negative trend
+Anova(appearance_mod_site_macro)
+#               Chisq Df Pr(>Chisq)    
+# year         45.8213  1  1.296e-11 ***
+# habitat       5.8339  3   0.119979    
+# year:habitat 13.7486  3   0.003268 ** 
+emtrends(appearance_mod_site_macro, pairwise ~ habitat, var = "year") # fringing different from fore, backreef from outer10
+appearance_emm_site_macro <- emmip(appearance_mod_site_macro,
+                        habitat ~ year,
+                        at = list(year = seq(2008,2023,1)), plotit = F, CIs = T)
+
+(turnover_site_macro_plot <- 
+  ggplot() +
+  geom_point(data = appearance_site_macro,
+             aes(x = year, y = turnover_trans, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = appearance_emm_site_macro, aes(x = year, y = inv_logit(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = appearance_emm_site_macro, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL),
+                                                  fill = habitat), alpha = 0.3) +
+  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+    labs(x = "Year", y = "Appearances\n") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) )
+
+```
+
+```{r site all disappearance}
+disappearance_site<-
+  turnover(
+  df = site_long_data,
+  time.var = "year",
+  species.var = "taxa",
+  abundance.var = "percent_cover",
+  replicate.var = "site_habitat",
+  metric = "disappearance"
+)
+disappearance_site$site_hab_year <- paste0(disappearance_site$site_habitat, "_", disappearance_site$year) # for merge
+disappearance_site <- merge(disappearance_site, site_meta_df, by = c("year", "site_habitat", "site_hab_year") )
+
+disappearance_site$turnover_trans <- sv_trans(disappearance_site$disappearance)
+
+disappearance_mod_site <- glmmTMB(turnover_trans ~ year*habitat + (1|site), family = beta_family(), data = disappearance_site)
+# no convergence issue
+summary(disappearance_mod_site)
+hist(residuals(disappearance_mod_site)) # positive skew
+plot(residuals(disappearance_mod_site) ~ fitted(disappearance_mod_site)) # negative trend
+Anova(disappearance_mod_site)
+#                Chisq Df Pr(>Chisq)    
+# year         10.402  1   0.001259 ** 
+# habitat      22.685  3  4.698e-05 ***
+# year:habitat 12.099  3   0.007053 ** 
+emtrends(disappearance_mod_site, pairwise ~ habitat, var = "year") # backrreef different from outer 10
+disappearance_emm_site <- emmip(disappearance_mod_site,
+                        habitat ~ year,
+                        at = list(year = seq(2008,2023,1)), plotit = F, CIs = T)
+
+(disappearance_site_plot <- 
+  ggplot() +
+  geom_point(data = disappearance_site,
+             aes(x = year, y = turnover_trans, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = disappearance_emm_site, aes(x = year, y = inv_logit(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = disappearance_emm_site, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL),
+                                                  fill = habitat), alpha = 0.3) +
+  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+    labs(x = "Year", y = "Disappearances\n") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) )
+
+```
+```{r site all turnover}
+disappearance_site_macro<-
+  turnover(
+  df = site_long_macro,
+  time.var = "year",
+  species.var = "taxa",
+  abundance.var = "percent_cover",
+  replicate.var = "site_habitat",
+  metric = "disappearance"
+)
+disappearance_site_macro$site_hab_year <- paste0(disappearance_site_macro$site_habitat, "_", disappearance_site_macro$year) # for merge
+disappearance_site_macro <- merge(disappearance_site_macro, site_meta_df, by = c("year", "site_habitat", "site_hab_year") )
+
+disappearance_site_macro$turnover_trans <- sv_trans(disappearance_site_macro$disappearance)
+
+disappearance_mod_site_macro <- glmmTMB(turnover_trans ~ year*habitat + (1|site), family = beta_family(), data = disappearance_site_macro) # same here
+summary(disappearance_mod_site_macro)
+hist(residuals(disappearance_mod_site_macro)) # not bad
+plot(residuals(disappearance_mod_site_macro) ~ fitted(disappearance_mod_site_macro)) # negative trend
+Anova(disappearance_mod_site_macro)
+#               Chisq Df Pr(>Chisq)    
+#                Chisq Df Pr(>Chisq)    
+# year         33.3709  1  7.616e-09 ***
+# habitat      11.5079  3   0.009274 ** 
+# year:habitat  1.5899  3   0.661677    
+emtrends(disappearance_mod_site_macro, pairwise ~ habitat, var = "year") # no sig difference
+disappearance_emm_site_macro <- emmip(disappearance_mod_site_macro,
+                        habitat ~ year,
+                        at = list(year = seq(2008,2023,1)), plotit = F, CIs = T)
+
+(turnover_site_macro_plot <- 
+  ggplot() +
+  geom_point(data = disappearance_site_macro,
+             aes(x = year, y = turnover_trans, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = disappearance_emm_site_macro, aes(x = year, y = inv_logit(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = disappearance_emm_site_macro, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL),
+                                                  fill = habitat), alpha = 0.3) +
+  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+    labs(x = "Year", y = "Disappearances\n") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) )
 

### 2024-09-06 noam_final_analyses.Rmd
added model selection for macroalgae, but I can't get stat summaries for forereef data even though there are no warnings in the model or the dredge.

@@
 alpha_diversity_site_macro <- merge(site_meta_df, alpha_diversity_site_macro, by= c("year", "site_hab_year"))
+alpha_diversity_site_macro$richness_area <- alpha_diversity_site_macro$richness/(alpha_diversity_site_macro$hard_cover/100)
 
@@
 fringing_all_site <- na.omit(subset(alpha_diversity_site, habitat == "fringing")) # only use rows that contain all factors
-
-fringing_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + mean_n + mean_temp_c +  herbivore_evenness + offset(log(hard_cover)) + (1|site), family = poisson, data = fringing_all_site, na.action = na.pass) # universal model has convergence issues
-fringing_all_richness_dredge <- dredge(fringing_all_richness_mod_universal) # most models have convergence issues
-model.sel(fringing_all_richness_dredge) # top model just includes year
-fringing_all_richness_top_mod <- glmmTMB(richness ~ year + offset(log(hard_cover)) + (1|site), family = poisson, data = fringing_all_site, na.action = na.pass)
-summary(fringing_all_richness_top_mod)
-hist(residuals(fringing_all_richness_top_mod)) # slight skew but not bad
-plot(residuals(fringing_all_richness_top_mod) ~ fitted(fringing_all_richness_top_mod)) # looks fine
-
-fringing_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = fringing_all_site, na.action = na.pass)
+# 
+# fringing_all_richness_mod_universal <- glmmTMB(richness ~ year + mean_n + mean_temp_c +  herbivore_evenness +
+# offset(log(hard_cover)) #+ (1|site), family = poisson, data = fringing_all_site, na.action = na.pass) 
+# fringing_all_richness_dredge <- dredge(fringing_all_richness_mod_universal) 
+# model.sel(fringing_all_richness_dredge) # top model just includes year
+# fringing_all_richness_top_mod <- glmmTMB(richness ~ year + offset(log(hard_cover)) + (1|site), family = poisson, data = # fringing_all_site, na.action = na.pass)
+# summary(fringing_all_richness_top_mod)
+# hist(residuals(fringing_all_richness_top_mod)) # slight skew but not bad
+# plot(residuals(fringing_all_richness_top_mod) ~ fitted(fringing_all_richness_top_mod)) # looks fine
+
+fringing_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = fringing_all_site, na.action = na.pass)
 fringing_all_shannon_dredge <- dredge(fringing_all_shannon_mod_universal) 
-model.sel(fringing_all_shannon_dredge) # tie for the top model - best includes year and coral cover or just nitrogen
-fringing_all_shannon_top_mod1 <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + (1|site), data = fringing_all_site, na.action = na.pass)
-fringing_all_shannon_top_mod2 <- glmmTMB(log(shannon + 1) ~ mean_n + mean_temp_c + (1|site), data = fringing_all_site, na.action = na.pass)
-hist(residuals(fringing_all_shannon_top_mod1)) # slightly bimodal
-hist(residuals(fringing_all_shannon_top_mod2)) # also uniform - neither are great
-plot(residuals(fringing_all_shannon_top_mod1) ~ fitted(fringing_all_shannon_top_mod1)) # slight patterning
-plot(residuals(fringing_all_shannon_top_mod2) ~ fitted(fringing_all_shannon_top_mod2)) # similar here
+model.sel(fringing_all_shannon_dredge) # tie for the top model just nitrogen
+fringing_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~ mean_n + (1|site), data = fringing_all_site, na.action = na.pass)
+
+qqnorm(residuals(fringing_all_shannon_top_mod))# not bad
+qqline(residuals(fringing_all_shannon_top_mod))
+plot(residuals(fringing_all_shannon_top_mod) ~ fitted(fringing_all_shannon_top_mod)) # slight patterning
 
 ## Backreef - includes all 
-backreef_all_site <- na.omit(subset(alpha_diversity_site, habitat == "backreef")) # only use rows that contain all factors
-
-backreef_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + offset(log(hard_cover)) + (1|site), family = poisson, data = backreef_all_site, na.action = na.pass) # universal model has convergence issues
-backreef_all_richness_dredge <- dredge(backreef_all_richness_mod_universal) # most models have convergence issues
-model.sel(backreef_all_richness_dredge) # top model includes year and coral 
-backreef_all_richness_top_mod <- glmmTMB(richness ~ year + coral_cover_hard + offset(log(hard_cover)) + (1|site), family = poisson, data = backreef_all_site, na.action = na.pass)
-summary(backreef_all_richness_top_mod)
-hist(residuals(backreef_all_richness_top_mod)) # slight skew but not bad
-plot(residuals(backreef_all_richness_top_mod) ~ fitted(backreef_all_richness_top_mod)) # looks fine
-
-backreef_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = backreef_all_site, na.action = na.pass)
+# backreef_all_site <- na.omit(subset(alpha_diversity_site, habitat == "backreef")) # only use rows that contain all factors
+# 
+# backreef_all_richness_mod_universal <- glmmTMB(richness ~ year + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + # offset(log(hard_cover)) + (1|site), family = poisson, data = backreef_all_site, na.action = na.pass) # universal model has # convergence issues
+# backreef_all_richness_dredge <- dredge(backreef_all_richness_mod_universal) # most models have convergence issues
+# model.sel(backreef_all_richness_dredge) # top model includes year and coral 
+# backreef_all_richness_top_mod <- glmmTMB(richness ~ year  + offset(log(hard_cover)) + (1|site), family = poisson, data = # backreef_all_site, na.action = na.pass)
+# summary(backreef_all_richness_top_mod)
+# hist(residuals(backreef_all_richness_top_mod)) # slight skew but not bad
+# plot(residuals(backreef_all_richness_top_mod) ~ fitted(backreef_all_richness_top_mod)) # looks fine
+
+backreef_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = backreef_all_site, na.action = na.pass)
 backreef_all_shannon_dredge <- dredge(backreef_all_shannon_mod_universal) 
-model.sel(backreef_all_shannon_dredge) #  best includes just coral cover 
-backreef_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  coral_cover_hard + (1|site), data = backreef_all_site, na.action = na.pass)
-hist(residuals(backreef_all_shannon_top_mod)) # pretty good
+model.sel(backreef_all_shannon_dredge) #  best includes just year, next best also includes herbivore evenness
+backreef_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  year + (1|site), data = backreef_all_site, na.action = na.pass)
+qqnorm(residuals(backreef_all_shannon_top_mod)) #  a little wonky on the tail but not bad
 plot(residuals(backreef_all_shannon_top_mod) ~ fitted(backreef_all_shannon_top_mod)) # looks fine
+summary(backreef_all_shannon_top_mod)
 
@@
 
-outer10_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c +  herbivore_evenness + offset(log(hard_cover)) + (1|site), family = poisson, data = outer10_all_site, na.action = na.pass) # universal model has convergence issues
-outer10_all_richness_dredge <- dredge(outer10_all_richness_mod_universal) # most models have convergence issues
-model.sel(outer10_all_richness_dredge) # top model includes just intercept... next best is coral 
-outer10_all_richness_top_mod <- glmmTMB(richness ~ coral_cover_hard + offset(log(hard_cover)) + (1|site), family = poisson, data = outer10_all_site, na.action = na.pass)
-summary(outer10_all_richness_top_mod)
-hist(residuals(outer10_all_richness_top_mod)) #  not bad
-plot(residuals(outer10_all_richness_top_mod) ~ fitted(outer10_all_richness_top_mod)) # looks fine
-
-outer10_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c +  herbivore_evenness + (1|site), data = outer10_all_site, na.action = na.pass)
-outer10_all_shannon_dredge <- dredge(outer10_all_shannon_mod_universal) # some convergence issues
-model.sel(outer10_all_shannon_dredge) #  best includes just herbivore evenness - first time there's been a decent gap in AICc
+# outer10_all_richness_mod_universal <- glmmTMB(richness ~ year  + wave_power_adcp  + mean_temp_c +  herbivore_evenness + # offset(log(hard_cover)) + (1|site), family = poisson, data = outer10_all_site, na.action = na.pass) # universal model has convergence # issues
+# outer10_all_richness_dredge <- dredge(outer10_all_richness_mod_universal) # most models have convergence issues
+# model.sel(outer10_all_richness_dredge)
+
+# 
+outer10_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp  + mean_temp_c +  herbivore_evenness + (1|site), data = outer10_all_site, na.action = na.pass)
+outer10_all_shannon_dredge <- dredge(outer10_all_shannon_mod_universal) 
+model.sel(outer10_all_shannon_dredge) #  best includes just herbivore evenness 
 outer10_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  herbivore_evenness + (1|site), data = outer10_all_site, na.action = na.pass)
-hist(residuals(outer10_all_shannon_top_mod)) # pretty good
+qqnorm(residuals(outer10_all_shannon_top_mod)) # pretty good
 plot(residuals(outer10_all_shannon_top_mod) ~ fitted(outer10_all_shannon_top_mod)) # looks fine
+summary(outer10_all_shannon_top_mod)
 
@@
 
-outer17_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c +  offset(log(hard_cover)) + (1|site), family = poisson, data = outer17_all_site, na.action = na.pass) # universal model has convergence issues
-outer17_all_richness_dredge <- dredge(outer17_all_richness_mod_universal) # most models have convergence issues
-model.sel(outer17_all_richness_dredge) # top model includes just intercept... next best is coral 
-outer17_all_richness_top_mod <- glmmTMB(richness ~ coral_cover_hard + offset(log(hard_cover)) + (1|site), family = poisson, data = outer17_all_site, na.action = na.pass)
-summary(outer17_all_richness_top_mod)
-hist(residuals(outer17_all_richness_top_mod)) #  not bad
-plot(residuals(outer17_all_richness_top_mod) ~ fitted(outer17_all_richness_top_mod)) # looks fine
-
-outer17_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c + (1|site), data = outer17_all_site, na.action = na.pass)
+# outer17_all_richness_mod_universal <- glmmTMB(richness ~ year  + wave_power_adcp  + mean_temp_c +  offset(log(hard_cover)) + # (1|site), family = poisson, data = outer17_all_site, na.action = na.pass) # universal model has convergence issues
+# outer17_all_richness_dredge <- dredge(outer17_all_richness_mod_universal) # most models have convergence issues
+# model.sel(outer17_all_richness_dredge) # top model includes just intercept...
+
+outer17_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp  + mean_temp_c + (1|site), data = outer17_all_site, na.action = na.pass)
 outer17_all_shannon_dredge <- dredge(outer17_all_shannon_mod_universal) # some convergence issues
-model.sel(outer17_all_shannon_dredge) #  best includes just year
+model.sel(outer17_all_shannon_dredge) #  best includes just year. next best also includes temperature and is almost equivalent AICc
 outer17_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  year + (1|site), data = outer17_all_site, na.action = na.pass)
-hist(residuals(outer17_all_shannon_top_mod)) # pretty good
+qqnorm(residuals(outer17_all_shannon_top_mod)) # pretty good
 plot(residuals(outer17_all_shannon_top_mod) ~ fitted(outer17_all_shannon_top_mod)) # looks fine
+summary(outer17_all_shannon_top_mod)
+
+```
+
+```{r stepwise models site level macroalgae}
+
+## FRINGING - exclude wave power
+fringing_macro_site <- na.omit(subset(alpha_diversity_site_macro, habitat == "fringing")) # only use rows that contain all factors
+
+fringing_macro_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = fringing_macro_site, na.action = na.pass)
+fringing_macro_shannon_dredge <- dredge(fringing_macro_shannon_mod_universal) 
+model.sel(fringing_macro_shannon_dredge) # top model just year, next best includes temp
+fringing_macro_shannon_top_mod <- glmmTMB(log(shannon + 1) ~ year + (1|site), data = fringing_macro_site, na.action = na.pass)
+qqnorm(residuals(fringing_macro_shannon_top_mod))# not bad
+plot(residuals(fringing_macro_shannon_top_mod) ~ fitted(fringing_macro_shannon_top_mod)) # slight negative trend
+
+## Backreef - includes all 
+backreef_macro_site <- na.omit(subset(alpha_diversity_site_macro, habitat == "backreef")) # only use rows that contain all factors
+
+backreef_macro_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = backreef_macro_site, na.action = na.pass)
+backreef_macro_shannon_dredge <- dredge(backreef_macro_shannon_mod_universal) 
+model.sel(backreef_macro_shannon_dredge) #  best includes just wave power but estimate is tiny
+backreef_macro_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  wave_power_adcp + (1|site), data = backreef_macro_site, na.action = na.pass)
+qqnorm(residuals(backreef_macro_shannon_top_mod)) #  a little wonky but not bad
+plot(residuals(backreef_macro_shannon_top_mod) ~ fitted(backreef_macro_shannon_top_mod)) # looks fine
+summary(backreef_macro_shannon_top_mod)
+Anova(backreef_macro_shannon_top_mod) # lol X2 = 2.56, P = 0.1
+
+## outer10 - no nitrogen 
+outer10_macro_site <- subset(alpha_diversity_site_macro, habitat == "outer_10") # only use rows that contain all factors
+# remove nitrogent columns
+outer10_macro_site<- outer10_macro_site[,-which(names(outer10_macro_site) %in% c("mean_n", "n_cv"))]
+# remove other na rows
+outer10_macro_site <- na.omit(outer10_macro_site)
+
+outer10_macro_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp  + mean_temp_c +  herbivore_evenness + (1|site), data = outer10_macro_site, na.action = na.pass)
+outer10_macro_shannon_dredge <- dredge(outer10_macro_shannon_mod_universal) 
+model.sel(outer10_macro_shannon_dredge) #  best includes wave power and year 
+outer10_macro_shannon_top_mod <- glmmTMB(log(shannon + 1) ~ year  + (1|site), data = outer10_macro_site, na.action = na.pass)
+qqnorm(residuals(outer10_macro_shannon_top_mod)) # pretty good, tails maybe off
+plot(residuals(outer10_macro_shannon_top_mod) ~ fitted(outer10_macro_shannon_top_mod)) # looks fine
+summary(outer10_macro_shannon_top_mod) # NaN?
+
+## outer17 - no nitrogen or herbivores
+outer17_macro_site<- subset(alpha_diversity_site_macro, habitat == "outer_17") # only use rows that contain all factors
+# remove nitrogent columns
+outer17_macro_site<- outer17_macro_site[,-which(names(outer17_macro_site) %in% c("mean_n", "n_cv", "herbivore_biomass", "herbivore_evenness", "herbivore_richness", "herbivore_shannon"))]
+# remove other na rows
+outer17_macro_site <- na.omit(outer17_macro_site)
+
+outer17_macro_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp  + mean_temp_c + (1|site), data = outer17_macro_site, na.action = na.pass)
+outer17_macro_shannon_dredge <- dredge(outer17_macro_shannon_mod_universal) # some convergence issues
+model.sel(outer17_macro_shannon_dredge) #  best includes all terms!
+qqnorm(residuals(outer17_macro_shannon_mod_universal)) # a little hump shaped
+plot(residuals(outer17_macro_shannon_mod_universal) ~ fitted(outer17_macro_shannon_mod_universal)) # looks fine
+summary(outer17_macro_shannon_mod_universal) # also NaNs...
+
 ```

### 2024-09-06 noam_beta_diversity.Rmd
added beta dispersion for quad level

@@
 plot(residuals(bray_mod_site) ~ fitted(bray_mod_site)) # some heteroskedasticity but not bad
-
+Anova(bray_mod_site)
 emtrends(bray_mod_site, pairwise ~ habitat , "time") # no sig differences
@@
 plot(residuals(bray_mod_site_baseline) ~ fitted(bray_mod_site_baseline) ) # heteroskedastic
-
-emtrends(bray_mod_site_baseline, pairwise ~ habitat , "time") # fringing slope significantly different from all others
+Anova(bray_mod_site_baseline)
+#                Chisq Df Pr(>Chisq)   
+# time          1.9842  1   0.158944   
+# habitat      12.3261  3   0.006346 **
+# time:habitat  0.7516  3   0.860994   
+emtrends(bray_mod_site_baseline, pairwise ~ habitat , "time") # no sig differences
 bray_trends_emm_site_baseline <- emmip(bray_mod_site_baseline, habitat ~ time , at = list(time = seq(1,16,1)), plotit = F, CIs = T)
@@
 plot(residuals(bray_mod_site_macro_baseline) ~ fitted(bray_mod_site_macro_baseline) ) # strong negative trend
-
+Anova(bray_mod_site_macro_baseline)
+Response: dissimilarity_trans
+#                Chisq Df Pr(>Chisq)    
+# time          0.7391  1    0.38994    
+# habitat      56.1360  3  3.929e-12 ***
+# time:habitat 14.4254  3    0.00238 ** 
 emtrends(bray_mod_site_macro_baseline, pairwise ~ habitat , "time") # fringing slope significantly different from all others
@@
 
-full_columns_to_keep <- apply(full_site_species_matrix, 2, function(x) any(x >= asin(sqrt(0.1) )))
+full_columns_to_keep <- apply(full_site_species_matrix, 2, function(x) any(x >= asin(sqrt(0.05) )))
 full_abundant_site_species_matrix <- as.data.frame(full_site_species_matrix)[ , full_columns_to_keep]
@@
   geom_point(data= full.site.scores_site, aes(x=NMDS1,y=NMDS2, colour = habitat), alpha = 0.05) + # add the point markers
-  geom_text_repel(data= full.abundant.species.scores_site,aes(x=NMDS1,y=NMDS2, label = species), size = 4, alpha = 0.5) +  # add the species labels
+  geom_text_repel(data= full.abundant.species.scores_site,aes(x=NMDS1,y=NMDS2, label = species), size = 6, alpha = 0.7) +  # add the species labels
 
   geom_path(data = full.cent_site, aes(x = NMDS1, y = NMDS2, group = habitat, colour = habitat),
-            arrow = arrow(length=unit(0.25,"cm"), ends="last", type = "closed"), size = 1) +
+            arrow = arrow(length=unit(0.4,"cm"), ends="last", type = "closed"), size = 1) +
   geom_point(data = full.cent_site, aes(x = NMDS1, y = NMDS2, group = habitat, colour = habitat, shape = disturbance), size = 3, stroke = 1.4) +
@@
 
-full_columns_to_keep <- apply(full_site_macro_species_matrix, 2, function(x) any(x >= asin(sqrt(0.1) )))
+full_columns_to_keep <- apply(full_site_macro_species_matrix, 2, function(x) any(x >= asin(sqrt(0.05) )))
 full_abundant_site_macro_species_matrix <- as.data.frame(full_site_macro_species_matrix)[ , full_columns_to_keep]
@@
 ----------------- QUADRAT BELOW
-```{r fringing bray curtis}
+```{r quad bray curtis setup}
 fringing_species_matrix <- as.matrix(fringing[ , -which(names(fringing) %in% meta_cols)])
-fringing_species_matrix <- fringing_species_matrix/100 # convert from percent to proportion
+fringing_species_matrix <- fringing_species_matrix/fringing$hard_cover # convert from percent to proportion
 rownames(fringing_species_matrix) <- paste(fringing$year,"-", fringing$location, sep = "")
@@
 fringing_common_species <- names(fringing_species_percentage[fringing_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
-fringing_species_matrix <- fringing_species_matrix[,colnames(fringing_species_matrix) %in% fringing_common_species] # filter original species matrix
+# fringing_species_matrix <- fringing_species_matrix[,colnames(fringing_species_matrix) %in% fringing_common_species] # filter original species matrix
 
@@
 fringing_species_matrix <- fringing_species_matrix[!fringing_rows_to_remove,] 
+fringing_species_matrix <- na.omit(fringing_species_matrix)
 
@@
 
-```
-
-```{r backreef bray curtis}
+## backreef
 backreef_species_matrix <- as.matrix(backreef[ , -which(names(backreef) %in% meta_cols)])
-backreef_species_matrix <- backreef_species_matrix/100 # convert from percent to proportion
+backreef_species_matrix <- backreef_species_matrix/backreef$hard_cover # convert from percent to proportion
 rownames(backreef_species_matrix) <- paste(backreef$year,"-", backreef$location, sep = "")
@@
 backreef_common_species <- names(backreef_species_percentage[backreef_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
-backreef_species_matrix <- backreef_species_matrix[,colnames(backreef_species_matrix) %in% backreef_common_species] # filter original species matrix
+# backreef_species_matrix <- backreef_species_matrix[,colnames(backreef_species_matrix) %in% backreef_common_species] # filter original species matrix
 
@@
 backreef_species_matrix <- backreef_species_matrix[!backreef_rows_to_remove,] 
+backreef_species_matrix <- na.omit(backreef_species_matrix)
 
@@
 
-```
-
-```{r outer10 bray curtis}
+## Outer 10
 outer10_species_matrix <- as.matrix(outer10[ , -which(names(outer10) %in% meta_cols)])
-outer10_species_matrix <- outer10_species_matrix/100 # convert from percent to proportion
+outer10_species_matrix <- outer10_species_matrix/outer10$hard_cover # convert from percent to proportion
 rownames(outer10_species_matrix) <- paste(outer10$year,"-", outer10$location, sep = "")
@@
 outer10_common_species <- names(outer10_species_percentage[outer10_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
-outer10_species_matrix <- outer10_species_matrix[,colnames(outer10_species_matrix) %in% outer10_common_species] # filter original species matrix
+# outer10_species_matrix <- outer10_species_matrix[,colnames(outer10_species_matrix) %in% outer10_common_species] # filter original species matrix
 
@@
 outer10_species_matrix <- outer10_species_matrix[!outer10_rows_to_remove,] 
+outer10_species_matrix <- na.omit(outer10_species_matrix)
 
@@
 
-```
-
-```{r outer17 bray curtis}
+## OUTER 17
 outer17_species_matrix <- as.matrix(outer17[ , -which(names(outer17) %in% meta_cols)])
-outer17_species_matrix <- outer17_species_matrix/100 # convert from percent to proportion
+outer17_species_matrix <- outer17_species_matrix/outer17$hard_cover # convert from percent to proportion
 rownames(outer17_species_matrix) <- paste(outer17$year,"-", outer17$location, sep = "")
@@
 outer17_common_species <- names(outer17_species_percentage[outer17_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...
-outer17_species_matrix <- outer17_species_matrix[,colnames(outer17_species_matrix) %in% outer17_common_species] # filter original species matrix
+# outer17_species_matrix <- outer17_species_matrix[,colnames(outer17_species_matrix) %in% outer17_common_species] # filter original species matrix
 
@@
 outer17_species_matrix <- outer17_species_matrix[!outer17_rows_to_remove,] 
+outer17_species_matrix <- na.omit(outer17_species_matrix)
 
@@
          )
+
+```
+
+```{r quad level beta dispersion}
+full_species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
+full_species_matrix <- full_species_matrix/data$hard_cover # convert from percent to proportion
+rownames(full_species_matrix) <- paste(data$year,"-", data$location, sep = "")
+full_rows_to_remove <- rowSums(full_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
+full_species_matrix <- full_species_matrix[!full_rows_to_remove,] 
+full_species_matrix <- na.omit(full_species_matrix)
+full_species_matrix <- asin(sqrt(full_species_matrix)) # arcsin squareroot transform
+data$stratum <- paste(data$year,"-", data$location, sep = "") # for merging with bray matrix later
+full_subset <- data[data$stratum %in% rownames(full_species_matrix),] # make sure rows are also removed from dataframe
+
+brayfull <- vegdist(full_species_matrix, "bray") # create bray dissimilarity matrix
+# full_quad_betadisper <- vegan::betadisper(brayfull, full_subset$hab_year)
+full_subset$betadispersion <- full_quad_betadisper$distances[match(full_subset$stratum, names(full_quad_betadisper$distances))]
+
+full_subset$betadispersion_trans <- sv_trans(full_subset$betadispersion, full_subset)
+
+betadisper_quad_mod <- glmmTMB(betadispersion_trans ~ year*habitat + (1|site), family = beta_family(link = "logit"), data = full_subset)
+summary(betadisper_quad_mod) # fair bit of within site variance - inv_logit(0.02) = 0.5
+hist(residuals(betadisper_quad_mod)) # positive skew
+plot(residuals(betadisper_quad_mod) ~ fitted(betadisper_quad_mod)) # slight heteroskedasticity
+Anova(betadisper_quad_mod)
+#                Chisq Df Pr(>Chisq)    
+# year           0.3156  1     0.5743    
+# habitat      333.7035  3     <2e-16 ***
+# year:habitat 720.3054  3     <2e-16 ***
+
+macro_species_matrix <- as.matrix(macroalgae[ , -which(names(macroalgae) %in% meta_cols)])
+macro_species_matrix <- macro_species_matrix/macroalgae$hard_cover # convert from percent to proportion
+rownames(macro_species_matrix) <- paste(macroalgae$year,"-", macroalgae$location, sep = "")
+macro_rows_to_remove <- rowSums(macro_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
+macro_species_matrix <- macro_species_matrix[!macro_rows_to_remove,] 
+macro_species_matrix <- na.omit(macro_species_matrix)
+macro_species_matrix <- asin(sqrt(macro_species_matrix)) # arcsin squareroot transform
+macroalgae$stratum <- paste(macroalgae$year,"-", macroalgae$location, sep = "") # for merging with bray matrix later
+macro_subset <- macroalgae[macroalgae$stratum %in% rownames(macro_species_matrix),] # make sure rows are also removed from macroalgaeframe
+
+braymacro <- vegdist(macro_species_matrix, "bray") # create bray dissimilarity matrix
+# macro_quad_betadisper <- vegan::betadisper(braymacro, macro_subset$hab_year)
+macro_subset$betadispersion <- macro_quad_betadisper$distances[match(macro_subset$stratum, names(macro_quad_betadisper$distances))]
 
@@
 
-full_species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
-full_species_matrix <- full_species_matrix/100 # convert from percent to proportion
-rownames(full_species_matrix) <- paste(data$year,"-", data$location, sep = "")
-
-full_species_matrix <- full_species_matrix[rownames(full_species_matrix) %in% relevant_quads,] # subset so only quads that didn't contain rare species are represented
-full_species_matrix <- asin(sqrt(full_species_matrix)) # arcsin squareroot transform
-data$stratum <- paste(data$year,"-", data$location, sep = "") # for merging with bray matrix later
-full_subset <- data[data$stratum %in% relevant_quads,] # make sure rows are also removed from dataframe
-
-brayfull <- vegdist(full_species_matrix, "bray") # create bray dissimilarity matrix
-# View(brayfull)
 full_braymat <- as.matrix(brayfull) # conver to matrix object

### 2024-09-06 noam_beta_diversity.Rmd
finished quad level beta dispersion models and figs

@@
 
-```{r quad level beta dispersion}
+```{r quad level beta dispersion, fig.height = 8, fig.width = 12}
 full_species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
@@
 data$stratum <- paste(data$year,"-", data$location, sep = "") # for merging with bray matrix later
-full_subset <- data[data$stratum %in% rownames(full_species_matrix),] # make sure rows are also removed from dataframe
-
-brayfull <- vegdist(full_species_matrix, "bray") # create bray dissimilarity matrix
-# full_quad_betadisper <- vegan::betadisper(brayfull, full_subset$hab_year)
-full_subset$betadispersion <- full_quad_betadisper$distances[match(full_subset$stratum, names(full_quad_betadisper$distances))]
-
-full_subset$betadispersion_trans <- sv_trans(full_subset$betadispersion, full_subset)
+# full_subset <- data[data$stratum %in% rownames(full_species_matrix),] # make sure rows are also removed from dataframe
+# 
+# brayfull <- vegdist(full_species_matrix, "bray") # create bray dissimilarity matrix
+# # full_quad_betadisper <- vegan::betadisper(brayfull, full_subset$hab_year)
+# full_subset$betadispersion <- full_quad_betadisper$distances[match(full_subset$stratum, names(full_quad_betadisper$distances))]
+# full_subset$betadispersion_trans <- sv_trans(full_subset$betadispersion, full_subset)
+# 
+# write_csv(full_subset, here("data", "quad_level_all_species_with_betadisper.csv"))
+
+full_subset <- read.csv(here("data", "quad_level_all_species_with_betadisper.csv")) # read back in data from csv so you don't have to re-run
 
@@
 
+emtrends(betadisper_quad_mod, pairwise ~ habitat , "year") # all different
+betadisper_quad_emm <- emmip(betadisper_quad_mod, habitat ~ year , at = list(year = seq(2007,2023,1)), plotit = F, CIs = T)
+
+ggplot()+
+   geom_point(data = full_subset, aes(x = year, y = betadispersion_trans, colour = habitat), alpha = 0.1 ) +
+   geom_line(data = betadisper_quad_emm, aes(x = year, y = inv_logit(yvar), colour = habitat), linewidth = 2) +
+   geom_ribbon(data = betadisper_quad_emm, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+   scale_fill_manual(values = habitat_colours) +
+   geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+   geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+   geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+   labs(y = "Beta dispersion\n", x = "Year") +
+   theme_bw() + 
+   theme(panel.grid.major = element_blank(), 
+         panel.grid.minor = element_blank(),
+         panel.background = element_blank(),
+         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+         axis.line.y = element_line(colour = "black"),
+         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         axis.line.x = element_line(colour = "black"),
+         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         strip.text.x = element_text(size = 32),
+         legend.key.size = unit(2, "cm"), # increase legend size
+         legend.text = element_text(size = rel(2)), # increase legend text
+         legend.title = element_text(size = rel(2.25)),
+         legend.position = 'bottom', # move legend to bottom
+         legend.box = 'horizontal',
+         legend.box.background = element_rect(colour = "black") )# make legend horizontal
+
 macro_species_matrix <- as.matrix(macroalgae[ , -which(names(macroalgae) %in% meta_cols)])
@@
 macroalgae$stratum <- paste(macroalgae$year,"-", macroalgae$location, sep = "") # for merging with bray matrix later
-macro_subset <- macroalgae[macroalgae$stratum %in% rownames(macro_species_matrix),] # make sure rows are also removed from macroalgaeframe
-
-braymacro <- vegdist(macro_species_matrix, "bray") # create bray dissimilarity matrix
-# macro_quad_betadisper <- vegan::betadisper(braymacro, macro_subset$hab_year)
-macro_subset$betadispersion <- macro_quad_betadisper$distances[match(macro_subset$stratum, names(macro_quad_betadisper$distances))]
-
+# macro_subset <- macroalgae[macroalgae$stratum %in% rownames(macro_species_matrix),] # make sure rows are also removed from macroalgaeframe
+# 
+# braymacro <- vegdist(macro_species_matrix, "bray") # create bray dissimilarity matrix
+# # macro_quad_betadisper <- vegan::betadisper(braymacro, macro_subset$hab_year)
+# macro_subset$betadispersion <- macro_quad_betadisper$distances[match(macro_subset$stratum, names(macro_quad_betadisper$distances))]
+# macro_subset$betadispersion_trans <- sv_trans(macro_subset$betadispersion, macro_subset)
+# write_csv(macro_subset, here("data", "quad_level_just_macroalgae_with_betadisper.csv"))
+macro_subset <- read.csv(here("data", "quad_level_just_macroalgae_with_betadisper.csv"))
+
+betadisper_macro_quad_mod <- glmmTMB(betadispersion_trans ~ year*habitat + (1|site), family = beta_family(link = "logit"), data = macro_subset)
+summary(betadisper_macro_quad_mod) # fair bit of within site variance - inv_logit(0.02) = 0.5
+qqnorm(residuals(betadisper_macro_quad_mod)) # a little lumpy but not bad
+plot(residuals(betadisper_macro_quad_mod) ~ fitted(betadisper_macro_quad_mod)) # negative trend
+Anova(betadisper_macro_quad_mod)
+
+#                 Chisq Df Pr(>Chisq)    
+# year           97.391  1  < 2.2e-16 ***
+# habitat      1973.896  3  < 2.2e-16 ***
+# year:habitat  511.722  3  < 2.2e-16 ***
+
+emtrends(betadisper_macro_quad_mod, pairwise ~ habitat , "year") # backreef, outer10, and fringing don't differ
+betadisper_quad_macro_emm <- emmip(betadisper_macro_quad_mod, habitat ~ year , at = list(year = seq(2007,2023,1)), plotit = F, CIs = T)
+
+ggplot()+
+   geom_point(data = macro_subset, aes(x = year, y = betadispersion_trans, colour = habitat), alpha = 0.1 ) +
+   geom_line(data = betadisper_quad_macro_emm, aes(x = year, y = inv_logit(yvar), colour = habitat), linewidth = 2) +
+   geom_ribbon(data = betadisper_quad_macro_emm, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+   scale_fill_manual(values = habitat_colours) +
+   geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+   geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+   geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+   labs(y = "Beta dispersion\n", x = "Year") +
+   theme_bw() + 
+   theme(panel.grid.major = element_blank(), 
+         panel.grid.minor = element_blank(),
+         panel.background = element_blank(),
+         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+         axis.line.y = element_line(colour = "black"),
+         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         axis.line.x = element_line(colour = "black"),
+         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         strip.text.x = element_text(size = 32),
+         legend.key.size = unit(2, "cm"), # increase legend size
+         legend.text = element_text(size = rel(2)), # increase legend text
+         legend.title = element_text(size = rel(2.25)),
+         legend.position = 'bottom', # move legend to bottom
+         legend.box = 'horizontal',
+         legend.box.background = element_rect(colour = "black") )# make legend horizontal
 ```

### 2024-09-06 noam_final_analyses.Rmd
potential fix for error lauren caught in allalgae metric

@@
 
-meta_df$all_algae_cover <- rowSums(data[startAlg:endAlg]) # percent cover of all algae
+meta_df$all_algae_cover <- rowSums(data[grep("acanthophora_spicifera", colnames(data)):
+                                        grep("valonia_ventricosa", colnames(data))]) # percent cover of all algae
 meta_df$all_algae_cover_hard <- case_when(
@@
 alpha_diversity_site_macro$richness_area <- alpha_diversity_site_macro$richness/(alpha_diversity_site_macro$hard_cover/100)
+
+```
+
+```{r beta dispersion}
+full_site_species_matrix <- as.matrix(site_full[ , -which(names(site_full) %in% meta_cols)])
+full_site_species_matrix <- full_site_species_matrix/site_full$hard_cover # convert from percent to proportion of hard cover
+rownames(full_site_species_matrix) <- paste(site_full$year,"-", site_full$site_habitat, sep = "")
+
+full_site_species_matrix <- asin(sqrt(full_site_species_matrix)) # arcsin squareroot transform
+
+full_site_macro_species_matrix <- as.matrix(site_macro[ , -which(names(site_macro) %in% colnames(site_meta_df))])
+full_site_macro_species_matrix <- full_site_macro_species_matrix/site_macro$hard_cover # convert from percent to proportion of hard cover
+rownames(full_site_macro_species_matrix) <- paste(site_macro$year,"-", site_macro$site_habitat, sep = "")
+
+full_site_macro_species_matrix <- asin(sqrt(full_site_macro_species_matrix)) # arcsin squareroot transform
+
+full_site_bray <- vegdist(full_site_species_matrix, "bray") 
+full_site_macro_bray <- vegdist(full_site_macro_species_matrix, "bray")
+
+full_site_betadisper <- vegan::betadisper(full_site_bray, site_full$hab_year)
+site_full$betadispersion <- full_site_betadisper$distances[match(site_full$stratum, names(full_site_betadisper$distances))]
+site_full$betadispersion_trans <- sv_trans(site_full$betadispersion, site_full)
+
+betadisper_site_mod <- glmmTMB(betadispersion_trans ~ year*habitat + (1|site), family = beta_family(link = "logit"), data = site_full)
+summary(betadisper_site_mod) # fair bit of within site variance - inv_logit(0.04) = 0.5
+hist(residuals(betadisper_site_mod)) # pretty good
+plot(residuals(betadisper_site_mod) ~ fitted(betadisper_site_mod)) # slight heteroskedasticity
+Anova(betadisper_site_mod)
+#                Chisq Df Pr(>Chisq)    
+# year         12.9561  1  0.0003189 ***
+# habitat      27.2191  3  5.296e-06 ***
+# year:habitat  6.8472  3  0.0769311 .
+
+emtrends(betadisper_site_mod, pairwise ~ habitat , "year") # fringing slope significantly different from outer17
+betadisper_site_emm <- emmip(betadisper_site_mod, habitat ~ year , at = list(year = seq(2007,2023,1)), plotit = F, CIs = T)
+
+ggplot()+
+   geom_point(data = site_full, aes(x = year, y = betadispersion_trans, colour = habitat), alpha = 0.1 ) +
+   geom_line(data = betadisper_site_emm, aes(x = year, y = inv_logit(yvar), colour = habitat), linewidth = 2) +
+   geom_ribbon(data = betadisper_site_emm, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+   scale_fill_manual(values = habitat_colours) +
+   geom_segment(aes(y = 0.6, yend = 0.6, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+   geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+   geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+   labs(y = "Beta dispersion\n", x = "Year") +
+   theme_bw() + 
+   theme(panel.grid.major = element_blank(), 
+         panel.grid.minor = element_blank(),
+         panel.background = element_blank(),
+         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+         axis.line.y = element_line(colour = "black"),
+         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         axis.line.x = element_line(colour = "black"),
+         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         strip.text.x = element_text(size = 32),
+         legend.key.size = unit(2, "cm"), # increase legend size
+         legend.text = element_text(size = rel(2)), # increase legend text
+         legend.title = element_text(size = rel(2.25)),
+         legend.position = 'bottom', # move legend to bottom
+         legend.box = 'horizontal',
+         legend.box.background = element_rect(colour = "black") )# make legend horizontal
+
+full_site_macro_betadisper <- vegan::betadisper(full_site_macro_bray, site_macro$hab_year)
+full_subset_site_macro$betadispersion <- full_site_macro_betadisper$distances[match(full_subset_site_macro$stratum, names(full_site_macro_betadisper$distances))]
+full_subset_site_macro$betadispersion_trans <- sv_trans(full_subset_site_macro$betadispersion, full_subset_site_macro)
+
+betadisper_site_macro_mod <- glmmTMB(betadispersion_trans ~ year*habitat + (1|site), family = beta_family(link = "logit"), data = full_subset_site_macro)
+summary(betadisper_site_macro_mod) # similar site variance - inv_logit(0.01) = 0.5
+hist(residuals(betadisper_site_macro_mod)) # pretty good
+plot(residuals(betadisper_site_macro_mod) ~ fitted(betadisper_site_macro_mod)) # slight heteroskedasticity and negative trend
+Anova(betadisper_site_macro_mod)
+#               Chisq Df Pr(>Chisq)    
+# year         36.591  1  1.457e-09 ***
+# habitat      46.029  3  5.590e-10 ***
+# year:habitat 31.232  3  7.597e-07 ***
+
+emtrends(betadisper_site_macro_mod, pairwise ~ habitat , "year") # backreef distinct from all other habitats
+betadisper_site_macro_emm <- emmip(betadisper_site_macro_mod, habitat ~ year , at = list(year = seq(2007,2023,1)), plotit = F, CIs = T)
+
+ggplot()+
+   geom_point(data = site_full, aes(x = year, y = betadispersion_trans, colour = habitat), alpha = 0.1 ) +
+   geom_line(data = betadisper_site_macro_emm, aes(x = year, y = inv_logit(yvar), colour = habitat), linewidth = 2) +
+   geom_ribbon(data = betadisper_site_macro_emm, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+   scale_fill_manual(values = habitat_colours) +
+   geom_segment(aes(y = 0.75, yend = 0.75, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+   geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+   geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+   labs(y = "Beta dispersion\n", x = "Year") +
+   theme_bw() + 
+   theme(panel.grid.major = element_blank(), 
+         panel.grid.minor = element_blank(),
+         panel.background = element_blank(),
+         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
+         axis.line.y = element_line(colour = "black"),
+         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
+         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         axis.line.x = element_line(colour = "black"),
+         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
+         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
+         strip.text.x = element_text(size = 32),
+         legend.key.size = unit(2, "cm"), # increase legend size
+         legend.text = element_text(size = rel(2)), # increase legend text
+         legend.title = element_text(size = rel(2.25)),
+         legend.position = 'bottom', # move legend to bottom
+         legend.box = 'horizontal',
+         legend.box.background = element_rect(colour = "black") )# make legend horizontal
 
@@
 ```
-```{r stepwise models site level}
+
+```{r stepwise models site level, fig.height = 8, fig.width=12}
 
@@
 plot(residuals(fringing_all_shannon_top_mod) ~ fitted(fringing_all_shannon_top_mod)) # slight patterning
+summary(fringing_all_shannon_top_mod)
+
+fringing_emm <-as.data.frame(emmeans(fringing_all_shannon_top_mod, ~mean_n,
+                                     at = list(mean_n = seq(min(fringing_all_site$mean_n),
+                                                            max(fringing_all_site$mean_n),
+                                                            0.01)))) 
+emmip(fringing_all_shannon_top_mod, 1~ mean_n)
+ggplot()+
+  geom_point(data = fringing_all_site, aes(x = mean_n, y = shannon))+
+  geom_line(data = fringing_emm, aes(x = mean_n, y = emmean)) +
+  geom_ribbon(data = fringing_emm, aes(x = mean_n, ymin = lower.CL, ymax = upper.CL, alpha = 0.2), show.legend = F) +
+  scale_y_log10() + 
+  labs(x = "Mean N", y = "Fringing Shannon Index (H')") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
 
@@
 
+outer10_emm <-as.data.frame(emmeans(outer10_all_shannon_top_mod, ~herbivore_evenness,
+                                     at = list(herbivore_evenness = seq(min(outer10_all_site$herbivore_evenness),
+                                                            max(outer10_all_site$herbivore_evenness),
+                                                            0.05))))
+
+ggplot()+
+  geom_point(data = outer10_all_site, aes(x = herbivore_evenness, y = shannon))+
+  geom_line(data = outer10_emm, aes(x = herbivore_evenness, y = emmean)) +
+  geom_ribbon(data = outer10_emm, aes(x = herbivore_evenness, ymin = lower.CL, ymax = upper.CL, alpha = 0.2), show.legend = F) +
+  scale_y_log10() + 
+  labs(x = "Herbivore evenness", y = "Outer 10 Shannon Index (H')") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
 ## outer17 - no nitrogen or herbivores
@@
 
+backreef_emm <-as.data.frame(emmeans(backreef_macro_shannon_top_mod, ~wave_power_adcp,
+                                     at = list(wave_power_adcp = seq(min(backreef_macro_site$wave_power_adcp),
+                                                            max(backreef_macro_site$wave_power_adcp),
+                                                            1000)))) 
+ggplot()+
+  geom_point(data = backreef_macro_site, aes(x = wave_power_adcp, y = shannon))+
+  geom_line(data = backreef_emm, aes(x = wave_power_adcp, y = emmean)) +
+  geom_ribbon(data = backreef_emm, aes(x = wave_power_adcp, ymin = lower.CL, ymax = upper.CL, alpha = 0.2), show.legend = F) +
+  scale_y_log10() + 
+  labs(x = "Wave power", y = "Backreef Shannon Index (H')") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
 ## outer10 - no nitrogen 
@@
 outer17_macro_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp  + mean_temp_c + (1|site), data = outer17_macro_site, na.action = na.pass)
-outer17_macro_shannon_dredge <- dredge(outer17_macro_shannon_mod_universal) # some convergence issues
+outer17_macro_shannon_dredge <- dredge(outer17_macro_shannon_mod_universal) 
 model.sel(outer17_macro_shannon_dredge) #  best includes all terms!

### 2024-09-10 julianna_figs_Aug29.Rmd
Want to get up to date for workshop

@@
 alltime %>% 
-  ggplot(aes(x = coral_cover_hard_mean, y = algae_cover_hard_cv)) +
+  ggplot(aes(x = coral_cover_hard_mean, y = macroalgae_cover_hard_cv)) +
   geom_point(aes(color = island_side)) +
@@
 alltime %>% 
-  ggplot(aes(x = coral_cover_hard_cv, y = algae_cover_hard_cv)) +
+  ggplot(aes(x = coral_cover_hard_cv, y = macroalgae_cover_hard_cv)) +
   geom_point(aes(color = island_side)) +

### 2024-09-10 noam_final_analyses.Rmd

@@
 ```{r beta dispersion}
-full_site_species_matrix <- as.matrix(site_full[ , -which(names(site_full) %in% meta_cols)])
+site_full$stratum <- paste(site_full$year,"-", site_full$site_habitat, sep = "") # for merge
+site_macro$stratum <- paste(site_macro$year,"-", site_macro$site_habitat, sep = "") # for merge
+
+full_site_species_matrix <- as.matrix(site_full[ , -which(names(site_full) %in% c(meta_cols, "stratum" ))])
 full_site_species_matrix <- full_site_species_matrix/site_full$hard_cover # convert from percent to proportion of hard cover
@@
 
-full_site_macro_species_matrix <- as.matrix(site_macro[ , -which(names(site_macro) %in% colnames(site_meta_df))])
+full_site_macro_species_matrix <- as.matrix(site_macro[ , -which(names(site_macro) %in% c(meta_cols, "stratum" ))])
 full_site_macro_species_matrix <- full_site_macro_species_matrix/site_macro$hard_cover # convert from percent to proportion of hard cover
@@
 full_site_macro_species_matrix <- asin(sqrt(full_site_macro_species_matrix)) # arcsin squareroot transform
-
+# some rows are all 0s
+full_site_macro_rows_to_remove <- rowSums(full_site_macro_species_matrix) == 0
+full_site_macro_species_matrix <- full_site_macro_species_matrix[!full_site_macro_rows_to_remove,] 
 full_site_bray <- vegdist(full_site_species_matrix, "bray") 
@@
 site_full$betadispersion <- full_site_betadisper$distances[match(site_full$stratum, names(full_site_betadisper$distances))]
-site_full$betadispersion_trans <- sv_trans(site_full$betadispersion, site_full)
+site_full$betadispersion_trans <- sv_trans(site_full$betadispersion)
 
@@
 
-full_site_macro_betadisper <- vegan::betadisper(full_site_macro_bray, site_macro$hab_year)
+full_subset_site_macro <- site_macro[site_macro$stratum %in% rownames(full_site_macro_species_matrix),]
+full_site_macro_betadisper <- vegan::betadisper(full_site_macro_bray, site_macro_subset$hab_year)
 full_subset_site_macro$betadispersion <- full_site_macro_betadisper$distances[match(full_subset_site_macro$stratum, names(full_site_macro_betadisper$distances))]
-full_subset_site_macro$betadispersion_trans <- sv_trans(full_subset_site_macro$betadispersion, full_subset_site_macro)
+full_subset_site_macro$betadispersion_trans <- sv_trans(full_subset_site_macro$betadispersion)
 
@@
 
+# add beta dispersion
+alpha_diversity_site <- full_join(alpha_diversity_site, site_full[,c("betadispersion_trans", "site_hab_year")], by = "site_hab_year")
+alpha_diversity_site$year_factor <- as.factor(as.character(alpha_diversity_site$year))
+
 alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro, nutrients, by = c("year", "site_hab_year", "site", "habitat"))
@@
 alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
-pairs(data = alpha_diversity_site_macro, ~ richness + shannon + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)
+
+alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro, full_subset_site_macro[,c("betadispersion_trans", "site_hab_year")], by = "site_hab_year")
+alpha_diversity_site_macro$year_factor <- as.factor(as.character(alpha_diversity_site_macro$year))
 
@@
 alpha_diversity_quad <- full_join(alpha_diversity_quad, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
-pairs(data = alpha_diversity_quad, ~ richness + shannon + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)
+alpha_diversity_quad$year_factor <- as.factor(as.character(alpha_diversity_quad$year))
 
@@
 alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
-pairs(data = alpha_diversity_quad_macro, ~ richness + shannon + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)
+alpha_diversity_quad_macro$year_factor <- as.factor(as.character(alpha_diversity_quad_macro$year))
+
+```
+
+```{r all algae cover mods}
+site_full_cover_mod <- glmmTMB(sv_trans( all_algae_cover_hard/100) ~ year*habitat + (1|site) , family = beta_family(),
+                             data = alpha_diversity_site) 
+# Warning in (function (start, objective, gradient = NULL, hessian = NULL,  : NA/NaN function evaluation
+# This warning occurs when the optimizer visits a region of parameter space that is invalid.
+It is not a problem as long as the optimizer has left that region of parameter space upon convergence, which is indicated by an absence of the model convergence warnings described above.
+summary(site_full_cover_mod) # decent variance in random effect
+Anova(site_full_cover_mod) 
+#                Chisq Df Pr(>Chisq)    
+# year          0.9983  1     0.3177    
+# habitat      43.0233  3  2.433e-09 ***
+# year:habitat 49.8146  3  8.750e-11 ***
+# habitat more important than time when time is numeric
+# check_mod
+hist(residuals(site_full_cover_mod)) # negative skew
+
+plot(residuals(site_full_cover_mod) ~ fitted(site_full_cover_mod)) # negative trend
+
+emtrends(site_full_cover_mod, pairwise ~  habitat, var = "year")
+# backreef not different from fringing, outer10 not different from outer17
+site_full_cover_emm <- emmip(site_full_cover_mod,habitat ~ year,
+                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)
+
+(site_base_cover_plot <- 
+  ggplot() +
+  geom_point(data = alpha_diversity_site,
+             aes(x = year, y = all_algae_cover_hard/100, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = site_full_cover_emm, aes(x = year, y = inv_logit(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = site_full_cover_emm, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
+  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+    labs(x = "Year", y = "Cover (prop)") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) ) # add box around legend
+
+site_full_cover_mod_factor <- glmmTMB(sv_trans( all_algae_cover_hard/100) ~ year_factor*habitat + 
+                                        (1|site) + ar1(year_factor + 0 | site),
+                                  family = beta_family(), data = alpha_diversity_site)
+
+summary(site_full_cover_mod_factor) # low variance in random effect but strong temporal autocorrelation
+Anova(site_full_cover_mod_factor) 
+#                 Chisq Df Pr(>Chisq)    
+# year_factor          93.081 16  6.745e-13 ***
+# habitat             113.174  3  < 2.2e-16 ***
+# year_factor:habitat  63.853 46    0.04171 *  
+hist(residuals(site_full_cover_mod_factor)) # still neg skew but not as bad
+
+plot(residuals(site_full_cover_mod_factor) ~ fitted(site_full_cover_mod_factor)) # strong patterning
+
+site_full_cover_factor_emm <- plot(emmeans(site_full_cover_mod_factor, ~  habitat*year_factor), plotit = F) 
+site_full_cover_factor_emm$year <- as.numeric(levels(site_full_cover_factor_emm$year_factor)) # for plotting
+
+(site_factor_cover_plot <- 
+  ggplot() +
+  geom_point(data = site_full_cover_factor_emm,
+             aes(x = year, y = inv_logit(the.emmean), colour = habitat), size = 2, position = position_dodge(0.4)) +
+  geom_line(data = site_full_cover_factor_emm, aes(x = year, y = inv_logit(the.emmean), colour = habitat, group = habitat),
+            position = position_dodge(0.4)) +
+  geom_errorbar(data = site_full_cover_factor_emm,
+                aes(x = year, ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), colour = habitat), width = 0.2,
+                position = position_dodge(0.4)) +
+  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+    labs(x = "Year", y = "cover") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) ) # add box around legend
+
+```
+
+```{r macroalgaealgae cover mods}
+site_macrofull_cover_mod <- glmmTMB(sv_trans( macroalgae_cover_hard/100) ~ year*habitat + (1|site) , family = beta_family(),
+                             data = alpha_diversity_site_macro) 
+# Warning in (function (start, objective, gradient = NULL, hessian = NULL,  : NA/NaN function evaluation
+# This warning occurs when the optimizer visits a region of parameter space that is invalid.
+It is not a problem as long as the optimizer has left that region of parameter space upon convergence, which is indicated by an absence of the model convergence warnings described above.
+summary(site_macrofull_cover_mod) # less variance in random effect than before
+Anova(site_macrofull_cover_mod) 
+#                Chisq Df Pr(>Chisq)    
+# year          3.9366  1   0.047247 *  
+# habitat      32.7335  3  3.666e-07 ***
+# year:habitat 13.0316  3   0.004569 ** 
+hist(residuals(site_macrofull_cover_mod)) # positive skew
+
+plot(residuals(site_macrofull_cover_mod) ~ fitted(site_macrofull_cover_mod)) # trend likely due to beta reg
+
+emtrends(site_macrofull_cover_mod, pairwise ~  habitat, var = "year")
+# backreef different from outer10
+site_macrofull_cover_emm <- emmip(site_macrofull_cover_mod,habitat ~ year,
+                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)
+
+(site_macrobase_cover_plot <- 
+  ggplot() +
+  geom_point(data = alpha_diversity_site_macro,
+             aes(x = year, y = macroalgae_cover_hard/100, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = site_macrofull_cover_emm, aes(x = year, y = inv_logit(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = site_macrofull_cover_emm, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
+  geom_segment(aes(y = 0.7, yend = 0.7, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+    labs(x = "Year", y = "Cover (prop)") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) ) # add box around legend
+
+site_macrofull_cover_mod_factor <- glmmTMB(sv_trans( macroalgae_cover_hard/100) ~ year_factor*habitat + 
+                                        (1|site) + ar1(year_factor + 0 | site),
+                                  family = beta_family(), data = alpha_diversity_site_macro)
+
+summary(site_macrofull_cover_mod_factor) # low variance in random effect but strong temporal autocorrelation
+Anova(site_macrofull_cover_mod_factor) 
+#                 Chisq Df Pr(>Chisq)    
+# year_factor         35.931 16   0.002958 ** 
+# habitat             36.929  3  4.764e-08 ***
+# year_factor:habitat 98.623 46  1.056e-05 ***
+hist(residuals(site_macrofull_cover_mod_factor)) # still pos skew but not as bad
+
+plot(residuals(site_macrofull_cover_mod_factor) ~ fitted(site_macrofull_cover_mod_factor)) # similar to numeric
+
+site_macrofull_cover_factor_emm <- plot(emmeans(site_macrofull_cover_mod_factor, ~  habitat*year_factor), plotit = F) 
+site_macrofull_cover_factor_emm$year <- as.numeric(levels(site_macrofull_cover_factor_emm$year_factor)) # for plotting
+
+(site_macrofactor_cover_plot <- 
+  ggplot() +
+  geom_point(data = site_macrofull_cover_factor_emm,
+             aes(x = year, y = inv_logit(the.emmean), colour = habitat), size = 2, position = position_dodge(0.4)) +
+  geom_line(data = site_macrofull_cover_factor_emm, aes(x = year, y = inv_logit(the.emmean), colour = habitat, group = habitat),
+            position = position_dodge(0.4)) +
+  geom_errorbar(data = site_macrofull_cover_factor_emm,
+                aes(x = year, ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), colour = habitat), width = 0.2,
+                position = position_dodge(0.4)) +
+  geom_segment(aes(y = 0.5, yend = 0.5, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+    labs(x = "Year", y = "cover") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) ) # add box around legend
 
@@
 
+```{r all algae alpha diversity site level year factor}
+
+site_full_richness_mod_factor <- glmmTMB(richness ~ year_factor*habitat + (1|site) + ar1(year_factor + 0 | site) + offset(log(hard_cover)),
+                                  family = poisson, data = alpha_diversity_site)
+
+summary(site_full_richness_mod_factor) # low variance in random effect but strong temporal autocorrelation
+Anova(site_full_richness_mod_factor) 
+#                 Chisq Df Pr(>Chisq)    
+# year_factor          93.081 16  6.745e-13 ***
+# habitat             113.174  3  < 2.2e-16 ***
+# year_factor:habitat  63.853 46    0.04171 *  
+hist(residuals(site_full_richness_mod_factor)) # looks great
+
+# striping/patterining expected because of discrete response var.
+plot(residuals(site_full_richness_mod_factor) ~ fitted(site_full_richness_mod_factor))
+
+site_full_richness_factor_emm <- plot(emmeans(site_full_richness_mod_factor, ~  habitat*year_factor), plotit = F) 
+site_full_richness_factor_emm$year <- as.numeric(levels(site_full_richness_factor_emm$year_factor)) # for plotting
+
+(site_factor_richness_plot <- 
+  ggplot() +
+  geom_point(data = site_full_richness_factor_emm,
+             aes(x = year, y = exp(the.emmean), colour = habitat), size = 2, position = position_dodge(0.4)) +
+  geom_line(data = site_full_richness_factor_emm, aes(x = year, y = exp(the.emmean), colour = habitat, group = habitat),
+            position = position_dodge(0.4)) +
+  geom_errorbar(data = site_full_richness_factor_emm,
+                aes(x = year, ymin = exp(asymp.LCL), ymax = exp(asymp.UCL), colour = habitat), width = 0.2,
+                position = position_dodge(0.4)) +
+  geom_segment(aes(y = 22, yend = 22, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  scale_y_log10() +
+    labs(x = "Year", y = "Richness") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) ) # add box around legend
+
+site_full_shannon_mod_factor <- glmmTMB(log(shannon+1) ~ year_factor*habitat + (1|site) + ar1(year_factor + 0 | site),
+                                        data = alpha_diversity_site) # model convergence - goes away if you drop site random effect
+# results are identical between including RE and not
+
+summary(site_full_shannon_mod_factor) # low variance in random effect but strong temporal autocorrelation
+Anova(site_full_shannon_mod_factor) 
+#                 Chisq Df Pr(>Chisq)    
+# year_factor         129.09 16  < 2.2e-16 ***
+# habitat             156.55  3  < 2.2e-16 ***
+# year_factor:habitat 152.84 46  2.162e-13 ***
+hist(residuals(site_full_shannon_mod_factor)) # a few points drive slight negative skew but good on the whole
+plot(residuals(site_full_shannon_mod_factor) ~ fitted(site_full_shannon_mod_factor))# looks good
+
+site_full_shannon_factor_emm <- plot(emmeans(site_full_shannon_mod_factor, ~  habitat*year_factor), plotit = F) 
+site_full_shannon_factor_emm$year <- as.numeric(levels(site_full_shannon_factor_emm$year_factor)) # for plotting
+
+(site_factor_shannon_plot <- 
+  ggplot() +
+  geom_point(data = site_full_shannon_factor_emm,
+             aes(x = year, y = exp(the.emmean), colour = habitat), size = 2, position = position_dodge(0.4)) +
+  geom_line(data = site_full_shannon_factor_emm, aes(x = year, y = exp(the.emmean), colour = habitat, group = habitat),
+            position = position_dodge(0.4)) +
+  geom_errorbar(data = site_full_shannon_factor_emm,
+                aes(x = year, ymin = exp(lower.CL), ymax = exp(upper.CL), colour = habitat), width = 0.2,
+                position = position_dodge(0.4)) +
+  geom_segment(aes(y = 3, yend = 3, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  scale_y_log10() +
+    labs(x = "Year", y = "Shannon Index (H')") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) ) # add box around legend
+
+```
+
 ```{r macroalgae alpha diversity site level}
@@
 site_macro_shannon_mod <- glmmTMB(log(shannon +1 ) ~ year*habitat + (1|site),
-                             data = alpha_diversity_site) 
+                             data = alpha_diversity_site_macro) 
 # sites with shannon div = 0 so log+1 transformed
@@
 #                Chisq Df Pr(>Chisq)    
-# year          0.0133  1   0.908348    
-# habitat      97.6274  3  < 2.2e-16 ***
-# year:habitat 14.0548  3   0.002831 ** 
-hist(residuals(site_macro_shannon_mod)) # looks good
+# year           6.7643  1    0.00930 ** 
+# habitat      104.2995  3    < 2e-16 ***
+# year:habitat   9.1657  3    0.02717 *  
+qqnorm(residuals(site_macro_shannon_mod)) # slight negative skew
 plot(residuals(site_macro_shannon_mod) ~ fitted(site_macro_shannon_mod)) # generally fine
@@
   geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_fill_manual(values = habitat_colours) +
+  scale_y_log10() +
+    labs(x = "Year", y = "Shannon Index (H')") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) ) # add box around legend
+
+```
+
+```{r macroalgae alpha diversity site level year factor}
+
+site_macrofull_richness_mod_factor <- glmmTMB(richness ~ year_factor*habitat + (1|site) + ar1(year_factor + 0 | site) +
+                                              offset(log(hard_cover)), family = poisson, data = alpha_diversity_site_macro)
+
+summary(site_macrofull_richness_mod_factor) # slightly lower temporal autocorrelation than before but still high
+Anova(site_macrofull_richness_mod_factor) 
+#                 Chisq Df Pr(>Chisq)    
+# year_factor         80.475 16  1.366e-10 ***
+# habitat             92.482  3  < 2.2e-16 ***
+# year_factor:habitat 54.467 46     0.1835    
+qqnorm(residuals(site_macrofull_richness_mod_factor)) # some positive skew
+
+# striping/patterining expected because of discrete response var.
+plot(residuals(site_macrofull_richness_mod_factor) ~ fitted(site_macrofull_richness_mod_factor))
+
+site_macrofull_richness_factor_emm <- plot(emmeans(site_macrofull_richness_mod_factor, ~  habitat*year_factor), plotit = F) 
+site_macrofull_richness_factor_emm$year <- as.numeric(levels(site_macrofull_richness_factor_emm$year_factor)) # for plotting
+
+(site_macrofactor_richness_plot <- 
+  ggplot() +
+  geom_point(data = site_macrofull_richness_factor_emm,
+             aes(x = year, y = exp(the.emmean), colour = habitat), size = 2, position = position_dodge(0.4)) +
+  geom_line(data = site_macrofull_richness_factor_emm, aes(x = year, y = exp(the.emmean), colour = habitat, group = habitat),
+            position = position_dodge(0.4)) +
+  geom_errorbar(data = site_macrofull_richness_factor_emm,
+                aes(x = year, ymin = exp(asymp.LCL), ymax = exp(asymp.UCL), colour = habitat), width = 0.2,
+                position = position_dodge(0.4)) +
+  geom_segment(aes(y = 22, yend = 22, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  scale_y_log10() +
+    labs(x = "Year", y = "Richness") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) ) # add box around legend
+
+site_macrofull_shannon_mod_factor <- glmmTMB(log(shannon+1) ~ year_factor*habitat + (1|site) + ar1(year_factor + 0 | site),
+                                        data = alpha_diversity_site_macro) # no convergence issues
+
+summary(site_macrofull_shannon_mod_factor) # fairly low temporal ac
+Anova(site_macrofull_shannon_mod_factor) 
+#                 Chisq Df Pr(>Chisq)    
+# year_factor          88.821 16  4.122e-12 ***
+# habitat             144.112  3  < 2.2e-16 ***
+# year_factor:habitat  58.923 46    0.09568 .  
+qqnorm(residuals(site_macrofull_shannon_mod_factor)) # a few points wonky but pretty good
+plot(residuals(site_macrofull_shannon_mod_factor) ~ fitted(site_macrofull_shannon_mod_factor))# looks good
+
+site_macrofull_shannon_factor_emm <- plot(emmeans(site_macrofull_shannon_mod_factor, ~  habitat*year_factor), plotit = F) 
+site_macrofull_shannon_factor_emm$year <- as.numeric(levels(site_macrofull_shannon_factor_emm$year_factor)) # for plotting
+
+(site_macrofactor_shannon_plot <- 
+  ggplot() +
+  geom_point(data = site_macrofull_shannon_factor_emm,
+             aes(x = year, y = exp(the.emmean), colour = habitat), size = 2, position = position_dodge(0.4)) +
+  geom_line(data = site_macrofull_shannon_factor_emm, aes(x = year, y = exp(the.emmean), colour = habitat, group = habitat),
+            position = position_dodge(0.4)) +
+  geom_errorbar(data = site_macrofull_shannon_factor_emm,
+                aes(x = year, ymin = exp(lower.CL), ymax = exp(upper.CL), colour = habitat), width = 0.2,
+                position = position_dodge(0.4)) +
+  geom_segment(aes(y = 3, yend = 3, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
   scale_fill_manual(values = habitat_colours) +
@@
 ## Backreef - includes all 
-# backreef_all_site <- na.omit(subset(alpha_diversity_site, habitat == "backreef")) # only use rows that contain all factors
+backreef_all_site <- na.omit(subset(alpha_diversity_site, habitat == "backreef")) # only use rows that contain all factors
 # 
@@
 summary(outer17_macro_shannon_mod_universal) # also NaNs...
+
+```
+
+```{r stepwise models beta dispersion site level, fig.height = 8, fig.width=12}
+
+## FRINGING - exclude wave power
+
+fringing_all_betadispersion_mod_universal <- glmmTMB(betadispersion_trans ~ year + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = fringing_all_site, na.action = na.pass, family = beta_family())
+fringing_all_betadispersion_dredge <- dredge(fringing_all_betadispersion_mod_universal) 
+model.sel(fringing_all_betadispersion_dredge) # top model is just intercept, next best is just year
+fringing_all_betadispersion_top_mod <- glmmTMB(betadispersion_trans ~ year + (1|site), data = fringing_all_site, na.action = na.pass, family = beta_family())
+
+qqnorm(residuals(fringing_all_betadispersion_top_mod))# good
+
+plot(residuals(fringing_all_betadispersion_top_mod) ~ fitted(fringing_all_betadispersion_top_mod)) # good
+summary(fringing_all_betadispersion_top_mod) # not even sig
+
+## Backreef - includes all
+
+backreef_all_betadispersion_mod_universal <- glmmTMB(betadispersion_trans ~ year + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = backreef_all_site, na.action = na.pass, family = beta_family())
+backreef_all_betadispersion_dredge <- dredge(backreef_all_betadispersion_mod_universal) 
+model.sel(backreef_all_betadispersion_dredge) #  best includes just nitrogen
+backreef_all_betadispersion_top_mod <- glmmTMB(betadispersion_trans ~ mean_n  + (1|site), data = backreef_all_site, na.action = na.pass, family = beta_family())
+qqnorm(residuals(backreef_all_betadispersion_top_mod)) #  a little wonky on the tail but not bad
+plot(residuals(backreef_all_betadispersion_top_mod) ~ fitted(backreef_all_betadispersion_top_mod)) # looks fine
+summary(backreef_all_betadispersion_top_mod) # borderline
+Anova(backreef_all_betadispersion_top_mod) # X2 = 3.5, p = 0.06
+
+backreef_emm <-as.data.frame(emmeans(backreef_all_betadispersion_top_mod, ~mean_n,
+                                     at = list(mean_n = seq(min(backreef_all_site$mean_n),
+                                                            max(backreef_all_site$mean_n),
+                                                            0.01))))
+
+ggplot()+
+  geom_point(data = backreef_all_site, aes(x = mean_n, y = betadispersion_trans))+
+  geom_line(data = backreef_emm, aes(x = mean_n, y = inv_logit(emmean))) +
+  geom_ribbon(data = backreef_emm, aes(x = mean_n, ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), alpha = 0.2), show.legend = F) +
+  labs(x = "Mean N", y = "Backreef beta dispersion") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+## outer10 - no nitrogen 
+outer10_all_site <- subset(alpha_diversity_site, habitat == "outer_10") # only use rows that contain all factors
+# remove nitrogent columns
+outer10_all_site <- outer10_all_site[,-which(names(outer10_all_site) %in% c("mean_n", "n_cv"))]
+# remove other na rows
+outer10_all_site <- na.omit(outer10_all_site)
+
+outer10_all_betadispersion_mod_universal <- glmmTMB(betadispersion_trans ~ year  + mean_temp_c + wave_power_adcp +  herbivore_evenness + (1|site), data = outer10_all_site, na.action = na.pass, family = beta_family())
+outer10_all_betadispersion_dredge <- dredge(outer10_all_betadispersion_mod_universal) 
+model.sel(outer10_all_betadispersion_dredge) #  best mean temp
+outer10_all_betadispersion_top_mod <- glmmTMB(betadispersion_trans ~  mean_temp_c + (1|site), data = outer10_all_site, na.action = na.pass, family = beta_family())
+qqnorm(residuals(outer10_all_betadispersion_top_mod)) # pretty good
+plot(residuals(outer10_all_betadispersion_top_mod) ~ fitted(outer10_all_betadispersion_top_mod)) # looks fine
+summary(outer10_all_betadispersion_top_mod)
+Anova(outer10_all_betadispersion_top_mod) # P = 0.01
+outer10_emm <-as.data.frame(emmeans(outer10_all_betadispersion_top_mod, ~mean_temp_c,
+                                     at = list(mean_temp_c = seq(min(outer10_all_site$mean_temp_c),
+                                                            max(outer10_all_site$mean_temp_c),
+                                                            0.01))))
+
+ggplot()+
+  geom_point(data = outer10_all_site, aes(x = mean_temp_c, y = betadispersion_trans))+
+  geom_line(data = outer10_emm, aes(x = mean_temp_c, y = inv_logit(emmean))) +
+  geom_ribbon(data = outer10_emm, aes(x = mean_temp_c, ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), alpha = 0.2), show.legend = F) +
+  labs(x = "Mean temp (C)", y = "Outer 10 betadispersion") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+## outer17 - no nitrogen or herbivores
+
+outer17_all_betadispersion_mod_universal <- glmmTMB(betadispersion_trans ~ year + wave_power_adcp +  + mean_temp_c + (1|site), data = outer17_all_site, na.action = na.pass, family = beta_family())
+outer17_all_betadispersion_dredge <- dredge(outer17_all_betadispersion_mod_universal) # some convergence issues
+model.sel(outer17_all_betadispersion_dredge) #  best includes just int, next best is temp
+outer17_all_betadispersion_top_mod <-glmmTMB(betadispersion_trans ~  mean_temp_c + (1|site), data = outer17_all_site, na.action = na.pass, family = beta_family())
+qqnorm(residuals(outer17_all_betadispersion_top_mod)) # pretty good
+plot(residuals(outer17_all_betadispersion_top_mod) ~ fitted(outer17_all_betadispersion_top_mod)) # looks fine
+summary(outer17_all_betadispersion_top_mod)
+
+```
+
+```{r stepwise models site level macroalgae}
+
+## FRINGING 
+fringing_macro_betadispersion_mod_universal <- glmmTMB(betadispersion_trans ~ year  + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = fringing_macro_site, na.action = na.pass, family = beta_family())
+fringing_macro_betadispersion_dredge <- dredge(fringing_macro_betadispersion_mod_universal) 
+model.sel(fringing_macro_betadispersion_dredge) # top model just int, next best includes n
+fringing_macro_betadispersion_top_mod <- glmmTMB(betadispersion_trans ~ mean_n + (1|site), data = fringing_macro_site, na.action = na.pass, family = beta_family())
+summary(fringing_macro_betadispersion_top_mod)
+qqnorm(residuals(fringing_macro_betadispersion_top_mod))# not bad
+plot(residuals(fringing_macro_betadispersion_top_mod) ~ fitted(fringing_macro_betadispersion_top_mod)) # good
+
+## Backreef - includes all
+
+backreef_macro_betadispersion_mod_universal <- glmmTMB(betadispersion_trans ~ year + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = backreef_macro_site, na.action = na.pass, family = beta_family()) # convergence issues
+backreef_macro_betadispersion_dredge <- dredge(backreef_macro_betadispersion_mod_universal) 
+model.sel(backreef_macro_betadispersion_dredge) #  herb, n, and year
+backreef_macro_betadispersion_top_mod <- glmmTMB(betadispersion_trans ~ year +  mean_n +  herbivore_evenness + (1|site), data = backreef_macro_site, na.action = na.pass, family = beta_family())
+qqnorm(residuals(backreef_macro_betadispersion_top_mod)) #  a little wonky but not bad
+plot(residuals(backreef_macro_betadispersion_top_mod) ~ fitted(backreef_macro_betadispersion_top_mod)) # negative trend
+summary(backreef_macro_betadispersion_top_mod)
+Anova(backreef_macro_betadispersion_top_mod)
+#                     Chisq Df Pr(>Chisq)   
+# year               4.5193  1    0.03351 * 
+# mean_n             2.4331  1    0.11880   
+# herbivore_evenness 6.8010  1    0.00911 **
+
+backreef_emm <-as.data.frame(emmeans(backreef_macro_betadispersion_top_mod, ~herbivore_evenness,
+                                     at = list(herbivore_evenness = seq(min(backreef_macro_site$herbivore_evenness),
+                                                            max(backreef_macro_site$herbivore_evenness),
+                                                            0.05)))) 
+ggplot()+
+  geom_point(data = backreef_macro_site, aes(x = herbivore_evenness, y = betadispersion_trans))+
+  geom_line(data = backreef_emm, aes(x = herbivore_evenness, y = inv_logit(emmean)) )+
+  geom_ribbon(data = backreef_emm, aes(x = herbivore_evenness, ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), alpha = 0.2), show.legend = F) +
+  labs(x = "Herbivore evenness", y = "Backreef betadispersion") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+## outer10 - no nitrogen
+
+outer10_macro_betadispersion_mod_universal <- glmmTMB(betadispersion_trans ~ year + wave_power_adcp + mean_temp_c +  herbivore_evenness + (1|site), data = outer10_macro_site, na.action = na.pass, family = beta_family()) # convergence
+outer10_macro_betadispersion_dredge <- dredge(outer10_macro_betadispersion_mod_universal) 
+model.sel(outer10_macro_betadispersion_dredge) #  best evenness and year
+outer10_macro_betadispersion_top_mod <- glmmTMB(betadispersion_trans ~ year + herbivore_evenness + (1|site), data = outer10_macro_site, na.action = na.pass, family = beta_family()) 
+qqnorm(residuals(outer10_macro_betadispersion_top_mod)) # pretty good
+plot(residuals(outer10_macro_betadispersion_top_mod) ~ fitted(outer10_macro_betadispersion_top_mod)) # looks fine
+summary(outer10_macro_betadispersion_top_mod) #
+Anova(outer10_macro_betadispersion_top_mod)
+
+outer10_emm <-as.data.frame(emmeans(outer10_macro_betadispersion_top_mod, ~herbivore_evenness,
+                                     at = list(herbivore_evenness = seq(min(outer10_macro_site$herbivore_evenness),
+                                                            max(outer10_macro_site$herbivore_evenness),
+                                                            0.05)))) 
+ggplot()+
+  geom_point(data = outer10_macro_site, aes(x = herbivore_evenness, y = betadispersion_trans))+
+  geom_line(data = outer10_emm, aes(x = herbivore_evenness, y = inv_logit(emmean)) )+
+  geom_ribbon(data = outer10_emm, aes(x = herbivore_evenness, ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), alpha = 0.2), show.legend = F) +
+  labs(x = "Herbivore evenness", y = "outer10 betadispersion") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+## outer17 - no nitrogen or herbivores
+outer17_macro_site<- subset(alpha_diversity_site_macro, habitat == "outer_17") # only use rows that contain all factors
+# remove nitrogent columns
+outer17_macro_site<- outer17_macro_site[,-which(names(outer17_macro_site) %in% c("mean_n", "n_cv", "herbivore_biomass", "herbivore_evenness", "herbivore_richness", "herbivore_betadispersion"))]
+# remove other na rows
+outer17_macro_site <- na.omit(outer17_macro_site)
+
+outer17_macro_betadispersion_mod_universal <- glmmTMB(betadispersion_trans ~ year + wave_power_adcp + mean_temp_c +  (1|site), data = outer10_macro_site, na.action = na.pass, family = beta_family()) # convergence
+outer17_macro_betadispersion_dredge <- dredge(outer17_macro_betadispersion_mod_universal) 
+model.sel(outer17_macro_betadispersion_dredge) #  best just year, then year and temp
+qqnorm(residuals(outer17_macro_betadispersion_mod_universal)) # good
+plot(residuals(outer17_macro_betadispersion_mod_universal) ~ fitted(outer17_macro_betadispersion_mod_universal)) # looks fine
+summary(outer17_macro_betadispersion_mod_universal) # NaNs...
 
@@
 ```
+
 ```{r site macro turnover}
@@
 ```
+
 ```{r site all turnover}
@@
 
-```{r all algae alpha diversity quadrat level}
+```{r all algae richness quadrat level}
 
 full_richness_mod <- glmmTMB(richness ~ year*habitat + (1|site/location) + offset(log(hard_cover)), family = poisson,
-                             data = alpha_diversity_quad[!is.na(alpha_diversity_quad$macroalgae_cover_hard),]) 
+                             data = alpha_diversity_quad) 
 summary(full_richness_mod)
-hist(residuals(full_richness_mod)) # slightly skewed but not terrible
+qqnorm(residuals(full_richness_mod)) # slight positive skew but not terrible
 
@@
 
-emtrends(full_richness_mod, pairwise ~  habitat, var = "year")
+emtrends(full_richness_mod, pairwise ~  habitat, var = "year") # fringing slope not different from forereef slopes
 full_richness_emm <- emmip(full_richness_mod,habitat ~ year,
@@
   geom_ribbon(data = full_richness_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
+  geom_segment(aes(y = 8, yend = 8, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_colour_manual(values = habitat_colours) +
@@
 
-```
-
-```{r macroalgae diversity quadrat level}
-
-```
+full_richness_mod_factor <- glmmTMB(richness ~ year_factor*habitat + (1|site/location) + ar1(year_factor + 0 | location) + offset(log(hard_cover)), family = poisson, data = alpha_diversity_quad) # convergence problem persists even after removing random effect
+
+summary(full_richness_mod_factor) # strong temporal autocorrelation
+Anova(full_richness_mod_factor) 
+#                 Chisq Df Pr(>Chisq)    
+# year_factor         594.41 16  < 2.2e-16 ***
+# habitat             563.42  3  < 2.2e-16 ***
+# year_factor:habitat 699.47 46  < 2.2e-16 ***
+qqnorm(residuals(full_richness_mod_factor)) # looks great - CLT?
+
+# striping/patterining expected because of discrete response var.
+plot(residuals(full_richness_mod_factor) ~ fitted(full_richness_mod_factor))
+
+full_richness_mod_factor_emm <- plot(emmeans(full_richness_mod_factor, ~  habitat*year_factor), plotit = F) 
+full_richness_mod_factor_emm$year <- as.numeric(levels(full_richness_mod_factor_emm$year_factor)) # for plotting
+
+(quad_factor_richness_plot <- 
+  ggplot() +
+  geom_point(data = full_richness_mod_factor_emm,
+             aes(x = year, y = exp(the.emmean), colour = habitat), size = 2, position = position_dodge(0.4)) +
+  geom_line(data = full_richness_mod_factor_emm, aes(x = year, y = exp(the.emmean), colour = habitat, group = habitat),
+            position = position_dodge(0.4)) +
+  geom_errorbar(data = full_richness_mod_factor_emm,
+                aes(x = year, ymin = exp(asymp.LCL), ymax = exp(asymp.UCL), colour = habitat), width = 0.2,
+                position = position_dodge(0.4)) +
+  geom_segment(aes(y = 8, yend = 8, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  scale_y_log10() +
+    labs(x = "Year", y = "Richness") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) ) # add box around legend
+```
+```{r algae shannon quadrat level}
+
+full_shannon_mod <- glmmTMB(log(shannon+1) ~ year*habitat + (1|site/location),
+                             data = alpha_diversity_quad) 
+summary(full_shannon_mod)
+qqnorm(residuals(full_shannon_mod)) 
+plot(residuals(full_shannon_mod) ~ fitted(full_shannon_mod)) # stripe at low div
+
+emtrends(full_shannon_mod, pairwise ~  habitat, var = "year") # all diff
+full_shannon_emm <- emmip(full_shannon_mod,habitat ~ year,
+                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)
+
+(base_shannon_plot <- 
+  ggplot() +
+  geom_point(data = alpha_diversity_quad,
+             aes(x = year, y = shannon, colour = habitat), size = 2, alpha = 0.05) + 
+  geom_line(data = full_shannon_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = full_shannon_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
+  geom_segment(aes(y = 3, yend = 3, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Shannon Index (H')") +
+  scale_y_log10() +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) ) # add box around legend
+
+full_shannon_mod_factor <- glmmTMB(log(shannon+1) ~ year_factor*habitat +  ar1(year_factor + 0 | location), data = alpha_diversity_quad) # convergence problem persists even after removing random effect
+
+summary(full_shannon_mod_factor) # strong temporal autocorrelation
+Anova(full_shannon_mod_factor) 
+# everything is 2.2e-16
+qqnorm(residuals(full_shannon_mod_factor)) # looks great - CLT?
+
+plot(residuals(full_shannon_mod_factor) ~ fitted(full_shannon_mod_factor)) # same stripe at low values
+
+full_shannon_mod_factor_emm <- plot(emmeans(full_shannon_mod_factor, ~  habitat*year_factor), plotit = F) # warning negative variance
+# could not compute 95% CIs unless random effect is removed
+full_shannon_mod_factor_emm$year <- as.numeric(levels(full_shannon_mod_factor_emm$year_factor)) # for plotting
+
+(quad_factor_shannon_plot <- 
+  ggplot() +
+  geom_point(data = full_shannon_mod_factor_emm,
+             aes(x = year, y = exp(the.emmean), colour = habitat), size = 2, position = position_dodge(0.4)) +
+  geom_line(data = full_shannon_mod_factor_emm, aes(x = year, y = exp(the.emmean), colour = habitat, group = habitat),
+            position = position_dodge(0.4)) +
+  geom_errorbar(data = full_shannon_mod_factor_emm,
+                aes(x = year, ymin = exp(lower.CL), ymax = exp(upper.CL), colour = habitat), width = 0.2,
+                position = position_dodge(0.4)) +
+  geom_segment(aes(y = 2, yend = 2, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  scale_y_log10() +
+    labs(x = "Year", y = "Shannon Index (H')") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) ) # add box around legend
+```
+
+```{r macroalgae richness quadrat level}
+
+macro_richness_mod <- glmmTMB(richness ~ year*habitat + (1|site/location) + offset(log(hard_cover)), family = poisson,
+                             data = alpha_diversity_quad_macro) 
+summary(macro_richness_mod)
+qqnorm(residuals(macro_richness_mod)) # slight positive skew but not terrible
+
+# striping/patterining expected because of discrete response var. otherwise fine
+plot(residuals(macro_richness_mod) ~ fitted(macro_richness_mod))
+
+emtrends(macro_richness_mod, pairwise ~  habitat, var = "year") # fringing slope not different from forereef slopes
+macro_richness_emm <- emmip(macro_richness_mod,habitat ~ year,
+                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)
+
+(base_richness_plot <- 
+  ggplot() +
+  geom_point(data = alpha_diversity_quad_macro,
+             aes(x = year, y = richness, colour = habitat), size = 2, alpha = 0.05) + 
+  geom_line(data = macro_richness_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = macro_richness_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
+  geom_segment(aes(y = 8, yend = 8, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Richness") +
+  scale_y_log10() +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) ) # add box around legend
+
+macro_richness_mod_factor <- glmmTMB(richness ~ year_factor*habitat + (1|site/location) + ar1(year_factor + 0 | location) + offset(log(hard_cover)), family = poisson, data = alpha_diversity_quad_macro) # convergence problem persists even after removing random effect
+
+summary(macro_richness_mod_factor) # strong temporal autocorrelation
+Anova(macro_richness_mod_factor) # useless metric
+qqnorm(residuals(macro_richness_mod_factor)) # some negative skew wonkiness but not bad
+plot(residuals(macro_richness_mod_factor) ~ fitted(macro_richness_mod_factor))
+
+macro_richness_mod_factor_emm <- plot(emmeans(macro_richness_mod_factor, ~  habitat*year_factor), plotit = F) 
+macro_richness_mod_factor_emm$year <- as.numeric(levels(macro_richness_mod_factor_emm$year_factor)) # for plotting
+
+(quad_factor_richness_plot <- 
+  ggplot() +
+  geom_point(data = macro_richness_mod_factor_emm,
+             aes(x = year, y = exp(the.emmean), colour = habitat), size = 2, position = position_dodge(0.4)) +
+  geom_line(data = macro_richness_mod_factor_emm, aes(x = year, y = exp(the.emmean), colour = habitat, group = habitat),
+            position = position_dodge(0.4)) +
+  geom_errorbar(data = macro_richness_mod_factor_emm,
+                aes(x = year, ymin = exp(asymp.LCL), ymax = exp(asymp.UCL), colour = habitat), width = 0.2,
+                position = position_dodge(0.4)) +
+  geom_segment(aes(y = 3, yend = 3, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  scale_y_log10() +
+    labs(x = "Year", y = "Richness") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) ) # add box around legend
+```
+
+```{r all algae shannon quadrat level}
+
+macro_shannon_mod <- glmmTMB(log(shannon+1) ~ year*habitat + (1|site/location),
+                             data = alpha_diversity_quad_macro) 
+summary(macro_shannon_mod)
+hist(residuals(macro_shannon_mod)) # weird pattern - almost bimodal.
+plot(residuals(macro_shannon_mod) ~ fitted(macro_shannon_mod)) # stronger patterning
+
+emtrends(macro_shannon_mod, pairwise ~  habitat, var = "year") # fringing no different from outer 10
+macro_shannon_emm <- emmip(macro_shannon_mod,habitat ~ year,
+                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)
+
+(base_shannon_plot <- 
+  ggplot() +
+  geom_point(data = alpha_diversity_quad_macro,
+             aes(x = year, y = shannon, colour = habitat), size = 2, alpha = 0.05) + 
+  geom_line(data = macro_shannon_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = macro_shannon_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
+  geom_segment(aes(y = 3, yend = 3, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Shannon Index (H')") +
+  scale_y_log10() +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) ) # add box around legend
+
+macro_shannon_mod_factor <- glmmTMB(log(shannon+1) ~ year_factor*habitat + (1|site/location) +  ar1(year_factor + 0 | location), data = alpha_diversity_quad_macro) # convergence problem persists even after removing random effect
+
+summary(macro_shannon_mod_factor) # strong temporal autocorrelation
+Anova(macro_shannon_mod_factor) 
+# everything is 2.2e-16
+qqnorm(residuals(macro_shannon_mod_factor)) # similar issue
+
+plot(residuals(macro_shannon_mod_factor) ~ fitted(macro_shannon_mod_factor)) # here too
+
+macro_shannon_mod_factor_emm <- plot(emmeans(macro_shannon_mod_factor, ~  habitat*year_factor), plotit = F) # warning negative variance
+macro_shannon_mod_factor_emm$year <- as.numeric(levels(macro_shannon_mod_factor_emm$year_factor)) # for plotting
+
+(quad_factor_shannon_plot <- 
+  ggplot() +
+  geom_point(data = macro_shannon_mod_factor_emm,
+             aes(x = year, y = exp(the.emmean), colour = habitat), size = 2, position = position_dodge(0.4)) +
+  geom_line(data = macro_shannon_mod_factor_emm, aes(x = year, y = exp(the.emmean), colour = habitat, group = habitat),
+            position = position_dodge(0.4)) +
+  geom_errorbar(data = macro_shannon_mod_factor_emm,
+                aes(x = year, ymin = exp(lower.CL), ymax = exp(upper.CL), colour = habitat), width = 0.2,
+                position = position_dodge(0.4)) +
+  geom_segment(aes(y = 2, yend = 2, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  scale_y_log10() +
+    labs(x = "Year", y = "Shannon Index (H')") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) ) # add box around legend
+```
+```{r read and mege beta dispersion}
+# beta dispersion was calculated in noam_beta_diversity.Rmd
+# written to csv
+quad_all_betadisper <- read.csv(here("data", "quad_level_all_species_with_betadisper.csv"))
+quad_macro_betadisper <- read.csv(here("data", "quad_level_just_macroalgae_with_betadisper.csv"))
+
+alpha_diversity_quad <- full_join(alpha_diversity_quad, quad_all_betadisper[,c("year", "betadispersion_trans", "location")], by = c("location", "year"))
+
+alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro, quad_macro_betadisper[,c("year", "betadispersion_trans", "location")], by = c("location", "year"))
+```

### 2024-09-10 noam_final_analyses.Rmd
up to current analyses

@@
 alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro, quad_macro_betadisper[,c("year", "betadispersion_trans", "location")], by = c("location", "year"))
-```
+
+fringing_all_quad <- na.omit(subset(alpha_diversity_quad, habitat == "fringing"))
+backreef_all_quad <- na.omit(subset(alpha_diversity_quad, habitat == "backreef"))
+outer10_all_quad <- subset(alpha_diversity_quad, habitat == "outer_10") 
+outer10_all_quad <- outer10_all_quad[,-which(names(outer10_all_quad) %in% c("mean_n", "n_cv"))]
+outer10_all_quad <- na.omit(outer10_all_quad)
+outer17_all_quad <- subset(alpha_diversity_quad, habitat == "outer_17") 
+outer17_all_quad <- outer17_all_quad[,-which(names(outer17_all_quad) %in% c("mean_n", "n_cv", "herbivore_biomass", "herbivore_evenness", "herbivore_richness", "herbivore_shannon"))]
+outer17_all_quad <- na.omit(outer17_all_quad)
+
+fringing_macro_quad <- na.omit(subset(alpha_diversity_quad_macro, habitat == "fringing"))
+backreef_macro_quad <- na.omit(subset(alpha_diversity_quad_macro, habitat == "backreef"))
+outer10_macro_quad <- subset(alpha_diversity_quad, habitat == "outer_10") 
+outer10_macro_quad <- outer10_macro_quad[,-which(names(outer10_macro_quad) %in% c("mean_n", "n_cv"))]
+outer10_macro_quad <- na.omit(outer10_macro_quad)
+outer17_macro_quad <- subset(alpha_diversity_quad, habitat == "outer_17") 
+outer17_macro_quad <- outer17_macro_quad[,-which(names(outer17_macro_quad) %in% c("mean_n", "n_cv", "herbivore_biomass", "herbivore_evenness", "herbivore_richness", "herbivore_shannon"))]
+outer17_macro_quad <- na.omit(outer17_macro_quad)
+```
+
+```{r stepwise model all shannon quad}
+fringing_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + mean_n +  mean_temp_c +  herbivore_evenness + (1|site/location), data = fringing_all_quad, na.action = na.pass)
+fringing_all_shannon_dredge <- dredge(fringing_all_shannon_mod_universal) 
+model.sel(fringing_all_shannon_dredge) # all terms
+qqnorm(residuals(fringing_all_shannon_mod_universal))# a little wonky
+qqline(residuals(fringing_all_shannon_mod_universal))
+plot(residuals(fringing_all_shannon_mod_universal) ~ fitted(fringing_all_shannon_mod_universal)) # sharp line again
+summary(fringing_all_shannon_mod_universal)
+Anova(fringing_all_shannon_mod_universal) # only year and mean temp sig. herbivore evenness is borderline.
+
+fringing_emm <-as.data.frame(emmeans(fringing_all_shannon_top_mod, ~mean_temp_c,
+                                     at = list(mean_temp_c = seq(min(fringing_all_quad$mean_temp_c),
+                                                            max(fringing_all_quad$mean_temp_c),
+                                                            0.01)))) 
+ggplot()+
+  geom_point(data = fringing_all_quad, aes(x = mean_temp_c, y = shannon))+
+  geom_line(data = fringing_emm, aes(x = mean_temp_c, y = emmean)) +
+  geom_ribbon(data = fringing_emm, aes(x = mean_temp_c, ymin = lower.CL, ymax = upper.CL, alpha = 0.2), show.legend = F) +
+  scale_y_log10() + 
+  labs(x = "Mean Temp (C)", y = "Fringing Shannon Index (H')") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+## Backreef - includes all
+
+backreef_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + (1|site/location), data = backreef_all_quad, na.action = na.pass)
+backreef_all_shannon_dredge <- dredge(backreef_all_shannon_mod_universal) # some have convergence issues
+model.sel(backreef_all_shannon_dredge) #  best includes all but results in only NaN results. Next best removes year and wave power
+backreef_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~ mean_n + mean_temp_c +  herbivore_evenness + (1|site/location), data = backreef_all_quad, na.action = na.pass) #  a little wonky on the tail but not bad
+hist(residuals(backreef_all_shannon_top_mod)) # still a little chunky
+plot(residuals(backreef_all_shannon_top_mod) ~ fitted(backreef_all_shannon_top_mod)) # same stripe
+summary(backreef_all_shannon_top_mod)
+Anova(backreef_all_shannon_top_mod)
+
+backreef_emm <-as.data.frame(emmeans(backreef_all_shannon_top_mod, ~mean_temp_c,
+                                     at = list(mean_temp_c = seq(min(backreef_all_quad$mean_temp_c),
+                                                            max(backreef_all_quad$mean_temp_c),
+                                                            0.01)))) 
+ggplot()+
+  geom_point(data = backreef_all_quad, aes(x = mean_temp_c, y = shannon))+
+  geom_line(data = backreef_emm, aes(x = mean_temp_c, y = emmean)) +
+  geom_ribbon(data = backreef_emm, aes(x = mean_temp_c, ymin = lower.CL, ymax = upper.CL, alpha = 0.2), show.legend = F) +
+  scale_y_log10() + 
+  labs(x = "Mean Temp (C)", y = "Backreef Shannon Index (H')") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+backreef_emm_2 <-as.data.frame(emmeans(backreef_all_shannon_top_mod, ~mean_n,
+                                     at = list(mean_n = seq(min(backreef_all_quad$mean_n),
+                                                            max(backreef_all_quad$mean_n),
+                                                            0.01)))) 
+ggplot()+
+  geom_point(data = backreef_all_quad, aes(x = mean_n, y = shannon))+
+  geom_line(data = backreef_emm_2, aes(x = mean_n, y = emmean)) +
+  geom_ribbon(data = backreef_emm_2, aes(x = mean_n, ymin = lower.CL, ymax = upper.CL, alpha = 0.2), show.legend = F) +
+  scale_y_log10() + 
+  labs(x = "Mean N", y = "Backreef Shannon Index (H')") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+backreef_emm_3 <-as.data.frame(emmeans(backreef_all_shannon_top_mod, ~herbivore_evenness,
+                                     at = list(herbivore_evenness = seq(min(backreef_all_quad$herbivore_evenness),
+                                                            max(backreef_all_quad$herbivore_evenness),
+                                                            0.01)))) 
+ggplot()+
+  geom_point(data = backreef_all_quad, aes(x = herbivore_evenness, y = shannon))+
+  geom_line(data = backreef_emm_3, aes(x = herbivore_evenness, y = emmean)) +
+  geom_ribbon(data = backreef_emm_3, aes(x = herbivore_evenness, ymin = lower.CL, ymax = upper.CL, alpha = 0.2), show.legend = F) +
+  scale_y_log10() + 
+  labs(x = "Herbivore evenness", y = "Backreef Shannon Index (H')") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+# Outer10
+outer10_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp  + mean_temp_c +  herbivore_evenness + (1|site/location), data = outer10_all_quad, na.action = na.pass)
+outer10_all_shannon_dredge <- dredge(outer10_all_shannon_mod_universal) 
+model.sel(outer10_all_shannon_dredge) #  best includes just herbivore evenness 
+outer10_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  herbivore_evenness + (1|site/location), data = outer10_all_quad, na.action = na.pass)
+qqnorm(residuals(outer10_all_shannon_top_mod)) # slight negative skew
+plot(residuals(outer10_all_shannon_top_mod) ~ fitted(outer10_all_shannon_top_mod)) # looks fine
+summary(outer10_all_shannon_top_mod)
+Anova(outer10_all_shannon_top_mod)
+outer10_emm <-as.data.frame(emmeans(outer10_all_shannon_top_mod, ~herbivore_evenness,
+                                     at = list(herbivore_evenness = seq(min(outer10_all_quad$herbivore_evenness),
+                                                            max(outer10_all_quad$herbivore_evenness),
+                                                            0.05))))
+
+ggplot()+
+  geom_point(data = outer10_all_quad, aes(x = herbivore_evenness, y = shannon))+
+  geom_line(data = outer10_emm, aes(x = herbivore_evenness, y = emmean)) +
+  geom_ribbon(data = outer10_emm, aes(x = herbivore_evenness, ymin = lower.CL, ymax = upper.CL, alpha = 0.2), show.legend = F) +
+  scale_y_log10() + 
+  labs(x = "Herbivore evenness", y = "Outer 10 Shannon Index (H')") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+## outer17
+
+outer17_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp  + mean_temp_c + (1|site/location), data = outer17_all_quad, na.action = na.pass)
+outer17_all_shannon_dredge <- dredge(outer17_all_shannon_mod_universal) 
+model.sel(outer17_all_shannon_dredge) # best includes all but results in NaNs
+# next best includes just year and mean_temp
+outer17_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~ year  + mean_temp_c + (1|site/location), data = outer17_all_quad, na.action = na.pass)
+hist(residuals(outer17_all_shannon_top_mod))
+plot(residuals(outer17_all_shannon_top_mod) ~ fitted(outer17_all_shannon_top_mod)) # looks fine
+summary(outer17_all_shannon_top_mod)
+Anova(outer17_all_shannon_top_mod)
+
+outer17_emm <-as.data.frame(emmeans(outer17_all_shannon_top_mod, ~mean_temp_c,
+                                     at = list(mean_temp_c = seq(min(outer17_all_quad$mean_temp_c),
+                                                            max(outer17_all_quad$mean_temp_c),
+                                                            0.01)))) 
+ggplot()+
+  geom_point(data = outer17_all_quad, aes(x = mean_temp_c, y = shannon))+
+  geom_line(data = outer17_emm, aes(x = mean_temp_c, y = emmean)) +
+  geom_ribbon(data = outer17_emm, aes(x = mean_temp_c, ymin = lower.CL, ymax = upper.CL, alpha = 0.2), show.legend = F) +
+  scale_y_log10() + 
+  labs(x = "Mean Temp (C)", y = "Outer 17 Shannon Index (H')") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+```
+```{r stepwise model all betadisp quad}
+fringing_all_betadisp_mod_universal <- glmmTMB(betadispersion_trans ~ year + mean_n +  mean_temp_c +  herbivore_evenness + (1|site/location), data = fringing_all_quad, na.action = na.pass, family = beta_family())
+fringing_all_betadisp_dredge <- dredge(fringing_all_betadisp_mod_universal) 
+model.sel(fringing_all_betadisp_dredge) # best includes evenness, nitrogen, and year
+fringing_all_betadisp_top_mod <- glmmTMB(betadispersion_trans ~ year +  mean_n +  herbivore_evenness + (1|site/location), data = fringing_all_quad, na.action = na.pass, family = beta_family())
+hist(residuals(fringing_all_betadisp_top_mod)) #  positive skew
+plot(residuals(fringing_all_betadisp_top_mod) ~ fitted(fringing_all_betadisp_top_mod)) # negative trend
+summary(fringing_all_betadisp_top_mod)
+Anova(fringing_all_betadisp_top_mod) # herbivore evenness not sig
+
+fringing_betadisp_emm <-as.data.frame(emmeans(fringing_all_betadisp_top_mod, ~mean_n,
+                                     at = list(mean_n = seq(min(fringing_all_quad$mean_n),
+                                                            max(fringing_all_quad$mean_n),
+                                                            0.01)))) 
+ggplot()+
+  geom_point(data = fringing_all_quad, aes(x = mean_n, y = betadispersion_trans))+
+  geom_line(data = fringing_betadisp_emm, aes(x = mean_n, y = inv_logit(emmean))) +
+  geom_ribbon(data = fringing_betadisp_emm, aes(x = mean_n, ymin = inv_logit(asymp.LCL),
+                                                ymax = inv_logit(asymp.UCL), alpha = 0.2), show.legend = F) +
+  labs(x = "Mean N", y = "Fringing Beta Dispersion") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+## Backreef - includes all
+
+backreef_all_betadisp_mod_universal <- glmmTMB(betadispersion_trans ~ year  + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + (1|site/location), data = backreef_all_quad, na.action = na.pass, family = beta_family())
+backreef_all_betadisp_dredge <- dredge(backreef_all_betadisp_mod_universal) # most have convergence issues
+model.sel(backreef_all_betadisp_dredge) #  best includes evenness nitrogen year
+backreef_all_betadisp_top_mod <- glmmTMB(betadispersion_trans ~ year  +  mean_n + herbivore_evenness + (1|site/location), data = backreef_all_quad, na.action = na.pass, family = beta_family()) 
+hist(residuals(backreef_all_betadisp_top_mod)) # positive skew
+plot(residuals(backreef_all_betadisp_top_mod) ~ fitted(backreef_all_betadisp_top_mod)) # negative trend
+summary(backreef_all_betadisp_top_mod)
+Anova(backreef_all_betadisp_top_mod) # evenness and year sig
+
+backreef_betadisp_emm <-as.data.frame(emmeans(backreef_all_betadisp_top_mod, ~herbivore_evenness,
+                                     at = list(herbivore_evenness = seq(min(backreef_all_quad$herbivore_evenness),
+                                                            max(backreef_all_quad$herbivore_evenness),
+                                                            0.01)))) 
+ggplot()+
+  geom_point(data = backreef_all_quad, aes(x = herbivore_evenness, y = betadispersion_trans))+
+  geom_line(data = backreef_betadisp_emm, aes(x = herbivore_evenness, y = inv_logit(emmean))) +
+  geom_ribbon(data = backreef_betadisp_emm, aes(x = herbivore_evenness, ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), alpha = 0.2), show.legend = F) +
+  labs(x = "Herbivore evenness", y = "Backreef Beta Dispersion") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+# Outer10
+outer10_all_betadisp_mod_universal <- glmmTMB(betadispersion_trans ~ year  + wave_power_adcp  + mean_temp_c +  herbivore_evenness + (1|site/location), data = outer10_all_quad, na.action = na.pass, family = beta_family()) # model convergence
+outer10_all_betadisp_dredge <- dredge(outer10_all_betadisp_mod_universal) # all have errors of some kind
+model.sel(outer10_all_betadisp_dredge) #  best includes herbivore evenness, temp, and year
+outer10_all_betadisp_top_mod <- glmmTMB(betadispersion_trans ~ year + mean_temp_c +  herbivore_evenness + (1|site/location), data = outer10_all_quad, na.action = na.pass, family = beta_family()) 
+qqnorm(residuals(outer10_all_betadisp_top_mod)) # positive skew
+plot(residuals(outer10_all_betadisp_top_mod) ~ fitted(outer10_all_betadisp_top_mod)) # less negative than before
+summary(outer10_all_betadisp_top_mod)
+Anova(outer10_all_betadisp_top_mod) # all sig
+outer10_betadisp_emm <-as.data.frame(emmeans(outer10_all_betadisp_top_mod, ~herbivore_evenness,
+                                     at = list(herbivore_evenness = seq(min(outer10_all_quad$herbivore_evenness),
+                                                            max(outer10_all_quad$herbivore_evenness),
+                                                            0.05))))
+
+ggplot()+
+  geom_point(data = outer10_all_quad, aes(x = herbivore_evenness, y = betadispersion_trans))+
+  geom_line(data = outer10_betadisp_emm, aes(x = herbivore_evenness, y = inv_logit(emmean))) +
+  geom_ribbon(data = outer10_betadisp_emm,
+              aes(x = herbivore_evenness, ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), alpha = 0.2), show.legend = F) +
+  labs(x = "Herbivore evenness", y = "Outer 10 Beta Dispersion") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+outer10_betadisp_emm2 <-as.data.frame(emmeans(outer10_all_betadisp_top_mod, ~mean_temp_c,
+                                     at = list(mean_temp_c = seq(min(outer10_all_quad$mean_temp_c),
+                                                            max(outer10_all_quad$mean_temp_c),
+                                                            0.05))))
+
+ggplot()+
+  geom_point(data = outer10_all_quad, aes(x = mean_temp_c, y = betadispersion_trans))+
+  geom_line(data = outer10_betadisp_emm2, aes(x = mean_temp_c, y = inv_logit(emmean))) +
+  geom_ribbon(data = outer10_betadisp_emm2,
+              aes(x = mean_temp_c, ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), alpha = 0.2), show.legend = F) +
+  labs(x = "Mean temp (C)", y = "Outer 10 Beta Dispersion") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+## outer17
+
+outer17_all_betadisp_mod_universal <- glmmTMB(betadispersion_trans ~ year  + wave_power_adcp  + mean_temp_c + (1|site/location), data = outer17_all_quad, na.action = na.pass, family = beta_family()) # convergence issue
+outer17_all_betadisp_dredge <- dredge(outer17_all_betadisp_mod_universal) 
+model.sel(outer17_all_betadisp_dredge) # best mean temp and year
+outer17_all_betadisp_top_mod <-  glmmTMB(betadispersion_trans ~ year  +  mean_temp_c + (1|site/location), data = outer17_all_quad, na.action = na.pass, family = beta_family())
+hist(residuals(outer17_all_betadisp_top_mod)) # similar issues
+plot(residuals(outer17_all_betadisp_top_mod) ~ fitted(outer17_all_betadisp_top_mod)) # similar issues
+summary(outer17_all_betadisp_top_mod)
+Anova(outer17_all_betadisp_top_mod) # sig
+
+outer17_betadisp_emm <-as.data.frame(emmeans(outer17_all_betadisp_top_mod, ~mean_temp_c,
+                                     at = list(mean_temp_c = seq(min(outer17_all_quad$mean_temp_c),
+                                                            max(outer17_all_quad$mean_temp_c),
+                                                            0.01)))) 
+ggplot()+
+  geom_point(data = outer17_all_quad, aes(x = mean_temp_c, y = betadispersion_trans))+
+  geom_line(data = outer17_betadisp_emm, aes(x = mean_temp_c, y = inv_logit(emmean))) +
+  geom_ribbon(data = outer17_betadisp_emm, aes(x = mean_temp_c, ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), alpha = 0.2), show.legend = F) +
+  labs(x = "Mean Temp (C)", y = "Outer 17 beta dispersion") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+```
+
+```{r stepwise model macro shannon quad}
+fringing_macro_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + mean_n +  mean_temp_c +  herbivore_evenness + (1|site/location), data = fringing_macro_quad, na.action = na.pass)
+fringing_macro_shannon_dredge <- dredge(fringing_macro_shannon_mod_universal) 
+model.sel(fringing_macro_shannon_dredge) # best just year
+fringing_macro_shannon_top_mod <- glmmTMB(log(shannon + 1) ~ year + (1|site/location), data = fringing_macro_quad, na.action = na.pass)
+hist(residuals(fringing_macro_shannon_top_mod))# almost bimodal
+
+plot(residuals(fringing_macro_shannon_top_mod) ~ fitted(fringing_macro_shannon_top_mod)) # sharp line again
+summary(fringing_macro_shannon_top_mod)
+Anova(fringing_macro_shannon_top_mod)
+
+## Backreef - includes macro
+
+backreef_macro_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + (1|site/location), data = backreef_macro_quad, na.action = na.pass)
+backreef_macro_shannon_dredge <- dredge(backreef_macro_shannon_mod_universal) # some have convergence issues
+model.sel(backreef_macro_shannon_dredge) #  best  removes wave power
+backreef_macro_shannon_top_mod <- glmmTMB(log(shannon + 1) ~ year + mean_n + mean_temp_c +  herbivore_evenness + (1|site/location), data = backreef_macro_quad, na.action = na.pass) #  a little wonky on the tail but not bad
+hist(residuals(backreef_macro_shannon_top_mod)) # bimodal again... zero inflation?
+plot(residuals(backreef_macro_shannon_top_mod) ~ fitted(backreef_macro_shannon_top_mod)) # same stripe
+summary(backreef_macro_shannon_top_mod)
+Anova(backreef_macro_shannon_top_mod) # just year and temp sig
+
+backreef_emm <-as.data.frame(emmeans(backreef_macro_shannon_top_mod, ~mean_temp_c,
+                                     at = list(mean_temp_c = seq(min(backreef_macro_quad$mean_temp_c),
+                                                            max(backreef_macro_quad$mean_temp_c),
+                                                            0.01)))) 
+ggplot()+
+  geom_point(data = backreef_macro_quad, aes(x = mean_temp_c, y = shannon))+
+  geom_line(data = backreef_emm, aes(x = mean_temp_c, y = emmean)) +
+  geom_ribbon(data = backreef_emm, aes(x = mean_temp_c, ymin = lower.CL, ymax = upper.CL, alpha = 0.2), show.legend = F) +
+  scale_y_log10() + 
+  labs(x = "Mean Temp (C)", y = "Backreef Shannon Index (H')") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+# Outer10
+outer10_macro_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp  + mean_temp_c +  herbivore_evenness + (1|site/location), data = outer10_macro_quad, na.action = na.pass)
+outer10_macro_shannon_dredge <- dredge(outer10_macro_shannon_mod_universal) 
+model.sel(outer10_macro_shannon_dredge) #  best includes just herbivore evenness 
+outer10_macro_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  herbivore_evenness + (1|site/location), data = outer10_macro_quad, na.action = na.pass)
+hist(residuals(outer10_macro_shannon_top_mod)) # slight negative skew but better than the others
+plot(residuals(outer10_macro_shannon_top_mod) ~ fitted(outer10_macro_shannon_top_mod)) # looks fine
+summary(outer10_macro_shannon_top_mod)
+Anova(outer10_macro_shannon_top_mod)
+outer10_emm <-as.data.frame(emmeans(outer10_macro_shannon_top_mod, ~herbivore_evenness,
+                                     at = list(herbivore_evenness = seq(min(outer10_macro_quad$herbivore_evenness),
+                                                            max(outer10_macro_quad$herbivore_evenness),
+                                                            0.05))))
+
+ggplot()+
+  geom_point(data = outer10_macro_quad, aes(x = herbivore_evenness, y = shannon))+
+  geom_line(data = outer10_emm, aes(x = herbivore_evenness, y = emmean)) +
+  geom_ribbon(data = outer10_emm, aes(x = herbivore_evenness, ymin = lower.CL, ymax = upper.CL, alpha = 0.2), show.legend = F) +
+  scale_y_log10() + 
+  labs(x = "Herbivore evenness", y = "Outer 10 Shannon Index (H')") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+## outer17
+
+outer17_macro_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp  + mean_temp_c + (1|site/location), data = outer17_macro_quad, na.action = na.pass)
+outer17_macro_shannon_dredge <- dredge(outer17_macro_shannon_mod_universal) 
+model.sel(outer17_macro_shannon_dredge) # best includes all but results in NaNs
+# next best includes just year and mean_temp
+outer17_macro_shannon_top_mod <- glmmTMB(log(shannon + 1) ~ year  + mean_temp_c + (1|site/location), data = outer17_macro_quad, na.action = na.pass)
+hist(residuals(outer17_macro_shannon_top_mod)) # negative skew
+plot(residuals(outer17_macro_shannon_top_mod) ~ fitted(outer17_macro_shannon_top_mod)) # looks fine
+summary(outer17_macro_shannon_top_mod)
+Anova(outer17_macro_shannon_top_mod)
+
+outer17_emm <-as.data.frame(emmeans(outer17_macro_shannon_top_mod, ~mean_temp_c,
+                                     at = list(mean_temp_c = seq(min(outer17_macro_quad$mean_temp_c),
+                                                            max(outer17_macro_quad$mean_temp_c),
+                                                            0.01)))) 
+ggplot()+
+  geom_point(data = outer17_macro_quad, aes(x = mean_temp_c, y = shannon))+
+  geom_line(data = outer17_emm, aes(x = mean_temp_c, y = emmean)) +
+  geom_ribbon(data = outer17_emm, aes(x = mean_temp_c, ymin = lower.CL, ymax = upper.CL, alpha = 0.2), show.legend = F) +
+  scale_y_log10() + 
+  labs(x = "Mean Temp (C)", y = "Outer 17 Shannon Index (H')") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+```
+```{r stepwise model macro betadisp quad}
+fringing_macro_betadisp_mod_universal <- glmmTMB(betadispersion_trans ~ year + mean_n +  mean_temp_c +  herbivore_evenness + (1|site/location), data = fringing_macro_quad, na.action = na.pass, family = beta_family())
+fringing_macro_betadisp_dredge <- dredge(fringing_macro_betadisp_mod_universal) 
+model.sel(fringing_macro_betadisp_dredge) # best includes all
+hist(residuals(fringing_macro_betadisp_mod_universal)) #  a little lumpy but not bad
+plot(residuals(fringing_macro_betadisp_mod_universal) ~ fitted(fringing_macro_betadisp_mod_universal)) # negative trend
+summary(fringing_macro_betadisp_mod_universal)
+Anova(fringing_macro_betadisp_mod_universal) # all sig
+
+fringing_betadisp_emm <-as.data.frame(emmeans(fringing_macro_betadisp_mod_universal, ~mean_n,
+                                     at = list(mean_n = seq(min(fringing_macro_quad$mean_n),
+                                                            max(fringing_macro_quad$mean_n),
+                                                            0.01)))) 
+ggplot()+
+  geom_point(data = fringing_macro_quad, aes(x = mean_n, y = betadispersion_trans))+
+  geom_line(data = fringing_betadisp_emm, aes(x = mean_n, y = inv_logit(emmean))) +
+  geom_ribbon(data = fringing_betadisp_emm, aes(x = mean_n, ymin = inv_logit(asymp.LCL),
+                                                ymax = inv_logit(asymp.UCL), alpha = 0.2), show.legend = F) +
+  labs(x = "Mean N", y = "Fringing Beta Dispersion") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+fringing_betadisp_emm2 <-as.data.frame(emmeans(fringing_macro_betadisp_mod_universal, ~herbivore_evenness,
+                                     at = list(herbivore_evenness = seq(min(fringing_macro_quad$herbivore_evenness),
+                                                            max(fringing_macro_quad$herbivore_evenness),
+                                                            0.01)))) 
+ggplot()+
+  geom_point(data = fringing_macro_quad, aes(x = herbivore_evenness, y = betadispersion_trans))+
+  geom_line(data = fringing_betadisp_emm2, aes(x = herbivore_evenness, y = inv_logit(emmean))) +
+  geom_ribbon(data = fringing_betadisp_emm2, aes(x = herbivore_evenness, ymin = inv_logit(asymp.LCL),
+                                                ymax = inv_logit(asymp.UCL), alpha = 0.2), show.legend = F) +
+  labs(x = "Herbivore evenness", y = "Fringing Beta Dispersion") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+fringing_betadisp_emm3 <-as.data.frame(emmeans(fringing_macro_betadisp_mod_universal, ~mean_temp_c,
+                                     at = list(mean_temp_c = seq(min(fringing_macro_quad$mean_temp_c),
+                                                            max(fringing_macro_quad$mean_temp_c),
+                                                            0.01)))) 
+ggplot()+
+  geom_point(data = fringing_macro_quad, aes(x = mean_temp_c, y = betadispersion_trans))+
+  geom_line(data = fringing_betadisp_emm3, aes(x = mean_temp_c, y = inv_logit(emmean))) +
+  geom_ribbon(data = fringing_betadisp_emm3, aes(x = mean_temp_c, ymin = inv_logit(asymp.LCL),
+                                                ymax = inv_logit(asymp.UCL), alpha = 0.2), show.legend = F) +
+  labs(x = "Mean Temp (C)", y = "Fringing Beta Dispersion") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+## Backreef - includes macro
+
+backreef_macro_betadisp_mod_universal <- glmmTMB(betadispersion_trans ~ year  + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + (1|site/location), data = backreef_macro_quad, na.action = na.pass, family = beta_family())
+backreef_macro_betadisp_dredge <- dredge(backreef_macro_betadisp_mod_universal) # most have convergence issues
+model.sel(backreef_macro_betadisp_dredge) #  best includes nitrogen and year
+backreef_macro_betadisp_top_mod <- glmmTMB(betadispersion_trans ~ year  +  mean_n  + (1|site/location), data = backreef_macro_quad, na.action = na.pass, family = beta_family()) 
+hist(residuals(backreef_macro_betadisp_top_mod)) # not bad
+plot(residuals(backreef_macro_betadisp_top_mod) ~ fitted(backreef_macro_betadisp_top_mod)) # weird cluster, slight negative trend
+summary(backreef_macro_betadisp_top_mod)
+Anova(backreef_macro_betadisp_top_mod)
+
+# Outer10
+outer10_macro_betadisp_mod_universal <- glmmTMB(betadispersion_trans ~ year  + wave_power_adcp  + mean_temp_c +  herbivore_evenness + (1|site/location), data = outer10_macro_quad, na.action = na.pass, family = beta_family()) # model convergence
+outer10_macro_betadisp_dredge <- dredge(outer10_macro_betadisp_mod_universal) # all have errors of some kind
+model.sel(outer10_macro_betadisp_dredge) #  best includes herbivore evenness, temp, and year
+outer10_macro_betadisp_top_mod <- glmmTMB(betadispersion_trans ~ year + mean_temp_c +  herbivore_evenness + (1|site/location), data = outer10_macro_quad, na.action = na.pass, family = beta_family()) 
+qqnorm(residuals(outer10_macro_betadisp_top_mod)) # positive skew
+plot(residuals(outer10_macro_betadisp_top_mod) ~ fitted(outer10_macro_betadisp_top_mod)) # less negative than before
+summary(outer10_macro_betadisp_top_mod)
+Anova(outer10_macro_betadisp_top_mod) # all sig
+outer10_betadisp_emm <-as.data.frame(emmeans(outer10_macro_betadisp_top_mod, ~herbivore_evenness,
+                                     at = list(herbivore_evenness = seq(min(outer10_macro_quad$herbivore_evenness),
+                                                            max(outer10_macro_quad$herbivore_evenness),
+                                                            0.05))))
+
+ggplot()+
+  geom_point(data = outer10_macro_quad, aes(x = herbivore_evenness, y = betadispersion_trans))+
+  geom_line(data = outer10_betadisp_emm, aes(x = herbivore_evenness, y = inv_logit(emmean))) +
+  geom_ribbon(data = outer10_betadisp_emm,
+              aes(x = herbivore_evenness, ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), alpha = 0.2), show.legend = F) +
+  labs(x = "Herbivore evenness", y = "Outer 10 Beta Dispersion") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+outer10_betadisp_emm2 <-as.data.frame(emmeans(outer10_macro_betadisp_top_mod, ~mean_temp_c,
+                                     at = list(mean_temp_c = seq(min(outer10_macro_quad$mean_temp_c),
+                                                            max(outer10_macro_quad$mean_temp_c),
+                                                            0.05))))
+
+ggplot()+
+  geom_point(data = outer10_macro_quad, aes(x = mean_temp_c, y = betadispersion_trans))+
+  geom_line(data = outer10_betadisp_emm2, aes(x = mean_temp_c, y = inv_logit(emmean))) +
+  geom_ribbon(data = outer10_betadisp_emm2,
+              aes(x = mean_temp_c, ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), alpha = 0.2), show.legend = F) +
+  labs(x = "Mean temp (C)", y = "Outer 10 Beta Dispersion") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+## outer17
+
+outer17_macro_betadisp_mod_universal <- glmmTMB(betadispersion_trans ~ year  + wave_power_adcp  + mean_temp_c + (1|site/location), data = outer17_macro_quad, na.action = na.pass, family = beta_family()) # convergence issue
+outer17_macro_betadisp_dredge <- dredge(outer17_macro_betadisp_mod_universal) 
+model.sel(outer17_macro_betadisp_dredge) # best mean temp and year
+outer17_macro_betadisp_top_mod <-  glmmTMB(betadispersion_trans ~ year  +  mean_temp_c + (1|site/location), data = outer17_macro_quad, na.action = na.pass, family = beta_family())
+hist(residuals(outer17_macro_betadisp_top_mod)) # similar positive skew, not bad
+plot(residuals(outer17_macro_betadisp_top_mod) ~ fitted(outer17_macro_betadisp_top_mod)) # similar issues
+summary(outer17_macro_betadisp_top_mod)
+Anova(outer17_macro_betadisp_top_mod) # sig
+
+outer17_betadisp_emm <-as.data.frame(emmeans(outer17_macro_betadisp_top_mod, ~mean_temp_c,
+                                     at = list(mean_temp_c = seq(min(outer17_macro_quad$mean_temp_c),
+                                                            max(outer17_macro_quad$mean_temp_c),
+                                                            0.01)))) 
+ggplot()+
+  geom_point(data = outer17_macro_quad, aes(x = mean_temp_c, y = betadispersion_trans))+
+  geom_line(data = outer17_betadisp_emm, aes(x = mean_temp_c, y = inv_logit(emmean))) +
+  geom_ribbon(data = outer17_betadisp_emm, aes(x = mean_temp_c, ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), alpha = 0.2), show.legend = F) +
+  labs(x = "Mean Temp (C)", y = "Backreef betadisp Index (H')") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank())
+
+```
+
+```{r turnover quad setup}
+# some quadrats don't have algae to run turnover and synchrony functions
+problem_children <- c(
+                      "lter_5_fringing_reef_algae_transect_3_quad_9",
+                      "lter_5_fringing_reef_algae_transect_3_quad_8",
+                      "lter_5_fringing_reef_algae_transect_3_quad_6",
+                      "lter_5_fringing_reef_algae_transect_3_quad_3",
+                      "lter_5_fringing_reef_algae_transect_3_quad_10")
+
+long_data_sub <- long_data[!long_data$location %in% problem_children, ] # remove problem children
+
+data_sub <- data[!data$location %in% problem_children, ]
+
+macroalgae_sub <- macroalgae[!macroalgae$location %in% problem_children, ]
+macro_long_data_sub <- macro_long_data[!macro_long_data$location %in% problem_children, ]
+```
+
+```{r turnover quad all}
+total_turnover<-
+  turnover(
+  df = long_data_sub,
+  time.var = "year",
+  species.var = "taxa",
+  abundance.var = "percent_cover",
+  replicate.var = "location",
+  metric = "total"
+)
+
+# merge with turnover df
+total_turnover <- merge(total_turnover, meta_df, by = c("year", "location"))
+total_turnover$year_factor <- as.factor(total_turnover$year)
+total_turnover$turnover_trans <- sv_trans(total_turnover$total)
+
+turnover_mod_quad <- glmmTMB(turnover_trans ~ year*habitat + (1|site/location), family = beta_family(), data = total_turnover) # same here
+summary(turnover_mod_quad)
+hist(residuals(turnover_mod_quad)) # almost bimodal but not terrible
+plot(residuals(turnover_mod_quad) ~ fitted(turnover_mod_quad)) # negative trend
+Anova(turnover_mod_quad)
+# all P < 2e-16  
+emtrends(turnover_mod_quad, pairwise ~ habitat, var = "year") # fringing not different from forereef
+turnover_quad_emm <- emmip(turnover_mod_quad,
+                        habitat ~ year,
+                        at = list(year = seq(2008,2023,1)), plotit = F, CIs = T)
+
+(turnover_quad_plot <- 
+  ggplot() +
+  geom_point(data = total_turnover,
+             aes(x = year, y = turnover_trans, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = turnover_quad_emm, aes(x = year, y = inv_logit(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = turnover_quad_emm, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL),
+                                                  fill = habitat), alpha = 0.3) +
+  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+    labs(x = "Year", y = "Total Turnover\n") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) )
+
+```
+```{r diversity synchrony quad all setup}
+lm_synchrony_quadrat <- codyn::synchrony(df = long_data_sub,
+                                   time.var = "year",
+                                   species.var = "taxa",
+                                   abundance.var = "percent_cover",
+                                   metric = "Loreau",
+                                   replicate.var = "location") # Loreau and Mazancourt synchrony
+colnames(lm_synchrony_quadrat) = c("location", "loreau_synchrony") # rename cols for merge later
+diversity_stability_quadrat <- alpha_diversity_quad %>%
+  dplyr::group_by(location, site, habitat, island_side, site_habitat, transect) %>%
+  dplyr::summarise(
+            algae_richness_mean = mean(algae_richness_mean), 
+            algae_shannon_mean = mean(algae_shannon_mean), 
+            algae_cover_mean = mean(algae_cover_mean),
+            algae_cover_cv = CV(algae_cover_hard_mean))
+
+diversity_stability_quadrat <- 
+  merge(diversity_stability_quadrat, lm_synchrony_quadrat, by = "location") # add in loreau and de Mazancourt synchrony
+diversity_stability_quadrat$loreau_synchrony_trans <- sv_trans(diversity_stability_quadrat$loreau_synchrony)
+
+# for plotting later
+diversity_ranges_quad <- diversity_stability_quadrat %>%
+  group_by(habitat) %>%
+  summarise(min_richness = min(algae_richness_mean),
+            max_richness = max(algae_richness_mean),
+            min_shannon = min(algae_shannon_mean),
+            max_shannon = max(algae_shannon_mean),
+            min_loreau = min(loreau_synchrony),
+            max_loreau = max(loreau_synchrony))
+
+```
+
+```{r diversity stability synchrony all quad, fig.height=8, fig.width = 8}
+diversity_cv_quad_mod <- glmmTMB(algae_cover_cv ~ algae_shannon_mean*habitat + (1|site), data = diversity_stability_quadrat, family = Gamma(link = "log"))
+hist(residuals(diversity_cv_quad_mod)) # a little chunky
+plot(residuals(diversity_cv_quad_mod) ~ fitted(diversity_cv_quad_mod)) # some heteroskedasticity
+summary(diversity_cv_quad_mod)
+Anova(diversity_cv_quad_mod)
+
+emtrends(diversity_cv_quad_mod, pairwise ~ habitat, var = "algae_shannon_mean") # backreef not different from forereef
+
+shannon_cv_emm_quad <- emmip(diversity_cv_quad_mod,
+                        habitat ~ algae_shannon_mean,
+                        at = list(algae_shannon_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_shannon_effects_quad <- shannon_cv_emm_quad %>%
+  left_join(diversity_ranges_quad, by = c("habitat")) %>%
+  filter(algae_shannon_mean >= min_shannon & algae_shannon_mean <= max_shannon)
+
+(shannon_cv_plot_quad = 
+  ggplot() +
+  geom_point(data = diversity_stability_quadrat,
+             aes(x = algae_shannon_mean, y = algae_cover_cv, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = filtered_shannon_effects_quad, aes(x = algae_shannon_mean, y = exp(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = filtered_shannon_effects_quad, aes(x = algae_shannon_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+diversity_synchrony_quad_mod <- glmmTMB(loreau_synchrony_trans ~ algae_shannon_mean*habitat + (1|site), data = diversity_stability_quadrat, family = beta_family())
+hist(residuals(diversity_synchrony_quad_mod)) # not bad
+plot(residuals(diversity_synchrony_quad_mod) ~ fitted(diversity_synchrony_quad_mod)) # heteroskedasticity
+summary(diversity_synchrony_quad_mod)
+Anova(diversity_synchrony_quad_mod)
+
+emtrends(diversity_synchrony_quad_mod, pairwise ~ habitat, var = "algae_shannon_mean") # backreef not different from forereef
+
+shannon_synchrony_emm_quad <- emmip(diversity_synchrony_quad_mod,
+                        habitat ~ algae_shannon_mean,
+                        at = list(algae_shannon_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_shannon_synchrony_effects_quad <- shannon_synchrony_emm_quad %>%
+  left_join(diversity_ranges_quad, by = c("habitat")) %>%
+  filter(algae_shannon_mean >= min_shannon & algae_shannon_mean <= max_shannon)
+
+(shannon_synchrony_plot_quad = 
+  ggplot() +
+  geom_point(data = diversity_stability_quadrat,
+             aes(x = algae_shannon_mean, y = loreau_synchrony_trans, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = filtered_shannon_synchrony_effects_quad, aes(x = algae_shannon_mean, y = inv_logit(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = filtered_shannon_synchrony_effects_quad, aes(x = algae_shannon_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
+              alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+synchrony_cv_quad_mod <- glmmTMB(algae_cover_cv ~ loreau_synchrony*habitat + (1|site), data = diversity_stability_quadrat, family = Gamma(link = "log"))
+hist(residuals(synchrony_cv_quad_mod)) # not bad
+plot(residuals(synchrony_cv_quad_mod) ~ fitted(synchrony_cv_quad_mod)) # sstripey
+summary(synchrony_cv_quad_mod)
+Anova(synchrony_cv_quad_mod)
+
+emtrends(synchrony_cv_quad_mod, pairwise ~ habitat, var = "loreau_synchrony") # backreef not different from forereef
+
+synchrony_cv_emm_quad <- emmip(synchrony_cv_quad_mod,
+                        habitat ~ algae_synchrony_mean,
+                        at = list(algae_synchrony_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_synchrony_effects_quad <- synchrony_cv_quad_mod %>%
+  left_join(diversity_ranges_quad, by = c("habitat")) %>%
+  filter(algae_synchrony_mean >= min_synchrony & algae_synchrony_mean <= max_synchrony)
+
+(synchrony_cv_plot_quad = 
+  ggplot() +
+  geom_point(data = diversity_stability_quadrat,
+             aes(x = algae_synchrony_mean, y = algae_cover_cv, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = filtered_synchrony_effects_quad, aes(x = algae_synchrony_mean, y = exp(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = filtered_synchrony_effects_quad, aes(x = algae_synchrony_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+
+```
+
+```{r total turnover macro}
+total_turnover_macro<-
+  turnover(
+  df = macro_long_data_sub,
+  time.var = "year",
+  species.var = "taxa",
+  abundance.var = "percent_cover",
+  replicate.var = "location",
+  metric = "total"
+)
+
+# merge with turnover df
+total_turnover_macro <- merge(total_turnover_macro, meta_df, by = c("year", "location"))
+total_turnover_macro$year_factor <- as.factor(total_turnover_macro$year)
+total_turnover_macro$turnover_trans <- sv_trans(total_turnover_macro$total)
+
+turnover_mod_quad_macro <- glmmTMB(turnover_trans ~ year*habitat + (1|site/location), family = beta_family(), data = total_turnover_macro) # same here
+summary(turnover_mod_quad_macro)
+hist(residuals(turnover_mod_quad_macro)) # almost bimodal but not terrible
+plot(residuals(turnover_mod_quad_macro) ~ fitted(turnover_mod_quad_macro)) # negative trend
+Anova(turnover_mod_quad_macro)
+# all P < 2e-16  
+emtrends(turnover_mod_quad_macro, pairwise ~ habitat, var = "year") # fringing not different from forereef
+turnover_quad_macro_emm <- emmip(turnover_mod_quad_macro,
+                        habitat ~ year,
+                        at = list(year = seq(2008,2023,1)), plotit = F, CIs = T)
+
+(turnover_quad_macro_plot <- 
+  ggplot() +
+  geom_point(data = total_turnover_macro,
+             aes(x = year, y = turnover_trans, colour = habitat), size = 2, alpha = 0.5) + 
+  geom_line(data = turnover_quad_macro_emm, aes(x = year, y = inv_logit(yvar), colour = habitat), size = 2) +
+  geom_ribbon(data = turnover_quad_macro_emm, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL),
+                                                  fill = habitat), alpha = 0.3) +
+  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+    labs(x = "Year", y = "Total Turnover\n") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) )
+
+```

### 2024-09-11 R/fish_plots.R
expecting a git issue

@@ -20,6 +20,7 @@ library(car)
 #data toime
 new_diversity <- read.csv(here("data", "algae_diversity_site.csv"))
 fish <- read.csv(here("data", "biomass_herb_summary.csv"))
+fish_fun <- read.csv(here("data", "biomass_functional_summary.csv"))
 
 #merge em
 #truly i will not forgive y'all for making all them things lowercase
@@ -122,10 +123,75 @@ herb_beta_model <- lmer(centroid_dist_habyear_mean ~ log10(Herbivore.Biomass) +
 
 Anova(herb_beta_model)
 
+
+### Repeat it with functional groups
+
+#merging
+#truly i will not forgive y'all for making all them things lowercase
+fish_fun[is.na(fish_fun)] <- 0
+fish_fun$year <- fish$Year
+fish_fun$habitat <- tolower(fish$Habitat)
+fish_fun$habitat <- gsub("outer 10", "outer_10", fish$habitat)
+fish_fun <- fish_fun %>% 
+  mutate(site = paste0("lter_", Site))
+alg_div_fun_fish <- left_join(new_diversity, fish_fun[,c(4:22)], by = join_by(year, site, habitat))
+merged_div <- alg_div_fun_fish[alg_div_fun_fish$habitat != "outer_17", ]
+
+
+axis_convert <- max(merged_div$Browser,na.rm = TRUE) / max(merged_div$algae_cover_hard_mean,na.rm = TRUE)
+browser_cover <- ggplot(merged_div, aes(x = year)) +
+  geom_line(aes(y = algae_cover_hard_mean, color = "Algae Cover of Hard Substrate")) +
+  geom_line(aes(y = Browser / axis_convert, color = "Herbivore Biomass")) + # Rescale Herbivore.Biomass
+  facet_grid(site ~ habitat) +
+  scale_y_continuous(
+    name = "Mean Algae Cover of Hard Substrate (%)",
+    sec.axis = sec_axis(~ . * axis_convert, name = "Browser Biomass") # Rescale back for labeling
+  ) +
+  scale_color_manual(values = c("Algae Cover of Hard Substrate" = "darkgreen", "Herbivore Biomass" = "darkblue")) +
+  labs(x = "Year", color = "Legend", title = "Trends in Algal Cover and Herbivore Biomass" # Add the plot title here
+  ) +
+  theme_bw() +
+  theme(legend.position = "bottom")
+
+browser_cover
 #YOU THOUGHT I WAS DONE?
 #no
 
[…]

### 2024-09-11 manuscript_1_code.Rmd
created figure 1 v1

---
title: "Manuscript 1 code"
author: "Noam Altman-Kurosaki"
date: "2024-09-11"
output: html_document
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(here)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(vegan)
library(glmmTMB)
library(car)
library(effects)
library(DHARMa)
library(stringr)
library(MuMIn)
library(codyn)
library(nlme)
library(mgcv)
library(emmeans)
```

## functions

```{r functions}
# coefficient of variation
CV<-function(x){
  return(sd(x,na.rm=T)/mean(x,na.rm=T))
}

inv_logit <- function(x){
  exp(x)/(1+exp(x))
}

# smithson verkuilen transformation for 0-1 continuous data
sv_trans <- function(prop, s = 0.000005){
  (prop*(length(prop) - 1) + s)/length(prop)
  # where prop is the vector of the proportional value you're transforming, 
  # N is the sample size, which is specified by taking the number of rows/observations from a given dataframe,
  # and s is a small offset 
}

se <- function(x){
  sd(x)/sqrt(n())
}

```

## Figure formatting
```{r figure formatting info}
habitat_colours <- c("fringing" = "#D81B60",
                     "backreef" = "#FFC107",
                     "outer_10" = "#1E88E5",
                     "outer_17" = "#004D40")

functional_group_colours <- c("coral" = "coral",
                         "cca" = "pink",
                         "maroalgae" = "forestgreen",
                         "turf" =  "darkolivegreen")

```

# read and clean data

```{r read and clean data}
data <- read.csv(here::here("data", "allbenthicdata_cleaned_colnames.csv"), stringsAsFactors = F)

not_algae <- c("ascidian", "bare_space", "coral", "coral_rubble", "corallimorpharia", "heteractis_sp", "millepora_platyphylla", "no_data", "sand", "sarcophyton_sp", "shell_debris", "soft_coral", "sponge", "tridacna_sp", "zooxanthellae", "row_sums")

algae <- data[ , -which(names(data) %in% not_algae)]
algae$algal_turf <- data$coral_rubble + data$bare_space + data$algal_turf

macroalgae <- algae %>% 
  select(-algal_turf, -crustose_corallines, -cyanophyta, -damselfish_turf, -lithophyllum_kotschyanum,
         -neogoniolithon_brassica_florida, -sporolithon_sp)

meta_cols <- c("year", "location", "site", "habitat", "site_habitat", "transect", "quadrat")

meta_df <- algae[,which(names(algae) %in% meta_cols)]

algae_species_cols <- algae[,-which(names(algae) %in% meta_cols)]

macro_species_cols <- macroalgae[,-which(names(macroalgae) %in% meta_cols)]

```

# Figure 1

## setup

```{r cover plot setup}
cover_df <- data.frame(
  year = data$year,
  habitat = data$habitat,
  site = data$site,
  location = data$location,
  coral = data$coral,
  cca = data$crustose_corallines + data$lithophyllum_kotschyanum + data$neogoniolithon_brassica_florida + data$sporolithon_sp,
  macroalgae = rowSums(macro_species_cols),
  turf = data$coral_rubble + data$algal_turf + data$bare_space + data$damselfish_turf + data$cyanophyta,
  sand = data$sand,
  other = data$ascidian + data$corallimorpharia + data$heteractis_sp + data$millepora_platyphylla + data$sarcophyton_sp + data$shell_debris +
    data$soft_coral + data$sponge + data$tridacna_sp + data$zooxanthellae + data$no_data
)

cover_df$total <- rowSums(cover_df[,-which(names(cover_df) %in% meta_cols)])

cover_means <- cover_df %>%
  group_by(year, habitat) %>%
  summarise(macroalgae_mean = mean(macroalgae),
            coral_mean = mean(coral),
            turf_mean = mean(turf),
            cca_mean = mean(cca),
            sand_mean = mean(sand),
            other_mean = mean(other)
            )

cover_se <- cover_df %>%
  group_by(year, habitat) %>%
  summarise(macroalgae_se = se(macroalgae),
            coral_se = se(coral),
            turf_se = se(turf),
            cca_se = se(cca),
            sand_se = se(sand),
            other_se = se(other)
            )

cover_long <- pivot_longer(
  data = cover_means,
  cols = 3:8,
  names_to = "functional_group",
  values_to = "percent_cover"
)

cover_long_se <- pivot_longer(
  data = cover_se,
  cols = 3:8,
  names_to = "functional_group",
  values_to = "se"
)

cover_long$se <- cover_long_se$se

simplified_cover <- data.frame(
  year = cover_df$year,
  habitat = cover_df$habitat,
  site = cover_df$site,
  location = cover_df$location,
  low_lying_algae = cover_df$cca + cover_df$turf + cover_df$cyanophyta,
  coral = cover_df$coral,
  macroalgae = cover_df$macroalgae
)

simplified_cover_means <- simplified_cover %>%
  group_by(year, habitat) %>%
  summarise(macroalgae_mean = mean(macroalgae),
            low_lying_algae_mean = mean(low_lying_algae),
            coral_mean = mean(coral),
            macroalgae_se = sd(macroalgae)/sqrt(n()),
            low_lying_algae_se = sd(low_lying_algae)/sqrt(n()),
            coral_se = sd(coral)/sqrt(n())
            )

simplified_cover_long <- pivot_longer(
  data = simplified_cover_means,
  cols = 3:5,
  names_to = "functional_group",
  values_to = "percent_cover"
)

simplified_cover_long_se <- pivot_longer(
  data = simplified_cover_means,
  cols = 6:8,
  names_to = "functional_group",
  values_to = "se"
)
simplified_cover_long$se <- simplified_cover_long_se$se
```

## plot

```{r figure 1, fig.height = 10, fig.width=14}
(figure_1 <- ggplot(data = subset(cover_long, functional_group != "sand_mean" & functional_group != "other_mean")) +
  geom_line( aes(x = year, y = percent_cover, colour = functional_group), size = 1) +
  geom_ribbon( aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, fill = functional_group), alpha = 0.3) +
  facet_wrap(~ habitat) +
  geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
#  scale_colour_manual(values = functional_group_colours) +
#  scale_fill_manual(values = functional_group_colours) +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

```

### 2024-09-11 manuscript_1_code.Rmd
Edits to Fig1 code

@@
 
-functional_group_colours <- c("coral" = "coral",
-                         "cca" = "pink",
-                         "maroalgae" = "forestgreen",
-                         "turf" =  "darkolivegreen")
+functional_group_colours <- c("coral_mean" = "#56B4E9",
+                         "cca_mean" = "#CC79A7",
+                         "macroalgae_mean" = "#009E73",
+                         "turf" =  "#D55E00")
 
@@
 
-simplified_cover <- data.frame(
-  year = cover_df$year,
-  habitat = cover_df$habitat,
-  site = cover_df$site,
-  location = cover_df$location,
-  low_lying_algae = cover_df$cca + cover_df$turf + cover_df$cyanophyta,
-  coral = cover_df$coral,
-  macroalgae = cover_df$macroalgae
-)
+cover_long$habitat <- factor(cover_long$habitat, 
+                             levels = c("fringing", "backreef", "outer_10","outer_17"),
+                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
 
-simplified_cover_means <- simplified_cover %>%
-  group_by(year, habitat) %>%
-  summarise(macroalgae_mean = mean(macroalgae),
-            low_lying_algae_mean = mean(low_lying_algae),
-            coral_mean = mean(coral),
-            macroalgae_se = sd(macroalgae)/sqrt(n()),
-            low_lying_algae_se = sd(low_lying_algae)/sqrt(n()),
-            coral_se = sd(coral)/sqrt(n())
-            )
+# simplified_cover <- data.frame(
+#   year = cover_df$year,
+#   habitat = cover_df$habitat,
+#   site = cover_df$site,
+#   location = cover_df$location,
+#   low_lying_algae = cover_df$cca + cover_df$turf + cover_df$cyanophyta,
+#   coral = cover_df$coral,
+#   macroalgae = cover_df$macroalgae
+# )
+# 
+# simplified_cover_means <- simplified_cover %>%
+#   group_by(year, habitat) %>%
+#   summarise(macroalgae_mean = mean(macroalgae),
+#             low_lying_algae_mean = mean(low_lying_algae),
+#             coral_mean = mean(coral),
+#             macroalgae_se = sd(macroalgae)/sqrt(n()),
+#             low_lying_algae_se = sd(low_lying_algae)/sqrt(n()),
+#             coral_se = sd(coral)/sqrt(n())
+#             )
+# 
+# simplified_cover_long <- pivot_longer(
+#   data = simplified_cover_means,
+#   cols = 3:5,
+#   names_to = "functional_group",
+#   values_to = "percent_cover"
+# )
+# 
+# simplified_cover_long_se <- pivot_longer(
+#   data = simplified_cover_means,
+#   cols = 6:8,
+#   names_to = "functional_group",
+#   values_to = "se"
+# )
+# simplified_cover_long$se <- simplified_cover_long_se$se
 
-simplified_cover_long <- pivot_longer(
-  data = simplified_cover_means,
-  cols = 3:5,
-  names_to = "functional_group",
-  values_to = "percent_cover"
-)
-
-simplified_cover_long_se <- pivot_longer(
-  data = simplified_cover_means,
-  cols = 6:8,
-  names_to = "functional_group",
-  values_to = "se"
-)
-simplified_cover_long$se <- simplified_cover_long_se$se
 ```
@@
 (figure_1 <- ggplot(data = subset(cover_long, functional_group != "sand_mean" & functional_group != "other_mean")) +
-  geom_line( aes(x = year, y = percent_cover, colour = functional_group), size = 1) +
+  geom_line(aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
   geom_ribbon( aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, fill = functional_group), alpha = 0.3) +
@@
   geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
-#  scale_colour_manual(values = functional_group_colours) +
-#  scale_fill_manual(values = functional_group_colours) +
+  scale_colour_manual(values = functional_group_colours) +
+  scale_fill_manual(values = functional_group_colours) +
   geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
   geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+   scale_x_continuous(breaks = unique(cover_long$year)) +
+   labs(y = "Percent Cover \n",
+        x = "Year") +
   theme_bw() + 
   theme(axis.title.x = element_text(size=18), 
+        axis.text.x = element_text(angle = 45, hjust = 1),
         axis.title.y = element_text(size=18), 
@@
 
+ggsave(filename = "output/figure_1.png", figure_1, height = 10, width = 14)
+
 ```

### 2024-09-11 manuscript_1_code.Rmd
changed fig 1 and how join cover_long

@@
 
-functional_group_colours <- c("coral_mean" = "#56B4E9",
-                         "cca_mean" = "#CC79A7",
-                         "macroalgae_mean" = "#009E73",
-                         "turf" =  "#D55E00")
+functional_group_colours <- c("coral" = "darkmagenta",
+                         "cca" = "pink",
+                         "macroalgae" = "#004D40",
+                         "turf" =  "#1E88E5")
 
@@
   group_by(year, habitat) %>%
-  summarise(macroalgae_mean = mean(macroalgae),
-            coral_mean = mean(coral),
-            turf_mean = mean(turf),
-            cca_mean = mean(cca),
-            sand_mean = mean(sand),
-            other_mean = mean(other)
+  summarise(macroalgae = mean(macroalgae),
+            coral = mean(coral),
+            turf = mean(turf),
+            cca = mean(cca),
+            sand = mean(sand),
+            other = mean(other)
             )
@@
   group_by(year, habitat) %>%
-  summarise(macroalgae_se = se(macroalgae),
-            coral_se = se(coral),
-            turf_se = se(turf),
-            cca_se = se(cca),
-            sand_se = se(sand),
-            other_se = se(other)
+  summarise(macroalgae = se(macroalgae),
+            coral = se(coral),
+            turf = se(turf),
+            cca = se(cca),
+            sand = se(sand),
+            other = se(other)
             )
+```
 
-cover_long <- pivot_longer(
+```{r cover long}
+cover_long_means <- pivot_longer(
   data = cover_means,
@@
 
-cover_long$se <- cover_long_se$se
+cover_long <- left_join(cover_long_means,cover_long_se)
 
+#make levels for habitat for plotting
 cover_long$habitat <- factor(cover_long$habitat, 
@@
                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+```
+
+```{r simplified cover - remove}
 
@@
 
-ggsave(filename = "output/figure_1.png", figure_1, height = 10, width = 14)
+#ggsave(filename = "output/figure_1.png", figure_1, height = 10, width = 14)
 

### 2024-09-11 manuscript_1_code.Rmd
whoops sorry before still had sand and other

@@
 ```{r figure 1, fig.height = 10, fig.width=14}
-(figure_1 <- ggplot(data = subset(cover_long, functional_group != "sand_mean" & functional_group != "other_mean")) +
+(figure_1 <- ggplot(data = subset(cover_long, functional_group != "sand" & functional_group != "other")) +
   geom_line(aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
@@
 #ggsave(filename = "output/figure_1.png", figure_1, height = 10, width = 14)
-
 ```

### 2024-09-11 manuscript_1_code.Rmd
fixed legend and colors (please make colors pretty)

@@
 
-functional_group_colours <- c("coral" = "darkmagenta",
-                         "cca" = "pink",
-                         "macroalgae" = "#004D40",
-                         "turf" =  "#1E88E5")
+functional_group_colours <- c("Coral" = "darkmagenta",
+                         "CCA" = "deeppink3",
+                         "Macroalgae" = "#004D40",
+                         "Turf" =  "chartreuse4")
 
@@
                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+
+cover_long$functional_group <- factor(cover_long$functional_group, 
+                             levels = c("macroalgae", "coral", "turf","cca", "sand", "other"),
+                             labels = c("Macroalgae", "Coral", "Turf", "CCA", "Sand", "Other"))
 ```
@@
 ```{r figure 1, fig.height = 10, fig.width=14}
-(figure_1 <- ggplot(data = subset(cover_long, functional_group != "sand" & functional_group != "other")) +
+(figure_1 <- ggplot(data = subset(cover_long, functional_group != "Sand" & functional_group != "Other")) +
   geom_line(aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +

### 2024-09-11 manuscript_1_code.Rmd
NAK incorporated discussion from 9/11 AM on functional group classification

@@
   macroalgae = rowSums(macro_species_cols),
-  turf = data$coral_rubble + data$algal_turf + data$bare_space + data$damselfish_turf + data$cyanophyta,
+  turf = data$algal_turf  + data$damselfish_turf, 
   sand = data$sand,
+  coral_rubble = data$coral_rubble,
+  cyanophyta = data$cyanophyta,
   other = data$ascidian + data$corallimorpharia + data$heteractis_sp + data$millepora_platyphylla + data$sarcophyton_sp + data$shell_debris +
-    data$soft_coral + data$sponge + data$tridacna_sp + data$zooxanthellae + data$no_data
+    data$soft_coral + data$sponge + data$tridacna_sp + data$zooxanthellae + data$no_data + data$bare_space
 )
@@
 
-#ggsave(filename = "output/figure_1.png", figure_1, height = 10, width = 14)
+ggsave(filename = "output/figure_1.png", figure_1, height = 10, width = 14)
 ```

### 2024-09-11 manuscript_1_code.Rmd
created fig 2

@@
 ```{r figure formatting info}
-habitat_colours <- c("fringing" = "#D81B60",
-                     "backreef" = "#FFC107",
-                     "outer_10" = "#1E88E5",
-                     "outer_17" = "#004D40")
+habitat_colours <- c("Fringing" = "#D81B60",
+                     "Backreef" = "#FFC107",
+                     "Forereef 10m" = "#1E88E5",
+                     "Forereef 17m" = "#004D40")
 
@@
                          "CCA" = "deeppink3",
-                         "Macroalgae" = "#004D40",
-                         "Turf" =  "chartreuse4")
+                         "Macroalgae" = "chartreuse4",
+                         "Turf" =  "#004D40")
 
@@
 algae <- data[ , -which(names(data) %in% not_algae)]
-algae$algal_turf <- data$coral_rubble + data$bare_space + data$algal_turf
 
@@
 
+```{r}
+figure_1_data <- subset(cover_long, functional_group != "Sand" & functional_group != "Other")
+```
+
 ## plot
 
-```{r figure 1, fig.height = 10, fig.width=14}
-(figure_1 <- ggplot(data = subset(cover_long, functional_group != "Sand" & functional_group != "Other")) +
-  geom_line(aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
-  geom_ribbon( aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, fill = functional_group), alpha = 0.3) +
+```{r figure 1, fig.height = 10, fig.width=14, warning=FALSE}
+(figure_1 <- ggplot() +
+  geom_line(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
+  geom_ribbon(data = figure_1_data, aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, fill = functional_group), alpha = 0.3) +
   facet_wrap(~ habitat) +
-  geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
+  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
+  annotate("segment", x = 2006, xend = 2010, y = 75, yend = 75, linewidth = 2, colour = "grey") +
   scale_colour_manual(values = functional_group_colours) +
@@
         legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
+        legend.title = element_blank(),
         legend.position = 'bottom', # move legend to bottom
@@
 
-ggsave(filename = "output/figure_1.png", figure_1, height = 10, width = 14)
+#ggsave(filename = "output/figure_1.png", figure_1, height = 10, width = 14)
 ```
+
+# Figure 2
+
+## Create Data Frames
+```{r create data}
+alpha_diversity_quad_macro <-
+  data.frame(
+    year = macroalgae$year,
+    location = macroalgae$location,
+    shannon = diversity(macroalgae[ , -which(names(macroalgae) %in% meta_cols)], "shannon"),
+    richness = rowSums(macroalgae[ , -which(names(macroalgae) %in% meta_cols)] > 0)
+  )
+
+alpha_diversity_quad_macro <- merge(meta_df, alpha_diversity_quad_macro, by = c("location", "year"))
+
+macroalgal_cover <- cover_df[,c("year", "habitat", "site", "location", "macroalgae")]
+
+alpha_diversity_quad_macro_with_cover <- left_join(alpha_diversity_quad_macro, macroalgal_cover,by = join_by(location, year, site, habitat)) %>% 
+  mutate(macroalgal_prop_cover = macroalgae / 100) %>% 
+  select(-macroalgae)
+```
+
+# Functional Group Prep
+
+```{r richness df}
+richness_df <- alpha_diversity_quad_macro %>% 
+  group_by(year, habitat) %>% 
+  summarise(richness_mean = mean(richness),
+            richness_se = se(richness))
+```
+
+```{r richness factors}
+richness_df$habitat <- factor(richness_df$habitat, 
+                             levels = c("fringing", "backreef", "outer_10","outer_17"),
+                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+```
+
+## plot
+```{r figure 2}
+(figure_2 <- ggplot() +
+  geom_line(data = richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
+  geom_ribbon(data = richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, fill = habitat), alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
+  annotate("segment", x = 2006, xend = 2010, y = 2, yend = 2, linewidth = 2, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+   scale_x_continuous(breaks = unique(richness_df$year)) +
+   labs(y = "Macroalgal Richness \n",
+        x = "Year") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) )
+
+ggsave(filename = "output/figure_2.png", figure_2, height = 10, width = 14)
+```
+
+## SYNCHRONY CODE
+```{r calculate syncrony}
+# make er long
+macro_long_data <- pivot_longer(
+  data = macroalgae,
+  cols = grep("acanthophora_spicifera", colnames(macroalgae)):grep("valonia_ventricosa", colnames(macroalgae)),
+  names_to = "taxa",
+  values_to = "percent_cover") %>% 
+  mutate(prop_cover = percent_cover/100) %>% 
+  select(-percent_cover)
+
+# write_csv(macro_long_data, here("Data", "macro_long_data.csv"))
+
+# CALCULATE SYNCHRONY CODE
+lm_synchrony <- codyn::synchrony(df = macro_long_data,
+                                   time.var = "year",
+                                   species.var = "taxa",
+                                   abundance.var = "prop_cover",
+                                   metric = "Loreau",
+                                   replicate.var = "location")
+```

### 2024-09-11 manuscript_1_code.Rmd
Models and figures for figure 4 :call_me_hand:

@@
 
-# Functional Group Prep
+# Figure 2
+
+## prep
 
@@
 
-## SYNCHRONY CODE
+# Figure 3
+
+## model
+
+```{r figure 3 model}
+alpha_diversity_quad_macro_with_cover$cover_trans <- sv_trans(alpha_diversity_quad_macro_with_cover$macroalgal_prop_cover)
+
+cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro_with_cover)
+summary(cover_mod)
+hist(residuals(cover_mod)) # looks fine
+plot(residuals(cover_mod) ~ fitted(cover_mod)) # negative trend but it can't be less than 0 so not too concerned. some wave pattern
+r.squaredGLMM(cover_mod) # 0.631, 0.650
+emtrends(cover_mod, pairwise ~ habitat, var = "richness") # backreef not different from forereef, forereef also not different
+
+```
+
+## figure 3 plot setup
+
+```{r figure 3 setup}
+diversity_ranges_quad <- alpha_diversity_quad_macro_with_cover %>%
+  group_by(habitat) %>%
+  summarise(min_richness = min(richness),
+            max_richness = max(richness),
+            min_cover= min(cover_trans),
+            max_cover = max(cover_trans))
+
+cover_emm <- emmip(cover_mod,
+                        habitat ~ richness,
+                        at = list(richness = seq(0,10,1)), plotit = F, CIs = T)
+
+filtered_cover_effects <- cover_emm %>%
+  left_join(diversity_ranges_quad, by = c("habitat")) %>%
+  filter(richness >= min_richness & richness <= max_richness)
+
+alpha_diversity_quad_macro_with_cover$habitat <- factor(alpha_diversity_quad_macro_with_cover$habitat, 
+                             levels = c("fringing", "backreef", "outer_10","outer_17"),
+                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+
+filtered_cover_effects$habitat <- factor(filtered_cover_effects$habitat, 
+                             levels = c("fringing", "backreef", "outer_10","outer_17"),
+                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+```
+
+## figure 3 plot
+
+```{r figure 3 plot}
+(figure_3 <- 
+  ggplot() +
+  geom_point(data = alpha_diversity_quad_macro_with_cover,
+             aes(x = richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
+  geom_line(data = filtered_cover_effects, aes(x = richness, y = inv_logit(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_cover_effects, aes(x = richness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Cover (Proprtion)\n", x = "Macroalgal Richness") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+```
+
+```{r figure 3 v2}
+(figure_3_v2 <- figure_3 + facet_wrap(~habitat))
+```
+# Figure 4
+
+## Synchrony setup
 ```{r calculate syncrony}
@@
                                    replicate.var = "location")
+
+diversity_stability <- alpha_diversity_quad_macro_with_cover %>%
+  dplyr::group_by(location, habitat, site) %>%
+  dplyr::summarise(
+            richness_mean = mean(richness), 
+            shannon_mean = mean(shannon), 
+            cover_mean = mean(macroalgal_prop_cover),
+            cover_cv = CV(macroalgal_prop_cover),
+            cover_stability = 1/CV(macroalgal_prop_cover))
+
+diversity_stability_synchrony <- left_join(diversity_stability, lm_synchrony, by = join_by(location))
+diversity_stability_synchrony$synchrony_trans <- sv_trans(diversity_stability_synchrony$synchrony)
+diversity_stability_synchrony <- na.omit(diversity_stability_synchrony) # remove quads that never had algae --> NaN calcs
+
+dss_ranges<- diversity_stability_synchrony %>%
+  group_by(habitat) %>%
+  summarise(min_richness = min(richness_mean),
+            max_richness = max(richness_mean),
+            min_stability= min(cover_stability),
+            max_stability = max(cover_stability),
+            min_synchrony = min(synchrony_trans),
+            max_synchrony = max(synchrony_trans))
 ```
+
+## figure 4a model - richness stability
+
+```{r richness synchrony mod}
+rich_stab_mod <- glmmTMB(cover_stability~ richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = diversity_stability_synchrony)
+summary(rich_stab_mod)
+hist(residuals(rich_stab_mod)) # looks good
+plot(residuals(rich_stab_mod) ~ fitted(rich_stab_mod)) # heteroskedastic - bring to group
+r.squaredGLMM(rich_stab_mod) # 0.803, 0.808
+emtrends(rich_stab_mod, pairwise ~ habitat, var = "richness_mean") # fringing not diff from outer17, backreef not different from outer10
+```
+
+```{r figure 4a setup}
+
+rich_stab_emm <- emmip(rich_stab_mod,
+                        habitat ~ richness_mean,
+                        at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_rich_stab_effects <- rich_stab_emm %>%
+  left_join(dss_ranges, by = c("habitat")) %>%
+  filter(richness_mean >= min_richness & richness_mean <= max_richness)
+
+```
+
+## plot figure 4a
+
+```{r figure 4a}
+(figure_4a <- 
+  ggplot() +
+  geom_point(data = diversity_stability_synchrony,
+             aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_rich_stab_effects, aes(x = richness_mean, y = exp(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_rich_stab_effects, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Stability\n", x = "Macroalgal Richness") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+```
+
+## figure 4c model - richness synchrony
+
+```{r richness synchrony mod}
+rich_synch_mod <- glmmTMB(synchrony_trans ~ richness_mean*habitat + (1|site), family = beta_family(), data = diversity_stability_synchrony)
+summary(rich_synch_mod)
+hist(residuals(rich_synch_mod)) # looks good
+plot(residuals(rich_synch_mod) ~ fitted(rich_synch_mod)) # same issue as always
+r.squaredGLMM(rich_synch_mod) # 0.701, 0.775
+emtrends(rich_synch_mod, pairwise ~ habitat, var = "richness_mean") # forereefs aren't different
+```
+
+```{r figure 4c setup}
+
+rich_synch_emm <- emmip(rich_synch_mod,
+                        habitat ~ richness_mean,
+                        at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_rich_synch_effects <- rich_synch_emm %>%
+  left_join(dss_ranges, by = c("habitat")) %>%
+  filter(richness_mean >= min_richness & richness_mean <= max_richness)
+
+```
+
+## plot figure 4c
+
+```{r figure 4c}
+(figure_4c <- 
+  ggplot() +
+  geom_point(data = diversity_stability_synchrony,
+             aes(x = richness_mean, y = synchrony_trans, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_rich_synch_effects, aes(x = richness_mean, y = inv_logit(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_rich_synch_effects, aes(x = richness_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Synchrony\n", x = "Macroalgal Richness") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+```

### 2024-09-11 manuscript_1_code.Rmd
adding ggsave

@@
          )
-
+ggsave(filename = "output/figure_3.png", figure_3, height = 10, width = 14)
 ```
@@
          )
+
+ggsave(filename = "output/figure_4a.png", figure_4a, height = 10, width = 14)
 ```
@@
          )
+ggsave(filename = "output/figure_4c.png", figure_4c, height = 10, width = 14)
 ```

### 2024-09-12 manuscript_1_code.Rmd
Changed figure 4c to synchrony stability instead of richness synchrony

@@
 ```{r richness synchrony mod}
-rich_synch_mod <- glmmTMB(synchrony_trans ~ richness_mean*habitat + (1|site), family = beta_family(), data = diversity_stability_synchrony)
-summary(rich_synch_mod)
-hist(residuals(rich_synch_mod)) # looks good
-plot(residuals(rich_synch_mod) ~ fitted(rich_synch_mod)) # same issue as always
-r.squaredGLMM(rich_synch_mod) # 0.701, 0.775
-emtrends(rich_synch_mod, pairwise ~ habitat, var = "richness_mean") # forereefs aren't different
+synch_stab_mod <- glmmTMB(cover_stability ~ synchrony*habitat + (1|site), family = Gamma("log"), data = diversity_stability_synchrony)
+summary(synch_stab_mod)
+hist(residuals(synch_stab_mod)) # looks good
+plot(residuals(synch_stab_mod) ~ fitted(synch_stab_mod)) # same issue as always
+r.squaredGLMM(synch_stab_mod) # 0.63, 0.71
+emtrends(synch_stab_mod, pairwise ~ habitat, var = "synchrony") # fringing and backreef not different
 ```
@@
 
-rich_synch_emm <- emmip(rich_synch_mod,
-                        habitat ~ richness_mean,
-                        at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+synch_stab_emm <- emmip(synch_stab_mod,
+                        habitat ~ synchrony,
+                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)
 
-filtered_rich_synch_effects <- rich_synch_emm %>%
+filtered_synch_stab_effects <- synch_stab_emm %>%
   left_join(dss_ranges, by = c("habitat")) %>%
-  filter(richness_mean >= min_richness & richness_mean <= max_richness)
+  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
 
@@
   geom_point(data = diversity_stability_synchrony,
-             aes(x = richness_mean, y = synchrony_trans, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_rich_synch_effects, aes(x = richness_mean, y = inv_logit(yvar), colour = habitat)) +
-  geom_ribbon(data = filtered_rich_synch_effects, aes(x = richness_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
+             aes(x = synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_synch_stab_effects, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_synch_stab_effects, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
               alpha = 0.3) +
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Synchrony\n", x = "Macroalgal Richness") +
+  labs(x = "Synchrony", y = "Stability\n") +
   theme_bw() + 

### 2024-09-12 manuscript_1_code.Rmd
added synchrony cover model and plot

@@
 
-## figure 4c model - richness synchrony
+## figure 4a no habitat
 
-```{r richness synchrony mod}
+```{r richness stability no hab}
+rich_stab_mod_2 <- glmmTMB(cover_stability~ richness_mean + (1|site), family = Gamma(link = "log"), data = diversity_stability_synchrony)
+summary(rich_stab_mod_2)
+hist(residuals(rich_stab_mod_2)) # a chunkier version
+plot(residuals(rich_stab_mod_2) ~ fitted(rich_stab_mod_2)) # heteroskedastic - bring to group
+r.squaredGLMM(rich_stab_mod_2) # 0.750, 0.757
+
+```
+
+```{r 4a no hab setup}
+rich_stab_emm_2 <-as.data.frame(emmeans(rich_stab_mod_2, ~richness_mean,
+                                     at = list(richness_mean = seq(min(diversity_stability_synchrony$richness_mean),
+                                                            max(diversity_stability_synchrony$richness_mean),
+                                                            0.01))))
+
+```
+
+## figure 4a no habitat plot
+
+```{r figure 4a no habitat}
+(figure_4a_no_hab <- 
+  ggplot() +
+  geom_point(data = diversity_stability_synchrony,
+             aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_rich_stab_effects, aes(x = richness_mean, y = exp(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_rich_stab_effects, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  geom_line(data = rich_stab_emm_2, aes(x = richness_mean, y = exp(emmean)), size = 2, colour = "black") +
+  geom_ribbon(data = rich_stab_emm_2, aes(x = richness_mean, ymin = exp(asymp.LCL), ymax = exp(asymp.UCL)), fill = "black",
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Stability\n", x = "Macroalgal Richness") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+ggsave(filename = "output/figure_4a_with_no_hab.png", figure_4a_no_hab, height = 10, width = 14)
+```
+
+## figure 4c model -  synchrony staability
+
+```{r synchrony staability mod}
 synch_stab_mod <- glmmTMB(cover_stability ~ synchrony*habitat + (1|site), family = Gamma("log"), data = diversity_stability_synchrony)
@@
 ```
+
+# Synchrony and cover
+
+## synchrony and cover model
+
+```{r synchrony cover mod}
+synch_cov_mod <- glmmTMB(cover_mean ~ synchrony*habitat + (1|site), family = beta_family(), data = diversity_stability_synchrony)
+summary(synch_cov_mod)
+
+emtrends(synch_cov_mod, pairwise ~ habitat, var = "synchrony") # forereef 17 distinct from all other habitats
+```
+
+```{r synch cov mod setup}
+synch_cov_emm <- emmip(synch_cov_mod,
+                        habitat ~ synchrony,
+                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)
+
+filtered_synch_cov_effects <- synch_cov_emm %>%
+  left_join(dss_ranges, by = c("habitat")) %>%
+  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
+```
+
+## plot
+
+```{r synchrony cover plot}
+(synchrony_cover_plot <- 
+  ggplot() +
+  geom_point(data = diversity_stability_synchrony,
+             aes(x = synchrony, y = cover_mean, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_synch_cov_effects, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_synch_cov_effects, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(x = "Synchrony", y = "Cover\n") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+ggsave(filename = "output/synchrony_cover_plot.png", synchrony_cover_plot, height = 10, width = 14)
+```

### 2024-09-12 manuscript_1_code.Rmd
starting site-level figures morning of 9/12 :tropical_fish:

--- 2024-09-12 manuscript_1_code.Rmd
@@
 title: "Manuscript 1 code"
-author: "Noam Altman-Kurosaki"
+author: "Noam Altman-Kurosaki, Callie Stephenson"
 date: "2024-09-11"
@@
 
+```
+
+```{r site macroalgal data}
+site_macro <- macroalgae %>%
+  dplyr::group_by(year, site, habitat,site_habitat) %>%
+  dplyr::summarize(across(acanthophora_spicifera:valonia_ventricosa, \(x) mean(x, na.rm = TRUE)))
+
+site_meta_cols <- c("year", "site", "habitat", "site_habitat")
+site_meta_df <- site_macro[,which(names(site_macro) %in% meta_cols)]
 ```
@@
         panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
+        plot.background = element_rect(fill = "white"),
         legend.text = element_text(size = rel(1)), # increase legend text
@@
 alpha_diversity_quad_macro <- merge(meta_df, alpha_diversity_quad_macro, by = c("location", "year"))
+```
 
+```{r site level richness df}
+alpha_diversity_site_macro <-
+  data.frame(
+    year = site_macro$year,
+    site_habitat = site_macro$site_habitat,
+    shannon = diversity(site_macro[ , -which(names(site_macro) %in% meta_cols)], "shannon"),
+    richness = rowSums(site_macro[ , -which(names(site_macro) %in% meta_cols)] > 0)
+  ) %>% 
+  left_join(site_meta_df, alpha_diversity_site_macro, by = c("site_habitat", "year"))
+```
+
+# Figure 2
+
+## Figure 2a
+
+### prep
+
+```{r richness df}
+quad_richness_df <- alpha_diversity_quad_macro %>% 
+  group_by(year, habitat) %>% 
+  summarise(richness_mean = mean(richness),
+            richness_se = se(richness))
+```
+
+```{r richness factors}
+quad_richness_df$habitat <- factor(quad_richness_df$habitat, 
+                             levels = c("fringing", "backreef", "outer_10","outer_17"),
+                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+```
+
+### plot
+```{r figure 2}
+(figure_2a <- ggplot() +
+  geom_line(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
+  geom_ribbon(data = quad_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, fill = habitat), alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
+  annotate("segment", x = 2006, xend = 2010, y = 2, yend = 2, linewidth = 2, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+   scale_x_continuous(breaks = unique(quad_richness_df$year)) +
+   labs(y = "Macroalgal Richness \n",
+        x = "Year",
+        title = "a. Quadrat-level Richness") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        axis.title.y = element_text(size=18), 
+        plot.title = element_text(size=18),
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) )
+
+ggsave(filename = "output/figure_2a.png", figure_2a, height = 10, width = 14)
+```
+
+## Figure 2b
+### prep
+
+```{r richness df}
+site_richness_df <- alpha_diversity_site_macro %>% 
+  group_by(year, habitat) %>% 
+  summarise(richness_mean = mean(richness),
+            richness_se = se(richness))
+
+```
+
+```{r richness factors}
+site_richness_df$habitat <- factor(site_richness_df$habitat, 
+                             levels = c("fringing", "backreef", "outer_10","outer_17"),
+                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+
+```
+
+```{r}
+(figure_2b <- ggplot() +
+  geom_line(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
+  geom_ribbon(data = site_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, fill = habitat), alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
+  annotate("segment", x = 2006, xend = 2010, y = 10, yend = 10, linewidth = 2, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+   scale_x_continuous(breaks = unique(site_richness_df$year)) +
+   labs(y = "Macroalgal Richness \n",
+        x = "Year",
+        title = "b. Site-level Richness") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        axis.title.y = element_text(size=18), 
+        plot.title = element_text(size=18),
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) )
+
+ggsave(filename = "output/figure_2b.png", figure_2b, height = 10, width = 14)
+```
+
+# Figure 3
+
+## set up data for model
+```{r quad data setup}
 macroalgal_cover <- cover_df[,c("year", "habitat", "site", "location", "macroalgae")]
@@
 
-# Figure 2
-
-## prep
-
-```{r richness df}
-richness_df <- alpha_diversity_quad_macro %>% 
-  group_by(year, habitat) %>% 
-  summarise(richness_mean = mean(richness),
-            richness_se = se(richness))
+```{r site data setup}
+alpha_diversity_site_macro_with_cover <- alpha_diversity_site_macro %>% 
+  mutate(percent_cover = rowSums(site_macro[,-which(names(site_macro)  %in% colnames(site_meta_df))])) %>% 
+  mutate(prop_cover = percent_cover / 100) %>% 
+  select(-percent_cover) %>% 
+  mutate(cover_trans = sv_trans(prop_cover))
 ```
 
-```{r richness factors}
-richness_df$habitat <- factor(richness_df$habitat, 
-                             levels = c("fringing", "backreef", "outer_10","outer_17"),
-                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
-```
+## models
 
-## plot
-```{r figure 2}
-(figure_2 <- ggplot() +
-  geom_line(data = richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
-  geom_ribbon(data = richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, fill = habitat), alpha = 0.3) +
-  scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
-  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
-  annotate("segment", x = 2006, xend = 2010, y = 2, yend = 2, linewidth = 2, colour = "grey") +
-  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
-  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
-   scale_x_continuous(breaks = unique(richness_df$year)) +
-   labs(y = "Macroalgal Richness \n",
-        x = "Year") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.text.x = element_text(angle = 45, hjust = 1),
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black")) )
+### quad
 
-ggsave(filename = "output/figure_2.png", figure_2, height = 10, width = 14)
-```
-
-# Figure 3
-
-## model
-
-```{r figure 3 model}
+```{r figure 3a model}
 alpha_diversity_quad_macro_with_cover$cover_trans <- sv_trans(alpha_diversity_quad_macro_with_cover$macroalgal_prop_cover)
@@
 emtrends(cover_mod, pairwise ~ habitat, var = "richness") # backreef not different from forereef, forereef also not different
-
 ```
 
-## figure 3 plot setup
+### site
+```{r site mod}
+site_cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover)
+summary(site_cover_mod)
+hist(residuals(site_cover_mod)) # some positive skew, not the worst thing ever
+plot(residuals(site_cover_mod) ~ fitted(site_cover_mod)) # a little heteroscedastic but can't be less than 0
+r.squaredGLMM(site_cover_mod) # 0.2212534, 0.5990825
+emtrends(site_cover_mod, pairwise ~ habitat, var = "richness") # no difference between the habitats
+```
 
-```{r figure 3 setup}
+## figure 3a plot setup
+
+```{r figure 3a setup}
 diversity_ranges_quad <- alpha_diversity_quad_macro_with_cover %>%
@@
 
-## figure 3 plot
+## figure 3a plot
 
-```{r figure 3 plot}
-(figure_3 <- 
+```{r figure 3a plot}
+(figure_3a <- 
   ggplot() +
@@
          )
-ggsave(filename = "output/figure_3.png", figure_3, height = 10, width = 14)
+ggsave(filename = "output/figure_3a.png", figure_3a, height = 10, width = 14)
 ```
 
-```{r figure 3 v2}
-(figure_3_v2 <- figure_3 + facet_wrap(~habitat))
+## Figure 3d
+
+### setup
+```{r}
+diversity_ranges_site <- alpha_diversity_site_macro_with_cover %>%
+  group_by(habitat) %>%
+  summarise(min_richness = min(richness),
+            max_richness = max(richness),
+            min_cover= min(cover_trans),
+            max_cover = max(cover_trans))
+
+site_cover_emm <- emmip(site_cover_mod,
+                        habitat ~ richness,
+                        at = list(richness = seq(0,17,1)), plotit = F, CIs = T)
+
+filtered_site_cover_effects <- site_cover_emm %>%
+  left_join(diversity_ranges_site, by = c("habitat")) %>%
+  filter(richness >= min_richness & richness <= max_richness)
+
+alpha_diversity_site_macro_with_cover$habitat <- factor(alpha_diversity_site_macro_with_cover$habitat, 
+                             levels = c("fringing", "backreef", "outer_10","outer_17"),
+                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+
+filtered_site_cover_effects$habitat <- factor(filtered_site_cover_effects$habitat, 
+                             levels = c("fringing", "backreef", "outer_10","outer_17"),
+                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
 ```
+
+### plot
+
+```{r}
+(figure_3d <- 
+  ggplot() +
+  geom_point(data = alpha_diversity_site_macro_with_cover,
+             aes(x = richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.3) + 
+  geom_line(data = filtered_site_cover_effects, aes(x = richness, y = inv_logit(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_site_cover_effects, aes(x = richness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Cover (Proprtion)\n", x = "Macroalgal Richness", title = "3d.") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        plot.title = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+ggsave(filename = "output/figure_3d.png", figure_3d, height = 10, width = 14)
+```
+
 # Figure 4
@@
 
-## figure 4c model - richness synchrony
+## figure 4a no habitat
 
-```{r richness synchrony mod}
+```{r richness stability no hab}
+rich_stab_mod_2 <- glmmTMB(cover_stability~ richness_mean + (1|site), family = Gamma(link = "log"), data = diversity_stability_synchrony)
+summary(rich_stab_mod_2)
+hist(residuals(rich_stab_mod_2)) # a chunkier version
+plot(residuals(rich_stab_mod_2) ~ fitted(rich_stab_mod_2)) # heteroskedastic - bring to group
+r.squaredGLMM(rich_stab_mod_2) # 0.750, 0.757
+
+```
+
+```{r 4a no hab setup}
+rich_stab_emm_2 <-as.data.frame(emmeans(rich_stab_mod_2, ~richness_mean,
+                                     at = list(richness_mean = seq(min(diversity_stability_synchrony$richness_mean),
+                                                            max(diversity_stability_synchrony$richness_mean),
+                                                            0.01))))
+
+```
+
+## figure 4a no habitat plot
+
+```{r figure 4a no habitat}
+(figure_4a_no_hab <- 
+  ggplot() +
+  geom_point(data = diversity_stability_synchrony,
+             aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_rich_stab_effects, aes(x = richness_mean, y = exp(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_rich_stab_effects, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  geom_line(data = rich_stab_emm_2, aes(x = richness_mean, y = exp(emmean)), size = 2, colour = "black") +
+  geom_ribbon(data = rich_stab_emm_2, aes(x = richness_mean, ymin = exp(asymp.LCL), ymax = exp(asymp.UCL)), fill = "black",
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Stability\n", x = "Macroalgal Richness") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+ggsave(filename = "output/figure_4a_with_no_hab.png", figure_4a_no_hab, height = 10, width = 14)
+```
+
+## figure 4c model -  synchrony staability
+
+```{r synchrony staability mod}
 synch_stab_mod <- glmmTMB(cover_stability ~ synchrony*habitat + (1|site), family = Gamma("log"), data = diversity_stability_synchrony)
@@
 ```
+
+# Figure 3c
+
+## synchrony and cover model
+
+```{r synchrony cover mod}
+synch_cov_mod <- glmmTMB(cover_mean ~ synchrony*habitat + (1|site), family = beta_family(), data = diversity_stability_synchrony)
+summary(synch_cov_mod)
+
+emtrends(synch_cov_mod, pairwise ~ habitat, var = "synchrony") # forereef 17 distinct from all other habitats
+```
+
+```{r synch cov mod setup}
+synch_cov_emm <- emmip(synch_cov_mod,
+                        habitat ~ synchrony,
+                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)
+
+filtered_synch_cov_effects <- synch_cov_emm %>%
+  left_join(dss_ranges, by = c("habitat")) %>%
+  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
+```
+
+## plot
+
+```{r synchrony cover plot}
+(figure_3c <- 
+  ggplot() +
+  geom_point(data = diversity_stability_synchrony,
+             aes(x = synchrony, y = cover_mean, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_synch_cov_effects, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_synch_cov_effects, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(x = "Synchrony", y = "Cover\n") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+ggsave(filename = "output/figure_3c.png", figure_3c, height = 10, width = 14)
+```
+
+## Synchrony setup site level
+```{r calculate syncrony site_level}
+# make er long
+macro_long_data_site <- pivot_longer(
+  data = site_macro,
+  cols = grep("acanthophora_spicifera", colnames(site_macro)):grep("valonia_ventricosa", colnames(site_macro)),
+  names_to = "taxa",
+  values_to = "percent_cover") %>% 
+  mutate(prop_cover = percent_cover/100) %>% 
+  select(-percent_cover)
+
+# write_csv(macro_long_data, here("Data", "macro_long_data.csv"))
+
+# CALCULATE SYNCHRONY CODE
+lm_synchrony_site <- codyn::synchrony(df = macro_long_data_site,
+                                   time.var = "year",
+                                   species.var = "taxa",
+                                   abundance.var = "prop_cover",
+                                   metric = "Loreau",
+                                   replicate.var = "site_habitat")
+
+diversity_stability_site <- alpha_diversity_site_macro_with_cover %>%
+  dplyr::group_by(site_habitat, habitat, site) %>%
+  dplyr::summarise(
+            richness_mean = mean(richness), 
+            shannon_mean = mean(shannon), 
+            cover_mean = mean(prop_cover),
+            cover_cv = CV(prop_cover),
+            cover_stability = 1/CV(prop_cover))
+
+diversity_stability_synchrony_site <- left_join(diversity_stability_site, lm_synchrony_site, by = join_by(site_habitat))
+diversity_stability_synchrony_site$synchrony_trans <- sv_trans(diversity_stability_synchrony_site$synchrony)
+# diversity_stability_synchrony_site <- na.omit(diversity_stability_synchrony_site) # remove quads that never had algae --> NaN calcs
+
+dss_ranges_site<- diversity_stability_synchrony_site %>%
+  group_by(habitat) %>%
+  summarise(min_richness = min(richness_mean),
+            max_richness = max(richness_mean),
+            min_stability= min(cover_stability),
+            max_stability = max(cover_stability),
+            min_synchrony = min(synchrony_trans),
+            max_synchrony = max(synchrony_trans))
+```
+
+## figure 4d model - richness stability
+
+```{r richness synchrony mod}
+rich_stab_mod_site <- glmmTMB(cover_stability~ richness_mean*habitat, family = Gamma(link = "log"), data = diversity_stability_synchrony_site)
+summary(rich_stab_mod_site)
+hist(residuals(rich_stab_mod_site)) # blocky but fine
+plot(residuals(rich_stab_mod_site) ~ fitted(rich_stab_mod_site)) # heteroskedastic - bring to group
+r.squaredGLMM(rich_stab_mod_site) # 0.671
+emtrends(rich_stab_mod_site, pairwise ~ habitat, var = "richness_mean") # no differences across habitats
+```
+
+```{r figure 4d setup}
+
+rich_stab_emm_site <- emmip(rich_stab_mod_site,
+                        habitat ~ richness_mean,
+                        at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_rich_stab_effects_site <- rich_stab_emm_site %>%
+  left_join(dss_ranges_site, by = c("habitat")) %>%
+  filter(richness_mean >= min_richness & richness_mean <= max_richness)
+
+```
+
+## plot figure 4d
+
+```{r figure 4d}
+(figure_4d <- 
+  ggplot() +
+  geom_point(data = diversity_stability_synchrony_site,
+             aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_rich_stab_effects_site, aes(x = richness_mean, y = exp(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_rich_stab_effects_site, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Stability\n", x = "Macroalgal Richness", title = "4d.") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        plot.title = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+ggsave(filename = "output/figure_4d.png", figure_4d, height = 10, width = 14)
+```

### 2024-09-12 manuscript_1_code.Rmd
Made figures 3f and 4f

@@
 summary(synch_cov_mod)
-
+hist(residuals(synch_cov_mod)) # slight skew
+plot(residuals(synch_cov_mod) ~ fitted(synch_cov_mod)) # same issues as always
 emtrends(synch_cov_mod, pairwise ~ habitat, var = "synchrony") # forereef 17 distinct from all other habitats
@@
 
-```{r synch cov mod setup}
+```{r synch cov plot setup}
 synch_cov_emm <- emmip(synch_cov_mod,
@@
 ```
+
+## figure 4f
+
+```{r synchrony staability mod site}
+synch_stab_mod_site <- glmmTMB(cover_stability ~ synchrony*habitat, family = Gamma("log"), data = diversity_stability_synchrony_site)
+summary(synch_stab_mod_site)
+hist(residuals(synch_stab_mod_site)) # almost uniform
+plot(residuals(synch_stab_mod_site) ~ fitted(synch_stab_mod_site)) # same issue as always
+r.squaredGLMM(synch_stab_mod_site) # 0.678
+emtrends(synch_stab_mod_site, pairwise ~ habitat, var = "synchrony") # fringing and backreef are different from forereef 17
+```
+
+```{r figure 4f setup}
+
+synch_stab_emm_site <- emmip(synch_stab_mod_site,
+                        habitat ~ synchrony,
+                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)
+
+filtered_synch_stab_effects_site <- synch_stab_emm_site %>%
+  left_join(dss_ranges_site, by = c("habitat")) %>%
+  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
+
+```
+
+## plot figure 4f
+
+```{r figure 4c}
+(figure_4f <- 
+  ggplot() +
+  geom_point(data = diversity_stability_synchrony_site,
+             aes(x = synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_synch_stab_effects_site, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_synch_stab_effects_site, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(x = "Synchrony", y = "Stability\n", main = "4f.") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18),
+        plot.title = element_text(size=18),
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+ggsave(filename = "output/figure_4f.png", figure_4f, height = 10, width = 14)
+```
+
+# figure 3f
+
+## model
+
+```{r synchrony cover mod}
+synch_cov_mod_site <- glmmTMB(cover_mean ~ synchrony*habitat, family = beta_family(), data = diversity_stability_synchrony_site)
+summary(synch_cov_mod_site)
+hist(residuals(synch_cov_mod_site)) # slight skew
+plot(residuals(synch_cov_mod_site) ~ fitted(synch_cov_mod_site)) # looks fine
+emtrends(synch_cov_mod_site, pairwise ~ habitat, var = "synchrony") # forereef 17 different from backreef
+```
+
+```{r synch cov site plot setup}
+synch_cov_emm_site<- emmip(synch_cov_mod_site,
+                        habitat ~ synchrony,
+                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)
+
+filtered_synch_cov_effects_site <- synch_cov_emm_site %>%
+  left_join(dss_ranges_site, by = c("habitat")) %>%
+  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
+```
+
+## plot figure 3f
+
+```{r synchrony cover plot site}
+(figure_3f <- 
+  ggplot() +
+  geom_point(data = diversity_stability_synchrony_site,
+             aes(x = synchrony, y = cover_mean, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_synch_cov_effects_site, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_synch_cov_effects_site, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(x = "Synchrony", y = "Cover\n", title = "3f.") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        plot.title = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_text(size = rel(1)),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+ggsave(filename = "output/figure_3f.png", figure_3f, height = 10, width = 14)
+```

### 2024-09-12 manuscript_1_code.Rmd
fixed figure 4f

@@
 
-```{r figure 4c}
+```{r figure 4f}
 (figure_4f <- 

### 2024-09-12 manuscript_1_code.Rmd
Fixed figures/formatting, added Anova() tables for all models

@@
 
+### if we want to make a missing or faulty data table
+# length(unique(data$location))
+# expected_data <- expand.grid(year = seq(2007,2023,1), location = unique(data$location))
+# missing_quadrats <- expected_data %>%
+#   anti_join(data, by = c("year", "location"))
+###
+
+# remove remaining backreef and fringing 2020 data because of covid extrapolation
+filtered_data <- data %>%
+  filter(!(year == 2020 & habitat %in% c("fringing", "backreef")))
+
 not_algae <- c("ascidian", "bare_space", "coral", "coral_rubble", "corallimorpharia", "heteractis_sp", "millepora_platyphylla", "no_data", "sand", "sarcophyton_sp", "shell_debris", "soft_coral", "sponge", "tridacna_sp", "zooxanthellae", "row_sums")
 
-algae <- data[ , -which(names(data) %in% not_algae)]
+algae <- filtered_data[ , -which(names(filtered_data) %in% not_algae)]
 
@@
 cover_df <- data.frame(
-  year = data$year,
-  habitat = data$habitat,
-  site = data$site,
-  location = data$location,
-  coral = data$coral,
-  cca = data$crustose_corallines + data$lithophyllum_kotschyanum + data$neogoniolithon_brassica_florida + data$sporolithon_sp,
+  year = filtered_data$year,
+  habitat = filtered_data$habitat,
+  site = filtered_data$site,
+  location = filtered_data$location,
+  coral = filtered_data$coral,
+  cca = filtered_data$crustose_corallines +
+    filtered_data$lithophyllum_kotschyanum +
+    filtered_data$neogoniolithon_brassica_florida +
+    filtered_data$sporolithon_sp,
   macroalgae = rowSums(macro_species_cols),
-  turf = data$algal_turf  + data$damselfish_turf, 
-  sand = data$sand,
-  coral_rubble = data$coral_rubble,
-  cyanophyta = data$cyanophyta,
-  other = data$ascidian + data$corallimorpharia + data$heteractis_sp + data$millepora_platyphylla + data$sarcophyton_sp + data$shell_debris +
-    data$soft_coral + data$sponge + data$tridacna_sp + data$zooxanthellae + data$no_data + data$bare_space
+  turf = filtered_data$algal_turf  + filtered_data$damselfish_turf, 
+  sand = filtered_data$sand,
+  coral_rubble = filtered_data$coral_rubble,
+  cyanophyta = filtered_data$cyanophyta,
+  other = filtered_data$ascidian + 
+    filtered_data$corallimorpharia +
+    filtered_data$heteractis_sp + 
+    filtered_data$millepora_platyphylla +
+    filtered_data$sarcophyton_sp + filtered_data$shell_debris +
+    filtered_data$soft_coral + filtered_data$sponge + filtered_data$tridacna_sp +
+    filtered_data$zooxanthellae + filtered_data$no_data + filtered_data$bare_space
 )
@@
 (figure_1 <- ggplot() +
-  geom_line(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
-  geom_ribbon(data = figure_1_data, aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, fill = functional_group), alpha = 0.3) +
+  geom_point(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), size = 3) +
+  geom_errorbar(data = figure_1_data, aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, colour = functional_group),
+                width = 0, size = 1.3) +
+   geom_line(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), size = 1) +
   facet_wrap(~ habitat) +
-  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
   annotate("segment", x = 2006, xend = 2010, y = 75, yend = 75, linewidth = 2, colour = "grey") +
@@
 
-#ggsave(filename = "output/figure_1.png", figure_1, height = 10, width = 14)
+# ggsave(filename = "output/figure_1.png", figure_1, height = 10, width = 14)
 ```
@@
 (figure_2a <- ggplot() +
+  geom_point(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 3) +
   geom_line(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
-  geom_ribbon(data = quad_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, fill = habitat), alpha = 0.3) +
+  geom_errorbar(data = quad_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se,
+                                             colour = habitat), width = 0, size = 1.3) +
   scale_colour_manual(values = habitat_colours) +
@@
 
-ggsave(filename = "output/figure_2a.png", figure_2a, height = 10, width = 14)
+# ggsave(filename = "output/figure_2a.png", figure_2a, height = 10, width = 14)
 ```
@@
 (figure_2b <- ggplot() +
+  geom_point(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 3) +
   geom_line(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
-  geom_ribbon(data = site_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, fill = habitat), alpha = 0.3) +
+  geom_errorbar(data = site_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, colour = habitat), size = 1.3, width = 0) +
   scale_colour_manual(values = habitat_colours) +
@@
   #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
-  annotate("segment", x = 2006, xend = 2010, y = 10, yend = 10, linewidth = 2, colour = "grey") +
+  annotate("segment", x = 2006, xend = 2010, y = 11, yend = 11, linewidth = 2, colour = "grey") +
   geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
@@
 
-ggsave(filename = "output/figure_2b.png", figure_2b, height = 10, width = 14)
+# ggsave(filename = "output/figure_2b.png", figure_2b, height = 10, width = 14)
 ```
@@
 summary(cover_mod)
+Anova(cover_mod)
+#                     Chisq Df Pr(>Chisq)    
+# richness         30916.73  1  < 2.2e-16 ***
+# habitat            141.78  3  < 2.2e-16 ***
+# richness:habitat   294.54  3  < 2.2e-16 ***
 hist(residuals(cover_mod)) # looks fine
@@
 summary(site_cover_mod)
+Anova(site_cover_mod)
+#                    Chisq Df Pr(>Chisq)    
+# richness         58.9195  1  1.643e-14 ***
+# habitat           6.6247  3    0.08487 .  
+# richness:habitat  3.2967  3    0.34811    
 hist(residuals(site_cover_mod)) # some positive skew, not the worst thing ever
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Cover (Proprtion)\n", x = "Macroalgal Richness") +
+  labs(y = "Cover\n", x = "Macroalgal Richness", title = "3a.") +
   theme_bw() + 
@@
          )
-ggsave(filename = "output/figure_3a.png", figure_3a, height = 10, width = 14)
+# ggsave(filename = "output/figure_3a.png", figure_3a, height = 10, width = 14)
 ```
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Cover (Proprtion)\n", x = "Macroalgal Richness", title = "3d.") +
+  labs(y = "Cover\n", x = "Macroalgal Richness", title = "3d.") +
   theme_bw() + 
@@
          )
-ggsave(filename = "output/figure_3d.png", figure_3d, height = 10, width = 14)
+# ggsave(filename = "output/figure_3d.png", figure_3d, height = 10, width = 14)
 ```
@@
 summary(rich_stab_mod)
+Anova(rich_stab_mod)
+#                          Chisq Df Pr(>Chisq)    
+# richness_mean         2721.438  1  < 2.2e-16 ***
+# habitat                325.876  3  < 2.2e-16 ***
+# richness_mean:habitat   33.926  3  2.054e-07 ***
 hist(residuals(rich_stab_mod)) # looks good
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Stability\n", x = "Macroalgal Richness") +
+  labs(y = "Stability\n", x = "Macroalgal Richness", title = "4a.") +
   theme_bw() + 
@@
         axis.title.y = element_text(size=18),
+        plot.title = element_text(size=18),
         panel.background = element_blank(), 
@@
 
-ggsave(filename = "output/figure_4a.png", figure_4a, height = 10, width = 14)
+# ggsave(filename = "output/figure_4a.png", figure_4a, height = 10, width = 14)
 ```
@@
 ```{r richness stability no hab}
-rich_stab_mod_2 <- glmmTMB(cover_stability~ richness_mean + (1|site), family = Gamma(link = "log"), data = diversity_stability_synchrony)
-summary(rich_stab_mod_2)
-hist(residuals(rich_stab_mod_2)) # a chunkier version
-plot(residuals(rich_stab_mod_2) ~ fitted(rich_stab_mod_2)) # heteroskedastic - bring to group
-r.squaredGLMM(rich_stab_mod_2) # 0.750, 0.757
+# rich_stab_mod_2 <- glmmTMB(cover_stability~ richness_mean + (1|site), family = Gamma(link = "log"), data = # diversity_stability_synchrony)
+# summary(rich_stab_mod_2)
+# hist(residuals(rich_stab_mod_2)) # a chunkier version
+# plot(residuals(rich_stab_mod_2) ~ fitted(rich_stab_mod_2)) # heteroskedastic - bring to group
+# r.squaredGLMM(rich_stab_mod_2) # 0.750, 0.757
 
@@
 ```{r 4a no hab setup}
-rich_stab_emm_2 <-as.data.frame(emmeans(rich_stab_mod_2, ~richness_mean,
-                                     at = list(richness_mean = seq(min(diversity_stability_synchrony$richness_mean),
-                                                            max(diversity_stability_synchrony$richness_mean),
-                                                            0.01))))
-
+# rich_stab_emm_2 <-as.data.frame(emmeans(rich_stab_mod_2, ~richness_mean,
+#                                      at = list(richness_mean = seq(min(diversity_stability_synchrony$richness_mean),
+#                                                             max(diversity_stability_synchrony$richness_mean),
+#                                                             0.01)))) 
+# 
 ```
@@
 ```{r figure 4a no habitat}
-(figure_4a_no_hab <- 
-  ggplot() +
-  geom_point(data = diversity_stability_synchrony,
-             aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_rich_stab_effects, aes(x = richness_mean, y = exp(yvar), colour = habitat)) +
-  geom_ribbon(data = filtered_rich_stab_effects, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
-              alpha = 0.3) +
-  geom_line(data = rich_stab_emm_2, aes(x = richness_mean, y = exp(emmean)), size = 2, colour = "black") +
-  geom_ribbon(data = rich_stab_emm_2, aes(x = richness_mean, ymin = exp(asymp.LCL), ymax = exp(asymp.UCL)), fill = "black",
-              alpha = 0.3) +
-  scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
-  labs(y = "Stability\n", x = "Macroalgal Richness") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_text(size = rel(1)),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-ggsave(filename = "output/figure_4a_with_no_hab.png", figure_4a_no_hab, height = 10, width = 14)
+# (figure_4a_no_hab <- 
+#   ggplot() +
+#   geom_point(data = diversity_stability_synchrony,
+#              aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+#   geom_line(data = filtered_rich_stab_effects, aes(x = richness_mean, y = exp(yvar), colour = habitat)) +
+#   geom_ribbon(data = filtered_rich_stab_effects, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+#               alpha = 0.3) +
+#   geom_line(data = rich_stab_emm_2, aes(x = richness_mean, y = exp(emmean)), size = 2, colour = "black") +
+#   geom_ribbon(data = rich_stab_emm_2, aes(x = richness_mean, ymin = exp(asymp.LCL), ymax = exp(asymp.UCL)), fill = "black",
+#               alpha = 0.3) +
+#   scale_colour_manual(values = habitat_colours) +
+#   scale_fill_manual(values = habitat_colours) +
+#   labs(y = "Stability\n", x = "Macroalgal Richness") +
+#   theme_bw() + 
+#   theme(axis.title.x = element_text(size=18), 
+#         axis.title.y = element_text(size=18), 
+#         panel.background = element_blank(), 
+#         panel.grid.major = element_blank(),  #remove major-grid labels
+#         panel.grid.minor = element_blank(),  #remove minor-grid labels
+#         plot.background = element_rect(fill = "white"),
+#         legend.text = element_text(size = rel(1)), # increase legend text
+#         legend.title = element_blank(),
+#         legend.position = 'bottom', # move legend to bottom
+#         legend.box = 'horizontal', # make legend horizontal
+#         legend.box.background = element_rect(colour = "black"))  # add box around legend
+#          )
+# 
+# ggsave(filename = "output/figure_4a_with_no_hab.png", figure_4a_no_hab, height = 10, width = 14)
 ```
@@
 summary(synch_stab_mod)
+Anova(synch_stab_mod)
+#                     Chisq Df Pr(>Chisq)    
+# synchrony         1308.33  1  < 2.2e-16 ***
+# habitat            103.13  3  < 2.2e-16 ***
+# synchrony:habitat  327.97  3  < 2.2e-16 ***
 hist(residuals(synch_stab_mod)) # looks good
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(x = "Synchrony", y = "Stability\n") +
+  labs(x = "Synchrony", y = "Stability\n", title = "4c.") +
   theme_bw() + 
@@
          )
-ggsave(filename = "output/figure_4c.png", figure_4c, height = 10, width = 14)
+# ggsave(filename = "output/figure_4c.png", figure_4c, height = 10, width = 14)
 ```
@@
 summary(synch_cov_mod)
+Anova(synch_cov_mod)
+#                     Chisq Df Pr(>Chisq)    
+# synchrony         209.545  1  < 2.2e-16 ***
+# habitat            52.512  3  2.330e-11 ***
+# synchrony:habitat  54.632  3  8.227e-12 ***
 hist(residuals(synch_cov_mod)) # slight skew
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(x = "Synchrony", y = "Cover\n") +
+  labs(x = "Synchrony", y = "Cover\n", title = "3c.") +
   theme_bw() + 
@@
 
-ggsave(filename = "output/figure_3c.png", figure_3c, height = 10, width = 14)
+# ggsave(filename = "output/figure_3c.png", figure_3c, height = 10, width = 14)
 ```
@@
 # diversity_stability_synchrony_site <- na.omit(diversity_stability_synchrony_site) # remove quads that never had algae --> NaN calcs
+# write_csv(diversity_stability_synchrony_site, here("data", "diversity_stability_synchrony_site.csv"))
 
@@
 summary(rich_stab_mod_site)
+Anova(rich_stab_mod_site)
+
+#                        Chisq Df Pr(>Chisq)    
+# richness_mean         17.645  1  2.662e-05 ***
+# habitat               32.229  3  4.682e-07 ***
+# richness_mean:habitat  8.075  3    0.04449 *  
 hist(residuals(rich_stab_mod_site)) # blocky but fine
@@
 
-ggsave(filename = "output/figure_4d.png", figure_4d, height = 10, width = 14)
+# ggsave(filename = "output/figure_4d.png", figure_4d, height = 10, width = 14)
 ```
@@
 summary(synch_stab_mod_site)
+Anova(synch_stab_mod_site)
+#                     Chisq Df Pr(>Chisq)    
+# synchrony         22.0686  1  2.631e-06 ***
+# habitat            9.3333  3    0.02517 *  
+# synchrony:habitat 10.5561  3    0.01439 *  
 hist(residuals(synch_stab_mod_site)) # almost uniform
@@
 r.squaredGLMM(synch_stab_mod_site) # 0.678
-emtrends(synch_stab_mod_site, pairwise ~ habitat, var = "synchrony") # fringing and backreef are different from forereef 17
+emtrends(synch_stab_mod_site, pairwise ~ habitat, var = "synchrony") # backreef is different from forereef 17, fringing is a p = 0.05
 ```
@@
          )
-ggsave(filename = "output/figure_4f.png", figure_4f, height = 10, width = 14)
+# ggsave(filename = "output/figure_4f.png", figure_4f, height = 10, width = 14)
 ```
@@
 summary(synch_cov_mod_site)
+Anova(synch_cov_mod_site)
+#                    Chisq Df Pr(>Chisq)  
+# synchrony         2.4184  1    0.11992  
+# habitat           7.8002  3    0.05033 .
+# synchrony:habitat 8.8866  3    0.03084 *
 hist(residuals(synch_cov_mod_site)) # slight skew
@@
 
-ggsave(filename = "output/figure_3f.png", figure_3f, height = 10, width = 14)
-```
+# ggsave(filename = "output/figure_3f.png", figure_3f, height = 10, width = 14)
+```

### 2024-09-13 manuscript_1_code.Rmd
Added in 2 a-d for site and quad richness and evenness by year :cow:

@@
 ## plot
-
 ```{r figure 1, fig.height = 10, fig.width=14, warning=FALSE}
@@
   scale_colour_manual(values = functional_group_colours) +
-  scale_fill_manual(values = functional_group_colours) +
+  # scale_fill_manual(values = functional_group_colours) +
   geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
@@
 ## Create Data Frames
-```{r create data}
+```{r create data for figure 2 quad}
 alpha_diversity_quad_macro <-
@@
     location = macroalgae$location,
-    shannon = diversity(macroalgae[ , -which(names(macroalgae) %in% meta_cols)], "shannon"),
-    richness = rowSums(macroalgae[ , -which(names(macroalgae) %in% meta_cols)] > 0)
+    shannon = diversity(macroalgae[, -which(names(macroalgae) %in% meta_cols)], "shannon"),
+    richness = rowSums(macroalgae[, -which(names(macroalgae) %in% meta_cols)] > 0)
+  )
+
+#JG Add Evenness
+alpha_diversity_quad_macro$ev <- alpha_diversity_quad_macro$shannon /
+  (log(alpha_diversity_quad_macro$richness))
+
+#makes 1 species 0 and 0 species NA for evenness metric
+alpha_diversity_quad_macro$evenness <- ifelse(
+  alpha_diversity_quad_macro$richness == 0,
+  NA,
+  # If richness == 0, ev is NA
+  ifelse(
+    alpha_diversity_quad_macro$richness == 1,
+    0,
+    # If richness == 1, ev is 0
+    alpha_diversity_quad_macro$ev
+  )
 )
@@
 
-# Figure 2
-
-## Figure 2a
+```{r site level evenness}
+#JG Add Evenness
+alpha_diversity_site_macro$ev <- alpha_diversity_site_macro$shannon /
+  (log(alpha_diversity_site_macro$richness))
+
+#makes 1 species 0 and 0 species NA for evenness metric
+alpha_diversity_site_macro$evenness <- ifelse(
+  alpha_diversity_site_macro$richness == 0,
+  NA,
+  # If richness == 0, ev is NA
+  ifelse(
+    alpha_diversity_site_macro$richness == 1,
+    0,
+    # If richness == 1, ev is 0
+    alpha_diversity_site_macro$ev
+  )
+)
+
+alpha_diversity_site_macro <- left_join(site_meta_df, alpha_diversity_site_macro,by = join_by(year, site, habitat, site_habitat))
+```
+
+```{r}
+macroalgal_cover <- cover_df[,c("year", "habitat", "site", "location", "macroalgae")]
+
+alpha_diversity_quad_macro_with_cover <- left_join(alpha_diversity_quad_macro, macroalgal_cover,by = join_by(location, year, site, habitat)) %>% 
+  mutate(macroalgal_prop_cover = macroalgae / 100) %>% 
+  select(-macroalgae)
+```
+
+## Figure 2a Richness @ Quad by Year
 
 ### prep
-
 ```{r richness df}
@@
   scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
+  # scale_fill_manual(values = habitat_colours) +
   #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
@@
 
-## Figure 2b
+## Figure 2b Richness @ Site by Year
+
 ### prep
@@
 
+## 2c Evenness @ Quad by Year
+
+```{r calculate evennes}
+#dear callie, why is this here?
+site_richness_df <- alpha_diversity_site_macro %>% 
+  group_by(year, habitat) %>% 
+  summarise(richness_mean = mean(richness),
+            richness_se = se(richness))
+
+evenness_df <- alpha_diversity_quad_macro_with_cover %>% 
+  drop_na(evenness) %>%
+  group_by(year, habitat) %>% 
+  summarise(eveness_mean = mean(evenness),
+            eveness_se = se(evenness))
+
+evenness_df$habitat <- factor(evenness_df$habitat, 
+                             levels = c("fringing", "backreef", "outer_10","outer_17"),
+                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+```
+
+```{r plot evenness}
+(figure_2c <- ggplot() +
+  geom_line(data = evenness_df, aes(x = year, y = eveness_mean, colour = habitat), linewidth = 1) +
+  geom_ribbon(data = evenness_df, aes(x = year, ymin = eveness_mean - eveness_se, ymax = eveness_mean + eveness_se, fill = habitat), alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
+  annotate("segment", x = 2006, xend = 2010, y = 0.7, yend = 0.7, linewidth = 2, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+   scale_x_continuous(breaks = unique(evenness_df$year)) +
+   labs(y = "Macroalgal Evenness \n",
+        x = "Year") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) )
+
+#ggsave(filename = "output/figure_2c.png", figure_2c, height = 10, width = 14)
+
+#hist(eveness_df$eveness_mean)
+```
+
+## 2d Evenness @ Site by Year
+
+```{r COME BACK HERE set up evenness}
+site_evenness_df <- alpha_diversity_site_macro %>% 
+  drop_na(evenness) %>%
+  group_by(year, habitat) %>% 
+  summarise(eveness_mean = mean(evenness),
+            eveness_se = se(evenness))
+
+site_evenness_df$habitat <- factor(site_evenness_df$habitat, 
+                             levels = c("fringing", "backreef", "outer_10","outer_17"),
+                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+```
+
+```{r plot evenness}
+(figure_2d <- ggplot() +
+  geom_line(data = site_evenness_df, aes(x = year, y = eveness_mean, colour = habitat), linewidth = 1) +
+  geom_ribbon(data = site_evenness_df, aes(x = year, ymin = eveness_mean - eveness_se, ymax = eveness_mean + eveness_se, fill = habitat), alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
+  annotate("segment", x = 2006, xend = 2010, y = 1, yend = 1, linewidth = 2, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+   scale_x_continuous(breaks = unique(evenness_df$year)) +
+   labs(y = "Macroalgal Evenness \n",
+        x = "Year") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_blank(),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) )
+
+#ggsave(filename = "output/figure_2d.png", figure_2d, height = 10, width = 14)
+```
+
 # Figure 3
@@
 ## set up data for model
-```{r quad data setup}
-macroalgal_cover <- cover_df[,c("year", "habitat", "site", "location", "macroalgae")]
-
-alpha_diversity_quad_macro_with_cover <- left_join(alpha_diversity_quad_macro, macroalgal_cover,by = join_by(location, year, site, habitat)) %>% 
-  mutate(macroalgal_prop_cover = macroalgae / 100) %>% 
-  select(-macroalgae)
-```
 

### 2024-09-13 manuscript_1_code.Rmd
Wrote models for quad and site evenness, made figure (proposed 3b) with quad-level evenness versus cover :cow2:

@@
 
-```{r COME BACK HERE set up evenness}
+```{r set up evenness}
 site_evenness_df <- alpha_diversity_site_macro %>% 
@@
 
-## set up data for model
-
-```{r site data setup}
+## set up site-level data for model
+```{r site cover by diversity data setup}
+# alpha_diversity_site_macro_with_cover <- alpha_diversity_site_macro %>% 
+#   dplyr::mutate(percent_cover = rowSums(site_macro[,-which(names(site_macro)  %in% colnames(site_meta_df))])) %>% 
+#   mutate(prop_cover = percent_cover / 100) %>% 
+#   select(-percent_cover) %>% 
+#   mutate(cover_trans = sv_trans(prop_cover))
+
+#there is no reasonable reason why the above chunk doesn't work, but it doesn't, so we did it this way
+alpha_diversity_site_macro$percent_cover <- rowSums(site_macro[,-which(names(site_macro)  %in% colnames(site_meta_df))])
+
 alpha_diversity_site_macro_with_cover <- alpha_diversity_site_macro %>% 
-  mutate(percent_cover = rowSums(site_macro[,-which(names(site_macro)  %in% colnames(site_meta_df))])) %>% 
-  mutate(prop_cover = percent_cover / 100) %>% 
-  select(-percent_cover) %>% 
-  mutate(cover_trans = sv_trans(prop_cover))
+  dplyr::mutate(prop_cover = percent_cover / 100) %>%
+  select(-percent_cover)
+
+alpha_diversity_site_macro_with_cover$cover_trans <- sv_trans(alpha_diversity_site_macro_with_cover$prop_cover)
 ```
@@
 ### quad
-
-```{r figure 3a model}
+```{r figure 3a richness quad model}
 alpha_diversity_quad_macro_with_cover$cover_trans <- sv_trans(alpha_diversity_quad_macro_with_cover$macroalgal_prop_cover)
@@
 
-### site
-```{r site mod}
+```{r 3b evenness quad level model}
+alpha_diversity_quad_macro_with_cover_evenness_mod <- alpha_diversity_quad_macro_with_cover %>% 
+  drop_na(evenness)
+
+evenness_cover_mod_quad <- glmmTMB(cover_trans ~ evenness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro_with_cover_evenness_mod)
+summary(evenness_cover_mod_quad)
+Anova(evenness_cover_mod_quad)
+#                     Chisq Df Pr(>Chisq)    
+# richness         30916.73  1  < 2.2e-16 ***
+# habitat            141.78  3  < 2.2e-16 ***
+# richness:habitat   294.54  3  < 2.2e-16 ***
+hist(residuals(evenness_cover_mod_quad)) # slight pos skeq
+plot(residuals(evenness_cover_mod_quad) ~ fitted(evenness_cover_mod_quad)) # negative trend but it can't be less than 0 so not too concerned. some wave pattern
+r.squaredGLMM(evenness_cover_mod_quad) # 0.1749907, 0.4042826
+emtrends(evenness_cover_mod_quad, pairwise ~ habitat, var = "evenness") #backreef is different from everything except outer 17, but just barely. everyone else is same same xoxo
+
+```
+
+### site richness
+```{r site richness mod}
 site_cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover)
@@
 
+### site evenness 
+```{r}
+alpha_diversity_site_macro_with_cover_evenness_mod <- alpha_diversity_site_macro_with_cover %>% 
+  drop_na(evenness)
+
+evenness_site_cover_mod <- glmmTMB(cover_trans ~ evenness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover_evenness_mod)
+summary(evenness_site_cover_mod)
+Anova(evenness_site_cover_mod)
+#                    Chisq Df Pr(>Chisq)    
+# richness         58.9195  1  1.643e-14 ***
+# habitat           6.6247  3    0.08487 .  
+# richness:habitat  3.2967  3    0.34811    
+hist(residuals(evenness_site_cover_mod)) # some positive skew, not the worst thing ever
+plot(residuals(evenness_site_cover_mod) ~ fitted(evenness_site_cover_mod)) # a little heteroscedastic but can't be less than 0
+r.squaredGLMM(evenness_site_cover_mod) #0.1206569, 0.5852773
+emtrends(evenness_site_cover_mod, pairwise ~ habitat, var = "evenness") # only the outers are different from each other
+```
+
 ## figure 3a plot setup
-
 ```{r figure 3a setup}
@@
 
-## figure 3a plot
+## figure 3a quad level richness versus cover
 
@@
 
-## Figure 3d
+## Figure 3B quad level evenness versus cover
+```{r COME BACK HERE figure 3b setup}
+diversity_ranges_quad_evenness <- alpha_diversity_quad_macro_with_cover_evenness_mod %>%
+  group_by(habitat) %>%
+  summarise(min_evenness = min(evenness),
+            max_evenness = max(evenness),
+            min_cover= min(cover_trans),
+            max_cover = max(cover_trans))
+
+cover_emm_evenness <- emmip(evenness_cover_mod_quad,
+                        habitat ~ evenness,
+                        at = list(evenness = seq(0,10,1)), plotit = F, CIs = T)
+
+filtered_cover_effects_evenness <- cover_emm %>%
+  left_join(diversity_ranges_quad_evenness, by = c("habitat")) %>%
+  filter(evenness >= min_evenness & evenness <= max_evenness)
+
+alpha_diversity_quad_macro_with_cover_evenness_mod$habitat <- factor(alpha_diversity_quad_macro_with_cover_evenness_mod$habitat, 
+                             levels = c("fringing", "backreef", "outer_10","outer_17"),
+                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+
+filtered_cover_effects_evenness$habitat <- factor(filtered_cover_effects_evenness$habitat, 
+                             levels = c("fringing", "backreef", "outer_10","outer_17"),
+                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+```
+
+```{r}
+(figure_3b <- 
+  ggplot() +
+  geom_point(data = alpha_diversity_quad_macro_with_cover_evenness_mod,
+             aes(x = evenness, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
+  geom_line(data = filtered_cover_effects_evenness, aes(x = evenness, y = inv_logit(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_cover_effects_evenness, aes(x = evenness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Cover\n", x = "Macroalgal evenness", title = "3a.") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+#ggsave(filename = "output/figure_3b.png", figure_3b, height = 10, width = 14)
+```
+
+## Figure 3E site level richness versus cover
 

### 2024-09-13 manuscript_1_code.Rmd
added site level evenness for figure 3

@@
 
-filtered_cover_effects_evenness <- cover_emm %>%
+filtered_cover_effects_evenness <- cover_emm_evenness %>%
   left_join(diversity_ranges_quad_evenness, by = c("habitat")) %>%
@@
 # ggsave(filename = "output/figure_3d.png", figure_3d, height = 10, width = 14)
+```
+
+# Figure 3f site level evenness versus cover
+
+## model
+
+```{r evenness cover model site level}
+evenness_cover_mod_site <- glmmTMB(cover_trans ~ evenness*habitat + (1|site) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover_evenness_mod)
+summary(evenness_cover_mod_site)
+hist(residuals(evenness_cover_mod_site)) # little wonky but not the worst
+plot(residuals(evenness_cover_mod_site) ~ fitted(evenness_cover_mod_site))
+Anova(evenness_cover_mod_site)
+#                   Chisq Df Pr(>Chisq)    
+# evenness         14.395  1  0.0001482 ***
+# habitat          45.636  3  6.777e-10 ***
+# evenness:habitat 16.126  3  0.0010687 ** 
+r.squaredGLMM(evenness_cover_mod_site) # 0.166, 0.258
+
+emtrends(evenness_cover_mod_site, pairwise ~ habitat, var = "evenness") # fringing different from backreef and outer10
+```
+
+## plot
+
+### setup
+
+```{r figure 3f evenness site plot setup}
+diversity_ranges_site_evenness <- alpha_diversity_site_macro_with_cover_evenness_mod %>%
+  group_by(habitat) %>%
+  summarise(min_evenness = min(evenness),
+            max_evenness = max(evenness),
+            min_cover= min(cover_trans),
+            max_cover = max(cover_trans))
+
+cover_emm_evenness_site <- emmip(evenness_cover_mod_site,
+                        habitat ~ evenness,
+                        at = list(evenness = seq(0,1,0.01)), plotit = F, CIs = T)
+
+filtered_cover_effects_evenness_site <- cover_emm_evenness_site %>%
+  left_join(diversity_ranges_site_evenness, by = c("habitat")) %>%
+  filter(evenness >= min_evenness & evenness <= max_evenness)
+
+alpha_diversity_site_macro_with_cover_evenness_mod$habitat <- factor(alpha_diversity_site_macro_with_cover_evenness_mod$habitat, 
+                             levels = c("fringing", "backreef", "outer_10","outer_17"),
+                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+
+```
+
+### plot
+
+```{r evenness cover plot site }
+(figure_3f <- 
+  ggplot() +
+  geom_point(data = alpha_diversity_site_macro_with_cover_evenness_mod,
+             aes(x = evenness, y = cover_trans, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_cover_effects_evenness_site, aes(x = evenness, y = inv_logit(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_cover_effects_evenness_site, aes(x = evenness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Cover\n", x = "Macroalgal evenness", title = "3f. Site Level Evenness") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+# ggsave(filename = "output/figure_3f.png", figure_3f, height = 10, width = 14)
 ```

### 2024-09-13 manuscript_1_code.Rmd
figure 4 evenness site level done. code is messy as fuck we need to clean the markdown.

@@
             cover_mean = mean(prop_cover),
+            evenness_mean = mean(evenness),
             cover_cv = CV(prop_cover),
@@
 
-dss_ranges_site<- diversity_stability_synchrony_site %>%
+dss_ranges_site <- diversity_stability_synchrony_site %>%
   group_by(habitat) %>%
@@
             max_richness = max(richness_mean),
+            min_evenness = min(evenness_mean, na.rm = T),
+            max_evenness = max(evenness_mean, na.rm = T),
             min_stability= min(cover_stability),
@@
 ```
+
+# evenness stability site
+
+## mod
+
+```{r evenness synchrony mod}
+even_stab_mod_site <- glmmTMB(cover_stability~ evenness_mean*habitat, family = Gamma(link = "log"), data = diversity_stability_synchrony_site)
+summary(even_stab_mod_site)
+Anova(even_stab_mod_site)
+
+#                         Chisq Df Pr(>Chisq)   
+# evenness_mean          0.1035  1   0.747623   
+# habitat               11.8743  3   0.007827 **
+# evenness_mean:habitat  9.6254  3   0.022034 *   
+hist(residuals(even_stab_mod_site)) # blocky with some pos skew
+plot(residuals(even_stab_mod_site) ~ fitted(even_stab_mod_site)) # heteroskedastic 
+r.squaredGLMM(even_stab_mod_site) # 0.505
+emtrends(even_stab_mod_site, pairwise ~ habitat, var = "evenness_mean") # no differences across habitats
+```
+
+```{r figure 4d setup}
+
+even_stab_emm_site <- emmip(even_stab_mod_site,
+                        habitat ~ evenness_mean,
+                        at = list(evenness_mean = seq(0,1,0.01)), plotit = F, CIs = T)
+#make levels for habitat for plotting
+
+filtered_even_stab_effects_site <- even_stab_emm_site %>%
+  left_join(dss_ranges_site, by = c("habitat")) %>%
+  filter(evenness_mean >= min_evenness & evenness_mean <= max_evenness)
+
+#make levels for habitat for plotting
+diversity_stability_synchrony_site$habitat <- factor(diversity_stability_synchrony_site$habitat, 
+                             levels = c("fringing", "backreef", "outer_10","outer_17"),
+                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+
+```
+
+## plot figure 4d
+
+```{r figure 4d}
+(figure_4_evenness <- 
+  ggplot() +
+  geom_point(data = diversity_stability_synchrony_site,
+             aes(x = evenness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_even_stab_effects_site, aes(x = evenness_mean, y = exp(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_even_stab_effects_site, aes(x = evenness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Stability\n", x = "Macroalgal evenness", title = "4f. Evenness site") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        plot.title = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+ggsave(filename = "output/figure_4_evenness_site.png", figure_4_evenness, height = 10, width = 14)
+```

### 2024-09-13 manuscript_1_code.Rmd
fixed figure 2

@@
   scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
   #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
@@
 
-evenness_df <- alpha_diversity_quad_macro_with_cover %>% 
+quad_evenness_df <- alpha_diversity_quad_macro_with_cover %>% 
   drop_na(evenness) %>%
   group_by(year, habitat) %>% 
-  summarise(eveness_mean = mean(evenness),
-            eveness_se = se(evenness))
-
-evenness_df$habitat <- factor(evenness_df$habitat, 
+  summarise(evenness_mean = mean(evenness),
+            evenness_se = se(evenness))
+
+quad_evenness_df$habitat <- factor(quad_evenness_df$habitat, 
                              levels = c("fringing", "backreef", "outer_10","outer_17"),
@@
 
-```{r plot evenness}
+```{r}
 (figure_2c <- ggplot() +
-  geom_line(data = evenness_df, aes(x = year, y = eveness_mean, colour = habitat), linewidth = 1) +
-  geom_ribbon(data = evenness_df, aes(x = year, ymin = eveness_mean - eveness_se, ymax = eveness_mean + eveness_se, fill = habitat), alpha = 0.3) +
+  geom_point(data = quad_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), size = 3) +
+  geom_line(data = quad_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), linewidth = 1) +
+  geom_errorbar(data = quad_evenness_df, aes(x = year, ymin = evenness_mean - evenness_se, ymax = evenness_mean + evenness_se,
+                                             colour = habitat), width = 0, size = 1.3) +
   scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
+  # scale_fill_manual(values = habitat_colours) +
   #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
-  annotate("segment", x = 2006, xend = 2010, y = 0.7, yend = 0.7, linewidth = 2, colour = "grey") +
+  annotate("segment", x = 2006, xend = 2010, y = .7, yend = .7, linewidth = 2, colour = "grey") +
   geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
   geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
-   scale_x_continuous(breaks = unique(evenness_df$year)) +
-   labs(y = "Macroalgal Evenness \n",
-        x = "Year") +
+   scale_x_continuous(breaks = unique(quad_evenness_df$year)) +
+   labs(y = "Macroalgal evenness \n",
+        x = "Year",
+        title = "c. Quadrat-level evenness") +
   theme_bw() + 
@@
         axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_blank(),
+        plot.title = element_text(size=18),
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
         legend.text = element_text(size = rel(1)), # increase legend text
@@
 
-#ggsave(filename = "output/figure_2c.png", figure_2c, height = 10, width = 14)
-
-#hist(eveness_df$eveness_mean)
+# ggsave(filename = "output/figure_2c.png", figure_2c, height = 10, width = 14)
 ```
@@
   group_by(year, habitat) %>% 
-  summarise(eveness_mean = mean(evenness),
-            eveness_se = se(evenness))
+  summarise(evenness_mean = mean(evenness),
+            evenness_se = se(evenness))
 
@@
 
-```{r plot evenness}
+```{r plot evenness site}
 (figure_2d <- ggplot() +
-  geom_line(data = site_evenness_df, aes(x = year, y = eveness_mean, colour = habitat), linewidth = 1) +
-  geom_ribbon(data = site_evenness_df, aes(x = year, ymin = eveness_mean - eveness_se, ymax = eveness_mean + eveness_se, fill = habitat), alpha = 0.3) +
+  geom_point(data = site_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), size = 3) +
+  geom_line(data = site_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), linewidth = 1) +
+  geom_errorbar(data = site_evenness_df, aes(x = year, ymin = evenness_mean - evenness_se, ymax = evenness_mean + evenness_se, colour = habitat), size = 1.3, width = 0) +
   scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
   #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
@@
   geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
-   scale_x_continuous(breaks = unique(evenness_df$year)) +
-   labs(y = "Macroalgal Evenness \n",
-        x = "Year") +
+   scale_x_continuous(breaks = unique(site_evenness_df$year)) +
+   labs(y = "Macroalgal evenness \n",
+        x = "Year",
+        title = "d. Site-level evenness") +
   theme_bw() + 
@@
 
-#ggsave(filename = "output/figure_2d.png", figure_2d, height = 10, width = 14)
+# ggsave(filename = "output/figure_2d.png", figure_2d, height = 10, width = 14)
 ```
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Cover\n", x = "Macroalgal evenness", title = "3a.") +
+  labs(y = "Cover\n", x = "Macroalgal evenness", title = "3b.") +
   theme_bw() + 

### 2024-09-13 manuscript_1_code.Rmd
figuring evenness out man

--- 2024-09-13 manuscript_1_code.Rmd
@@
             cover_mean = mean(macroalgal_prop_cover),
+            evenness_mean = mean(evenness, na.rm = T),
             cover_cv = CV(macroalgal_prop_cover),
@@
 diversity_stability_synchrony$synchrony_trans <- sv_trans(diversity_stability_synchrony$synchrony)
-diversity_stability_synchrony <- na.omit(diversity_stability_synchrony) # remove quads that never had algae --> NaN calcs
+# diversity_stability_synchrony <- na.omit(diversity_stability_synchrony) # remove quads that never had algae --> NaN calcs
 
@@
             max_richness = max(richness_mean),
-            min_stability= min(cover_stability),
-            max_stability = max(cover_stability),
-            min_synchrony = min(synchrony_trans),
-            max_synchrony = max(synchrony_trans))
+            min_evenness = min(richness_mean, na.rm = TRUE),
+            max_evenness = max(richness_mean, na.rm = TRUE),
+            min_stability= min(cover_stability, na.rm = TRUE),
+            max_stability = max(cover_stability, na.rm = TRUE),
+            min_synchrony = min(synchrony_trans, na.rm = TRUE),
+            max_synchrony = max(synchrony_trans, na.rm = TRUE))
 ```

### 2024-09-13 julianna_gut_check.Rmd
Added a script to gut check site vs. quad level comparisons

---
title: "Gut check: site-level analyses"
output: html_document
date: "2024-09-13"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clean up
rm(list=ls())

# load packages
require(here) # relative file paths
require(tidyverse) # data wrangling
require(vegan) # community/diversity analyses
```

# Load Callie's data

```{r}
synchrony <- read_csv(here("data/diversity_stability_synchrony.csv"))
alpha <- read_csv(here("data/alpha_diversity_quad_macro_with_cover.csv"))
```

# Summarize

```{r synchrony summary}
synchrony %>% 
  group_by(habitat, site) %>% 
  summarize(richness_mean_mean = mean(richness_mean), 
            cover_mean_mean = mean(cover_mean), 
            cover_cv_mean = mean(cover_cv), 
            cover_stability_mean = mean(cover_stability), 
            synchrony_trans_mean = mean(synchrony_trans)) -> synchrony_mean
```

```{r alpha summary}
alpha %>% 
  group_by(year, habitat, site) %>% 
  summarize(richness_mean = mean(richness), 
            macroalgal_prop_cover_mean = mean(macroalgal_prop_cover), 
            cover_trans_mean = mean(cover_trans)) -> alpha_mean
```

# Plot synchrony

```{r}
synchrony_mean
```

### 2024-09-13 manuscript_1_code.Rmd

@@
 
+### plot
 ```{r}
@@
 ## 2c Evenness @ Quad by Year
-
+### prep
 ```{r calculate evennes}
@@
 ## 2d Evenness @ Site by Year
-
+### prep
 ```{r set up evenness}
@@
 
+### plot
 ```{r plot evenness site}

### 2024-09-14 julianna_gut_check.Rmd
Finished gut check code

@@
   group_by(habitat, site) %>% 
-  summarize(richness_mean_mean = mean(richness_mean), 
-            cover_mean_mean = mean(cover_mean), 
-            cover_cv_mean = mean(cover_cv), 
-            cover_stability_mean = mean(cover_stability), 
-            synchrony_trans_mean = mean(synchrony_trans)) -> synchrony_mean
+  summarize(richness_mean_mean = mean(richness_mean, na.rm = TRUE), 
+            cover_mean_mean = mean(cover_mean, na.rm = TRUE), 
+            cover_cv_mean = mean(cover_cv, na.rm = TRUE), 
+            cover_stability_mean = mean(cover_stability, na.rm = TRUE), 
+            synchrony_trans_mean = mean(synchrony_trans, na.rm = TRUE)) -> synchrony_mean
 ```
@@
   group_by(year, habitat, site) %>% 
-  summarize(richness_mean = mean(richness), 
-            macroalgal_prop_cover_mean = mean(macroalgal_prop_cover), 
-            cover_trans_mean = mean(cover_trans)) -> alpha_mean
+  summarize(richness_mean = mean(richness, na.rm = TRUE), 
+            macroalgal_prop_cover_mean = mean(macroalgal_prop_cover, na.rm = TRUE), 
+            cover_trans_mean = mean(cover_trans, na.rm = TRUE)) -> alpha_mean
 ```
 
-# Plot synchrony
+# Fig 3
 
 ```{r}
-synchrony_mean
+synchrony_mean %>% 
+  ggplot(aes(x = richness_mean_mean, y = cover_mean_mean, color = habitat, fill = habitat)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  theme_classic()
 ```
+
+```{r}
+synchrony_mean %>% 
+  ggplot(aes(x = synchrony_trans_mean, y = cover_mean_mean, color = habitat, fill = habitat)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  theme_classic()
+```
+
+# Fig 4
+
+Stability on y, richness/synchrony on x
+
+```{r richness}
+synchrony_mean %>% 
+  ggplot(aes(x = richness_mean_mean, y = 1/cover_cv_mean, color = habitat, fill = habitat)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  theme_classic()
+```
+
+```{r synchrony}
+synchrony_mean %>% 
+  ggplot(aes(x = synchrony_trans_mean, y = 1/cover_cv_mean, color = habitat, fill = habitat)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  theme_classic()
+```

### 2024-09-14 manuscript_1_code.Rmd
Just trying to avoid problems here

@@
 plot(residuals(synch_cov_mod) ~ fitted(synch_cov_mod)) # same issues as always
+
+r.squaredGLMM(synch_cov_mod) # error??
 emtrends(synch_cov_mod, pairwise ~ habitat, var = "synchrony") # forereef 17 distinct from all other habitats

### 2024-09-20 manuscript_1_code.Rmd
fixed habitat levels to be in the original dataframe :tropical_fish:

@@
 
+#THIS IS THE BIG CHANGE HERE: 9/19
+filtered_data$habitat <- factor(filtered_data$habitat, 
+                             levels = c("fringing", "backreef", "outer_10","outer_17"),
+                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+
+#DID NOT TOUCH BELOW:
 not_algae <- c("ascidian", "bare_space", "coral", "coral_rubble", "corallimorpharia", "heteractis_sp", "millepora_platyphylla", "no_data", "sand", "sarcophyton_sp", "shell_debris", "soft_coral", "sponge", "tridacna_sp", "zooxanthellae", "row_sums")
@@
 macro_species_cols <- macroalgae[,-which(names(macroalgae) %in% meta_cols)]
-
 ```
@@
 
-cover_long <- left_join(cover_long_means,cover_long_se)
-
-#make levels for habitat for plotting
-cover_long$habitat <- factor(cover_long$habitat, 
-                             levels = c("fringing", "backreef", "outer_10","outer_17"),
-                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
-
+cover_long <- left_join(cover_long_means,cover_long_se, by = join_by(year, habitat, functional_group))
+
+#make levels of functional group for plotting
 cover_long$functional_group <- factor(cover_long$functional_group, 
@@
 
-```{r simplified cover - remove}
-
-# simplified_cover <- data.frame(
-#   year = cover_df$year,
-#   habitat = cover_df$habitat,
-#   site = cover_df$site,
-#   location = cover_df$location,
-#   low_lying_algae = cover_df$cca + cover_df$turf + cover_df$cyanophyta,
-#   coral = cover_df$coral,
-#   macroalgae = cover_df$macroalgae
-# )
-# 
-# simplified_cover_means <- simplified_cover %>%
-#   group_by(year, habitat) %>%
-#   summarise(macroalgae_mean = mean(macroalgae),
-#             low_lying_algae_mean = mean(low_lying_algae),
-#             coral_mean = mean(coral),
-#             macroalgae_se = sd(macroalgae)/sqrt(n()),
-#             low_lying_algae_se = sd(low_lying_algae)/sqrt(n()),
-#             coral_se = sd(coral)/sqrt(n())
-#             )
-# 
-# simplified_cover_long <- pivot_longer(
-#   data = simplified_cover_means,
-#   cols = 3:5,
-#   names_to = "functional_group",
-#   values_to = "percent_cover"
-# )
-# 
-# simplified_cover_long_se <- pivot_longer(
-#   data = simplified_cover_means,
-#   cols = 6:8,
-#   names_to = "functional_group",
-#   values_to = "se"
-# )
-# simplified_cover_long$se <- simplified_cover_long_se$se
-
-```
-
-```{r}
+```{r make figure1 data}
 figure_1_data <- subset(cover_long, functional_group != "Sand" & functional_group != "Other")
@@
                 width = 0, size = 1.3) +
-   geom_line(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), size = 1) +
+   geom_line(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
   facet_wrap(~ habitat) +
@@
 
-```{r richness factors}
-quad_richness_df$habitat <- factor(quad_richness_df$habitat, 
-                             levels = c("fringing", "backreef", "outer_10","outer_17"),
-                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
-```
-
 ### plot
@@
 
-```{r richness factors}
-site_richness_df$habitat <- factor(site_richness_df$habitat, 
-                             levels = c("fringing", "backreef", "outer_10","outer_17"),
-                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
-
-```
-
 ### plot
@@
             evenness_se = se(evenness))
-
-quad_evenness_df$habitat <- factor(quad_evenness_df$habitat, 
-                             levels = c("fringing", "backreef", "outer_10","outer_17"),
-                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
 ```
@@
             evenness_se = se(evenness))
-
-site_evenness_df$habitat <- factor(site_evenness_df$habitat, 
-                             levels = c("fringing", "backreef", "outer_10","outer_17"),
-                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
 ```
@@
 
-alpha_diversity_quad_macro_with_cover$habitat <- factor(alpha_diversity_quad_macro_with_cover$habitat, 
-                             levels = c("fringing", "backreef", "outer_10","outer_17"),
-                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
-
-filtered_cover_effects$habitat <- factor(filtered_cover_effects$habitat, 
-                             levels = c("fringing", "backreef", "outer_10","outer_17"),
-                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+# alpha_diversity_quad_macro_with_cover$habitat <- factor(alpha_diversity_quad_macro_with_cover$habitat, 
+#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
+#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+# 
+# 
+# filtered_cover_effects$habitat <- factor(filtered_cover_effects$habitat, 
+#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
+#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
 ```
@@
 ## Figure 3B quad level evenness versus cover
-```{r COME BACK HERE figure 3b setup}
+```{r figure 3b setup}
 diversity_ranges_quad_evenness <- alpha_diversity_quad_macro_with_cover_evenness_mod %>%
@@
 
-alpha_diversity_quad_macro_with_cover_evenness_mod$habitat <- factor(alpha_diversity_quad_macro_with_cover_evenness_mod$habitat, 
-                             levels = c("fringing", "backreef", "outer_10","outer_17"),
-                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
-
-filtered_cover_effects_evenness$habitat <- factor(filtered_cover_effects_evenness$habitat, 
-                             levels = c("fringing", "backreef", "outer_10","outer_17"),
-                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+# alpha_diversity_quad_macro_with_cover_evenness_mod$habitat <- factor(alpha_diversity_quad_macro_with_cover_evenness_mod$habitat, 
+#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
+#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+# 
+# 
+# filtered_cover_effects_evenness$habitat <- factor(filtered_cover_effects_evenness$habitat, 
+#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
+#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
 ```
@@
 
-alpha_diversity_site_macro_with_cover$habitat <- factor(alpha_diversity_site_macro_with_cover$habitat, 
-                             levels = c("fringing", "backreef", "outer_10","outer_17"),
-                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
-
-filtered_site_cover_effects$habitat <- factor(filtered_site_cover_effects$habitat, 
-                             levels = c("fringing", "backreef", "outer_10","outer_17"),
-                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+# alpha_diversity_site_macro_with_cover$habitat <- factor(alpha_diversity_site_macro_with_cover$habitat, 
+#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
+#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+# 
+# 
+# filtered_site_cover_effects$habitat <- factor(filtered_site_cover_effects$habitat, 
+#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
+#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
 ```
@@
 
-alpha_diversity_site_macro_with_cover_evenness_mod$habitat <- factor(alpha_diversity_site_macro_with_cover_evenness_mod$habitat, 
-                             levels = c("fringing", "backreef", "outer_10","outer_17"),
-                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
-
+# alpha_diversity_site_macro_with_cover_evenness_mod$habitat <- factor(alpha_diversity_site_macro_with_cover_evenness_mod$habitat, 
+#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
+#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+# 
+# 
+# 
 ```
@@
 ```{r figure 4a setup}
-
 rich_stab_emm <- emmip(rich_stab_mod,
@@
   filter(richness_mean >= min_richness & richness_mean <= max_richness)
-
 ```
@@
 ```{r figure 4c setup}
-
 synch_stab_emm <- emmip(synch_stab_mod,
@@
   filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
-
 ```
@@
 ```{r figure 4d setup}
-
 rich_stab_emm_site <- emmip(rich_stab_mod_site,
@@
 ```{r figure 4f setup}
-
 synch_stab_emm_site <- emmip(synch_stab_mod_site,
@@
 ```{r figure 4d setup}
-
 even_stab_emm_site <- emmip(even_stab_mod_site,
@@
 #make levels for habitat for plotting
-diversity_stability_synchrony_site$habitat <- factor(diversity_stability_synchrony_site$habitat, 
-                             levels = c("fringing", "backreef", "outer_10","outer_17"),
-                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
-
+# diversity_stability_synchrony_site$habitat <- factor(diversity_stability_synchrony_site$habitat, 
+#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
+#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
 ```

### 2024-09-20 manuscript_1_code.Rmd
cleaned up figure 1 and 2, need to add code for functional richness on 3 and 4. :tropical_fish:

@@
 
+```{r Figure 1 v2}
+plot_list <- lapply(levels(figure_1_data$habitat), function(hab) {
+  ggplot(data = subset(figure_1_data, habitat == hab)) +
+    geom_point(aes(x = year, y = percent_cover, colour = functional_group), size = 3) +
+    geom_errorbar(aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, colour = functional_group), 
+                  width = 0, size = 1.3) +
+    geom_line(aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
+    annotate("segment", x = 2006, xend = 2010, y = 75, yend = 75, linewidth = 2, colour = "grey") +
+    geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+    geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+    scale_colour_manual(values = functional_group_colours) +
+    scale_x_continuous(breaks = unique(figure_1_data$year)) +
+    labs(y = "Percent Cover",
+         x = "Year",
+         title = hab) +  # Add habitat as title
+    theme_bw() +
+    theme(axis.title.x = element_text(size = 18),
+          axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
+          axis.title.y = element_text(size = 18),
+          axis.text.y = element_text(size = 12),
+          plot.title = element_text(size = 20, hjust = 0),
+          legend.text = element_text(size = 14), 
+          legend.title = element_blank(),
+          legend.position = 'bottom',
+          legend.box = 'horizontal',
+          legend.box.background = element_rect(colour = "black"))
+})
+
+figure_1 <- ggarrange(plotlist = plot_list, 
+                      ncol = 2, nrow = 2, 
+                      labels = c("a.", "b.", "c.", "d."),
+                      common.legend = TRUE,
+                      legend = "bottom", 
+                      font.label = list(size = 20, color = "black", face = "plain"))
+# 
+# figure_1_caption <- "Figure 1.
+Benthic coverage of four major space holders in Moorea across four habitats (Fringing, Backreef, Forereef at 10m depth, Forereef  at 17m depth).
+Annual fluctuations in percent cover of macroalgae, coral, turf, and crustose coralline algae (CCA) are displayed as points for the average cover for each benthic group across sites at each habitat with lines showing standard error.
+Dotted lines show disturbance events for Cyclone Oli (2010) in black, and the 2019 bleaching event in red, and the gray horizontal line shows crown of thorns (COTs) outbreak which occurred from 2007-2010."
+# 
+# figure_1_with_caption <- annotate_figure(figure_1,
+#                 bottom = text_grob(figure_1_caption, hjust = 0.5, vjust = 1, size = 10))
+
+#ggsave(filename = "output/figure_1v2.png", figure_1, height = 10, width = 14)
+```
+
 # Figure 2
@@
 (figure_2a <- ggplot() +
-  geom_point(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 3) +
+  geom_point(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4) +
   geom_line(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
@@
    scale_x_continuous(breaks = unique(quad_richness_df$year)) +
-   labs(y = "Macroalgal Richness \n",
+   labs(y = "Macroalgal Richness",
         x = "Year",
-        title = "a. Quadrat-level Richness") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.text.x = element_text(angle = 45, hjust = 1),
+        title = "Quadrat-level Richness") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.text.x = element_text(angle = 45, hjust = 1, size=14),
+        axis.text.y = element_text(size=12),
         axis.title.y = element_text(size=18), 
@@
 (figure_2b <- ggplot() +
-  geom_point(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 3) +
+  geom_point(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4) +
+  geom_jitter(data = alpha_diversity_site_macro, aes(x = year, y = richness, colour = habitat, alpha = 0.05), size = 2, width = 0.4, height = 0) +
   geom_line(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
@@
   #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
-  annotate("segment", x = 2006, xend = 2010, y = 11, yend = 11, linewidth = 2, colour = "grey") +
+  annotate("segment", x = 2006, xend = 2010, y = 14, yend = 14, linewidth = 2, colour = "grey") +
   geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
@@
    scale_x_continuous(breaks = unique(site_richness_df$year)) +
-   labs(y = "Macroalgal Richness \n",
+   labs(y = "Macroalgal Richness",
         x = "Year",
-        title = "b. Site-level Richness") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.text.x = element_text(angle = 45, hjust = 1),
-        axis.title.y = element_text(size=18), 
+        title = "Site-level Richness") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.text.x = element_text(angle = 45, hjust = 1, size=14),
+        axis.title.y = element_text(size=18), 
+        axis.text.y = element_text(size=14),
         plot.title = element_text(size=18),
@@
 
-## 2c Evenness @ Quad by Year
-### prep
-```{r calculate evennes}
-#dear callie, why is this here?
-site_richness_df <- alpha_diversity_site_macro %>% 
-  group_by(year, habitat) %>% 
-  summarise(richness_mean = mean(richness),
-            richness_se = se(richness))
-
-quad_evenness_df <- alpha_diversity_quad_macro_with_cover %>% 
-  drop_na(evenness) %>%
-  group_by(year, habitat) %>% 
-  summarise(evenness_mean = mean(evenness),
-            evenness_se = se(evenness))
-```
-
-### plot
-```{r}
-(figure_2c <- ggplot() +
-  geom_point(data = quad_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), size = 3) +
-  geom_line(data = quad_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), linewidth = 1) +
-  geom_errorbar(data = quad_evenness_df, aes(x = year, ymin = evenness_mean - evenness_se, ymax = evenness_mean + evenness_se,
-                                             colour = habitat), width = 0, size = 1.3) +
-  scale_colour_manual(values = habitat_colours) +
-  # scale_fill_manual(values = habitat_colours) +
-  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
-  annotate("segment", x = 2006, xend = 2010, y = .7, yend = .7, linewidth = 2, colour = "grey") +
-  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
-  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
-   scale_x_continuous(breaks = unique(quad_evenness_df$year)) +
-   labs(y = "Macroalgal evenness \n",
-        x = "Year",
-        title = "c. Quadrat-level evenness") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.text.x = element_text(angle = 45, hjust = 1),
-        axis.title.y = element_text(size=18), 
-        plot.title = element_text(size=18),
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black")) )
-
-# ggsave(filename = "output/figure_2c.png", figure_2c, height = 10, width = 14)
-```
-
-## 2d Evenness @ Site by Year
-### prep
-```{r set up evenness}
-site_evenness_df <- alpha_diversity_site_macro %>% 
-  drop_na(evenness) %>%
-  group_by(year, habitat) %>% 
-  summarise(evenness_mean = mean(evenness),
-            evenness_se = se(evenness))
-```
-
-### plot
-```{r plot evenness site}
-(figure_2d <- ggplot() +
-  geom_point(data = site_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), size = 3) +
-  geom_line(data = site_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), linewidth = 1) +
-  geom_errorbar(data = site_evenness_df, aes(x = year, ymin = evenness_mean - evenness_se, ymax = evenness_mean + evenness_se, colour = habitat), size = 1.3, width = 0) +
-  scale_colour_manual(values = habitat_colours) +
-  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
-  annotate("segment", x = 2006, xend = 2010, y = 1, yend = 1, linewidth = 2, colour = "grey") +
-  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
-  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
-   scale_x_continuous(breaks = unique(site_evenness_df$year)) +
-   labs(y = "Macroalgal evenness \n",
-        x = "Year",
-        title = "d. Site-level evenness") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.text.x = element_text(angle = 45, hjust = 1),
-        axis.title.y = element_text(size=18), 
-        plot.title = element_text(size=18),
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black")) )
-
-# ggsave(filename = "output/figure_2d.png", figure_2d, height = 10, width = 14)
+## Figure 2 Full Plot
+```{r ggpubr fig2}
+fig2_plot_list <- list(figure_2a, figure_2b)
+figure_2 <- ggarrange(plotlist = fig2_plot_list, 
+                      ncol = 2, nrow = 1, 
+                      labels = c("a.", "b."),
+                      common.legend = TRUE,
+                      legend = "bottom", 
+                      font.label = list(size = 20, color = "black", face = "plain"))
+
+#ggsave(filename = "output/figure_2.png", figure_2, height = 10, width = 20)
 ```
@@
 
+## Figure 3c
+
+### synchrony set up
+```{r calculate syncrony}
+# make er long
+macro_long_data <- pivot_longer(
+  data = macroalgae,
+  cols = grep("acanthophora_spicifera", colnames(macroalgae)):grep("valonia_ventricosa", colnames(macroalgae)),
+  names_to = "taxa",
+  values_to = "percent_cover") %>% 
+  mutate(prop_cover = percent_cover/100) %>% 
+  select(-percent_cover)
+
+# write_csv(macro_long_data, here("Data", "macro_long_data.csv"))
+
+# CALCULATE SYNCHRONY CODE
+lm_synchrony <- codyn::synchrony(df = macro_long_data,
+                                   time.var = "year",
+                                   species.var = "taxa",
+                                   abundance.var = "prop_cover",
+                                   metric = "Loreau",
+                                   replicate.var = "location")
+
+diversity_stability <- alpha_diversity_quad_macro_with_cover %>%
+  dplyr::group_by(location, habitat, site) %>%
+  dplyr::summarise(
+            richness_mean = mean(richness), 
+            shannon_mean = mean(shannon), 
+            cover_mean = mean(macroalgal_prop_cover),
+            evenness_mean = mean(evenness, na.rm = T),
+            cover_cv = CV(macroalgal_prop_cover),
+            cover_stability = 1/CV(macroalgal_prop_cover))
+
+diversity_stability_synchrony <- left_join(diversity_stability, lm_synchrony, by = join_by(location))
+diversity_stability_synchrony$synchrony_trans <- sv_trans(diversity_stability_synchrony$synchrony)
+# diversity_stability_synchrony <- na.omit(diversity_stability_synchrony) # remove quads that never had algae --> NaN calcs
+
+dss_ranges<- diversity_stability_synchrony %>%
+  group_by(habitat) %>%
+  summarise(min_richness = min(richness_mean),
+            max_richness = max(richness_mean),
+            min_evenness = min(richness_mean, na.rm = TRUE),
+            max_evenness = max(richness_mean, na.rm = TRUE),
+            min_stability= min(cover_stability, na.rm = TRUE),
+            max_stability = max(cover_stability, na.rm = TRUE),
+            min_synchrony = min(synchrony_trans, na.rm = TRUE),
+            max_synchrony = max(synchrony_trans, na.rm = TRUE))
+```
+
+### synchrony and cover model
+
+```{r synchrony cover mod}
+synch_cov_mod <- glmmTMB(cover_mean ~ synchrony*habitat + (1|site), family = beta_family(), data = diversity_stability_synchrony)
+summary(synch_cov_mod)
+Anova(synch_cov_mod)
+#                     Chisq Df Pr(>Chisq)    
+# synchrony         209.545  1  < 2.2e-16 ***
+# habitat            52.512  3  2.330e-11 ***
+# synchrony:habitat  54.632  3  8.227e-12 ***
+hist(residuals(synch_cov_mod)) # slight skew
+plot(residuals(synch_cov_mod) ~ fitted(synch_cov_mod)) # same issues as always
+
+r.squaredGLMM(synch_cov_mod) # error??
+emtrends(synch_cov_mod, pairwise ~ habitat, var = "synchrony") # forereef 17 distinct from all other habitats
+```
+
+```{r synch cov plot setup}
+synch_cov_emm <- emmip(synch_cov_mod,
+                        habitat ~ synchrony,
+                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)
+
+filtered_synch_cov_effects <- synch_cov_emm %>%
+  left_join(dss_ranges, by = c("habitat")) %>%
+  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
+```
+
+### plot
+
+```{r synchrony cover plot}
+(figure_3c <- 
+  ggplot() +
+  geom_point(data = diversity_stability_synchrony,
+             aes(x = synchrony, y = cover_mean, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_synch_cov_effects, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_synch_cov_effects, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(x = "Synchrony", y = "Cover\n", title = "3c.") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18),
+        plot.title = element_text(size=18),
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+# ggsave(filename = "output/figure_3c.png", figure_3c, height = 10, width = 14)
+```
+
+### Synchrony setup site level
+```{r calculate syncrony site_level}
+# make er long
+macro_long_data_site <- pivot_longer(
+  data = site_macro,
+  cols = grep("acanthophora_spicifera", colnames(site_macro)):grep("valonia_ventricosa", colnames(site_macro)),
+  names_to = "taxa",
+  values_to = "percent_cover") %>% 
+  mutate(prop_cover = percent_cover/100) %>% 
+  select(-percent_cover)
+
+# write_csv(macro_long_data, here("Data", "macro_long_data.csv"))
+
+# CALCULATE SYNCHRONY CODE
+lm_synchrony_site <- codyn::synchrony(df = macro_long_data_site,
+                                   time.var = "year",
+                                   species.var = "taxa",
+                                   abundance.var = "prop_cover",
+                                   metric = "Loreau",
+                                   replicate.var = "site_habitat")
+
+diversity_stability_site <- alpha_diversity_site_macro_with_cover %>%
+  dplyr::group_by(site_habitat, habitat, site) %>%
+  dplyr::summarise(
+            richness_mean = mean(richness), 
+            shannon_mean = mean(shannon), 
+            cover_mean = mean(prop_cover),
+            evenness_mean = mean(evenness),
+            cover_cv = CV(prop_cover),
+            cover_stability = 1/CV(prop_cover))
+
+diversity_stability_synchrony_site <- left_join(diversity_stability_site, lm_synchrony_site, by = join_by(site_habitat))
+diversity_stability_synchrony_site$synchrony_trans <- sv_trans(diversity_stability_synchrony_site$synchrony)
+# diversity_stability_synchrony_site <- na.omit(diversity_stability_synchrony_site) # remove quads that never had algae --> NaN calcs
+# write_csv(diversity_stability_synchrony_site, here("data", "diversity_stability_synchrony_site.csv"))
+
+dss_ranges_site <- diversity_stability_synchrony_site %>%
+  group_by(habitat) %>%
+  summarise(min_richness = min(richness_mean),
+            max_richness = max(richness_mean),
+            min_evenness = min(evenness_mean, na.rm = T),
+            max_evenness = max(evenness_mean, na.rm = T),
+            min_stability= min(cover_stability),
+            max_stability = max(cover_stability),
+            min_synchrony = min(synchrony_trans),
+            max_synchrony = max(synchrony_trans))
+```
+
 ## Figure 3E site level richness versus cover
@@
 
-# Figure 3f site level evenness versus cover
-
-## model
+## figure 3f
+
+### model
+
+```{r synchrony cover mod}
+synch_cov_mod_site <- glmmTMB(cover_mean ~ synchrony*habitat, family = beta_family(), data = diversity_stability_synchrony_site)
+summary(synch_cov_mod_site)
+Anova(synch_cov_mod_site)
+#                    Chisq Df Pr(>Chisq)  
+# synchrony         2.4184  1    0.11992  
+# habitat           7.8002  3    0.05033 .
+# synchrony:habitat 8.8866  3    0.03084 *
+hist(residuals(synch_cov_mod_site)) # slight skew
+plot(residuals(synch_cov_mod_site) ~ fitted(synch_cov_mod_site)) # looks fine
+emtrends(synch_cov_mod_site, pairwise ~ habitat, var = "synchrony") # forereef 17 different from backreef
+```
+
+```{r synch cov site plot setup}
+synch_cov_emm_site<- emmip(synch_cov_mod_site,
+                        habitat ~ synchrony,
+                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)
+
+filtered_synch_cov_effects_site <- synch_cov_emm_site %>%
+  left_join(dss_ranges_site, by = c("habitat")) %>%
+  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
+```
+
+### plot figure 3f
+
+```{r synchrony cover plot site}
+(figure_3f <- 
+  ggplot() +
+  geom_point(data = diversity_stability_synchrony_site,
+             aes(x = synchrony, y = cover_mean, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_synch_cov_effects_site, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_synch_cov_effects_site, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(x = "Synchrony", y = "Cover\n", title = "3f.") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        plot.title = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+# ggsave(filename = "output/figure_3f.png", figure_3f, height = 10, width = 14)
+```
+
+## Figure 3f site level evenness versus cover
+
+### model
 
@@
 
-## plot
+### plot
 
@@
 
-## Synchrony setup
-```{r calculate syncrony}
-# make er long
-macro_long_data <- pivot_longer(
-  data = macroalgae,
-  cols = grep("acanthophora_spicifera", colnames(macroalgae)):grep("valonia_ventricosa", colnames(macroalgae)),
-  names_to = "taxa",
-  values_to = "percent_cover") %>% 
-  mutate(prop_cover = percent_cover/100) %>% 
-  select(-percent_cover)
-
-# write_csv(macro_long_data, here("Data", "macro_long_data.csv"))
-
-# CALCULATE SYNCHRONY CODE
-lm_synchrony <- codyn::synchrony(df = macro_long_data,
-                                   time.var = "year",
-                                   species.var = "taxa",
-                                   abundance.var = "prop_cover",
-                                   metric = "Loreau",
-                                   replicate.var = "location")
-
-diversity_stability <- alpha_diversity_quad_macro_with_cover %>%
-  dplyr::group_by(location, habitat, site) %>%
-  dplyr::summarise(
-            richness_mean = mean(richness), 
-            shannon_mean = mean(shannon), 
-            cover_mean = mean(macroalgal_prop_cover),
-            evenness_mean = mean(evenness, na.rm = T),
-            cover_cv = CV(macroalgal_prop_cover),
-            cover_stability = 1/CV(macroalgal_prop_cover))
-
-diversity_stability_synchrony <- left_join(diversity_stability, lm_synchrony, by = join_by(location))
-diversity_stability_synchrony$synchrony_trans <- sv_trans(diversity_stability_synchrony$synchrony)
-# diversity_stability_synchrony <- na.omit(diversity_stability_synchrony) # remove quads that never had algae --> NaN calcs
-
-dss_ranges<- diversity_stability_synchrony %>%
-  group_by(habitat) %>%
-  summarise(min_richness = min(richness_mean),
-            max_richness = max(richness_mean),
-            min_evenness = min(richness_mean, na.rm = TRUE),
-            max_evenness = max(richness_mean, na.rm = TRUE),
-            min_stability= min(cover_stability, na.rm = TRUE),
-            max_stability = max(cover_stability, na.rm = TRUE),
-            min_synchrony = min(synchrony_trans, na.rm = TRUE),
-            max_synchrony = max(synchrony_trans, na.rm = TRUE))
-```
-
-## figure 4a model - richness stability
+## Figure 4a
+
+###model - richness stability
 
@@
 
-## plot figure 4a
+### plot figure 4a
 
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Stability\n", x = "Macroalgal Richness", title = "4a.") +
+  labs(y = "Stability\n", x = "Taxonomic Richness", title = "4a.") +
   theme_bw() + 
@@
 
-## figure 4a no habitat
+#### no habitat
 
@@
 
-## figure 4a no habitat plot
+#### no habitat plot
 
@@
 
-## figure 4c model -  synchrony staability
+## figure 4c 
+### model -  synchrony staability
 
@@
 
-## plot figure 4c
+### plot figure 4c
 
@@
 
-# Figure 3c
-
-## synchrony and cover model
-
-```{r synchrony cover mod}
-synch_cov_mod <- glmmTMB(cover_mean ~ synchrony*habitat + (1|site), family = beta_family(), data = diversity_stability_synchrony)
-summary(synch_cov_mod)
-Anova(synch_cov_mod)
-#                     Chisq Df Pr(>Chisq)    
-# synchrony         209.545  1  < 2.2e-16 ***
-# habitat            52.512  3  2.330e-11 ***
-# synchrony:habitat  54.632  3  8.227e-12 ***
-hist(residuals(synch_cov_mod)) # slight skew
-plot(residuals(synch_cov_mod) ~ fitted(synch_cov_mod)) # same issues as always
-
-r.squaredGLMM(synch_cov_mod) # error??
-emtrends(synch_cov_mod, pairwise ~ habitat, var = "synchrony") # forereef 17 distinct from all other habitats
-```
-
-```{r synch cov plot setup}
-synch_cov_emm <- emmip(synch_cov_mod,
-                        habitat ~ synchrony,
-                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)
-
-filtered_synch_cov_effects <- synch_cov_emm %>%
-  left_join(dss_ranges, by = c("habitat")) %>%
-  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
-```
-
-## plot
-
-```{r synchrony cover plot}
-(figure_3c <- 
-  ggplot() +
-  geom_point(data = diversity_stability_synchrony,
-             aes(x = synchrony, y = cover_mean, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_synch_cov_effects, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
-  geom_ribbon(data = filtered_synch_cov_effects, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
-              alpha = 0.3) +
-  scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
-  labs(x = "Synchrony", y = "Cover\n", title = "3c.") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18),
-        plot.title = element_text(size=18),
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-# ggsave(filename = "output/figure_3c.png", figure_3c, height = 10, width = 14)
-```
-
-## Synchrony setup site level
-```{r calculate syncrony site_level}
-# make er long
-macro_long_data_site <- pivot_longer(
-  data = site_macro,
-  cols = grep("acanthophora_spicifera", colnames(site_macro)):grep("valonia_ventricosa", colnames(site_macro)),
-  names_to = "taxa",
-  values_to = "percent_cover") %>% 
-  mutate(prop_cover = percent_cover/100) %>% 
-  select(-percent_cover)
-
-# write_csv(macro_long_data, here("Data", "macro_long_data.csv"))
-
-# CALCULATE SYNCHRONY CODE
-lm_synchrony_site <- codyn::synchrony(df = macro_long_data_site,
-                                   time.var = "year",
-                                   species.var = "taxa",
-                                   abundance.var = "prop_cover",
-                                   metric = "Loreau",
-                                   replicate.var = "site_habitat")
-
-diversity_stability_site <- alpha_diversity_site_macro_with_cover %>%
-  dplyr::group_by(site_habitat, habitat, site) %>%
-  dplyr::summarise(
-            richness_mean = mean(richness), 
-            shannon_mean = mean(shannon), 
-            cover_mean = mean(prop_cover),
-            evenness_mean = mean(evenness),
-            cover_cv = CV(prop_cover),
-            cover_stability = 1/CV(prop_cover))
-
-diversity_stability_synchrony_site <- left_join(diversity_stability_site, lm_synchrony_site, by = join_by(site_habitat))
-diversity_stability_synchrony_site$synchrony_trans <- sv_trans(diversity_stability_synchrony_site$synchrony)
-# diversity_stability_synchrony_site <- na.omit(diversity_stability_synchrony_site) # remove quads that never had algae --> NaN calcs
-# write_csv(diversity_stability_synchrony_site, here("data", "diversity_stability_synchrony_site.csv"))
-
-dss_ranges_site <- diversity_stability_synchrony_site %>%
-  group_by(habitat) %>%
-  summarise(min_richness = min(richness_mean),
-            max_richness = max(richness_mean),
-            min_evenness = min(evenness_mean, na.rm = T),
-            max_evenness = max(evenness_mean, na.rm = T),
-            min_stability= min(cover_stability),
-            max_stability = max(cover_stability),
-            min_synchrony = min(synchrony_trans),
-            max_synchrony = max(synchrony_trans))
-```
-
-## figure 4d model - richness stability
+## figure 4d
+
+###model - richness stability
 
@@
 
-## plot figure 4d
+### plot figure 4d
 
@@
 
+### model
 ```{r synchrony staability mod site}
@@
 
-## plot figure 4f
+### plot figure 4f
 
@@
 
-# figure 3f
-
-## model
-
-```{r synchrony cover mod}
-synch_cov_mod_site <- glmmTMB(cover_mean ~ synchrony*habitat, family = beta_family(), data = diversity_stability_synchrony_site)
-summary(synch_cov_mod_site)
-Anova(synch_cov_mod_site)
-#                    Chisq Df Pr(>Chisq)  
-# synchrony         2.4184  1    0.11992  
-# habitat           7.8002  3    0.05033 .
-# synchrony:habitat 8.8866  3    0.03084 *
-hist(residuals(synch_cov_mod_site)) # slight skew
-plot(residuals(synch_cov_mod_site) ~ fitted(synch_cov_mod_site)) # looks fine
-emtrends(synch_cov_mod_site, pairwise ~ habitat, var = "synchrony") # forereef 17 different from backreef
-```
-
-```{r synch cov site plot setup}
-synch_cov_emm_site<- emmip(synch_cov_mod_site,
-                        habitat ~ synchrony,
-                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)
-
-filtered_synch_cov_effects_site <- synch_cov_emm_site %>%
-  left_join(dss_ranges_site, by = c("habitat")) %>%
-  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
-```
-
-## plot figure 3f
-
-```{r synchrony cover plot site}
-(figure_3f <- 
-  ggplot() +
-  geom_point(data = diversity_stability_synchrony_site,
-             aes(x = synchrony, y = cover_mean, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_synch_cov_effects_site, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
-  geom_ribbon(data = filtered_synch_cov_effects_site, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
-              alpha = 0.3) +
-  scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
-  labs(x = "Synchrony", y = "Cover\n", title = "3f.") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        plot.title = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-# ggsave(filename = "output/figure_3f.png", figure_3f, height = 10, width = 14)
-```
-
 # evenness stability site
@@
 ```
+
+## Figure 4 Full Plot
+```{r ggpubr fig4}
+fig4_plot_list <- list(figure_4a, figure_4c, figure_4d, figure_4f)
+figure_4 <- ggarrange(plotlist = fig4_plot_list, 
+                      ncol = 2, nrow = 2, 
+                      labels = c("a.", "b."),
+                      common.legend = TRUE,
+                      legend = "bottom", 
+                      font.label = list(size = 20, color = "black", face = "plain"))
+
+#ggsave(filename = "output/figure_4.png", figure_4, height = 10, width = 20)
+```
+
+# SUPPLEMENTAL FIGURE CODE:
+
+## 2c Evenness @ Quad by Year
+### prep
+```{r calculate evennes}
+#dear callie, why is this here?
+site_richness_df <- alpha_diversity_site_macro %>% 
+  group_by(year, habitat) %>% 
+  summarise(richness_mean = mean(richness),
+            richness_se = se(richness))
+
+quad_evenness_df <- alpha_diversity_quad_macro_with_cover %>% 
+  drop_na(evenness) %>%
+  group_by(year, habitat) %>% 
+  summarise(evenness_mean = mean(evenness),
+            evenness_se = se(evenness))
+```
+
+### plot
+```{r}
+(figure_2c <- ggplot() +
+  geom_point(data = quad_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), size = 3) +
+  geom_line(data = quad_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), linewidth = 1) +
+  geom_errorbar(data = quad_evenness_df, aes(x = year, ymin = evenness_mean - evenness_se, ymax = evenness_mean + evenness_se,
+                                             colour = habitat), width = 0, size = 1.3) +
+  scale_colour_manual(values = habitat_colours) +
+  # scale_fill_manual(values = habitat_colours) +
+  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
+  annotate("segment", x = 2006, xend = 2010, y = .7, yend = .7, linewidth = 2, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+   scale_x_continuous(breaks = unique(quad_evenness_df$year)) +
+   labs(y = "Macroalgal evenness \n",
+        x = "Year",
+        title = "c. Quadrat-level evenness") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        axis.title.y = element_text(size=18), 
+        plot.title = element_text(size=18),
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) )
+
+# ggsave(filename = "output/figure_2c.png", figure_2c, height = 10, width = 14)
+```
+
+## 2d Evenness @ Site by Year
+### prep
+```{r set up evenness}
+site_evenness_df <- alpha_diversity_site_macro %>% 
+  drop_na(evenness) %>%
+  group_by(year, habitat) %>% 
+  summarise(evenness_mean = mean(evenness),
+            evenness_se = se(evenness))
+```
+
+### plot
+```{r plot evenness site}
+(figure_2d <- ggplot() +
+  geom_point(data = site_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), size = 3) +
+  geom_line(data = site_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), linewidth = 1) +
+  geom_errorbar(data = site_evenness_df, aes(x = year, ymin = evenness_mean - evenness_se, ymax = evenness_mean + evenness_se, colour = habitat), size = 1.3, width = 0) +
+  scale_colour_manual(values = habitat_colours) +
+  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
+  annotate("segment", x = 2006, xend = 2010, y = 1, yend = 1, linewidth = 2, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+   scale_x_continuous(breaks = unique(site_evenness_df$year)) +
+   labs(y = "Macroalgal evenness \n",
+        x = "Year",
+        title = "d. Site-level evenness") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        axis.title.y = element_text(size=18), 
+        plot.title = element_text(size=18),
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black")) )
+
+# ggsave(filename = "output/figure_2d.png", figure_2d, height = 10, width = 14)
+```

### 2024-09-24 manuscript_1_code.Rmd
added 'mytheme' code to shorten some of the ggplots :tropical_fish:

@@
 library(here)
-library(dplyr)
 library(tidyverse)
@@
 library(emmeans)
+library(dplyr)
 ```
@@
 ```{r figure formatting info}
+mytheme <- theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        axis.title.y = element_text(size=18), 
+        panel.background = element_rect(fill = "white"), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))
+
 habitat_colours <- c("Fringing" = "#D81B60",
@@
   theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.text.x = element_text(angle = 45, hjust = 1),
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black")) )
+  mytheme)
 
@@
 
+## plot after aes changes
 ```{r Figure 1 v2}
@@
     theme_bw() +
-    theme(axis.title.x = element_text(size = 18),
-          axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
-          axis.title.y = element_text(size = 18),
-          axis.text.y = element_text(size = 12),
-          plot.title = element_text(size = 20, hjust = 0),
-          legend.text = element_text(size = 14), 
-          legend.title = element_blank(),
-          legend.position = 'bottom',
-          legend.box = 'horizontal',
-          legend.box.background = element_rect(colour = "black"))
+    mytheme
 })
@@
   theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.text.x = element_text(angle = 45, hjust = 1, size=14),
+  mytheme +
+  theme(axis.text.x = element_text(size=14),
         axis.text.y = element_text(size=12),
-        axis.title.y = element_text(size=18), 
-        plot.title = element_text(size=18),
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black")) )
+        plot.title = element_text(size=18)))
 
@@
         title = "Site-level Richness") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.text.x = element_text(angle = 45, hjust = 1, size=14),
-        axis.title.y = element_text(size=18), 
+  mytheme +
+   theme(axis.text.x = element_text(size=14),
       axis.text.y = element_text(size=14),
       plot.title = element_text(size=18),
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black")) )
+      ))
 

### 2024-09-24 Personal_Prelim_Code/dataexploration.Rmd
I HAVE MOVED YOUR PERSONAL CODE.
I did not delete it, but it is in a new folder. your file paths may not be happy, but the code is still there.

---
title: "data_exploration"
author: "L. Enright"
date: "2024-05-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(dplyr)
library(tidyverse)
#install.packages("fishualize")
library(fishualize)

data <- read.csv(here("data", "allbenthicdata_cleaned.csv"), stringsAsFactors = F) # replacements don't work when strings are factors

str(data)

#pivot long

long_data <- pivot_longer(
  data = data,
  cols = 8:92,
  names_to = "taxa",
  values_to = "percent_cover")

```

#look at sites by location (back, fringing, forereef)
```{r backreef}

transect_level <- long_data %>%
  #dplyr::filter(Habitat == "Backreef") %>%
  dplyr::group_by(Year, Site, Habitat, Transect, taxa) %>%
  summarise(mean_cover = mean(percent_cover), 
            se = sd(percent_cover)/sqrt(n()))

site_level <- transect_level %>%
  dplyr::group_by(Year, Site, Habitat, taxa) %>%
  summarise(mean_site_cover = mean(mean_cover), 
            site_se = sd(mean_cover)/sqrt(n()))

#backreef
site_level %>%
  dplyr::filter(taxa %in% c("Coral", "Algal.Turf", "Crustose.Corallines", "Lobophora.variegata", 
                            "Asparagopsis.taxiformis")) %>%
  dplyr::filter(Habitat == "Backreef") %>%
  #dplyr::filter(Site == "LTER 1") %>%
  ggplot(aes(x = Year, y = mean_site_cover, color = taxa, group = taxa)) +
  geom_point() +
  scale_color_fish_d(option = "Coryphaena_hippurus")+
  geom_line() +
  geom_errorbar(aes(ymax = mean_site_cover + site_se,
                ymin = mean_site_cover - site_se)) +
  facet_wrap(~Site)

site_level %>%
  dplyr::filter(taxa %in% c("Algal.Turf")) %>%
  dplyr::filter(Habitat == "Backreef") %>%
   dplyr::filter(Site == "LTER 1")
```

```{r fringing}

#fringing

site_level %>%
  dplyr::filter(taxa %in% c("Coral", "Algal.Turf", "Crustose.Corallines", "Lobophora.variegata", 
                            "Asparagopsis.taxiformis")) %>%
  dplyr::filter(Habitat == "Fringing") %>%
  #dplyr::filter(Site == "LTER 1") %>%
  ggplot(aes(x = Year, y = mean_site_cover, color = taxa, group = taxa)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymax = mean_site_cover + site_se,
                ymin = mean_site_cover - site_se)) +
  facet_wrap(~Site) +
  theme_classic()
```

```{r forereef 10}

#forereef

site_level %>%
  dplyr::filter(taxa %in% c("Coral", "Algal.Turf", "Crustose.Corallines", "Lobophora.variegata", 
                            "Asparagopsis.taxiformis")) %>%
  dplyr::filter(Habitat == "Outer 10") %>%
  #dplyr::filter(Site == "LTER 1") %>%
  ggplot(aes(x = Year, y = mean_site_cover, color = taxa, group = taxa)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymax = mean_site_cover + site_se,
                ymin = mean_site_cover - site_se)) +
  facet_wrap(~Site)

```

```{r forereef 17}

#forereef

site_level %>%
  dplyr::filter(taxa %in% c("Coral", "Algal.Turf", "Crustose.Corallines", "Lobophora.variegata", 
                            "Asparagopsis.taxiformis", "Bare.Space")) %>%
  dplyr::filter(Habitat == "Outer 17") %>%
  #dplyr::filter(Site == "LTER 1") %>%
  ggplot(aes(x = Year, y = mean_site_cover, color = taxa, group = taxa)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymax = mean_site_cover + site_se,
                ymin = mean_site_cover - site_se)) +
  facet_wrap(~Site)

```

### 2024-09-24 Personal_Prelim_Code/julianna_checkLTER1fringe.Rmd
I HAVE MOVED YOUR PERSONAL CODE.
I did not delete it, but it is in a new folder. your file paths may not be happy, but the code is still there.

---
title: "julianna_check2017fringeLTER1"
output: html_document
date: "2024-09-10"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clean up
rm(list=ls())

# load packages
require(here) # relative file paths
require(tidyverse) # data wrangling
require(vegan) # community/diversity analyses
```

This script summarizes the coral and macroalgae subsets to be joinable with the full algae dataset

# Bring in the data

```{r}
# algal cover 
algae_raw <- read_csv(here("data/MCR benthic algae data cover RAW.csv"))

```

```{r}
algae_raw %>% 
  filter(Taxonomy_Substrate_Functional_Group %in% c("Ascidian", "Sponge", "Millepora platyphylla", "Shell Debris", "Soft Coral", "Corallimorpharia", "Cyanophyta", "Coral Rubble", "No data", "Heteractis sp.", "Tridacna sp.", "Bare Space", "Sarcophyton sp.", "Zooxanthellae")) %>% 
  group_by(Year, Site, Location, Taxonomy_Substrate_Functional_Group) %>% 
  summarize(Total_other_cover = sum(Percent_Cover)) %>% 
  mutate(Habitat = case_when(str_detect(string = Location, pattern = "Fring") ~ "Fringing",
                              TRUE ~ "nawh")) %>% 
  filter(Year == 2017,
         Habitat == "Fringing",
         Site == "LTER 1"
         ) %>% view()

```

### 2024-09-24 Personal_Prelim_Code/julianna_figs_Aug29.Rmd
I HAVE MOVED YOUR PERSONAL CODE.
I did not delete it, but it is in a new folder. your file paths may not be happy, but the code is still there.

--- 2024-09-10 julianna_figs_Aug29.Rmd
@@
          macroalgae_cover, macroalgae_cover_hard, coral_cover, coral_cover_hard, sand_cover, 
-         turf_cover, turf_cover_hard, CCA_cover, CCA_cover_hard
+         turf_cover, turf_cover_hard, CCA_cover, CCA_cover_hard, sand_cover
          ) %>% 
@@
   select(year, location, site, habitat, transect, quadrat, island_side, 
-         macroalgae_cover, turf_cover, CCA_cover, coral_cover) %>% 
+         macroalgae_cover, turf_cover, CCA_cover, coral_cover, sand_cover) %>% 
   # pivot key columns longer
-  pivot_longer(cols = c(macroalgae_cover, turf_cover, CCA_cover, coral_cover),
+  pivot_longer(cols = c(macroalgae_cover, turf_cover, CCA_cover, coral_cover, sand_cover),
                names_to = "Substrate",
@@
   theme_bw() +
-   scale_fill_manual(values = c(
-    moma.colors("Klein", type = "discrete")[2], # CCA
-    moma.colors("Klein", type = "discrete")[1], # coral
-    moma.colors("Klein", type = "discrete")[5], # macro
-    moma.colors("Klein", type = "discrete")[4] # turf
-  ),
-  labels = c("CCA", "Coral", "Macroalgae", "Turf algae")) +
+  # scale_fill_manual(values = c(
+  #  moma.colors("Klein", type = "discrete")[2], # CCA
+  #  moma.colors("Klein", type = "discrete")[1], # coral
+  #  moma.colors("Klein", type = "discrete")[5], # macro
+  #  moma.colors("Klein", type = "discrete")[4] # turf
+  #),
+  #labels = c("CCA", "Coral", "Macroalgae", "Turf algae")) +
   facet_grid(site ~ habitat) + # not sure which direction is best here

### 2024-09-24 Personal_Prelim_Code/mcr_algal_data_cleaning_final.Rmd
I HAVE MOVED YOUR PERSONAL CODE.
I did not delete it, but it is in a new folder. your file paths may not be happy, but the code is still there.

--- 2024-07-03 mcr_algal_data_cleaning_final.Rmd
@@
   wide_alldata_filtered %>%
-  janitor::clean_names()
-
-# Julianna is adding this code here IF we want it:
-wide_alldata_filtered_cleaned_column_names %>% 
+  janitor::clean_names() %>% 
+  # Julianna is adding this code here IF we want it to clean the inside:
   mutate(location = gsub(" ", "_", str_to_lower(location)),
@@
          site_habitat = gsub(" ", "_", str_to_lower(site_habitat)),
-         ) # would want to save this then do it for the algae dataset as well
+         )
 
-write.csv(wide_alldata_filtered_cleaned_column_names, "allbenthicdata_cleaned_colnames.csv", row.names = FALSE)
+write.csv(wide_alldata_filtered_cleaned_column_names, "data/allbenthicdata_cleaned_colnames.csv", row.names = FALSE)
 
@@
 
-wide_algaeonly_filtered <- wide_alldata_filtered[ , -which(names(wide_alldata_filtered) 
+wide_algaeonly_filtered <- wide_alldatsa_filtered[ , -which(names(wide_alldata_filtered) 
                                                            %in% not_algae)]
@@
   wide_algaeonly_filtered %>%
-  janitor::clean_names()
+  # clean column names
+  janitor::clean_names() %>% 
+  # clean the inside as well
+  mutate(location = gsub(" ", "_", str_to_lower(location)),
+         site = gsub(" ", "_", str_to_lower(site)),
+         habitat = gsub(" ", "_", str_to_lower(habitat)),
+         site_habitat = gsub(" ", "_", str_to_lower(site_habitat)),
+         )
 
-write.csv(wide_algaeonly_cleaned_column_names, "algaeonly_cleaned_colnames.csv", row.names = FALSE)
+write.csv(wide_algaeonly_cleaned_column_names, "data/algaeonly_cleaned_colnames.csv", row.names = FALSE)
 

### 2024-09-30 manuscript_1_code.Rmd
Moved functions to their own script

@@
 
-## functions
-
-```{r functions}
-# coefficient of variation
-CV<-function(x){
-  return(sd(x,na.rm=T)/mean(x,na.rm=T))
-}
-
-inv_logit <- function(x){
-  exp(x)/(1+exp(x))
-}
-
-# smithson verkuilen transformation for 0-1 continuous data
-sv_trans <- function(prop, s = 0.000005){
-  (prop*(length(prop) - 1) + s)/length(prop)
-  # where prop is the vector of the proportional value you're transforming, 
-  # N is the sample size, which is specified by taking the number of rows/observations from a given dataframe,
-  # and s is a small offset 
-}
-
-se <- function(x){
-  sd(x)/sqrt(n())
-}
-
+## Functions
+```{r}
+source("R/functions.R")
 ```

### 2024-09-30 R/functions.R

@@ -1,0 +1,24 @@
+#### 
+# Functions for Manuscript 1 #
+####
+
+# coefficient of variation
+CV<-function(x){
+  return(sd(x,na.rm=T)/mean(x,na.rm=T))
+}
+
+inv_logit <- function(x){
+  exp(x)/(1+exp(x))
+}
+
+# smithson verkuilen transformation for 0-1 continuous data
+sv_trans <- function(prop, s = 0.000005){
+  (prop*(length(prop) - 1) + s)/length(prop)
+  # where prop is the vector of the proportional value you're transforming, 
+  # N is the sample size, which is specified by taking the number of rows/observations from a given dataframe,
+  # and s is a small offset 
+}
+
+se <- function(x){
+  sd(x)/sqrt(n())
+}

### 2024-10-01 manuscript_1_code.Rmd
updated figure 2

@@
 ### plot
-```{r figure 2}
+```{r figure 2a}
 (figure_2a <- ggplot() +
@@
         x = "Year",
-        title = "Quadrat-level Richness") +
+        title = "a. Quadrat-level Richness") +
   theme_bw() + 
   mytheme +
-  theme(axis.text.x = element_text(size=14),
-        axis.text.y = element_text(size=12),
-        plot.title = element_text(size=18)))
+   theme(axis.text.x = element_text(size=20),
+         axis.title.x = element_text(size=22),
+         axis.title.y = element_text(size=22),
+      axis.text.y = element_text(size=20),
+      plot.title = element_text(size=24),
+      legend.text = element_text(size=20),
+      legend.title = element_text(size=22)))
 
@@
 
-```{r richness df}
+```{r site level richness df}
 site_richness_df <- alpha_diversity_site_macro %>% 
@@
 ### plot
-```{r}
+```{r plot figure 2b}
 (figure_2b <- ggplot() +
   geom_point(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4) +
-  geom_jitter(data = alpha_diversity_site_macro, aes(x = year, y = richness, colour = habitat, alpha = 0.05), size = 2, width = 0.4, height = 0) +
+#  geom_jitter(data = alpha_diversity_site_macro, 
+#  aes(x = year, y = richness, colour = habitat, alpha = 0.05), size = 2, width = 0.4, height = 0) +
   geom_line(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
@@
    scale_x_continuous(breaks = unique(site_richness_df$year)) +
-   labs(y = "Macroalgal Richness",
+   labs(y = "",
         x = "Year",
-        title = "Site-level Richness") +
+        title = "b. Site-level Richness") +
   mytheme +
-   theme(axis.text.x = element_text(size=14),
-      axis.text.y = element_text(size=14),
-      plot.title = element_text(size=18),
+   theme(axis.text.x = element_text(size=20),
+         axis.title.x = element_text(size=22),
+      axis.text.y = element_text(size=20),
+      plot.title = element_text(size=24),
+      legend.text = element_text(size=20),
+      legend.title = element_text(size=22),
       ))
@@
 ## Figure 2 Full Plot
-```{r ggpubr fig2}
+```{r ggpubr fig2, fig.height = 10, fig.width=20}
 fig2_plot_list <- list(figure_2a, figure_2b)
-figure_2 <- ggarrange(plotlist = fig2_plot_list, 
+(figure_2 <- ggarrange(plotlist = fig2_plot_list, 
                       ncol = 2, nrow = 1, 
-                      labels = c("a.", "b."),
                       common.legend = TRUE,
                       legend = "bottom", 
-                      font.label = list(size = 20, color = "black", face = "plain"))
-
-#ggsave(filename = "output/figure_2.png", figure_2, height = 10, width = 20)
+                      font.label = list(size = 20, color = "black", face = "plain")))
+
+# ggsave(filename = "output/figure_2v2.png", figure_2, height = 10, width = 20)
 ```

### 2024-10-01 manuscript_1_code.Rmd
updated plot aesthetics for figures 3 and 4. Pending new functional data.

@@
       legend.text = element_text(size=20),
-      legend.title = element_text(size=22)))
+      legend.title = element_blank()))
 
@@
   #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
-  annotate("segment", x = 2006, xend = 2010, y = 14, yend = 14, linewidth = 2, colour = "grey") +
+  annotate("segment", x = 2006, xend = 2010, y = 12, yend = 12, linewidth = 2, colour = "grey") +
   geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
@@
       legend.text = element_text(size=20),
-      legend.title = element_text(size=22),
+      legend.title = element_blank(),
       ))
@@
 ## Figure 2 Full Plot
-```{r ggpubr fig2, fig.height = 10, fig.width=20}
+```{r ggpubr fig2, fig.height = 10, fig.width=22}
 fig2_plot_list <- list(figure_2a, figure_2b)
@@
 
-# ggsave(filename = "output/figure_2v2.png", figure_2, height = 10, width = 20)
+# ggsave(filename = "output/figure_2v2.png", figure_2, height = 10, width = 22)
 ```
@@
 emtrends(cover_mod, pairwise ~ habitat, var = "richness") # backreef not different from forereef, forereef also not different
-```
-
-```{r 3b evenness quad level model}
-alpha_diversity_quad_macro_with_cover_evenness_mod <- alpha_diversity_quad_macro_with_cover %>% 
-  drop_na(evenness)
-
-evenness_cover_mod_quad <- glmmTMB(cover_trans ~ evenness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro_with_cover_evenness_mod)
-summary(evenness_cover_mod_quad)
-Anova(evenness_cover_mod_quad)
-#                     Chisq Df Pr(>Chisq)    
-# richness         30916.73  1  < 2.2e-16 ***
-# habitat            141.78  3  < 2.2e-16 ***
-# richness:habitat   294.54  3  < 2.2e-16 ***
-hist(residuals(evenness_cover_mod_quad)) # slight pos skeq
-plot(residuals(evenness_cover_mod_quad) ~ fitted(evenness_cover_mod_quad)) # negative trend but it can't be less than 0 so not too concerned. some wave pattern
-r.squaredGLMM(evenness_cover_mod_quad) # 0.1749907, 0.4042826
-emtrends(evenness_cover_mod_quad, pairwise ~ habitat, var = "evenness") #backreef is different from everything except outer 17, but just barely. everyone else is same same xoxo
-
 ```
@@
 ### site evenness 
-```{r}
+```{r site level evenness}
 alpha_diversity_site_macro_with_cover_evenness_mod <- alpha_diversity_site_macro_with_cover %>% 
@@
   geom_ribbon(data = filtered_cover_effects, aes(x = richness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
-              alpha = 0.3) +
+              alpha = 0.5) +
   scale_colour_manual(values = habitat_colours) +
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Cover\n", x = "Macroalgal Richness", title = "3a.") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
+  labs(y = "Cover\n", x = "", title = "3a.") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.title.y = element_text(size=24), 
+        axis.text.x = element_text(size=22), 
+        axis.text.y = element_text(size=22),
+        legend.text = element_text(size=22),
+        plot.title = element_text(size=26), 
         panel.background = element_blank(), 
@@
         plot.background = element_rect(fill = "white"),
-        legend.text = element_text(size = rel(1)), # increase legend text
         legend.title = element_blank(),
@@
 # ggsave(filename = "output/figure_3a.png", figure_3a, height = 10, width = 14)
-```
-
-## Figure 3B quad level evenness versus cover
-```{r figure 3b setup}
-diversity_ranges_quad_evenness <- alpha_diversity_quad_macro_with_cover_evenness_mod %>%
-  group_by(habitat) %>%
-  summarise(min_evenness = min(evenness),
-            max_evenness = max(evenness),
-            min_cover= min(cover_trans),
-            max_cover = max(cover_trans))
-
-cover_emm_evenness <- emmip(evenness_cover_mod_quad,
-                        habitat ~ evenness,
-                        at = list(evenness = seq(0,10,1)), plotit = F, CIs = T)
-
-filtered_cover_effects_evenness <- cover_emm_evenness %>%
-  left_join(diversity_ranges_quad_evenness, by = c("habitat")) %>%
-  filter(evenness >= min_evenness & evenness <= max_evenness)
-
-# alpha_diversity_quad_macro_with_cover_evenness_mod$habitat <- factor(alpha_diversity_quad_macro_with_cover_evenness_mod$habitat, 
-#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
-#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
-# 
-# 
-# filtered_cover_effects_evenness$habitat <- factor(filtered_cover_effects_evenness$habitat, 
-#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
-#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
-```
-
-```{r}
-(figure_3b <- 
-  ggplot() +
-  geom_point(data = alpha_diversity_quad_macro_with_cover_evenness_mod,
-             aes(x = evenness, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
-  geom_line(data = filtered_cover_effects_evenness, aes(x = evenness, y = inv_logit(yvar), colour = habitat)) +
-  geom_ribbon(data = filtered_cover_effects_evenness, aes(x = evenness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
-              alpha = 0.3) +
-  scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
-  labs(y = "Cover\n", x = "Macroalgal evenness", title = "3b.") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-#ggsave(filename = "output/figure_3b.png", figure_3b, height = 10, width = 14)
 ```
@@
   geom_point(data = diversity_stability_synchrony,
-             aes(x = synchrony, y = cover_mean, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_synch_cov_effects, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
+             aes(x = synchrony, y = cover_mean, colour = habitat), size = 1, alpha = 0.3) + 
+  geom_line(data = filtered_synch_cov_effects, aes(x = synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
   geom_ribbon(data = filtered_synch_cov_effects, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(x = "Synchrony", y = "Cover\n", title = "3c.") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18),
-        plot.title = element_text(size=18),
+  labs(x = "", y = "", title = "3c.") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.title.y = element_text(size=24), 
+        axis.text.x = element_text(size=22), 
+        axis.text.y = element_text(size=22),
+        legend.text = element_text(size=22),
+        plot.title = element_text(size=26), 
         panel.background = element_blank(), 
@@
 
-## Figure 3E site level richness versus cover
+## Figure 3d site level richness versus cover
 
 ### setup
-```{r}
+```{r site richness cover setup}
 diversity_ranges_site <- alpha_diversity_site_macro_with_cover %>%
@@
 
-```{r}
+```{r plot figure 3d}
 (figure_3d <- 
@@
              aes(x = richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.3) + 
-  geom_line(data = filtered_site_cover_effects, aes(x = richness, y = inv_logit(yvar), colour = habitat)) +
+  geom_line(data = filtered_site_cover_effects, aes(x = richness, y = inv_logit(yvar), colour = habitat), size = 1.5) +
   geom_ribbon(data = filtered_site_cover_effects, aes(x = richness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
@@
   theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        plot.title = element_text(size=18), 
+  theme(axis.title.x = element_text(size=24), 
+        axis.title.y = element_text(size=24), 
+        axis.text.x = element_text(size=22), 
+        axis.text.y = element_text(size=22),
+        legend.text = element_text(size=22),
+        plot.title = element_text(size=26), 
         panel.background = element_blank(), 
@@
 
-## figure 3f
+## figure 3f - synchrony site level
 
@@
              aes(x = synchrony, y = cover_mean, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_synch_cov_effects_site, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
+  geom_line(data = filtered_synch_cov_effects_site, aes(x = synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
   geom_ribbon(data = filtered_synch_cov_effects_site, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(x = "Synchrony", y = "Cover\n", title = "3f.") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        plot.title = element_text(size=18), 
+  labs(x = "Synchrony", y = "", title = "3f.") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.title.y = element_text(size=24), 
+        axis.text.x = element_text(size=22), 
+        axis.text.y = element_text(size=22),
+        legend.text = element_text(size=22),
+        plot.title = element_text(size=26), 
         panel.background = element_blank(), 
@@
         plot.background = element_rect(fill = "white"),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-# ggsave(filename = "output/figure_3f.png", figure_3f, height = 10, width = 14)
-```
-
-## Figure 3f site level evenness versus cover
-
-### model
-
-```{r evenness cover model site level}
-evenness_cover_mod_site <- glmmTMB(cover_trans ~ evenness*habitat + (1|site) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover_evenness_mod)
-summary(evenness_cover_mod_site)
-hist(residuals(evenness_cover_mod_site)) # little wonky but not the worst
-plot(residuals(evenness_cover_mod_site) ~ fitted(evenness_cover_mod_site))
-Anova(evenness_cover_mod_site)
-#                   Chisq Df Pr(>Chisq)    
-# evenness         14.395  1  0.0001482 ***
-# habitat          45.636  3  6.777e-10 ***
-# evenness:habitat 16.126  3  0.0010687 ** 
-r.squaredGLMM(evenness_cover_mod_site) # 0.166, 0.258
-
-emtrends(evenness_cover_mod_site, pairwise ~ habitat, var = "evenness") # fringing different from backreef and outer10
-```
-
-### plot
-
-### setup
-
-```{r figure 3f evenness site plot setup}
-diversity_ranges_site_evenness <- alpha_diversity_site_macro_with_cover_evenness_mod %>%
-  group_by(habitat) %>%
-  summarise(min_evenness = min(evenness),
-            max_evenness = max(evenness),
-            min_cover= min(cover_trans),
-            max_cover = max(cover_trans))
-
-cover_emm_evenness_site <- emmip(evenness_cover_mod_site,
-                        habitat ~ evenness,
-                        at = list(evenness = seq(0,1,0.01)), plotit = F, CIs = T)
-
-filtered_cover_effects_evenness_site <- cover_emm_evenness_site %>%
-  left_join(diversity_ranges_site_evenness, by = c("habitat")) %>%
-  filter(evenness >= min_evenness & evenness <= max_evenness)
-
-# alpha_diversity_site_macro_with_cover_evenness_mod$habitat <- factor(alpha_diversity_site_macro_with_cover_evenness_mod$habitat, 
-#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
-#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
-# 
-# 
-# 
-```
-
-### plot
-
-```{r evenness cover plot site }
-(figure_3f <- 
-  ggplot() +
-  geom_point(data = alpha_diversity_site_macro_with_cover_evenness_mod,
-             aes(x = evenness, y = cover_trans, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_cover_effects_evenness_site, aes(x = evenness, y = inv_logit(yvar), colour = habitat)) +
-  geom_ribbon(data = filtered_cover_effects_evenness_site, aes(x = evenness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
-              alpha = 0.3) +
-  scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
-  labs(y = "Cover\n", x = "Macroalgal evenness", title = "3f. Site Level Evenness") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.text = element_text(size = rel(1)), # increase legend text
         legend.title = element_blank(),
@@
              aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_rich_stab_effects, aes(x = richness_mean, y = exp(yvar), colour = habitat)) +
+  geom_line(data = filtered_rich_stab_effects, aes(x = richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
   geom_ribbon(data = filtered_rich_stab_effects, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Stability\n", x = "Taxonomic Richness", title = "4a.") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18),
-        plot.title = element_text(size=18),
+  labs(y = "Stability\n", x = "", title = "4a.") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.title.y = element_text(size=24), 
+        axis.text.x = element_text(size=22), 
+        axis.text.y = element_text(size=22),
+        legend.text = element_text(size=22),
+        plot.title = element_text(size=26), 
         panel.background = element_blank(), 
@@
              aes(x = synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_synch_stab_effects, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
+  geom_line(data = filtered_synch_stab_effects, aes(x = synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
   geom_ribbon(data = filtered_synch_stab_effects, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(x = "Synchrony", y = "Stability\n", title = "4c.") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18),
-        plot.title = element_text(size=18),
+  labs(x = "", y = "", title = "4c.") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.title.y = element_text(size=24), 
+        axis.text.x = element_text(size=22), 
+        axis.text.y = element_text(size=22),
+        legend.text = element_text(size=22),
+        plot.title = element_text(size=26), 
         panel.background = element_blank(), 
@@
              aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_rich_stab_effects_site, aes(x = richness_mean, y = exp(yvar), colour = habitat)) +
+  geom_line(data = filtered_rich_stab_effects_site, aes(x = richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
   geom_ribbon(data = filtered_rich_stab_effects_site, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Stability\n", x = "Macroalgal Richness", title = "4d.") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        plot.title = element_text(size=18), 
+  labs(y = "Stability\n", x = "Taxonomic Richness", title = "4d.") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.title.y = element_text(size=24), 
+        axis.text.x = element_text(size=22), 
+        axis.text.y = element_text(size=22),
+        legend.text = element_text(size=22),
+        plot.title = element_text(size=26), 
         panel.background = element_blank(), 
@@
              aes(x = synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_synch_stab_effects_site, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
+  geom_line(data = filtered_synch_stab_effects_site, aes(x = synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
   geom_ribbon(data = filtered_synch_stab_effects_site, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(x = "Synchrony", y = "Stability\n", main = "4f.") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18),
-        plot.title = element_text(size=18),
+  labs(x = "Synchrony", y = "", main = "4f.") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.title.y = element_text(size=24), 
+        axis.text.x = element_text(size=22), 
+        axis.text.y = element_text(size=22),
+        legend.text = element_text(size=22),
+        plot.title = element_text(size=26), 
         panel.background = element_blank(), 
@@
 # ggsave(filename = "output/figure_4f.png", figure_4f, height = 10, width = 14)
-```
-
-# evenness stability site
-
-## mod
-
-```{r evenness synchrony mod}
-even_stab_mod_site <- glmmTMB(cover_stability~ evenness_mean*habitat, family = Gamma(link = "log"), data = diversity_stability_synchrony_site)
-summary(even_stab_mod_site)
-Anova(even_stab_mod_site)
-
-#                         Chisq Df Pr(>Chisq)   
-# evenness_mean          0.1035  1   0.747623   
-# habitat               11.8743  3   0.007827 **
-# evenness_mean:habitat  9.6254  3   0.022034 *   
-hist(residuals(even_stab_mod_site)) # blocky with some pos skew
-plot(residuals(even_stab_mod_site) ~ fitted(even_stab_mod_site)) # heteroskedastic 
-r.squaredGLMM(even_stab_mod_site) # 0.505
-emtrends(even_stab_mod_site, pairwise ~ habitat, var = "evenness_mean") # no differences across habitats
-```
-
-```{r figure 4d setup}
-even_stab_emm_site <- emmip(even_stab_mod_site,
-                        habitat ~ evenness_mean,
-                        at = list(evenness_mean = seq(0,1,0.01)), plotit = F, CIs = T)
-#make levels for habitat for plotting
-
-filtered_even_stab_effects_site <- even_stab_emm_site %>%
-  left_join(dss_ranges_site, by = c("habitat")) %>%
-  filter(evenness_mean >= min_evenness & evenness_mean <= max_evenness)
-
-#make levels for habitat for plotting
-# diversity_stability_synchrony_site$habitat <- factor(diversity_stability_synchrony_site$habitat, 
-#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
-#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
-```
-
-## plot figure 4d
-
-```{r figure 4d}
-(figure_4_evenness <- 
-  ggplot() +
-  geom_point(data = diversity_stability_synchrony_site,
-             aes(x = evenness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_even_stab_effects_site, aes(x = evenness_mean, y = exp(yvar), colour = habitat)) +
-  geom_ribbon(data = filtered_even_stab_effects_site, aes(x = evenness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
-              alpha = 0.3) +
-  scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
-  labs(y = "Stability\n", x = "Macroalgal evenness", title = "4f. Evenness site") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        plot.title = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-ggsave(filename = "output/figure_4_evenness_site.png", figure_4_evenness, height = 10, width = 14)
 ```
@@
 ```
+
+## evenness and cover relationships
+
+```{r evenness cover quad level model}
+alpha_diversity_quad_macro_with_cover_evenness_mod <- alpha_diversity_quad_macro_with_cover %>% 
+  drop_na(evenness)
+
+evenness_cover_mod_quad <- glmmTMB(cover_trans ~ evenness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro_with_cover_evenness_mod)
+summary(evenness_cover_mod_quad)
+Anova(evenness_cover_mod_quad)
+#                     Chisq Df Pr(>Chisq)    
+# richness         30916.73  1  < 2.2e-16 ***
+# habitat            141.78  3  < 2.2e-16 ***
+# richness:habitat   294.54  3  < 2.2e-16 ***
+hist(residuals(evenness_cover_mod_quad)) # slight pos skeq
+plot(residuals(evenness_cover_mod_quad) ~ fitted(evenness_cover_mod_quad)) # negative trend but it can't be less than 0 so not too concerned. some wave pattern
+r.squaredGLMM(evenness_cover_mod_quad) # 0.1749907, 0.4042826
+emtrends(evenness_cover_mod_quad, pairwise ~ habitat, var = "evenness") #backreef is different from everything except outer 17, but just barely. everyone else is same same xoxo
+
+```
+
+```{r evenness cover plot setup}
+diversity_ranges_quad_evenness <- alpha_diversity_quad_macro_with_cover_evenness_mod %>%
+  group_by(habitat) %>%
+  summarise(min_evenness = min(evenness),
+            max_evenness = max(evenness),
+            min_cover= min(cover_trans),
+            max_cover = max(cover_trans))
+
+cover_emm_evenness <- emmip(evenness_cover_mod_quad,
+                        habitat ~ evenness,
+                        at = list(evenness = seq(0,10,1)), plotit = F, CIs = T)
+
+filtered_cover_effects_evenness <- cover_emm_evenness %>%
+  left_join(diversity_ranges_quad_evenness, by = c("habitat")) %>%
+  filter(evenness >= min_evenness & evenness <= max_evenness)
+
+# alpha_diversity_quad_macro_with_cover_evenness_mod$habitat <- factor(alpha_diversity_quad_macro_with_cover_evenness_mod$habitat, 
+#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
+#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+# 
+# 
+# filtered_cover_effects_evenness$habitat <- factor(filtered_cover_effects_evenness$habitat, 
+#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
+#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+```
+
+```{r evenness cover plot}
+(figure_3b <- 
+  ggplot() +
+  geom_point(data = alpha_diversity_quad_macro_with_cover_evenness_mod,
+             aes(x = evenness, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
+  geom_line(data = filtered_cover_effects_evenness, aes(x = evenness, y = inv_logit(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_cover_effects_evenness, aes(x = evenness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Cover\n", x = "Macroalgal evenness", title = "3b.") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+#ggsave(filename = "output/figure_3b.png", figure_3b, height = 10, width = 14)
+```
+
+## site level evenness versus cover
+
+### model
+
+```{r evenness cover model site level}
+evenness_cover_mod_site <- glmmTMB(cover_trans ~ evenness*habitat + (1|site) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover_evenness_mod)
+summary(evenness_cover_mod_site)
+hist(residuals(evenness_cover_mod_site)) # little wonky but not the worst
+plot(residuals(evenness_cover_mod_site) ~ fitted(evenness_cover_mod_site))
+Anova(evenness_cover_mod_site)
+#                   Chisq Df Pr(>Chisq)    
+# evenness         14.395  1  0.0001482 ***
+# habitat          45.636  3  6.777e-10 ***
+# evenness:habitat 16.126  3  0.0010687 ** 
+r.squaredGLMM(evenness_cover_mod_site) # 0.166, 0.258
+
+emtrends(evenness_cover_mod_site, pairwise ~ habitat, var = "evenness") # fringing different from backreef and outer10
+```
+
+### plot setup
+
+```{r figure 3f evenness site plot setup}
+diversity_ranges_site_evenness <- alpha_diversity_site_macro_with_cover_evenness_mod %>%
+  group_by(habitat) %>%
+  summarise(min_evenness = min(evenness),
+            max_evenness = max(evenness),
+            min_cover= min(cover_trans),
+            max_cover = max(cover_trans))
+
+cover_emm_evenness_site <- emmip(evenness_cover_mod_site,
+                        habitat ~ evenness,
+                        at = list(evenness = seq(0,1,0.01)), plotit = F, CIs = T)
+
+filtered_cover_effects_evenness_site <- cover_emm_evenness_site %>%
+  left_join(diversity_ranges_site_evenness, by = c("habitat")) %>%
+  filter(evenness >= min_evenness & evenness <= max_evenness)
+
+# alpha_diversity_site_macro_with_cover_evenness_mod$habitat <- factor(alpha_diversity_site_macro_with_cover_evenness_mod$habitat, 
+#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
+#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+# 
+# 
+# 
+```
+
+### plot
+
+```{r evenness cover plot site }
+(figure_3f <- 
+  ggplot() +
+  geom_point(data = alpha_diversity_site_macro_with_cover_evenness_mod,
+             aes(x = evenness, y = cover_trans, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_cover_effects_evenness_site, aes(x = evenness, y = inv_logit(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_cover_effects_evenness_site, aes(x = evenness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Cover\n", x = "Macroalgal evenness", title = "3f. Site Level Evenness") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=18), 
+        axis.title.y = element_text(size=18), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+# ggsave(filename = "output/figure_3f.png", figure_3f, height = 10, width = 14)
+```

### 2024-10-01 manuscript_1_code.Rmd
jittered points for figure 2

@@
 (figure_2a <- ggplot() +
-  geom_point(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4) +
+  geom_point(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4,
+             position = position_dodge(0.4)) +
   geom_line(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
   geom_errorbar(data = quad_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se,
-                                             colour = habitat), width = 0, size = 1.3) +
+                                             colour = habitat), width = 0, size = 1.3, position = position_dodge(0.4)) +
   scale_colour_manual(values = habitat_colours) +
@@
 (figure_2b <- ggplot() +
-  geom_point(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4) +
+  geom_point(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4,
+             position = position_dodge(0.4)) +
 #  geom_jitter(data = alpha_diversity_site_macro, 
@@
   geom_line(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
-  geom_errorbar(data = site_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, colour = habitat), size = 1.3, width = 0) +
+  geom_errorbar(data = site_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, colour = habitat), size = 1.3, width = 0,  position = position_dodge(0.4)) +
   scale_colour_manual(values = habitat_colours) +

### 2024-10-02 functional_groups_v3.Rmd
Add files via upload

---
title: "functional_groups"
author: "L. Enright"
date: "2024-09-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(janitor)
library(stringr)
library(vegan)
library(car)
library(glmmTMB)
library(MuMIn)
library(emmeans)
library(performance)
library(lme4)
library(DHARMa)
library(ggplot2)
```

```{r}
#read in groupings
#these are the OG groupings 
algae_fg_raw <- read_csv(here("data","algae_functional_groupings.csv"))

#these are our best guesses based on Bob's feedback (subject to change)
algae_fg_raw2 <- read_csv(here("data","algae_functional_groupings_092724.csv"))

#clean column names
algae_fg <- algae_fg_raw2 %>%
  clean_names()

#change species names to be lowercase & have underscore 
algae_fg$species <- gsub(" ", "_", tolower(algae_fg$species))
algae_fg$morphology <- trimws(algae_fg$morphology)
algae_fg$morphology <- gsub(" ", "_", algae_fg$morphology)

#rename species column to taxa

algae_fg <- algae_fg %>%
  select(species, morphology) %>%
  dplyr::rename(taxa = species) %>%
   dplyr::rename(functional_group = morphology)

unique(algae_fg$functional_group)

#9 fg in v1 (8 excluding turf)
#12 fg in v2 (10 excluding turf), added spherical and branched calcareous

```

```{r species distribution across functional groups}
algae_fg_raw %>%
  group_by(Morphology) %>%
  summarise(count = n())

algae_fg_raw2 %>%
  group_by(Morphology) %>%
  summarise(count = n())
```

```{r functions}
#this is needed for figures

inv_logit <- function(x){
  exp(x)/(1+exp(x))
}
```

```{r read in macroalgal data}
#read in long macro data 
macro_long <- read_csv(here("data", "macro_long_data.csv"))

#merge functional groups with 
macro_functional_groups <- merge(macro_long, algae_fg, all.x = TRUE)

unique(macro_functional_groups$taxa)

#macro_functional_groups$functional_group[NA]

#look for mistakes in merging 
macro_functional_groups %>% filter(is.na(functional_group))
#none, should be fine!

```

```{r summarizing long macroalgae data}

taxa_sum_check <- macro_functional_groups %>%
  group_by(location, year) %>%
  summarise(sum_macrocover = sum(prop_cover))

macro_functional_groups %>%
  filter(location == "lter_2_backreef_algae_transect_1_quad_1") %>%
  filter(year == "2022")

#summarizing functional groups at the quad level
fg_summary <- macro_functional_groups %>%
  group_by(location, year, functional_group, habitat, site) %>%
  summarise(sum_fg_cover = sum(prop_cover))

max(fg_summary$sum_fg_cover)

#summarizing functional groups at the site level
#NOTE that this is ADDED and not averaged. This data is only used to find richness at the site level. LE did this as adding and as a sum, both end up the same from a richness perspective. 
fg_summary_site_SUM <- macro_functional_groups %>%
  group_by(site, year, habitat, functional_group) %>%
  summarise(sum_fg_cover = sum(prop_cover))

#don't need to use this, was just to check both methods
#fg_summary_site_AVG <- macro_functional_groups %>%
  #group_by(site, year, habitat, functional_group) %>%
  #summarise(mean_fg_cover = mean(prop_cover))

#fg_summary %>%
  #filter(sum_fg_cover==1)

#QAQC
fg_summary_qaqc <- macro_functional_groups %>%
  group_by(location, year, functional_group) %>%
  summarise(sum_fg_cover = sum(prop_cover))

fg_qaqc <- fg_summary_qaqc %>%
  group_by(location, year) %>%
  summarise(sum_fg_check = sum(sum_fg_cover))

check_them <- merge(taxa_sum_check, fg_qaqc)

check_them %>%
  filter(all.equal(sum_macrocover, sum_fg_check) != TRUE)

#should be 0 :)

```

```{r info for themes and colors for figures}
mytheme <- theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        panel.background = element_rect(fill = "white"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))

habitat_colours <- c("Fringing" = "#D81B60",
                     "Backreef" = "#FFC107",
                     "Forereef 10m" = "#1E88E5",
                     "Forereef 17m" = "#004D40")
```

```{r functional group data, pivot wide for quad}
#pivot wide at quad level
fg_summary_wide <- fg_summary %>%
  pivot_wider(names_from = "functional_group", values_from = "sum_fg_cover")

#richness at every quad/year combination 
fg_summary_wide$fg_richness_quad <- vegan::specnumber(fg_summary_wide[5:14])

unique(fg_summary_wide$fg_richness_quad)
#0 to 4

#pull out the habitat and site, still at quad level 
fg_summary_wide_meta <- fg_summary_wide %>%
  mutate(habitat = str_extract(location, 
                                 "(fringing|outer_10|outer_17|backreef)")) %>%
  mutate(site = str_extract(location, "lter_[123456]"))
```

```{r functional group data, pivot wide for site}
#SUM
#pivot wide at site level
fg_summary_site_wide_SUM <- fg_summary_site_SUM %>%
  pivot_wider(names_from = "functional_group", values_from = "sum_fg_cover")

#richness at every site/year combination ---> NOT AVERAGED,
fg_summary_site_wide_SUM$fg_richness_site <- vegan::specnumber(fg_summary_site_wide_SUM[4:13])

fg_summary_site_wide <- fg_summary_site_wide_SUM %>%
   mutate(site_habitat = paste(site, habitat, sep = "_"))

#AVG -- don't need to use this, just was checking both methods 
#fg_summary_site_wide_AVG <- fg_summary_site_AVG %>%
  #pivot_wider(names_from = "functional_group", values_from = "mean_fg_cover")

#richness at every site/year combination ---> NOT AVERAGED 
#fg_summary_site_wide_AVG$fg_richness_site <- vegan::specnumber(fg_summary_site_wide_AVG[4:13])

#setdiff(fg_summary_site_wide_SUM$fg_richness_site, fg_summary_site_wide_AVG$fg_richness_site)

```

```{r functional richness through time at quad level}
fg_summary_wide_meta_through_time <- fg_summary_wide_meta %>%
  group_by(location, habitat, site) %>%
  summarise(fg_richness_mean = mean(fg_richness_quad),
            fg_richness_se = sd(fg_richness_quad)/n())

#each quad only has ONE value, averaged through time 
#1200 quads in this data set
#this is used in figure 4 things
```

# Figure 3

```{r figure 3b - setup dataframe quad}

#read in Noam's cover data 
alpha_div_cover <- read_csv(here("data", "alpha_diversity_quad_macro_with_cover_updated.csv"))

#merge Noam's data with functional group summary 
fig3df_quad <- merge(alpha_div_cover, fg_summary_wide, by = c("location", "year", "site"))

#choose habitat column & rename from Noam's cover data

fig3df_quad <- fig3df_quad %>%
  select(-habitat.y) %>%
  rename(habitat = habitat.x)

```

## Figure 3b model (functional group)

```{r figure 3b model - fg VS cover - quad}
cover_mod_fg <- glmmTMB(cover_trans ~ fg_richness_quad*habitat + (1|site/location) + (1|year), family = beta_family(), data = fig3df_quad)

hist(fig3df_quad$cover_trans)

summary(cover_mod_fg)
car::Anova(cover_mod_fg)
#                            Chisq Df Pr(>Chisq)    
#fg_richness_quad         31615.090  1  < 2.2e-16 ***
#habitat                     71.048  3  2.546e-15 ***
#fg_richness_quad:habitat   383.667  3  < 2.2e-16 ***
hist(residuals(cover_mod_fg)) # looks fine

# Let's check model diagnostics 
#step 1: linearity
#flat and horizontal --> linearity 
scatter.smooth(residuals(cover_mod_fg) ~ fitted(cover_mod_fg), col = "red")

#looks... fine?

#step 2: homosketadicity..

#standarized pearson's residualizuals

standard_mod3b <- (resid(cover_mod_fg, type = 'pearson')/sd(resid(cover_mod_fg, type = 'pearson')))

#took sqrt

sqrt_residuals_mod3b <-  sqrt(abs(standard_mod3b))

#plotted against fitted 
scatter.smooth(sqrt_residuals_mod3b ~ fitted(cover_mod_fg), col = "red")

#we have determined that this is ok, despite small hump and upward trend

#step 3: check uniformity of residuals plot

plot(performance::simulate_residuals(cover_mod_fg))

#gosh, these seems ... questionsable

r.squaredGLMM(cover_mod_fg) # 0.625, 0.645
emtrends(cover_mod_fg, pairwise ~ habitat, var = "fg_richness_quad") 
#all pairwise comparisons are different

```

## figure 3b plot setup
```{r figure 3b setup}
diversity_ranges_quad_fg <- fig3df_quad %>%
  group_by(habitat) %>%
  summarise(min_fg_richness_quad = min(fg_richness_quad),
            max_fg_richness_quad = max(fg_richness_quad),
            min_cover= min(cover_trans),
            max_cover = max(cover_trans))

cover_emm <- emmip(cover_mod_fg,
                        habitat ~ fg_richness_quad,
                        at = list(fg_richness_quad = seq(0,10,1)), plotit = F, CIs = T)

filtered_cover_effects <- cover_emm %>%
  left_join(diversity_ranges_quad_fg, by = c("habitat")) %>%
  filter(fg_richness_quad >= min_fg_richness_quad & fg_richness_quad <= max_fg_richness_quad)

```

## figure 3b plot

```{r figure 3b plot}
(figure_3b <- 
  ggplot() +
  geom_point(data = fig3df_quad,
             aes(x = fg_richness_quad, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
  geom_line(data = filtered_cover_effects, aes(x = fg_richness_quad, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_cover_effects, aes(x = fg_richness_quad, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Macroalgal Cover\n", x = "Functional Richness", title = "3b.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

#ggsave(filename = "output/figure_3b_quad_v2.png", figure_3b, height = 10, width = 14)
```

```{r figure 3e - setup dataframe site}

#read in Noam's site cover data

alpha_div_site_macro <- read_csv(here("data", "alpha_diversity_site_macro_with_cover.csv"))

#merge site level cover with site functional group richness

site_cover_fg_richness <- merge(alpha_div_site_macro, fg_summary_site_wide, by = c("year", "site_habitat", "site"))

#choose habitat column & rename from Noam's cover data

site_cover_fg_richness_df <- site_cover_fg_richness %>%
  select(-habitat.y) %>%
  rename(habitat = habitat.x)

```

### fig 3e site functional richness mod

```{r site functional richness mod - fig3e}
site_cover_mod_fg <- glmmTMB(cover_trans ~ fg_richness_site*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = site_cover_fg_richness_df)

summary(site_cover_mod_fg)
Anova(site_cover_mod_fg)

#                           Chisq Df Pr(>Chisq)    
#fg_richness_site         31.5959  1  1.898e-08 ***
#habitat                   4.9665  3     0.1743    
#fg_richness_site:habitat  4.3903  3     0.2223

hist(residuals(site_cover_mod_fg)) #looks bad
plot(residuals(site_cover_mod_fg) ~ fitted(site_cover_mod_fg)) # a little heteroscedastic but can't be less than 0
r.squaredGLMM(site_cover_mod_fg) # 0.1826, 0.5729
emtrends(site_cover_mod_fg, pairwise ~ habitat, var = "fg_richness_site") # no difference between the habitats

#step 1: linearity

#flat and horizontal --> linearity 
scatter.smooth(residuals(site_cover_mod_fg) ~ fitted(site_cover_mod_fg), col = "red")
#scatter.smooth(resid(site_cover_mod_fg, type = 'pearson') ~ fitted(site_cover_mod_fg), col = "red")

#looks good

#step 2: homosketadicity..

#standarized residualized

standard_mod3e <- (resid(site_cover_mod_fg, type = 'pearson')/sd(resid(site_cover_mod_fg, type = 'pearson')))

#took sqrt
#https://boostedml.com/2019/03/linear-regression-plots-scale-location-plot.html

sqrt_residuals_mod3e <- sqrt(abs(standard_mod3e))

#plotted against fitted 
scatter.smooth(sqrt_residuals_mod3e ~ fitted(site_cover_mod_fg), col = "red")

#this is ok

#tells us these are uniform
plot(performance::simulate_residuals(site_cover_mod_fg))

#looks okay?

```

### 3e plot setup
```{r figure 3e setup }
diversity_ranges_site <- site_cover_fg_richness_df %>%
  group_by(habitat) %>%
  summarise(min_fg_richness_site = min(fg_richness_site),
            max_fg_richness_site = max(fg_richness_site),
            min_cover= min(cover_trans),
            max_cover = max(cover_trans))

site_cover_emm <- emmip(site_cover_mod_fg,
                        habitat ~ fg_richness_site,
                        at = list(fg_richness_site = seq(0,17,1)), plotit = F, CIs = T)

filtered_site_cover_effects <- site_cover_emm %>%
  left_join(diversity_ranges_site, by = c("habitat")) %>%
  filter(fg_richness_site >= min_fg_richness_site & fg_richness_site <= max_fg_richness_site)

```

### 3e plot

```{r figure 3e plot }
(figure_3e <- 
  ggplot() +
  geom_point(data = site_cover_fg_richness_df,
             aes(x = fg_richness_site, y = cover_trans, colour = habitat), size = 1, alpha = 0.3) + 
  geom_line(data = filtered_site_cover_effects, aes(x = fg_richness_site, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_site_cover_effects, aes(x = fg_richness_site, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Cover\n", x = "Functional Richness", title = "3e.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
#ggsave(filename = "output/figure_3e_v2.png", figure_3e, height = 10, width = 14)
```
#FIGURE 4

#Stability Quad Level

```{r figure 4b stabililty quad level}
#read in noam's stability data 
stability_data <- read_csv(here("data", "diversity_stability_synchrony.csv"))

#merge with fg data 
stability_quad_level <- merge(fg_summary_wide_meta_through_time, stability_data, 
                              by = 
                                c("location","site"))

#choose habitat column & rename from Noam's cover data

stability_quad_level_df <- stability_quad_level %>%
  select(-habitat.x) %>%
  rename(habitat = habitat.y)

#we lose 16 quads for this, 1184 instead of 1200 
```

```{r DSS - Stability}
dss_ranges_fg <- stability_quad_level_df %>%
  group_by(habitat) %>%
  summarise(min_richness_fg = min(fg_richness_mean),
            max_richness_fg = max(fg_richness_mean),
            min_stability= min(cover_stability, na.rm = TRUE),
            max_stability = max(cover_stability, na.rm = TRUE))
```

#figure 4b model
```{r fig4b fg richness stability mod}

rich_stab_mod_fg <- glmmTMB(cover_stability ~ fg_richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = stability_quad_level_df)

summary(rich_stab_mod_fg)
car::Anova(rich_stab_mod_fg)

#                            Chisq Df Pr(>Chisq)    
#fg_richness_mean         3025.235  1  < 2.2e-16 ***
#habitat                   122.635  3  < 2.2e-16 ***
#fg_richness_mean:habitat   20.721  3  0.0001203 ***

hist(residuals(rich_stab_mod_fg)) # ??
plot(residuals(rich_stab_mod_fg) ~ fitted(rich_synch_mod_fg)) #ugh 
r.squaredGLMM(rich_stab_mod_fg) # 0.815, 0.823
emtrends(rich_stab_mod_fg, pairwise ~ habitat, var = "fg_richness_mean") 
# backreef and fringing are different, fringing and outer10 are different, fringing and outer17 are different. (so, fringing is always differnt )

#step 1: linearity

#flat and horizontal --> linearity 
scatter.smooth(residuals(rich_stab_mod_fg) ~ fitted(rich_stab_mod_fg), col = "red")
#scatter.smooth(resid(site_cover_mod_fg, type = 'pearson') ~ fitted(site_cover_mod_fg), col = "red")

#looks good

#step 2: homosketadicity..

#standarized residualized

standard_mod4b <- (resid(rich_stab_mod_fg, type = 'pearson')/sd(resid(rich_stab_mod_fg, type = 'pearson')))

#took sqrt
#https://boostedml.com/2019/03/linear-regression-plots-scale-location-plot.html

sqrt_residuals_mod4b <- sqrt(abs(standard_mod4b))

#plotted against fitted 
scatter.smooth(sqrt_residuals_mod4b ~ fitted(rich_stab_mod_fg), col = "red")

#this is ok

#tells us these are uniform
plot(performance::simulate_residuals(rich_stab_mod_fg))

#yay, looks good!

performance::check_model(rich_stab_mod_fg)

```

```{r figure 4b setup}
rich_stab_emm_fg <- emmip(rich_stab_mod_fg,
                        habitat ~ fg_richness_mean,
                        at = list(fg_richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_rich_stab_effects_fg <- rich_stab_emm_fg %>%
  left_join(dss_ranges_fg, by = c("habitat")) %>%
  filter(fg_richness_mean >= min_richness_fg & fg_richness_mean <= max_richness_fg)
```

```{r}
(figure_4b <- 
  ggplot() +
  geom_point(data = stability_quad_level_df,
             aes(x = fg_richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects_fg, aes(x = fg_richness_mean, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_rich_stab_effects_fg, aes(x = fg_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "Functional Richness", title = "4b.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18),
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

#ggsave(filename = "output/figure_4b_v2.png", figure_4b, height = 10, width = 14)
```

```{r functional group -- stability site}

stability_site <- read_csv(here("data", "diversity_stability_synchrony_site.csv"))
#n = 24

#need to get average functional richness through time! 
fg_summary_site_wide_summary <- fg_summary_site_wide %>%
  group_by(habitat, site, site_habitat) %>%
  summarize(avg_richness_fg = mean(fg_richness_site),
            se_richness_fg = sd(fg_richness_site)/sqrt(n()),
            n = n())

#n = 24!

stability_site_level_fg <- merge(stability_site, fg_summary_site_wide_summary, by = c("site_habitat", "site"))

#need to remove my habitat column and rename Noam's

stability_site_level_fg_df <- stability_site_level_fg %>%
  select(-habitat.y) %>%
  rename(habitat = habitat.x)

```

```{r}
dss_ranges_site_fg_stab <- stability_site_level_fg_df %>%
  group_by(habitat) %>%
  summarise(min_richness_fg = min(avg_richness_fg),
            max_richness_fg = max(avg_richness_fg),
            min_stability= min(cover_stability),
            max_stability = max(cover_stability))
```

```{r fig 4e fg richness stability mod - site level }

rich_stab_mod_site_fg <- glmmTMB(cover_stability~ avg_richness_fg*habitat, family = Gamma(link = "log"), data = stability_site_level_fg_df)
summary(rich_stab_mod_site_fg)
Anova(rich_stab_mod_site_fg)

#                          Chisq Df Pr(>Chisq)   
#avg_richness_fg          3.8463  1   0.049856 * 
#habitat                 12.5935  3   0.005603 **
#avg_richness_fg:habitat  7.8799  3   0.048560 *

hist(residuals(rich_stab_mod_site_fg)) # blocky but fine
plot(residuals(rich_stab_mod_site_fg) ~ fitted(rich_stab_mod_site_fg)) # heteroskedastic - bring to group
r.squaredGLMM(rich_stab_mod_site_fg) # 0.671
emtrends(rich_stab_mod_site_fg, pairwise ~ habitat, var = "avg_richness_fg") 
# forereef 10 and fringing are different

#step 1: linearity

#flat and horizontal --> linearity 
scatter.smooth(residuals(rich_stab_mod_site_fg) ~ fitted(rich_stab_mod_site_fg), col = "red")
#scatter.smooth(resid(site_cover_mod_fg, type = 'pearson') ~ fitted(site_cover_mod_fg), col = "red")

#looks good

#step 2: homosketadicity..

#standarized residualized

standard_mod4e <- (resid(rich_stab_mod_site_fg, type = 'pearson')/sd(resid(rich_stab_mod_site_fg, type = 'pearson')))

#took sqrt
#https://boostedml.com/2019/03/linear-regression-plots-scale-location-plot.html

sqrt_residuals_mod4e <- sqrt(abs(standard_mod4e))

#plotted against fitted 
scatter.smooth(sqrt_residuals_mod4e ~ fitted(rich_stab_mod_site_fg), col = "red")

#this is ok

#tells us these are uniform
plot(performance::simulate_residuals(rich_stab_mod_site_fg))

#yay, looks good!

performance::check_model(rich_stab_mod_site_fg)

```

```{r figure 4e setup}
rich_stab_emm_site_fg <- emmip(rich_stab_mod_site_fg,
                        habitat ~ avg_richness_fg,
                        at = list(avg_richness_fg = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_rich_stab_effects_site <- rich_stab_emm_site_fg %>%
  left_join(dss_ranges_site_fg_stab, by = c("habitat")) %>%
  filter((avg_richness_fg >= min_richness_fg) & (avg_richness_fg <= max_richness_fg))

```

```{r figure 4e}
(figure_4e <- 
  ggplot() +
  geom_point(data = stability_site_level_fg_df,
             aes(x = avg_richness_fg, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects_site, aes(x = avg_richness_fg, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_rich_stab_effects_site, aes(x = avg_richness_fg, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "Functional Richness", title = "4e.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

#ggsave(filename = "output/figure_4e_v2.png", figure_4e, height = 10, width = 14)
```

#STOP HERE

### 2024-10-02 manuscript_1_code.Rmd
added functional group work from Lauren up to figure 3b

@@
 library(dplyr)
+library(janitor)
 ```
@@
 
-### site richness
-```{r site richness mod}
-site_cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover)
-summary(site_cover_mod)
-Anova(site_cover_mod)
-#                    Chisq Df Pr(>Chisq)    
-# richness         58.9195  1  1.643e-14 ***
-# habitat           6.6247  3    0.08487 .  
-# richness:habitat  3.2967  3    0.34811    
-hist(residuals(site_cover_mod)) # some positive skew, not the worst thing ever
-plot(residuals(site_cover_mod) ~ fitted(site_cover_mod)) # a little heteroscedastic but can't be less than 0
-r.squaredGLMM(site_cover_mod) # 0.2212534, 0.5990825
-emtrends(site_cover_mod, pairwise ~ habitat, var = "richness") # no difference between the habitats
-```
-
-### site evenness 
-```{r site level evenness}
-alpha_diversity_site_macro_with_cover_evenness_mod <- alpha_diversity_site_macro_with_cover %>% 
-  drop_na(evenness)
-
-evenness_site_cover_mod <- glmmTMB(cover_trans ~ evenness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover_evenness_mod)
-summary(evenness_site_cover_mod)
-Anova(evenness_site_cover_mod)
-#                    Chisq Df Pr(>Chisq)    
-# richness         58.9195  1  1.643e-14 ***
-# habitat           6.6247  3    0.08487 .  
-# richness:habitat  3.2967  3    0.34811    
-hist(residuals(evenness_site_cover_mod)) # some positive skew, not the worst thing ever
-plot(residuals(evenness_site_cover_mod) ~ fitted(evenness_site_cover_mod)) # a little heteroscedastic but can't be less than 0
-r.squaredGLMM(evenness_site_cover_mod) #0.1206569, 0.5852773
-emtrends(evenness_site_cover_mod, pairwise ~ habitat, var = "evenness") # only the outers are different from each other
-```
-
 ## figure 3a plot setup
@@
 
+## Figure 3b - functional groups
+
+### Setup
+```{r read and prep functional group data}
+algae_fg_raw2 <- read_csv(here("data","algae_functional_groupings_092724.csv"))
+
+#clean column names
+algae_fg <- algae_fg_raw2 %>%
+  janitor::clean_names()
+
+#change species names to be lowercase & have underscore 
+algae_fg$species <- gsub(" ", "_", tolower(algae_fg$species))
+algae_fg$morphology <- trimws(algae_fg$morphology)
+algae_fg$morphology <- gsub(" ", "_", algae_fg$morphology)
+
+#rename species column to taxa
+
+algae_fg <- algae_fg %>%
+  select(species, morphology) %>%
+  dplyr::rename(taxa = species) %>%
+   dplyr::rename(functional_group = morphology)
+
+unique(algae_fg$functional_group)
+
+#merge functional groups with 
+macro_functional_groups <- merge(macro_long_data, algae_fg, all.x = TRUE)
+
+taxa_sum_check <- macro_functional_groups %>%
+  group_by(location, year) %>%
+  summarise(sum_macrocover = sum(prop_cover))
+
+macro_functional_groups %>%
+  filter(location == "lter_2_backreef_algae_transect_1_quad_1") %>%
+  filter(year == "2022")
+
+#summarizing functional groups at the quad level
+fg_summary <- macro_functional_groups %>%
+  group_by(location, year, functional_group, habitat, site) %>%
+  summarise(sum_fg_cover = sum(prop_cover))
+
+max(fg_summary$sum_fg_cover)
+
+#summarizing functional groups at the site level
+#NOTE that this is ADDED and not averaged. This data is only used to find richness at the site level. LE did this as adding and as a sum, both end up the same from a richness perspective. 
+fg_summary_site_SUM <- macro_functional_groups %>%
+  group_by(site, year, habitat, functional_group) %>%
+  summarise(sum_fg_cover = sum(prop_cover))
+
+#don't need to use this, was just to check both methods
+#fg_summary_site_AVG <- macro_functional_groups %>%
+  #group_by(site, year, habitat, functional_group) %>%
+  #summarise(mean_fg_cover = mean(prop_cover))
+
+#fg_summary %>%
+  #filter(sum_fg_cover==1)
+
+#QAQC
+fg_summary_qaqc <- macro_functional_groups %>%
+  group_by(location, year, functional_group) %>%
+  summarise(sum_fg_cover = sum(prop_cover))
+
+fg_qaqc <- fg_summary_qaqc %>%
+  group_by(location, year) %>%
+  summarise(sum_fg_check = sum(sum_fg_cover))
+
+check_them <- merge(taxa_sum_check, fg_qaqc)
+
+check_them %>%
+  filter(all.equal(sum_macrocover, sum_fg_check) != TRUE)
+
+#pivot wide at quad level
+fg_summary_wide <- fg_summary %>%
+  pivot_wider(names_from = "functional_group", values_from = "sum_fg_cover")
+
+#richness at every quad/year combination 
+fg_summary_wide$fg_richness_quad <- vegan::specnumber(fg_summary_wide[5:14])
+
+unique(fg_summary_wide$fg_richness_quad)
+#0 to 4
+
+##### REPEAT AT SITE LEVEL
+#pivot wide at site level
+fg_summary_site_wide_SUM <- fg_summary_site_SUM %>%
+  pivot_wider(names_from = "functional_group", values_from = "sum_fg_cover")
+
+#richness at every site/year combination ---> NOT AVERAGED,
+fg_summary_site_wide_SUM$fg_richness_site <- vegan::specnumber(fg_summary_site_wide_SUM[4:13])
+
+fg_summary_site_wide <- fg_summary_site_wide_SUM %>%
+   mutate(site_habitat = paste(site, habitat, sep = "_"))
+```
+
+### model
+```{r functional richness cover mod quad}
+#merge Noam's data with functional group summary 
+functional_richness_quad_cover <- merge(alpha_diversity_quad_macro_with_cover, fg_summary_wide,
+                                        by = c("location", "year", "site", "habitat"))
+
+cover_mod_fg <- glmmTMB(cover_trans ~ fg_richness_quad*habitat + (1|site/location) + (1|year), data = functional_richness_quad_cover)
+
+summary(cover_mod_fg)
+car::Anova(cover_mod_fg)
+#                              Chisq Df Pr(>Chisq)    
+# fg_richness_quad         10323.133  1  < 2.2e-16 ***
+# habitat                     87.166  3  < 2.2e-16 ***
+# fg_richness_quad:habitat    77.758  3  < 2.2e-16 ***
+hist(residuals(cover_mod_fg)) # a little positive skew, not the worst
+
+r.squaredGLMM(cover_mod_fg) # 0.39, 0.51
+emtrends(cover_mod_fg, pairwise ~ habitat, var = "fg_richness_quad") 
+#all habitats are different from forereef 17
+
+```
+
+## figure 3b plot setup
+```{r figure 3b setup}
+diversity_ranges_quad_fg <- functional_richness_quad_cover %>%
+  group_by(habitat) %>%
+  summarise(min_fg_richness_quad = min(fg_richness_quad),
+            max_fg_richness_quad = max(fg_richness_quad),
+            min_cover= min(cover_trans),
+            max_cover = max(cover_trans))
+
+cover_emm <- emmip(cover_mod_fg,
+                        habitat ~ fg_richness_quad,
+                        at = list(fg_richness_quad = seq(0,10,1)), plotit = F, CIs = T)
+
+filtered_cover_effects <- cover_emm %>%
+  left_join(diversity_ranges_quad_fg, by = c("habitat")) %>%
+  filter(fg_richness_quad >= min_fg_richness_quad & fg_richness_quad <= max_fg_richness_quad)
+
+```
+
+## figure 3b plot
+
+```{r figure 3b plot}
+(figure_3b <- 
+  ggplot() +
+  geom_point(data = functional_richness_quad_cover,
+             aes(x = fg_richness_quad, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
+  geom_line(data = filtered_cover_effects, aes(x = fg_richness_quad, y = inv_logit(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_cover_effects, aes(x = fg_richness_quad, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "", x = "Functional Richness", title = "3b.") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.title.y = element_text(size=24), 
+        axis.text.x = element_text(size=22), 
+        axis.text.y = element_text(size=22),
+        legend.text = element_text(size=22),
+        plot.title = element_text(size=26), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+#ggsave(filename = "output/figure_3b_quad_v2.png", figure_3b, height = 10, width = 14)
+```
+
 ## Figure 3c
@@
 ## Figure 3d site level richness versus cover
+
+### model at site level
+```{r site richness mod}
+site_cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover)
+summary(site_cover_mod)
+Anova(site_cover_mod)
+#                    Chisq Df Pr(>Chisq)    
+# richness         58.9195  1  1.643e-14 ***
+# habitat           6.6247  3    0.08487 .  
+# richness:habitat  3.2967  3    0.34811    
+hist(residuals(site_cover_mod)) # some positive skew, not the worst thing ever
+plot(residuals(site_cover_mod) ~ fitted(site_cover_mod)) # a little heteroscedastic but can't be less than 0
+r.squaredGLMM(site_cover_mod) # 0.2212534, 0.5990825
+emtrends(site_cover_mod, pairwise ~ habitat, var = "richness") # no difference between the habitats
+```
 
@@
 # ggsave(filename = "output/figure_4a_with_no_hab.png", figure_4a_no_hab, height = 10, width = 14)
+```
+
+## Figure 4b - functional group richness stability
+### Setup
+```{r functional richness through time at quad level}
+fg_summary_wide_meta <- merge(fg_summary_wide, meta_df, by = c("location", "site", "habitat", "year") )
+fg_summary_wide_meta_through_time <- fg_summary_wide_meta %>%
+  group_by(location, habitat, site) %>%
+  summarise(fg_richness_mean = mean(fg_richness_quad),
+            fg_richness_se = sd(fg_richness_quad)/n())
+
+#each quad only has ONE value, averaged through time 
+#1200 quads in this data set
+#this is used in figure 4 things
 ```
@@
 
+### site evenness 
+```{r site level evenness}
+alpha_diversity_site_macro_with_cover_evenness_mod <- alpha_diversity_site_macro_with_cover %>% 
+  drop_na(evenness)
+
+evenness_site_cover_mod <- glmmTMB(cover_trans ~ evenness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover_evenness_mod)
+summary(evenness_site_cover_mod)
+Anova(evenness_site_cover_mod)
+#                    Chisq Df Pr(>Chisq)    
+# richness         58.9195  1  1.643e-14 ***
+# habitat           6.6247  3    0.08487 .  
+# richness:habitat  3.2967  3    0.34811    
+hist(residuals(evenness_site_cover_mod)) # some positive skew, not the worst thing ever
+plot(residuals(evenness_site_cover_mod) ~ fitted(evenness_site_cover_mod)) # a little heteroscedastic but can't be less than 0
+r.squaredGLMM(evenness_site_cover_mod) #0.1206569, 0.5852773
+emtrends(evenness_site_cover_mod, pairwise ~ habitat, var = "evenness") # only the outers are different from each other
+```
+
 ```{r evenness cover model site level}

### 2024-10-02 manuscript_1_code.Rmd
completed figure 3 panel

@@
 
-#don't need to use this, was just to check both methods
-#fg_summary_site_AVG <- macro_functional_groups %>%
-  #group_by(site, year, habitat, functional_group) %>%
-  #summarise(mean_fg_cover = mean(prop_cover))
-
-#fg_summary %>%
-  #filter(sum_fg_cover==1)
-
 #QAQC
@@
 
-fg_summary_site_wide <- fg_summary_site_wide_SUM %>%
-   mutate(site_habitat = paste(site, habitat, sep = "_"))
+fg_summary_site_wide <- fg_summary_site_wide_SUM 
 ```
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "", x = "Functional Richness", title = "3b.") +
+  labs(y = "", x = "", title = "3b.") +
   theme_bw() + 
@@
 
+## Figure 3e - functional richness at site level
+### model
+
+```{r site level fg richness model}
+#merge site level cover with site functional group richness
+
+functional_richness_site_cover<- merge(alpha_diversity_site_macro_with_cover, fg_summary_site_wide,
+                                by = c("year", "site",  "habitat"))
+
+site_cover_mod_fg <- glmmTMB(cover_trans ~ fg_richness_site*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = functional_richness_site_cover)
+
+summary(site_cover_mod_fg)
+Anova(site_cover_mod_fg)
+
+#                            Chisq Df Pr(>Chisq)    
+# fg_richness_site         30.9412  1   2.66e-08 ***
+# habitat                   4.8929  3     0.1798    
+# fg_richness_site:habitat  4.2193  3     0.2387
+
+hist(residuals(site_cover_mod_fg)) # positive skew
+plot(residuals(site_cover_mod_fg) ~ fitted(site_cover_mod_fg)) # a little heteroscedastic but can't be less than 0
+r.squaredGLMM(site_cover_mod_fg) # 0.180, 0.563
+emtrends(site_cover_mod_fg, pairwise ~ habitat, var = "fg_richness_site") # no difference between the habitats
+
+```
+```{r site fg richness cover plot setup}
+diversity_ranges_site <- functional_richness_site_cover %>%
+  group_by(habitat) %>%
+  summarise(min_fg_richness_site = min(fg_richness_site),
+            max_fg_richness_site = max(fg_richness_site),
+            min_cover= min(cover_trans),
+            max_cover = max(cover_trans))
+
+site_cover_emm <- emmip(site_cover_mod_fg,
+                        habitat ~ fg_richness_site,
+                        at = list(fg_richness_site = seq(0,17,1)), plotit = F, CIs = T)
+
+filtered_site_cover_effects <- site_cover_emm %>%
+  left_join(diversity_ranges_site, by = c("habitat")) %>%
+  filter(fg_richness_site >= min_fg_richness_site & fg_richness_site <= max_fg_richness_site)
+```
+
+### 3e plot
+
+```{r figure 3e plot }
+(figure_3e <- 
+  ggplot() +
+  geom_point(data = functional_richness_site_cover,
+             aes(x = fg_richness_site, y = cover_trans, colour = habitat), size = 1, alpha = 0.3) + 
+  geom_line(data = filtered_site_cover_effects, aes(x = fg_richness_site, y = inv_logit(yvar), colour = habitat),
+            size = 1.5) +
+  geom_ribbon(data = filtered_site_cover_effects, aes(x = fg_richness_site, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "", x = "Functional Richness", title = "3e.") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.title.y = element_text(size=24), 
+        axis.text.x = element_text(size=22), 
+        axis.text.y = element_text(size=22),
+        legend.text = element_text(size=22),
+        plot.title = element_text(size=26), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+#ggsave(filename = "output/figure_3e_v2.png", figure_3e, height = 10, width = 14)
+```
+
 ## figure 3f - synchrony site level
@@
 # ggsave(filename = "output/figure_3f.png", figure_3f, height = 10, width = 14)
+```
+
+## Figure 3 Full Plot
+```{r ggpubr fig3, fig.height = 8, fig.width=16}
+
+(figure_3 <- ggarrange(figure_3a, figure_3b, figure_3c, figure_3d, figure_3e, figure_3f, 
+                      ncol = 3, nrow = 2, 
+                      common.legend = TRUE,
+                      legend = "bottom", 
+                      font.label = list(size = 20, color = "black", face = "plain")))
+
+# ggsave(filename = "output/figure_3.png", figure_3, height = 8, width = 16)
 ```

### 2024-10-02 manuscript_1_code.Rmd
Completed figure 4 with functional richness

@@
 #this is used in figure 4 things
-```
-
-## figure 4c 
-### model -  synchrony staability
-
-```{r synchrony staability mod}
-synch_stab_mod <- glmmTMB(cover_stability ~ synchrony*habitat + (1|site), family = Gamma("log"), data = diversity_stability_synchrony)
-summary(synch_stab_mod)
-Anova(synch_stab_mod)
+
+#merge with fg data 
+fg_stability_quad_level <- merge(fg_summary_wide_meta_through_time, diversity_stability_synchrony, 
+                              by = 
+                                c("location","site", "habitat"))
+
+dss_ranges_fg <- fg_stability_quad_level %>%
+  group_by(habitat) %>%
+  summarise(min_richness_fg = min(fg_richness_mean),
+            max_richness_fg = max(fg_richness_mean),
+            min_stability= min(cover_stability, na.rm = TRUE),
+            max_stability = max(cover_stability, na.rm = TRUE))
+```
+
+### figure 4b model
+```{r fig4b fg richness stability mod}
+
+rich_stab_mod_fg <- glmmTMB(cover_stability ~ fg_richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = fg_stability_quad_level)
+
+summary(rich_stab_mod_fg)
+car::Anova(rich_stab_mod_fg)
+
 #                             Chisq Df Pr(>Chisq)    
-# synchrony         1308.33  1  < 2.2e-16 ***
-# habitat            103.13  3  < 2.2e-16 ***
-# synchrony:habitat  327.97  3  < 2.2e-16 ***
-hist(residuals(synch_stab_mod)) # looks good
-plot(residuals(synch_stab_mod) ~ fitted(synch_stab_mod)) # same issue as always
-r.squaredGLMM(synch_stab_mod) # 0.63, 0.71
-emtrends(synch_stab_mod, pairwise ~ habitat, var = "synchrony") # fringing and backreef not different
-```
-
-```{r figure 4c setup}
-synch_stab_emm <- emmip(synch_stab_mod,
-                        habitat ~ synchrony,
-                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)
-
-filtered_synch_stab_effects <- synch_stab_emm %>%
-  left_join(dss_ranges, by = c("habitat")) %>%
-  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
-```
-
-### plot figure 4c
-
-```{r figure 4c}
-(figure_4c <- 
+# fg_richness_mean         3079.229  1  < 2.2e-16 ***
+# habitat                   118.605  3  < 2.2e-16 ***
+# fg_richness_mean:habitat   25.692  3  1.107e-05 ***
+
+hist(residuals(rich_stab_mod_fg)) # looks fine
+plot(residuals(rich_stab_mod_fg) ~ fitted(rich_stab_mod_fg)) # same as usual 
+r.squaredGLMM(rich_stab_mod_fg) # 0.819, 0.828
+emtrends(rich_stab_mod_fg, pairwise ~ habitat, var = "fg_richness_mean") 
+#  fringing different from all others
+
+```
+```{r figure 4b plot setup}
+rich_stab_emm_fg <- emmip(rich_stab_mod_fg,
+                        habitat ~ fg_richness_mean,
+                        at = list(fg_richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_rich_stab_effects_fg <- rich_stab_emm_fg %>%
+  left_join(dss_ranges_fg, by = c("habitat")) %>%
+  filter(fg_richness_mean >= min_richness_fg & fg_richness_mean <= max_richness_fg)
+```
+
+### plot fig 4b
+
+```{r plot figure 4b}
+(figure_4b <- 
   ggplot() +
-  geom_point(data = diversity_stability_synchrony,
-             aes(x = synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_synch_stab_effects, aes(x = synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
-  geom_ribbon(data = filtered_synch_stab_effects, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+  geom_point(data = fg_stability_quad_level,
+             aes(x = fg_richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_rich_stab_effects_fg, aes(x = fg_richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
+  geom_ribbon(data = filtered_rich_stab_effects_fg, aes(x = fg_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
               alpha = 0.3) +
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(x = "", y = "", title = "4c.") +
+  labs(y = "", x = "", title = "4b.") +
   theme_bw() + 
@@
          )
-# ggsave(filename = "output/figure_4c.png", figure_4c, height = 10, width = 14)
-```
-
-## figure 4d
-
-###model - richness stability
-
-```{r richness synchrony mod}
-rich_stab_mod_site <- glmmTMB(cover_stability~ richness_mean*habitat, family = Gamma(link = "log"), data = diversity_stability_synchrony_site)
-summary(rich_stab_mod_site)
-Anova(rich_stab_mod_site)
-
+
+#ggsave(filename = "output/figure_4b_v2.png", figure_4b, height = 10, width = 14)
+```
+
+## figure 4c 
+### model -  synchrony staability
+
+```{r synchrony staability mod}
+synch_stab_mod <- glmmTMB(cover_stability ~ synchrony*habitat + (1|site), family = Gamma("log"), data = diversity_stability_synchrony)
+summary(synch_stab_mod)
+Anova(synch_stab_mod)
 #                     Chisq Df Pr(>Chisq)    
-# richness_mean         17.645  1  2.662e-05 ***
-# habitat               32.229  3  4.682e-07 ***
-# richness_mean:habitat  8.075  3    0.04449 *  
-hist(residuals(rich_stab_mod_site)) # blocky but fine
-plot(residuals(rich_stab_mod_site) ~ fitted(rich_stab_mod_site)) # heteroskedastic - bring to group
-r.squaredGLMM(rich_stab_mod_site) # 0.671
-emtrends(rich_stab_mod_site, pairwise ~ habitat, var = "richness_mean") # no differences across habitats
-```
-
-```{r figure 4d setup}
-rich_stab_emm_site <- emmip(rich_stab_mod_site,
-                        habitat ~ richness_mean,
-                        at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)
-
-filtered_rich_stab_effects_site <- rich_stab_emm_site %>%
-  left_join(dss_ranges_site, by = c("habitat")) %>%
-  filter(richness_mean >= min_richness & richness_mean <= max_richness)
-```
-
-### plot figure 4d
-
-```{r figure 4d}
-(figure_4d <- 
+# synchrony         1308.33  1  < 2.2e-16 ***
+# habitat            103.13  3  < 2.2e-16 ***
+# synchrony:habitat  327.97  3  < 2.2e-16 ***
+hist(residuals(synch_stab_mod)) # looks good
+plot(residuals(synch_stab_mod) ~ fitted(synch_stab_mod)) # same issue as always
+r.squaredGLMM(synch_stab_mod) # 0.63, 0.71
+emtrends(synch_stab_mod, pairwise ~ habitat, var = "synchrony") # fringing and backreef not different
+```
+
+```{r figure 4c setup}
+synch_stab_emm <- emmip(synch_stab_mod,
+                        habitat ~ synchrony,
+                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)
+
+filtered_synch_stab_effects <- synch_stab_emm %>%
+  left_join(dss_ranges, by = c("habitat")) %>%
+  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
+```
+
+### plot figure 4c
+
+```{r figure 4c}
+(figure_4c <- 
   ggplot() +
-  geom_point(data = diversity_stability_synchrony_site,
-             aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_rich_stab_effects_site, aes(x = richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
-  geom_ribbon(data = filtered_rich_stab_effects_site, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+  geom_point(data = diversity_stability_synchrony,
+             aes(x = synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_synch_stab_effects, aes(x = synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
+  geom_ribbon(data = filtered_synch_stab_effects, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
               alpha = 0.3) +
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Stability\n", x = "Taxonomic Richness", title = "4d.") +
+  labs(x = "", y = "", title = "4c.") +
   theme_bw() + 
@@
          )
-
-# ggsave(filename = "output/figure_4d.png", figure_4d, height = 10, width = 14)
-```
-
-## figure 4f
-
-### model
-```{r synchrony staability mod site}
-synch_stab_mod_site <- glmmTMB(cover_stability ~ synchrony*habitat, family = Gamma("log"), data = diversity_stability_synchrony_site)
-summary(synch_stab_mod_site)
-Anova(synch_stab_mod_site)
+# ggsave(filename = "output/figure_4c.png", figure_4c, height = 10, width = 14)
+```
+
+## figure 4d
+
+###model - richness stability
+
+```{r richness synchrony mod}
+rich_stab_mod_site <- glmmTMB(cover_stability~ richness_mean*habitat, family = Gamma(link = "log"), data = diversity_stability_synchrony_site)
+summary(rich_stab_mod_site)
+Anova(rich_stab_mod_site)
+
 #                        Chisq Df Pr(>Chisq)    
-# synchrony         22.0686  1  2.631e-06 ***
-# habitat            9.3333  3    0.02517 *  
-# synchrony:habitat 10.5561  3    0.01439 *  
-hist(residuals(synch_stab_mod_site)) # almost uniform
-plot(residuals(synch_stab_mod_site) ~ fitted(synch_stab_mod_site)) # same issue as always
-r.squaredGLMM(synch_stab_mod_site) # 0.678
-emtrends(synch_stab_mod_site, pairwise ~ habitat, var = "synchrony") # backreef is different from forereef 17, fringing is a p = 0.05
-```
-
-```{r figure 4f setup}
-synch_stab_emm_site <- emmip(synch_stab_mod_site,
-                        habitat ~ synchrony,
-                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)
-
-filtered_synch_stab_effects_site <- synch_stab_emm_site %>%
+# richness_mean         17.645  1  2.662e-05 ***
+# habitat               32.229  3  4.682e-07 ***
+# richness_mean:habitat  8.075  3    0.04449 *  
+hist(residuals(rich_stab_mod_site)) # blocky but fine
+plot(residuals(rich_stab_mod_site) ~ fitted(rich_stab_mod_site)) # heteroskedastic - bring to group
+r.squaredGLMM(rich_stab_mod_site) # 0.671
+emtrends(rich_stab_mod_site, pairwise ~ habitat, var = "richness_mean") # no differences across habitats
+```
+
+```{r figure 4d setup}
+rich_stab_emm_site <- emmip(rich_stab_mod_site,
+                        habitat ~ richness_mean,
+                        at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_rich_stab_effects_site <- rich_stab_emm_site %>%
   left_join(dss_ranges_site, by = c("habitat")) %>%
-  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
-```
-
-### plot figure 4f
-
-```{r figure 4f}
-(figure_4f <- 
+  filter(richness_mean >= min_richness & richness_mean <= max_richness)
+```
+
+### plot figure 4d
+
+```{r figure 4d}
+(figure_4d <- 
   ggplot() +
   geom_point(data = diversity_stability_synchrony_site,
-             aes(x = synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_synch_stab_effects_site, aes(x = synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
-  geom_ribbon(data = filtered_synch_stab_effects_site, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+             aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_rich_stab_effects_site, aes(x = richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
+  geom_ribbon(data = filtered_rich_stab_effects_site, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
               alpha = 0.3) +
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(x = "Synchrony", y = "", main = "4f.") +
+  labs(y = "Stability\n", x = "Taxonomic Richness", title = "4d.") +
   theme_bw() + 
@@
          )
+
+# ggsave(filename = "output/figure_4d.png", figure_4d, height = 10, width = 14)
+```
+
+## figure 4e
+
+### setup and model
+
+```{r functional group -- stability site}
+
+#need to get average functional richness through time! 
+fg_summary_site_wide_summary <- fg_summary_site_wide %>%
+  group_by(habitat, site) %>%
+  summarize(avg_richness_fg = mean(fg_richness_site),
+            se_richness_fg = sd(fg_richness_site)/sqrt(n()))
+
+#n = 24!
+
+# merge
+stability_site_level_fg <- merge(diversity_stability_synchrony_site, fg_summary_site_wide_summary, by = c("site", "habitat"))
+
+# extract relevant ranges for plotting
+dss_ranges_site_fg_stab <- stability_site_level_fg %>%
+  group_by(habitat) %>%
+  summarise(min_richness_fg = min(avg_richness_fg),
+            max_richness_fg = max(avg_richness_fg),
+            min_stability= min(cover_stability),
+            max_stability = max(cover_stability))
+
+rich_stab_mod_site_fg <- glmmTMB(cover_stability~ avg_richness_fg*habitat, family = Gamma(link = "log"), data = stability_site_level_fg)
+summary(rich_stab_mod_site_fg)
+Anova(rich_stab_mod_site_fg)
+
+#                           Chisq Df Pr(>Chisq)   
+# avg_richness_fg          3.4019  1   0.065122 . 
+# habitat                 12.8194  3   0.005044 **
+# avg_richness_fg:habitat  7.9921  3   0.046176 *
+
+hist(residuals(rich_stab_mod_site_fg)) # blocky but fine
+plot(residuals(rich_stab_mod_site_fg) ~ fitted(rich_stab_mod_site_fg)) # heteroskedastic
+r.squaredGLMM(rich_stab_mod_site_fg) # 0.563
+emtrends(rich_stab_mod_site_fg, pairwise ~ habitat, var = "avg_richness_fg") 
+# Fringing different from outer 10
+
+```
+
+```{r figure 4e setup}
+rich_stab_emm_site_fg <- emmip(rich_stab_mod_site_fg,
+                        habitat ~ avg_richness_fg,
+                        at = list(avg_richness_fg = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_rich_stab_effects_site <- rich_stab_emm_site_fg %>%
+  left_join(dss_ranges_site_fg_stab, by = c("habitat")) %>%
+  filter((avg_richness_fg >= min_richness_fg) & (avg_richness_fg <= max_richness_fg))
+
+```
+
+### plot figure 4e
+
+```{r figure 4e}
+(figure_4e <- 
+  ggplot() +
+  geom_point(data = stability_site_level_fg,
+             aes(x = avg_richness_fg, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_rich_stab_effects_site, aes(x = avg_richness_fg, y = exp(yvar), colour = habitat), size = 1.5) +
+  geom_ribbon(data = filtered_rich_stab_effects_site, aes(x = avg_richness_fg, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "", x = "Functional Richness", title = "4e.") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.title.y = element_text(size=24), 
+        axis.text.x = element_text(size=22), 
+        axis.text.y = element_text(size=22),
+        legend.text = element_text(size=22),
+        plot.title = element_text(size=26), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+#ggsave(filename = "output/figure_4e_v2.png", figure_4e, height = 10, width = 14)
+```
+
+## figure 4f
+
+### model
+```{r synchrony staability mod site}
+synch_stab_mod_site <- glmmTMB(cover_stability ~ synchrony*habitat, family = Gamma("log"), data = diversity_stability_synchrony_site)
+summary(synch_stab_mod_site)
+Anova(synch_stab_mod_site)
+#                     Chisq Df Pr(>Chisq)    
+# synchrony         22.0686  1  2.631e-06 ***
+# habitat            9.3333  3    0.02517 *  
+# synchrony:habitat 10.5561  3    0.01439 *  
+hist(residuals(synch_stab_mod_site)) # almost uniform
+plot(residuals(synch_stab_mod_site) ~ fitted(synch_stab_mod_site)) # same issue as always
+r.squaredGLMM(synch_stab_mod_site) # 0.678
+emtrends(synch_stab_mod_site, pairwise ~ habitat, var = "synchrony") # backreef is different from forereef 17, fringing is a p = 0.05
+```
+
+```{r figure 4f setup}
+synch_stab_emm_site <- emmip(synch_stab_mod_site,
+                        habitat ~ synchrony,
+                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)
+
+filtered_synch_stab_effects_site <- synch_stab_emm_site %>%
+  left_join(dss_ranges_site, by = c("habitat")) %>%
+  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
+```
+
+### plot figure 4f
+
+```{r figure 4f}
+(figure_4f <- 
+  ggplot() +
+  geom_point(data = diversity_stability_synchrony_site,
+             aes(x = synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_synch_stab_effects_site, aes(x = synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
+  geom_ribbon(data = filtered_synch_stab_effects_site, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(x = "Synchrony", y = "", main = "4f.") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.title.y = element_text(size=24), 
+        axis.text.x = element_text(size=22), 
+        axis.text.y = element_text(size=22),
+        legend.text = element_text(size=22),
+        plot.title = element_text(size=26), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
 # ggsave(filename = "output/figure_4f.png", figure_4f, height = 10, width = 14)
@@
 ## Figure 4 Full Plot
-```{r ggpubr fig4}
-fig4_plot_list <- list(figure_4a, figure_4c, figure_4d, figure_4f)
-figure_4 <- ggarrange(plotlist = fig4_plot_list, 
-                      ncol = 2, nrow = 2, 
-                      labels = c("a.", "b."),
+```{r ggpubr fig4, fig.height = 8, fig.width=16}
+
+(figure_4 <- ggarrange(figure_4a, figure_4b, figure_4c, figure_4d, figure_4e, figure_4f, 
+                      ncol = 3, nrow = 2, 
                       common.legend = TRUE,
                       legend = "bottom", 
-                      font.label = list(size = 20, color = "black", face = "plain"))
-
-#ggsave(filename = "output/figure_4.png", figure_4, height = 10, width = 20)
+                      font.label = list(size = 20, color = "black", face = "plain")))
+
+# ggsave(filename = "output/figure_4.png", figure_4, height = 8, width = 16)
 ```

### 2024-10-03 manuscript_1_code.Rmd
corrected figure for 3b

@@
 
-cover_mod_fg <- glmmTMB(cover_trans ~ fg_richness_quad*habitat + (1|site/location) + (1|year), data = functional_richness_quad_cover)
+cover_mod_fg <- glmmTMB(cover_trans ~ fg_richness_quad*habitat + (1|site/location) + (1|year), family = beta_family(), data = functional_richness_quad_cover)
 
@@
 #                              Chisq Df Pr(>Chisq)    
-# fg_richness_quad         10323.133  1  < 2.2e-16 ***
-# habitat                     87.166  3  < 2.2e-16 ***
-# fg_richness_quad:habitat    77.758  3  < 2.2e-16 ***
-hist(residuals(cover_mod_fg)) # a little positive skew, not the worst
-
-r.squaredGLMM(cover_mod_fg) # 0.39, 0.51
+# fg_richness_quad         31341.052  1  < 2.2e-16 ***
+# habitat                     49.887  3  8.444e-11 ***
+# fg_richness_quad:habitat   447.780  3  < 2.2e-16 ***
+hist(residuals(cover_mod_fg)) # good
+
+r.squaredGLMM(cover_mod_fg) # 0.64, 0.65
 emtrends(cover_mod_fg, pairwise ~ habitat, var = "fg_richness_quad") 

### 2024-10-10 manuscript_1_code.Rmd
Updated figs, created spatial synchrony figs/analyses

@@
 library(janitor)
+library(knitr)
 ```
@@
 mytheme <- theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
+  theme(axis.title.x = element_text(size=24), 
         axis.text.x = element_text(angle = 45, hjust = 1),
-        axis.title.y = element_text(size=18), 
+        axis.title.y = element_text(size=24), 
+        plot.title = element_text(size = 24),
         panel.background = element_rect(fill = "white"), 
@@
 macro_species_cols <- macroalgae[,-which(names(macroalgae) %in% meta_cols)]
+
+# make er long
+macro_long_data <- pivot_longer(
+  data = macroalgae,
+  cols = grep("acanthophora_spicifera", colnames(macroalgae)):grep("valonia_ventricosa", colnames(macroalgae)),
+  names_to = "taxa",
+  values_to = "percent_cover") %>% 
+  mutate(prop_cover = percent_cover/100) %>% 
+  select(-percent_cover)
+
+# write_csv(macro_long_data, here("Data", "macro_long_data.csv"))
 ```
@@
 
-# Figure 1
+# Figure 1 - benthic cover trends among habitats
 
@@
 ## plot after aes changes
-```{r Figure 1 v2}
+```{r Figure 1 v2, fig.height = 10, fig.width=14}
 plot_list <- lapply(levels(figure_1_data$habitat), function(hab) {
@@
 
-figure_1 <- ggarrange(plotlist = plot_list, 
+(figure_1 <- ggarrange(plotlist = plot_list, 
                       ncol = 2, nrow = 2, 
@@
                       legend = "bottom", 
-                      font.label = list(size = 20, color = "black", face = "plain"))
+                      font.label = list(size = 24, color = "black", face = "plain")))
 # 
@@
 
-#ggsave(filename = "output/figure_1v2.png", figure_1, height = 10, width = 14)
-```
-
-# Figure 2
+# ggsave(filename = "output/figure_1v3.png", figure_1, height = 10, width = 14)
+```
+
+# Figure 2 - macroalgal richness trends among habitats
 
@@
 
-```{r}
-macroalgal_cover <- cover_df[,c("year", "habitat", "site", "location", "macroalgae")]
-
-alpha_diversity_quad_macro_with_cover <- left_join(alpha_diversity_quad_macro, macroalgal_cover,by = join_by(location, year, site, habitat)) %>% 
+```{r add cover to quad df}
+macroalgal_cover_quad <- cover_df[,c("year", "habitat", "site", "location", "macroalgae")]
+
+alpha_diversity_quad_macro <-
+  left_join(alpha_diversity_quad_macro, macroalgal_cover_quad,by = join_by(location, year, site, habitat)) %>% 
   mutate(macroalgal_prop_cover = macroalgae / 100) %>% 
@@
         x = "Year",
-        title = "a. Quadrat-level Richness") +
+        title = "Quadrat-level Richness") +
   theme_bw() + 
@@
         x = "Year",
-        title = "b. Site-level Richness") +
+        title = "Site-level Richness") +
   mytheme +
@@
                       ncol = 2, nrow = 1, 
+                      labels = c("a.", "b."),
                       common.legend = TRUE,
                       legend = "bottom", 
-                      font.label = list(size = 20, color = "black", face = "plain")))
-
-# ggsave(filename = "output/figure_2v2.png", figure_2, height = 10, width = 22)
-```
-
-# Figure 3
-
-## set up site-level data for model
-```{r site cover by diversity data setup}
-# alpha_diversity_site_macro_with_cover <- alpha_diversity_site_macro %>% 
-#   dplyr::mutate(percent_cover = rowSums(site_macro[,-which(names(site_macro)  %in% colnames(site_meta_df))])) %>% 
-#   mutate(prop_cover = percent_cover / 100) %>% 
-#   select(-percent_cover) %>% 
-#   mutate(cover_trans = sv_trans(prop_cover))
-
-#there is no reasonable reason why the above chunk doesn't work, but it doesn't, so we did it this way
-alpha_diversity_site_macro$percent_cover <- rowSums(site_macro[,-which(names(site_macro)  %in% colnames(site_meta_df))])
-
-alpha_diversity_site_macro_with_cover <- alpha_diversity_site_macro %>% 
-  dplyr::mutate(prop_cover = percent_cover / 100) %>%
-  select(-percent_cover)
-
-alpha_diversity_site_macro_with_cover$cover_trans <- sv_trans(alpha_diversity_site_macro_with_cover$prop_cover)
-```
+                      font.label = list(size = 24, color = "black", face = "plain")))
+
+# ggsave(filename = "output/figure_2v3.png", figure_2, height = 10, width = 22)
+```
+
+# Figure 3 - ecosystem function
 
@@
 ```{r figure 3a richness quad model}
-alpha_diversity_quad_macro_with_cover$cover_trans <- sv_trans(alpha_diversity_quad_macro_with_cover$macroalgal_prop_cover)
-
-cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro_with_cover)
+alpha_diversity_quad_macro$cover_trans <- sv_trans(alpha_diversity_quad_macro$macroalgal_prop_cover)
+
+cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro)
 summary(cover_mod)
@@
 emtrends(cover_mod, pairwise ~ habitat, var = "richness") # backreef not different from forereef, forereef also not different
+# contrast                    estimate     SE  df z.ratio p.value
+# Fringing - Backreef          0.34720 0.0241 Inf  14.409  <.0001
+# Fringing - Forereef 10m      0.34865 0.0240 Inf  14.514  <.0001
+# Fringing - Forereef 17m      0.38953 0.0250 Inf  15.599  <.0001
+# Backreef - Forereef 10m      0.00144 0.0205 Inf   0.071  0.9999
+# Backreef - Forereef 17m      0.04232 0.0214 Inf   1.979  0.1957
+# Forereef 10m - Forereef 17m  0.04088 0.0209 Inf   1.952  0.2063
 ```
@@
 ```{r figure 3a setup}
-diversity_ranges_quad <- alpha_diversity_quad_macro_with_cover %>%
+diversity_ranges_quad <- alpha_diversity_quad_macro %>%
   group_by(habitat) %>%
@@
 
-# alpha_diversity_quad_macro_with_cover$habitat <- factor(alpha_diversity_quad_macro_with_cover$habitat, 
-#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
-#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
-# 
-# 
-# filtered_cover_effects$habitat <- factor(filtered_cover_effects$habitat, 
-#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
-#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
 ```
@@
   ggplot() +
-  geom_point(data = alpha_diversity_quad_macro_with_cover,
+  geom_point(data = alpha_diversity_quad_macro,
              aes(x = richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Cover\n", x = "", title = "3a.") +
+  labs(y = "Cover\n", x = "", title = "") +
   theme_bw() + 
@@
 
-#merge functional groups with 
+#merge functional groups  
 macro_functional_groups <- merge(macro_long_data, algae_fg, all.x = TRUE)
@@
 #merge Noam's data with functional group summary 
-functional_richness_quad_cover <- merge(alpha_diversity_quad_macro_with_cover, fg_summary_wide,
+functional_richness_quad_cover <- merge(alpha_diversity_quad_macro, fg_summary_wide,
                                         by = c("location", "year", "site", "habitat"))
@@
 emtrends(cover_mod_fg, pairwise ~ habitat, var = "fg_richness_quad") 
-#all habitats are different from forereef 17
+# contrast                    estimate     SE  df z.ratio p.value
+# Fringing - Backreef            0.214 0.0266 Inf   8.051  <.0001
+# Fringing - Forereef 10m        0.346 0.0261 Inf  13.252  <.0001
+# Fringing - Forereef 17m        0.530 0.0262 Inf  20.244  <.0001
+# Backreef - Forereef 10m        0.132 0.0231 Inf   5.708  <.0001
+# Backreef - Forereef 17m        0.316 0.0232 Inf  13.617  <.0001
+# Forereef 10m - Forereef 17m    0.184 0.0223 Inf   8.241  <.0001
 
@@
 
-cover_emm <- emmip(cover_mod_fg,
+cover_emm_fg <- emmip(cover_mod_fg,
                         habitat ~ fg_richness_quad,
@@
 
-filtered_cover_effects <- cover_emm %>%
+filtered_cover_effects_fg <- cover_emm_fg %>%
   left_join(diversity_ranges_quad_fg, by = c("habitat")) %>%
@@
              aes(x = fg_richness_quad, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
-  geom_line(data = filtered_cover_effects, aes(x = fg_richness_quad, y = inv_logit(yvar), colour = habitat)) +
-  geom_ribbon(data = filtered_cover_effects, aes(x = fg_richness_quad, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
+  geom_line(data = filtered_cover_effects_fg, aes(x = fg_richness_quad, y = inv_logit(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_cover_effects_fg, aes(x = fg_richness_quad, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
               alpha = 0.3) +
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "", x = "", title = "3b.") +
+  labs(y = "", x = "", title = "") +
   theme_bw() + 
@@
 ```{r calculate syncrony}
-# make er long
-macro_long_data <- pivot_longer(
-  data = macroalgae,
-  cols = grep("acanthophora_spicifera", colnames(macroalgae)):grep("valonia_ventricosa", colnames(macroalgae)),
-  names_to = "taxa",
-  values_to = "percent_cover") %>% 
-  mutate(prop_cover = percent_cover/100) %>% 
-  select(-percent_cover)
-
-# write_csv(macro_long_data, here("Data", "macro_long_data.csv"))
 
@@
 
-diversity_stability <- alpha_diversity_quad_macro_with_cover %>%
+diversity_stability <- alpha_diversity_quad_macro %>%
   dplyr::group_by(location, habitat, site) %>%
@@
 diversity_stability_synchrony$synchrony_trans <- sv_trans(diversity_stability_synchrony$synchrony)
-# diversity_stability_synchrony <- na.omit(diversity_stability_synchrony) # remove quads that never had algae --> NaN calcs
 
@@
 emtrends(synch_cov_mod, pairwise ~ habitat, var = "synchrony") # forereef 17 distinct from all other habitats
+
+# contrast                    estimate    SE  df z.ratio p.value
+# Fringing - Backreef            0.101 0.290 Inf   0.348  0.9856
+# Fringing - Forereef 10m        0.383 0.365 Inf   1.048  0.7211
+# Fringing - Forereef 17m        3.458 0.484 Inf   7.140  <.0001
+# Backreef - Forereef 10m        0.282 0.378 Inf   0.746  0.8784
+# Backreef - Forereef 17m        3.357 0.492 Inf   6.820  <.0001
+# Forereef 10m - Forereef 17m    3.075 0.501 Inf   6.142  <.0001
+
 ```
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(x = "", y = "", title = "3c.") +
+  labs(x = "", y = "", title = "") +
   theme_bw() + 
@@
 
-diversity_stability_site <- alpha_diversity_site_macro_with_cover %>%
+diversity_stability_site <- alpha_diversity_site_macro %>%
   dplyr::group_by(site_habitat, habitat, site) %>%
@@
 diversity_stability_synchrony_site$synchrony_trans <- sv_trans(diversity_stability_synchrony_site$synchrony)
-# diversity_stability_synchrony_site <- na.omit(diversity_stability_synchrony_site) # remove quads that never had algae --> NaN calcs
+
 # write_csv(diversity_stability_synchrony_site, here("data", "diversity_stability_synchrony_site.csv"))
@@
 
+## set up site-level data for model
+```{r site cover by diversity data setup}
+
+alpha_diversity_site_macro$percent_cover <- rowSums(site_macro[,-which(names(site_macro)  %in% colnames(site_meta_df))])
+
+alpha_diversity_site_macro$prop_cover <- alpha_diversity_site_macro$percent_cover/100
+
+alpha_diversity_site_macro$cover_trans <- sv_trans(alpha_diversity_site_macro$prop_cover)
+```
+
 ### model at site level
 ```{r site richness mod}
-site_cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover)
+site_cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro)
 summary(site_cover_mod)
@@
 emtrends(site_cover_mod, pairwise ~ habitat, var = "richness") # no difference between the habitats
+# contrast                    estimate     SE  df z.ratio p.value
+# Fringing - Backreef          -0.0163 0.0394 Inf  -0.413  0.9762
+# Fringing - Forereef 10m       0.0639 0.0503 Inf   1.271  0.5818
+# Fringing - Forereef 17m       0.0334 0.0612 Inf   0.546  0.9476
+# Backreef - Forereef 10m       0.0801 0.0460 Inf   1.742  0.3020
+# Backreef - Forereef 17m       0.0497 0.0583 Inf   0.853  0.8289
+# Forereef 10m - Forereef 17m  -0.0304 0.0651 Inf  -0.467  0.9662
 ```
@@
 ```{r site richness cover setup}
-diversity_ranges_site <- alpha_diversity_site_macro_with_cover %>%
+diversity_ranges_site <- alpha_diversity_site_macro %>%
   group_by(habitat) %>%
@@
 
-# alpha_diversity_site_macro_with_cover$habitat <- factor(alpha_diversity_site_macro_with_cover$habitat, 
-#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
-#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
-# 
-# 
-# filtered_site_cover_effects$habitat <- factor(filtered_site_cover_effects$habitat, 
-#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
-#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
 ```
@@
   ggplot() +
-  geom_point(data = alpha_diversity_site_macro_with_cover,
+  geom_point(data = alpha_diversity_site_macro,
              aes(x = richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.3) + 
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Cover\n", x = "Macroalgal Richness", title = "3d.") +
+  labs(y = "Cover\n", x = "Macroalgal Richness", title = "") +
   theme_bw() + 
@@
 
-functional_richness_site_cover<- merge(alpha_diversity_site_macro_with_cover, fg_summary_site_wide,
+functional_richness_site_cover<- merge(alpha_diversity_site_macro, fg_summary_site_wide,
                                 by = c("year", "site",  "habitat"))
@@
 emtrends(site_cover_mod_fg, pairwise ~ habitat, var = "fg_richness_site") # no difference between the habitats
+# Fringing - Backreef           0.0379 0.0825 Inf   0.459  0.9679
+# Fringing - Forereef 10m       0.1846 0.0930 Inf   1.985  0.1935
+# Fringing - Forereef 17m       0.0933 0.0960 Inf   0.972  0.7655
+# Backreef - Forereef 10m       0.1467 0.0986 Inf   1.488  0.4447
+# Backreef - Forereef 17m       0.0554 0.1018 Inf   0.545  0.9480
+# Forereef 10m - Forereef 17m  -0.0913 0.1093 Inf  -0.835  0.8377
 
@@
 ```{r site fg richness cover plot setup}
-diversity_ranges_site <- functional_richness_site_cover %>%
+fg_ranges_site <- functional_richness_site_cover %>%
   group_by(habitat) %>%
@@
 
-site_cover_emm <- emmip(site_cover_mod_fg,
+site_cover_fg_emm <- emmip(site_cover_mod_fg,
                         habitat ~ fg_richness_site,
@@
 
-filtered_site_cover_effects <- site_cover_emm %>%
-  left_join(diversity_ranges_site, by = c("habitat")) %>%
+filtered_site_fg_cover_effects <- site_cover_fg_emm %>%
+  left_join(fg_ranges_site, by = c("habitat")) %>%
   filter(fg_richness_site >= min_fg_richness_site & fg_richness_site <= max_fg_richness_site)
@@
              aes(x = fg_richness_site, y = cover_trans, colour = habitat), size = 1, alpha = 0.3) + 
-  geom_line(data = filtered_site_cover_effects, aes(x = fg_richness_site, y = inv_logit(yvar), colour = habitat),
+  geom_line(data = filtered_site_fg_cover_effects, aes(x = fg_richness_site, y = inv_logit(yvar), colour = habitat),
             size = 1.5) +
-  geom_ribbon(data = filtered_site_cover_effects, aes(x = fg_richness_site, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
+  geom_ribbon(data = filtered_site_fg_cover_effects, aes(x = fg_richness_site, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
               alpha = 0.3) +
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "", x = "Functional Richness", title = "3e.") +
+  labs(y = "", x = "Functional Richness", title = "") +
   theme_bw() + 
@@
 emtrends(synch_cov_mod_site, pairwise ~ habitat, var = "synchrony") # forereef 17 different from backreef
+# contrast                    estimate   SE  df z.ratio p.value
+# Fringing - Backreef           -4.299 3.38 Inf  -1.270  0.5821
+# Fringing - Forereef 10m       -0.279 3.22 Inf  -0.087  0.9998
+# Fringing - Forereef 17m        3.675 3.78 Inf   0.973  0.7649
+# Backreef - Forereef 10m        4.019 1.98 Inf   2.031  0.1764
+# Backreef - Forereef 17m        7.973 2.80 Inf   2.848  0.0228
+# Forereef 10m - Forereef 17m    3.954 2.59 Inf   1.526  0.4217
 ```
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(x = "Synchrony", y = "", title = "3f.") +
+  labs(x = "Synchrony", y = "", title = "") +
   theme_bw() + 
@@
                       common.legend = TRUE,
+                      labels = c("a.", "b.", "c.", "d.", "e.", "f."),
                       legend = "bottom", 
-                      font.label = list(size = 20, color = "black", face = "plain")))
-
-# ggsave(filename = "output/figure_3.png", figure_3, height = 8, width = 16)
+                      font.label = list(size = 26, color = "black", face = "plain")))
+
+# ggsave(filename = "output/figure_3v2.png", figure_3, height = 8, width = 16)
 ```
@@
 emtrends(rich_stab_mod, pairwise ~ habitat, var = "richness_mean") # fringing not diff from outer17, backreef not different from outer10
+# contrast                     estimate     SE  df z.ratio p.value
+# Fringing - Backreef          0.186670 0.0414 Inf   4.512  <.0001
+# Fringing - Forereef 10m      0.187196 0.0455 Inf   4.114  0.0002
+# Fringing - Forereef 17m      0.027112 0.0463 Inf   0.586  0.9364
+# Backreef - Forereef 10m      0.000525 0.0405 Inf   0.013  1.0000
+# Backreef - Forereef 17m     -0.159559 0.0411 Inf  -3.885  0.0006
+# Forereef 10m - Forereef 17m -0.160084 0.0439 Inf  -3.647  0.0015
 ```
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Stability\n", x = "", title = "4a.") +
+  labs(y = "Stability\n", x = "", title = "") +
   theme_bw() + 
@@
 
-#each quad only has ONE value, averaged through time 
-#1200 quads in this data set
-#this is used in figure 4 things
-
 #merge with fg data 
@@
 #  fringing different from all others
-
-```
+# contrast                    estimate     SE  df z.ratio p.value
+# Fringing - Backreef           0.1872 0.0468 Inf   3.999  0.0004
+# Fringing - Forereef 10m       0.2348 0.0519 Inf   4.524  <.0001
+# Fringing - Forereef 17m       0.2040 0.0489 Inf   4.173  0.0002
+# Backreef - Forereef 10m       0.0477 0.0472 Inf   1.011  0.7432
+# Backreef - Forereef 17m       0.0169 0.0439 Inf   0.385  0.9806
+# Forereef 10m - Forereef 17m  -0.0308 0.0475 Inf  -0.648  0.9164
+#
+```
+
 ```{r figure 4b plot setup}
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "", x = "", title = "4b.") +
+  labs(y = "", x = "", title = "") +
   theme_bw() + 
@@
 emtrends(synch_stab_mod, pairwise ~ habitat, var = "synchrony") # fringing and backreef not different
+# contrast                    estimate    SE  df z.ratio p.value
+# Fringing - Backreef            0.181 0.107 Inf   1.695  0.3262
+# Fringing - Forereef 10m        1.492 0.142 Inf  10.487  <.0001
+# Fringing - Forereef 17m        2.436 0.149 Inf  16.336  <.0001
+# Backreef - Forereef 10m        1.311 0.157 Inf   8.346  <.0001
+# Backreef - Forereef 17m        2.255 0.161 Inf  14.008  <.0001
+# Forereef 10m - Forereef 17m    0.944 0.181 Inf   5.222  <.0001
 ```
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(x = "", y = "", title = "4c.") +
+  labs(x = "", y = "", title = "") +
   theme_bw() + 
@@
 emtrends(rich_stab_mod_site, pairwise ~ habitat, var = "richness_mean") # no differences across habitats
+# contrast                    estimate    SE  df z.ratio p.value
+# Fringing - Backreef           0.1854 0.120 Inf   1.539  0.4138
+# Fringing - Forereef 10m      -0.2155 0.145 Inf  -1.484  0.4471
+# Fringing - Forereef 17m      -0.2876 0.214 Inf  -1.344  0.5349
+# Backreef - Forereef 10m      -0.4009 0.165 Inf  -2.432  0.0711
+# Backreef - Forereef 17m      -0.4730 0.228 Inf  -2.077  0.1607
+# Forereef 10m - Forereef 17m  -0.0721 0.242 Inf  -0.298  0.9908
 ```
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Stability\n", x = "Taxonomic Richness", title = "4d.") +
+  labs(y = "Stability\n", x = "Taxonomic Richness", title = "") +
   theme_bw() + 
@@
 # Fringing different from outer 10
-
+# contrast                    estimate    SE  df z.ratio p.value
+# Fringing - Backreef           0.2552 0.254 Inf   1.005  0.7463
+# Fringing - Forereef 10m       1.2611 0.469 Inf   2.686  0.0364
+# Fringing - Forereef 17m       0.3543 0.287 Inf   1.235  0.6045
+# Backreef - Forereef 10m       1.0059 0.492 Inf   2.044  0.1720
+# Backreef - Forereef 17m       0.0991 0.323 Inf   0.307  0.9900
+# Forereef 10m - Forereef 17m  -0.9068 0.510 Inf  -1.778  0.2838
 ```
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "", x = "Functional Richness", title = "4e.") +
+  labs(y = "", x = "Functional Richness", title = "") +
   theme_bw() + 
@@
 emtrends(synch_stab_mod_site, pairwise ~ habitat, var = "synchrony") # backreef is different from forereef 17, fringing is a p = 0.05
+#
+# contrast                    estimate    SE  df z.ratio p.value
+# Fringing - Backreef           -0.256 0.971 Inf  -0.263  0.9936
+# Fringing - Forereef 10m        1.324 0.916 Inf   1.445  0.4709
+# Fringing - Forereef 17m        2.930 1.142 Inf   2.567  0.0503
+# Backreef - Forereef 10m        1.580 0.858 Inf   1.842  0.2534
+# Backreef - Forereef 17m        3.186 1.095 Inf   2.909  0.0190
+# Forereef 10m - Forereef 17m    1.606 1.047 Inf   1.534  0.4172
 ```
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(x = "Synchrony", y = "", main = "4f.") +
+  labs(x = "Synchrony", y = "", title = "") +
   theme_bw() + 
@@
                       common.legend = TRUE,
+                      heights = c(1,1),
+                      widths = c(1,1,1),
                       legend = "bottom", 
-                      font.label = list(size = 20, color = "black", face = "plain")))
-
-# ggsave(filename = "output/figure_4.png", figure_4, height = 8, width = 16)
+                      labels = c("a.", "b.", "c.", "d.", "e.", "f."),
+                      font.label = list(size = 26, color = "black", face = "plain")))
+
+# ggsave(filename = "output/figure_4v2.png", figure_4, height = 8, width = 16)
+```
+
+# Figure 5 - spatial synchrony
+
+## create new dfs
+```{r mean and spatial synchrony}
+# calculate spatial synchrony - degree to which quadrats vary together within a site
+spatial_synchrony <- codyn::synchrony(df = alpha_diversity_quad_macro,
+                                   time.var = "year",
+                                   species.var = "location",
+                                   abundance.var = "macroalgal_prop_cover",
+                                   metric = "Loreau",
+                                   replicate.var = "site_habitat")
+colnames(spatial_synchrony) <- c("site_habitat", "spatial_synchrony") # rename to avoid overwrite
+
+# calculate mean synchrony within a site (!= synchrony of species at the site level!)
+mean_synchrony <- diversity_stability_synchrony %>% 
+  group_by(habitat, site) %>% 
+  summarise(synchrony_mean = mean(synchrony))
+
+# merge with site level dss metrics
+dss_spatial <- merge(diversity_stability_site, spatial_synchrony, by = "site_habitat")
+dss_spatial_2 <- merge(dss_spatial, mean_synchrony, by = c("habitat", "site"))
+
+dss_spatial_ranges <- dss_spatial_2 %>%
+  group_by(habitat) %>%
+  summarise(min_spatial_synchrony = min(dss_spatial_2$spatial_synchrony, na.rm = TRUE),
+            max_spatial_synchrony = max(dss_spatial_2$spatial_synchrony, na.rm = TRUE),
+            min_mean_synchrony= min(synchrony_mean, na.rm = TRUE),
+            max_mean_synchrony = max(synchrony_mean, na.rm  = TRUE))
+
+```
+
+## models
+```{r model spatial synchrony}
+spatial_synchrony_mod <- glmmTMB(cover_stability ~ spatial_synchrony*habitat, family = Gamma("log"), data = dss_spatial_2)
+hist(residuals(spatial_synchrony_mod)) # a little skew
+plot(residuals(spatial_synchrony_mod) ~ predict(spatial_synchrony_mod))
+summary(spatial_synchrony_mod)
+Anova(spatial_synchrony_mod)
+#                             Chisq Df Pr(>Chisq)    
+# spatial_synchrony         16.2914  1  5.431e-05 ***
+# habitat                   19.5303  3  0.0002124 ***
+# spatial_synchrony:habitat  6.7239  3  0.0812368 .  
+emtrends(spatial_synchrony_mod, pairwise ~ habitat, var = "spatial_synchrony") # no sig differences but close
+# contrast                    estimate    SE  df z.ratio p.value
+# Fringing - Backreef            0.404 1.765 Inf   0.229  0.9958
+# Fringing - Forereef 10m        1.365 1.754 Inf   0.778  0.8643
+# Fringing - Forereef 17m        3.112 1.882 Inf   1.654  0.3484
+# Backreef - Forereef 10m        0.961 0.857 Inf   1.121  0.6766
+# Backreef - Forereef 17m        2.708 1.095 Inf   2.472  0.0644
+# Forereef 10m - Forereef 17m    1.747 1.078 Inf   1.621  0.3668
+```
+```{r plot spatial synchrony}
+spat_stab_emm <- emmip(spatial_synchrony_mod,
+                        habitat ~ spatial_synchrony,
+                        at = list(spatial_synchrony = seq(0,1,0.01)), plotit = F, CIs = T)
+
+filtered_spat_stab_effects <- spat_stab_emm %>%
+  left_join(dss_spatial_ranges, by = c("habitat")) %>%
+  filter(spatial_synchrony >= min_spatial_synchrony & 
+           spatial_synchrony <= max_spatial_synchrony)
+
+(figure_5a <- 
+  ggplot() +
+  geom_point(data = dss_spatial_2,
+             aes(x = spatial_synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
+  geom_ribbon(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Stability\n", x = "Spatial synchrony", title = "") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.title.y = element_text(size=24), 
+        axis.text.x = element_text(size=22), 
+        axis.text.y = element_text(size=22),
+        legend.text = element_text(size=22),
+        plot.title = element_text(size=26), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+```
+```{r mean synchrony}
+mean_synchrony_mod <- glmmTMB(cover_stability ~ synchrony_mean*habitat, family = Gamma("log"), data = dss_spatial_2)
+hist(residuals(mean_synchrony_mod)) # chunky
+plot(residuals(mean_synchrony_mod) ~ predict(mean_synchrony_mod)) # heteroskedastic
+summary(mean_synchrony_mod)
+Anova(mean_synchrony_mod)
+#                         Chisq Df Pr(>Chisq)    
+# synchrony_mean         30.134  1  4.031e-08 ***
+# habitat                22.008  3  6.498e-05 ***
+# synchrony_mean:habitat 17.251  3  0.0006275 ***
+emtrends(mean_synchrony_mod, pairwise ~ habitat, var = "synchrony_mean") # no sig differences but close
+# contrast                    estimate   SE  df z.ratio p.value
+# Fringing - Backreef            -2.75 1.79 Inf  -1.538  0.4147
+# Fringing - Forereef 10m         2.57 1.05 Inf   2.444  0.0691
+# Fringing - Forereef 17m         3.73 1.27 Inf   2.946  0.0170
+# Backreef - Forereef 10m         5.33 1.85 Inf   2.875  0.0211
+# Backreef - Forereef 17m         6.48 1.98 Inf   3.271  0.0059
+# Forereef 10m - Forereef 17m     1.15 1.35 Inf   0.853  0.8287
+```
+```{r plot mean synchrony}
+mean_synch_emm <- emmip(mean_synchrony_mod,
+                        habitat ~ synchrony_mean,
+                        at = list(synchrony_mean = seq(0,1,0.01)), plotit = F, CIs = T)
+
+filtered_mean_synch_effects <- mean_synch_emm %>%
+  left_join(dss_spatial_ranges, by = c("habitat")) %>%
+  filter(synchrony_mean >= min_spatial_synchrony & 
+           synchrony_mean <= max_spatial_synchrony)
+
+(figure_5b <- 
+  ggplot() +
+  geom_point(data = dss_spatial_2,
+             aes(x = synchrony_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_mean_synch_effects, aes(x = synchrony_mean, y = exp(yvar), colour = habitat), size = 1.5) +
+  geom_ribbon(data = filtered_mean_synch_effects, aes(x = synchrony_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Stability\n", x = "Mean synchrony", title = "") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.title.y = element_text(size=24), 
+        axis.text.x = element_text(size=22), 
+        axis.text.y = element_text(size=22),
+        legend.text = element_text(size=22),
+        plot.title = element_text(size=26), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+```
+
+## panel graph
+```{r figure 5 panel, fig.height = 8, fig.width=16}
+(figure_5 <- ggarrange(figure_5a, figure_5b, 
+                      ncol = 2, nrow = 1, 
+                      common.legend = TRUE,
+                      legend = "bottom", 
+                      labels = c("a.", "b."),
+                      font.label = list(size = 26, color = "black", face = "plain")))
+
+ggsave(filename = "output/figure_5.png", figure_5, height = 8, width = 16)
 ```
@@
 
-quad_evenness_df <- alpha_diversity_quad_macro_with_cover %>% 
+quad_evenness_df <- alpha_diversity_quad_macro %>% 
   drop_na(evenness) %>%
@@
 ```{r evenness cover quad level model}
-alpha_diversity_quad_macro_with_cover_evenness_mod <- alpha_diversity_quad_macro_with_cover %>% 
+alpha_diversity_quad_macro_evenness_mod <- alpha_diversity_quad_macro %>% 
   drop_na(evenness)
 
-evenness_cover_mod_quad <- glmmTMB(cover_trans ~ evenness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro_with_cover_evenness_mod)
+evenness_cover_mod_quad <- glmmTMB(cover_trans ~ evenness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro_evenness_mod)
 summary(evenness_cover_mod_quad)
@@
 ```{r evenness cover plot setup}
-diversity_ranges_quad_evenness <- alpha_diversity_quad_macro_with_cover_evenness_mod %>%
+diversity_ranges_quad_evenness <- alpha_diversity_quad_macro_evenness_mod %>%
   group_by(habitat) %>%
@@
 
-# alpha_diversity_quad_macro_with_cover_evenness_mod$habitat <- factor(alpha_diversity_quad_macro_with_cover_evenness_mod$habitat, 
+# alpha_diversity_quad_macro_evenness_mod$habitat <- factor(alpha_diversity_quad_macro_evenness_mod$habitat, 
 #                              levels = c("fringing", "backreef", "outer_10","outer_17"),
@@
   ggplot() +
-  geom_point(data = alpha_diversity_quad_macro_with_cover_evenness_mod,
+  geom_point(data = alpha_diversity_quad_macro_evenness_mod,
              aes(x = evenness, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
@@
 ```{r site level evenness}
-alpha_diversity_site_macro_with_cover_evenness_mod <- alpha_diversity_site_macro_with_cover %>% 
+alpha_diversity_site_macro_evenness_mod <- alpha_diversity_site_macro %>% 
   drop_na(evenness)
 
-evenness_site_cover_mod <- glmmTMB(cover_trans ~ evenness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover_evenness_mod)
+evenness_site_cover_mod <- glmmTMB(cover_trans ~ evenness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_evenness_mod)
 summary(evenness_site_cover_mod)
@@
 ```{r evenness cover model site level}
-evenness_cover_mod_site <- glmmTMB(cover_trans ~ evenness*habitat + (1|site) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover_evenness_mod)
+evenness_cover_mod_site <- glmmTMB(cover_trans ~ evenness*habitat + (1|site) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_evenness_mod)
 summary(evenness_cover_mod_site)
@@
 ```{r figure 3f evenness site plot setup}
-diversity_ranges_site_evenness <- alpha_diversity_site_macro_with_cover_evenness_mod %>%
+diversity_ranges_site_evenness <- alpha_diversity_site_macro_evenness_mod %>%
   group_by(habitat) %>%
@@
 
-# alpha_diversity_site_macro_with_cover_evenness_mod$habitat <- factor(alpha_diversity_site_macro_with_cover_evenness_mod$habitat, 
+# alpha_diversity_site_macro_evenness_mod$habitat <- factor(alpha_diversity_site_macro_evenness_mod$habitat, 
 #                              levels = c("fringing", "backreef", "outer_10","outer_17"),
@@
   ggplot() +
-  geom_point(data = alpha_diversity_site_macro_with_cover_evenness_mod,
+  geom_point(data = alpha_diversity_site_macro_evenness_mod,
              aes(x = evenness, y = cover_trans, colour = habitat), size = 1, alpha = 0.5) + 

### 2024-10-10 manuscript_1_code.Rmd
commented out unwanted code
fixed ranges on figure 5

@@
   group_by(habitat) %>%
-  summarise(min_spatial_synchrony = min(dss_spatial_2$spatial_synchrony, na.rm = TRUE),
-            max_spatial_synchrony = max(dss_spatial_2$spatial_synchrony, na.rm = TRUE),
+  summarise(min_spatial_synchrony = min(spatial_synchrony, na.rm = TRUE),
+            max_spatial_synchrony = max(spatial_synchrony, na.rm = TRUE),
             min_mean_synchrony= min(synchrony_mean, na.rm = TRUE),
@@
 ```
+
 ```{r plot spatial synchrony}
@@
 ```
+
 ```{r mean synchrony}
@@
 
-ggsave(filename = "output/figure_5.png", figure_5, height = 8, width = 16)
+# ggsave(filename = "output/figure_5.png", figure_5, height = 8, width = 16)
 ```

### 2024-10-11 manuscript_1_code.Rmd
fixed range on mean synchrony

@@
   group_by(habitat, site) %>% 
-  summarise(synchrony_mean = mean(synchrony))
+  summarise(synchrony_mean = mean(synchrony, na.rm = T))
 
@@
 hist(residuals(mean_synchrony_mod)) # chunky
-plot(residuals(mean_synchrony_mod) ~ predict(mean_synchrony_mod)) # heteroskedastic
+plot(residuals(mean_synchrony_mod) ~ predict(mean_synchrony_mod)) # fine
 summary(mean_synchrony_mod)
@@
 #                         Chisq Df Pr(>Chisq)    
-# synchrony_mean         30.134  1  4.031e-08 ***
-# habitat                22.008  3  6.498e-05 ***
-# synchrony_mean:habitat 17.251  3  0.0006275 ***
+# synchrony_mean         32.914  1  9.634e-09 ***
+# habitat                18.523  3  0.0003430 ***
+# synchrony_mean:habitat 19.062  3  0.0002654 ***
 emtrends(mean_synchrony_mod, pairwise ~ habitat, var = "synchrony_mean") # no sig differences but close
 # contrast                    estimate    SE  df z.ratio p.value
-# Fringing - Backreef            -2.75 1.79 Inf  -1.538  0.4147
-# Fringing - Forereef 10m         2.57 1.05 Inf   2.444  0.0691
-# Fringing - Forereef 17m         3.73 1.27 Inf   2.946  0.0170
-# Backreef - Forereef 10m         5.33 1.85 Inf   2.875  0.0211
-# Backreef - Forereef 17m         6.48 1.98 Inf   3.271  0.0059
-# Forereef 10m - Forereef 17m     1.15 1.35 Inf   0.853  0.8287
+# Fringing - Backreef            -3.26 1.503 Inf  -2.166  0.1327
+# Fringing - Forereef 10m         2.22 0.995 Inf   2.231  0.1149
+# Fringing - Forereef 17m         3.37 1.220 Inf   2.764  0.0291
+# Backreef - Forereef 10m         5.48 1.622 Inf   3.377  0.0041
+# Backreef - Forereef 17m         6.63 1.769 Inf   3.748  0.0010
+# Forereef 10m - Forereef 17m     1.15 1.363 Inf   0.846  0.8322
 ```
@@
   left_join(dss_spatial_ranges, by = c("habitat")) %>%
-  filter(synchrony_mean >= min_spatial_synchrony & 
-           synchrony_mean <= max_spatial_synchrony)
+  filter(synchrony_mean >= min_mean_synchrony & 
+           synchrony_mean <= max_mean_synchrony)
 

### 2024-10-11 manuscript_1_code.Rmd
created option for fig 5 that includes plot level richness-synchrony relationships

@@
 
-```{r richness synchrony mod}
+```{r richness stability mod}
 rich_stab_mod <- glmmTMB(cover_stability~ richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = diversity_stability_synchrony)
@@
 
+## diversity-synchrony plot level
+
+```{r ds plot}
+rich_synch_mod <- glmmTMB(synchrony_trans~ richness_mean*habitat + (1|site), family = beta_family(), data = diversity_stability_synchrony)
+summary(rich_synch_mod)
+Anova(rich_synch_mod)
+#                         Chisq Df Pr(>Chisq)    
+# richness_mean         395.416  1  < 2.2e-16 ***
+# habitat               136.859  3  < 2.2e-16 ***
+# richness_mean:habitat  90.031  3  < 2.2e-16 ***
+hist(residuals(rich_synch_mod)) # looks good
+plot(residuals(rich_synch_mod) ~ fitted(rich_synch_mod)) 
+r.squaredGLMM(rich_synch_mod) # 0.71, 0.78
+emtrends(rich_synch_mod, pairwise ~ habitat, var = "richness_mean") # forereefs not different from each other
+# contrast                    estimate    SE  df z.ratio p.value
+# Fringing - Backreef           -0.823 0.161 Inf  -5.113  <.0001
+# Fringing - Forereef 10m       -1.443 0.183 Inf  -7.879  <.0001
+# Fringing - Forereef 17m       -1.628 0.187 Inf  -8.723  <.0001
+# Backreef - Forereef 10m       -0.620 0.151 Inf  -4.103  0.0002
+# Backreef - Forereef 17m       -0.805 0.154 Inf  -5.240  <.0001
+# Forereef 10m - Forereef 17m   -0.185 0.165 Inf  -1.124  0.6745
+
+rich_synch_emm <- emmip(rich_synch_mod,
+                        habitat ~ richness_mean,
+                        at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_rich_synch_effects <- rich_synch_emm %>%
+  left_join(dss_ranges, by = c("habitat")) %>%
+  filter(richness_mean >= min_richness & richness_mean <= max_richness)
+
+(rich_synch_plot <- 
+  ggplot() +
+  geom_point(data = diversity_stability_synchrony,
+             aes(x = richness_mean, y = synchrony_trans, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_rich_synch_effects, aes(x = richness_mean, y = inv_logit(yvar), colour = habitat), size = 1.5) +
+  geom_ribbon(data = filtered_rich_synch_effects, aes(x = richness_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Species synchrony\n", x = "Taxonomic richness", title = "") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.title.y = element_text(size=24), 
+        axis.text.x = element_text(size=22), 
+        axis.text.y = element_text(size=22),
+        legend.text = element_text(size=22),
+        plot.title = element_text(size=26), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+```
+
 ## create new dfs
@@
 ```
+
 ```{r plot mean synchrony}
@@
 ## panel graph
-```{r figure 5 panel, fig.height = 8, fig.width=16}
-(figure_5 <- ggarrange(figure_5a, figure_5b, 
-                      ncol = 2, nrow = 1, 
+```{r figure 5 panel, fig.height = 6, fig.width=22}
+(figure_5 <- ggarrange(rich_synch_plot, figure_5b, figure_5a, 
+                      ncol = 3, nrow = 1, 
                       common.legend = TRUE,
                       legend = "bottom", 
-                      labels = c("a.", "b."),
+                      labels = c("a.", "b.", "c."),
                       font.label = list(size = 26, color = "black", face = "plain")))
 
-# ggsave(filename = "output/figure_5.png", figure_5, height = 8, width = 16)
+# ggsave(filename = "output/figure_5_option2.png", figure_5, height = 8, width = 16)
 ```

### 2024-10-11 manuscript_1_code.Rmd
Added figs and mods for evenness. Most are usesless and should be ignored.

@@
 alpha_diversity_quad_macro <- merge(meta_df, alpha_diversity_quad_macro, by = c("location", "year"))
+alpha_diversity_quad_macro$evenness[is.na(alpha_diversity_quad_macro$evenness)] <- 0
 ```
@@
 alpha_diversity_site_macro <- left_join(site_meta_df, alpha_diversity_site_macro,by = join_by(year, site, habitat, site_habitat))
+# prior transformation did not properly set 0 values for evenness
+alpha_diversity_site_macro$evenness[is.na(alpha_diversity_site_macro$evenness)] <- 0
 ```
@@
 
-## 2c Evenness @ Quad by Year
-### prep
-```{r calculate evennes}
-#dear callie, why is this here?
-site_richness_df <- alpha_diversity_site_macro %>% 
-  group_by(year, habitat) %>% 
-  summarise(richness_mean = mean(richness),
-            richness_se = se(richness))
+## evenness over time
+```{r evenness over time quad,  fig.height = 10, fig.width = 14}
 
@@
             evenness_se = se(evenness))
-```
-
-### plot
-```{r}
-(figure_2c <- ggplot() +
+
+(evenness_time_quad <- ggplot() +
   geom_point(data = quad_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), size = 3) +
@@
   theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.text.x = element_text(angle = 45, hjust = 1),
-        axis.title.y = element_text(size=18), 
-        plot.title = element_text(size=18),
+  theme(axis.title.x = element_text(size=24), 
+        axis.text.x = element_text(angle = 45, hjust = 1, size = 22),
+        axis.text.y = element_text( size = 22),
+        axis.title.y = element_text(size=24), 
+        plot.title = element_text(size=26),
         panel.background = element_blank(), 
@@
 
-# ggsave(filename = "output/figure_2c.png", figure_2c, height = 10, width = 14)
-```
-
-## 2d Evenness @ Site by Year
-### prep
-```{r set up evenness}
+# ggsave(filename = "output/evenness_time_quad.png", evenness_time_quad, height = 10, width = 14)
+```
+
+##  Evenness @ Site by Year
+```{r evenness over time quad,  fig.height = 10, fig.width = 14}
 site_evenness_df <- alpha_diversity_site_macro %>% 
@@
             evenness_se = se(evenness))
-```
-
-### plot
-```{r plot evenness site}
-(figure_2d <- ggplot() +
+
+(evenness_time_site <- ggplot() +
   geom_point(data = site_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), size = 3) +
@@
         x = "Year",
-        title = "d. Site-level evenness") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.text.x = element_text(angle = 45, hjust = 1),
-        axis.title.y = element_text(size=18), 
-        plot.title = element_text(size=18),
+        title = "Site-level evenness") +
+  theme_bw() + 
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.text.x = element_text(angle = 45, hjust = 1, size = 22),
+        axis.text.y = element_text( size = 22),
+        axis.title.y = element_text(size=24), 
+        plot.title = element_text(size=26),
         panel.background = element_blank(), 
@@
 
-# ggsave(filename = "output/figure_2d.png", figure_2d, height = 10, width = 14)
+# ggsave(filename = "output/evenness_time_site.png", evenness_time_site, height = 10, width = 14)
 ```
@@
 ```{r evenness cover quad level model}
-alpha_diversity_quad_macro_evenness_mod <- alpha_diversity_quad_macro %>% 
-  drop_na(evenness)
-
-evenness_cover_mod_quad <- glmmTMB(cover_trans ~ evenness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro_evenness_mod)
+
+# NOTE: I tried transforming evenness and the trends/plot were basically identical, just on a different scale. There's no fixing a shit sandwich - Noam
+evenness_cover_mod_quad <- glmmTMB(cover_trans ~ evenness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro)
 summary(evenness_cover_mod_quad)
@@
 #                    Chisq Df Pr(>Chisq)    
-# richness         30916.73  1  < 2.2e-16 ***
-# habitat            141.78  3  < 2.2e-16 ***
-# richness:habitat   294.54  3  < 2.2e-16 ***
+# evenness         9226.79  1  < 2.2e-16 ***
+# habitat           270.78  3  < 2.2e-16 ***
+# evenness:habitat  411.03  3  < 2.2e-16 ***
 hist(residuals(evenness_cover_mod_quad)) # slight pos skeq
-plot(residuals(evenness_cover_mod_quad) ~ fitted(evenness_cover_mod_quad)) # negative trend but it can't be less than 0 so not too concerned. some wave pattern
-r.squaredGLMM(evenness_cover_mod_quad) # 0.1749907, 0.4042826
-emtrends(evenness_cover_mod_quad, pairwise ~ habitat, var = "evenness") #backreef is different from everything except outer 17, but just barely. everyone else is same same xoxo
+plot(residuals(evenness_cover_mod_quad) ~ fitted(evenness_cover_mod_quad))
+r.squaredGLMM(evenness_cover_mod_quad) # 0.38, 0.46
+emtrends(evenness_cover_mod_quad, pairwise ~ habitat, var = "evenness") # everything different
+#  contrast                    estimate     SE  df z.ratio p.value
+#  Fringing - Backreef           -0.206 0.0708 Inf  -2.904  0.0193
+#  Fringing - Forereef 10m        0.491 0.0679 Inf   7.228  <.0001
+#  Fringing - Forereef 17m        0.920 0.0689 Inf  13.361  <.0001
+#  Backreef - Forereef 10m        0.696 0.0581 Inf  11.994  <.0001
+#  Backreef - Forereef 17m        1.126 0.0595 Inf  18.925  <.0001
+#  Forereef 10m - Forereef 17m    0.429 0.0551 Inf   7.794  <.0001
 
@@
 ```{r evenness cover plot setup}
-diversity_ranges_quad_evenness <- alpha_diversity_quad_macro_evenness_mod %>%
+diversity_ranges_quad_evenness <- alpha_diversity_quad_macro %>%
   group_by(habitat) %>%
@@
 
-# alpha_diversity_quad_macro_evenness_mod$habitat <- factor(alpha_diversity_quad_macro_evenness_mod$habitat, 
-#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
-#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
-# 
-# 
-# filtered_cover_effects_evenness$habitat <- factor(filtered_cover_effects_evenness$habitat, 
-#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
-#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
 ```
@@
 ```{r evenness cover plot}
-(figure_3b <- 
+(evenness_cover_quad_plot <- 
   ggplot() +
-  geom_point(data = alpha_diversity_quad_macro_evenness_mod,
-             aes(x = evenness, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
+  geom_point(data = alpha_diversity_quad_macro,
+             aes(x = evenness, y = cover_trans, colour = habitat), size = 1, alpha = 1) + 
   geom_line(data = filtered_cover_effects_evenness, aes(x = evenness, y = inv_logit(yvar), colour = habitat)) +
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Cover\n", x = "Macroalgal evenness", title = "3b.") +
+  labs(y = "Cover\n", x = "Macroalgal evenness", title = "") +
   theme_bw() + 
@@
 
-#ggsave(filename = "output/figure_3b.png", figure_3b, height = 10, width = 14)
+# ggsave(filename = "output/evenness_cover_quad.png", evenness_cover_quad_plot, height = 10, width = 14)
 ```
@@
 
-### model
-
-### site evenness 
 ```{r site level evenness}
-alpha_diversity_site_macro_evenness_mod <- alpha_diversity_site_macro %>% 
-  drop_na(evenness)
-
-evenness_site_cover_mod <- glmmTMB(cover_trans ~ evenness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_evenness_mod)
+
+evenness_site_cover_mod <- glmmTMB(cover_trans ~ evenness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro)
 summary(evenness_site_cover_mod)
@@
 #                    Chisq Df Pr(>Chisq)   
-# richness         58.9195  1  1.643e-14 ***
-# habitat           6.6247  3    0.08487 .  
-# richness:habitat  3.2967  3    0.34811    
+# evenness          8.1578  1   0.004288 **
+# habitat           6.6305  3   0.084654 . 
+# evenness:habitat 14.6557  3   0.002136 ** 
 hist(residuals(evenness_site_cover_mod)) # some positive skew, not the worst thing ever
 plot(residuals(evenness_site_cover_mod) ~ fitted(evenness_site_cover_mod)) # a little heteroscedastic but can't be less than 0
-r.squaredGLMM(evenness_site_cover_mod) #0.1206569, 0.5852773
+r.squaredGLMM(evenness_site_cover_mod) # 0.1300791 0.5988565
 emtrends(evenness_site_cover_mod, pairwise ~ habitat, var = "evenness") # only the outers are different from each other
-```
-
-```{r evenness cover model site level}
-evenness_cover_mod_site <- glmmTMB(cover_trans ~ evenness*habitat + (1|site) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_evenness_mod)
-summary(evenness_cover_mod_site)
-hist(residuals(evenness_cover_mod_site)) # little wonky but not the worst
-plot(residuals(evenness_cover_mod_site) ~ fitted(evenness_cover_mod_site))
-Anova(evenness_cover_mod_site)
-#                   Chisq Df Pr(>Chisq)    
-# evenness         14.395  1  0.0001482 ***
-# habitat          45.636  3  6.777e-10 ***
-# evenness:habitat 16.126  3  0.0010687 ** 
-r.squaredGLMM(evenness_cover_mod_site) # 0.166, 0.258
-
-emtrends(evenness_cover_mod_site, pairwise ~ habitat, var = "evenness") # fringing different from backreef and outer10
-```
-
-### plot setup
-
-```{r figure 3f evenness site plot setup}
-diversity_ranges_site_evenness <- alpha_diversity_site_macro_evenness_mod %>%
+# contrast                    estimate    SE  df z.ratio p.value
+# Fringing - Backreef            0.967 0.545 Inf   1.774  0.2860
+# Fringing - Forereef 10m        1.258 0.409 Inf   3.073  0.0114
+# Fringing - Forereef 17m       -0.228 0.459 Inf  -0.497  0.9597
+# Backreef - Forereef 10m        0.292 0.563 Inf   0.517  0.9549
+# Backreef - Forereef 17m       -1.195 0.595 Inf  -2.007  0.1854
+# Forereef 10m - Forereef 17m   -1.486 0.472 Inf  -3.146  0.0090
+```
+### plot
+
+```{r  evenness site plot }
+diversity_ranges_site_evenness <- alpha_diversity_site_macro %>%
   group_by(habitat) %>%
@@
 
-cover_emm_evenness_site <- emmip(evenness_cover_mod_site,
+cover_emm_evenness_site <- emmip(evenness_site_cover_mod,
                         habitat ~ evenness,
@@
 
-# alpha_diversity_site_macro_evenness_mod$habitat <- factor(alpha_diversity_site_macro_evenness_mod$habitat, 
-#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
-#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
-# 
-# 
-# 
-```
-
-### plot
-
-```{r evenness cover plot site }
-(figure_3f <- 
+(evenness_cover_site_plot <- 
   ggplot() +
-  geom_point(data = alpha_diversity_site_macro_evenness_mod,
+  geom_point(data = alpha_diversity_site_macro,
              aes(x = evenness, y = cover_trans, colour = habitat), size = 1, alpha = 0.5) + 
@@
 
-# ggsave(filename = "output/figure_3f.png", figure_3f, height = 10, width = 14)
-```
+# ggsave(filename = "output/evenness_cover_site.png", evenness_cover_site_plot, height = 10, width = 14)
+```
+
+## evenness stability
+
+### quad level mod
+
+```{r evenness stability quad mod}
+even_stab_mod_quad <- glmmTMB(cover_stability~ evenness_mean*habitat + (1|site), family = Gamma("log"), data = diversity_stability_synchrony)
+summary(even_stab_mod_quad)
+Anova(even_stab_mod_quad)
+
+#                          Chisq Df Pr(>Chisq)    
+# evenness_mean         1235.417  1    < 2e-16 ***
+# habitat                292.670  3    < 2e-16 ***
+# evenness_mean:habitat   13.518  3    0.00364 ** 
+hist(residuals(even_stab_mod_quad)) # slight positive skew
+plot(residuals(even_stab_mod_quad) ~ fitted(even_stab_mod_quad)) 
+r.squaredGLMM(even_stab_mod_quad) # 0.6582154 0.6582154
+emtrends(even_stab_mod_quad, pairwise ~ habitat, var = "evenness_mean") # forereef 10 diff from fringe and backreef
+# contrast                    estimate    SE  df z.ratio p.value
+# Fringing - Backreef          -0.0123 0.151 Inf  -0.082  0.9998
+# Fringing - Forereef 10m       0.4039 0.148 Inf   2.725  0.0326
+# Fringing - Forereef 17m       0.3023 0.140 Inf   2.152  0.1368
+# Backreef - Forereef 10m       0.4162 0.140 Inf   2.979  0.0153
+# Backreef - Forereef 17m       0.3146 0.131 Inf   2.394  0.0782
+# Forereef 10m - Forereef 17m  -0.1016 0.129 Inf  -0.790  0.8589
+
+even_stab_emm <- emmip(even_stab_mod_quad,
+                        habitat ~ evenness_mean,
+                        at = list(evenness_mean = seq(0,1,0.01)), plotit = F, CIs = T)
+
+filtered_even_stab_effects <- even_stab_emm %>%
+  left_join(dss_ranges, by = c("habitat")) %>%
+  filter(evenness_mean >= min_evenness & evenness_mean <= max_evenness)
+
+(even_stab_plot <- 
+  ggplot() +
+  geom_point(data = diversity_stability_synchrony,
+             aes(x = evenness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_even_stab_effects, aes(x = evenness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
+  geom_ribbon(data = filtered_even_stab_effects, aes(x = evenness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Stability\n", x = "Evenness", title = "") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.title.y = element_text(size=24), 
+        axis.text.x = element_text(size=22), 
+        axis.text.y = element_text(size=22),
+        legend.text = element_text(size=22),
+        plot.title = element_text(size=26), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+
+# ggsave(filename = "output/evenness_stability_quad.png", even_stab_plot, height = 10, width = 14)
+```
+
+## site level evenness stability
+
+```{r site level evenness}
+even_stab_mod_site <- glmmTMB(cover_stability~ evenness_mean*habitat, family = Gamma("log"), data = diversity_stability_synchrony_site)
+summary(even_stab_mod_site)
+Anova(even_stab_mod_site)
+
+#                         Chisq Df Pr(>Chisq)   
+# evenness_mean          0.1035  1   0.747623   
+# habitat               11.8743  3   0.007827 **
+# evenness_mean:habitat  9.6254  3   0.022034 * 
+hist(residuals(even_stab_mod_site)) # a little chunky
+plot(residuals(even_stab_mod_site) ~ fitted(even_stab_mod_site)) 
+r.squaredGLMM(even_stab_mod_site) # 0.51
+emtrends(even_stab_mod_site, pairwise ~ habitat, var = "evenness_mean") # no diff
+# contrast                    estimate   SE  df z.ratio p.value
+# Fringing - Backreef             2.82 1.66 Inf   1.705  0.3210
+# Fringing - Forereef 10m         8.30 3.70 Inf   2.244  0.1115
+# Fringing - Forereef 17m         4.27 1.94 Inf   2.205  0.1219
+# Backreef - Forereef 10m         5.48 3.82 Inf   1.432  0.4793
+# Backreef - Forereef 17m         1.45 2.17 Inf   0.668  0.9093
+# Forereef 10m - Forereef 17m    -4.03 3.95 Inf  -1.019  0.7385
+
+even_stab_emm_site <- emmip(even_stab_mod_site,
+                        habitat ~ evenness_mean,
+                        at = list(evenness_mean = seq(0,1,0.01)), plotit = F, CIs = T)
+
+filtered_even_stab_effects_site <- even_stab_emm_site %>%
+  left_join(dss_ranges, by = c("habitat")) %>%
+  filter(evenness_mean >= min_evenness & evenness_mean <= max_evenness)
+# This is useless
+(even_stab_plot_site <- 
+  ggplot() +
+  geom_point(data = diversity_stability_synchrony_site,
+             aes(x = evenness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_even_stab_effects_site, aes(x = evenness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
+  geom_ribbon(data = filtered_even_stab_effects_site, aes(x = evenness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Stability\n", x = "Evenness", title = "") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.title.y = element_text(size=24), 
+        axis.text.x = element_text(size=22), 
+        axis.text.y = element_text(size=22),
+        legend.text = element_text(size=22),
+        plot.title = element_text(size=26), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+ggsave(filename = "output/evenness_stability_site.png", even_stab_plot_site, height = 10, width = 14)
+```

### 2024-10-11 manuscript_1_code.Rmd
comment out ggsave

@@
          )
-ggsave(filename = "output/evenness_stability_site.png", even_stab_plot_site, height = 10, width = 14)
-```
+# ggsave(filename = "output/evenness_stability_site.png", even_stab_plot_site, height = 10, width = 14)
+```

### 2024-10-11 manuscript_1_code.Rmd
Added option for alpha and gamma stability

@@
 
+```{r add cover to site df}
+
+alpha_diversity_site_macro$percent_cover <- rowSums(site_macro[,-which(names(site_macro)  %in% colnames(site_meta_df))])
+
+alpha_diversity_site_macro$prop_cover <- alpha_diversity_site_macro$percent_cover/100
+
+alpha_diversity_site_macro$cover_trans <- sv_trans(alpha_diversity_site_macro$prop_cover)
+```
+
 ## Figure 2a Richness @ Quad by Year
@@
 ## Figure 3d site level richness versus cover
-
-## set up site-level data for model
-```{r site cover by diversity data setup}
-
-alpha_diversity_site_macro$percent_cover <- rowSums(site_macro[,-which(names(site_macro)  %in% colnames(site_meta_df))])
-
-alpha_diversity_site_macro$prop_cover <- alpha_diversity_site_macro$percent_cover/100
-
-alpha_diversity_site_macro$cover_trans <- sv_trans(alpha_diversity_site_macro$prop_cover)
-```
 
@@
 # calculate mean synchrony within a site (!= synchrony of species at the site level!)
-mean_synchrony <- diversity_stability_synchrony %>% 
+mean_synchrony_stability <- diversity_stability_synchrony %>% 
   group_by(habitat, site) %>% 
-  summarise(synchrony_mean = mean(synchrony, na.rm = T))
+  summarise(synchrony_mean = mean(synchrony, na.rm = T),
+            stability_mean = mean(cover_stability, na.rm = T))
 
@@
 dss_spatial <- merge(diversity_stability_site, spatial_synchrony, by = "site_habitat")
-dss_spatial_2 <- merge(dss_spatial, mean_synchrony, by = c("habitat", "site"))
+dss_spatial_2 <- merge(dss_spatial, mean_synchrony_stability, by = c("habitat", "site"))
 
@@
             min_mean_synchrony= min(synchrony_mean, na.rm = TRUE),
-            max_mean_synchrony = max(synchrony_mean, na.rm  = TRUE))
+            max_mean_synchrony = max(synchrony_mean, na.rm  = TRUE),
+            min_mean_stability = min(stability_mean, na.rm = TRUE),
+            max_mean_stability = max(stability_mean, na.rm  = TRUE))
 
@@
 
+```{r alpha gamma stability mod plot}
+alpha_gamma_stab_mod <- glmmTMB(cover_stability ~ stability_mean*habitat, family = Gamma("log"), data = dss_spatial_2)
+summary(alpha_gamma_stab_mod)
+Anova(alpha_gamma_stab_mod)
+#                          Chisq Df Pr(>Chisq)    
+# stability_mean         61.2725  1   4.97e-15 ***
+# habitat                12.9323  3   0.004785 ** 
+# stability_mean:habitat  9.4832  3   0.023511 *
+
+emtrends(alpha_gamma_stab_mod, pairwise ~ habitat, var = "stability_mean") # backreef different from fringe and forereef 10
+# Fringing - Backreef            1.078 0.381 Inf   2.828  0.0242
+# Fringing - Forereef 10m        0.175 0.345 Inf   0.509  0.9571
+# Fringing - Forereef 17m        0.380 0.305 Inf   1.248  0.5965
+# Backreef - Forereef 10m       -0.903 0.349 Inf  -2.589  0.0475
+# Backreef - Forereef 17m       -0.698 0.309 Inf  -2.256  0.1085
+# Forereef 10m - Forereef 17m    0.205 0.263 Inf   0.779  0.8638
+
+alpha_gamma_emm <- emmip(alpha_gamma_stab_mod,
+                        habitat ~ stability_mean,
+                        at = list(stability_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_ag_effects <- alpha_gamma_emm %>%
+  left_join(dss_spatial_ranges, by = c("habitat")) %>%
+  filter(stability_mean >= min_mean_stability & 
+           stability_mean <= max_mean_stability)
+
+(alpha_gamma_stability_plot <- 
+  ggplot() +
+  geom_point(data = dss_spatial_2,
+             aes(x = stability_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_ag_effects, aes(x = stability_mean, y = exp(yvar), colour = habitat), size = 1.5) +
+  geom_ribbon(data = filtered_ag_effects, aes(x = stability_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Site stability\n", x = "Plot stability", title = "") +
+  theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.title.y = element_text(size=24), 
+        axis.text.x = element_text(size=22), 
+        axis.text.y = element_text(size=22),
+        legend.text = element_text(size=22),
+        plot.title = element_text(size=26), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+         )
+```
+
 ## panel graph
 ```{r figure 5 panel, fig.height = 6, fig.width=22}
-(figure_5 <- ggarrange(rich_synch_plot, figure_5b, figure_5a, 
+(figure_5 <- ggarrange(figure_5b, figure_5a, alpha_gamma_stability_plot, 
                       ncol = 3, nrow = 1, 
@@
 
-# ggsave(filename = "output/figure_5_option2.png", figure_5, height = 8, width = 16)
+# ggsave(filename = "output/figure_5_option3.png", figure_5, height = 8, width = 16)
 ```

### 2024-10-25 manuscript_1_code.Rmd
fixed f5v3 aes
cleaned plot themes, re-did figure 1 without panel titles

@@
 ```{r figure formatting info}
-mytheme <- theme_bw() + 
+trends_theme <- theme_bw() + 
   theme(axis.title.x = element_text(size=24), 
@@
 
+model_themes <-   theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.title.y = element_text(size=24), 
+        axis.text.x = element_text(size=22), 
+        axis.text.y = element_text(size=22),
+        legend.text = element_text(size=22),
+        plot.title = element_text(size=26), 
+        panel.background = element_blank(), 
+        panel.grid.major = element_blank(),  #remove major-grid labels
+        panel.grid.minor = element_blank(),  #remove minor-grid labels
+        plot.background = element_rect(fill = "white"),
+        legend.title = element_blank(),
+        legend.position = 'bottom', # move legend to bottom
+        legend.box = 'horizontal', # make legend horizontal
+        legend.box.background = element_rect(colour = "black"))  # add box around legend
+
 habitat_colours <- c("Fringing" = "#D81B60",
@@
 
-#THIS IS THE BIG CHANGE HERE: 9/19
+# re-factor for figures
 filtered_data$habitat <- factor(filtered_data$habitat, 
@@
 
-#DID NOT TOUCH BELOW:
+# note non-algae species
 not_algae <- c("ascidian", "bare_space", "coral", "coral_rubble", "corallimorpharia", "heteractis_sp", "millepora_platyphylla", "no_data", "sand", "sarcophyton_sp", "shell_debris", "soft_coral", "sponge", "tridacna_sp", "zooxanthellae", "row_sums")
 
+# remove
 algae <- filtered_data[ , -which(names(filtered_data) %in% not_algae)]
@@
 
-# make er long
+# make long dataset
 macro_long_data <- pivot_longer(
@@
 ## plot
-```{r figure 1, fig.height = 10, fig.width=14, warning=FALSE}
-(figure_1 <- ggplot() +
-  geom_point(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), size = 3) +
-  geom_errorbar(data = figure_1_data, aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, colour = functional_group),
-                width = 0, size = 1.3) +
-   geom_line(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
-  facet_wrap(~ habitat) +
-  annotate("segment", x = 2006, xend = 2010, y = 75, yend = 75, linewidth = 2, colour = "grey") +
-  scale_colour_manual(values = functional_group_colours) +
-  # scale_fill_manual(values = functional_group_colours) +
-  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
-  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
-   scale_x_continuous(breaks = unique(cover_long$year)) +
-   labs(y = "Percent Cover \n",
-        x = "Year") +
-  theme_bw() + 
-  mytheme)
-
-# ggsave(filename = "output/figure_1.png", figure_1, height = 10, width = 14)
-```
-
-## plot after aes changes
-```{r Figure 1 v2, fig.height = 10, fig.width=14}
+
+```{r Figure 1, fig.height = 10, fig.width=14}
 plot_list <- lapply(levels(figure_1_data$habitat), function(hab) {
@@
     labs(y = "Percent Cover",
-         x = "Year",
-         title = hab) +  # Add habitat as title
-    theme_bw() +
-    mytheme
+         x = "Year", title = "") +  
+    trends_theme
 })
@@
                       font.label = list(size = 24, color = "black", face = "plain")))
-# 
-# figure_1_caption <- "Figure 1.
-Benthic coverage of four major space holders in Moorea across four habitats (Fringing, Backreef, Forereef at 10m depth, Forereef  at 17m depth).
-Annual fluctuations in percent cover of macroalgae, coral, turf, and crustose coralline algae (CCA) are displayed as points for the average cover for each benthic group across sites at each habitat with lines showing standard error.
-Dotted lines show disturbance events for Cyclone Oli (2010) in black, and the 2019 bleaching event in red, and the gray horizontal line shows crown of thorns (COTs) outbreak which occurred from 2007-2010."
-# 
-# figure_1_with_caption <- annotate_figure(figure_1,
-#                 bottom = text_grob(figure_1_caption, hjust = 0.5, vjust = 1, size = 10))
 
@@
         title = "Quadrat-level Richness") +
-  theme_bw() + 
-  mytheme +
-   theme(axis.text.x = element_text(size=20),
-         axis.title.x = element_text(size=22),
-         axis.title.y = element_text(size=22),
-      axis.text.y = element_text(size=20),
-      plot.title = element_text(size=24),
-      legend.text = element_text(size=20),
-      legend.title = element_blank()))
+   trends_theme)
 
@@
         title = "Site-level Richness") +
-  mytheme +
-   theme(axis.text.x = element_text(size=20),
-         axis.title.x = element_text(size=22),
-      axis.text.y = element_text(size=20),
-      plot.title = element_text(size=24),
-      legend.text = element_text(size=20),
-      legend.title = element_blank(),
-      ))
+  trends_theme)
 
@@
   labs(y = "Cover\n", x = "", title = "") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=24), 
-        axis.title.y = element_text(size=24), 
-        axis.text.x = element_text(size=22), 
-        axis.text.y = element_text(size=22),
-        legend.text = element_text(size=22),
-        plot.title = element_text(size=26), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
+  model_themes 
          )
@@
   labs(y = "", x = "", title = "") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=24), 
-        axis.title.y = element_text(size=24), 
-        axis.text.x = element_text(size=22), 
-        axis.text.y = element_text(size=22),
-        legend.text = element_text(size=22),
-        plot.title = element_text(size=26), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
+  trends_theme
          )
@@
   labs(x = "", y = "", title = "") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=24), 
-        axis.title.y = element_text(size=24), 
-        axis.text.x = element_text(size=22), 
-        axis.text.y = element_text(size=22),
-        legend.text = element_text(size=22),
-        plot.title = element_text(size=26), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
+  trends_theme
          )
@@
   labs(y = "Cover\n", x = "Macroalgal Richness", title = "") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=24), 
-        axis.title.y = element_text(size=24), 
-        axis.text.x = element_text(size=22), 
-        axis.text.y = element_text(size=22),
-        legend.text = element_text(size=22),
-        plot.title = element_text(size=26), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
+  trends_theme
          )
@@
   labs(y = "", x = "Functional Richness", title = "") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=24), 
-        axis.title.y = element_text(size=24), 
-        axis.text.x = element_text(size=22), 
-        axis.text.y = element_text(size=22),
-        legend.text = element_text(size=22),
-        plot.title = element_text(size=26), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
+  trends_theme
          )
@@
   labs(x = "Synchrony", y = "", title = "") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=24), 
-        axis.title.y = element_text(size=24), 
-        axis.text.x = element_text(size=22), 
-        axis.text.y = element_text(size=22),
-        legend.text = element_text(size=22),
-        plot.title = element_text(size=26), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
+  trends_theme
          )
@@
   labs(y = "Stability\n", x = "", title = "") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=24), 
-        axis.title.y = element_text(size=24), 
-        axis.text.x = element_text(size=22), 
-        axis.text.y = element_text(size=22),
-        legend.text = element_text(size=22),
-        plot.title = element_text(size=26), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
+  trends_theme
          )
@@
   labs(y = "Stability\n", x = "Taxonomic Richness", title = "") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=24), 
-        axis.title.y = element_text(size=24), 
-        axis.text.x = element_text(size=22), 
-        axis.text.y = element_text(size=22),
-        legend.text = element_text(size=22),
-        plot.title = element_text(size=26), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
+  trends_theme
          )
@@
   labs(y = "Species synchrony\n", x = "Taxonomic richness", title = "") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=24), 
-        axis.title.y = element_text(size=24), 
-        axis.text.x = element_text(size=22), 
-        axis.text.y = element_text(size=22),
-        legend.text = element_text(size=22),
-        plot.title = element_text(size=26), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
+  trends_theme
          )
@@
   labs(y = "Stability\n", x = "Spatial synchrony", title = "") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=24), 
-        axis.title.y = element_text(size=24), 
-        axis.text.x = element_text(size=22), 
-        axis.text.y = element_text(size=22),
-        legend.text = element_text(size=22),
-        plot.title = element_text(size=26), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
+  trends_theme
          )
@@
   labs(y = "Stability\n", x = "Mean synchrony", title = "") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=24), 
-        axis.title.y = element_text(size=24), 
-        axis.text.x = element_text(size=22), 
-        axis.text.y = element_text(size=22),
-        legend.text = element_text(size=22),
-        plot.title = element_text(size=26), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
+  trends_theme
          )
@@
   labs(y = "Site stability\n", x = "Plot stability", title = "") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=24), 
-        axis.title.y = element_text(size=24), 
-        axis.text.x = element_text(size=22), 
-        axis.text.y = element_text(size=22),
-        legend.text = element_text(size=22),
-        plot.title = element_text(size=26), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
+  trends_theme
          )
@@
 
-# ggsave(filename = "output/figure_5_option3.png", figure_5, height = 8, width = 16)
+# ggsave(filename = "output/figure_5_option3.png", figure_5, height = 6, width = 22)
 ```

### 2024-10-25 R/functions.R
added extract_ranges and filter_ranges function for extracting range of data and filtering for plotting of trends

@@ -19,6 +19,31 @@ sv_trans <- function(prop, s = 0.000005){
   # and s is a small offset 
 }
 
+# standard error
 se <- function(x){
   sd(x)/sqrt(n())
+}
+
+# function to extract the ranges of the data for plotting
+extract_ranges <- function(df, # a dataframe
+                           group, # grouping variable (e.g. habitat)
+                           columns # columns you want to summarize (e.g. richness, cover, synchrony)
+){
+  df %>% # in your dataframe
+    group_by(across(all_of(group))) %>% 
+    summarise(across(all_of(columns), # summarize the columns in your group
+                     list(min = ~min(.x, na.rm = TRUE), max = ~max(.x, na.rm = TRUE)), # by taking the min/max of them
+                     .names = "{.fn}_{.col}")) # and assigning them to the columns name "min/max_value"
+}
+
+# function for filtering ranges of emtrends to only fit range of actual data
+filter_ranges <- function(trend, # emtrends df created from emmip
+                          range_obj, # output df from extract_ranges() custom function
+                          group, # group over which to summarize (e.g. habitat)
+                          value # value to filter range (e.g. richness)
+){
+  trend %>% # take emtrend
+    left_join(range_obj, by = group) %>% # join with range object
+    filter(.data[[value]] >= .data[[paste0("min_", value)]] & # filter so emtrend only covers range of actual values
+             .data[[value]] <= .data[[paste0("max_", value)]])
 }

### 2024-10-25 manuscript_1_code.Rmd
removed evenness. Further streamlined themes and code.

@@
   theme(axis.title.x = element_text(size=24), 
-        axis.text.x = element_text(angle = 45, hjust = 1),
+        axis.text.x = element_text(size = 22, angle = 45, hjust = 1),
         axis.title.y = element_text(size=24), 
+        axis.text.y = element_text(size = 22),
         plot.title = element_text(size = 24),
@@
 
-# ggsave(filename = "output/figure_1v3.png", figure_1, height = 10, width = 14)
+# ggsave(filename = "output/figure_1v4.png", figure_1, height = 10, width = 14)
 ```
@@
 
-#JG Add Evenness
-alpha_diversity_quad_macro$ev <- alpha_diversity_quad_macro$shannon /
-  (log(alpha_diversity_quad_macro$richness))
-
-#makes 1 species 0 and 0 species NA for evenness metric
-alpha_diversity_quad_macro$evenness <- ifelse(
-  alpha_diversity_quad_macro$richness == 0,
-  NA,
-  # If richness == 0, ev is NA
-  ifelse(
-    alpha_diversity_quad_macro$richness == 1,
-    0,
-    # If richness == 1, ev is 0
-    alpha_diversity_quad_macro$ev
-  )
-)
-
 alpha_diversity_quad_macro <- merge(meta_df, alpha_diversity_quad_macro, by = c("location", "year"))
-alpha_diversity_quad_macro$evenness[is.na(alpha_diversity_quad_macro$evenness)] <- 0
+
 ```
@@
 
-```{r site level evenness}
-#JG Add Evenness
-alpha_diversity_site_macro$ev <- alpha_diversity_site_macro$shannon /
-  (log(alpha_diversity_site_macro$richness))
-
-#makes 1 species 0 and 0 species NA for evenness metric
-alpha_diversity_site_macro$evenness <- ifelse(
-  alpha_diversity_site_macro$richness == 0,
-  NA,
-  # If richness == 0, ev is NA
-  ifelse(
-    alpha_diversity_site_macro$richness == 1,
-    0,
-    # If richness == 1, ev is 0
-    alpha_diversity_site_macro$ev
-  )
-)
-
-alpha_diversity_site_macro <- left_join(site_meta_df, alpha_diversity_site_macro,by = join_by(year, site, habitat, site_habitat))
-# prior transformation did not properly set 0 values for evenness
-alpha_diversity_site_macro$evenness[is.na(alpha_diversity_site_macro$evenness)] <- 0
-```
-
 ```{r add cover to quad df}
@@
   scale_colour_manual(values = habitat_colours) +
-  # scale_fill_manual(values = habitat_colours) +
-  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
   annotate("segment", x = 2006, xend = 2010, y = 2, yend = 2, linewidth = 2, colour = "grey") +
@@
         x = "Year",
-        title = "Quadrat-level Richness") +
+        title = "") +
    trends_theme)
@@
   scale_colour_manual(values = habitat_colours) +
-  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
   annotate("segment", x = 2006, xend = 2010, y = 12, yend = 12, linewidth = 2, colour = "grey") +
@@
         x = "Year",
-        title = "Site-level Richness") +
+        title = "") +
   trends_theme)
@@
 fig2_plot_list <- list(figure_2a, figure_2b)
-(figure_2 <- ggarrange(plotlist = fig2_plot_list, 
+(figure_2 <- ggarrange(figure_2a, figure_2b,
                       ncol = 2, nrow = 1, 
@@
 
-# ggsave(filename = "output/figure_2v3.png", figure_2, height = 10, width = 22)
+# ggsave(filename = "output/figure_2v4.png", figure_2, height = 10, width = 22)
 ```
@@
 ```{r figure 3a setup}
-diversity_ranges_quad <- alpha_diversity_quad_macro %>%
-  group_by(habitat) %>%
-  summarise(min_richness = min(richness),
-            max_richness = max(richness),
-            min_cover= min(cover_trans),
-            max_cover = max(cover_trans))
+
+diversity_ranges_quad <- extract_ranges(alpha_diversity_quad_macro, "habitat", c("richness", "cover_trans"))
 
@@
 
-filtered_cover_effects <- cover_emm %>%
-  left_join(diversity_ranges_quad, by = c("habitat")) %>%
-  filter(richness >= min_richness & richness <= max_richness)
+filtered_cover_effects <- filter_ranges(trend = cover_emm, range_obj = diversity_ranges_quad, group = "habitat", value = "richness")
 
@@
 fg_summary_site_wide <- fg_summary_site_wide_SUM
-```
-
-### model
-```{r functional richness cover mod quad}
+
 #merge Noam's data with functional group summary 
@@
                                         by = c("location", "year", "site", "habitat"))
+
+#merge site level cover with site functional group richness
+functional_richness_site_cover<- merge(alpha_diversity_site_macro, fg_summary_site_wide,
+                                by = c("year", "site",  "habitat"))
+```
+
+### model
+```{r functional richness cover mod quad}
 
@@
 ```{r site level fg richness model}
-#merge site level cover with site functional group richness
-
-functional_richness_site_cover<- merge(alpha_diversity_site_macro, fg_summary_site_wide,
-                                by = c("year", "site",  "habitat"))
 
@@
 # SUPPLEMENTAL FIGURE CODE:
-
-## evenness over time
-```{r evenness over time quad,  fig.height = 10, fig.width = 14}
-
-quad_evenness_df <- alpha_diversity_quad_macro %>% 
-  drop_na(evenness) %>%
-  group_by(year, habitat) %>% 
-  summarise(evenness_mean = mean(evenness),
-            evenness_se = se(evenness))
-
-(evenness_time_quad <- ggplot() +
-  geom_point(data = quad_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), size = 3) +
-  geom_line(data = quad_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), linewidth = 1) +
-  geom_errorbar(data = quad_evenness_df, aes(x = year, ymin = evenness_mean - evenness_se, ymax = evenness_mean + evenness_se,
-                                             colour = habitat), width = 0, size = 1.3) +
-  scale_colour_manual(values = habitat_colours) +
-  # scale_fill_manual(values = habitat_colours) +
-  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
-  annotate("segment", x = 2006, xend = 2010, y = .7, yend = .7, linewidth = 2, colour = "grey") +
-  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
-  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
-   scale_x_continuous(breaks = unique(quad_evenness_df$year)) +
-   labs(y = "Macroalgal evenness \n",
-        x = "Year",
-        title = "c. Quadrat-level evenness") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=24), 
-        axis.text.x = element_text(angle = 45, hjust = 1, size = 22),
-        axis.text.y = element_text( size = 22),
-        axis.title.y = element_text(size=24), 
-        plot.title = element_text(size=26),
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black")) )
-
-# ggsave(filename = "output/evenness_time_quad.png", evenness_time_quad, height = 10, width = 14)
-```
-
-##  Evenness @ Site by Year
-```{r evenness over time quad,  fig.height = 10, fig.width = 14}
-site_evenness_df <- alpha_diversity_site_macro %>% 
-  drop_na(evenness) %>%
-  group_by(year, habitat) %>% 
-  summarise(evenness_mean = mean(evenness),
-            evenness_se = se(evenness))
-
-(evenness_time_site <- ggplot() +
-  geom_point(data = site_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), size = 3) +
-  geom_line(data = site_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), linewidth = 1) +
-  geom_errorbar(data = site_evenness_df, aes(x = year, ymin = evenness_mean - evenness_se, ymax = evenness_mean + evenness_se, colour = habitat), size = 1.3, width = 0) +
-  scale_colour_manual(values = habitat_colours) +
-  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
-  annotate("segment", x = 2006, xend = 2010, y = 1, yend = 1, linewidth = 2, colour = "grey") +
-  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
-  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
-   scale_x_continuous(breaks = unique(site_evenness_df$year)) +
-   labs(y = "Macroalgal evenness \n",
-        x = "Year",
-        title = "Site-level evenness") +
-  theme_bw() + 
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=24), 
-        axis.text.x = element_text(angle = 45, hjust = 1, size = 22),
-        axis.text.y = element_text( size = 22),
-        axis.title.y = element_text(size=24), 
-        plot.title = element_text(size=26),
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black")) )
-
-# ggsave(filename = "output/evenness_time_site.png", evenness_time_site, height = 10, width = 14)
-```
-
-## evenness and cover relationships
-
-```{r evenness cover quad level model}
-
-# NOTE: I tried transforming evenness and the trends/plot were basically identical, just on a different scale. There's no fixing a shit sandwich - Noam
-evenness_cover_mod_quad <- glmmTMB(cover_trans ~ evenness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro)
-summary(evenness_cover_mod_quad)
-Anova(evenness_cover_mod_quad)
-#                    Chisq Df Pr(>Chisq)    
-# evenness         9226.79  1  < 2.2e-16 ***
-# habitat           270.78  3  < 2.2e-16 ***
-# evenness:habitat  411.03  3  < 2.2e-16 ***
-hist(residuals(evenness_cover_mod_quad)) # slight pos skeq
-plot(residuals(evenness_cover_mod_quad) ~ fitted(evenness_cover_mod_quad))
-r.squaredGLMM(evenness_cover_mod_quad) # 0.38, 0.46
-emtrends(evenness_cover_mod_quad, pairwise ~ habitat, var = "evenness") # everything different
-#  contrast                    estimate     SE  df z.ratio p.value
-#  Fringing - Backreef           -0.206 0.0708 Inf  -2.904  0.0193
-#  Fringing - Forereef 10m        0.491 0.0679 Inf   7.228  <.0001
-#  Fringing - Forereef 17m        0.920 0.0689 Inf  13.361  <.0001
-#  Backreef - Forereef 10m        0.696 0.0581 Inf  11.994  <.0001
-#  Backreef - Forereef 17m        1.126 0.0595 Inf  18.925  <.0001
-#  Forereef 10m - Forereef 17m    0.429 0.0551 Inf   7.794  <.0001
-
-```
-
-```{r evenness cover plot setup}
-diversity_ranges_quad_evenness <- alpha_diversity_quad_macro %>%
-  group_by(habitat) %>%
-  summarise(min_evenness = min(evenness),
-            max_evenness = max(evenness),
-            min_cover= min(cover_trans),
-            max_cover = max(cover_trans))
-
-cover_emm_evenness <- emmip(evenness_cover_mod_quad,
-                        habitat ~ evenness,
-                        at = list(evenness = seq(0,10,1)), plotit = F, CIs = T)
-
-filtered_cover_effects_evenness <- cover_emm_evenness %>%
-  left_join(diversity_ranges_quad_evenness, by = c("habitat")) %>%
-  filter(evenness >= min_evenness & evenness <= max_evenness)
-
-```
-
-```{r evenness cover plot}
-(evenness_cover_quad_plot <- 
-  ggplot() +
-  geom_point(data = alpha_diversity_quad_macro,
-             aes(x = evenness, y = cover_trans, colour = habitat), size = 1, alpha = 1) + 
-  geom_line(data = filtered_cover_effects_evenness, aes(x = evenness, y = inv_logit(yvar), colour = habitat)) +
-  geom_ribbon(data = filtered_cover_effects_evenness, aes(x = evenness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
-              alpha = 0.3) +
-  scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
-  labs(y = "Cover\n", x = "Macroalgal evenness", title = "") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-# ggsave(filename = "output/evenness_cover_quad.png", evenness_cover_quad_plot, height = 10, width = 14)
-```
-
-## site level evenness versus cover
-
-```{r site level evenness}
-
-evenness_site_cover_mod <- glmmTMB(cover_trans ~ evenness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro)
-summary(evenness_site_cover_mod)
-Anova(evenness_site_cover_mod)
-#                    Chisq Df Pr(>Chisq)   
-# evenness          8.1578  1   0.004288 **
-# habitat           6.6305  3   0.084654 . 
-# evenness:habitat 14.6557  3   0.002136 ** 
-hist(residuals(evenness_site_cover_mod)) # some positive skew, not the worst thing ever
-plot(residuals(evenness_site_cover_mod) ~ fitted(evenness_site_cover_mod)) # a little heteroscedastic but can't be less than 0
-r.squaredGLMM(evenness_site_cover_mod) # 0.1300791 0.5988565
-emtrends(evenness_site_cover_mod, pairwise ~ habitat, var = "evenness") # only the outers are different from each other
-# contrast                    estimate    SE  df z.ratio p.value
-# Fringing - Backreef            0.967 0.545 Inf   1.774  0.2860
-# Fringing - Forereef 10m        1.258 0.409 Inf   3.073  0.0114
-# Fringing - Forereef 17m       -0.228 0.459 Inf  -0.497  0.9597
-# Backreef - Forereef 10m        0.292 0.563 Inf   0.517  0.9549
-# Backreef - Forereef 17m       -1.195 0.595 Inf  -2.007  0.1854
-# Forereef 10m - Forereef 17m   -1.486 0.472 Inf  -3.146  0.0090
-```
-### plot
-
-```{r  evenness site plot }
-diversity_ranges_site_evenness <- alpha_diversity_site_macro %>%
-  group_by(habitat) %>%
-  summarise(min_evenness = min(evenness),
-            max_evenness = max(evenness),
-            min_cover= min(cover_trans),
-            max_cover = max(cover_trans))
-
-cover_emm_evenness_site <- emmip(evenness_site_cover_mod,
-                        habitat ~ evenness,
-                        at = list(evenness = seq(0,1,0.01)), plotit = F, CIs = T)
-
-filtered_cover_effects_evenness_site <- cover_emm_evenness_site %>%
-  left_join(diversity_ranges_site_evenness, by = c("habitat")) %>%
-  filter(evenness >= min_evenness & evenness <= max_evenness)
-
-(evenness_cover_site_plot <- 
-  ggplot() +
-  geom_point(data = alpha_diversity_site_macro,
-             aes(x = evenness, y = cover_trans, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_cover_effects_evenness_site, aes(x = evenness, y = inv_logit(yvar), colour = habitat)) +
-  geom_ribbon(data = filtered_cover_effects_evenness_site, aes(x = evenness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
-              alpha = 0.3) +
-  scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
-  labs(y = "Cover\n", x = "Macroalgal evenness", title = "3f. Site Level Evenness") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=18), 
-        axis.title.y = element_text(size=18), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.text = element_text(size = rel(1)), # increase legend text
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-# ggsave(filename = "output/evenness_cover_site.png", evenness_cover_site_plot, height = 10, width = 14)
-```
-
-## evenness stability
-
-### quad level mod
-
-```{r evenness stability quad mod}
-even_stab_mod_quad <- glmmTMB(cover_stability~ evenness_mean*habitat + (1|site), family = Gamma("log"), data = diversity_stability_synchrony)
-summary(even_stab_mod_quad)
-Anova(even_stab_mod_quad)
-
-#                          Chisq Df Pr(>Chisq)    
-# evenness_mean         1235.417  1    < 2e-16 ***
-# habitat                292.670  3    < 2e-16 ***
-# evenness_mean:habitat   13.518  3    0.00364 ** 
-hist(residuals(even_stab_mod_quad)) # slight positive skew
-plot(residuals(even_stab_mod_quad) ~ fitted(even_stab_mod_quad)) 
-r.squaredGLMM(even_stab_mod_quad) # 0.6582154 0.6582154
-emtrends(even_stab_mod_quad, pairwise ~ habitat, var = "evenness_mean") # forereef 10 diff from fringe and backreef
-# contrast                    estimate    SE  df z.ratio p.value
-# Fringing - Backreef          -0.0123 0.151 Inf  -0.082  0.9998
-# Fringing - Forereef 10m       0.4039 0.148 Inf   2.725  0.0326
-# Fringing - Forereef 17m       0.3023 0.140 Inf   2.152  0.1368
-# Backreef - Forereef 10m       0.4162 0.140 Inf   2.979  0.0153
-# Backreef - Forereef 17m       0.3146 0.131 Inf   2.394  0.0782
-# Forereef 10m - Forereef 17m  -0.1016 0.129 Inf  -0.790  0.8589
-
-even_stab_emm <- emmip(even_stab_mod_quad,
-                        habitat ~ evenness_mean,
-                        at = list(evenness_mean = seq(0,1,0.01)), plotit = F, CIs = T)
-
-filtered_even_stab_effects <- even_stab_emm %>%
-  left_join(dss_ranges, by = c("habitat")) %>%
-  filter(evenness_mean >= min_evenness & evenness_mean <= max_evenness)
-
-(even_stab_plot <- 
-  ggplot() +
-  geom_point(data = diversity_stability_synchrony,
-             aes(x = evenness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_even_stab_effects, aes(x = evenness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
-  geom_ribbon(data = filtered_even_stab_effects, aes(x = evenness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
-              alpha = 0.3) +
-  scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
-  labs(y = "Stability\n", x = "Evenness", title = "") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=24), 
-        axis.title.y = element_text(size=24), 
-        axis.text.x = element_text(size=22), 
-        axis.text.y = element_text(size=22),
-        legend.text = element_text(size=22),
-        plot.title = element_text(size=26), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-
-# ggsave(filename = "output/evenness_stability_quad.png", even_stab_plot, height = 10, width = 14)
-```
-
-## site level evenness stability
-
-```{r site level evenness}
-even_stab_mod_site <- glmmTMB(cover_stability~ evenness_mean*habitat, family = Gamma("log"), data = diversity_stability_synchrony_site)
-summary(even_stab_mod_site)
-Anova(even_stab_mod_site)
-
-#                         Chisq Df Pr(>Chisq)   
-# evenness_mean          0.1035  1   0.747623   
-# habitat               11.8743  3   0.007827 **
-# evenness_mean:habitat  9.6254  3   0.022034 * 
-hist(residuals(even_stab_mod_site)) # a little chunky
-plot(residuals(even_stab_mod_site) ~ fitted(even_stab_mod_site)) 
-r.squaredGLMM(even_stab_mod_site) # 0.51
-emtrends(even_stab_mod_site, pairwise ~ habitat, var = "evenness_mean") # no diff
-# contrast                    estimate   SE  df z.ratio p.value
-# Fringing - Backreef             2.82 1.66 Inf   1.705  0.3210
-# Fringing - Forereef 10m         8.30 3.70 Inf   2.244  0.1115
-# Fringing - Forereef 17m         4.27 1.94 Inf   2.205  0.1219
-# Backreef - Forereef 10m         5.48 3.82 Inf   1.432  0.4793
-# Backreef - Forereef 17m         1.45 2.17 Inf   0.668  0.9093
-# Forereef 10m - Forereef 17m    -4.03 3.95 Inf  -1.019  0.7385
-
-even_stab_emm_site <- emmip(even_stab_mod_site,
-                        habitat ~ evenness_mean,
-                        at = list(evenness_mean = seq(0,1,0.01)), plotit = F, CIs = T)
-
-filtered_even_stab_effects_site <- even_stab_emm_site %>%
-  left_join(dss_ranges, by = c("habitat")) %>%
-  filter(evenness_mean >= min_evenness & evenness_mean <= max_evenness)
-# This is useless
-(even_stab_plot_site <- 
-  ggplot() +
-  geom_point(data = diversity_stability_synchrony_site,
-             aes(x = evenness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_even_stab_effects_site, aes(x = evenness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
-  geom_ribbon(data = filtered_even_stab_effects_site, aes(x = evenness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
-              alpha = 0.3) +
-  scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
-  labs(y = "Stability\n", x = "Evenness", title = "") +
-  theme_bw() + 
-  theme(axis.title.x = element_text(size=24), 
-        axis.title.y = element_text(size=24), 
-        axis.text.x = element_text(size=22), 
-        axis.text.y = element_text(size=22),
-        legend.text = element_text(size=22),
-        plot.title = element_text(size=26), 
-        panel.background = element_blank(), 
-        panel.grid.major = element_blank(),  #remove major-grid labels
-        panel.grid.minor = element_blank(),  #remove minor-grid labels
-        plot.background = element_rect(fill = "white"),
-        legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
-        legend.box = 'horizontal', # make legend horizontal
-        legend.box.background = element_rect(colour = "black"))  # add box around legend
-         )
-# ggsave(filename = "output/evenness_stability_site.png", even_stab_plot_site, height = 10, width = 14)
-```

### 2024-10-25 manuscript_1_code.Rmd
cleaned code up through figure 3 with new functions for extracting ranges of data (see functions.R)

@@
 
+```{r read and prep functional group data}
+algae_fg_raw2 <- read_csv(here("data","algae_functional_groupings_092724.csv"))
+
+#clean column names
+algae_fg <- algae_fg_raw2 %>%
+  janitor::clean_names()
+
+#change species names to be lowercase & have underscore 
+algae_fg$species <- gsub(" ", "_", tolower(algae_fg$species))
+algae_fg$morphology <- trimws(algae_fg$morphology)
+algae_fg$morphology <- gsub(" ", "_", algae_fg$morphology)
+
+#rename species column to taxa
+
+algae_fg <- algae_fg %>%
+  select(species, morphology) %>%
+  dplyr::rename(taxa = species) %>%
+   dplyr::rename(functional_group = morphology)
+
+unique(algae_fg$functional_group)
+
+#merge functional groups  
+macro_functional_groups <- merge(macro_long_data, algae_fg, all.x = TRUE)
+
+taxa_sum_check <- macro_functional_groups %>%
+  group_by(location, year) %>%
+  summarise(sum_macrocover = sum(prop_cover))
+
+macro_functional_groups %>%
+  filter(location == "lter_2_backreef_algae_transect_1_quad_1") %>%
+  filter(year == "2022")
+
+#summarizing functional groups at the quad level
+fg_summary <- macro_functional_groups %>%
+  group_by(location, year, functional_group, habitat, site) %>%
+  summarise(sum_fg_cover = sum(prop_cover))
+
+max(fg_summary$sum_fg_cover)
+
+#summarizing functional groups at the site level
+#NOTE that this is ADDED and not averaged. This data is only used to find richness at the site level. LE did this as adding and as a sum, both end up the same from a richness perspective. 
+fg_summary_site_SUM <- macro_functional_groups %>%
+  group_by(site, year, habitat, functional_group) %>%
+  summarise(sum_fg_cover = sum(prop_cover))
+
+#QAQC
+fg_summary_qaqc <- macro_functional_groups %>%
+  group_by(location, year, functional_group) %>%
+  summarise(sum_fg_cover = sum(prop_cover))
+
+fg_qaqc <- fg_summary_qaqc %>%
+  group_by(location, year) %>%
+  summarise(sum_fg_check = sum(sum_fg_cover))
+
+check_them <- merge(taxa_sum_check, fg_qaqc)
+
+check_them %>%
+  filter(all.equal(sum_macrocover, sum_fg_check) != TRUE)
+
+#pivot wide at quad level
+fg_summary_wide <- fg_summary %>%
+  pivot_wider(names_from = "functional_group", values_from = "sum_fg_cover")
+
+#richness at every quad/year combination 
+fg_summary_wide$functional_richness <- vegan::specnumber(fg_summary_wide[5:14])
+
+unique(fg_summary_wide$functional_richness)
+#0 to 4
+
+##### REPEAT AT SITE LEVEL
+#pivot wide at site level
+fg_summary_site_wide_SUM <- fg_summary_site_SUM %>%
+  pivot_wider(names_from = "functional_group", values_from = "sum_fg_cover")
+
+#richness at every site/year combination ---> NOT AVERAGED,
+fg_summary_site_wide_SUM$functional_richness <- vegan::specnumber(fg_summary_site_wide_SUM[4:13])
+
+#merge Noam's data with functional group summary 
+alpha_diversity_quad_macro <- merge(alpha_diversity_quad_macro,
+                                    fg_summary_wide[,c("location", "year", "site",  "habitat", "functional_richness")],
+                                    by = c("location", "year", "site", "habitat"))
+
+#merge site level cover with site functional group richness
+alpha_diversity_site_macro<-merge(alpha_diversity_site_macro,
+                                  fg_summary_site_wide_SUM[,c("year", "site",  "habitat", "functional_richness")],
+                                  by = c("year", "site",  "habitat"))
+```
+
 ## models
@@
 
-diversity_ranges_quad <- extract_ranges(alpha_diversity_quad_macro, "habitat", c("richness", "cover_trans"))
+diversity_ranges_quad <- extract_ranges(alpha_diversity_quad_macro, "habitat", c("richness", "cover_trans", "functional_richness"))
 
@@
 ### Setup
-```{r read and prep functional group data}
-algae_fg_raw2 <- read_csv(here("data","algae_functional_groupings_092724.csv"))
-
-#clean column names
-algae_fg <- algae_fg_raw2 %>%
-  janitor::clean_names()
-
-#change species names to be lowercase & have underscore 
-algae_fg$species <- gsub(" ", "_", tolower(algae_fg$species))
-algae_fg$morphology <- trimws(algae_fg$morphology)
-algae_fg$morphology <- gsub(" ", "_", algae_fg$morphology)
-
-#rename species column to taxa
-
-algae_fg <- algae_fg %>%
-  select(species, morphology) %>%
-  dplyr::rename(taxa = species) %>%
-   dplyr::rename(functional_group = morphology)
-
-unique(algae_fg$functional_group)
-
-#merge functional groups  
-macro_functional_groups <- merge(macro_long_data, algae_fg, all.x = TRUE)
-
-taxa_sum_check <- macro_functional_groups %>%
-  group_by(location, year) %>%
-  summarise(sum_macrocover = sum(prop_cover))
-
-macro_functional_groups %>%
-  filter(location == "lter_2_backreef_algae_transect_1_quad_1") %>%
-  filter(year == "2022")
-
-#summarizing functional groups at the quad level
-fg_summary <- macro_functional_groups %>%
-  group_by(location, year, functional_group, habitat, site) %>%
-  summarise(sum_fg_cover = sum(prop_cover))
-
-max(fg_summary$sum_fg_cover)
-
-#summarizing functional groups at the site level
-#NOTE that this is ADDED and not averaged. This data is only used to find richness at the site level. LE did this as adding and as a sum, both end up the same from a richness perspective. 
-fg_summary_site_SUM <- macro_functional_groups %>%
-  group_by(site, year, habitat, functional_group) %>%
-  summarise(sum_fg_cover = sum(prop_cover))
-
-#QAQC
-fg_summary_qaqc <- macro_functional_groups %>%
-  group_by(location, year, functional_group) %>%
-  summarise(sum_fg_cover = sum(prop_cover))
-
-fg_qaqc <- fg_summary_qaqc %>%
-  group_by(location, year) %>%
-  summarise(sum_fg_check = sum(sum_fg_cover))
-
-check_them <- merge(taxa_sum_check, fg_qaqc)
-
-check_them %>%
-  filter(all.equal(sum_macrocover, sum_fg_check) != TRUE)
-
-#pivot wide at quad level
-fg_summary_wide <- fg_summary %>%
-  pivot_wider(names_from = "functional_group", values_from = "sum_fg_cover")
-
-#richness at every quad/year combination 
-fg_summary_wide$fg_richness_quad <- vegan::specnumber(fg_summary_wide[5:14])
-
-unique(fg_summary_wide$fg_richness_quad)
-#0 to 4
-
-##### REPEAT AT SITE LEVEL
-#pivot wide at site level
-fg_summary_site_wide_SUM <- fg_summary_site_SUM %>%
-  pivot_wider(names_from = "functional_group", values_from = "sum_fg_cover")
-
-#richness at every site/year combination ---> NOT AVERAGED,
-fg_summary_site_wide_SUM$fg_richness_site <- vegan::specnumber(fg_summary_site_wide_SUM[4:13])
-
-fg_summary_site_wide <- fg_summary_site_wide_SUM
-
-#merge Noam's data with functional group summary 
-functional_richness_quad_cover <- merge(alpha_diversity_quad_macro, fg_summary_wide,
-                                        by = c("location", "year", "site", "habitat"))
-
-#merge site level cover with site functional group richness
-functional_richness_site_cover<- merge(alpha_diversity_site_macro, fg_summary_site_wide,
-                                by = c("year", "site",  "habitat"))
-```
 
@@
 
-cover_mod_fg <- glmmTMB(cover_trans ~ fg_richness_quad*habitat + (1|site/location) + (1|year), family = beta_family(), data = functional_richness_quad_cover)
+cover_mod_fg <- glmmTMB(cover_trans ~ functional_richness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro)
 
@@
 r.squaredGLMM(cover_mod_fg) # 0.64, 0.65
-emtrends(cover_mod_fg, pairwise ~ habitat, var = "fg_richness_quad") 
+emtrends(cover_mod_fg, pairwise ~ habitat, var = "functional_richness") 
 # contrast                    estimate     SE  df z.ratio p.value
@@
 ```{r figure 3b setup}
-diversity_ranges_quad_fg <- functional_richness_quad_cover %>%
-  group_by(habitat) %>%
-  summarise(min_fg_richness_quad = min(fg_richness_quad),
-            max_fg_richness_quad = max(fg_richness_quad),
-            min_cover= min(cover_trans),
-            max_cover = max(cover_trans))
 
 cover_emm_fg <- emmip(cover_mod_fg,
-                        habitat ~ fg_richness_quad,
-                        at = list(fg_richness_quad = seq(0,10,1)), plotit = F, CIs = T)
-
-filtered_cover_effects_fg <- cover_emm_fg %>%
-  left_join(diversity_ranges_quad_fg, by = c("habitat")) %>%
-  filter(fg_richness_quad >= min_fg_richness_quad & fg_richness_quad <= max_fg_richness_quad)
+                        habitat ~ functional_richness,
+                        at = list(functional_richness = seq(0,10,1)), plotit = F, CIs = T)
+
+filtered_cover_effects_fg <- filter_ranges(cover_emm_fg, diversity_ranges_quad, "habitat", "functional_richness")
 
@@
   ggplot() +
-  geom_point(data = functional_richness_quad_cover,
-             aes(x = fg_richness_quad, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
-  geom_line(data = filtered_cover_effects_fg, aes(x = fg_richness_quad, y = inv_logit(yvar), colour = habitat)) +
-  geom_ribbon(data = filtered_cover_effects_fg, aes(x = fg_richness_quad, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
+  geom_point(data = alpha_diversity_quad_macro,
+             aes(x = functional_richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
+  geom_line(data = filtered_cover_effects_fg, aes(x = functional_richness, y = inv_logit(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_cover_effects_fg, aes(x = functional_richness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
               alpha = 0.3) +
@@
             cover_mean = mean(macroalgal_prop_cover),
-            evenness_mean = mean(evenness, na.rm = T),
+            functional_richness_mean = mean(functional_richness),
             cover_cv = CV(macroalgal_prop_cover),
@@
 
-dss_ranges<- diversity_stability_synchrony %>%
-  group_by(habitat) %>%
-  summarise(min_richness = min(richness_mean),
-            max_richness = max(richness_mean),
-            min_evenness = min(richness_mean, na.rm = TRUE),
-            max_evenness = max(richness_mean, na.rm = TRUE),
-            min_stability= min(cover_stability, na.rm = TRUE),
-            max_stability = max(cover_stability, na.rm = TRUE),
-            min_synchrony = min(synchrony_trans, na.rm = TRUE),
-            max_synchrony = max(synchrony_trans, na.rm = TRUE))
+dss_ranges <- extract_ranges(diversity_stability_synchrony, "habitat",
+                             c("richness_mean", "functional_richness_mean", "cover_stability", "synchrony"))
+
 ```
@@
 
-filtered_synch_cov_effects <- synch_cov_emm %>%
-  left_join(dss_ranges, by = c("habitat")) %>%
-  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
+filtered_synch_cov_effects <- filter_ranges(synch_cov_emm, dss_ranges, "habitat", "synchrony")
+
 ```
@@
 ```{r calculate syncrony site_level}
-# make er long
+# pivot longer
 macro_long_data_site <- pivot_longer(
@@
             cover_mean = mean(prop_cover),
-            evenness_mean = mean(evenness),
             cover_cv = CV(prop_cover),
@@
 
-dss_ranges_site <- diversity_stability_synchrony_site %>%
-  group_by(habitat) %>%
-  summarise(min_richness = min(richness_mean),
-            max_richness = max(richness_mean),
-            min_evenness = min(evenness_mean, na.rm = T),
-            max_evenness = max(evenness_mean, na.rm = T),
-            min_stability= min(cover_stability),
-            max_stability = max(cover_stability),
-            min_synchrony = min(synchrony_trans),
-            max_synchrony = max(synchrony_trans))
+dss_ranges_site <- extract_ranges(diversity_stability_synchrony_site, "habitat",
+                                  c("richness_mean", "cover_stability", "synchrony"))
+
 ```
@@
 ```{r site richness cover setup}
-diversity_ranges_site <- alpha_diversity_site_macro %>%
-  group_by(habitat) %>%
-  summarise(min_richness = min(richness),
-            max_richness = max(richness),
-            min_cover= min(cover_trans),
-            max_cover = max(cover_trans))
+diversity_ranges_site <- extract_ranges(alpha_diversity_site_macro, "habitat", c("richness", "functional_richness", "cover_trans"))
 
@@
 
-filtered_site_cover_effects <- site_cover_emm %>%
-  left_join(diversity_ranges_site, by = c("habitat")) %>%
-  filter(richness >= min_richness & richness <= max_richness)
+filtered_site_cover_effects <- filter_ranges(site_cover_emm, diversity_ranges_site, "habitat", "richness")
 
@@
 
-site_cover_mod_fg <- glmmTMB(cover_trans ~ fg_richness_site*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = functional_richness_site_cover)
+site_cover_mod_fg <- glmmTMB(cover_trans ~ functional_richness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro)
 
@@
 r.squaredGLMM(site_cover_mod_fg) # 0.180, 0.563
-emtrends(site_cover_mod_fg, pairwise ~ habitat, var = "fg_richness_site") # no difference between the habitats
+emtrends(site_cover_mod_fg, pairwise ~ habitat, var = "functional_richness") # no difference between the habitats
 # Fringing - Backreef           0.0379 0.0825 Inf   0.459  0.9679
@@
 ```{r site fg richness cover plot setup}
-fg_ranges_site <- functional_richness_site_cover %>%
-  group_by(habitat) %>%
-  summarise(min_fg_richness_site = min(fg_richness_site),
-            max_fg_richness_site = max(fg_richness_site),
-            min_cover= min(cover_trans),
-            max_cover = max(cover_trans))
 
 site_cover_fg_emm <- emmip(site_cover_mod_fg,
-                        habitat ~ fg_richness_site,
-                        at = list(fg_richness_site = seq(0,17,1)), plotit = F, CIs = T)
-
-filtered_site_fg_cover_effects <- site_cover_fg_emm %>%
-  left_join(fg_ranges_site, by = c("habitat")) %>%
-  filter(fg_richness_site >= min_fg_richness_site & fg_richness_site <= max_fg_richness_site)
+                        habitat ~ functional_richness,
+                        at = list(functional_richness = seq(0,17,1)), plotit = F, CIs = T)
+
+filtered_site_fg_cover_effects <- filter_ranges(site_cover_fg_emm, diversity_ranges_site, "habitat", "functional_richness")
+
 ```
@@
   ggplot() +
-  geom_point(data = functional_richness_site_cover,
-             aes(x = fg_richness_site, y = cover_trans, colour = habitat), size = 1, alpha = 0.3) + 
-  geom_line(data = filtered_site_fg_cover_effects, aes(x = fg_richness_site, y = inv_logit(yvar), colour = habitat),
+  geom_point(data = alpha_diversity_site_macro,
+             aes(x = functional_richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.3) + 
+  geom_line(data = filtered_site_fg_cover_effects, aes(x = functional_richness, y = inv_logit(yvar), colour = habitat),
             size = 1.5) +
-  geom_ribbon(data = filtered_site_fg_cover_effects, aes(x = fg_richness_site, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
+  geom_ribbon(data = filtered_site_fg_cover_effects, aes(x = functional_richness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
               alpha = 0.3) +
@@
 
-filtered_synch_cov_effects_site <- synch_cov_emm_site %>%
-  left_join(dss_ranges_site, by = c("habitat")) %>%
-  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
+filtered_synch_cov_effects_site <- filter_ranges(synch_cov_emm_site, dss_ranges_site, "habitat", "synchrony")
 ```

### 2024-10-25 manuscript_1_code.Rmd
cleaned code up to figure 5. Moved figure 2 to supplements. moved diversity-synhcrony plot to supplements.

@@
 
-## Figure 2a Richness @ Quad by Year
-
-### prep
-```{r richness df}
-quad_richness_df <- alpha_diversity_quad_macro %>% 
-  group_by(year, habitat) %>% 
-  summarise(richness_mean = mean(richness),
-            richness_se = se(richness))
-```
-
-### plot
-```{r figure 2a}
-(figure_2a <- ggplot() +
-  geom_point(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4,
-             position = position_dodge(0.4)) +
-  geom_line(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
-  geom_errorbar(data = quad_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se,
-                                             colour = habitat), width = 0, size = 1.3, position = position_dodge(0.4)) +
-  scale_colour_manual(values = habitat_colours) +
-  annotate("segment", x = 2006, xend = 2010, y = 2, yend = 2, linewidth = 2, colour = "grey") +
-  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
-  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
-   scale_x_continuous(breaks = unique(quad_richness_df$year)) +
-   labs(y = "Macroalgal Richness",
-        x = "Year",
-        title = "") +
-   trends_theme)
-
-# ggsave(filename = "output/figure_2a.png", figure_2a, height = 10, width = 14)
-```
-
-## Figure 2b Richness @ Site by Year
-
-### prep
-
-```{r site level richness df}
-site_richness_df <- alpha_diversity_site_macro %>% 
-  group_by(year, habitat) %>% 
-  summarise(richness_mean = mean(richness),
-            richness_se = se(richness))
-
-```
-
-### plot
-```{r plot figure 2b}
-(figure_2b <- ggplot() +
-  geom_point(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4,
-             position = position_dodge(0.4)) +
-#  geom_jitter(data = alpha_diversity_site_macro, 
-#  aes(x = year, y = richness, colour = habitat, alpha = 0.05), size = 2, width = 0.4, height = 0) +
-  geom_line(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
-  geom_errorbar(data = site_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, colour = habitat), size = 1.3, width = 0,  position = position_dodge(0.4)) +
-  scale_colour_manual(values = habitat_colours) +
-  annotate("segment", x = 2006, xend = 2010, y = 12, yend = 12, linewidth = 2, colour = "grey") +
-  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
-  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
-   scale_x_continuous(breaks = unique(site_richness_df$year)) +
-   labs(y = "",
-        x = "Year",
-        title = "") +
-  trends_theme)
-
-# ggsave(filename = "output/figure_2b.png", figure_2b, height = 10, width = 14)
-```
-
-## Figure 2 Full Plot
-```{r ggpubr fig2, fig.height = 10, fig.width=22}
-fig2_plot_list <- list(figure_2a, figure_2b)
-(figure_2 <- ggarrange(figure_2a, figure_2b,
-                      ncol = 2, nrow = 1, 
-                      labels = c("a.", "b."),
-                      common.legend = TRUE,
-                      legend = "bottom", 
-                      font.label = list(size = 24, color = "black", face = "plain")))
-
-# ggsave(filename = "output/figure_2v4.png", figure_2, height = 10, width = 22)
-```
-
-# Figure 3 - ecosystem function
+# prep functional group data
 
@@
             richness_mean = mean(richness), 
-            shannon_mean = mean(shannon), 
+            functional_richness_mean = mean(functional_richness), 
             cover_mean = mean(prop_cover),
@@
 dss_ranges_site <- extract_ranges(diversity_stability_synchrony_site, "habitat",
-                                  c("richness_mean", "cover_stability", "synchrony"))
+                                  c("richness_mean", "functional_richness_mean", "cover_stability", "synchrony"))
 
@@
 
-filtered_rich_stab_effects <- rich_stab_emm %>%
-  left_join(dss_ranges, by = c("habitat")) %>%
-  filter(richness_mean >= min_richness & richness_mean <= max_richness)
+filtered_rich_stab_effects <- filter_ranges(rich_stab_emm, dss_ranges, "habitat", "richness_mean")
 ```
@@
 
-#### no habitat
-
-```{r richness stability no hab}
-# rich_stab_mod_2 <- glmmTMB(cover_stability~ richness_mean + (1|site), family = Gamma(link = "log"), data = # diversity_stability_synchrony)
-# summary(rich_stab_mod_2)
-# hist(residuals(rich_stab_mod_2)) # a chunkier version
-# plot(residuals(rich_stab_mod_2) ~ fitted(rich_stab_mod_2)) # heteroskedastic - bring to group
-# r.squaredGLMM(rich_stab_mod_2) # 0.750, 0.757
-
-```
-
-```{r 4a no hab setup}
-# rich_stab_emm_2 <-as.data.frame(emmeans(rich_stab_mod_2, ~richness_mean,
-#                                      at = list(richness_mean = seq(min(diversity_stability_synchrony$richness_mean),
-#                                                             max(diversity_stability_synchrony$richness_mean),
-#                                                             0.01)))) 
-# 
-```
-
-#### no habitat plot
-
-```{r figure 4a no habitat}
-# (figure_4a_no_hab <- 
-#   ggplot() +
-#   geom_point(data = diversity_stability_synchrony,
-#              aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
-#   geom_line(data = filtered_rich_stab_effects, aes(x = richness_mean, y = exp(yvar), colour = habitat)) +
-#   geom_ribbon(data = filtered_rich_stab_effects, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
-#               alpha = 0.3) +
-#   geom_line(data = rich_stab_emm_2, aes(x = richness_mean, y = exp(emmean)), size = 2, colour = "black") +
-#   geom_ribbon(data = rich_stab_emm_2, aes(x = richness_mean, ymin = exp(asymp.LCL), ymax = exp(asymp.UCL)), fill = "black",
-#               alpha = 0.3) +
-#   scale_colour_manual(values = habitat_colours) +
-#   scale_fill_manual(values = habitat_colours) +
-#   labs(y = "Stability\n", x = "Macroalgal Richness") +
-#   theme_bw() + 
-#   theme(axis.title.x = element_text(size=18), 
-#         axis.title.y = element_text(size=18), 
-#         panel.background = element_blank(), 
-#         panel.grid.major = element_blank(),  #remove major-grid labels
-#         panel.grid.minor = element_blank(),  #remove minor-grid labels
-#         plot.background = element_rect(fill = "white"),
-#         legend.text = element_text(size = rel(1)), # increase legend text
-#         legend.title = element_blank(),
-#         legend.position = 'bottom', # move legend to bottom
-#         legend.box = 'horizontal', # make legend horizontal
-#         legend.box.background = element_rect(colour = "black"))  # add box around legend
-#          )
-# 
-# ggsave(filename = "output/figure_4a_with_no_hab.png", figure_4a_no_hab, height = 10, width = 14)
-```
-
 ## Figure 4b - functional group richness stability
-### Setup
-```{r functional richness through time at quad level}
-fg_summary_wide_meta <- merge(fg_summary_wide, meta_df, by = c("location", "site", "habitat", "year") )
-fg_summary_wide_meta_through_time <- fg_summary_wide_meta %>%
-  group_by(location, habitat, site) %>%
-  summarise(fg_richness_mean = mean(fg_richness_quad),
-            fg_richness_se = sd(fg_richness_quad)/n())
-
-#merge with fg data 
-fg_stability_quad_level <- merge(fg_summary_wide_meta_through_time, diversity_stability_synchrony, 
-                              by = 
-                                c("location","site", "habitat"))
-
-dss_ranges_fg <- fg_stability_quad_level %>%
-  group_by(habitat) %>%
-  summarise(min_richness_fg = min(fg_richness_mean),
-            max_richness_fg = max(fg_richness_mean),
-            min_stability= min(cover_stability, na.rm = TRUE),
-            max_stability = max(cover_stability, na.rm = TRUE))
-```
-
-### figure 4b model
+
+### model
 ```{r fig4b fg richness stability mod}
 
-rich_stab_mod_fg <- glmmTMB(cover_stability ~ fg_richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = fg_stability_quad_level)
+rich_stab_mod_fg <- glmmTMB(cover_stability ~ functional_richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = diversity_stability_synchrony)
 
@@
 r.squaredGLMM(rich_stab_mod_fg) # 0.819, 0.828
-emtrends(rich_stab_mod_fg, pairwise ~ habitat, var = "fg_richness_mean") 
+emtrends(rich_stab_mod_fg, pairwise ~ habitat, var = "functional_richness_mean") 
 #  fringing different from all others
@@
 rich_stab_emm_fg <- emmip(rich_stab_mod_fg,
-                        habitat ~ fg_richness_mean,
-                        at = list(fg_richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)
-
-filtered_rich_stab_effects_fg <- rich_stab_emm_fg %>%
-  left_join(dss_ranges_fg, by = c("habitat")) %>%
-  filter(fg_richness_mean >= min_richness_fg & fg_richness_mean <= max_richness_fg)
+                        habitat ~ functional_richness_mean,
+                        at = list(functional_richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_rich_stab_effects_fg <- filter_ranges(rich_stab_emm_fg, dss_ranges, "habitat", "functional_richness_mean")
 ```
@@
   ggplot() +
-  geom_point(data = fg_stability_quad_level,
-             aes(x = fg_richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_rich_stab_effects_fg, aes(x = fg_richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
-  geom_ribbon(data = filtered_rich_stab_effects_fg, aes(x = fg_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+  geom_point(data = diversity_stability_synchrony,
+             aes(x = functional_richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_rich_stab_effects_fg, aes(x = functional_richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
+  geom_ribbon(data = filtered_rich_stab_effects_fg,
+              aes(x = functional_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
               alpha = 0.3) +
@@
 
-filtered_synch_stab_effects <- synch_stab_emm %>%
-  left_join(dss_ranges, by = c("habitat")) %>%
-  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
+filtered_synch_stab_effects <- filter_ranges(synch_stab_emm, dss_ranges, "habitat", "synchrony")
 ```
@@
 
-filtered_rich_stab_effects_site <- rich_stab_emm_site %>%
-  left_join(dss_ranges_site, by = c("habitat")) %>%
-  filter(richness_mean >= min_richness & richness_mean <= max_richness)
+filtered_rich_stab_effects_site <- filter_ranges(rich_stab_emm_site, dss_ranges_site, "habitat", "richness_mean")
 ```
@@
 
-#need to get average functional richness through time! 
-fg_summary_site_wide_summary <- fg_summary_site_wide %>%
-  group_by(habitat, site) %>%
-  summarize(avg_richness_fg = mean(fg_richness_site),
-            se_richness_fg = sd(fg_richness_site)/sqrt(n()))
-
-#n = 24!
-
-# merge
-stability_site_level_fg <- merge(diversity_stability_synchrony_site, fg_summary_site_wide_summary, by = c("site", "habitat"))
-
-# extract relevant ranges for plotting
-dss_ranges_site_fg_stab <- stability_site_level_fg %>%
-  group_by(habitat) %>%
-  summarise(min_richness_fg = min(avg_richness_fg),
-            max_richness_fg = max(avg_richness_fg),
-            min_stability= min(cover_stability),
-            max_stability = max(cover_stability))
-
-rich_stab_mod_site_fg <- glmmTMB(cover_stability~ avg_richness_fg*habitat, family = Gamma(link = "log"), data = stability_site_level_fg)
+rich_stab_mod_site_fg <- glmmTMB(cover_stability~ functional_richness_mean*habitat, family = Gamma(link = "log"), data = diversity_stability_site)
 summary(rich_stab_mod_site_fg)
@@
 r.squaredGLMM(rich_stab_mod_site_fg) # 0.563
-emtrends(rich_stab_mod_site_fg, pairwise ~ habitat, var = "avg_richness_fg") 
+emtrends(rich_stab_mod_site_fg, pairwise ~ habitat, var = "functional_richness_mean") 
 # Fringing different from outer 10
@@
 rich_stab_emm_site_fg <- emmip(rich_stab_mod_site_fg,
-                        habitat ~ avg_richness_fg,
-                        at = list(avg_richness_fg = seq(0,10,0.01)), plotit = F, CIs = T)
-
-filtered_rich_stab_effects_site <- rich_stab_emm_site_fg %>%
-  left_join(dss_ranges_site_fg_stab, by = c("habitat")) %>%
-  filter((avg_richness_fg >= min_richness_fg) & (avg_richness_fg <= max_richness_fg))
+                        habitat ~ functional_richness_mean,
+                        at = list(functional_richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_rich_stab_effects_site <- filter_ranges(rich_stab_emm_site_fg, dss_ranges_site, "habitat", "functional_richness_mean")
 
@@
   ggplot() +
-  geom_point(data = stability_site_level_fg,
-             aes(x = avg_richness_fg, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_rich_stab_effects_site, aes(x = avg_richness_fg, y = exp(yvar), colour = habitat), size = 1.5) +
-  geom_ribbon(data = filtered_rich_stab_effects_site, aes(x = avg_richness_fg, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+  geom_point(data = diversity_stability_site,
+             aes(x = functional_richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_rich_stab_effects_site, aes(x = functional_richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
+  geom_ribbon(data = filtered_rich_stab_effects_site, 
+              aes(x = functional_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
               alpha = 0.3) +
@@
 
-filtered_synch_stab_effects_site <- synch_stab_emm_site %>%
-  left_join(dss_ranges_site, by = c("habitat")) %>%
-  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
+filtered_synch_stab_effects_site <- filter_ranges(synch_stab_emm_site, dss_ranges_site, "habitat", "synchrony")
 ```
@@
 
-## diversity-synchrony plot level
+## create new dfs
+```{r mean and spatial synchrony}
+# calculate spatial synchrony - degree to which quadrats vary together within a site
+spatial_synchrony <- codyn::synchrony(df = alpha_diversity_quad_macro,
+                                   time.var = "year",
+                                   species.var = "location",
+                                   abundance.var = "macroalgal_prop_cover",
+                                   metric = "Loreau",
+                                   replicate.var = "site_habitat")
+colnames(spatial_synchrony) <- c("site_habitat", "spatial_synchrony") # rename to avoid overwrite
+
+# calculate mean synchrony within a site (!= synchrony of species at the site level!)
+mean_synchrony_stability <- diversity_stability_synchrony %>% 
+  group_by(habitat, site) %>% 
+  summarise(synchrony_mean = mean(synchrony, na.rm = T),
+            stability_mean = mean(cover_stability, na.rm = T))
+
+# merge with site level dss metrics
+dss_spatial <- merge(diversity_stability_site, spatial_synchrony, by = "site_habitat")
+dss_spatial_2 <- merge(dss_spatial, mean_synchrony_stability, by = c("habitat", "site"))
+
+dss_spatial_ranges <- extract_ranges(dss_spatial_2, "habitat", c("spatial_synchrony", "synchrony_mean", "stability_mean"))
+
+```
+
+## models
+
+### spatial synchrony
+```{r model spatial synchrony}
+spatial_synchrony_mod <- glmmTMB(cover_stability ~ spatial_synchrony*habitat, family = Gamma("log"), data = dss_spatial_2)
+hist(residuals(spatial_synchrony_mod)) # a little skew
+plot(residuals(spatial_synchrony_mod) ~ predict(spatial_synchrony_mod))
+summary(spatial_synchrony_mod)
+Anova(spatial_synchrony_mod)
+#                             Chisq Df Pr(>Chisq)    
+# spatial_synchrony         16.2914  1  5.431e-05 ***
+# habitat                   19.5303  3  0.0002124 ***
+# spatial_synchrony:habitat  6.7239  3  0.0812368 .  
+emtrends(spatial_synchrony_mod, pairwise ~ habitat, var = "spatial_synchrony") # no sig differences but close
+# contrast                    estimate    SE  df z.ratio p.value
+# Fringing - Backreef            0.404 1.765 Inf   0.229  0.9958
+# Fringing - Forereef 10m        1.365 1.754 Inf   0.778  0.8643
+# Fringing - Forereef 17m        3.112 1.882 Inf   1.654  0.3484
+# Backreef - Forereef 10m        0.961 0.857 Inf   1.121  0.6766
+# Backreef - Forereef 17m        2.708 1.095 Inf   2.472  0.0644
+# Forereef 10m - Forereef 17m    1.747 1.078 Inf   1.621  0.3668
+```
+
+```{r plot spatial synchrony}
+spat_stab_emm <- emmip(spatial_synchrony_mod,
+                        habitat ~ spatial_synchrony,
+                        at = list(spatial_synchrony = seq(0,1,0.01)), plotit = F, CIs = T)
+
+filtered_spat_stab_effects <- filter_ranges(spat_stab_emm, dss_spatial_ranges, "habitat", "spatial_synchrony")
+
+(spatial_synchrony_plot <- 
+  ggplot() +
+  geom_point(data = dss_spatial_2,
+             aes(x = spatial_synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
+  geom_ribbon(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Stability\n", x = "Spatial synchrony", title = "") +
+  trends_theme
+         )
+
+```
+
+### mean synchrony
+```{r mean synchrony}
+mean_synchrony_mod <- glmmTMB(cover_stability ~ synchrony_mean*habitat, family = Gamma("log"), data = dss_spatial_2)
+hist(residuals(mean_synchrony_mod)) # chunky
+plot(residuals(mean_synchrony_mod) ~ predict(mean_synchrony_mod)) # fine
+summary(mean_synchrony_mod)
+Anova(mean_synchrony_mod)
+#                         Chisq Df Pr(>Chisq)    
+# synchrony_mean         32.914  1  9.634e-09 ***
+# habitat                18.523  3  0.0003430 ***
+# synchrony_mean:habitat 19.062  3  0.0002654 ***
+emtrends(mean_synchrony_mod, pairwise ~ habitat, var = "synchrony_mean") # no sig differences but close
+# contrast                    estimate    SE  df z.ratio p.value
+# Fringing - Backreef            -3.26 1.503 Inf  -2.166  0.1327
+# Fringing - Forereef 10m         2.22 0.995 Inf   2.231  0.1149
+# Fringing - Forereef 17m         3.37 1.220 Inf   2.764  0.0291
+# Backreef - Forereef 10m         5.48 1.622 Inf   3.377  0.0041
+# Backreef - Forereef 17m         6.63 1.769 Inf   3.748  0.0010
+# Forereef 10m - Forereef 17m     1.15 1.363 Inf   0.846  0.8322
+```
+
+```{r plot mean synchrony}
+mean_synch_emm <- emmip(mean_synchrony_mod,
+                        habitat ~ synchrony_mean,
+                        at = list(synchrony_mean = seq(0,1,0.01)), plotit = F, CIs = T)
+
+filtered_mean_synch_effects <- filter_ranges(mean_synch_emm, dss_spatial_ranges, "habitat", "synchrony_mean")
+
+(mean_synchrony_plot <- 
+  ggplot() +
+  geom_point(data = dss_spatial_2,
+             aes(x = synchrony_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_mean_synch_effects, aes(x = synchrony_mean, y = exp(yvar), colour = habitat), size = 1.5) +
+  geom_ribbon(data = filtered_mean_synch_effects, aes(x = synchrony_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Stability\n", x = "Mean synchrony", title = "") +
+  trends_theme
+         )
+```
+
+### mean stability plot
+```{r alpha gamma stability mod plot}
+alpha_gamma_stab_mod <- glmmTMB(cover_stability ~ stability_mean*habitat, family = Gamma("log"), data = dss_spatial_2)
+summary(alpha_gamma_stab_mod)
+Anova(alpha_gamma_stab_mod)
+#                          Chisq Df Pr(>Chisq)    
+# stability_mean         61.2725  1   4.97e-15 ***
+# habitat                12.9323  3   0.004785 ** 
+# stability_mean:habitat  9.4832  3   0.023511 *
+
+emtrends(alpha_gamma_stab_mod, pairwise ~ habitat, var = "stability_mean") # backreef different from fringe and forereef 10
+# Fringing - Backreef            1.078 0.381 Inf   2.828  0.0242
+# Fringing - Forereef 10m        0.175 0.345 Inf   0.509  0.9571
+# Fringing - Forereef 17m        0.380 0.305 Inf   1.248  0.5965
+# Backreef - Forereef 10m       -0.903 0.349 Inf  -2.589  0.0475
+# Backreef - Forereef 17m       -0.698 0.309 Inf  -2.256  0.1085
+# Forereef 10m - Forereef 17m    0.205 0.263 Inf   0.779  0.8638
+
+alpha_gamma_emm <- emmip(alpha_gamma_stab_mod,
+                        habitat ~ stability_mean,
+                        at = list(stability_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_ag_effects <- filter_ranges(alpha_gamma_emm, dss_spatial_ranges, "habitat", "stability_mean")
+
+(alpha_gamma_stability_plot <- 
+  ggplot() +
+  geom_point(data = dss_spatial_2,
+             aes(x = stability_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_ag_effects, aes(x = stability_mean, y = exp(yvar), colour = habitat), size = 1.5) +
+  geom_ribbon(data = filtered_ag_effects, aes(x = stability_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Site stability\n", x = "Plot stability", title = "") +
+  trends_theme
+         )
+```
+
+## panel graph
+```{r figure 5 panel, fig.height = 6, fig.width=22}
+(figure_5 <- ggarrange(mean_synchrony_plot, alpha_gamma_stability_plot, spatial_synchrony_plot,
+                      ncol = 3, nrow = 1, 
+                      common.legend = TRUE,
+                      legend = "bottom", 
+                      labels = c("a.", "b.", "c."),
+                      font.label = list(size = 26, color = "black", face = "plain")))
+
+# ggsave(filename = "output/figure_5_option3.png", figure_5, height = 6, width = 22)
+```
+
+# SUPPLEMENTAL FIGURE CODE:
+
+## Trends in richness over time
+
+### prep
+```{r richness df}
+quad_richness_df <- alpha_diversity_quad_macro %>% 
+  group_by(year, habitat) %>% 
+  summarise(richness_mean = mean(richness),
+            richness_se = se(richness))
+```
+
+### plot
+```{r figure 2a}
+(figure_s1a <- ggplot() +
+  geom_point(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4,
+             position = position_dodge(0.4)) +
+  geom_line(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
+  geom_errorbar(data = quad_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se,
+                                             colour = habitat), width = 0, size = 1.3, position = position_dodge(0.4)) +
+  scale_colour_manual(values = habitat_colours) +
+  annotate("segment", x = 2006, xend = 2010, y = 2, yend = 2, linewidth = 2, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+   scale_x_continuous(breaks = unique(quad_richness_df$year)) +
+   labs(y = "Macroalgal Richness",
+        x = "Year",
+        title = "") +
+   trends_theme)
+
+# ggsave(filename = "output/figure_s1a.png", figure_s1a, height = 10, width = 14)
+```
+
+## Figure 2b Richness @ Site by Year
+
+### prep
+
+```{r site level richness df}
+site_richness_df <- alpha_diversity_site_macro %>% 
+  group_by(year, habitat) %>% 
+  summarise(richness_mean = mean(richness),
+            richness_se = se(richness))
+
+```
+
+### plot
+```{r plot figure 2b}
+(figure_s1b <- ggplot() +
+  geom_point(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4,
+             position = position_dodge(0.4)) +
+#  geom_jitter(data = alpha_diversity_site_macro, 
+#  aes(x = year, y = richness, colour = habitat, alpha = 0.05), size = 2, width = 0.4, height = 0) +
+  geom_line(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
+  geom_errorbar(data = site_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, colour = habitat), size = 1.3, width = 0,  position = position_dodge(0.4)) +
+  scale_colour_manual(values = habitat_colours) +
+  annotate("segment", x = 2006, xend = 2010, y = 12, yend = 12, linewidth = 2, colour = "grey") +
+  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
+  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
+   scale_x_continuous(breaks = unique(site_richness_df$year)) +
+   labs(y = "",
+        x = "Year",
+        title = "") +
+  trends_theme)
+
+# ggsave(filename = "output/figure_s1b.png", figure_s1b, height = 10, width = 14)
+```
+
+## Figure 2 Full Plot
+```{r ggpubr fig2, fig.height = 10, fig.width=22}
+fig2_plot_list <- list(figure_s1a, figure_s1b)
+(figure_s1 <- ggarrange(figure_s1a, figure_s1b,
+                      ncol = 2, nrow = 1, 
+                      labels = c("a.", "b."),
+                      common.legend = TRUE,
+                      legend = "bottom", 
+                      font.label = list(size = 24, color = "black", face = "plain")))
+
+# ggsave(filename = "output/figure_s1v4.png", figure_s1, height = 10, width = 22)
+```
+
+## diversity-synchrony relationships: plot level
 
@@
 ```
-
-## create new dfs
-```{r mean and spatial synchrony}
-# calculate spatial synchrony - degree to which quadrats vary together within a site
-spatial_synchrony <- codyn::synchrony(df = alpha_diversity_quad_macro,
-                                   time.var = "year",
-                                   species.var = "location",
-                                   abundance.var = "macroalgal_prop_cover",
-                                   metric = "Loreau",
-                                   replicate.var = "site_habitat")
-colnames(spatial_synchrony) <- c("site_habitat", "spatial_synchrony") # rename to avoid overwrite
-
-# calculate mean synchrony within a site (!= synchrony of species at the site level!)
-mean_synchrony_stability <- diversity_stability_synchrony %>% 
-  group_by(habitat, site) %>% 
-  summarise(synchrony_mean = mean(synchrony, na.rm = T),
-            stability_mean = mean(cover_stability, na.rm = T))
-
-# merge with site level dss metrics
-dss_spatial <- merge(diversity_stability_site, spatial_synchrony, by = "site_habitat")
-dss_spatial_2 <- merge(dss_spatial, mean_synchrony_stability, by = c("habitat", "site"))
-
-dss_spatial_ranges <- dss_spatial_2 %>%
-  group_by(habitat) %>%
-  summarise(min_spatial_synchrony = min(spatial_synchrony, na.rm = TRUE),
-            max_spatial_synchrony = max(spatial_synchrony, na.rm = TRUE),
-            min_mean_synchrony= min(synchrony_mean, na.rm = TRUE),
-            max_mean_synchrony = max(synchrony_mean, na.rm  = TRUE),
-            min_mean_stability = min(stability_mean, na.rm = TRUE),
-            max_mean_stability = max(stability_mean, na.rm  = TRUE))
-
-```
-
-## models
-```{r model spatial synchrony}
-spatial_synchrony_mod <- glmmTMB(cover_stability ~ spatial_synchrony*habitat, family = Gamma("log"), data = dss_spatial_2)
-hist(residuals(spatial_synchrony_mod)) # a little skew
-plot(residuals(spatial_synchrony_mod) ~ predict(spatial_synchrony_mod))
-summary(spatial_synchrony_mod)
-Anova(spatial_synchrony_mod)
-#                             Chisq Df Pr(>Chisq)    
-# spatial_synchrony         16.2914  1  5.431e-05 ***
-# habitat                   19.5303  3  0.0002124 ***
-# spatial_synchrony:habitat  6.7239  3  0.0812368 .  
-emtrends(spatial_synchrony_mod, pairwise ~ habitat, var = "spatial_synchrony") # no sig differences but close
-# contrast                    estimate    SE  df z.ratio p.value
-# Fringing - Backreef            0.404 1.765 Inf   0.229  0.9958
-# Fringing - Forereef 10m        1.365 1.754 Inf   0.778  0.8643
-# Fringing - Forereef 17m        3.112 1.882 Inf   1.654  0.3484
-# Backreef - Forereef 10m        0.961 0.857 Inf   1.121  0.6766
-# Backreef - Forereef 17m        2.708 1.095 Inf   2.472  0.0644
-# Forereef 10m - Forereef 17m    1.747 1.078 Inf   1.621  0.3668
-```
-
-```{r plot spatial synchrony}
-spat_stab_emm <- emmip(spatial_synchrony_mod,
-                        habitat ~ spatial_synchrony,
-                        at = list(spatial_synchrony = seq(0,1,0.01)), plotit = F, CIs = T)
-
-filtered_spat_stab_effects <- spat_stab_emm %>%
-  left_join(dss_spatial_ranges, by = c("habitat")) %>%
-  filter(spatial_synchrony >= min_spatial_synchrony & 
-           spatial_synchrony <= max_spatial_synchrony)
-
-(figure_5a <- 
-  ggplot() +
-  geom_point(data = dss_spatial_2,
-             aes(x = spatial_synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
-  geom_ribbon(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
-              alpha = 0.3) +
-  scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
-  labs(y = "Stability\n", x = "Spatial synchrony", title = "") +
-  trends_theme
-         )
-
-```
-
-```{r mean synchrony}
-mean_synchrony_mod <- glmmTMB(cover_stability ~ synchrony_mean*habitat, family = Gamma("log"), data = dss_spatial_2)
-hist(residuals(mean_synchrony_mod)) # chunky
-plot(residuals(mean_synchrony_mod) ~ predict(mean_synchrony_mod)) # fine
-summary(mean_synchrony_mod)
-Anova(mean_synchrony_mod)
-#                         Chisq Df Pr(>Chisq)    
-# synchrony_mean         32.914  1  9.634e-09 ***
-# habitat                18.523  3  0.0003430 ***
-# synchrony_mean:habitat 19.062  3  0.0002654 ***
-emtrends(mean_synchrony_mod, pairwise ~ habitat, var = "synchrony_mean") # no sig differences but close
-# contrast                    estimate    SE  df z.ratio p.value
-# Fringing - Backreef            -3.26 1.503 Inf  -2.166  0.1327
-# Fringing - Forereef 10m         2.22 0.995 Inf   2.231  0.1149
-# Fringing - Forereef 17m         3.37 1.220 Inf   2.764  0.0291
-# Backreef - Forereef 10m         5.48 1.622 Inf   3.377  0.0041
-# Backreef - Forereef 17m         6.63 1.769 Inf   3.748  0.0010
-# Forereef 10m - Forereef 17m     1.15 1.363 Inf   0.846  0.8322
-```
-
-```{r plot mean synchrony}
-mean_synch_emm <- emmip(mean_synchrony_mod,
-                        habitat ~ synchrony_mean,
-                        at = list(synchrony_mean = seq(0,1,0.01)), plotit = F, CIs = T)
-
-filtered_mean_synch_effects <- mean_synch_emm %>%
-  left_join(dss_spatial_ranges, by = c("habitat")) %>%
-  filter(synchrony_mean >= min_mean_synchrony & 
-           synchrony_mean <= max_mean_synchrony)
-
-(figure_5b <- 
-  ggplot() +
-  geom_point(data = dss_spatial_2,
-             aes(x = synchrony_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_mean_synch_effects, aes(x = synchrony_mean, y = exp(yvar), colour = habitat), size = 1.5) +
-  geom_ribbon(data = filtered_mean_synch_effects, aes(x = synchrony_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
-              alpha = 0.3) +
-  scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
-  labs(y = "Stability\n", x = "Mean synchrony", title = "") +
-  trends_theme
-         )
-```
-
-```{r alpha gamma stability mod plot}
-alpha_gamma_stab_mod <- glmmTMB(cover_stability ~ stability_mean*habitat, family = Gamma("log"), data = dss_spatial_2)
-summary(alpha_gamma_stab_mod)
-Anova(alpha_gamma_stab_mod)
-#                          Chisq Df Pr(>Chisq)    
-# stability_mean         61.2725  1   4.97e-15 ***
-# habitat                12.9323  3   0.004785 ** 
-# stability_mean:habitat  9.4832  3   0.023511 *
-
-emtrends(alpha_gamma_stab_mod, pairwise ~ habitat, var = "stability_mean") # backreef different from fringe and forereef 10
-# Fringing - Backreef            1.078 0.381 Inf   2.828  0.0242
-# Fringing - Forereef 10m        0.175 0.345 Inf   0.509  0.9571
-# Fringing - Forereef 17m        0.380 0.305 Inf   1.248  0.5965
-# Backreef - Forereef 10m       -0.903 0.349 Inf  -2.589  0.0475
-# Backreef - Forereef 17m       -0.698 0.309 Inf  -2.256  0.1085
-# Forereef 10m - Forereef 17m    0.205 0.263 Inf   0.779  0.8638
-
-alpha_gamma_emm <- emmip(alpha_gamma_stab_mod,
-                        habitat ~ stability_mean,
-                        at = list(stability_mean = seq(0,10,0.01)), plotit = F, CIs = T)
-
-filtered_ag_effects <- alpha_gamma_emm %>%
-  left_join(dss_spatial_ranges, by = c("habitat")) %>%
-  filter(stability_mean >= min_mean_stability & 
-           stability_mean <= max_mean_stability)
-
-(alpha_gamma_stability_plot <- 
-  ggplot() +
-  geom_point(data = dss_spatial_2,
-             aes(x = stability_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
-  geom_line(data = filtered_ag_effects, aes(x = stability_mean, y = exp(yvar), colour = habitat), size = 1.5) +
-  geom_ribbon(data = filtered_ag_effects, aes(x = stability_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
-              alpha = 0.3) +
-  scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
-  labs(y = "Site stability\n", x = "Plot stability", title = "") +
-  trends_theme
-         )
-```
-
-## panel graph
-```{r figure 5 panel, fig.height = 6, fig.width=22}
-(figure_5 <- ggarrange(figure_5b, figure_5a, alpha_gamma_stability_plot, 
-                      ncol = 3, nrow = 1, 
-                      common.legend = TRUE,
-                      legend = "bottom", 
-                      labels = c("a.", "b.", "c."),
-                      font.label = list(size = 26, color = "black", face = "plain")))
-
-# ggsave(filename = "output/figure_5_option3.png", figure_5, height = 6, width = 22)
-```
-
-# SUPPLEMENTAL FIGURE CODE:

### 2024-11-20 Personal_Prelim_Code/mcr_algal_data_cleaning_final.Rmd

@@
 # Caulerpa peltata is outdated --> Caulerpa racemosa
-data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Caulerpa peltata"] <- "Caulerpa racemosa"
+data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Caulerpa peltata"] <- "Caulerpa chemnitzia"
 

### 2024-11-20 Personal_Prelim_Code/noam_data_exploration_v3.Rmd

--- 2024-08-23 noam_data_exploration_v3.Rmd
@@
 ```{r import and format data}
+# coefficient of variation
+CV<-function(x){
+  return(sd(x,na.rm=T)/mean(x,na.rm=T))
+}
+
 # starting with all data now because of 0 algae transects/quadrats
@@
 
-meta_cols <- c(colnames(data[1:7] ), "island_side", "island_side_habitat", "stratum")
+meta_cols <- c("year", "location", "site", "habitat", "site_habitat", "transect", "quadrat", "island_side", "island_side_habitat", "stratum")
 species_cols <- colnames(data)[!colnames(data) %in% meta_cols]
@@
   data = data,
-  cols = 8:77,
+  cols = 8:76,
   names_to = "taxa",
@@
   # fit gls for current time series
-  loop_mod <- gls(logRich ~ year, correlation =corAR1(form=~year), 
+  loop_mod <- gls(richness ~ year, correlation =corAR1(form=~year), 
                   control=glsControl(msVerbose=T), data = newData)
@@
   theme_classic() +
-  labs(x = "Year", y = "Estimated marginal mean richness\n") +
+  labs(x = "Year", y = "Richness\n") +
   theme(panel.grid.major = element_blank(), 
@@
 # Julianna's model structure
-j_rich_mod <- glmmTMB(richness ~ year + habitat + (1|site/transect),
-               family = poisson,
-               data = alpha_diversity_quad)
-summary(j_rich_mod)
-plot(allEffects(j_rich_mod))
-
-grid <- expand_grid(year = unique(alpha_diversity_quad$year),
-                       habitat = unique(alpha_diversity_quad$habitat),
-                       site = unique(alpha_diversity_quad$site),
-                       island_side = unique(alpha_diversity_quad$island_side),
-                       transect = unique(alpha_diversity_quad$transect))
-grid$Preds <- predict(j_rich_mod, newdata = grid)
-
-grid %>% 
-  # get transect means
-  group_by(year, habitat, island_side) %>% 
-  summarize(Pred_mean_richness = mean(Preds),
-            SD_richness = sd(Preds)) %>% 
-  ggplot(aes(x = year, y = Pred_mean_richness, group = habitat, color = habitat)) +
-  geom_line() +
-  theme_bw() +
-  facet_wrap(~island_side)
+# j_rich_mod <- glmmTMB(richness ~ year + habitat + (1|site/transect),
+#                family = poisson,
+#                data = alpha_diversity_quad)
+# summary(j_rich_mod)
+# plot(allEffects(j_rich_mod))
+# 
+# grid <- expand_grid(year = unique(alpha_diversity_quad$year),
+#                        habitat = unique(alpha_diversity_quad$habitat),
+#                        site = unique(alpha_diversity_quad$site),
+#                        island_side = unique(alpha_diversity_quad$island_side),
+#                        transect = unique(alpha_diversity_quad$transect))
+# grid$Preds <- predict(j_rich_mod, newdata = grid)
+# 
+# grid %>% 
+#   # get transect means
+#   group_by(year, habitat, island_side) %>% 
+#   summarize(Pred_mean_richness = mean(Preds),
+#             SD_richness = sd(Preds)) %>% 
+#   ggplot(aes(x = year, y = Pred_mean_richness, group = habitat, color = habitat)) +
+#   geom_line() +
+#   theme_bw() +
+#   facet_wrap(~island_side)
 
@@
 Anova(shannon_mod)
-plot(allEffects(shannon_mod_2))
+
 mod_emm <- emmeans(shannon_mod, ~ habitat * island_side * year_factor) # if we want to compare habitats year to year
@@
 
-```{r mean rank shift, fig.height = 24, fig.width=16}
+```{r mean rank shift, fig.height = 24, fig.width=16, eval = FALSE}
 # rank shift
@@
 
-```{r rate of change, fig.height=15, fig.width=15}
+```{r rate of change, fig.height=15, fig.width=15, eval = FALSE}
 # rate change
@@
 
-```{r rate change vis method 2, fig.height=15, fig.width=15}
+```{r rate change vis method 2, fig.height=15, fig.width=15, eval = FALSE}
 rate_change_untransformed<-
@@
 ```{r CV, fig.height = 8, fig.width = 10}
-# coefficient of variation
-CV<-function(x){
-  return(sd(x,na.rm=T)/mean(x,na.rm=T))
-}
 

### 2024-11-20 manuscript_1_code.Rmd

@@
 library(knitr)
+library(performance)
 ```
@@
 r.squaredGLMM(cover_mod) # 0.631, 0.650
+performance::r2(cover_mod) # 0.634, 0.654
 emtrends(cover_mod, pairwise ~ habitat, var = "richness") # backreef not different from forereef, forereef also not different
@@
 r.squaredGLMM(cover_mod_fg) # 0.64, 0.65
+performance::r2(cover_mod_fg) # 0.63, 0.65
 emtrends(cover_mod_fg, pairwise ~ habitat, var = "functional_richness") 
@@
 
-r.squaredGLMM(synch_cov_mod) # error??
+# can't compute r2 - manually fit null model. Null model won't fit without sv_trans() for some reason
+null_synch_mod <- glmmTMB(sv_trans(cover_mean) ~ 1, family = beta_family(), data = diversity_stability_synchrony)
+performance::r2(model = synch_cov_mod, null_model = null_synch_mod)
 emtrends(synch_cov_mod, pairwise ~ habitat, var = "synchrony") # forereef 17 distinct from all other habitats
@@
 r.squaredGLMM(site_cover_mod) # 0.2212534, 0.5990825
+performance::r2(site_cover_mod) # 0.22, 0.59
 emtrends(site_cover_mod, pairwise ~ habitat, var = "richness") # no difference between the habitats
@@
 r.squaredGLMM(site_cover_mod_fg) # 0.180, 0.563
+performance::r2(site_cover_mod_fg) # 0.180, 0.563
 emtrends(site_cover_mod_fg, pairwise ~ habitat, var = "functional_richness") # no difference between the habitats
@@
 # synchrony:habitat 8.8866  3    0.03084 *
+performance::r2(synch_cov_mod_site)
 hist(residuals(synch_cov_mod_site)) # slight skew
@@
 r.squaredGLMM(rich_stab_mod) # 0.803, 0.808
+performance::r2(rich_stab_mod) # 0.812, 0.816
 emtrends(rich_stab_mod, pairwise ~ habitat, var = "richness_mean") # fringing not diff from outer17, backreef not different from outer10
@@
 r.squaredGLMM(rich_stab_mod_fg) # 0.819, 0.828
+performance::r2(rich_stab_mod_fg)
 emtrends(rich_stab_mod_fg, pairwise ~ habitat, var = "functional_richness_mean") 
@@
 r.squaredGLMM(synch_stab_mod) # 0.63, 0.71
+performance::r2(synch_stab_mod)
 emtrends(synch_stab_mod, pairwise ~ habitat, var = "synchrony") # fringing and backreef not different
@@
 r.squaredGLMM(rich_stab_mod_site) # 0.671
+performance::r2(rich_stab_mod_site)
 emtrends(rich_stab_mod_site, pairwise ~ habitat, var = "richness_mean") # no differences across habitats
@@
 r.squaredGLMM(rich_stab_mod_site_fg) # 0.563
+performance::r2(rich_stab_mod_site_fg) # 0.551
 emtrends(rich_stab_mod_site_fg, pairwise ~ habitat, var = "functional_richness_mean") 
@@
 r.squaredGLMM(synch_stab_mod_site) # 0.678
+performance::r2(synch_stab_mod_site) # 0.702
 emtrends(synch_stab_mod_site, pairwise ~ habitat, var = "synchrony") # backreef is different from forereef 17, fringing is a p = 0.05
@@
 # spatial_synchrony:habitat  6.7239  3  0.0812368 .  
+performance::r2(spatial_synchrony_mod) # 0.64
 emtrends(spatial_synchrony_mod, pairwise ~ habitat, var = "spatial_synchrony") # no sig differences but close
@@
 # synchrony_mean:habitat 19.062  3  0.0002654 ***
+performance::r2(mean_synchrony_mod)
 emtrends(mean_synchrony_mod, pairwise ~ habitat, var = "synchrony_mean") # no sig differences but close
@@
 # stability_mean:habitat  9.4832  3   0.023511 *  
+performance::r2(alpha_gamma_stab_mod)
 

### 2024-11-20 Personal_Prelim_Code/mcr_algal_data_cleaning_final.Rmd
Added taxa information for folks to see but changed Caulerpa racemosa back so Naom can decide

--- 2024-09-24 Personal_Prelim_Code/mcr_algal_data_cleaning_final.Rmd
@@
 #### REPLACE OUTDATED TAXA
+
+# Julianna notes on taxa:
+# Changed Caulerpa peltata to Caulerpa chemnitzia 
+# Changed Chnoospora implexa to Pseudochnoospora implexa
+# Changed Neogoniolithon frutescens to Neogoniolithon brassica-florida
+# Changed Caulerpa pickeringii to Caulerpa webbiana
+# Changed Boodlea kaeneana to Cladophoropsis membranacea
+# Changed Dictyota divaricata to Dictyota cervicornis
 

### 2024-11-26 Personal_Prelim_Code/mcr_algal_data_cleaning_final.Rmd
updated new species names for data with Julianna's most recent check

@@
 
-# Caulerpa peltata is outdated --> Caulerpa racemosa
-data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Caulerpa peltata"] <- "Caulerpa racemosa"
+### ADDING JULIANNA UPDATES HERE ###
+data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Chnoospora implexa"] <- "Pseudochnoospora implexa"
+
+data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Caulerpa pickeringii"] <- "Caulerpa webbiana"
+
+data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Boodlea kaeneana"] <- "Cladophoropsis membranacea"
+
+data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Dictyota divaricata"] <- "Dictyota cervicornis"
+
+# Caulerpa peltata is outdated --> Caulerpa racemosa JULIANNA UPDATE = Caulerpa chemnitzia
+data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Caulerpa peltata"] <- "Caulerpa chemnitzia"
+
+### EVERYTHING ELSE BELOW AS BEFORE ###
 
@@
 data <- droplevels(data)
-levels(as.factor(data$Taxonomy_Substrate_Functional_Group)) # 84 levels
+levels(as.factor(data$Taxonomy_Substrate_Functional_Group)) # 84 levels --> 85 AFTER JULIANNA UPDATES
 
@@
 
-not_algae <- c("Ascidian", "Bare Space", "Coral", "Coral Rubble", "Corallimorpharia", "Heteractis sp.", "Millepora platyphylla", "No data", "Sand", "Sacrophyton sp.", "Shell Debris", "Soft Coral", "Sponge", "Tridacna sp.", "Zooxanthellae")
+not_algae <- c("Ascidian", "Bare Space", "Coral", "Coral Rubble", "Corallimorpharia", "Heteractis sp.", "Millepora platyphylla", "No data", "Sand", "Sarcophyton sp.", "Shell Debris", "Soft Coral", "Sponge", "Tridacna sp.", "Zooxanthellae")
 
@@
 #writing one file so column names will work with Noam's code
-write.csv(wide_alldata_filtered, "allbenthicdata_cleaned.csv", row.names = FALSE)
+write.csv(wide_alldata_filtered, "data/allbenthicdata_cleaned.csv", row.names = FALSE)
 

### 2024-11-27 R/functions.R
Updated functions to ensure proper pathing

@@ -30,8 +30,8 @@ extract_ranges <- function(df, # a dataframe
                            columns # columns you want to summarize (e.g. richness, cover, synchrony)
 ){
   df %>% # in your dataframe
-    group_by(across(all_of(group))) %>% 
-    summarise(across(all_of(columns), # summarize the columns in your group
+    dplyr::group_by(across(all_of(group))) %>% 
+    dplyr::summarise(across(all_of(columns), # summarize the columns in your group
                      list(min = ~min(.x, na.rm = TRUE), max = ~max(.x, na.rm = TRUE)), # by taking the min/max of them
                      .names = "{.fn}_{.col}")) # and assigning them to the columns name "min/max_value"
 }

### 2024-11-27 manuscript_1_code.Rmd
Updated stat info for new species changes discovered in Nov 2024 by J Renzi

@@
 cover_means <- cover_df %>%
-  group_by(year, habitat) %>%
-  summarise(macroalgae = mean(macroalgae),
+  dplyr::group_by(year, habitat) %>%
+  dplyr::summarise(macroalgae = mean(macroalgae),
             coral = mean(coral),
@@
 cover_se <- cover_df %>%
-  group_by(year, habitat) %>%
-  summarise(macroalgae = se(macroalgae),
+  dplyr::group_by(year, habitat) %>%
+  dplyr::summarise(macroalgae = se(macroalgae),
             coral = se(coral),
@@
 taxa_sum_check <- macro_functional_groups %>%
-  group_by(location, year) %>%
-  summarise(sum_macrocover = sum(prop_cover))
+  dplyr::group_by(location, year) %>%
+  dplyr::summarise(sum_macrocover = sum(prop_cover))
 
@@
 fg_summary <- macro_functional_groups %>%
-  group_by(location, year, functional_group, habitat, site) %>%
-  summarise(sum_fg_cover = sum(prop_cover))
+  dplyr::group_by(location, year, functional_group, habitat, site) %>%
+  dplyr::summarise(sum_fg_cover = sum(prop_cover))
 
@@
 fg_summary_site_SUM <- macro_functional_groups %>%
-  group_by(site, year, habitat, functional_group) %>%
-  summarise(sum_fg_cover = sum(prop_cover))
+  dplyr::group_by(site, year, habitat, functional_group) %>%
+  dplyr::summarise(sum_fg_cover = sum(prop_cover))
 
@@
 fg_summary_qaqc <- macro_functional_groups %>%
-  group_by(location, year, functional_group) %>%
-  summarise(sum_fg_cover = sum(prop_cover))
+  dplyr::group_by(location, year, functional_group) %>%
+  dplyr::summarise(sum_fg_cover = sum(prop_cover))
 
 fg_qaqc <- fg_summary_qaqc %>%
-  group_by(location, year) %>%
-  summarise(sum_fg_check = sum(sum_fg_cover))
+  dplyr::group_by(location, year) %>%
+  dplyr::summarise(sum_fg_check = sum(sum_fg_cover))
 
@@
 fg_summary_wide <- fg_summary %>%
-  pivot_wider(names_from = "functional_group", values_from = "sum_fg_cover")
+  tidyr::pivot_wider(names_from = "functional_group", values_from = "sum_fg_cover")
 
@@
 #                     Chisq Df Pr(>Chisq)    
-# richness         30916.73  1  < 2.2e-16 ***
-# habitat            141.78  3  < 2.2e-16 ***
-# richness:habitat   294.54  3  < 2.2e-16 ***
+# richness         30918.96  1  < 2.2e-16 ***
+# habitat            141.68  3  < 2.2e-16 ***
+# richness:habitat   294.44  3  < 2.2e-16 ***
 hist(residuals(cover_mod)) # looks fine
@@
 r.squaredGLMM(cover_mod) # 0.631, 0.650
-performance::r2(cover_mod) # 0.634, 0.654
+performance::r2(cover_mod) # 0.635, 0.654
 emtrends(cover_mod, pairwise ~ habitat, var = "richness") # backreef not different from forereef, forereef also not different
 # contrast                    estimate     SE  df z.ratio p.value
-# Fringing - Backreef          0.34720 0.0241 Inf  14.409  <.0001
-# Fringing - Forereef 10m      0.34865 0.0240 Inf  14.514  <.0001
-# Fringing - Forereef 17m      0.38953 0.0250 Inf  15.599  <.0001
-# Backreef - Forereef 10m      0.00144 0.0205 Inf   0.071  0.9999
-# Backreef - Forereef 17m      0.04232 0.0214 Inf   1.979  0.1957
-# Forereef 10m - Forereef 17m  0.04088 0.0209 Inf   1.952  0.2063
+# Fringing - Backreef          0.34715 0.0241 Inf  14.407  <.0001
+# Fringing - Forereef 10m      0.34862 0.0240 Inf  14.513  <.0001
+# Fringing - Forereef 17m      0.38937 0.0250 Inf  15.593  <.0001
+# Backreef - Forereef 10m      0.00147 0.0205 Inf   0.072  0.9999
+# Backreef - Forereef 17m      0.04223 0.0214 Inf   1.975  0.1974
+# Forereef 10m - Forereef 17m  0.04076 0.0209 Inf   1.947  0.2087
+
 ```
@@
 #                                 Chisq Df Pr(>Chisq)    
-# fg_richness_quad         31341.052  1  < 2.2e-16 ***
-# habitat                     49.887  3  8.444e-11 ***
-# fg_richness_quad:habitat   447.780  3  < 2.2e-16 ***
+# functional_richness         31362.191  1  < 2.2e-16 ***
+# habitat                        49.343  3  1.103e-10 ***
+# functional_richness:habitat   451.974  3  < 2.2e-16 ***
 hist(residuals(cover_mod_fg)) # good
@@
 # contrast                    estimate     SE  df z.ratio p.value
-# Fringing - Backreef            0.214 0.0266 Inf   8.051  <.0001
-# Fringing - Forereef 10m        0.346 0.0261 Inf  13.252  <.0001
-# Fringing - Forereef 17m        0.530 0.0262 Inf  20.244  <.0001
-# Backreef - Forereef 10m        0.132 0.0231 Inf   5.708  <.0001
-# Backreef - Forereef 17m        0.316 0.0232 Inf  13.617  <.0001
-# Forereef 10m - Forereef 17m    0.184 0.0223 Inf   8.241  <.0001
+# Fringing - Backreef            0.211 0.0266 Inf   7.937  <.0001
+# Fringing - Forereef 10m        0.348 0.0262 Inf  13.285  <.0001
+# Fringing - Forereef 17m        0.531 0.0262 Inf  20.280  <.0001
+# Backreef - Forereef 10m        0.136 0.0231 Inf   5.888  <.0001
+# Backreef - Forereef 17m        0.320 0.0232 Inf  13.812  <.0001
+# Forereef 10m - Forereef 17m    0.184 0.0223 Inf   8.253  <.0001
 
@@
 #                     Chisq Df Pr(>Chisq)    
-# synchrony         209.545  1  < 2.2e-16 ***
-# habitat            52.512  3  2.330e-11 ***
-# synchrony:habitat  54.632  3  8.227e-12 ***
+# synchrony         209.521  1  < 2.2e-16 ***
+# habitat            52.498  3  2.345e-11 ***
+# synchrony:habitat  54.627  3  8.246e-12 ***
 hist(residuals(synch_cov_mod)) # slight skew
@@
 null_synch_mod <- glmmTMB(sv_trans(cover_mean) ~ 1, family = beta_family(), data = diversity_stability_synchrony)
-performance::r2(model = synch_cov_mod, null_model = null_synch_mod)
+performance::r2(model = synch_cov_mod, null_model = null_synch_mod) # 0.352, 0.534
 emtrends(synch_cov_mod, pairwise ~ habitat, var = "synchrony") # forereef 17 distinct from all other habitats
@@
 #                    Chisq Df Pr(>Chisq)    
-# richness         58.9195  1  1.643e-14 ***
-# habitat           6.6247  3    0.08487 .  
-# richness:habitat  3.2967  3    0.34811    
+# richness         58.9196  1  1.642e-14 ***
+# habitat           6.6248  3    0.08487 .  
+# richness:habitat  3.2966  3    0.34811    
 hist(residuals(site_cover_mod)) # some positive skew, not the worst thing ever
@@
 #                               Chisq Df Pr(>Chisq)    
-# fg_richness_site         30.9412  1   2.66e-08 ***
-# habitat                   4.8929  3     0.1798    
-# fg_richness_site:habitat  4.2193  3     0.2387
+# functional_richness         30.3776  1  3.556e-08 ***
+# habitat                      4.9314  3     0.1769    
+# functional_richness:habitat  4.1429  3     0.2464
 
@@
 emtrends(site_cover_mod_fg, pairwise ~ habitat, var = "functional_richness") # no difference between the habitats
-# Fringing - Backreef           0.0379 0.0825 Inf   0.459  0.9679
-# Fringing - Forereef 10m       0.1846 0.0930 Inf   1.985  0.1935
-# Fringing - Forereef 17m       0.0933 0.0960 Inf   0.972  0.7655
-# Backreef - Forereef 10m       0.1467 0.0986 Inf   1.488  0.4447
-# Backreef - Forereef 17m       0.0554 0.1018 Inf   0.545  0.9480
-# Forereef 10m - Forereef 17m  -0.0913 0.1093 Inf  -0.835  0.8377
+# contrast                    estimate     SE  df z.ratio p.value
+# Fringing - Backreef          -0.0163 0.0394 Inf  -0.413  0.9762
+# Fringing - Forereef 10m       0.0639 0.0503 Inf   1.271  0.5818
+# Fringing - Forereef 17m       0.0334 0.0612 Inf   0.546  0.9476
+# Backreef - Forereef 10m       0.0801 0.0460 Inf   1.742  0.3020
+# Backreef - Forereef 17m       0.0497 0.0583 Inf   0.853  0.8289
+# Forereef 10m - Forereef 17m  -0.0304 0.0651 Inf  -0.467  0.9662
 
@@
 #                    Chisq Df Pr(>Chisq)  
-# synchrony         2.4184  1    0.11992  
-# habitat           7.8002  3    0.05033 .
-# synchrony:habitat 8.8866  3    0.03084 *
-performance::r2(synch_cov_mod_site)
+# synchrony         2.4156  1    0.12013  
+# habitat           7.7966  3    0.05041 .
+# synchrony:habitat 8.8876  3    0.03082 *
+performance::r2(synch_cov_mod_site) # 0.48
 hist(residuals(synch_cov_mod_site)) # slight skew
@@
 # contrast                    estimate   SE  df z.ratio p.value
-# Fringing - Backreef           -4.299 3.38 Inf  -1.270  0.5821
-# Fringing - Forereef 10m       -0.279 3.22 Inf  -0.087  0.9998
-# Fringing - Forereef 17m        3.675 3.78 Inf   0.973  0.7649
-# Backreef - Forereef 10m        4.019 1.98 Inf   2.031  0.1764
-# Backreef - Forereef 17m        7.973 2.80 Inf   2.848  0.0228
-# Forereef 10m - Forereef 17m    3.954 2.59 Inf   1.526  0.4217
+# Fringing - Backreef            -4.31 3.39 Inf  -1.269  0.5824
+# Fringing - Forereef 10m        -0.29 3.23 Inf  -0.090  0.9997
+# Fringing - Forereef 17m         3.66 3.79 Inf   0.968  0.7679
+# Backreef - Forereef 10m         4.02 1.98 Inf   2.031  0.1764
+# Backreef - Forereef 17m         7.97 2.80 Inf   2.848  0.0228
+# Forereef 10m - Forereef 17m     3.95 2.59 Inf   1.526  0.4218
 ```
@@
 #                          Chisq Df Pr(>Chisq)    
-# richness_mean         2721.438  1  < 2.2e-16 ***
-# habitat                325.876  3  < 2.2e-16 ***
-# richness_mean:habitat   33.926  3  2.054e-07 ***
+# richness_mean         2720.311  1  < 2.2e-16 ***
+# habitat                325.657  3  < 2.2e-16 ***
+# richness_mean:habitat   33.857  3  2.124e-07 ***
 hist(residuals(rich_stab_mod)) # looks good
@@
 emtrends(rich_stab_mod, pairwise ~ habitat, var = "richness_mean") # fringing not diff from outer17, backreef not different from outer10
-# contrast                     estimate     SE  df z.ratio p.value
-# Fringing - Backreef          0.186670 0.0414 Inf   4.512  <.0001
-# Fringing - Forereef 10m      0.187196 0.0455 Inf   4.114  0.0002
-# Fringing - Forereef 17m      0.027112 0.0463 Inf   0.586  0.9364
-# Backreef - Forereef 10m      0.000525 0.0405 Inf   0.013  1.0000
-# Backreef - Forereef 17m     -0.159559 0.0411 Inf  -3.885  0.0006
-# Forereef 10m - Forereef 17m -0.160084 0.0439 Inf  -3.647  0.0015
+# $contrasts
+#  contrast                     estimate     SE  df z.ratio p.value
+#  Fringing - Backreef          0.186640 0.0414 Inf   4.511  <.0001
+#  Fringing - Forereef 10m      0.187248 0.0455 Inf   4.114  0.0002
+#  Fringing - Forereef 17m      0.027482 0.0463 Inf   0.594  0.9340
+#  Backreef - Forereef 10m      0.000608 0.0405 Inf   0.015  1.0000
+#  Backreef - Forereef 17m     -0.159158 0.0411 Inf  -3.875  0.0006
+#  Forereef 10m - Forereef 17m -0.159766 0.0439 Inf  -3.640  0.0016
+
 ```
@@
 #                                    Chisq Df Pr(>Chisq)    
-# fg_richness_mean         3079.229  1  < 2.2e-16 ***
-# habitat                   118.605  3  < 2.2e-16 ***
-# fg_richness_mean:habitat   25.692  3  1.107e-05 ***
+# functional_richness_mean         3082.22  1  < 2.2e-16 ***
+# habitat                           116.63  3  < 2.2e-16 ***
+# functional_richness_mean:habitat   25.79  3  1.055e-05 ***
 
@@
 # contrast                    estimate     SE  df z.ratio p.value
-# Fringing - Backreef           0.1872 0.0468 Inf   3.999  0.0004
-# Fringing - Forereef 10m       0.2348 0.0519 Inf   4.524  <.0001
-# Fringing - Forereef 17m       0.2040 0.0489 Inf   4.173  0.0002
-# Backreef - Forereef 10m       0.0477 0.0472 Inf   1.011  0.7432
-# Backreef - Forereef 17m       0.0169 0.0439 Inf   0.385  0.9806
-# Forereef 10m - Forereef 17m  -0.0308 0.0475 Inf  -0.648  0.9164
+# Fringing - Backreef           0.1825 0.0468 Inf   3.903  0.0006
+# Fringing - Forereef 10m       0.2372 0.0518 Inf   4.575  <.0001
+# Fringing - Forereef 17m       0.2045 0.0488 Inf   4.187  0.0002
+# Backreef - Forereef 10m       0.0547 0.0472 Inf   1.158  0.6534
+# Backreef - Forereef 17m       0.0219 0.0439 Inf   0.500  0.9592
+# Forereef 10m - Forereef 17m  -0.0327 0.0475 Inf  -0.689  0.9015
 #
@@
 #                     Chisq Df Pr(>Chisq)    
-# synchrony         1308.33  1  < 2.2e-16 ***
+# synchrony         1308.34  1  < 2.2e-16 ***
 # habitat            103.13  3  < 2.2e-16 ***
-# synchrony:habitat  327.97  3  < 2.2e-16 ***
+# synchrony:habitat  327.96  3  < 2.2e-16 ***
 hist(residuals(synch_stab_mod)) # looks good
@@
 r.squaredGLMM(synch_stab_mod) # 0.63, 0.71
-performance::r2(synch_stab_mod)
+performance::r2(synch_stab_mod) # 0.66, 0.74
 emtrends(synch_stab_mod, pairwise ~ habitat, var = "synchrony") # fringing and backreef not different
@@
 
-#                           Chisq Df Pr(>Chisq)   
-# avg_richness_fg          3.4019  1   0.065122 . 
-# habitat                 12.8194  3   0.005044 **
-# avg_richness_fg:habitat  7.9921  3   0.046176 *
+# 
+#                                    Chisq Df Pr(>Chisq)   
+# functional_richness_mean          3.3605  1    0.06678 . 
+# habitat                          12.9894  3    0.00466 **
+# functional_richness_mean:habitat  7.9219  3    0.04765 *
 
@@
 # Fringing different from outer 10
-# contrast                    estimate    SE  df z.ratio p.value
-# Fringing - Backreef           0.2552 0.254 Inf   1.005  0.7463
-# Fringing - Forereef 10m       1.2611 0.469 Inf   2.686  0.0364
-# Fringing - Forereef 17m       0.3543 0.287 Inf   1.235  0.6045
-# Backreef - Forereef 10m       1.0059 0.492 Inf   2.044  0.1720
-# Backreef - Forereef 17m       0.0991 0.323 Inf   0.307  0.9900
-# Forereef 10m - Forereef 17m  -0.9068 0.510 Inf  -1.778  0.2838
+#$contrasts
+# contrast                    estimate    SE  df z.ratio p.value
+# Fringing - Backreef            0.244 0.252 Inf   0.966  0.7686
+# Fringing - Forereef 10m        1.250 0.469 Inf   2.663  0.0388
+# Fringing - Forereef 17m        0.354 0.280 Inf   1.264  0.5860
+# Backreef - Forereef 10m        1.006 0.493 Inf   2.040  0.1735
+# Backreef - Forereef 17m        0.110 0.318 Inf   0.346  0.9858
+# Forereef 10m - Forereef 17m   -0.896 0.508 Inf  -1.763  0.2911
 ```
@@
 #                    Chisq Df Pr(>Chisq)    
-# synchrony         22.0686  1  2.631e-06 ***
-# habitat            9.3333  3    0.02517 *  
-# synchrony:habitat 10.5561  3    0.01439 *  
+# synchrony         22.117  1  2.565e-06 ***
+# habitat            9.345  3    0.02504 *  
+# synchrony:habitat 10.534  3    0.01453 *   
 hist(residuals(synch_stab_mod_site)) # almost uniform
@@
 # contrast                    estimate    SE  df z.ratio p.value
-# Fringing - Backreef           -0.256 0.971 Inf  -0.263  0.9936
-# Fringing - Forereef 10m        1.324 0.916 Inf   1.445  0.4709
-# Fringing - Forereef 17m        2.930 1.142 Inf   2.567  0.0503
-# Backreef - Forereef 10m        1.580 0.858 Inf   1.842  0.2534
-# Backreef - Forereef 17m        3.186 1.095 Inf   2.909  0.0190
-# Forereef 10m - Forereef 17m    1.606 1.047 Inf   1.534  0.4172
+# Fringing - Backreef           -0.262 0.972 Inf  -0.269  0.9932
+# Fringing - Forereef 10m        1.318 0.918 Inf   1.437  0.4764
+# Fringing - Forereef 17m        2.924 1.143 Inf   2.559  0.0513
+# Backreef - Forereef 10m        1.580 0.857 Inf   1.843  0.2533
+# Backreef - Forereef 17m        3.186 1.095 Inf   2.910  0.0189
+# Forereef 10m - Forereef 17m    1.606 1.047 Inf   1.534  0.4170
 ```
@@
 mean_synchrony_stability <- diversity_stability_synchrony %>% 
-  group_by(habitat, site) %>% 
-  summarise(synchrony_mean = mean(synchrony, na.rm = T),
+  dplyr::group_by(habitat, site) %>% 
+  dplyr::summarise(synchrony_mean = mean(synchrony, na.rm = T),
             stability_mean = mean(cover_stability, na.rm = T))
@@
 #                             Chisq Df Pr(>Chisq)    
-# spatial_synchrony         16.2914  1  5.431e-05 ***
-# habitat                   19.5303  3  0.0002124 ***
-# spatial_synchrony:habitat  6.7239  3  0.0812368 .  
+# spatial_synchrony         16.2942  1  5.423e-05 ***
+# habitat                   19.5302  3  0.0002124 ***
+# spatial_synchrony:habitat  6.7253  3  0.0811882 .  
 performance::r2(spatial_synchrony_mod) # 0.64
@@
 # contrast                    estimate    SE  df z.ratio p.value
-# Fringing - Backreef            0.404 1.765 Inf   0.229  0.9958
+# Fringing - Backreef            0.404 1.764 Inf   0.229  0.9958
 # Fringing - Forereef 10m        1.365 1.754 Inf   0.778  0.8643
@@
 # Backreef - Forereef 10m        0.961 0.857 Inf   1.121  0.6766
-# Backreef - Forereef 17m        2.708 1.095 Inf   2.472  0.0644
-# Forereef 10m - Forereef 17m    1.747 1.078 Inf   1.621  0.3668
+# Backreef - Forereef 17m        2.708 1.095 Inf   2.472  0.0643
+# Forereef 10m - Forereef 17m    1.747 1.078 Inf   1.621  0.3667
 ```
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Stability\n", x = "Spatial synchrony", title = "") +
+  labs(y = "", x = "Spatial synchrony", title = "") +
   trends_theme
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Stability\n", x = "Mean synchrony", title = "") +
+  labs(y = "Site-level Stability\n", x = "Mean plot-level species synchrony", title = "") +
   trends_theme
@@
 #                          Chisq Df Pr(>Chisq)    
-# stability_mean         61.2725  1   4.97e-15 ***
-# habitat                12.9323  3   0.004785 ** 
-# stability_mean:habitat  9.4832  3   0.023511 *  
-performance::r2(alpha_gamma_stab_mod)
+# stability_mean         61.2742  1  4.965e-15 ***
+# habitat                12.9328  3   0.004784 ** 
+# stability_mean:habitat  9.4834  3   0.023508 * 
+performance::r2(alpha_gamma_stab_mod) # 0.838
 
 emtrends(alpha_gamma_stab_mod, pairwise ~ habitat, var = "stability_mean") # backreef different from fringe and forereef 10
+# $contrasts
+#  contrast                    estimate    SE  df z.ratio p.value
 #  Fringing - Backreef            1.078 0.381 Inf   2.828  0.0242
@@
 #  Fringing - Forereef 17m        0.380 0.305 Inf   1.248  0.5965
-# Backreef - Forereef 10m       -0.903 0.349 Inf  -2.589  0.0475
+#  Backreef - Forereef 10m       -0.903 0.349 Inf  -2.589  0.0474
 #  Backreef - Forereef 17m       -0.698 0.309 Inf  -2.256  0.1085
-# Forereef 10m - Forereef 17m    0.205 0.263 Inf   0.779  0.8638
+#  Forereef 10m - Forereef 17m    0.205 0.263 Inf   0.779  0.8639
 
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Site stability\n", x = "Plot stability", title = "") +
+  labs(y = "", x = "Plot stability", title = "") +
   trends_theme
@@
 
-# ggsave(filename = "output/figure_5_option3.png", figure_5, height = 6, width = 22)
-```
-
-# SUPPLEMENTAL FIGURE CODE:
+# ggsave(filename = "output/figure_5_option3_v2.png", figure_5, height = 6, width = 22)
+```
 
@@
 quad_richness_df <- alpha_diversity_quad_macro %>% 
-  group_by(year, habitat) %>% 
-  summarise(richness_mean = mean(richness),
+  dplyr::group_by(year, habitat) %>% 
+  dplyr::summarise(richness_mean = mean(richness),
             richness_se = se(richness))
@@
 site_richness_df <- alpha_diversity_site_macro %>% 
-  group_by(year, habitat) %>% 
-  summarise(richness_mean = mean(richness),
+  dplyr::group_by(year, habitat) %>% 
+  dplyr::summarise(richness_mean = mean(richness),
             richness_se = se(richness))

### 2024-11-27 manuscript_1_code.Rmd
Still more axis corrections

@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "", x = "Plot stability", title = "") +
+  labs(y = "", x = "Plot-level stability", title = "") +
   trends_theme

### 2025-01-21 SEM_quad.Rmd
Added code for piecewiseSEMs

---
title: "SEM with quad-level algae data"
output: html_document
date: "2025-01-17"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(tidyverse)
require(here)
require(piecewiseSEM)
require(lme4) # glmer() function
require(performance)
```

# Read in data

```{r}
# quad-level data summarized with one point for all time
algae <- read_csv(here("data/diversity_stability_synchrony.csv"))
```

# Scale data

Gamma families are currently not supported and we're not able to get glmmTMB to work for the time being.

```{r}
algae %>% 
  mutate(
    richness_mean_scale = scale(algae$richness_mean)[,1],
    cover_mean_scale = scale(algae$cover_mean)[,1],
    cover_stability_scale = scale(algae$cover_stability)[,1],
    synchrony_scale = scale(algae$synchrony)[,1]
  ) -> algae_scaled
```

# Make a simple SEM

First, fit individual models

```{r}
# effect of richness on synchrony
m.sprich.synch <- glmer(synchrony_scale ~ richness_mean_scale + (1|site) + (1|habitat),
                        data = algae_scaled)
  performance::check_model(m.sprich.synch)

# effect of synchrony on stability
m.synch.stab.cov <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site) + (1|habitat),
                      data = algae_scaled)  
  performance::check_model(m.synch.stab)

# effect of richness on cover
m.sprich.cov <- glmer(cover_mean_scale ~ richness_mean_scale + (1|site) + (1|habitat),
                      data = algae_scaled)
  performance::check_model(m.sprich.cov)
```

Try it with piecewiseSEM

```{r}
# Build SEM
sem <- psem(
  m.sprich.synch,
  m.synch.stab.cov,
  m.sprich.cov,
  data = as.data.frame(algae)
)

summary(sem)
```

Plot

```{r}
# Extract coefficients
#coefficients <- summary(sem)$coefficients

# Add a column for color based on the sign of the path coefficient
#coefficients$color <- ifelse(coefficients$Estimate > 0, "blue", #"red")

plot(sem)
```

# Frining reef: bracketed

This is trying to make these models without rescaling within the fringing reef to see if it matters (i.e., we'll just subset from the original scaled dataframe)

```{r}
# effect of richness on synchrony
m.sprich.synch.FR <- glmer(synchrony_scale ~ richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Fringing", ])
  performance::check_model(m.sprich.synch.FR)

# effect of synchrony on stability
m.synch.stab.cov.FR <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Fringing", ])  
  performance::check_model(m.synch.stab.cov.FR)

# effect of richness on cover
m.sprich.cov.FR <- glmer(cover_mean_scale ~ richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Fringing", ])
  performance::check_model(m.sprich.cov.FR)
```

Try it with piecewiseSEM

```{r}
# Build SEM
sem_fringe_brac <- psem(
  m.sprich.synch.FR,
  m.synch.stab.cov.FR,
  m.sprich.cov.FR,
  data = as.data.frame(algae_scaled[algae_scaled$habitat == "Fringing", ])
)

summary(sem_fringe_brac)
```

Plot

```{r}
plot(sem_fringe_brac)
```

# Backreef: bracketed

This is trying to make these models without rescaling within the fringing reef to see if it matters (i.e., we'll just subset from the original scaled dataframe)

```{r}
# effect of richness on synchrony
m.sprich.synch.BA <- glmer(synchrony_scale ~ richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Backreef", ])
  performance::check_model(m.sprich.synch.BA)

# effect of synchrony on stability
m.synch.stab.cov.BA <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Backreef", ])  
  performance::check_model(m.synch.stab.cov.BA)

# effect of richness on cover
m.sprich.cov.BA <- glmer(cover_mean_scale ~ richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Backreef", ])
  performance::check_model(m.sprich.cov.BA)
```

Try it with piecewiseSEM

```{r}
# Build SEM
sem_back_brac <- psem(
  m.sprich.synch.BA,
  m.synch.stab.cov.BA,
  m.sprich.cov.BA,
  data = as.data.frame(algae_scaled[algae_scaled$habitat == "Backreef", ])
)

summary(sem_back_brac)
```

Plot

```{r}
plot(sem_back_brac)
```

# Forereef 10: bracketed

This is trying to make these models without rescaling within the fringing reef to see if it matters (i.e., we'll just subset from the original scaled dataframe)

```{r}
# effect of richness on synchrony
m.sprich.synch.FO10 <- glmer(synchrony_scale ~ richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Forereef 10m", ])
  performance::check_model(m.sprich.synch.FO10)

# effect of synchrony on stability
m.synch.stab.cov.FO10 <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Forereef 10m", ])  
  performance::check_model(m.synch.stab.cov.FO10)

# effect of richness on cover
m.sprich.cov.FO10 <- glmer(cover_mean_scale ~ richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Forereef 10m", ])
  performance::check_model(m.sprich.cov.FO10)
```

Try it with piecewiseSEM

```{r}
# Build SEM
sem_for10_brac <- psem(
  m.sprich.synch.FO10,
  m.synch.stab.cov.FO10,
  m.sprich.cov.FO10,
  data = as.data.frame(algae_scaled[algae_scaled$habitat == "Forereef 10m", ])
)

summary(sem_for10_brac)
```

Plot

```{r}
plot(sem_for10_brac)
```

# Forereef 17: bracketed

This is trying to make these models without rescaling within the fringing reef to see if it matters (i.e., we'll just subset from the original scaled dataframe)

```{r}
# effect of richness on synchrony
m.sprich.synch.FO17 <- glmer(synchrony_scale ~ richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Forereef 17m", ])
  performance::check_model(m.sprich.synch.FO17)

# effect of synchrony on stability
m.synch.stab.cov.FO17 <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Forereef 17m", ])  
  performance::check_model(m.synch.stab.cov.FO17)

# effect of richness on cover
m.sprich.cov.FO17 <- glmer(cover_mean_scale ~ richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Forereef 17m", ])
  performance::check_model(m.sprich.cov.FO17)
```

Try it with piecewiseSEM

```{r}
# Build SEM
sem_for17_brac <- psem(
  m.sprich.synch.FO17,
  m.synch.stab.cov.FO17,
  m.sprich.cov.FO17,
  data = as.data.frame(algae_scaled[algae_scaled$habitat == "Forereef 17m", ])
)

summary(sem_for17_brac)
```

Plot

```{r}
plot(sem_for17_brac)
```

# Try re-scaling within a habitat: Frining

```{r}
algae %>% 
  filter(habitat == "Fringing") -> fringe

fringe %>% 
  mutate(
    richness_mean_scale = scale(fringe$richness_mean)[,1],
    cover_mean_scale = scale(fringe$cover_mean)[,1],
    cover_stability_scale = scale(fringe$cover_stability)[,1],
    synchrony_scale = scale(fringe$synchrony)[,1]
  ) -> algae_scaled_fringe

```

```{r}
# effect of richness on synchrony
m.sprich.synch.FRsc <- glmer(synchrony_scale ~ richness_mean_scale + (1|site),
                        data = algae_scaled_fringe)
  performance::check_model(m.sprich.synch.FRsc)

# effect of synchrony on stability
m.synch.stab.cov.FRsc <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site),
                      data = algae_scaled_fringe)  
  performance::check_model(m.synch.stab.cov.FRsc)

# effect of richness on cover
m.sprich.cov.FRsc <- glmer(cover_mean_scale ~ richness_mean_scale + (1|site),
                      data = algae_scaled_fringe)
  performance::check_model(m.sprich.cov.FRsc)
```

Try it with piecewiseSEM

```{r}
# Build SEM
sem.FRsc <- psem(
  m.sprich.synch.FRsc,
  m.synch.stab.cov.FRsc,
  m.sprich.cov.FRsc,
  data = as.data.frame(algae_scaled_fringe)
)

summary(sem.FRsc)
```

Plot

```{r}
plot(sem.FRsc)
```

### 2025-01-21 manuscript_1_code.Rmd

@@
 
+### mean plot-level alpha diversity metrics
+
+```{r mean plot level adiv}
+mean_adiv_quad <- alpha_diversity_quad_macro %>% 
+  dplyr::group_by(year, habitat, site) %>% 
+  dplyr::summarise(
+    richness_mean = mean(richness, na.rm = T),
+    functional_richness_mean = mean(functional_richness, na.rm = T),
+    cover_mean = mean(macroalgal_prop_cover, na.rm = T))
+
+# sv transform cover due to 0s in data
+mean_adiv_quad$cover_trans <- sv_trans(mean_adiv_quad$cover_mean)
+```
+
 ## models
@@
 
+```{r mean richness quad model}
+cover_mod_2 <- glmmTMB(cover_trans ~ richness_mean*habitat + (1|site) + (1|year), family = beta_family(), data = mean_adiv_quad)
+summary(cover_mod_2)
+Anova(cover_mod_2)
+#                           Chisq Df Pr(>Chisq)    
+# richness_mean         1112.4314  1  < 2.2e-16 ***
+# habitat                  4.0269  3     0.2586    
+# richness_mean:habitat   25.9539  3  9.752e-06 ***
+hist(residuals(cover_mod_2)) # not terrible not great
+plot(residuals(cover_mod_2) ~ predict(cover_mod_2)) # heteroskedastic?
+performance::r2(cover_mod_2) # 0.774, 0.849
+emtrends(cover_mod_2, pairwise ~ habitat, var = "richness_mean") # Fringing different from backreef and FR10, backreef from FR17
+# $emtrends
+#  habitat      richness_mean.trend     SE  df asymp.LCL asymp.UCL
+#  Fringing                    1.77 0.0924 Inf      1.59      1.95
+#  Backreef                    1.19 0.0680 Inf      1.06      1.33
+#  Forereef 10m                1.44 0.0798 Inf      1.28      1.59
+#  Forereef 17m                1.55 0.1020 Inf      1.35      1.75
+# 
+# $contrasts
+#  contrast                    estimate    SE  df z.ratio p.value
+#  Fringing - Backreef            0.579 0.116 Inf   4.999  <.0001
+#  Fringing - Forereef 10m        0.334 0.120 Inf   2.792  0.0269
+#  Fringing - Forereef 17m        0.220 0.132 Inf   1.668  0.3406
+#  Backreef - Forereef 10m       -0.245 0.106 Inf  -2.322  0.0929
+#  Backreef - Forereef 17m       -0.360 0.124 Inf  -2.896  0.0198
+#  Forereef 10m - Forereef 17m   -0.114 0.121 Inf  -0.946  0.7799
+```
+
 ## figure 3a plot setup
@@
 
+### V2 SETUP
+diversity_ranges_quad_2 <- extract_ranges(mean_adiv_quad, "habitat", c("richness_mean", "cover_trans", "functional_richness_mean"))
+
+cover_emm_2 <- emmip(cover_mod_2,
+                   habitat ~ richness_mean,
+                   at = list(richness_mean = seq(0,10,0.1)), plotit = F, CIs = T)
+
+filtered_cover_effects_2 <- filter_ranges(trend = cover_emm_2, range_obj = diversity_ranges_quad_2, group = "habitat", value = "richness_mean")
+
 ```
@@
   geom_point(data = alpha_diversity_quad_macro,
-             aes(x = richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
+             aes(x = richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.5) + 
   geom_line(data = filtered_cover_effects, aes(x = richness, y = inv_logit(yvar), colour = habitat)) +
@@
 # ggsave(filename = "output/figure_3a.png", figure_3a, height = 10, width = 14)
+
+## V2
+(figure_3a_2 <- 
+  ggplot() +
+  geom_point(data = mean_adiv_quad,
+             aes(x = richness_mean, y = cover_trans, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_cover_effects_2, aes(x = richness_mean, y = inv_logit(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_cover_effects_2, aes(x = richness_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
+              alpha = 0.5) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Cover\n", x = "", title = "") +
+  model_themes 
+         )
+
+# ggsave(filename = "output/figure_3a_v2.png", figure_3a_2, height = 10, width = 14)
 ```
@@
 
+```{r mean fr quad model}
+
+cover_mod_fg_2 <- glmmTMB(cover_trans ~ functional_richness_mean*habitat + (1|site) + (1|year), family = beta_family(), data = mean_adiv_quad)
+summary(cover_mod_fg_2)
+car::Anova(cover_mod_fg_2)
+#                                     Chisq Df Pr(>Chisq)    
+# functional_richness_mean         1067.254  1  < 2.2e-16 ***
+# habitat                            43.997  3  1.512e-09 ***
+# functional_richness_mean:habitat   16.364  3  0.0009549 ***
+hist(residuals(cover_mod_fg_2)) # basically the same as the richness mod
+plot(residuals(cover_mod_fg_2) ~ predict(cover_mod_fg_2)) # here too
+
+performance::r2(cover_mod_fg_2) # 0.80, 0.84
+emtrends(cover_mod_fg_2, pairwise ~ habitat, var = "functional_richness_mean") # fringing different from both FRs
+# $emtrends
+#  habitat      functional_richness_mean.trend     SE  df asymp.LCL asymp.UCL
+#  Fringing                               2.01 0.1070 Inf      1.80      2.22
+#  Backreef                               1.81 0.1010 Inf      1.61      2.00
+#  Forereef 10m                           1.57 0.0918 Inf      1.39      1.75
+#  Forereef 17m                           1.49 0.1040 Inf      1.29      1.69
+# $contrasts
+#  contrast                    estimate    SE  df z.ratio p.value
+#  Fringing - Backreef           0.2035 0.148 Inf   1.374  0.5157
+#  Fringing - Forereef 10m       0.4398 0.138 Inf   3.182  0.0080
+#  Fringing - Forereef 17m       0.5183 0.142 Inf   3.647  0.0015
+#  Backreef - Forereef 10m       0.2363 0.135 Inf   1.752  0.2968
+#  Backreef - Forereef 17m       0.3149 0.147 Inf   2.149  0.1378
+#  Forereef 10m - Forereef 17m   0.0786 0.131 Inf   0.600  0.9321
+```
+
 ## figure 3b plot setup
@@
 
+### V2
+cover_emm_fg_2 <- emmip(cover_mod_fg_2,
+                        habitat ~ functional_richness_mean,
+                        at = list(functional_richness_mean = seq(0,10,0.1)), plotit = F, CIs = T)
+
+filtered_cover_effects_fg_2 <- filter_ranges(cover_emm_fg_2, diversity_ranges_quad_2, "habitat", "functional_richness_mean")
 ```
@@
   geom_point(data = alpha_diversity_quad_macro,
-             aes(x = functional_richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
+             aes(x = functional_richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.5) + 
   geom_line(data = filtered_cover_effects_fg, aes(x = functional_richness, y = inv_logit(yvar), colour = habitat)) +
@@
          )
-
 #ggsave(filename = "output/figure_3b_quad_v2.png", figure_3b, height = 10, width = 14)
+
+# V2
+(figure_3b_2 <- 
+  ggplot() +
+  geom_point(data = mean_adiv_quad,
+             aes(x = functional_richness_mean, y = cover_trans, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_cover_effects_fg_2, aes(x = functional_richness_mean, y = inv_logit(yvar), colour = habitat)) +
+  geom_ribbon(data = filtered_cover_effects_fg_2, aes(x = functional_richness_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "", x = "", title = "") +
+  trends_theme
+         )
+
+# ggsave(filename = "output/figure_3b_quad_v3.png", figure_3b_2, height = 10, width = 14)
 ```
@@
 
-```
-
+# repeat for the mean of each plot
+dss_ranges_2 <- extract_ranges(mean_synchrony_stability, "habitat",
+                             c("richness_mean", "functional_richness_mean", "synchrony_mean", "stability_mean", "cover_mean"))
+```
+
+### mean plot-level dss setup
+```{r mean plot dss}
+# calculate mean synchrony within a site (!= synchrony of species at the site level!)
+mean_synchrony_stability <- diversity_stability_synchrony %>% 
+  dplyr::group_by(habitat, site) %>% 
+  dplyr::summarise(
+    richness_mean = mean(richness_mean, na.rm = T),
+    functional_richness_mean = mean(functional_richness_mean, na.rm = T),
+    synchrony_mean = mean(synchrony, na.rm = T),
+    stability_mean = mean(cover_stability, na.rm = T),
+    cover_mean = mean(cover_mean, na.rm = T))
+```
 ### synchrony and cover model
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Cover\n", x = "Macroalgal Richness", title = "") +
+  labs(y = "Cover\n", x = "Taxonomic Richness", title = "") +
   trends_theme
@@
 
+### full vs mean plot comparison plot
+```{r ggpubr meanfull, fig.height = 8, fig.width=9}
+
+(mean_full_fxn_plot <- ggarrange(figure_3a, figure_3a_2, figure_3b, figure_3b_2, 
+                      ncol = 2, nrow = 2, 
+                      common.legend = TRUE,
+                      legend = "bottom", 
+                      font.label = list(size = 26, color = "black", face = "plain")))
+
+# ggsave(filename = "output/figure_3v2.png", figure_3, height = 8, width = 16)
+```
+```{r figure 3 with mean, fig.height = 8, fig.width=9}
+(figure_3_2 <- ggarrange(figure_3a_2, figure_3b_2, figure_3d, figure_3e, 
+                      ncol = 2, nrow = 2, 
+                      common.legend = TRUE,
+                      labels = c("a.", "b.", "c.", "d."),
+                      legend = "bottom", 
+                      font.label = list(size = 26, color = "black", face = "plain")))
+
+# ggsave(filename = "output/figure_3v3.png", figure_3_2, height = 8, width = 16)
+
+```
 # Figure 4
@@
 
+```{r mean plot richstab mod}
+rich_stab_mod_2 <- glmmTMB(stability_mean~ richness_mean*habitat , family = Gamma(link = "log"), data = mean_synchrony_stability)
+summary(rich_stab_mod_2)
+Anova(rich_stab_mod_2)
+#                          Chisq Df Pr(>Chisq)    
+# richness_mean         213.0187  1  < 2.2e-16 ***
+# habitat                29.9628  3  1.405e-06 ***
+# richness_mean:habitat   2.8556  3     0.4144    
+hist(residuals(rich_stab_mod_2)) #  little chunky
+plot(residuals(rich_stab_mod_2) ~ fitted(rich_stab_mod_2)) 
+performance::r2(rich_stab_mod_2) # 0.95
+emtrends(rich_stab_mod_2, pairwise ~ habitat, var = "richness_mean")  # no differences between sites
+# $emtrends
+#  habitat      richness_mean.trend     SE  df asymp.LCL asymp.UCL
+#  Fringing                   0.796 0.1060 Inf     0.588     1.004
+#  Backreef                   0.687 0.0924 Inf     0.506     0.868
+#  Forereef 10m               0.749 0.1130 Inf     0.528     0.971
+#  Forereef 17m               0.940 0.1210 Inf     0.703     1.178
+# $contrasts
+#  contrast                    estimate    SE  df z.ratio p.value
+#  Fringing - Backreef           0.1086 0.141 Inf   0.773  0.8669
+#  Fringing - Forereef 10m       0.0468 0.155 Inf   0.302  0.9905
+#  Fringing - Forereef 17m      -0.1446 0.161 Inf  -0.898  0.8057
+#  Backreef - Forereef 10m      -0.0619 0.146 Inf  -0.424  0.9745
+#  Backreef - Forereef 17m      -0.2532 0.152 Inf  -1.662  0.3442
+#  Forereef 10m - Forereef 17m  -0.1913 0.166 Inf  -1.155  0.6555
+```
+
 ```{r figure 4a setup}
@@
 filtered_rich_stab_effects <- filter_ranges(rich_stab_emm, dss_ranges, "habitat", "richness_mean")
+
+# V2
+rich_stab_emm_2 <- emmip(rich_stab_mod_2,
+                        habitat ~ richness_mean,
+                        at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_rich_stab_effects_2 <- filter_ranges(rich_stab_emm_2, dss_ranges_2, "habitat", "richness_mean")
+
 ```
@@
 # ggsave(filename = "output/figure_4a.png", figure_4a, height = 10, width = 14)
+
+(figure_4a_2 <- 
+  ggplot() +
+  geom_point(data = mean_synchrony_stability,
+             aes(x = richness_mean, y = stability_mean, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_rich_stab_effects_2, aes(x = richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
+  geom_ribbon(data = filtered_rich_stab_effects_2, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "Stability\n", x = "", title = "") +
+  trends_theme
+         )
+
+# ggsave(filename = "output/figure_4a_2.png", figure_4a_2, height = 10, width = 14)
 ```
@@
 
+```{r mean plot fg mod}
+rich_stab_mod_fg_2 <- glmmTMB(stability_mean ~ functional_richness_mean*habitat, family = Gamma(link = "log"), data = mean_synchrony_stability)
+summary(rich_stab_mod_fg_2)
+car::Anova(rich_stab_mod_fg_2)
+#                                     Chisq Df Pr(>Chisq)    
+# functional_richness_mean         194.7617  1    < 2e-16 ***
+# habitat                            8.9415  3    0.03008 *  
+# functional_richness_mean:habitat   0.8120  3    0.84659    
+hist(residuals(rich_stab_mod_fg_2)) # looks fine
+plot(residuals(rich_stab_mod_fg_2) ~ fitted(rich_stab_mod_fg_2)) # same as usual
+
+performance::r2(rich_stab_mod_fg_2) # 0.94
+emtrends(rich_stab_mod_fg_2, pairwise ~ habitat, var = "functional_richness_mean") # no differences among habitats
+# $emtrends
+#  habitat      functional_richness_mean.trend    SE  df asymp.LCL asymp.UCL
+#  Fringing                              0.945 0.128 Inf     0.694      1.20
+#  Backreef                              0.833 0.116 Inf     0.606      1.06
+#  Forereef 10m                          0.858 0.146 Inf     0.571      1.14
+#  Forereef 17m                          0.969 0.130 Inf     0.713      1.22
+# 
+# Confidence level used: 0.95 
+# 
+# $contrasts
+#  contrast                    estimate    SE  df z.ratio p.value
+#  Fringing - Backreef           0.1117 0.173 Inf   0.646  0.9170
+#  Fringing - Forereef 10m       0.0871 0.194 Inf   0.448  0.9700
+#  Fringing - Forereef 17m      -0.0235 0.183 Inf  -0.129  0.9992
+#  Backreef - Forereef 10m      -0.0246 0.187 Inf  -0.132  0.9992
+#  Backreef - Forereef 17m      -0.1353 0.175 Inf  -0.774  0.8661
+#  Forereef 10m - Forereef 17m  -0.1106 0.196 Inf  -0.565  0.9426
+
+```
+
 ```{r figure 4b plot setup}
@@
 filtered_rich_stab_effects_fg <- filter_ranges(rich_stab_emm_fg, dss_ranges, "habitat", "functional_richness_mean")
+
+## V2
+rich_stab_emm_fg_2 <- emmip(rich_stab_mod_fg_2,
+                        habitat ~ functional_richness_mean,
+                        at = list(functional_richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_rich_stab_effects_fg_2 <- filter_ranges(rich_stab_emm_fg_2, dss_ranges_2, "habitat", "functional_richness_mean")
 ```
@@
 #ggsave(filename = "output/figure_4b_v2.png", figure_4b, height = 10, width = 14)
+
+## V2
+(figure_4b_2 <- 
+  ggplot() +
+  geom_point(data = mean_synchrony_stability,
+             aes(x = functional_richness_mean, y = stability_mean, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_rich_stab_effects_fg_2, aes(x = functional_richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
+  geom_ribbon(data = filtered_rich_stab_effects_fg_2,
+              aes(x = functional_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "", x = "", title = "") +
+  trends_theme
+         )
 ```
@@
 
+```{r plot mean synchrony mod}
+synch_stab_mod_2 <- glmmTMB(stability_mean ~ synchrony_mean*habitat, family = Gamma("log"), data = mean_synchrony_stability)
+summary(synch_stab_mod_2)
+Anova(synch_stab_mod_2)
+#                          Chisq Df Pr(>Chisq)    
+# synchrony_mean         28.1431  1  1.127e-07 ***
+# habitat                 1.2421  3     0.7429    
+# synchrony_mean:habitat  5.3778  3     0.1461    
+hist(residuals(synch_stab_mod_2)) # a little chunky but looks good
+plot(residuals(synch_stab_mod_2) ~ fitted(synch_stab_mod_2)) # looks good
+
+performance::r2(synch_stab_mod_2) # 0.77
+emtrends(synch_stab_mod_2, pairwise ~ habitat, var = "synchrony_mean") 
+# $emtrends
+#  habitat      synchrony_mean.trend    SE  df asymp.LCL asymp.UCL
+#  Fringing                    -1.64 0.635 Inf     -2.89    -0.398
+#  Backreef                    -3.95 1.660 Inf     -7.19    -0.700
+#  Forereef 10m                -2.37 0.900 Inf     -4.13    -0.602
+#  Forereef 17m                -4.49 1.190 Inf     -6.82    -2.157
+# 
+# Confidence level used: 0.95 
+# 
+# $contrasts
+#  contrast                    estimate   SE  df z.ratio p.value
+#  Fringing - Backreef            2.304 1.77 Inf   1.299  0.5636
+#  Fringing - Forereef 10m        0.724 1.10 Inf   0.657  0.9131
+#  Fringing - Forereef 17m        2.846 1.35 Inf   2.111  0.1495
+#  Backreef - Forereef 10m       -1.580 1.89 Inf  -0.838  0.8363
+#  Backreef - Forereef 17m        0.542 2.04 Inf   0.266  0.9934
+#  Forereef 10m - Forereef 17m    2.122 1.49 Inf   1.423  0.4851
+```
+
 ```{r figure 4c setup}
@@
 filtered_synch_stab_effects <- filter_ranges(synch_stab_emm, dss_ranges, "habitat", "synchrony")
+
+## V2
+synch_stab_emm_2 <- emmip(synch_stab_mod_2,
+                        habitat ~ synchrony_mean,
+                        at = list(synchrony_mean = seq(0,1,0.01)), plotit = F, CIs = T)
+
+filtered_synch_stab_effects_2 <- filter_ranges(synch_stab_emm_2, dss_ranges_2, "habitat", "synchrony_mean")
 ```
@@
 # ggsave(filename = "output/figure_4c.png", figure_4c, height = 10, width = 14)
+
+## V2
+(figure_4c_2 <- 
+  ggplot() +
+  geom_point(data = mean_synchrony_stability,
+             aes(x = synchrony_mean, y = stability_mean, colour = habitat), size = 1, alpha = 0.5) + 
+  geom_line(data = filtered_synch_stab_effects_2, aes(x = synchrony_mean, y = exp(yvar), colour = habitat), size = 1.5) +
+  geom_ribbon(data = filtered_synch_stab_effects_2, aes(x = synchrony_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+              alpha = 0.3) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(x = "", y = "", title = "") +
+  trends_theme
+         )
 ```
@@
 # ggsave(filename = "output/figure_4v2.png", figure_4, height = 8, width = 16)
+```
+
+### compare plot level graphs
+```{r ggpubr fig4, fig.height = 8, fig.width=16}
+
+(mean_plot_graphs <- ggarrange(figure_4a, figure_4b, figure_4c, figure_4a_2, figure_4b_2, figure_4c_2, 
+                      ncol = 3, nrow = 2, 
+                      common.legend = TRUE,
+                      heights = c(1,1),
+                      widths = c(1,1,1),
+                      legend = "bottom", 
+                      labels = c("a.", "b.", "c.", "d.", "e.", "f."),
+                      font.label = list(size = 26, color = "black", face = "plain")))
+
+# ggsave(filename = "output/plot_dss_graphs.png", mean_plot_graphs, height = 8, width = 16)
+```
+### DSS graph with mean plot
+```{r ggpubr fig4, fig.height = 8, fig.width=16}
+(figure_4_2 <- ggarrange(figure_4a_2, figure_4b_2, figure_4c_2, figure_4d, figure_4e, figure_4f, 
+                      ncol = 3, nrow = 2, 
+                      common.legend = TRUE,
+                      heights = c(1,1),
+                      widths = c(1,1,1),
+                      legend = "bottom", 
+                      labels = c("a.", "b.", "c.", "d.", "e.", "f."),
+                      font.label = list(size = 26, color = "black", face = "plain")))
+
+# ggsave(filename = "output/figure_4v3.png", figure_4_2, height = 8, width = 16)
+
 ```
@@
 
-# calculate mean synchrony within a site (!= synchrony of species at the site level!)
-mean_synchrony_stability <- diversity_stability_synchrony %>% 
-  dplyr::group_by(habitat, site) %>% 
-  dplyr::summarise(synchrony_mean = mean(synchrony, na.rm = T),
-            stability_mean = mean(cover_stability, na.rm = T))
-
 # merge with site level dss metrics

### 2025-01-21 manuscript_1_code.Rmd
orgranized code and wrote new csvs for potential SEM

@@
 diversity_stability_synchrony$synchrony_trans <- sv_trans(diversity_stability_synchrony$synchrony)
-
-dss_ranges <- extract_ranges(diversity_stability_synchrony, "habitat",
-                             c("richness_mean", "functional_richness_mean", "cover_stability", "synchrony"))
-
-# repeat for the mean of each plot
-dss_ranges_2 <- extract_ranges(mean_synchrony_stability, "habitat",
-                             c("richness_mean", "functional_richness_mean", "synchrony_mean", "stability_mean", "cover_mean"))
-```
-
-### mean plot-level dss setup
-```{r mean plot dss}
+# write.csv(file = here::here("data", "full_plot_level_dss.csv"), diversity_stability_synchrony)
+
 # calculate mean synchrony within a site (!= synchrony of species at the site level!)
@@
     cover_mean = mean(cover_mean, na.rm = T))
-```
+
+# write.csv(file = here::here("data", "mean_plot_level_dss.csv"), mean_synchrony_stability)
+dss_ranges <- extract_ranges(diversity_stability_synchrony, "habitat",
+                             c("richness_mean", "functional_richness_mean", "cover_stability", "synchrony"))
+
+# repeat for the mean of each plot
+dss_ranges_2 <- extract_ranges(mean_synchrony_stability, "habitat",
+                             c("richness_mean", "functional_richness_mean", "synchrony_mean", "stability_mean", "cover_mean"))
+```
+
 ### synchrony and cover model
@@
 dss_spatial <- merge(diversity_stability_site, spatial_synchrony, by = "site_habitat")
-dss_spatial_2 <- merge(dss_spatial, mean_synchrony_stability, by = c("habitat", "site"))
-
+dss_spatial_2 <- merge(dss_spatial, mean_synchrony_stability[, which(names(mean_synchrony_stability) %in%
+                                                               c("habitat", "site", "synchrony_mean", "stability_mean"))],
+                       by = c("habitat", "site"))
+# write.csv(file = here::here("data", "spatial_synchrony.csv"), dss_spatial_2)
 dss_spatial_ranges <- extract_ranges(dss_spatial_2, "habitat", c("spatial_synchrony", "synchrony_mean", "stability_mean"))

### 2025-01-21 SEM_site.Rmd

---
title: "SEM with site-level algae data"
output: html_document
date: "2025-01-21"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clear workspace
rm(list = ls())

# load needed packages
require(tidyverse)
require(here)
require(piecewiseSEM)
require(lme4) # glmer() function
require(performance)
```

# Read in data

# Read in data

```{r}
# spatial synchrony data from Noam & functional richness--need to double check that FR is site-level
spatial_synch <- read_csv(here("data/spatial_synchrony.csv")) %>% 
  select(habitat, site, functional_richness_mean, spatial_synchrony)

# site-level data summarized with one point for all time per site
algae_site <- read_csv(here("data/diversity_stability_synchrony_site.csv")) %>% 
  full_join(spatial_synch)
```

# Scale data

Gamma families are currently not supported and we're not able to get glmmTMB to work for the time being.

```{r}
algae_site %>% 
  mutate(
    richness_mean_scale = scale(algae_site$richness_mean)[,1],
    functional_richness_mean_scale = scale(algae_site$functional_richness_mean)[,1],
    cover_mean_scale = scale(algae_site$cover_mean)[,1],
    cover_stability_scale = scale(algae_site$cover_stability)[,1],
    synchrony_scale = scale(algae_site$synchrony)[,1],
    spatial_synchrony_scale = scale(algae_site$spatial_synchrony)[,1]
  ) -> algae_site_scaled
```

# Make a simple SEM

First, fit individual models

```{r}
# effect of richness on synchrony
m.sprich.synch <- glmer(synchrony_mean_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site) + (1|habitat),
                        data = algae_site_scaled)
  performance::check_model(m.sprich.synch)

# effect of synchrony on stability
m.synch.stab.cov <- glmer(cover_stability_scale ~ synchrony_scale + functional_richness_mean_scale + cover_mean_scale + richness_mean_scale + spatial_synchrony_scale + (1|site) + (1|habitat),
                      data = algae_site_scaled)  
  performance::check_model(m.synch.stab.cov)

# effect of richness on cover
m.sprich.cov <- glmer(cover_mean_scale ~ richness_mean_scale + (1|site) + (1|habitat),
                      data = algae_site_scaled)
  performance::check_model(m.sprich.cov)
```

Make it with piecewiseSEM

```{r}
# Build SEM
sem <- psem(
  m.sprich.synch,
  m.synch.stab.cov,
  m.sprich.cov,
  data = as.data.frame(algae_site_scaled)
)

summary(sem)
```

Plot

```{r}
plot(sem)
```

### 2025-01-24 manuscript_1_code.Rmd
starting to add cover as area calculations

@@
 
+# convert cover to area
+macro_long_data$area <- macro_long_data$prop_cover*0.25
 # write_csv(macro_long_data, here("Data", "macro_long_data.csv"))
@@
 ```{r site macroalgal data}
+# NOTE - FRI JAN 24 - updated mean() in summarize across to sum() to reflect new method of calculating cover as area
 site_macro <- macroalgae %>%
   dplyr::group_by(year, site, habitat,site_habitat) %>%
-  dplyr::summarize(across(acanthophora_spicifera:valonia_ventricosa, \(x) mean(x, na.rm = TRUE)))
+  dplyr::summarize(across(acanthophora_spicifera:valonia_ventricosa, \(x) sum(x, na.rm = TRUE)))
 
@@
   select(-macroalgae)
+
+alpha_diversity_quad_macro$area <- alpha_diversity_quad_macro$macroalgal_prop_cover*0.25
 ```
@@
 alpha_diversity_site_macro$cover_trans <- sv_trans(alpha_diversity_site_macro$prop_cover)
+
+alpha_diversity_site_macro$area <- alpha_diversity_site_macro$prop_cover*0.25
 ```

### 2025-01-24 SEM_quad.Rmd

@@
 
+# clear workspace
+rm(list = ls())
+
 require(tidyverse)
@@
 ```{r}
+# functional richness from Noam
+functrich <- read_csv(here("data/full_plot_level_dss.csv"))
+
 # quad-level data summarized with one point for all time
-algae <- read_csv(here("data/diversity_stability_synchrony.csv"))
+algae_1 <- read_csv(here("data/diversity_stability_synchrony.csv"))
+```
+
+These two guys have different dimensions. See where they differ:
+
+```{r}
+functrich[!(functrich$location %in% algae_1$location),] 
+```
+
+Noam says we can just inner join/exclude these--just used to calculate averages
+
+```{r}
+functrich %>% 
+  select(location, habitat, site, functional_richness_mean) %>% 
+  inner_join(algae_1) -> algae
 ```
@@
     richness_mean_scale = scale(algae$richness_mean)[,1],
+    functional_richness_mean_scale = scale(algae$functional_richness_mean)[,1],
     cover_mean_scale = scale(algae$cover_mean)[,1],
@@
 
-# Make a simple SEM
+# Make an overall SEM with functional richness
 
@@
 # effect of richness on synchrony
-m.sprich.synch <- glmer(synchrony_scale ~ richness_mean_scale + (1|site) + (1|habitat),
+m.sprich.synch <- glmer(synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site) + (1|habitat),
                         data = algae_scaled)
@@
 
+# effect of taxonomic richness on functional richness
+m.sprich.funrich <- glmer(functional_richness_mean_scale ~ richness_mean_scale + (1|site) + (1|habitat),
+                        data = algae_scaled)
+  performance::check_model(m.sprich.funrich)
+
 # effect of synchrony on stability
-m.synch.stab.cov <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site) + (1|habitat),
+m.synch.stab.cov <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + functional_richness_mean_scale + (1|site) + (1|habitat),
                       data = algae_scaled)  
-  performance::check_model(m.synch.stab)
+  performance::check_model(m.synch.stab.cov)
 
@@
   m.sprich.synch,
+  m.sprich.funrich,
   m.synch.stab.cov,
   m.sprich.cov,
-  data = as.data.frame(algae)
+  data = as.data.frame(algae_scaled)
 )
@@
 ```{r}
-# Extract coefficients
-#coefficients <- summary(sem)$coefficients
+plot(sem)
+```
 
-# Add a column for color based on the sign of the path coefficient
-#coefficients$color <- ifelse(coefficients$Estimate > 0, "blue", #"red")
+# Make an overall SEM withOUT functional richness
 
-plot(sem)
+First, fit individual models
+
+```{r}
+# effect of richness on synchrony
+m.sprich.synch.noFR <- glmer(synchrony_scale ~ richness_mean_scale  + (1|site) + (1|habitat),
+                        data = algae_scaled)
+  performance::check_model(m.sprich.synch.noFR)
+
+# effect of synchrony on stability
+m.synch.stab.cov.noFR <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site) + (1|habitat),
+                      data = algae_scaled)  
+  performance::check_model(m.synch.stab.cov.noFR)
+
+# effect of richness on cover
+m.sprich.cov.noFR <- glmer(cover_mean_scale ~ richness_mean_scale + (1|site) + (1|habitat),
+                      data = algae_scaled)
+  performance::check_model(m.sprich.cov.noFR)
+```
+
+Try it with piecewiseSEM
+
+```{r}
+# Build SEM
+sem.noFR <- psem(
+  m.sprich.synch,
+  m.synch.stab.cov,
+  m.sprich.cov,
+  data = as.data.frame(algae_scaled)
+)
+
+summary(sem.noFR)
+```
+
+Plot
+
+```{r}
+plot(sem.noFR)
 ```
@@
 # effect of richness on synchrony
-m.sprich.synch.FR <- glmer(synchrony_scale ~ richness_mean_scale + (1|site),
+m.sprich.synch.FR <- glmer(synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site),
                         data = algae_scaled[algae_scaled$habitat == "Fringing", ])
@@
 
+# effect of taxonomic richness on functional richness
+m.sprich.funrich.FR <- glmer(functional_richness_mean_scale ~ richness_mean_scale + (1|site),
+                        data = algae_scaled[algae_scaled$habitat == "Fringing", ])
+  performance::check_model(m.sprich.funrich.FR)
+
 # effect of synchrony on stability
-m.synch.stab.cov.FR <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site),
+m.synch.stab.cov.FR <- glmer(cover_stability_scale ~ synchrony_scale + functional_richness_mean_scale +  cover_mean_scale + richness_mean_scale + (1|site),
                       data = algae_scaled[algae_scaled$habitat == "Fringing", ])  
@@
   m.sprich.synch.FR,
+  m.sprich.funrich.FR,
   m.synch.stab.cov.FR,
@@
 # effect of richness on synchrony
-m.sprich.synch.BA <- glmer(synchrony_scale ~ richness_mean_scale + (1|site),
+m.sprich.synch.BA <- glmer(synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site),
                         data = algae_scaled[algae_scaled$habitat == "Backreef", ])
@@
 
+# effect of taxonomic richness on functional richness
+m.sprich.funrich.BA <- glmer(functional_richness_mean_scale ~ richness_mean_scale + (1|site),
+                        data = algae_scaled[algae_scaled$habitat == "Backreef", ])
+  performance::check_model(m.sprich.funrich.BA)
+
 # effect of synchrony on stability
-m.synch.stab.cov.BA <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site),
+m.synch.stab.cov.BA <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + functional_richness_mean_scale + richness_mean_scale + (1|site),
                       data = algae_scaled[algae_scaled$habitat == "Backreef", ])  
@@
   m.sprich.synch.BA,
+  m.sprich.funrich.BA,
   m.synch.stab.cov.BA,
@@
 # effect of richness on synchrony
-m.sprich.synch.FO10 <- glmer(synchrony_scale ~ richness_mean_scale + (1|site),
+m.sprich.synch.FO10 <- glmer(synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site),
                         data = algae_scaled[algae_scaled$habitat == "Forereef 10m", ])
@@
 
+# effect of taxonomic richness on functional richness
+m.sprich.funrich.FO10 <- glmer(functional_richness_mean_scale ~ richness_mean_scale + (1|site),
+                        data = algae_scaled[algae_scaled$habitat == "Forereef 10m", ])
+  performance::check_model(m.sprich.funrich.FO10)
+
 # effect of synchrony on stability
-m.synch.stab.cov.FO10 <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site),
+m.synch.stab.cov.FO10 <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + functional_richness_mean_scale + (1|site),
                       data = algae_scaled[algae_scaled$habitat == "Forereef 10m", ])  
@@
   m.sprich.synch.FO10,
+  m.sprich.funrich.FO10,
   m.synch.stab.cov.FO10,
@@
 # effect of richness on synchrony
-m.sprich.synch.FO17 <- glmer(synchrony_scale ~ richness_mean_scale + (1|site),
+m.sprich.synch.FO17 <- glmer(synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site),
                         data = algae_scaled[algae_scaled$habitat == "Forereef 17m", ])
@@
 
+# effect of taxonomic richness on functional richness
+m.sprich.funrich.FO17 <- glmer(functional_richness_mean_scale ~ richness_mean_scale + (1|site),
+                        data = algae_scaled[algae_scaled$habitat == "Forereef 17m", ])
+  performance::check_model(m.sprich.funrich.FO17)
+
 # effect of synchrony on stability
-m.synch.stab.cov.FO17 <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site),
+m.synch.stab.cov.FO17 <- glmer(cover_stability_scale ~ synchrony_scale + functional_richness_mean_scale + cover_mean_scale + richness_mean_scale + (1|site),
                       data = algae_scaled[algae_scaled$habitat == "Forereef 17m", ])  
@@
   m.sprich.synch.FO17,
+  m.sprich.funrich.FO17,
   m.synch.stab.cov.FO17,

### 2025-01-24 SEM_quad_lavaan.Rmd

---
title: "SEM for quad-level data using lavaan"
output: html_document
date: "2025-01-22"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clear workspace
rm(list = ls())

require(tidyverse)
require(here)
require(lavaan)
require(lavaanPlot)
```

# Read in data

```{r}
# functional richness from Noam
functrich <- read_csv(here("data/full_plot_level_dss.csv")) %>% 
  select(location, habitat, site, functional_richness_mean)

# quad-level data summarized with one point for all time
algae <- read_csv(here("data/diversity_stability_synchrony.csv")) %>% 
  inner_join(functrich)
```

# Scale data

```{r}
algae %>% 
  mutate(
    richness_mean_scale = scale(algae$richness_mean)[,1],
    functional_richness_mean_scale = scale(algae$functional_richness_mean)[,1],
    cover_mean_scale = scale(algae$cover_mean)[,1],
    cover_stability_scale = scale(algae$cover_stability)[,1],
    synchrony_scale = scale(algae$synchrony)[,1]
  ) %>% 
  select(site, habitat, richness_mean_scale, functional_richness_mean_scale,
         cover_mean_scale, cover_stability_scale, synchrony_scale
         ) -> algae_scaled
```

# Overall model

```{r define model}
# define the model
m.all <- '
  # Structural model/regressions
  synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale
  functional_richness_mean_scale ~ richness_mean_scale
  cover_stability_scale ~ synchrony_scale + functional_richness_mean_scale + cover_mean_scale + richness_mean_scale
  cover_mean_scale ~ richness_mean_scale
'   
```

```{r examine fit}
fit.all <- lavaan::sem(m.all, cluster = "site", data = algae_scaled) 
  lavaan::summary(fit.all, standardized = TRUE, fit.measures = TRUE)
```

```{r}
# make labels
labels1 <- list(synchrony_scale = "Synchrony", richness_mean_scale = "Spp richness", functional_richness_mean_scale = "Func. Richness", cover_stability_scale = "Stability", cover_mean_scale = "Cover")

# plot
lavaanPlot(model = fit.all,
           labels = labels1,
           coefs = TRUE,
           stand = TRUE,
           sig = 0.05 # standardized regression paths, showing only paths with p<= .05
           )
```

# Overall modell: no functional richness

```{r define model}
# define the model
m.all.noFR <- '
  # Structural model/regressions
  synchrony_scale ~ richness_mean_scale 
  cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale
  cover_mean_scale ~ richness_mean_scale
'   
```

```{r examine fit}
fit.noFR <- lavaan::sem(m.all.noFR, cluster = "site", data = algae_scaled) 
  lavaan::summary(fit.noFR, standardized = TRUE, fit.measures = TRUE)
```

```{r}
# make labels
labels2 <- list(synchrony_scale = "Synchrony", richness_mean_scale = "Spp richness", cover_stability_scale = "Stability", cover_mean_scale = "Cover")

# plot
lavaanPlot(model = fit.noFR,
           labels = labels2,
           coefs = TRUE,
           stand = TRUE,
           sig = 0.05 # standardized regression paths, showing only paths with p<= .05
           )
```

# Overall model: no cover

```{r define model}
# define the model
m.noC <- '
  # Structural model/regressions
  synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale
  functional_richness_mean_scale ~ richness_mean_scale
  cover_stability_scale ~ synchrony_scale + functional_richness_mean_scale + richness_mean_scale
'   
```

```{r examine fit}
fit.noC <- lavaan::sem(m.noC, cluster = "site", data = algae_scaled) 
  lavaan::summary(fit.noC, standardized = TRUE, fit.measures = TRUE)
```

```{r}
# make labels
labels3 <- list(synchrony_scale = "Synchrony", richness_mean_scale = "Spp richness", functional_richness_mean_scale = "Func. Richness", cover_stability_scale = "Stability")

# plot
lavaanPlot(model = fit.noC,
           labels = labels3,
           coefs = TRUE,
           stand = TRUE,
           sig = 0.05 # standardized regression paths, showing only paths with p<= .05
           )
```

# Overall model: no cover or functional richness

```{r define model}
# define the model
m.noCorFR <- '
  # Structural model/regressions
  synchrony_scale ~ richness_mean_scale 
  cover_stability_scale ~ synchrony_scale + richness_mean_scale
'   
```

```{r examine fit}
fit.noCorF <- lavaan::sem(m.noCorFR, cluster = "site", data = algae_scaled) 
  lavaan::summary(fit.noCorF, standardized = TRUE, fit.measures = TRUE)
```

```{r}
# make labels
labels4 <- list(synchrony_scale = "Synchrony", richness_mean_scale = "Spp richness",  cover_stability_scale = "Stability")

# plot
lavaanPlot(model = fit.noCorF,
           labels = labels4,
           coefs = TRUE,
           stand = TRUE,
           sig = 0.05 # standardized regression paths, showing only paths with p<= .05
           )
```

# Fringing model: no cover or functional richness

```{r define model}
# define the model
m.noCorFR.FR <- '
  # Structural model/regressions
  synchrony_scale ~ richness_mean_scale 
  cover_stability_scale ~  richness_mean_scale + synchrony_scale
'   
```

```{r examine fit}
fit.noCorF.FR <- lavaan::sem(m.noCorFR.FR, cluster = "site", data = algae_scaled) # [algae_scaled$habitat == "Fringing", ] 
  lavaan::summary(fit.noCorF.FR, standardized = TRUE, fit.measures = TRUE)
```

```{r}
# make labels
labels4 <- list(synchrony_scale = "Synchrony", richness_mean_scale = "Spp richness", cover_stability_scale = "Stability")

# plot
lavaanPlot(model = fit.noCorF.FR,
           labels = labels4,
           coefs = TRUE,
           stand = TRUE,
           sig = 0.05 # standardized regression paths, showing only paths with p<= .05
           )
```

### 2025-01-24 SEM_site.Rmd

@@
 ```
-
-# Read in data
 
@@
 
-# Make a simple SEM
+# Make an overall SEM
 
@@
 # effect of richness on synchrony
-m.sprich.synch <- glmer(synchrony_mean_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site) + (1|habitat),
+m.sprich.synch <- glmer(synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site) + (1|habitat),
                         data = algae_site_scaled)
   performance::check_model(m.sprich.synch)
+
+# effect of taxonomic richness on functional richness
+m.sprich.funrich <- glmer(functional_richness_mean_scale ~ richness_mean_scale + (1|site) + (1|habitat),
+                        data = algae_site_scaled)
+  performance::check_model(m.sprich.funrich)
 
@@
   m.sprich.cov,
+  m.sprich.funrich,
   data = as.data.frame(algae_site_scaled)
@@
 ```
+
+Poor convergence issues/singularity
+
+# What about no random effect for habitat?
+
+```{r}
+# effect of richness on synchrony
+m.sprich.synch.nR <- glmer(synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site),
+                        data = algae_site_scaled)
+  performance::check_model(m.sprich.synch.nR)
+
+# effect of taxonomic richness on functional richness
+m.sprich.funrich.nR <- glmer(functional_richness_mean_scale ~ richness_mean_scale + (1|site),
+                        data = algae_site_scaled)
+  performance::check_model(m.sprich.funrich.nR)
+
+# effect of synchrony on stability
+m.synch.stab.cov.nR <- glmer(cover_stability_scale ~ synchrony_scale + functional_richness_mean_scale + cover_mean_scale + richness_mean_scale + spatial_synchrony_scale + (1|site),
+                      data = algae_site_scaled)  
+  performance::check_model(m.synch.stab.cov.nR)
+
+# effect of richness on cover
+m.sprich.cov.nR <- glmer(cover_mean_scale ~ richness_mean_scale + (1|site),
+                      data = algae_site_scaled)
+  performance::check_model(m.sprich.cov.nR)
+```
+
+Make it with piecewiseSEM
+
+```{r}
+# Build SEM
+sem.nR <- psem(
+  m.sprich.synch.nR,
+  m.synch.stab.cov.nR,
+  m.sprich.cov.nR,
+  m.sprich.funrich.nR,
+  data = as.data.frame(algae_site_scaled)
+)
+
+summary(sem.nR)
+```
+
+Plot
+
+```{r}
+plot(sem.nR)
+```

### 2025-01-28 manuscript_1_code.Rmd
added initial area model at quad level

@@
                                     by = c("location", "year", "site", "habitat"))
+# add small amount to avoid 0s in data for gamma modeling
+alpha_diversity_quad_macro$area_trans <- alpha_diversity_quad_macro$area + 10E-10
 
@@
     functional_richness_mean = mean(functional_richness, na.rm = T),
-    cover_mean = mean(macroalgal_prop_cover, na.rm = T))
+    cover_mean = mean(macroalgal_prop_cover, na.rm = T),
+    area_mean = mean(area_trans, na.rm = T))
 
@@
 # Forereef 10m - Forereef 17m  0.04076 0.0209 Inf   1.947  0.2087
+
+# area model - play around with area vs. area_trans
+area_mod <- glmmTMB(area ~ richness*habitat + (1|site/location) + (1|year), family = ziGamma("log"), ziformula = ~1, data = alpha_diversity_quad_macro)
+summary(area_mod)
+Anova(area_mod)
+hist(residuals(area_mod))
+plot(simulateResiduals(area_mod))
+plot(residuals(area_mod) ~ fitted(area_mod)) # heteroskedastic - because of the few 100% cover plots?
+r.squaredGLMM(area_mod) # 0.631, 0.650
+performance::r2(area_mod) # 0.635, 0.654
+emtrends(area_mod, pairwise ~ habitat, var = "richness")
 

### 2025-01-28 manuscript_1_code.Rmd
further specified how to run model

@@
 
-# area model - play around with area vs. area_trans
+# area model - play around with area vs. area_trans. If you use area_trans, family = Gamma("log") and remove ziformula argument
 area_mod <- glmmTMB(area ~ richness*habitat + (1|site/location) + (1|year), family = ziGamma("log"), ziformula = ~1, data = alpha_diversity_quad_macro)

### 2025-02-05 manuscript_1_code.Rmd
removed area calculations, updated functional group csv

--- 2025-01-24 manuscript_1_code.Rmd
@@
 
-# convert cover to area
-macro_long_data$area <- macro_long_data$prop_cover*0.25
-# write_csv(macro_long_data, here("Data", "macro_long_data.csv"))
 ```
@@
 ```{r site macroalgal data}
-# NOTE - FRI JAN 24 - updated mean() in summarize across to sum() to reflect new method of calculating cover as area
+
 site_macro <- macroalgae %>%
   dplyr::group_by(year, site, habitat,site_habitat) %>%
-  dplyr::summarize(across(acanthophora_spicifera:valonia_ventricosa, \(x) sum(x, na.rm = TRUE)))
+  dplyr::summarize(across(acanthophora_spicifera:valonia_ventricosa, \(x) mean(x, na.rm = TRUE)))
 
@@
 
-alpha_diversity_quad_macro$area <- alpha_diversity_quad_macro$macroalgal_prop_cover*0.25
 ```
@@
 
-alpha_diversity_site_macro$area <- alpha_diversity_site_macro$prop_cover*0.25
 ```
@@
 ```{r read and prep functional group data}
-algae_fg_raw2 <- read_csv(here("data","algae_functional_groupings_092724.csv"))
+algae_fg_raw2 <- read_csv(here("data","MCR_Algal_Functional_Groups_02042025_updatedtax.csv"))
 
@@
 algae_fg$species <- gsub(" ", "_", tolower(algae_fg$species))
-algae_fg$morphology <- trimws(algae_fg$morphology)
+algae_fg$morphology <- trimws(algae_fg$functional_grouping_updated)
 algae_fg$morphology <- gsub(" ", "_", algae_fg$morphology)

### 2025-03-06 manuscript_1_code.Rmd
Corrected main figure aesthetics

@@
         axis.text.y = element_text(size=22),
-        legend.text = element_text(size=22),
+        legend.text = element_text(size=24),
         plot.title = element_text(size=26), 
@@
   geom_point(data = alpha_diversity_quad_macro,
-             aes(x = richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.5) + 
+             aes(x = richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.75) + 
   geom_line(data = filtered_cover_effects, aes(x = richness, y = inv_logit(yvar), colour = habitat)) +
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Cover\n", x = "", title = "") +
+  labs(y = "Cover\n", x = "Taxonomic richness", title = "") +
   model_themes 
@@
   geom_point(data = alpha_diversity_quad_macro,
-             aes(x = functional_richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.5) + 
+             aes(x = functional_richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.75) + 
   geom_line(data = filtered_cover_effects_fg, aes(x = functional_richness, y = inv_logit(yvar), colour = habitat)) +
   geom_ribbon(data = filtered_cover_effects_fg, aes(x = functional_richness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
-              alpha = 0.3) +
-  scale_colour_manual(values = habitat_colours) +
-  scale_fill_manual(values = habitat_colours) +
-  labs(y = "", x = "", title = "") +
-  trends_theme
+              alpha = 0.5) +
+  scale_colour_manual(values = habitat_colours) +
+  scale_fill_manual(values = habitat_colours) +
+  labs(y = "", x = "Functional richness", title = "") +
+  model_themes
          )
@@
 ## Figure 3 Full Plot
-```{r ggpubr fig3, fig.height = 8, fig.width=16}
-
-(figure_3 <- ggarrange(figure_3a, figure_3b, figure_3c, figure_3d, figure_3e, figure_3f, 
-                      ncol = 3, nrow = 2, 
+```{r ggpubr fig3 main, fig.height = 8, fig.width=16}
+
+(figure_3 <- ggarrange(figure_3a, figure_3b,  
+                      ncol = 2, nrow = 1, 
                       common.legend = TRUE,
-                      labels = c("a.", "b.", "c.", "d.", "e.", "f."),
+                      labels = c("a.", "b."),
+                      legend = "bottom", 
+                      font.label = list(size = 26, color = "black", face = "plain")) )
+
+# ggsave(filename = "output/figure_3v2.png", figure_3, height = 8, width = 16)
+```
+
+```{r figure 3 site supplement, fig.height = 8, fig.width = 16}
+(figure_3_supplement <- ggarrange(figure_3d, figure_3e,  
+                      ncol = 2, nrow = 1, 
+                      common.legend = TRUE,
+                      labels = c("a.", "b."),
                       legend = "bottom", 
@@
 
-# ggsave(filename = "output/figure_3v2.png", figure_3, height = 8, width = 16)
-```
-
-### full vs mean plot comparison plot
-```{r ggpubr meanfull, fig.height = 8, fig.width=9}
-
-(mean_full_fxn_plot <- ggarrange(figure_3a, figure_3a_2, figure_3b, figure_3b_2, 
-                      ncol = 2, nrow = 2, 
-                      common.legend = TRUE,
-                      legend = "bottom", 
-                      font.label = list(size = 26, color = "black", face = "plain")))
-
-# ggsave(filename = "output/figure_3v2.png", figure_3, height = 8, width = 16)
-```
-```{r figure 3 with mean, fig.height = 8, fig.width=9}
-(figure_3_2 <- ggarrange(figure_3a_2, figure_3b_2, figure_3d, figure_3e, 
-                      ncol = 2, nrow = 2, 
-                      common.legend = TRUE,
-                      labels = c("a.", "b.", "c.", "d."),
-                      legend = "bottom", 
-                      font.label = list(size = 26, color = "black", face = "plain")))
-
-# ggsave(filename = "output/figure_3v3.png", figure_3_2, height = 8, width = 16)
-
-```
+# ggsave(filename = "output/figure_3_site.png", figure_3_supplement, height = 8, width = 16)
+```
+
 # Figure 4
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Stability\n", x = "", title = "") +
-  trends_theme
+  labs(y = "Stability\n", x = "Taxonomic richness", title = "") +
+  model_themes
          )
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "", x = "", title = "") +
-  trends_theme
+  scale_x_continuous(breaks = c(0,1,2)) +
+  labs(y = "", x = "Functional richness", title = "") +
+  model_themes
          )
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(x = "", y = "", title = "") +
-  trends_theme
+  labs(x = "Synchrony", y = "", title = "") +
+  model_themes
          )
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "Stability\n", x = "Taxonomic Richness", title = "") +
-  trends_theme
+  scale_y_continuous(limits = c(0.5, 4.2), breaks = c(1,2,3,4)) +
+  labs(y = "Site-level Stability\n", x = "Taxonomic richness", title = "") +
+  model_themes
          )
@@
   scale_fill_manual(values = habitat_colours) +
-  labs(y = "", x = "Functional Richness", title = "") +
-  trends_theme
+  scale_y_continuous(limits = c(0.5, 4.2), breaks = c(1,2,3,4)) +
+  scale_x_continuous(limits = c(2, 4.2), breaks = c(1,2,3,4)) +
+  labs(y = "", x = "Functional richness", title = "") +
+  model_themes
          )
@@
   labs(x = "Synchrony", y = "", title = "") +
-  trends_theme
+  scale_y_continuous(limits = c(0.5, 4.2), breaks = c(1,2,3,4)) +
+  model_themes
          )
@@
 ## Figure 4 Full Plot
-```{r ggpubr fig4, fig.height = 8, fig.width=16}
-
-(figure_4 <- ggarrange(figure_4a, figure_4b, figure_4c, figure_4d, figure_4e, figure_4f, 
-                      ncol = 3, nrow = 2, 
+```{r ggpubr fig4, fig.height = 6, fig.width=18}
+
+(figure_4 <- ggarrange(figure_4a, figure_4b, figure_4c, 
+                      ncol = 3, nrow = 1, 
                       common.legend = TRUE,
@@
 
-# ggsave(filename = "output/figure_4v2.png", figure_4, height = 8, width = 16)
+# ggsave(filename = "output/figure_4v3.png", figure_4, height = 6, width = 18)
 ```
@@
   labs(y = "", x = "Spatial synchrony", title = "") +
-  trends_theme
+  scale_y_continuous(limits = c(0.5, 4.2), breaks = c(1,2,3,4)) +
+  model_themes
          )
@@
   labs(y = "Site-level Stability\n", x = "Mean plot-level species synchrony", title = "") +
-  trends_theme
+  scale_y_continuous(limits = c(0.5, 4.2), breaks = c(1,2,3,4)) +
+  model_themes
          )
@@
   labs(y = "", x = "Plot-level stability", title = "") +
-  trends_theme
+  scale_y_continuous(limits = c(0.5, 4.2), breaks = c(1,2,3,4)) +
+  model_themes
          )
@@
 ## panel graph
-```{r figure 5 panel, fig.height = 6, fig.width=22}
-(figure_5 <- ggarrange(mean_synchrony_plot, alpha_gamma_stability_plot, spatial_synchrony_plot,
-                      ncol = 3, nrow = 1, 
+```{r figure 5 panel, fig.height = 10, fig.width=18}
+(figure_5 <- ggarrange(figure_4d, figure_4e, figure_4f, 
+                       mean_synchrony_plot, alpha_gamma_stability_plot, spatial_synchrony_plot,
+                      ncol = 3, nrow = 2, 
                       common.legend = TRUE,
                       legend = "bottom", 
-                      labels = c("a.", "b.", "c."),
+                      labels = c("a.", "b.", "c.", "d.", "e.", "f."),
                       font.label = list(size = 26, color = "black", face = "plain")))
 
-# ggsave(filename = "output/figure_5_option3_v2.png", figure_5, height = 6, width = 22)
+# ggsave(filename = "output/figure_5_v2.png", figure_5, height = 10, width = 18)
 ```

### 2025-03-10 supplemental_table_figures.Rmd
Code for turning supplemental tables into figures

---
title: "Supplemental tables to figures"
output: html_document
date: "2025-03-10"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(here) # relative file paths
require(tidyverse) # data wrangling
require(patchwork) # putting plots together
```

# Load data

```{r}
supp_tables <- read_csv(here("data/supplemental_tables.csv"))
```

# Set plotting colors and plotting theme

```{r}
habitat_colours <- c("Fringing" = "#D81B60",
                     "Backreef" = "#FFC107",
                     "Forereef 10m" = "#1E88E5",
                     "Forereef 17m" = "#004D40")

model_themes <-   theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.title.y = element_text(size=24), 
        axis.text.x = element_text(size=22), 
        axis.text.y = element_text(size=22),
        legend.text = element_text(size=24),
        # add the same size for facet_wrap titles
        strip.text = element_text(size = 24),
        plot.title = element_text(size=26), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
```

# Plot S2

First make the plot-level panel:

```{r}
supp_tables %>% 
  filter(Table == "S2") %>% 
  filter(`Spatial scale` == "Plot-level") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  facet_wrap(~ Predictor) +
  scale_colour_manual(values = habitat_colours) +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(a) Plot-level") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.03 , y = Habitat, 
                label = Letter), colour = "black", size = 10) +
  theme(legend.position = "none") -> s2.plt.level
```

Then the site-level:

```{r}
supp_tables %>% 
  filter(Table == "S2") %>% 
  filter(`Spatial scale` == "Site-level") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  facet_wrap(~ Predictor) +
  scale_colour_manual(values = habitat_colours) +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(b) Site-level") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.03 , y = Habitat, 
                label = Letter), colour = "black", size = 10) -> s2.site.level
```

Plot together using `patchwork`

```{r}
pdf(here("output/tableS2toFig.pdf"), height = 12, width = 13)

s2.plt.level / s2.site.level

dev.off()
```

# Plot S3

Taxonomic richness:

```{r}
supp_tables %>% 
  filter(Table == "S3") %>% 
  filter(Predictor == "Taxonomic richness") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  facet_wrap(~ Predictor) +
  scale_colour_manual(values = habitat_colours) +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(a) Taxonomic richness") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.01 , y = Habitat, 
                label = Letter), colour = "black", size = 10) +
  theme(legend.position = "none") -> s3.taxon
```

Functional richness:

```{r}
supp_tables %>% 
  filter(Table == "S3") %>% 
  filter(Predictor == "Functional richness") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  scale_colour_manual(values = habitat_colours) +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(b) Functional richness") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.01 , y = Habitat, 
                label = Letter), colour = "black", size = 10) +
  theme(legend.position = "none") -> s3.functional
```

Synchrony:

```{r}
supp_tables %>% 
  filter(Table == "S3") %>% 
  filter(Predictor == "Synchrony") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  scale_colour_manual(values = habitat_colours) +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(c) Synchrony") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.10 , y = Habitat, 
                label = Letter), colour = "black", size = 10) -> s3.synch
```

Combine using `patchwork`:

```{r}
pdf(here("output/tableS3toFig.pdf"), height = 17, width = 11)

s3.taxon / s3.functional / s3.synch

dev.off()
```

# Plot S4

Site-level taxonomic richness

```{r}
supp_tables %>% 
  filter(Table == "S4") %>% 
  filter(Predictor == "Site-level taxonomic-richness") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  scale_colour_manual(values = habitat_colours) +
  # if you want the 0 reference:
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 1, color = "gray") +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(a) Site-level taxonomic-richness") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.05 , y = Habitat, 
                label = Letter), colour = "black", size = 10) +
  theme(legend.position = "none") -> s4.taxon
```

```{r}
supp_tables %>% 
  filter(Table == "S4") %>% 
  filter(Predictor == "Site-level functional-richness") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  scale_colour_manual(values = habitat_colours) +
    # if you want the 0 reference:
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 1, color = "gray") +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(b) Site-level functional-richness") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.1 , y = Habitat, 
                label = Letter), colour = "black", size = 10) +
  theme(legend.position = "none") -> s4.functional
```

Site-level species synchrony

```{r}
supp_tables %>% 
  filter(Table == "S4") %>% 
  filter(Predictor == "Site-level species synchrony") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  scale_colour_manual(values = habitat_colours) +
    # if you want the 0 reference:
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 1, color = "gray") +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(c) Site-level species synchrony") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.4 , y = Habitat, 
                label = Letter), colour = "black", size = 10) +
  theme(legend.position = "none") -> s4.synch
```

Mean plot-level species synchrony

```{r}
supp_tables %>% 
  filter(Table == "S4") %>% 
  filter(Predictor == "Mean plot-level species synchrony") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  scale_colour_manual(values = habitat_colours) +
    # if you want the 0 reference:
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 1, color = "gray") +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(d) Mean plot-level species synchrony") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.6 , y = Habitat, 
                label = Letter), colour = "black", size = 10) -> s4.plt.synch
```

Spatial synchrony

```{r}
supp_tables %>% 
  filter(Table == "S4") %>% 
  filter(Predictor == "Spatial synchrony") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  scale_colour_manual(values = habitat_colours) +
    # if you want the 0 reference:
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 1, color = "gray") +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(f) Spatial synchrony") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.4 , y = Habitat, 
                label = Letter), colour = "black", size = 10) +
  theme(legend.position = "none") -> s4.plt.synch
```

Plot together using `patchwork`

```{r}
pdf(here("output/tableS4toFig.pdf"), height = 15, width = 32)

(s4.taxon + s4.functional + s4.synch) /
  (s4.plt.stab + s4.plt.synch + s4.spat.synch)

dev.off()
```

### 2025-03-13 manuscript_1_code.Rmd
Added scripts to make supplemental figures

@@
 ```
+
+# Supplemental tables
+
+Save `emtrends()` output as a table for supplemental figures
+
+## Table S2
+
+For supplemental table S2: cover vs. diversity
+
+```{r}
+# get estimates: taxonomic richness
+t.s2.a <- as_tibble(emtrends(cover_mod, pairwise ~ habitat, var = "richness")$emtrends) %>% 
+  mutate(Predictor = "Taxonomic richness",
+         `Spatial scale` = "Plot-level") %>% 
+  # rename this so it matches for joining purposes
+  rename_if(str_detect(names(.), ".trend"), ~"Mean")
+
+# just want to make sure this^ isn't this (below)?
+# emtrends(cover_mod_2, pairwise ~ habitat, var = "richness_mean")
+
+t.s2.b <- as_tibble(emtrends(site_cover_mod, pairwise ~ habitat, var = "richness")$emtrends) %>% 
+  mutate(Predictor = "Taxonomic richness",
+         `Spatial scale` = "Site-level") %>% 
+  rename_if(str_detect(names(.), ".trend"), ~"Mean")
+
+# functional richness
+t.s2.c <- as.tibble(emtrends(cover_mod_fg, pairwise ~ habitat, var = "functional_richness")$emtrends) %>% 
+  mutate(Predictor = "Functional richness",
+         `Spatial scale` = "Plot-level") %>% 
+  rename_if(str_detect(names(.), ".trend"), ~"Mean")
+
+# not this right?
+# emtrends(cover_mod_fg_2, pairwise ~ habitat, var = "functional_richness_mean")
+
+t.s2.d <- as.tibble(emtrends(site_cover_mod_fg, pairwise ~ habitat, var = "functional_richness")$emtrends) %>% 
+  mutate(Predictor = "Functional richness",
+         `Spatial scale` = "Site-level") %>% 
+  rename_if(str_detect(names(.), ".trend"), ~"Mean")
+
+# not this right?
+# emtrends(cover_mod_fg, pairwise ~ habitat, var = "functional_richness")
+```
+
+Join:
+
+```{r}
+rbind(t.s2.a, t.s2.b, t.s2.c, t.s2.d) %>% 
+  mutate(Table = "S2") %>% 
+  select(-SE, -df) -> table_s2
+```
+
+## Table S3
+
+Stability
+
+```{r}
+# get estimates for taxonomic richness
+t.s3.a <- as.tibble(emtrends(rich_stab_mod, pairwise ~ habitat, var = "richness_mean")$emtrends) %>% 
+  mutate(Predictor = "Taxonomic richness") %>% 
+  rename_if(str_detect(names(.), ".trend"), ~"Mean")
+
+# get estimates for fucntional richness
+t.s3.b <- as.tibble(emtrends(rich_stab_mod_fg, pairwise ~ habitat, var = "functional_richness_mean")$emtrends) %>% 
+  mutate(Predictor = "Functional richness") %>% 
+  rename_if(str_detect(names(.), ".trend"), ~"Mean")
+
+# get estimates for synchrony
+t.s3.c <- as.tibble(emtrends(synch_stab_mod, pairwise ~ habitat, var = "synchrony")$emtrends) %>% 
+  mutate(Predictor = "Synchrony") %>% 
+  rename_if(str_detect(names(.), ".trend"), ~"Mean")
+
+```
+
+Join:
+
+```{r}
+rbind(t.s3.a, t.s3.b, t.s3.c) %>% 
+  mutate(`Spatial scale` = "Plot-level", 
+         Table = "S3") %>% 
+  select(-SE, -df) -> table_s3
+```
+
+## Table S4
+
+For supplemental table S4: scale figure
+
+```{r}
+# get estimates: site-level richness
+t.s4.a <- as.tibble(emtrends(rich_stab_mod_site, pairwise ~ habitat, var = "richness_mean")$emtrends) %>% 
+  mutate(Predictor = "Site-level taxonomic-richness") %>% 
+  rename_if(str_detect(names(.), ".trend"), ~"Mean")
+
+# not this right?
+# emtrends(rich_stab_mod, pairwise ~ habitat, var = "richness_mean"
+
+t.s4.b <- as.tibble(emtrends(rich_stab_mod_site_fg, pairwise ~ habitat, var = "functional_richness_mean")$emtrends) %>% 
+  mutate(Predictor = "Site-level functional richness") %>% 
+  rename_if(str_detect(names(.), ".trend"), ~"Mean")
+
+## CHECK ^^^^^^ -- doesn't match??
+
+t.s4.c <- as.tibble(emtrends(synch_stab_mod_site, pairwise ~ habitat, var = "synchrony")$emtrends) %>% 
+  mutate(Predictor = "Site-level species synchrony") %>% 
+  rename_if(str_detect(names(.), ".trend"), ~"Mean")
+
+t.s4.d <- as.tibble(emtrends(mean_synchrony_mod, pairwise ~ habitat, var = "synchrony_mean")$emtrends) %>% 
+  mutate(Predictor = "Mean plot-level species synchrony") %>% 
+  rename_if(str_detect(names(.), ".trend"), ~"Mean")
+
+t.s4.e <- as.tibble(emtrends(alpha_gamma_stab_mod, pairwise ~ habitat, var = "stability_mean")$emtrends) %>% 
+  mutate(Predictor = "Plot-level stability") %>% 
+  rename_if(str_detect(names(.), ".trend"), ~"Mean")
+
+t.s4.f <- as.tibble(emtrends(spatial_synchrony_mod, pairwise ~ habitat, var = "spatial_synchrony")$emtrends) %>% 
+  mutate(Predictor = "Spatial synchrony") %>% 
+  rename_if(str_detect(names(.), ".trend"), ~"Mean")
+```
+
+Join:
+
+```{r}
+rbind(t.s4.a, t.s4.b, t.s4.c,
+      t.s4.d, t.s4.e, t.s4.f) %>% 
+  mutate(Table = "S4") %>% 
+  select(-SE, -df) -> table_s4
+```
+
+### Save as a CSV
+
+Join them all together
+
+```{r}
+table_s2 %>% 
+  full_join(table_s3) %>% 
+  full_join(table_s4) %>% 
+  # rename to match previous version/code
+  rename(Habitat = habitat,
+         Lower_CI = asymp.LCL,
+         Upper_CI = asymp.UCL) %>%
+  # make a blank column to fill in the letter designations manually
+  mutate(Letter = NA) %>%
+  # save as CSV
+  write_csv(here("data/supplemental_tables.csv"))
+```

### 2025-03-13 supplemental_table_figures.Rmd
Added scripts to make supplemental figures

@@
 ```{r}
-supp_tables <- read_csv(here("data/supplemental_tables.csv"))
+supp_tables <- read_csv(here("data/supplemental_tables_manual_letters.csv"))
 ```
@@
   ggtitle("(b) Site-level") +
+  # if you want the 0 reference:
+  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 1, color = "gray") +
   model_themes + 
@@
   ylab("") +
-  ggtitle("(a) Site-level taxonomic-richness") +
+  ggtitle("(a) Site-level taxonomic richness") +
   model_themes + 
@@
   filter(Table == "S4") %>% 
-  filter(Predictor == "Site-level functional-richness") %>% 
+  filter(Predictor == "Site-level functional richness") %>% 
   # order the habitats by distance from shore
@@
   ylab("") +
-  ggtitle("(b) Site-level functional-richness") +
+  ggtitle("(b) Site-level functional richness") +
   model_themes + 
@@
 
+Plot-level stability
+
+```{r}
+supp_tables %>% 
+  filter(Table == "S4") %>% 
+  filter(Predictor == "Plot-level stability") %>% 
+  # order the habitats by distance from shore
+  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
+  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
+  geom_point(size = 6) +
+  # add confidence intervals:
+  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
+  scale_colour_manual(values = habitat_colours) +
+  # if you want the 0 reference:
+  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 1, color = "gray") +
+  xlab("Coefficient") +
+  ylab("") +
+  ggtitle("(e) Plot-level stability") +
+  model_themes + 
+  # add letters denoting significance
+  geom_text(aes(x = Upper_CI + 0.1 , y = Habitat, 
+                label = Letter), colour = "black", size = 10) -> s4.plt.stab
+```
+
 Spatial synchrony
@@
                 label = Letter), colour = "black", size = 10) +
-  theme(legend.position = "none") -> s4.plt.synch
+  theme(legend.position = "none") -> s4.spat.synch
 ```

### 2025-03-17 manuscript_1_code.Rmd
implemented plot aesthetic changes

@@
         plot.background = element_rect(fill = "white"),
-        legend.text = element_text(size = rel(1)), # increase legend text
+        legend.text = element_text(size=24),
         legend.title = element_blank(),
-        legend.position = 'bottom', # move legend to bottom
+        legend.position = 'bottom', # move legend to bottom,
+        legend.key.size = unit(2, 'cm'), # increase legend size
         legend.box = 'horizontal', # make legend horizontal
@@
         legend.title = element_blank(),
+        legend.key.size = unit(2, 'cm'), # increase legend size
         legend.position = 'bottom', # move legend to bottom
@@
     scale_colour_manual(values = functional_group_colours) +
-    scale_x_continuous(breaks = unique(figure_1_data$year)) +
+    scale_x_continuous(breaks = seq(2007, 2023, 2)) +
     labs(y = "Percent Cover",
@@
                       legend = "bottom", 
+                      legend.grob = get_legend(figure_3a),
                       font.label = list(size = 26, color = "black", face = "plain")))

### 2025-07-31 beta_div.Rmd
beta diversity code
This verison includes transformations and removal of rare species

---
title: "beta diversity"
author: "L. Enright"
date: "2025-06-10"
output: html_document
---

# read in the data. these are written out from previous scripts, need to rerun all of these scripts and then read back in the data, to make sure they are up to date.

if someone else other than lauren is running these, need to update data folder name
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(vegan)

macro_long_data <- read.csv(here::here("data_le", "macro_long_data.csv"), stringsAsFactors = F) 
diversity_stability_synchrony <- read.csv(here::here("data_le", "full_plot_level_dss.csv"), stringsAsFactors = F) 
diversity_stability_synchrony_site <- read.csv(here::here("data_le", "diversity_stability_synchrony_site.csv"), stringsAsFactors = F) 
spatial_synchrony <- read.csv(here::here("data_le", "spatial_synchrony.csv"), stringsAsFactors = F) 
mean_plot_level_dss <- read.csv(here::here("data_le", "mean_plot_level_dss.csv"), stringsAsFactors = F)

taxo <- read.csv(here::here("data_le", "MCR_Algal_Functional_Groups_02042025_updatedtax.csv"), stringsAsFactors = F)
```

# Data Prep
*** is it okay to leave the data in prop cover? instead of percent cover? leaving in prop cover for now. --> Noam says yes!

```{r}
# first, need to SUM the data across years for each quad 
macro_long_data_summed <- macro_long_data %>%
  group_by(location, site, habitat, site_habitat, transect, quadrat, taxa) %>%
  summarise(sum_taxa = sum(prop_cover)) %>%
  ungroup()
#should have 1 row for each site/habitat/transect/quad/taxa, summarized across years. 
```

# beta diversity - intital attempt
```{r}

#from long to wide
macro_wide_data_summed <- macro_long_data_summed %>%
   pivot_wider(names_from = taxa, values_from = sum_taxa, values_fill = 0)

#calculating beta diversity for individual sites/habitats so that I can check the foreloop is working.
#LTER 1 BR

macro_wide_data_summed_LTER1BR <- macro_wide_data_summed %>%
  filter(site_habitat == "lter_1_backreef") %>%
   filter(rowSums(.[, 7:68]) != 0)
# lost 1 :) --> this would be a quad that has no macroalgae in it 
#(lter_1_backreef_algae_transect_5_quad_7)

#for checking row sums
#macro_wide_data_summed_LTER1BR$row_sum <- rowSums(macro_wide_data_summed_LTER1BR[, 7:68])

sp_community_LTER1BR <- as.matrix(macro_wide_data_summed_LTER1BR[,c(7:68)])
meta_LTER1BR <-macro_wide_data_summed_LTER1BR[,c(1:6)]
sp_community_LTER1BR_transformed <- wisconsin(sp_community_LTER1BR)

rowSums(sp_community_LTER1BR)

#pairwise comparisons 
bray_LTER1BR <- vegdist(sp_community_LTER1BR, method = "bray")
bray_LTER1BR_transformed <- vegdist(sp_community_LTER1BR_transformed, method = "bray")

median_vegdist <- median(bray_LTER1BR)
#0.4913037
median_vegdist_tranformed <- median(bray_LTER1BR_transformed)
# 0.6590098
mean_vegdist <- mean(bray_LTER1BR)
#0.5195
mean_vegdist_tranformed <- mean(bray_LTER1BR_transformed)
# 0.6451
betadisp_LTER1BR <- betadisper(bray_LTER1BR, meta_LTER1BR$site_habitat)
#0.3596  
betadisp_LTER1BR_transformed <- betadisper(bray_LTER1BR_transformed, meta_LTER1BR$site_habitat)
#0.4521

#LTER 2 Outer 10

macro_wide_data_summed_LTER2OUT10 <- macro_wide_data_summed %>%
  filter(site_habitat == "lter_2_outer_10") %>%
   filter(rowSums(.[, 7:68]) != 0)
# lost 0  :)

sp_community_LTER2O10 <- as.matrix(macro_wide_data_summed_LTER2OUT10[,c(7:68)])
meta_LTER2O10 <-macro_wide_data_summed_LTER2OUT10[,c(1:6)]

rowSums(sp_community_LTER2O10)

#pairwise comparisons 
bray_LTER2O10 <- vegdist(sp_community_LTER2O10, method = "bray")
median_vegdist <- median(bray_LTER2O10)
#0.283019
betadisp_LTER2O10 <- betadisper(bray_LTER2O10, meta_LTER2O10$site_habitat)
#0.2041

```
## Good forloop for beta
this is good for calculating beta diversity for all 24 sites 
```{r}

# Initialize a results data frame
results_beta <- data.frame(
  site_habitat = character(),
  median_bray = numeric(),
   mean_bray = numeric(),
  mean_betadisp = numeric(),
  stringsAsFactors = FALSE
)

betadisp_list <- list()

# Get unique site_habitat combinations
sites <- unique(macro_wide_data_summed$site_habitat)

for (sh in sites) {
  # Filter for one site/habitat
  df_sub <- macro_wide_data_summed %>%
    filter(site_habitat == sh) %>%
    filter(rowSums(.[, 7:68]) != 0)

  # Skip if no rows left
  if (nrow(df_sub) < 2) {
    warning(paste("Skipping", sh, "- not enough data"))
    next
  }

  # Community and metadata
  sp_community <- as.matrix(df_sub[, 7:68])
  meta <- df_sub[, 1:6]

  # Bray-Curtis distances
  bray <- vegdist(sp_community, method = "bray")
  median_bray <- median(bray) # i used median not mean and I am not sure why? should it be mean? 
  mean_bray <- mean(bray)

  # Beta dispersion
  betadisp <- betadisper(bray, meta$site_habitat)
  mean_disp <- mean(betadisp$distances)

  betadisp_list[[sh]] <- betadisp

  # Store result
  results_beta <- rbind(results_beta, data.frame(
    site_habitat = sh,
    #betadisp_results = betadisp_results,
    median_bray = median_bray,
    mean_bray = mean_bray,
    mean_betadisp = mean_disp,
    stringsAsFactors = FALSE
  ))
}

# View results
print(results_beta)
betadisp_list

#betadisp list =       0.3596 LTER 1 BR
#betadisp list =       0.2041  LTER 2 FR10
#okay, these match with the single sites I did above, so I think the loop is working well. 
```

## plotting for beta 
```{r merge and plot}

# will eventually want to match Noam's nice themes, but just focusing on colors for now. 
habitat_colours <- c("Fringing" = "#D81B60",
                     "Backreef" = "#FFC107",
                     "Forereef 10m" = "#1E88E5",
                     "Forereef 17m" = "#004D40")

# merge together the beta div with the spatial synchrony output
beta_spatialsync <- merge(spatial_synchrony, results_beta, by = "site_habitat")

# some exploratory plots
beta_spatialsync %>%
  ggplot(aes(x = median_bray, y = spatial_synchrony, color = site)) +
  geom_point() +
  #scale_colour_manual(values = habitat_colours) +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

beta_spatialsync %>%
  ggplot(aes(x = mean_bray, y = spatial_synchrony, color = site)) +
  geom_point() +
  #scale_colour_manual(values = habitat_colours) +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

beta_spatialsync %>%
  ggplot(aes(x = mean_betadisp, y = spatial_synchrony, color = site)) +
  geom_point() +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

#patterns are similar whether we use median bray or mean bray or mean betadisp ... do we need to try with median bray?

#NO FACET
beta_spatialsync %>%
  ggplot(aes(x = median_bray, y = spatial_synchrony, color = habitat, shape = site)) +
  geom_point() +
   geom_smooth(method = "lm", se = TRUE, aes(group = habitat)) +
  scale_colour_manual(values = habitat_colours) +
  theme_minimal()
  #geom_line(method = lm)

# negative relationship between median bray and stability 
beta_spatialsync %>%
  ggplot(aes(x = median_bray, y = stability_mean, color = habitat, shape = site)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, aes(group = habitat)) +
  scale_colour_manual(values = habitat_colours) +
  theme_minimal()

   #geom_line(method = lm)

```

## probably the figure we want to use based on discussion 
shows a negative relationship, don't need to touch on habitat really 
```{r}
# need to make pretty
beta_spatialsync %>%
  ggplot(aes(x = mean_bray, y = spatial_synchrony)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  #scale_colour_manual(values = habitat_colours) +
  theme_minimal()
```
# Trying transformation with no removal of rare species

```{r}

#since we have values greater than 1, cannot use arcsin sqrt, need to use Wisconsin transformation

# Initialize a results data frame
results_beta_transformed <- data.frame(
  site_habitat = character(),
  median_bray = numeric(),
  median_bray_trans = numeric(),
  mean_bray = numeric(),
  mean_bray_trans = numeric(),
  mean_betadisp = numeric(),
   mean_disp_transformed = numeric(),
  stringsAsFactors = FALSE
)

betadisp_list_transformed <- list()

# Get unique site_habitat combinations
sites <- unique(macro_wide_data_summed$site_habitat)

for (sh in sites) {
  # Filter for one site/habitat
  df_sub <- macro_wide_data_summed %>%
    filter(site_habitat == sh) %>%
    filter(rowSums(.[, 7:68]) != 0)  # Remove rows with all zeros

  # Community and metadata
  sp_community <- as.matrix(df_sub[, 7:68])
  meta <- df_sub[, 1:6]

  # Arc-sine square root transformation for the community matrix (required before vegdist)
  sp_community_transformed <- wisconsin(sp_community)

  # Confirm it's working
  print(paste("Range after transform for", sh, ":", paste(range(sp_community_transformed, na.rm = TRUE), collapse = " - ")))

  # Bray-Curtis distances
   bray_raw <- vegdist(sp_community, method = "bray")
  bray_transformed <- vegdist(sp_community_transformed, method = "bray")

  correlation <- cor(as.numeric(bray_raw), as.numeric(bray_transformed))
  print(paste("Correlation between raw and transformed Bray-Curtis for", sh, ":", round(correlation, 4)))

  # Calculate median and mean Bray-Curtis distance
  median_bray <- median(bray_raw)
  median_bray_trans <- median(bray_transformed)
  mean_bray <- mean(bray_raw)
  mean_bray_trans <- mean(bray_transformed)

  # Beta dispersion
  betadisp <- betadisper(bray_raw, meta$site_habitat)
  betadisp_transformed <- betadisper(bray_transformed, meta$site_habitat)
  mean_disp <- mean(betadisp$distances)
  mean_disp_transformed <- mean(betadisp_transformed$distances)

  betadisp_list[[sh]] <- betadisp
  betadisp_list_transformed[[sh]] <- betadisp_transformed

  # Store results
  results_beta_transformed <- bind_rows(results_beta_transformed, data.frame(
    site_habitat = sh,
    median_bray = median_bray,
    median_bray_trans = median_bray_trans,
    mean_bray = mean_bray,
    mean_bray_trans = mean_bray_trans,
    mean_betadisp = mean_disp,
    mean_disp_transformed = mean_disp_transformed,
    stringsAsFactors = FALSE
  ))
}

# View results
print(results_beta_transformed)
betadisp_list_transformed

```

```{r}
beta_spatialsync_transformed <- merge(spatial_synchrony, results_beta_transformed, by = "site_habitat")

# some exploratory plots
# not transformed
beta_spatialsync_transformed %>%
  ggplot(aes(x = mean_bray, y = spatial_synchrony, color = site)) +
  geom_point() +
  #scale_colour_manual(values = habitat_colours) +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

#transformed
beta_spatialsync_transformed %>%
  ggplot(aes(x = mean_bray_trans, y = spatial_synchrony, color = site)) +
  geom_point() +
  #scale_colour_manual(values = habitat_colours) +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

beta_spatialsync_transformed %>%
  ggplot(aes(x = median_bray_trans, y = spatial_synchrony, color = site)) +
  geom_point() +
  #scale_colour_manual(values = habitat_colours) +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

beta_spatialsync_transformed %>%
  ggplot(aes(x = mean_disp_transformed, y = spatial_synchrony, color = site)) +
  geom_point() +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

#patterns are similar whether we use transformed median bray or transformed mean bray or transformed mean betadisp
# slightly different pattersn with the transformation

```

#Again, I think this is the figure we want to show. 
```{r}

beta_spatialsync_transformed %>%
  ggplot(aes(x = mean_bray, y = spatial_synchrony)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  #scale_colour_manual(values = habitat_colours) +
  theme_minimal()

beta_spatialsync_transformed %>%
  ggplot(aes(x = mean_bray_trans, y = spatial_synchrony)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  #scale_colour_manual(values = habitat_colours) +
  theme_minimal()
```
# Trying removing the species.

let's make a matrix for each habitat

## Fringing
```{r}

fringing_macro_wide_data_summed <- macro_wide_data_summed %>%
  filter(habitat == "Fringing")

#300 rows.. 
#make into a matrix
fringing_macro_wide_data_summed_matrix <- as.matrix(fringing_macro_wide_data_summed[, 7:68])

fringing_macro_wide_data_summed_meta <- fringing_macro_wide_data_summed[, 1:6]

#filter rare species
#how many quads out of 300 have a species
fringing_site_species_presence <- colSums(fringing_macro_wide_data_summed_matrix > 0) # Presence absence 
#divide by # of rows totla
fringing_site_species_percentage <- (fringing_site_species_presence/nrow(fringing_macro_wide_data_summed_matrix)) # proportion of rows where a species is present

fringing_site_common_species <- names(fringing_site_species_percentage[fringing_site_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...

fringing_site_species_percentage
fringing_site_common_species

#17 species

fringing_site_common_species_matrix <- fringing_macro_wide_data_summed_matrix[,colnames(fringing_macro_wide_data_summed_matrix) %in% fringing_site_common_species] # filter original species matrix
#okay this has 17

#now need to make it back into a dataframe for the forloop
fringing_site_common_species_df <- cbind(fringing_macro_wide_data_summed_meta, fringing_site_common_species_matrix)
#fringing_site_rows_to_remove <- rowSums(fringing_site_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
```

will need to do the forloops for each habitat - fringing
```{r}

#since we have values greater than 1, cannot use arcsin sqrt, need to use Wisconsin transformation

# Initialize a results data frame
results_beta_transformed_fringing_commonspecies <- data.frame(
  site_habitat = character(),
  median_bray = numeric(),
  median_bray_trans = numeric(),
  mean_bray = numeric(),
  mean_bray_trans = numeric(),
  mean_betadisp = numeric(),
   mean_disp_transformed = numeric(),
  stringsAsFactors = FALSE
)

betadisp_list_fringing_commonspecies <- list()
betadisp_list_transformed_fringing_commonspecies <- list()

# Get unique site_habitat combinations
sites_fring <- unique(fringing_site_common_species_df$site_habitat)

for (sh in sites_fring) {
  # Filter for one site/habitat
  df_sub <- fringing_site_common_species_df %>%
    filter(site_habitat == sh) %>%
    filter(rowSums(.[, 7:23]) != 0)  # Remove rows with all zeros

  # Community and metadata
  sp_community <- as.matrix(df_sub[, 7:23])
  meta <- df_sub[, 1:6]

  # Arc-sine square root transformation for the community matrix (required before vegdist)
  sp_community_transformed <- wisconsin(sp_community)

  # Confirm it's working
  print(paste("Range after transform for", sh, ":", paste(range(sp_community_transformed, na.rm = TRUE), collapse = " - ")))

  # Bray-Curtis distances
   bray_raw <- vegdist(sp_community, method = "bray")
  bray_transformed <- vegdist(sp_community_transformed, method = "bray")

  correlation <- cor(as.numeric(bray_raw), as.numeric(bray_transformed))
  print(paste("Correlation between raw and transformed Bray-Curtis for", sh, ":", round(correlation, 4)))

  # Calculate median and mean Bray-Curtis distance
  median_bray <- median(bray_raw)
  median_bray_trans <- median(bray_transformed)
  mean_bray <- mean(bray_raw)
  mean_bray_trans <- mean(bray_transformed)

  # Beta dispersion
  betadisp <- betadisper(bray_raw, meta$site_habitat)
  betadisp_transformed <- betadisper(bray_transformed, meta$site_habitat)
  mean_disp <- mean(betadisp$distances)
  mean_disp_transformed <- mean(betadisp_transformed$distances)

  betadisp_list_fringing_commonspecies[[sh]] <- betadisp
  betadisp_list_transformed_fringing_commonspecies[[sh]] <- betadisp_transformed

  # Store results
  results_beta_transformed_fringing_commonspecies <- bind_rows(results_beta_transformed_fringing_commonspecies, data.frame(
    site_habitat = sh,
    median_bray = median_bray,
    median_bray_trans = median_bray_trans,
    mean_bray = mean_bray,
    mean_bray_trans = mean_bray_trans,
    mean_betadisp = mean_disp,
    mean_disp_transformed = mean_disp_transformed,
    stringsAsFactors = FALSE
  ))
}

# View results
print(results_beta_transformed_fringing_commonspecies)
betadisp_list_transformed_fringing_commonspecies

```
again, trying a one off to make sure foreloop is working 
```{r}
commonspecies_macro_LTER1FRI <- fringing_site_common_species_df %>%
  filter(site_habitat == "lter_1_fringing") %>%
   filter(rowSums(.[, 7:23]) != 0)

sp_community_LTER1FRI <- as.matrix(commonspecies_macro_LTER1FRI[,c(7:23)])
meta_LTER1FR <-commonspecies_macro_LTER1FRI[,c(1:6)]
sp_community_LTER1FRI_transformed <- wisconsin(sp_community_LTER1FRI)

rowSums(sp_community_LTER1BR)

#pairwise comparisons 
bray_LTER1FRI <- vegdist(sp_community_LTER1FRI, method = "bray")
bray_LTER1FRI_transformed <- vegdist(sp_community_LTER1FRI_transformed, method = "bray")

mean_vegdist <- mean(bray_LTER1FRI)
#0.3833
mean_vegdist_tranformed <- mean(bray_LTER1FRI_transformed)
#0.5812

#yay, they match with forloop output
```

## Backreef
```{r}

backreef_macro_wide_data_summed <- macro_wide_data_summed %>%
  filter(habitat == "Backreef")

#300 rows.. 
#make into a matrix
backreef_macro_wide_data_summed_matrix <- as.matrix(backreef_macro_wide_data_summed[, 7:68])

backreef_macro_wide_data_summed_meta <- backreef_macro_wide_data_summed[, 1:6]

#filter rare species
#how many quads out of 300 have a species
backreef_site_species_presence <- colSums(backreef_macro_wide_data_summed_matrix > 0) # Presence absence 
#divide by # of rows totla
backreef_site_species_percentage <- (backreef_site_species_presence/nrow(backreef_macro_wide_data_summed_matrix)) # proportion of rows where a species is present

backreef_site_common_species <- names(backreef_site_species_percentage[backreef_site_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...

backreef_site_species_percentage
backreef_site_common_species

#18 species

backreef_site_common_species_matrix <- backreef_macro_wide_data_summed_matrix[,colnames(backreef_macro_wide_data_summed_matrix) %in% backreef_site_common_species] # filter original species matrix
#okay this has 18

#now need to make it back into a dataframe for the forloop
backreef_site_common_species_df <- cbind(backreef_macro_wide_data_summed_meta, backreef_site_common_species_matrix)

```

will need to do the forloops for each habitat - backreef
```{r}

#since we have values greater than 1, cannot use arcsin sqrt, need to use Wisconsin transformation

# Initialize a results data frame
results_beta_transformed_backreef_commonspecies <- data.frame(
  site_habitat = character(),
  median_bray = numeric(),
  median_bray_trans = numeric(),
  mean_bray = numeric(),
  mean_bray_trans = numeric(),
  mean_betadisp = numeric(),
   mean_disp_transformed = numeric(),
  stringsAsFactors = FALSE
)

betadisp_list_backreef_commonspecies <- list()
betadisp_list_transformed_backreef_commonspecies <- list()

# Get unique site_habitat combinations
sites_back <- unique(backreef_site_common_species_df$site_habitat)

for (sh in sites_back) {
  # Filter for one site/habitat
  df_sub <- backreef_site_common_species_df %>%
    filter(site_habitat == sh) %>%
    filter(rowSums(.[, 7:24]) != 0)  # Remove rows with all zeros

  # Community and metadata
  sp_community <- as.matrix(df_sub[, 7:24])
  meta <- df_sub[, 1:6]

  # Arc-sine square root transformation for the community matrix (required before vegdist)
  sp_community_transformed <- wisconsin(sp_community)

  # Confirm it's working
  print(paste("Range after transform for", sh, ":", paste(range(sp_community_transformed, na.rm = TRUE), collapse = " - ")))

  # Bray-Curtis distances
   bray_raw <- vegdist(sp_community, method = "bray")
  bray_transformed <- vegdist(sp_community_transformed, method = "bray")

  correlation <- cor(as.numeric(bray_raw), as.numeric(bray_transformed))
  print(paste("Correlation between raw and transformed Bray-Curtis for", sh, ":", round(correlation, 4)))

  # Calculate median and mean Bray-Curtis distance
  median_bray <- median(bray_raw)
  median_bray_trans <- median(bray_transformed)
  mean_bray <- mean(bray_raw)
  mean_bray_trans <- mean(bray_transformed)

  # Beta dispersion
  betadisp <- betadisper(bray_raw, meta$site_habitat)
  betadisp_transformed <- betadisper(bray_transformed, meta$site_habitat)
  mean_disp <- mean(betadisp$distances)
  mean_disp_transformed <- mean(betadisp_transformed$distances)

  betadisp_list_backreef_commonspecies[[sh]] <- betadisp
  betadisp_list_transformed_backreef_commonspecies[[sh]] <- betadisp_transformed

  # Store results
  results_beta_transformed_backreef_commonspecies <- bind_rows(results_beta_transformed_backreef_commonspecies, data.frame(
    site_habitat = sh,
    median_bray = median_bray,
    median_bray_trans = median_bray_trans,
    mean_bray = mean_bray,
    mean_bray_trans = mean_bray_trans,
    mean_betadisp = mean_disp,
    mean_disp_transformed = mean_disp_transformed,
    stringsAsFactors = FALSE
  ))
}

# View results
print(results_beta_transformed_backreef_commonspecies)
betadisp_list_transformed_backreef_commonspecies

```

## Forereef 10
```{r}

forereef10_macro_wide_data_summed <- macro_wide_data_summed %>%
  filter(habitat == "Forereef 10m")

#300 rows.. 
#make into a matrix
forereef10_macro_wide_data_summed_matrix <- as.matrix(forereef10_macro_wide_data_summed[, 7:68])

forereef10_macro_wide_data_summed_meta <- forereef10_macro_wide_data_summed[, 1:6]

#filter rare species
#how many quads out of 300 have a species
forereef10_site_species_presence <- colSums(forereef10_macro_wide_data_summed_matrix > 0) # Presence absence 
#divide by # of rows totla
forereef10_site_species_percentage <- (forereef10_site_species_presence/nrow(forereef10_macro_wide_data_summed_matrix)) # proportion of rows where a species is present

forereef10_site_common_species <- names(forereef10_site_species_percentage[forereef10_site_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...

forereef10_site_species_percentage
forereef10_site_common_species

#15 species

forereef10_site_common_species_matrix <- forereef10_macro_wide_data_summed_matrix[,colnames(forereef10_macro_wide_data_summed_matrix) %in% forereef10_site_common_species] # filter original species matrix
#okay this has 15

#now need to make it back into a dataframe for the forloop
forereef10_site_common_species_df <- cbind(forereef10_macro_wide_data_summed_meta, forereef10_site_common_species_matrix)

```

will need to do the forloops for each habitat - forereef10
```{r}

#since we have values greater than 1, cannot use arcsin sqrt, need to use Wisconsin transformation

# Initialize a results data frame
results_beta_transformed_forereef10_commonspecies <- data.frame(
  site_habitat = character(),
  median_bray = numeric(),
  median_bray_trans = numeric(),
  mean_bray = numeric(),
  mean_bray_trans = numeric(),
  mean_betadisp = numeric(),
   mean_disp_transformed = numeric(),
  stringsAsFactors = FALSE
)

betadisp_list_forereef10_commonspecies <- list()
betadisp_list_transformed_forereef10_commonspecies <- list()

# Get unique site_habitat combinations
sites_FO10 <- unique(forereef10_site_common_species_df$site_habitat)

for (sh in sites_FO10) {
  # Filter for one site/habitat
  df_sub <- forereef10_site_common_species_df %>%
    filter(site_habitat == sh) %>%
    filter(rowSums(.[, 7:21]) != 0)  # Remove rows with all zeros

  # Community and metadata
  sp_community <- as.matrix(df_sub[, 7:21])
  meta <- df_sub[, 1:6]

  # Arc-sine square root transformation for the community matrix (required before vegdist)
  sp_community_transformed <- wisconsin(sp_community)

  # Confirm it's working
  print(paste("Range after transform for", sh, ":", paste(range(sp_community_transformed, na.rm = TRUE), collapse = " - ")))

  # Bray-Curtis distances
   bray_raw <- vegdist(sp_community, method = "bray")
  bray_transformed <- vegdist(sp_community_transformed, method = "bray")

  correlation <- cor(as.numeric(bray_raw), as.numeric(bray_transformed))
  print(paste("Correlation between raw and transformed Bray-Curtis for", sh, ":", round(correlation, 4)))

  # Calculate median and mean Bray-Curtis distance
  median_bray <- median(bray_raw)
  median_bray_trans <- median(bray_transformed)
  mean_bray <- mean(bray_raw)
  mean_bray_trans <- mean(bray_transformed)

  # Beta dispersion
  betadisp <- betadisper(bray_raw, meta$site_habitat)
  betadisp_transformed <- betadisper(bray_transformed, meta$site_habitat)
  mean_disp <- mean(betadisp$distances)
  mean_disp_transformed <- mean(betadisp_transformed$distances)

  betadisp_list_forereef10_commonspecies[[sh]] <- betadisp
  betadisp_list_transformed_forereef10_commonspecies[[sh]] <- betadisp_transformed

  # Store results
  results_beta_transformed_forereef10_commonspecies <- bind_rows(results_beta_transformed_forereef10_commonspecies, data.frame(
    site_habitat = sh,
    median_bray = median_bray,
    median_bray_trans = median_bray_trans,
    mean_bray = mean_bray,
    mean_bray_trans = mean_bray_trans,
    mean_betadisp = mean_disp,
    mean_disp_transformed = mean_disp_transformed,
    stringsAsFactors = FALSE
  ))
}

# View results
print(results_beta_transformed_forereef10_commonspecies)
betadisp_list_transformed_forereef10_commonspecies

```
## Forereef 17
```{r}

forereef17_macro_wide_data_summed <- macro_wide_data_summed %>%
  filter(habitat == "Forereef 17m")

#300 rows.. 
#make into a matrix
forereef17_macro_wide_data_summed_matrix <- as.matrix(forereef17_macro_wide_data_summed[, 7:68])

forereef17_macro_wide_data_summed_meta <- forereef17_macro_wide_data_summed[, 1:6]

#filter rare species
#how many quads out of 300 have a species
forereef17_site_species_presence <- colSums(forereef17_macro_wide_data_summed_matrix > 0) # Presence absence 
#divide by # of rows totla
forereef17_site_species_percentage <- (forereef17_site_species_presence/nrow(forereef17_macro_wide_data_summed_matrix)) # proportion of rows where a species is present

forereef17_site_common_species <- names(forereef17_site_species_percentage[forereef17_site_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...

forereef17_site_species_percentage
forereef17_site_common_species

#13 species

forereef17_site_common_species_matrix <- forereef17_macro_wide_data_summed_matrix[,colnames(forereef17_macro_wide_data_summed_matrix) %in% forereef17_site_common_species] # filter original species matrix
#okay this has 13

#now need to make it back into a dataframe for the forloop
forereef17_site_common_species_df <- cbind(forereef17_macro_wide_data_summed_meta, forereef17_site_common_species_matrix)

```

will need to do the forloops for each habitat - forereef17
```{r}

#since we have values greater than 1, cannot use arcsin sqrt, need to use Wisconsin transformation

# Initialize a results data frame
results_beta_transformed_forereef17_commonspecies <- data.frame(
  site_habitat = character(),
  median_bray = numeric(),
  median_bray_trans = numeric(),
  mean_bray = numeric(),
  mean_bray_trans = numeric(),
  mean_betadisp = numeric(),
   mean_disp_transformed = numeric(),
  stringsAsFactors = FALSE
)

betadisp_list_forereef17_commonspecies <- list()
betadisp_list_transformed_forereef17_commonspecies <- list()

# Get unique site_habitat combinations
sites_FO17 <- unique(forereef17_site_common_species_df$site_habitat)

for (sh in sites_FO17) {
  # Filter for one site/habitat
  df_sub <- forereef17_site_common_species_df %>%
    filter(site_habitat == sh) %>%
    filter(rowSums(.[, 7:19]) != 0)  # Remove rows with all zeros

  # Community and metadata
  sp_community <- as.matrix(df_sub[, 7:19])
  meta <- df_sub[, 1:6]

  # Arc-sine square root transformation for the community matrix (required before vegdist)
  sp_community_transformed <- wisconsin(sp_community)

  # Confirm it's working
  print(paste("Range after transform for", sh, ":", paste(range(sp_community_transformed, na.rm = TRUE), collapse = " - ")))

  # Bray-Curtis distances
   bray_raw <- vegdist(sp_community, method = "bray")
  bray_transformed <- vegdist(sp_community_transformed, method = "bray")

  correlation <- cor(as.numeric(bray_raw), as.numeric(bray_transformed))
  print(paste("Correlation between raw and transformed Bray-Curtis for", sh, ":", round(correlation, 4)))

  # Calculate median and mean Bray-Curtis distance
  median_bray <- median(bray_raw)
  median_bray_trans <- median(bray_transformed)
  mean_bray <- mean(bray_raw)
  mean_bray_trans <- mean(bray_transformed)

  # Beta dispersion
  betadisp <- betadisper(bray_raw, meta$site_habitat)
  betadisp_transformed <- betadisper(bray_transformed, meta$site_habitat)
  mean_disp <- mean(betadisp$distances)
  mean_disp_transformed <- mean(betadisp_transformed$distances)

  betadisp_list_forereef17_commonspecies[[sh]] <- betadisp
  betadisp_list_transformed_forereef17_commonspecies[[sh]] <- betadisp_transformed

  # Store results
  results_beta_transformed_forereef17_commonspecies <- bind_rows(results_beta_transformed_forereef17_commonspecies, data.frame(
    site_habitat = sh,
    median_bray = median_bray,
    median_bray_trans = median_bray_trans,
    mean_bray = mean_bray,
    mean_bray_trans = mean_bray_trans,
    mean_betadisp = mean_disp,
    mean_disp_transformed = mean_disp_transformed,
    stringsAsFactors = FALSE
  ))
}

# View results
print(results_beta_transformed_forereef17_commonspecies)
betadisp_list_transformed_forereef17_commonspecies

```
# add all habitats together
```{r}
commonspecies_allhabs <- 
  bind_rows(results_beta_transformed_fringing_commonspecies, 
            results_beta_transformed_backreef_commonspecies,
            results_beta_transformed_forereef10_commonspecies,
            results_beta_transformed_forereef17_commonspecies)

# yay, 24 rows!

# now combine with spatial synchrony data

beta_spatialsync_commonspecies <- merge(spatial_synchrony, commonspecies_allhabs, by = "site_habitat")
```
Plotting time
```{r }
# some exploratory plots
# not transformed, all species
beta_spatialsync_transformed %>%
  ggplot(aes(x = mean_bray, y = spatial_synchrony, color = site)) +
  geom_point() +
  #scale_colour_manual(values = habitat_colours) +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

#transformed, all species
beta_spatialsync_transformed %>%
  ggplot(aes(x = mean_bray_trans, y = spatial_synchrony, color = site)) +
  geom_point() +
  #scale_colour_manual(values = habitat_colours) +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

## not transformed, common species
beta_spatialsync_commonspecies %>%
  ggplot(aes(x = mean_bray, y = spatial_synchrony, color = site)) +
  geom_point() +
  #scale_colour_manual(values = habitat_colours) +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

# transformed, common species
beta_spatialsync_commonspecies %>%
  ggplot(aes(x = mean_bray_trans, y = spatial_synchrony, color = site)) +
  geom_point() +
  #scale_colour_manual(values = habitat_colours) +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

#comparison between removing rare species and not removing them look very similar. looking below to make sure they not exactly the same, which would indicate I did something wrong

beta_spatialsync_transformed %>%
  select(site_habitat, site, mean_bray, mean_bray_trans) %>%
  filter(site == "lter_6")

beta_spatialsync_commonspecies %>%
   select(site_habitat, site, mean_bray, mean_bray_trans) %>%
  filter(site == "lter_6")

#nope, just look different but similar! good sign?

```

#Again, I think this is the figure we want to show. 
```{r}

beta_spatialsync_transformed %>%
  ggplot(aes(x = mean_bray, y = spatial_synchrony)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  #scale_colour_manual(values = habitat_colours) +
  theme_minimal()

beta_spatialsync_transformed %>%
  ggplot(aes(x = mean_bray_trans, y = spatial_synchrony)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  #scale_colour_manual(values = habitat_colours) +
  theme_minimal()

beta_spatialsync_commonspecies %>%
  ggplot(aes(x = mean_bray, y = spatial_synchrony)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  #scale_colour_manual(values = habitat_colours) +
  theme_minimal()

beta_spatialsync_commonspecies %>%
  ggplot(aes(x = mean_bray_trans, y = spatial_synchrony)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  #scale_colour_manual(values = habitat_colours) +
  theme_minimal()
```
# looking for an R2, just for kicks
```{r}
model_mean_bray_allspecies <- lm(spatial_synchrony ~ mean_bray, data = beta_spatialsync_transformed)
summary(model_mean_bray_allspecies)$r.squared
#.358

model_mean_bray_trans_allspecies <- lm(spatial_synchrony ~ mean_bray_trans, data = beta_spatialsync_transformed)
summary(model_mean_bray_trans_allspecies)$r.squared
#.28

model_mean_bray_common <- lm(spatial_synchrony ~ mean_bray, data = beta_spatialsync_commonspecies)
summary(model_mean_bray_common)$r.squared
#.354

model_mean_bray_trans_common <- lm(spatial_synchrony ~ mean_bray_trans, data = beta_spatialsync_commonspecies)
summary(model_mean_bray_trans_common)$r.squared
#.270
```

### 2025-09-03 supplemental_abundance_figures.Rmd
supplemental abundance figures for Figure S1

---
title: "supplemental abundance figures"
author: "L. Enright"
date: "2025-06-10"
output: html_document
---

# here we read in the data produced from other scripts. should rerun other scripts and make sure these are the most up to date csvs

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(codyn)

macro_long_data <- read.csv(here::here("data_le", "macro_long_data.csv"), stringsAsFactors = F) 
diversity_stability_synchrony <- read.csv(here::here("data_le", "full_plot_level_dss.csv"), stringsAsFactors = F) 
diversity_stability_synchrony_site <- read.csv(here::here("data_le", "diversity_stability_synchrony_site.csv"), stringsAsFactors = F) 
spatial_synchrony <- read.csv(here::here("data_le", "spatial_synchrony.csv"), stringsAsFactors = F) 
mean_plot_level_dss <- read.csv(here::here("data_le", "mean_plot_level_dss.csv"), stringsAsFactors = F)

taxo <- read.csv(here::here("data_le", "MCR_Algal_Functional_Groups_02042025_updatedtax.csv"), stringsAsFactors = F)
```

```{r set colors and labels}

colors_abundance <- c(
    "sargassum_pacificum" = "#543005",
    #"turbinaria_ornata" =  "#7B3F00",
    "turbinaria_ornata" = "#8C510A",
   # "lobophora_variegata" = "#8C510A",
     "lobophora_variegata" = "#BF812D",
    #"dictyota_bartayresiana" =  "#BF812D",
    "dictyota_bartayresiana" = "#DFC27D",
    "peyssonnelia_inamoena" = "#E6B88A",
    "peyssonnelia_sp" = "#F6E8C3",
    "halimeda_sp" = "#C7EAE5",
    "halimeda_minima" = "#80CDC1" ,
     "halimeda_opuntia" = "#35978F",
    "amansia_rhodantha" = "#01665E", 
    "asparagopsis_taxiformis" = "#003C30",
    "Other" = "grey70")

#other colors I considered
"#5AA6D1"
"#7BAFD4"
"#2B83BA"
"#F1D6A7"
"#D2C48D"
"#E6B88A"
"#E9A972"
"#704214"
"#8B4513"
"#6C3A08"
"#75420B"
"#64330A"

# Use a named LIST of expressions for italics
labs_fill <- list(
  "Other" = "Other",
  "sargassum_pacificum" = expression(italic("Sargassum pacificum")),
  "turbinaria_ornata"   = expression(italic("Turbinaria ornata")),
  "lobophora_variegata" = expression(italic("Lobophora variegata")),
  "dictyota_bartayresiana" = expression(italic("Dictyota bartayresiana")),
  "peyssonnelia_inamoena"  = expression(italic("Peyssonnelia inamoena")),
  "peyssonnelia_sp" = expression(italic("Peyssonnelia") ~ "sp."),
 # "peyssonnelia_sp"        = expression(italic("Peyssonnelia sp.")),
    "halimeda_sp" = expression(italic("Halimeda") ~ "sp."),
  "halimeda_minima"        = expression(italic("Halimeda minima")),
  "halimeda_opuntia"       = expression(italic("Halimeda opuntia")),
  "amansia_rhodantha"      = expression(italic("Amansia rhodantha")),
  "asparagopsis_taxiformis"= expression(italic("Asparagopsis taxiformis"))
)

```

# Clock Circle Absolute Abundance Figure 
```{r}

#pull out top 10 abundance, otherwise is messy with all ofthem 
aggdat_hab <- aggregate(prop_cover ~ taxa * year * habitat,  data = subset(macro_long_data,
      taxa == "sargassum_pacificum" |
      taxa == "turbinaria_ornata" | 
      taxa == "dictyota_bartayresiana"| 
      taxa == "peyssonnelia_inamoena"| 
      taxa == "peyssonnelia_sp"| 
      taxa == "halimeda_minima"| 
      taxa == "halimeda_sp"| 
      taxa == "lobophora_variegata"| 
      taxa == "halimeda_opuntia"| 
      taxa == "amansia_rhodantha"| 
      taxa == "asparagopsis_taxiformis"), FUN = mean)

# order the habitats so figure is correct
aggdat_hab$habitat <- factor(aggdat_hab$habitat, levels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

aggdat_hab$habitat

aggdat_hab_supp <- ggplot(aggdat_hab, aes(year, prop_cover, color = taxa)) + 
  # plot species lines
  geom_line(size = 3) + # 3 or 4 works well. 
  # faceted by species
  facet_wrap(~habitat) +
  # on polor coordinates
  coord_polar()  +
  scale_color_manual(values = colors_abundance, labels = labs_fill, drop = FALSE) +
  theme_minimal() +
  labs(x = "", y = "proportional cover", color = "Taxa") +
   theme(
    strip.text   = element_text(size = 25, face = "bold"),   # facet labels
    axis.title   = element_text(size = 25),                  # axis titles
    axis.text    = element_text(size = 20),                  # axis tick labels
    legend.title = element_text(size = 25, face = "bold"),   # legend title
    legend.text  = element_text(size = 20)                   # legend items
  )

#ggsave(filename = "output/supplement/abundanceclock_supplement_09032025v3.jpg", aggdat_hab_supp, height = 15, width = 20)

```
#Relative Abundance Bar Graph

```{r}
macro_long_data

TAXA_sum_TAXA_YR <- macro_long_data %>%
  group_by(habitat, taxa, year) %>%
   summarise(sum_TAXA = sum(prop_cover))

MA_sum_TAXA_YR <- macro_long_data %>%
  group_by(habitat, year) %>%
   summarise(sum_ALLMA = sum(prop_cover))

relative_abundance_TAXA_YR <- merge(TAXA_sum_TAXA_YR, MA_sum_TAXA_YR, by = c("habitat", "year"))

relative_abundance_cover_TAXA_YR <- relative_abundance_TAXA_YR %>%
  #lter 5 fringing does not have any MA in 2018 or 2023
  mutate(relative_abundance = ifelse(sum_ALLMA > 0, sum_TAXA/sum_ALLMA, 0))

highlight_species <- c("sargassum_pacificum", "turbinaria_ornata", "lobophora_variegata", 
                       "dictyota_bartayresiana", "peyssonnelia_inamoena", "peyssonnelia_sp",
                       "halimeda_minima", "halimeda_opuntia", "halimeda_sp",
                       "amansia_rhodantha", "asparagopsis_taxiformis" 
                        )

data_RA <- relative_abundance_cover_TAXA_YR %>%
  mutate(color_group = ifelse(taxa %in% highlight_species, taxa, "Other")) %>%
  group_by(year, habitat, color_group) %>%  # Group by year, habitat, and color category
  summarize(relative_abundance = sum(relative_abundance), .groups = "drop") %>%  # Sum abundance for "Other"
  mutate(color_group = factor(color_group, levels = c("Other", highlight_species)))

# order the habitats so figure is correct
data_RA$habitat <- factor(data_RA$habitat, levels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

#figure option 1

relative_abundance_supp <- data_RA %>%
ggplot(aes(x = year, y = relative_abundance, fill = color_group)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = colors_abundance, labels = labs_fill, drop = FALSE) +
  labs(
    x = "Year",
    y = "Relative Abundance",
    fill = "Taxa"
  ) +
  scale_y_continuous(labels = scales::percent) +  
  theme_classic() +
  facet_wrap(~habitat) +
  theme(
    strip.text   = element_text(size = 25, face = "bold"),   # facet labels
    axis.title   = element_text(size = 25),                  # axis titles
    axis.text    = element_text(size = 20),                  # axis tick labels
    legend.title = element_text(size = 25, face = "bold"),   # legend title
    legend.text  = element_text(size = 18)                   # legend items
  )

#ggsave(filename = "output/supplement/relative_abundance_supplement_09032025.jpg", relative_abundance_supp , height = 15, width = 20)
```

#Absolute Abundance Bar Graph

```{r}

#11 highlight species

#average cover in each habitat/year

TAXA_mean_TAXA_YR <- macro_long_data %>%
  group_by(habitat, taxa, year) %>%
   summarise(mean_TAXA = mean(prop_cover))

#add together averages for "other species", avoids the little gray lines
TAXA_mean_TAXA_YR_plot <- TAXA_mean_TAXA_YR %>%
  mutate(color_group = if_else(taxa %in% highlight_species, taxa, "Other")) %>%
  group_by(year, habitat, color_group) %>%
  summarise(mean_TAXA_plot = sum(mean_TAXA, na.rm = TRUE), .groups = "drop") %>%
  mutate(color_group = factor(color_group, levels = c("Other", highlight_species)))

# order the habitats so figure is correct
TAXA_mean_TAXA_YR_plot$habitat <- factor(TAXA_mean_TAXA_YR_plot$habitat, levels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

absolute_abundance_supp <- TAXA_mean_TAXA_YR_plot %>%
ggplot(aes(x = year, y = mean_TAXA_plot, fill = color_group)) +
  geom_bar(stat = "identity") +
    scale_fill_manual(values = colors_abundance, labels = labs_fill, drop = FALSE) +
  labs(
    x = "Site",
    y = "Average Percent Cover",
    fill = "Taxa"
  ) +
  scale_y_continuous(labels = scales::percent) +  
  theme_classic() +
  facet_wrap(~habitat) +
  theme(
    strip.text   = element_text(size = 25, face = "bold"),   # facet labels
    axis.title   = element_text(size = 25),                  # axis titles
    axis.text    = element_text(size = 20),                  # axis tick labels
    legend.title = element_text(size = 25, face = "bold"),   # legend title
    legend.text  = element_text(size = 18)                   # legend items
  )

#ggsave(filename = "output/supplement/absolute_abundance_supplement_09032025.jpg", relative_abundance_supp , height = 15, width = 20)

```

### 2025-09-03 figure4.Rmd
figure 4 models and plots

---
title: "figure 4 - models and plots"
author: "L. Enright"
date: "2025-06-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(glmmTMB)
library(car)
library(emmeans)
library(ggpubr)

#don't think we need these df in this script
macro_long_data <- read.csv(here::here("data_le", "macro_long_data.csv"), stringsAsFactors = F) 
diversity_stability_synchrony <- read.csv(here::here("data_le", "full_plot_level_dss.csv"), stringsAsFactors = F) 
diversity_stability_synchrony_site <- read.csv(here::here("data_le", "diversity_stability_synchrony_site.csv"), stringsAsFactors = F) 
mean_plot_level_dss <- read.csv(here::here("data_le", "mean_plot_level_dss.csv"), stringsAsFactors = F) 
taxo <- read.csv(here::here("data_le", "MCR_Algal_Functional_Groups_02042025_updatedtax.csv"), stringsAsFactors = F)

# DO NEED THIS ONE
spatial_synchrony <- read.csv(here::here("data_le", "spatial_synchrony.csv"), stringsAsFactors = F)

```

```{r}
source("R/functions.R")
```

```{r plotting standards set here}

habitat_colours <- c("Fringing" = "#D81B60",
                     "Backreef" = "#FFC107",
                     "Forereef 10m" = "#1E88E5",
                     "Forereef 17m" = "#004D40")

trends_theme <- theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.text.x = element_text(size = 22, angle = 45, hjust = 1),
        axis.title.y = element_text(size=24), 
        axis.text.y = element_text(size = 22),
        plot.title = element_text(size = 24),
        panel.background = element_rect(fill = "white"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size=24),
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom,
        legend.key.size = unit(2, 'cm'), # increase legend size
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))

model_themes <-   theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.title.y = element_text(size=24), 
        axis.text.x = element_text(size=22), 
        axis.text.y = element_text(size=22),
        legend.text = element_text(size=24),
        plot.title = element_text(size=26), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.title = element_blank(),
        legend.key.size = unit(2, 'cm'), # increase legend size
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend

```

```{r}
dss_spatial_ranges <- extract_ranges(spatial_synchrony, "habitat", c("spatial_synchrony", "synchrony_mean", "stability_mean"))

dss_spatial_ranges$habitat <- factor(dss_spatial_ranges$habitat, levels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

#running this factor is important for contrasts estimates outputs 
spatial_synchrony$habitat <- factor(spatial_synchrony$habitat, levels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

```
### Figure 4A - plot level stability X site level stability

```{r alpha gamma stability mod plot}
figure4Amodel_alpha_gamma <- glmmTMB(cover_stability ~ stability_mean*habitat, family = Gamma("log"), data = spatial_synchrony)
summary(figure4Amodel_alpha_gamma)
Anova(figure4Amodel_alpha_gamma)
#slightly different than what Noam got? 
#NOAM
#                          Chisq Df Pr(>Chisq)    
# stability_mean         61.2742  1  4.965e-15 ***
# habitat                12.9328  3   0.004784 ** 
# stability_mean:habitat  9.4834  3   0.023508 *

#ME:
#                         Chisq Df Pr(>Chisq)    
#stability_mean         61.2725  1   4.97e-15 ***
#habitat                12.9323  3   0.004785 ** 
#stability_mean:habitat  9.4832  3   0.023511 *

performance::r2(figure4Amodel_alpha_gamma) # 0.838 #nice, matches w/ Noam

emtrends(figure4Amodel_alpha_gamma, pairwise ~ habitat, var = "stability_mean") # backreef different from fringe and forereef 10
# $contrasts
#  contrast                    estimate    SE  df z.ratio p.value
#  Fringing - Backreef            1.078 0.381 Inf   2.828  0.0242
#  Fringing - Forereef 10m        0.175 0.345 Inf   0.509  0.9571
#  Fringing - Forereef 17m        0.380 0.305 Inf   1.248  0.5965
#  Backreef - Forereef 10m       -0.903 0.349 Inf  -2.589  0.0474
#  Backreef - Forereef 17m       -0.698 0.309 Inf  -2.256  0.1085
#  Forereef 10m - Forereef 17m    0.205 0.263 Inf   0.779  0.8639
#nice, this matches w/ Noam
```

```{r plot site level stability X site level stability - Figure 4A}
alpha_gamma_figure4A_emm <- emmip(figure4Amodel_alpha_gamma,
                        habitat ~ stability_mean,
                        at = list(stability_mean = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_ag_effects <- filter_ranges(alpha_gamma_figure4A_emm, dss_spatial_ranges, "habitat", "stability_mean")

(alpha_gamma_stability_plot_figure4A <- 
  ggplot() +
  geom_point(data = spatial_synchrony,
             aes(x = stability_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_ag_effects, aes(x = stability_mean, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_ag_effects, aes(x = stability_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Site-Level Stability", x = "Mean Plot-Level Stability", title = "") +
  scale_y_continuous(limits = c(0.5, 4.2), breaks = c(1,2,3,4)) +
  model_themes
         )

#what is wrong with error ribbon for forereef 17 m?
```

### Figure 4B - spatial sychrony X site level stability 
```{r model site level stability X spatial synchrony - Figure 4B}
spatial_synchrony_mod_Fig4B <- glmmTMB(cover_stability ~ spatial_synchrony*habitat, family = Gamma("log"), data = spatial_synchrony)
hist(residuals(spatial_synchrony_mod_Fig4B)) # a little skew
plot(residuals(spatial_synchrony_mod_Fig4B) ~ predict(spatial_synchrony_mod_Fig4B))
summary(spatial_synchrony_mod_Fig4B)
car::Anova(spatial_synchrony_mod_Fig4B)
#I got slightly different than Noam's version on Github
#ME:
#                            Chisq Df Pr(>Chisq)    
#spatial_synchrony         16.2914  1  5.431e-05 ***
#habitat                   19.5303  3  0.0002124 ***
#spatial_synchrony:habitat  6.7239  3  0.0812369 .

#NOAM:
#                             Chisq Df Pr(>Chisq)    
# spatial_synchrony         16.2942  1  5.423e-05 ***
# habitat                   19.5302  3  0.0002124 ***
# spatial_synchrony:habitat  6.7253  3  0.0811882 .

performance::r2(spatial_synchrony_mod_Fig4B) # Noam = 0.64; Me: Nagelkerke's R2: 0.638 (matches)

emtrends(spatial_synchrony_mod_Fig4B, pairwise ~ habitat, var = "spatial_synchrony")
#NOAM OUTPUT
# no sig differences but close
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef            0.404 1.764 Inf   0.229  0.9958
# Fringing - Forereef 10m        1.365 1.754 Inf   0.778  0.8643
# Fringing - Forereef 17m        3.112 1.882 Inf   1.654  0.3484
# Backreef - Forereef 10m        0.961 0.857 Inf   1.121  0.6766
# Backreef - Forereef 17m        2.708 1.095 Inf   2.472  0.0643
# Forereef 10m - Forereef 17m    1.747 1.078 Inf   1.621  0.3667

```

```{r plot site level stability X spatial synchrony - Figure 4B}
spat_stab_emm <- emmip(spatial_synchrony_mod_Fig4B,
                        habitat ~ spatial_synchrony,
                        at = list(spatial_synchrony = seq(0,1,0.01)), plotit = F, CIs = T)

filtered_spat_stab_effects <- filter_ranges(spat_stab_emm, dss_spatial_ranges, "habitat", "spatial_synchrony")

(spatial_synchrony_plot_figure4b <- 
  ggplot() +
  geom_point(data = spatial_synchrony,
             aes(x = spatial_synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Site-Level Stability", x = "Spatial Synchrony", title = "", y = "Site Level Stability") +
  scale_y_continuous(limits = c(0.5, 4.2), breaks = c(1,2,3,4)) +
  model_themes
         )

```

### Figure 4C - Ratio

```{r Figure 4C - Ratio} 
#spatial_synchrony dataframe has average plot level stability AND site level stability columns, so using this one.

#add stability ratio 
# site / plot 
# Values > 1 indicate that the site is more stable than individual plots — very intuitive.
#cover_stability = SITE LEVEL
#stability_mean = MEAN PLOT LEVEL

figure4c_df<- spatial_synchrony %>%
  mutate(ratio = cover_stability/stability_mean)

# plot / site
# values < 1 suggest that the site is more stable than individual plots → indicative of a portfolio effect or spatial insurance.
#another option, but NOT USING

#shape for site
figure4c_df %>%
  ggplot(aes(x = spatial_synchrony, y = ratio, color = habitat, shape = site)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, aes(group = habitat)) +
  scale_colour_manual(values = habitat_colours) +
  labs(y = "Ratio", x = "Spatial Synchrony", title = "") +
  trends_theme #this was set at beginning of script
 # geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "red")

#no shape for site, probably using this one. 
figure4c <- figure4c_df %>%
  ggplot(aes(x = spatial_synchrony, y = ratio, color = habitat)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, aes(group = habitat, fill = habitat), alpha = 0.3) +
#  geom_ribbon(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
            #  alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Ratio", x = "Spatial Synchrony", title = "") +
  trends_theme  #this was set at beginning of script

```

```{r figure 4 panel, fig.height = 10, fig.width=18}
(figure_4 <- ggarrange(alpha_gamma_stability_plot_figure4A, spatial_synchrony_plot_figure4b, figure4c, 
                      ncol = 3, nrow = 1, 
                      common.legend = TRUE,
                      legend = "bottom", 
                      labels = c("a.", "b.", "c."),
                      font.label = list(size = 26, color = "black", face = "plain")))

#ggsave(filename = "output/figure4_09032025.jpg", figure_4, height = 15, width = 30)
```

### 2025-09-05 00_functions_and_aes.R
Began reorganizing pipeline into more workable chunks

@@ -1,0 +1,95 @@
+#### 
+# Functions for Manuscript 1 #
+####
+
+# coefficient of variation
+CV<-function(x){
+  return(sd(x,na.rm=T)/mean(x,na.rm=T))
+}
+
+inv_logit <- function(x){
+  exp(x)/(1+exp(x))
+}
+
+# smithson verkuilen transformation for 0-1 continuous data
+sv_trans <- function(prop, s = 0.000005){
+  (prop*(length(prop) - 1) + s)/length(prop)
+  # where prop is the vector of the proportional value you're transforming, 
+  # N is the sample size, which is specified by taking the number of rows/observations from a given dataframe,
+  # and s is a small offset 
+}
+
+# standard error
+se <- function(x){
+  sd(x)/sqrt(n())
+}
+
+# function to extract the ranges of the data for plotting
+extract_ranges <- function(df, # a dataframe
+                           group, # grouping variable (e.g. habitat)
+                           columns # columns you want to summarize (e.g. richness, cover, synchrony)
+){
+  df %>% # in your dataframe
+    dplyr::group_by(across(all_of(group))) %>% 
+    dplyr::summarise(across(all_of(columns), # summarize the columns in your group
+                            list(min = ~min(.x, na.rm = TRUE), max = ~max(.x, na.rm = TRUE)), # by taking the min/max of them
+                            .names = "{.fn}_{.col}")) # and assigning them to the columns name "min/max_value"
+}
+
+# function for filtering ranges of emtrends to only fit range of actual data
+filter_ranges <- function(trend, # emtrends df created from emmip
+                          range_obj, # output df from extract_ranges() custom function
+                          group, # group over which to summarize (e.g. habitat)
+                          value # value to filter range (e.g. richness)
+){
+  trend %>% # take emtrend
+    left_join(range_obj, by = group) %>% # join with range object
+    filter(.data[[value]] >= .data[[paste0("min_", value)]] & # filter so emtrend only covers range of actual values
+             .data[[value]] <= .data[[paste0("max_", value)]])
+}
+
+
+#### PLOT AESTHETICS ####
+trends_theme <- theme_bw() + 
+  theme(axis.title.x = element_text(size=24), 
+        axis.text.x = element_text(size = 22, angle = 45, hjust = 1),
+        axis.title.y = element_text(size=24), 
+        axis.text.y = element_text(size = 22),
+        plot.title = element_text(size = 24),
+        panel.background = element_rect(fill = "white"), 
[…]

### 2025-09-05 01a_data_qaqc.R

@@ -1,0 +1,264 @@
+rm(list=ls()) # clean up
+library(tidyverse)
+library(knitr)
+library(dplyr)
+library(plyr)
+library(car)
+library(reshape)
+library(reshape2)
+library(tidyr)
+library(effects)
+library(devtools)
+library(here)
+
+data <- read.csv(here("data", "MCR_LTER_Annual_Survey_Benthic_Cover_20231211.csv"), stringsAsFactors = F) # replacements don't work when strings are factors
+
+str(data)
+
+levels(as.factor(data$Taxonomy_Substrate_Functional_Group)) # 86 levels
+data$Site_habitat <- as.factor( paste(data$Site, data$Habitat))
+
+#### REPLACE OUTDATED TAXA
+
+# Julianna notes on taxa:
+# Changed Caulerpa peltata to Caulerpa chemnitzia 
+# Changed Chnoospora implexa to Pseudochnoospora implexa
+# Changed Neogoniolithon frutescens to Neogoniolithon brassica-florida
+# Changed Caulerpa pickeringii to Caulerpa webbiana
+# Changed Boodlea kaeneana to Cladophoropsis membranacea
+# Changed Dictyota divaricata to Dictyota cervicornis
+
+### ADDING JULIANNA UPDATES HERE ###
+data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Chnoospora implexa"] <- "Pseudochnoospora implexa"
+
+data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Caulerpa pickeringii"] <- "Caulerpa webbiana"
+
+data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Boodlea kaeneana"] <- "Cladophoropsis membranacea"
+
+data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Dictyota divaricata"] <- "Dictyota cervicornis"
+
+# Caulerpa peltata is outdated --> Caulerpa racemosa JULIANNA UPDATE = Caulerpa chemnitzia
+data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Caulerpa peltata"] <- "Caulerpa chemnitzia"
+
+
+### EVERYTHING ELSE BELOW AS BEFORE ###
+
+
+# Boodlea kaeneana --> Cladophoropsis membranacea
[…]

### 2025-09-05 01b_data_prep.R

@@ -1,0 +1,221 @@
+rm(list = ls())
+library(here)
+library(tidyverse)
+library(vegan)
+
+source("00_functions_and_aes.R")
+
+#### MACROALGAE DATA ####
+data <- read.csv(here::here("data", "allbenthicdata_cleaned_colnames.csv"), stringsAsFactors = F) 
+
+### if we want to make a missing or faulty data table
+# length(unique(data$location))
+# expected_data <- expand.grid(year = seq(2007,2023,1), location = unique(data$location))
+# missing_quadrats <- expected_data %>%
+#   anti_join(data, by = c("year", "location"))
+###
+
+# remove remaining backreef and fringing 2020 data because of covid extrapolation
+filtered_data <- data %>%
+  filter(!(year == 2020 & habitat %in% c("fringing", "backreef")))
+
+# re-factor for figures
+filtered_data$habitat <- factor(filtered_data$habitat, 
+                                levels = c("fringing", "backreef", "outer_10","outer_17"),
+                                labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
+
+# note non-algae species
+not_algae <- c("ascidian", "bare_space", "coral", "coral_rubble", "corallimorpharia", "heteractis_sp", "millepora_platyphylla", "no_data", "sand", "sarcophyton_sp", "shell_debris", "soft_coral", "sponge", "tridacna_sp", "zooxanthellae", "row_sums")
+
+
+# remove
+algae <- filtered_data[ , -which(names(filtered_data) %in% not_algae)]
+
+
+
+macroalgae <- algae %>% 
+  select(-algal_turf, -crustose_corallines, -cyanophyta, -damselfish_turf, -lithophyllum_kotschyanum,
+         -neogoniolithon_brassica_florida, -sporolithon_sp)
+
+meta_cols <- c("year", "location", "site", "habitat", "site_habitat", "transect", "quadrat")
+
+meta_df <- algae[,which(names(algae) %in% meta_cols)]
+
+algae_species_cols <- algae[,-which(names(algae) %in% meta_cols)]
+
+macro_species_cols <- macroalgae[,-which(names(macroalgae) %in% meta_cols)]
+
+# make long dataset
+macro_long_data <- pivot_longer(
+  data = macroalgae,
[…]

### 2025-09-05 02_time_series_plot.R

@@ -1,0 +1,83 @@
+rm(list = ls())
+library(here)
+library(tidyverse)
+library(ggplot2)
+library(ggpubr)
+
+source("01b_data_prep.R")
+
+#### CREATE MEAN AND SE DATAFRAMES ####
+
+
+cover_means <- cover_df %>%
+  dplyr::group_by(year, habitat) %>%
+  dplyr::summarise(macroalgae = mean(macroalgae),
+                   coral = mean(coral),
+                   turf = mean(turf),
+                   cca = mean(cca),
+                   sand = mean(sand),
+                   other = mean(other)
+  )
+
+cover_se <- cover_df %>%
+  dplyr::group_by(year, habitat) %>%
+  dplyr::summarise(macroalgae = se(macroalgae),
+                   coral = se(coral),
+                   turf = se(turf),
+                   cca = se(cca),
+                   sand = se(sand),
+                   other = se(other)
+  )
+
+# pivot long
+cover_long_means <- pivot_longer(
+  data = cover_means,
+  cols = 3:8,
+  names_to = "functional_group",
+  values_to = "percent_cover"
+)
+
+cover_long_se <- pivot_longer(
+  data = cover_se,
+  cols = 3:8,
+  names_to = "functional_group",
+  values_to = "se"
+)
+
+cover_long <- left_join(cover_long_means,cover_long_se, by = join_by(year, habitat, functional_group))
+
+# re-label benthic groupings for plotting
+cover_long$functional_group <- factor(cover_long$functional_group, 
+                                      levels = c("macroalgae", "coral", "turf","cca", "sand", "other"),
+                                      labels = c("Macroalgae", "Coral", "Turf", "CCA", "Sand", "Other"))
+
+# subset relevant data
+figure_1_data <- subset(cover_long, functional_group != "Sand" & functional_group != "Other")
+
+
+#### PLOT FIGURE 1 ####
+plot_list <- lapply(levels(figure_1_data$habitat), function(hab) {
+  ggplot(data = subset(figure_1_data, habitat == hab)) +
+    geom_point(aes(x = year, y = percent_cover, colour = functional_group), size = 3) +
+    geom_errorbar(aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, colour = functional_group), 
+                  width = 0, size = 1.3) +
+    geom_line(aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
[…]

### 2025-09-05 03a_diversity_cover_models.R

@@ -1,0 +1,53 @@
+rm(list = ls())
+library(here)
+library(tidyverse)
+library(glmmTMB)
+library(emmeans)
+library(car)
+library(effects)
+library(DHARMa)
+
+source("01b_data_prep.R")
+
+#### COVER ~ RICHNESS ####
+cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro)
+summary(cover_mod)
+Anova(cover_mod)
+#                     Chisq Df Pr(>Chisq)    
+# richness         30918.96  1  < 2.2e-16 ***
+# habitat            141.68  3  < 2.2e-16 ***
+# richness:habitat   294.44  3  < 2.2e-16 ***
+hist(residuals(cover_mod)) # looks fine
+plot(residuals(cover_mod) ~ fitted(cover_mod)) # negative trend but it can't be less than 0 so not too concerned. some wave pattern
+performance::r2(cover_mod) # 0.635, 0.654
+emtrends(cover_mod, pairwise ~ habitat, var = "richness") # backreef not different from forereef, forereef also not different
+# contrast                    estimate     SE  df z.ratio p.value
+# Fringing - Backreef          0.34715 0.0241 Inf  14.407  <.0001
+# Fringing - Forereef 10m      0.34862 0.0240 Inf  14.513  <.0001
+# Fringing - Forereef 17m      0.38937 0.0250 Inf  15.593  <.0001
+# Backreef - Forereef 10m      0.00147 0.0205 Inf   0.072  0.9999
+# Backreef - Forereef 17m      0.04223 0.0214 Inf   1.975  0.1974
+# Forereef 10m - Forereef 17m  0.04076 0.0209 Inf   1.947  0.2087
+
+#### COVER ~ FUNCTIONAL RICHNESS ####
+
+cover_mod_fg <- glmmTMB(cover_trans ~ functional_richness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro)
+
+
+summary(cover_mod_fg)
+car::Anova(cover_mod_fg)
+#                                 Chisq Df Pr(>Chisq)    
+# functional_richness         31565.55  1  < 2.2e-16 ***
+# habitat                        49.36  3  1.093e-10 ***
+# functional_richness:habitat   444.61  3  < 2.2e-16 ***
[…]

### 2025-09-05 03b_diversity_cover_figures.R

@@ -1,0 +1,75 @@
+rm(list = ls())
+source("03a_diversity_cover_models.R")
+library(here)
+library(tidyverse)
+library(glmmTMB)
+library(emmeans)
+library(car)
+library(effects)
+library(DHARMa)
+library(ggplot2)
+library(ggpubr)
+
+
+
+#### COVER ~ RICHNESS ####
+# SETUP
+
+diversity_ranges_quad <- extract_ranges(alpha_diversity_quad_macro, "habitat", c("richness", "cover_trans", "functional_richness"))
+
+cover_emm <- emmip(cover_mod,
+                   habitat ~ richness,
+                   at = list(richness = seq(0,10,1)), plotit = F, CIs = T)
+
+
+
+filtered_cover_effects <- filter_ranges(trend = cover_emm, range_obj = diversity_ranges_quad, group = "habitat", value = "richness")
+
+# PLOT
+(figure_3a <- 
+    ggplot() +
+#    geom_point(data = alpha_diversity_quad_macro,
+#               aes(x = richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.75) + 
+    geom_line(data = filtered_cover_effects, aes(x = richness, y = inv_logit(yvar), colour = habitat)) +
+    geom_ribbon(data = filtered_cover_effects, aes(x = richness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
+                alpha = 0.5) +
+    scale_colour_manual(values = habitat_colours) +
+    scale_fill_manual(values = habitat_colours) +
+    labs(y = "Cover\n", x = "Taxonomic richness", title = "") +
+    model_themes 
+)
+# ggsave(filename = "output/figure_3a.png", figure_3a, height = 10, width = 14)
+
+#### COVER ~ FUNCTIONAL RICHNESS ####
+# SETUP
+cover_emm_fg <- emmip(cover_mod_fg,
+                      habitat ~ functional_richness,
+                      at = list(functional_richness = seq(0,10,1)), plotit = F, CIs = T)
+
+filtered_cover_effects_fg <- filter_ranges(cover_emm_fg, diversity_ranges_quad, "habitat", "functional_richness")
+
+# PLOT
+(figure_3b <- 
+    ggplot() +
+    geom_jitter(data = alpha_diversity_quad_macro,
+               aes(x = functional_richness, y = cover_trans, colour = habitat), width = 0.2, size = 1, alpha = 0.1) + 
[…]

### 2025-09-05 04a_calculate_synchrony.R

@@ -1,0 +1,75 @@
+rm(list = ls())
+library(here)
+library(tidyverse)
+library(glmmTMB)
+library(emmeans)
+library(car)
+library(effects)
+library(DHARMa)
+library(codyn)
+
+source("01b_data_prep.R")
+
+#### SPECIES SYNCHRONY ####
+lm_synchrony <- codyn::synchrony(df = macro_long_data,
+                                 time.var = "year",
+                                 species.var = "taxa",
+                                 abundance.var = "prop_cover",
+                                 metric = "Loreau",
+                                 replicate.var = "location")
+
+
+diversity_stability <- alpha_diversity_quad_macro %>%
+  dplyr::group_by(location, habitat, site) %>%
+  dplyr::summarise(
+    richness_mean = mean(richness), 
+    shannon_mean = mean(shannon), 
+    cover_mean = mean(macroalgal_prop_cover),
+    functional_richness_mean = mean(functional_richness),
+    cover_cv = CV(macroalgal_prop_cover),
+    cover_stability = 1/CV(macroalgal_prop_cover)) 
+
+diversity_stability_synchrony <- left_join(diversity_stability, lm_synchrony, by = join_by(location))
+diversity_stability_synchrony$synchrony_trans <- sv_trans(diversity_stability_synchrony$synchrony)
+# write.csv(file = here::here("data", "full_plot_level_dss.csv"), diversity_stability_synchrony)
+
+# EXTRACT RANGES
+dss_ranges <- extract_ranges(diversity_stability_synchrony, "habitat",
+                             c("richness_mean", "functional_richness_mean", "cover_stability", "synchrony"))
+
+#### SITE LEVEL SPECIES SYNCHRONY ####
+macro_long_data_site <- pivot_longer(
+  data = site_macro,
+  cols = grep("acanthophora_spicifera", colnames(site_macro)):grep("valonia_ventricosa", colnames(site_macro)),
+  names_to = "taxa",
+  values_to = "percent_cover") %>% 
+  mutate(prop_cover = percent_cover/100) %>% 
+  select(-percent_cover)
+
+# write_csv(macro_long_data, here("Data", "macro_long_data.csv"))
+
+# CALCULATE SYNCHRONY CODE
+lm_synchrony_site <- codyn::synchrony(df = macro_long_data_site,
+                                      time.var = "year",
+                                      species.var = "taxa",
+                                      abundance.var = "prop_cover",
+                                      metric = "Loreau",
[…]

### 2025-09-05 04b_dss_models.R

@@ -1,0 +1,77 @@
+rm(list = ls())
+library(glmmTMB)
+library(emmeans)
+library(car)
+library(effects)
+library(DHARMa)
+library(codyn)
+
+source("04a_calculate_synchrony.R")
+
+#### STABILITY ~ RICHNESS ####
+rich_stab_mod <- glmmTMB(cover_stability~ richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = diversity_stability_synchrony)
+summary(rich_stab_mod)
+Anova(rich_stab_mod)
+#                          Chisq Df Pr(>Chisq)    
+# richness_mean         2720.311  1  < 2.2e-16 ***
+# habitat                325.657  3  < 2.2e-16 ***
+# richness_mean:habitat   33.857  3  2.124e-07 ***
+hist(residuals(rich_stab_mod)) # looks good
+plot(residuals(rich_stab_mod) ~ fitted(rich_stab_mod)) # heteroskedastic - bring to group
+r.squaredGLMM(rich_stab_mod) # 0.803, 0.808
+performance::r2(rich_stab_mod) # 0.812, 0.816
+emtrends(rich_stab_mod, pairwise ~ habitat, var = "richness_mean") # fringing not diff from outer17, backreef not different from outer10
+# $contrasts
+#  contrast                     estimate     SE  df z.ratio p.value
+#  Fringing - Backreef          0.186640 0.0414 Inf   4.511  <.0001
+#  Fringing - Forereef 10m      0.187248 0.0455 Inf   4.114  0.0002
+#  Fringing - Forereef 17m      0.027482 0.0463 Inf   0.594  0.9340
+#  Backreef - Forereef 10m      0.000608 0.0405 Inf   0.015  1.0000
+#  Backreef - Forereef 17m     -0.159158 0.0411 Inf  -3.875  0.0006
+#  Forereef 10m - Forereef 17m -0.159766 0.0439 Inf  -3.640  0.0016
+
+#### STABILITY ~ FUNCTIONAL RICHNESS ####
+rich_stab_mod_fg <- glmmTMB(cover_stability ~ functional_richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = diversity_stability_synchrony)
+
+summary(rich_stab_mod_fg)
+car::Anova(rich_stab_mod_fg)
+
+#                                    Chisq Df Pr(>Chisq)    
[…]

### 2025-09-05 04c_dss_figures.R

@@ -1,0 +1,88 @@
+source("04b_dss_models.R")
+library(ggplot2)
+library(ggpubr)
+
+#### STABILITY ~ RICHNESS ####
+# SETUP
+rich_stab_emm <- emmip(rich_stab_mod,
+                       habitat ~ richness_mean,
+                       at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_rich_stab_effects <- filter_ranges(rich_stab_emm, dss_ranges, "habitat", "richness_mean")
+
+# PLOT
+(figure_4a <- 
+    ggplot() +
+    geom_point(data = diversity_stability_synchrony,
+               aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+    geom_line(data = filtered_rich_stab_effects, aes(x = richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
+    geom_ribbon(data = filtered_rich_stab_effects, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+                alpha = 0.3) +
+    scale_colour_manual(values = habitat_colours) +
+    scale_fill_manual(values = habitat_colours) +
+    labs(y = "Stability\n", x = "Taxonomic richness", title = "") +
+    model_themes
+)
+
+# ggsave(filename = "output/figure_4a.png", figure_4a, height = 10, width = 14)
+
+#### STABILITY ~ FUNCTIONAL RICHNESS ####
+# SETUP
+rich_stab_emm_fg <- emmip(rich_stab_mod_fg,
+                          habitat ~ functional_richness_mean,
+                          at = list(functional_richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_rich_stab_effects_fg <- filter_ranges(rich_stab_emm_fg, dss_ranges, "habitat", "functional_richness_mean")
+
+# PLOT
+(figure_4b <- 
+    ggplot() +
+    geom_point(data = diversity_stability_synchrony,
+               aes(x = functional_richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+    geom_line(data = filtered_rich_stab_effects_fg, aes(x = functional_richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
+    geom_ribbon(data = filtered_rich_stab_effects_fg,
[…]

### 2025-09-05 04d_site-level_dss_models.R

@@ -1,0 +1,78 @@
+rm(list = ls())
+library(glmmTMB)
+library(emmeans)
+library(car)
+library(effects)
+library(DHARMa)
+library(codyn)
+
+source("04a_calculate_synchrony.R")
+
+#### STABILITY ~ RICHNESS ####
+rich_stab_mod_site <- glmmTMB(cover_stability~ richness_mean*habitat, family = Gamma(link = "log"), data = diversity_stability_synchrony_site)
+summary(rich_stab_mod_site)
+Anova(rich_stab_mod_site)
+
+#                        Chisq Df Pr(>Chisq)    
+# richness_mean         17.645  1  2.662e-05 ***
+# habitat               32.229  3  4.682e-07 ***
+# richness_mean:habitat  8.075  3    0.04449 *  
+hist(residuals(rich_stab_mod_site)) # blocky but fine
+plot(residuals(rich_stab_mod_site) ~ fitted(rich_stab_mod_site)) # heteroskedastic - bring to group
+performance::r2(rich_stab_mod_site)
+emtrends(rich_stab_mod_site, pairwise ~ habitat, var = "richness_mean") # no differences across habitats
+# contrast                    estimate    SE  df z.ratio p.value
+# Fringing - Backreef           0.1854 0.120 Inf   1.539  0.4138
+# Fringing - Forereef 10m      -0.2155 0.145 Inf  -1.484  0.4471
+# Fringing - Forereef 17m      -0.2876 0.214 Inf  -1.344  0.5349
+# Backreef - Forereef 10m      -0.4009 0.165 Inf  -2.432  0.0711
+# Backreef - Forereef 17m      -0.4730 0.228 Inf  -2.077  0.1607
+# Forereef 10m - Forereef 17m  -0.0721 0.242 Inf  -0.298  0.9908
+
+#### STABILITY ~ FUNCTIONAL RICHNESS ####
+
+rich_stab_mod_site_fg <- glmmTMB(cover_stability~ functional_richness_mean*habitat, family = Gamma(link = "log"), data = diversity_stability_site)
+summary(rich_stab_mod_site_fg)
+Anova(rich_stab_mod_site_fg)
+
+# 
+#                                    Chisq Df Pr(>Chisq)   
+# functional_richness_mean          2.2098  1   0.137136   
+# habitat                          14.5000  3   0.002298 **
+# functional_richness_mean:habitat 12.6102  3   0.005560 **
+
[…]

### 2025-09-05 04e_site-level_dss_figures.R

@@ -1,0 +1,91 @@
+source("04d_dss_models.R")
+library(ggplot2)
+library(ggpubr)
+
+#### STABILITY ~ RICHNESS ####
+# SETUP
+rich_stab_emm_site <- emmip(rich_stab_mod_site,
+                            habitat ~ richness_mean,
+                            at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_rich_stab_effects_site <- filter_ranges(rich_stab_emm_site, dss_ranges_site, "habitat", "richness_mean")
+
+# PLOT
+(figure_4d <- 
+   ggplot() +
+   geom_point(data = diversity_stability_synchrony_site,
+              aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+   geom_line(data = filtered_rich_stab_effects_site, aes(x = richness_mean, y = exp(yvar), colour = habitat), size = 1.5) +
+   geom_ribbon(data = filtered_rich_stab_effects_site, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+               alpha = 0.3) +
+   scale_colour_manual(values = habitat_colours) +
+   scale_fill_manual(values = habitat_colours) +
+   scale_y_continuous(limits = c(0.5, 4.2), breaks = c(1,2,3,4)) +
+   labs(y = "Site-level Stability\n", x = "Taxonomic richness", title = "") +
+   model_themes
+)
+
+# ggsave(filename = "output/figure_4d.png", figure_4d, height = 10, width = 14)
+
+#### STABILITY ~ FUNCTIONAL RICHNESS ####
+# SETUP
+rich_stab_emm_site_fg <- emmip(rich_stab_mod_site_fg,
+                               habitat ~ functional_richness_mean,
+                               at = list(functional_richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)
+
+filtered_rich_stab_effects_site <- filter_ranges(rich_stab_emm_site_fg, dss_ranges_site, "habitat", "functional_richness_mean")
+
+# PLOT
+(figure_4e <- 
+    ggplot() +
+    geom_point(data = diversity_stability_site,
+               aes(x = functional_richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
[…]

### 2025-09-05 05a_calculate_spatial_synchrony_and_bray.R

@@ -1,0 +1,196 @@
+rm(list = ls())
+source("04a_calculate_synchrony.R")
+library(tidyverse)
+library(vegan)
+library(codyn)
+
+#### SPATIAL SYNCHRONY ####
+# calculate mean synchrony and stability within a site
+mean_synchrony_stability <- diversity_stability_synchrony %>% 
+  dplyr::group_by(habitat, site) %>% 
+  dplyr::summarise(
+    richness_mean = mean(richness_mean, na.rm = T),
+    functional_richness_mean = mean(functional_richness_mean, na.rm = T),
+    synchrony_mean = mean(synchrony, na.rm = T),
+    stability_mean = mean(cover_stability, na.rm = T),
+    cover_mean = mean(cover_mean, na.rm = T))
+
+# calculate spatial synchrony - degree to which quadrats vary together within a site
+spatial_synchrony <- codyn::synchrony(df = alpha_diversity_quad_macro,
+                                      time.var = "year",
+                                      species.var = "location",
+                                      abundance.var = "macroalgal_prop_cover",
+                                      metric = "Loreau",
+                                      replicate.var = "site_habitat")
+colnames(spatial_synchrony) <- c("site_habitat", "spatial_synchrony") # rename to avoid overwrite
+
+
+
+# merge with site level dss metrics
+dss_spatial <- merge(diversity_stability_site, spatial_synchrony, by = "site_habitat")
+dss_spatial_2 <- merge(dss_spatial, mean_synchrony_stability[, which(names(mean_synchrony_stability) %in%
+                                                                       c("habitat", "site", "synchrony_mean", "stability_mean"))],
+                       by = c("habitat", "site"))
+
+
+dss_spatial_2$ratio <- dss_spatial_2$cover_stability/dss_spatial_2$stability_mean
+
+# write.csv(file = here::here("data", "spatial_synchrony.csv"), dss_spatial_2)
+
+# calculate ranges
+dss_spatial_ranges <- extract_ranges(dss_spatial_2, "habitat", c("spatial_synchrony", "synchrony_mean", "stability_mean"))
+
+#### BRAY ####
+##### setup #####
+# SUM the data across years for each quad 
+macro_long_data_summed <- macro_long_data %>%
[…]

### 2025-09-05 05b_spatial_synchrony_and_bray_models.R

@@ -1,0 +1,95 @@
+rm(list = ls())
+library(glmmTMB)
+library(emmeans)
+library(car)
+library(effects)
+library(DHARMa)
+library(codyn)
+
+source("05a_calculate_spatial_synchrony_and_bray.R")
+
+#### REGIONAL STABILITY ~ LOCAL STABILITY ####
+alpha_gamma_stab_mod <- glmmTMB(cover_stability ~ stability_mean*habitat, family = Gamma("log"), data = dss_spatial_2)
+summary(alpha_gamma_stab_mod)
+Anova(alpha_gamma_stab_mod)
+#                          Chisq Df Pr(>Chisq)    
+# stability_mean         61.2742  1  4.965e-15 ***
+# habitat                12.9328  3   0.004784 ** 
+# stability_mean:habitat  9.4834  3   0.023508 * 
+performance::r2(alpha_gamma_stab_mod) # 0.838
+
+emtrends(alpha_gamma_stab_mod, pairwise ~ habitat, var = "stability_mean") # backreef different from fringe and forereef 10
+# $contrasts
+#  contrast                    estimate    SE  df z.ratio p.value
+#  Fringing - Backreef            1.078 0.381 Inf   2.828  0.0242
+#  Fringing - Forereef 10m        0.175 0.345 Inf   0.509  0.9571
+#  Fringing - Forereef 17m        0.380 0.305 Inf   1.248  0.5965
+#  Backreef - Forereef 10m       -0.903 0.349 Inf  -2.589  0.0474
+#  Backreef - Forereef 17m       -0.698 0.309 Inf  -2.256  0.1085
+#  Forereef 10m - Forereef 17m    0.205 0.263 Inf   0.779  0.8639
+
+
+#### REGIONAL STABILITY ~ SPATIAL SYNCHRONY ####
+spatial_synchrony_mod <- glmmTMB(cover_stability ~ spatial_synchrony*habitat, family = Gamma("log"), data = dss_spatial_2)
+hist(residuals(spatial_synchrony_mod)) # a little skew
+plot(residuals(spatial_synchrony_mod) ~ predict(spatial_synchrony_mod))
+summary(spatial_synchrony_mod)
+Anova(spatial_synchrony_mod)
+#                             Chisq Df Pr(>Chisq)    
+# spatial_synchrony         16.2942  1  5.423e-05 ***
[…]

### 2025-09-05 05c_spatial_synchrony_and_bray_figures.R

@@ -1,0 +1,107 @@
+rm(list = ls())
+source("05b_spatial_synchrony_and_bray_models.R")
+library(ggplot2)
+library(ggpubr)
+
+#### REGIONAL STABILITY ~ LOCAL STABILITY ####
+# SETUP
+alpha_gamma_emm <- emmip(alpha_gamma_stab_mod,
+                         habitat ~ stability_mean,
+                         at = list(stability_mean = seq(0,10,0.01)), plotit = F, CIs = T)  
+
+filtered_ag_effects <- filter_ranges(alpha_gamma_emm, dss_spatial_ranges, "habitat", "stability_mean")
+
+# PLOT
+(alpha_gamma_stability_plot <- 
+    ggplot() +
+    geom_point(data = dss_spatial_2,
+               aes(x = stability_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+    geom_line(data = filtered_ag_effects, aes(x = stability_mean, y = exp(yvar), colour = habitat), size = 1.5) +
+    geom_ribbon(data = filtered_ag_effects, aes(x = stability_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
+                alpha = 0.3) +
+    scale_colour_manual(values = habitat_colours) +
+    scale_fill_manual(values = habitat_colours) +
+    labs(y = "Regional stability", x = "Mean plot-level stability", title = "") +
+#    scale_y_continuous(limits = c(0.5, 4.2), breaks = c(1,2,3,4)) +
+    model_themes
+)
+
+#### REGIONAL STABILITY ~ SPATIAL SYNCHRONY ####
+# SETUP
+spat_stab_emm <- emmip(spatial_synchrony_mod,
+                       habitat ~ spatial_synchrony,
+                       at = list(spatial_synchrony = seq(0,1,0.01)), plotit = F, CIs = T)  
+
+filtered_spat_stab_effects <- filter_ranges(spat_stab_emm, dss_spatial_ranges, "habitat", "spatial_synchrony")
+
+# PLOT
+(spatial_synchrony_plot <- 
+    ggplot() +
+    geom_point(data = dss_spatial_2,
+               aes(x = spatial_synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
+    geom_line(data = filtered_spat_stab_effects, aes(x = spatial_synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
[…]

### 2025-09-05 05c_spatial_synchrony_and_bray_figures.R
corrected label

@@ -21,7 +21,7 @@ filtered_ag_effects <- filter_ranges(alpha_gamma_emm, dss_spatial_ranges, "habit
                 alpha = 0.3) +
     scale_colour_manual(values = habitat_colours) +
     scale_fill_manual(values = habitat_colours) +
-    labs(y = "Regional stability", x = "Mean plot-level stability", title = "") +
+    labs(y = "Site-level stability", x = "Mean plot-level stability", title = "") +
 #    scale_y_continuous(limits = c(0.5, 4.2), breaks = c(1,2,3,4)) +
     model_themes
 )
@@ -44,7 +44,7 @@ filtered_spat_stab_effects <- filter_ranges(spat_stab_emm, dss_spatial_ranges, "
                 alpha = 0.3) +
     scale_colour_manual(values = habitat_colours) +
     scale_fill_manual(values = habitat_colours) +
-    labs(y = "Regional stability", x = "Spatial synchrony", title = "") +
+    labs(y = "Site-level stability", x = "Spatial synchrony", title = "") +
 #    scale_y_continuous(limits = c(0.5, 4.2), breaks = c(1,2,3,4)) +
     model_themes
 )
@@ -68,7 +68,7 @@ filtered_ratio_effects <- filter_ranges(ratio_emm, dss_spatial_ranges, "habitat"
     scale_colour_manual(values = habitat_colours) +
     scale_fill_manual(values = habitat_colours) +
     labs(y = "", x = "Spatial synchrony", title = "") +
-    ylab(expression(frac(Regional~stability, Mean~plot-level~stability))) +
+    ylab(expression(frac(Site-level~stability, Mean~plot-level~stability))) +
 #    scale_y_continuous(limits = c(0.5, 4.2), breaks = c(1,2,3,4)) +
     model_themes
 )
